### 2018-10-03 raw_notes_questions.rmd
Update raw_notes_questions.rmd

# Question
##When I write to the Master, I am writing to the forked copy of this website, correct? 
So if that is correct, and Mike is still listed as a collaborator:
Are MY edits being kept here until I send a Pull/Push request?
My edits only affect this repo, correct?

### 2018-10-03 raw_notes_questions.rmd
Update raw_notes_questions.rmd

@@
 # Question
-##When I write to the Master, I am writing to the forked copy of this website, correct? 
+#### When I write to the Master, I am writing to the forked copy of this website, correct? 
 So if that is correct, and Mike is still listed as a collaborator:
@@
 My edits only affect this repo, correct?
+
+# Question
+#### Anyone else need a QUICK tour of the branch and pull request map?
+
+# Question
+#### Still unclear on difference between a fork and a branch. 
+Branch is an alternative version of a single file, but fork is a copy of the original repo in total. Y/N?
+
+# Question
+#### When I am working in my own repo, does it propagate? How do I connect to it in R?
+When I created my default website repo via fork, I assume I can edit that to my liking, and it will not be propagated back to whomever I forked it from.
+I also can presumably download a ZIP of a repo and install it to my local machine (Dropbox, actually).
+
+So how do I tell R to use my GitHub repo's as their source for data?
+
+# Question
+#### Having a slight problem with Git on laptop.
+I can use my desktop and create repos to Dropbox or local machine from GitHub. Laptop throws in an extra step that seems to corrupt process.

### 2018-10-03 raw_notes_questions.rmd
Update

@@
 # Question
-#### When I write to the Master, I am writing to the forked copy of this website, correct? 
-So if that is correct, and Mike is still listed as a collaborator:
-Are MY edits being kept here until I send a Pull/Push request?
-My edits only affect this repo, correct?
+#### When I write to the Master, I am writing to my forked copy of the website, correct? 
+If Mike is still listed as a collaborator, MY edits are only pushed to my version in my repo, and affect only my copy. If I send a Push request, Mike can pull in my changes IF HE CHOOSES?
 
@@
 # Question
-#### Still unclear on difference between a fork and a branch. 
+#### Still not sure I know difference between a fork and a branch. 
 Branch is an alternative version of a single file, but fork is a copy of the original repo in total. Y/N?
@@
 # Question
-#### When I am working in my own repo, does it propagate? How do I connect to it in R?
-When I created my default website repo via fork, I assume I can edit that to my liking, and it will not be propagated back to whomever I forked it from.
-I also can presumably download a ZIP of a repo and install it to my local machine (Dropbox, actually).
+#### Can I build multiple files in R, then push project to GitHub as a complete repo?
+Yep. Found instructions for doing it on this website.
+https://cfss.uchicago.edu/git05.html
+
+Is this correct, that if need be, I could download a ZIP of a project from someone else, unpack in my local machine (Dropbox, actually), then upload back to GitHub just the same as if I cloned it directly inside GH?
+
+I know it is the long way around, but if I have a folder of resources someone gives me, it seems rude to ask, "can you put that on GitHub first for me?"
 
@@
 I can use my desktop and create repos to Dropbox or local machine from GitHub. Laptop throws in an extra step that seems to corrupt process.
+
+**SOLVED**. The path to Git was wrong. Needs to be very specific: 
+/usr/local/git/bin/git
+
+Other Git and Git-GUI paths do not work.

### 2018-10-22 index.Rmd
Update

---
title: "Data Analysis in R: Research & Learning"
---

```{r, child="_setup.Rmd"}
```

## So It Begins

This is a general landing page for your website for your project. This is just a launching pad for our community. Feel free to customise as much as you would like.

You can use this page to give introductions about what you are doing and your aims.

The structure of the website can be changed by manipulating the `_site.yml`. Font Awesome icons are available to use. You can check those out [here](https://fontawesome.com/v4.7.0/icons/).

## General Structure

We have provided a rough template for your website
- **course_notes.Rmd** - Write your notes for the class, try some code, work some exercises and use this as a virtual notepad for the items we review. This is a great way to showcase new skills and serve as a reference when you can't quite remember how you did something

- **pre_work.Rmd** - Use this document to complete any pre-work assignments.

- **project.Rmd** - As you might have guessed, use this section to write up your project

- **library.bib** - is text file that is available for you to put citations. You can then use these citations in the other Rmd documents. Also note that if you want to cite a package and want the bibtex you can do this by using the command `citation("packagename")`.

- **footer.html** - Go ahead and put your name in there!

When you are happy with what you have written for the day, go ahead and make a website out of it by using the following command in the console `rmarkdown::render_site()`

### 2018-10-24 questions.rmd
update

--- 2018-10-03 raw_notes_questions.rmd
@@
-# Question
+# R Code
+
+-----
+
+# Project
+
+-----
+
+# Github
 #### When I write to the Master, I am writing to my forked copy of the website, correct? 
@@
 
-# Question
-#### Anyone else need a QUICK tour of the branch and pull request map?
+-----
 
-# Question
 #### Still not sure I know difference between a fork and a branch. 
@@
 
-# Question
+-----
+
 #### Can I build multiple files in R, then push project to GitHub as a complete repo?
@@
 
-# Question
+-----
+
 #### Having a slight problem with Git on laptop.

### 2018-10-27 project.Rmd
test live shiny

---
title: "Project"
author: "Dan Johnson"
date: "10/26/2018"
runtime: shiny
output: html_document
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

# Abstract

v### Here are two Shiny widgets

```{r echo = FALSE}
selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20)

sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
```

### ...that build a histogram.

```{r echo = FALSE}
renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")

  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

# Introduction

# Literature Review

# Methodology

## Data

## Analysis

# Results

# Discussion

# Conclusion

# References

### 2018-10-31 project.Rmd
update

@@
 ---
-
 ```{r, child="_setup.Rmd"}
@@
 
-v### Here are two Shiny widgets
+### Here are two Shiny widgets
 

### 2018-10-31 questions.rmd
update

@@
 # R Code
+#### I can write code for Shiny and build pages locally, but they do not build for a website. 
+The build stalls here, with the error message:
+
+>label: unnamed-chunk-1 (with options) 
+>List of 1  
+> $ echo: logi FALSE  
+>Error in selectInput("n_breaks", label = "Number of bins:", choices = c(10,  :   
+>  could not find function "selectInput"  
+>Calls: <Anonymous> ... handle -> withCallingHandlers -> withVisible -> eval -> eval  
+>Execution halted  
+>Exited with status 1.
+
+Ideas?
 
@@
 # Github
-#### When I write to the Master, I am writing to my forked copy of the website, correct? 
-If Mike is still listed as a collaborator, MY edits are only pushed to my version in my repo, and affect only my copy. If I send a Push request, Mike can pull in my changes IF HE CHOOSES?
+#### When I work on Github, I see no easy way to create new file folders.
+
+If I empty a folder, it is deleted automatically. Am I overlooking tools, or is goal to be working locally to create files and use web interface less often?
+
+-----
+
+#### I am not sure how to clone and blend my OWN repos. 
+I was trying to build a starting point with both data code and web site pieces by cloning the web site repo, then adding the standard data layout folder.
+Could not see an obvious way to clone one of my own, or to blend two into one.
 

### 2018-11-19 project.Rmd
modified

@@
 date: "10/26/2018"
-runtime: shiny
 output: html_document
@@
 
-### Here are two Shiny widgets
+# Introduction
+## Background
+In the context of introductory STEM courses, technical writing training for undergraduates often is poorly designed, implemented, and executed.
+Often it is treated as a mechanical skill that each student develops "along the way." Also, faculty often are not directly responsible for developing this skill.
+Technical writing training usually is part of a lab course associated with a lecture class.
+These lab sections may be be led by graduate teaching assistants or even advanced undergraduates.
+These class leaders may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.
 
-```{r echo = FALSE}
-selectInput("n_breaks", label = "Number of bins:",
-              choices = c(10, 20, 35, 50), selected = 20)
+We have been evaluating standardized approaches and techniques for teaching technical writing that:
 
-sliderInput("bw_adjust", label = "Bandwidth adjustment:",
-              min = 0.2, max = 2, value = 1, step = 0.2)
-```
+*  Focus on holistic coaching, 
+*  Embody best practices from the writing in the disciplines literature
+*  Can be implemented at large scale by less experienced teaching assistants
+*  Focus student and instructor attention on larger issues rather than simple copy-correction.
 
-### ...that build a histogram.
+As part of this project, we have assembled a structured database of laboratory reports written by undergraduate students for introductory biology courses.
+In addition to the students' original texts we have collected these meta-data:
 
-```{r echo = FALSE}
-renderPlot({
-  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
-       xlab = "Duration (minutes)", main = "Geyser eruption duration")
+*  Circumstances: first or second report of semester; initial submission or revision; course; topic; year of student author in college
+*  Instructor information: identity of TA providing feedback; score assigned; rationale behind score
+*  All comments that a TA made on the individual reports they graded.
 
-  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
-  lines(dens, col = "blue")
-})
-```
+Using text-mining, we are searching for features of student writing, as well as connections and correlations between those features, that can inform how we teach technical writing.
 
-# Introduction
+## Context for This Project
+It is well established that students make greater gains as writers if instructors adopt a coaching- centric approach and provide holistic, growth-oriented guidance rather than just telling students what to correct.
+To that end, we train graduate teaching assistants use a **global first, specific later** approach to grading undergraduate technical writing.
+They are instructed not to spend endless time copy editing.
+Instead we ask them to spend a limited block of time writing fewer comments that focus on broader issues.
+Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection.
+As students' writing improves, the TA can focus on smaller issues.
+
+While this approach is sound in theory, we are uncertain how well TAs can implement the recommended strategy.
+One of our ongoing questions has been, "what do TAs ACTUALLY spend the most time and effort making comments about as they grade student writing?"
+
+In Summer 2018, we began trying to answer this question by analyzing TAs' comments from student papers.
+We extracted and tabulated all TA comments from approximately 500 pairs of reports (initial submissions and revisions) in 3 different courses collected during a single semester.
+Initial data cleaning involved removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging the intent.
+In all, the final tabulated dataset included ~11,000 separate comments.
+
+We next used a subset of 100 randomly selected comments to build an initial qualitative codebook for sorting comments into types.
+The initial codebook was re-tested using 1100 additional randomly selected comments.
+Using the final codebook, we sorted all 11,000 comments into separate categories (see LINK for complete codebook.)
+
+Coded TA comments were sorted on two axes:
+
+The resulting sorted table provided us with a relative estimate of TA effort.
+
+## The General Problem
+While the qualitative coding of TA comments is informative, it is unsustainable.
+A single investigator worked for nearly 2 months to develop the codebook then rate each of 11,000 comments.
+Even with the codebook, individual coders will require training to achieve any accuracy.
+Given that 10-12,000 comments are the normal output EACH SEMESTER, an automated sorting method is needed.
+
+### Specific Problems in the Analysis
+I have 11,000 comments from TAs on student reports. I want to parse them into topic groups more or less automatically, using pre-defined criteria.
+
+#### Approach Option: Latent Dirichlet allocation
+In text mining, we often have collections of documents, such as blog posts or news articles, that we’d like to divide into natural groups so that we can understand them separately.
+Topic modeling is a method for unsupervised classification of such documents, similar to clustering on numeric data, which finds natural groups of items even when we’re not sure what we’re looking for.
+
+Latent Dirichlet allocation (LDA) is a particularly popular method for fitting a topic model.
+It treats each document as a mixture of topics, and each topic as a mixture of words.
+This allows documents to “overlap” each other in terms of content, rather than being separated into discrete groups, in a way that mirrors typical use of natural language.
+
+	https://www.tidytextmining.com/topicmodeling.html
+
+Challenge is that I have groups **already based on the existing coding book.** The individual groups are already labeled. LDA uses clustering based on its own analysis.
+
+#### Approach Option: Comparison Modeling Based on Twitter Strategies
+What if I treated each of the pre-defined sub-sets from primary codebook analysis as if they are tweets from separate users? 
+	https://www.tidytextmining.com/twitter.html
+
+Strategy would be:
+*  Un-nest tokens to create a tidy frame and remove stop words using anti-join
+*  Calculate word frequencies for each sub-type
+*  Compare frequencies by calculating log-odds ratios
 
@@
 
-# Methodology
+# Methods
+## Preliminary Proof of Concept
 

### 2018-11-22 course_notes.Rmd
change

---
title: "Course Notes"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

You can detail what you learned in the session within these pages. You can also use the Rmd markdown language and insert code chunks with `CMD/CTRL + SHIFT + I`.

# Session 1
Change

# Session 2

# Session 3

# Session 4

# Session 5

# Session 6

# Session 7

# Session 8

# References

### 2018-11-22 project.Rmd
chn

@@
 
-*  Focus on holistic coaching, 
+*  Focus on holistic coaching 
 *  Embody best practices from the writing in the disciplines literature

### 2018-11-22 project.Rmd
change

@@
 ## Background
-In the context of introductory STEM courses, technical writing training for undergraduates often is poorly designed, implemented, and executed.
-Often it is treated as a mechanical skill that each student develops "along the way." Also, faculty often are not directly responsible for developing this skill.
+Undergraduate introductory STEM courses typically include technical writing training.
+Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed.
+Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills.
+
 Technical writing training usually is part of a lab course associated with a lecture class.
@@
 
-*  Focus on holistic coaching 
-*  Embody best practices from the writing in the disciplines literature
+*  Embody best practices from nearly 40 years of "writing in the disciplines"" literature
 *  Can be implemented at large scale by less experienced teaching assistants
+*  Emphasize student development through holistic coaching 
 *  Focus student and instructor attention on larger issues rather than simple copy-correction.
 
-As part of this project, we have assembled a structured database of laboratory reports written by undergraduate students for introductory biology courses.
-In addition to the students' original texts we have collected these meta-data:
+We have amassed a structured database of laboratory reports written by undergraduate students for introductory biology courses. Besides the original texts we have collected these meta-data:
 
@@
 
-Using text-mining, we are searching for features of student writing, as well as connections and correlations between those features, that can inform how we teach technical writing.
+We are using text-mining approaches to map key features of student writing, as well as connections and correlations between those features, that can inform how we teach technical writing.
 
 ## Context for This Project
-It is well established that students make greater gains as writers if instructors adopt a coaching- centric approach and provide holistic, growth-oriented guidance rather than just telling students what to correct.
+Students make greater gains as writers when instructors avoid just telling students what to correct ("copy editing"), and adopt a coaching- oriented approach instead.
 To that end, we train graduate teaching assistants use a **global first, specific later** approach to grading undergraduate technical writing.
-They are instructed not to spend endless time copy editing.
-Instead we ask them to spend a limited block of time writing fewer comments that focus on broader issues.
+We ask them to spend a limited block of time writing fewer comments that focus on broader issues.
 Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection.
@@
 
-While this approach is sound in theory, we are uncertain how well TAs can implement the recommended strategy.
+While this approach is well-grounded in research, we are uncertain how well TAs can implement the recommended strategy.
 One of our ongoing questions has been, "what do TAs ACTUALLY spend the most time and effort making comments about as they grade student writing?"

### 2018-11-29 course_notes.Rmd
updates

@@
 ---
-title: "Course Notes"
+title: "Notes & Ideas"
 bibliography: library.bib
@@
 
-You can detail what you learned in the session within these pages. You can also use the Rmd markdown language and insert code chunks with `CMD/CTRL + SHIFT + I`.
-
-# Session 1
+# Session 1 (Oct. 1, 2018)
 Change
 
-# Session 2
+####\
 
-# Session 3
+# Session 2 (Oct. 22, 2018)
 
-# Session 4
+####\
 
-# Session 5
+# Session 3 (Nov. 12, 2018)
 
-# Session 6
+####\
 
-# Session 7
+# Session 4 (Dec. 3, 2018)
 
-# Session 8
+####\
+
+# Session 5 (Jan. 14, 2019)
+
+####\
+
+# Session 6 (Feb. 4, 2019)
+
+####\
+
+# Session 7 (Feb. 25, 2019)
+
+####\
+
+# Session 8 (Mar. 18, 2019)
+
+####\
 
 # References
+
+####\

### 2018-11-29 index.Rmd
updates

@@
 ---
-title: "Data Analysis in R: Research & Learning"
+title: "Analyzing TA Comments on Student Writing in R"
 ---
@@
 
-## So It Begins
+## Background
+I am interested in understanding how students learn technical writing skills in biology, and how we can use data science methods to improve that process.
 
-This is a general landing page for your website for your project. This is just a launching pad for our community. Feel free to customise as much as you would like.
+Our NSF-sponsored project asks if and how scripted instruction, automated support, and holistic feedback accelerates student development as writers.
+One of our specific aims is to see if changing the structure and focus of instructors' comments affects how students' writing develops over time.
+To do this, we need to extract instructor comments from student reports, classify those comments, and correlate the types of comments with student performance.
 
-You can use this page to give introductions about what you are doing and your aims.
+#### \   
+## Prior Work
+In Summer 2018, we extracted and tabulated ~11,000 TA comments from ~500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses collected during a single semester.
+Using an iteratively generated codebook, all 11,000 comments were tagged in two categories:
 
-The structure of the website can be changed by manipulating the `_site.yml`. Font Awesome icons are available to use. You can check those out [here](https://fontawesome.com/v4.7.0/icons/).
+* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
+* __Structure__: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?
 
-## General Structure
+#### \   
+Counts of the coded TA comments were sorted into a two-axis contingency table. This table provided two pieces of information:
 
-We have provided a rough template for your website
-- **course_notes.Rmd** - Write your notes for the class, try some code, work some exercises and use this as a virtual notepad for the items we review.
-This is a great way to showcase new skills and serve as a reference when you can't quite remember how you did something
+* A snapshot of the most common types of comments TAs made, and 
+* A relative estimate of where TAs are placing their greatest effort.
 
-- **pre_work.Rmd** - Use this document to complete any pre-work assignments.
+#### \   
+## Specific Project Goals
+An existing Unix-based script can extract TA comments and meta-data for each student report then assemble the data into a CSV file.
+Classifying TA comments by hand is impractical for routine use though; 10-12,000 score-able comments are generated EACH SEMESTER.
+An automated classification method is needed both for speed and consistency.
 
-- **project.Rmd** - As you might have guessed, use this section to write up your project
+For my project I will be building an R-based text analysis workflow that:
 
-- **library.bib** - is text file that is available for you to put citations.
-You can then use these citations in the other Rmd documents.
-Also note that if you want to cite a package and want the bibtex you can do this by using the command `citation("packagename")`.
+1.
+**Creates a de-identified dataset.** Required data will be extracted from the master dataset and organized in an anonymized data frame where all names, comments, and confidential metadata are re-keyed to unique random identifiers.
+2.
+**Classifies TA comments according to our existing codebook**.
+Initially I will be evaluating whether terms frequencies can be used to classify comments.
+The goal is to be able to assign instructor comments into our pre-defined Subject and Structure codes with 90% accuracy or greater
+3.
+**Sorts TA comments using unsupervised grouping methods**.
+The codebook for classifying TA comments makes prior assumptions about most relevant features.
+Unsupervised grouping may uncover more robust or informative classification methods.
 
-- **footer.html** - Go ahead and put your name in there!
-
-When you are happy with what you have written for the day, go ahead and make a website out of it by using the following command in the console `rmarkdown::render_site()`
+#### \

### 2018-11-29 questions.rmd
updates

@@
 # R Code
-#### I can write code for Shiny and build pages locally, but they do not build for a website. 
-The build stalls here, with the error message:
-
->label: unnamed-chunk-1 (with options) 
->List of 1  
-> $ echo: logi FALSE  
->Error in selectInput("n_breaks", label = "Number of bins:", choices = c(10,  :   
->  could not find function "selectInput"  
->Calls: <Anonymous> ... handle -> withCallingHandlers -> withVisible -> eval -> eval  
->Execution halted  
->Exited with status 1.
+#### I cannot get numbers on CSV extracted data frames.
 
@@
 
-#### Still not sure I know difference between a fork and a branch. 
-Branch is an alternative version of a single file, but fork is a copy of the original repo in total. Y/N?
+#### Question about fork and branch. 
+When I fork another repo, I get comments saying I am "#" steps ahead of it. Should I clone instead?
 
@@
 Other Git and Git-GUI paths do not work.
+
+####\

### 2018-11-29 course_notes.Rmd
updates

@@
 # Session 3 (Nov. 12, 2018)
+I have 11,000 comments from TAs on student reports. I want to parse them into topic groups more or less automatically, using pre-defined criteria. What are the options?
+
+**Option: Latent Dirichlet Allocation**
+
+Found this:  
+"In text mining, we often have collections of documents, such as blog posts or news articles, that we’d like to divide into natural groups so that we can understand them separately.
+Topic modeling is a method for unsupervised classification of such documents, similar to clustering on numeric data, which finds natural groups of items even when we’re not sure what we’re looking for."
+
+Latent Dirichlet allocation (LDA) is a particularly popular method for fitting a topic model.
+It treats each document as a mixture of topics, and each topic as a mixture of words.
+This allows documents to “overlap” each other in terms of content, rather than being separated into discrete groups, in a way that mirrors typical use of natural language.
+
+	https://www.tidytextmining.com/topicmodeling.html
+
+Challenge is that I have groups **already based on the existing coding book.** The individual groups are already labeled. LDA uses clustering based on its own analysis.
+
+**Option: Comparison Modeling Based on Twitter Strategies**
+
+What if I treated each of the pre-defined sub-sets from primary codebook analysis as if they are tweets from separate users? 
+	https://www.tidytextmining.com/twitter.html
+
+Strategy would be:
+
+*  Un-nest tokens to create a tidy frame and remove stop words using anti-join
+*  Calculate word frequencies for each sub-type
+*  Compare frequencies by calculating log-odds ratios
 

### 2018-11-29 project.Rmd
updates

@@
 # Abstract
+
+####\
 
@@
 
-The resulting sorted table provided us with a relative estimate of TA effort.
+* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
+* __Structure__: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?
 
-## The General Problem
-While the qualitative coding of TA comments is informative, it is unsustainable.
-A single investigator worked for nearly 2 months to develop the codebook then rate each of 11,000 comments.
+The resulting sorted contingency table provided a relative estimate of TA effort.
+
+####\
+
+## The Central Problem
+While qualitative coding of TA comments is informative, it is unsustainable.
+A single investigator needed 2 months to develop the codebook then rate 11,000 comments.
 Even with the codebook, individual coders will require training to achieve any accuracy.
-Given that 10-12,000 comments are the normal output EACH SEMESTER, an automated sorting method is needed.
+Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.
 
-### Specific Problems in the Analysis
-I have 11,000 comments from TAs on student reports. I want to parse them into topic groups more or less automatically, using pre-defined criteria.
-
-#### Approach Option: Latent Dirichlet allocation
-In text mining, we often have collections of documents, such as blog posts or news articles, that we’d like to divide into natural groups so that we can understand them separately.
-Topic modeling is a method for unsupervised classification of such documents, similar to clustering on numeric data, which finds natural groups of items even when we’re not sure what we’re looking for.
-
-Latent Dirichlet allocation (LDA) is a particularly popular method for fitting a topic model.
-It treats each document as a mixture of topics, and each topic as a mixture of words.
-This allows documents to “overlap” each other in terms of content, rather than being separated into discrete groups, in a way that mirrors typical use of natural language.
-
-	https://www.tidytextmining.com/topicmodeling.html
-
-Challenge is that I have groups **already based on the existing coding book.** The individual groups are already labeled. LDA uses clustering based on its own analysis.
-
-#### Approach Option: Comparison Modeling Based on Twitter Strategies
-What if I treated each of the pre-defined sub-sets from primary codebook analysis as if they are tweets from separate users? 
-	https://www.tidytextmining.com/twitter.html
-
-Strategy would be:
-*  Un-nest tokens to create a tidy frame and remove stop words using anti-join
-*  Calculate word frequencies for each sub-type
-*  Compare frequencies by calculating log-odds ratios
+####\
 
@@
 
+####\
+
 ## Data
+
+####\
 
@@
 
+####\
+
 # Results
+
+####\
 
@@
 
+####\
+
 # Conclusion
 
+####\
+
 # References
+
+####\

### 2018-11-30 course_notes.Rmd
updates

@@
 # Session 1 (Oct. 1, 2018)
-Change
 
@@
 # Session 3 (Nov. 12, 2018)
-I have 11,000 comments from TAs on student reports. I want to parse them into topic groups more or less automatically, using pre-defined criteria. What are the options?
+I have 11,000 comments from TAs on student reports. Process for bringing in the tidied data and structuring it for analysis is not done, but is coming along.
+
+I want to sort TA comments into topic groups more or less automatically, using pre-defined criteria. What are the options?
 
@@
 
+The approach is described here:
 	https://www.tidytextmining.com/topicmodeling.html
 
-Challenge is that I have groups **already based on the existing coding book.** The individual groups are already labeled. LDA uses clustering based on its own analysis.
+Challenge is that my groups are **already based on the existing coding book.** The individual groups are already labeled. LDA clusters items based on its own analysis.
 
@@
 
-What if I treated each of the pre-defined sub-sets from primary codebook analysis as if they are tweets from separate users? 
+What if I treated each of the pre-defined sub-sets of comments I created using codebook analysis as if they were tweets from separate users? That approach is described here: 
 	https://www.tidytextmining.com/twitter.html
@@
 
+**Option: Comparison Modeling Via Text Classification**
+
+What if I treated each of the comments as text blocks where pre-defined words are more or less likely to be associated with the category. Jerid developed a Shiny app that does something similar:
+https://francojc.github.io/2015/08/15/generating-annotated-text-in-shiny/
+
 ####\
@@
 # Session 4 (Dec. 3, 2018)
+Looking at Jerid's Intro to Statistical Thinking clarified some issues in my approach. I definitely need to be thinking about this problem more in terms of unsupervised exploration.
+
+That leads to my follow-up question: training with 80% seems excessive.
+If there is a large dataset, is there any cost or benefit to splitting to 33/33/34%, or even 25% x 4?
+The 80/20 and 60/20/20% seem out of balance, but there is probably logic I do not see.
+
+In the exploration phase, is re-sampling such a major concern?
+
+####\
+
+**Confidentiality Problem**
+My initial dataset contains confidential information that needs to be de-identified. How do I convert the data so it can be displayed on a public site?
+
+Most direct solution is to write bridge scripts.
+The versions that actually read my data in private repo.
+The public versions of these scripts would have 1+ empty placeholder spaces in the key commands that read in the data.
+Script documentation would explain how others can modify the bridge script file paths.
+
+I assume that without correct system paths, outside users would not be able to access them. I can provide the de-identified data publicly, without exposing the confidential data.
+
+Any gaps or problems with this approach?
+
+####\
+
+**What Are the Primary Data Keys For Structuring the Dataset?**
+These are values that are present already in the dataset.
+
+*record.id*  
+Column is a semester ID and 5-digit number, like sp18.00001.
+This particular record.id indicates the data line is from Spring 2018 CSV file, and is line 00001 of the data table.
+This should be the unique primary key value for each record.
+Once assigned, it should not change.
+
+*report.id*  
+Format is R\_nnn(...).
+The string is assigned to each entry made in Qualtrics.
+This particular string is assigned when a report is collected by our system.
+It is anonymous, and serves as a unique id for each REPORT document that we collect.
+
+*student*  
+The email address of the SUBMITTING STUDENT.
+Unique to each student, and assigned by WFU.
+The email address connects multiple reports to a SINGLE STUDENT AUTHOR.
+It is NOT confidential though, so needs to be switched out.
+
+*ta*  
+The name of the TA assigning scores. Unique to instructor, and connects multiple reports scored by SINGLE TA. Again it is NOT confidential, so also needs to be switched out.
+
+Consider either:
+
+* Creating lists of non-repeating IDs with S\_nnn(...) = Student name, T\_nnn(...) = TA name. Then assign the names via a "lookup" style function
+* Using a global names generator to build 25K list of randomly generated non-repeating first_last names. In practice, random names are easier to scan for visually.
+
+####\
+
+**Lagniappe**
+Maybe it is just me, but I've been struggling with some of the outside instructional materials becuase they use generic terms in ways that make code structure more confusing.
+
+For example, using either "text" or "word" as a variable_name is confusing if the same word is used as part of commands.
+An example I found of this is when I tried to use the "unnest_tokens" command.
+Look at the language below from one of the exercises; are the terms inside the parentheses variable_names, or modifiers for the command?
+
+```
+(text_tibble) %>% 
+  unnest_tokens(word,text)
+```
+
+ Just adding a numeral to show it is a variable name would help. For example, "text_1" clearly is not a command or modifier term.
+
+Maybe this goes back to developing a habit of adopting one approach (like snake_case, or never using "_" as the first character in a filename) and sticking to it.
+
+>Thank you x^y times over for the Regex101 link! Getting Regex right has been a pain for me! This greatly simplifies the process.
+
+**Miscellaneous To Do**
+
+* Need to review map, cast and bind commands again. Not sticking in my head.
+* Try visualizing the 2-D array of the TA comments as a mosaic plot.
+
+**Marking Lab Reports (random thoughts on a different project)**
+
+Text files for processed lab reports already are structured so we can use them like corpus files with inline meta-data. All lab reports have this as embedded metadata up front:
+
+* Number of lab report in semester: first
+* This is a revision.
+* Date submitted: 2018-11-06 12:17:40
+* Author of this lab report is: *Student's NAME*
+* ID: R_1et5dvwqfl2ZCX1
+* Course number: 214
+* TA: *TA's NAME*
+
+Again, the main change we would need to make is to de-identify the student and TA. In theory I should be able to build some kind of direct replace functions for the TA-name and the Student-name.
 

### 2018-11-30 index.Rmd
updates

@@
 ## Specific Project Goals
-An existing Unix-based script can extract TA comments and meta-data for each student report then assemble the data into a CSV file.
-Classifying TA comments by hand is impractical for routine use though; 10-12,000 score-able comments are generated EACH SEMESTER.
+TA comments and meta-data for each student report can be extracted and compiled into a CSV datafile using an existing Unix script.
+However routinely classifying TA comments by hand is impractical; 10-12,000 comments need to be scored EACH SEMESTER.
 An automated classification method is needed both for speed and consistency.
@@
 1.
-**Creates a de-identified dataset.** Required data will be extracted from the master dataset and organized in an anonymized data frame where all names, comments, and confidential metadata are re-keyed to unique random identifiers.
+**Creates a de-identified dataset.** Required data will be extracted from the master CSV data set and organized in an anonymized data frame where all names, comments, and confidential metadata are re-keyed to unique random identifiers.
 2.
 **Classifies TA comments according to our existing codebook**.
+The goal is to be able to assign instructor comments into our pre-defined Subject and Structure codes with 90% accuracy or greater.
 Initially I will be evaluating whether terms frequencies can be used to classify comments.
-The goal is to be able to assign instructor comments into our pre-defined Subject and Structure codes with 90% accuracy or greater
+Other methods will be evaluated as they come to light.
 3.
 **Sorts TA comments using unsupervised grouping methods**.
-The codebook for classifying TA comments makes prior assumptions about most relevant features.
-Unsupervised grouping may uncover more robust or informative classification methods.
+The codebook for classifying TA comments makes prior assumptions about most relevant text features.
+Unsupervised learning may uncover more robust or informative classification groups.
 

### 2018-11-30 questions.rmd
updates

@@
 
-Ideas?
+Ideas? Detailed notes are in "Data_intake_1.rmd", starting Line 102.
+
+#### Why can I not append the read.csv command?
+This works fine:
+```{r}
+base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18.csv')
+head(base_data)
+```
+
+This fails:
+```{r}
+base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18.csv', header=FALSE)
+head(base_data)
+```
+I get:
+```	
+	Error in read_csv(file = "data/coded_full_comments_dataset_Spring18.csv",  : 
+  unused argument (header = FALSE)
+```
 
@@
 # Github
-#### When I work on Github, I see no easy way to create new file folders.
-
-If I empty a folder, it is deleted automatically. Am I overlooking tools, or is goal to be working locally to create files and use web interface less often?
-
------
-
-#### I am not sure how to clone and blend my OWN repos. 
-I was trying to build a starting point with both data code and web site pieces by cloning the web site repo, then adding the standard data layout folder.
-Could not see an obvious way to clone one of my own, or to blend two into one.
-
------
-
 #### Question about fork and branch. 
@@
 
-#### Having a slight problem with Git on laptop.
-I can use my desktop and create repos to Dropbox or local machine from GitHub. Laptop throws in an extra step that seems to corrupt process.
-
-**SOLVED**. The path to Git was wrong. Needs to be very specific: 
-/usr/local/git/bin/git
-
-Other Git and Git-GUI paths do not work.
-
 ####\

### 2018-11-30 course_notes.Rmd
Update

@@
 # Session 3 (Nov. 12, 2018)
-I have 11,000 comments from TAs on student reports. Process for bringing in the tidied data and structuring it for analysis is not done, but is coming along.
+I have 11,000 comments from TAs on student reports.
+Working out a process for bringing in the tidied data and structuring it for analysis is not finished, but is coming along.
+Some of the challenges and questions I have are outlined in Session 4.
 
-I want to sort TA comments into topic groups more or less automatically, using pre-defined criteria. What are the options?
+While learning the coding, I have been thinking about how to approach the larger problem once data are in a reasonable form.
+I want to sort TA comments into topic groups more or less automatically, using pre-defined criteria.
+What are my options?
+
+####\
 
@@
 
-Found this:  
-"In text mining, we often have collections of documents, such as blog posts or news articles, that we’d like to divide into natural groups so that we can understand them separately.
-Topic modeling is a method for unsupervised classification of such documents, similar to clustering on numeric data, which finds natural groups of items even when we’re not sure what we’re looking for."
+I found this summary, and a description of the approach at
+[https://www.tidytextmining.com/topicmodeling.html]():
 
-Latent Dirichlet allocation (LDA) is a particularly popular method for fitting a topic model.
+>"In text mining, we often have collections of documents, such as blog posts or news articles, that we’d like to divide into natural groups so that we can understand them separately.
+Topic modeling is a method for unsupervised classification of such documents, similar to clustering on numeric data, which finds natural groups of items even when we’re not sure what we’re looking for.
+
+>"Latent Dirichlet allocation (LDA) is a particularly popular method for fitting a topic model.
 It treats each document as a mixture of topics, and each topic as a mixture of words.
-This allows documents to “overlap” each other in terms of content, rather than being separated into discrete groups, in a way that mirrors typical use of natural language.
+This allows documents to “overlap” each other in terms of content, rather than being separated into discrete groups, in a way that mirrors typical use of natural language."
 
-The approach is described here:
-	https://www.tidytextmining.com/topicmodeling.html
+My challenge is that LDA clusters items based on its own analysis. My groups are already sorted based on the **existing coding book.** The individual groups are already labeled.
 
-Challenge is that my groups are **already based on the existing coding book.** The individual groups are already labeled. LDA clusters items based on its own analysis.
+####\
 
@@
 
+####\
+
 **Option: Comparison Modeling Via Text Classification**
@@
 # Session 4 (Dec. 3, 2018)
-Looking at Jerid's Intro to Statistical Thinking clarified some issues in my approach. I definitely need to be thinking about this problem more in terms of unsupervised exploration.
+Looking at Jerid's *Intro to Statistical Thinking* clarified some points in my approach.
+I definitely need to be thinking about this problem more in terms of unsupervised exploration.
+It did raise some follow-up questions:
 
-That leads to my follow-up question: training with 80% seems excessive.
+*  A training set with 80% of the data seems excessive.
 If there is a large dataset, is there any cost or benefit to splitting to 33/33/34%, or even 25% x 4?
-The 80/20 and 60/20/20% seem out of balance, but there is probably logic I do not see.
+The 80/20 and 60/20/20% also seems out of balance, but there is probably logic I do not see.
+*  In the exploration phase, is re-sampling such a major concern?
 
-In the exploration phase, is re-sampling such a major concern?
+**Separate note:** thank you many times over for the Regex101 link! Writing regex correctly has long been a pain for me! This greatly simplifies the process.
 
@@
 
-**Confidentiality Problem**
+####Confidentiality Problem
 My initial dataset contains confidential information that needs to be de-identified. How do I convert the data so it can be displayed on a public site?
@@
 
-**What Are the Primary Data Keys For Structuring the Dataset?**
+####Primary Data Keys For Structuring the Dataset?
 These are values that are present already in the dataset.
@@
 *record.id*  
-Column is a semester ID and 5-digit number, like sp18.00001.
+>Column is a semester ID and 5-digit number, like sp18.00001.
 This particular record.id indicates the data line is from Spring 2018 CSV file, and is line 00001 of the data table.
@@
 *report.id*  
-Format is R\_nnn(...).
+>Format is R\_nnn(...).
 The string is assigned to each entry made in Qualtrics.
@@
 *student*  
-The email address of the SUBMITTING STUDENT.
+>The email address of the SUBMITTING STUDENT.
 Unique to each student, and assigned by WFU.
@@
 *ta*  
-The name of the TA assigning scores. Unique to instructor, and connects multiple reports scored by SINGLE TA. Again it is NOT confidential, so also needs to be switched out.
+>The name of the TA assigning scores. Unique to instructor, and connects multiple reports scored by SINGLE TA. Again it is NOT confidential, so also needs to be switched out.
 
@@
 
-**Lagniappe**
+####Lagniappe
 Maybe it is just me, but I've been struggling with some of the outside instructional materials becuase they use generic terms in ways that make code structure more confusing.
@@
 
->Thank you x^y times over for the Regex101 link! Getting Regex right has been a pain for me! This greatly simplifies the process.
+####\
 
-**Miscellaneous To Do**
-
+####Miscellaneous To Do
 * Need to review map, cast and bind commands again. Not sticking in my head.

### 2018-11-30 course_notes.Rmd
update

@@
 *record.id*
+
 >Column is a semester ID and 5-digit number, like sp18.00001.
@@
 *report.id*
+
 >Format is R\_nnn(...).
@@
 *student*
+
 >The email address of the SUBMITTING STUDENT.
@@
 *ta*
+
 >The name of the TA assigning scores. Unique to instructor, and connects multiple reports scored by SINGLE TA. Again it is NOT confidential, so also needs to be switched out.

### 2018-11-30 course_notes.Rmd
update

@@
 
-**Marking Lab Reports (random thoughts on a different project)**
+####\
 
-Text files for processed lab reports already are structured so we can use them like corpus files with inline meta-data. All lab reports have this as embedded metadata up front:
+####Corpus Creation Using Student Lab Reports
+(Random thoughts on a different project)
+
+Text files for our processed lab reports already are structured so we can process them using Jerid's model for corpus files with inline meta-data.
+
+All lab reports have these metadata embedded at the front end of the document:
 
@@
 
-Again, the main change we would need to make is to de-identify the student and TA. In theory I should be able to build some kind of direct replace functions for the TA-name and the Student-name.
+Again, the we would need to de-identify the student and TA using some kind of firewalled search/replace functions.
 

### 2018-12-16 course_notes.Rmd
update

@@
 ####\
+
+**Option: Naive Bayes Text Classification**
+
+I found this summary, and a description of the approach at
 
@@
 
-*record.id*
-
->Column is a semester ID and 5-digit number, like sp18.00001.
+*record.id*: Column is a semester ID and 5-digit number, like sp18.00001.
 This particular record.id indicates the data line is from Spring 2018 CSV file, and is line 00001 of the data table.
@@
 
-*report.id*
-
->Format is R\_nnn(...).
+*report.id*: Format is R\_nnn(...).
 The string is assigned to each entry made in Qualtrics.
@@
 
-*student*
-
->The email address of the SUBMITTING STUDENT.
+*student*: The email address of the SUBMITTING STUDENT.
 Unique to each student, and assigned by WFU.
@@
 
-*ta*
+*ta*: The name of the TA assigning scores. Unique to instructor, and connects multiple reports scored by SINGLE TA. Again it is NOT confidential, so also needs to be switched out.
 
->The name of the TA assigning scores. Unique to instructor, and connects multiple reports scored by SINGLE TA. Again it is NOT confidential, so also needs to be switched out.
-
-Consider either:
-
-* Creating lists of non-repeating IDs with S\_nnn(...) = Student name, T\_nnn(...) = TA name. Then assign the names via a "lookup" style function
+Strategy we decided is to create ing lists of non-repeating IDs with S\_nnn(...) = Student name, T\_nnn(...) = TA name. Then assign the names via a "mutate" and join by row_number function
 * Using a global names generator to build 25K list of randomly generated non-repeating first_last names. In practice, random names are easier to scan for visually.

### 2018-12-16 draft.Rmd
update

---
title: "Project"
author: "Dan Johnson"
date: "10/26/2018"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```

# Abstract

####\

# Introduction
## Background
Undergraduate introductory STEM courses typically include technical writing training. Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed. Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills.

Technical writing training usually is part of a lab course associated with a lecture class. These lab sections may be be led by graduate teaching assistants or even advanced undergraduates. These instructors may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.

In 2014 we launched an automated data collection process we call SAWHET that collects systematic observations and sample data from student writers and graduate instructors. Our resarch goal is to better understand student and instructor behaviors within the technical writing instructional space. Our practical goal is to develop techniques for teaching technical writing that:

*  Can be implemented at large scale by less experienced teaching assistants;
*  Emphasize student development through holistic coaching;
*  Shift student and instructor attention away from simple copy-correction towards larger writing issues; and
*  Embody other best practices outlined in nearly 40 years of research literature on "writing in the disciplines."

We have a structured database of >5000 laboratory reports written by undergraduate students for introductory biology courses. In addition we have all comments that a TA made on the individual reports, plus associated metadata:

*  Context. Was it first or second report of the semester? Initial submission or revision? What course and topic in the course? What year in college was the author?
*  Instructor information. Which TA provided the feedback? What score was assigned, and what what the rationale behind that score?

We are using multiple text-mining methods to map key features of student writing, as well as connections and correlations between those features. These insights then inform how we teach technical writing and train graduate instructors.

####\

## Context for This Project & Prior Work
Students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead. To that end, we train graduate teaching assistants to use a **global first, specific later** approach to grading undergraduate technical writing. We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues. Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection. As students' writing improves, the TA can focus on smaller issues.

This strategy is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy. One of our ongoing questions has been, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**

In Summer 2018, we began analyzing TAs' comments from student papers. We extracted and tabulated all comments from approximately 500 pairs of reports (initial submissions and revisions) collected during a single semester in 3 different courses. Initial data cleaning involved removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent. In all, the final tabulated dataset included ~11,000 separate comments.

We next used a subset of 100 randomly selected comments to build an initial qualitative codebook for sorting comments into types. The initial codebook was re-tested using 1100 additional randomly selected comments. Using the final codebook, we sorted all 11,000 comments into separate categories (see LINK for complete codebook.)

Coded TA comments were sorted on two axes:

* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
* __Structure__: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?

The sorted contingency table provided us with a relative estimate of TA effort.

####\

## The Central Problem of This Project
Qualitative coding of TA comments is informative but unsustainable. Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort. Even with the codebook, individual coders will require training to achieve any accuracy. Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.

####\

# Literature Review
##General Approach to Initial Dataset
The following excerpt from *Text Mining with R: A Tidy Approach* (Julia Silge and David Robinson) summarizes how extended text (in this case, TA comments) can be organized as a tidy matrix.*All emphases are mine.*

>We thus define the tidy text format as being a table with one-token-per-row. A token is a **meaningful unit of text**, such as a word, that we are interested in using for analysis, and tokenization is the process of splitting text into tokens.

>This one-token-per-row structure is in contrast to the ways text is often stored in current analyses, perhaps as strings or in a document-term matrix. For tidy text mining, the token that is stored in each row is most often a single word, **but can also be an n-gram, sentence, or paragraph** . In the tidytext package, we provide functionality to tokenize by commonly used units of text like these and convert to a one-term-per-row format.

Using tidy format (vs. other common text data formats) better maintains relationships between individual comments and metadata, which simplifies subsequent analysis.

####\

# Methods

####\

## Preliminary Analysis
###Log-Odds Text Comparison

####Comparing Pearson Correlations of Word Frequencies

####Calculating and Plotting Log-Odds Pairs

###Direct Comparison Using Only Words Appearing in All Subsets of Comments

###Other Analysis Approaches

####\

## Data

####\

## Analysis

####\

# Results

####\

# Discussion

####\

# Conclusion

####\

# References

####\

### 2018-12-16 project.Rmd
Update
update

@@
 These lab sections may be be led by graduate teaching assistants or even advanced undergraduates.
-These class leaders may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.
+These instructors may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.
 
-We have been evaluating standardized approaches and techniques for teaching technical writing that:
+In 2014 we launched an automated data collection process we call SAWHET that collects systematic observations and sample data from student writers and graduate instructors.
+Our resarch goal is to better understand student and instructor behaviors within the technical writing instructional space.
+Our practical goal is to develop techniques for teaching technical writing that:
 
-*  Embody best practices from nearly 40 years of "writing in the disciplines"" literature
-*  Can be implemented at large scale by less experienced teaching assistants
-*  Emphasize student development through holistic coaching 
-*  Focus student and instructor attention on larger issues rather than simple copy-correction.
+*  Can be implemented at large scale by less experienced teaching assistants;
+*  Emphasize student development through holistic coaching;
+*  Shift student and instructor attention away from simple copy-correction towards larger writing issues; and
+*  Embody other best practices outlined in nearly 40 years of research literature on "writing in the disciplines."
 
-We have amassed a structured database of laboratory reports written by undergraduate students for introductory biology courses. Besides the original texts we have collected these meta-data:
+We have a structured database of >5000 laboratory reports written by undergraduate students for introductory biology courses.
+In addition we have all comments that a TA made on the individual reports, plus associated metadata:
 
-*  Circumstances: first or second report of semester; initial submission or revision; course; topic; year of student author in college
-*  Instructor information: identity of TA providing feedback; score assigned; rationale behind score
-*  All comments that a TA made on the individual reports they graded.
+*  Context. Was it first or second report of the semester? Initial submission or revision? What course and topic in the course? What year in college was the author?
+*  Instructor information. Which TA provided the feedback? What score was assigned, and what what the rationale behind that score?
 
-We are using text-mining approaches to map key features of student writing, as well as connections and correlations between those features, that can inform how we teach technical writing.
+We are using multiple text-mining methods to map key features of student writing, as well as connections and correlations between those features.
+These insights then inform how we teach technical writing and train graduate instructors.
 
-## Context for This Project
-Students make greater gains as writers when instructors avoid just telling students what to correct ("copy editing"), and adopt a coaching- oriented approach instead.
-To that end, we train graduate teaching assistants use a **global first, specific later** approach to grading undergraduate technical writing.
-We ask them to spend a limited block of time writing fewer comments that focus on broader issues.
+####\
+
+## Context for This Project & Prior Work
+Students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead.
+To that end, we train graduate teaching assistants to use a **global first, specific later** approach to grading undergraduate technical writing.
+We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues.
 Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection.
@@
 
-While this approach is well-grounded in research, we are uncertain how well TAs can implement the recommended strategy.
-One of our ongoing questions has been, "what do TAs ACTUALLY spend the most time and effort making comments about as they grade student writing?"
+This strategy is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy.
+One of our ongoing questions has been, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**
 
-In Summer 2018, we began trying to answer this question by analyzing TAs' comments from student papers.
-We extracted and tabulated all TA comments from approximately 500 pairs of reports (initial submissions and revisions) in 3 different courses collected during a single semester.
-Initial data cleaning involved removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging the intent.
+In Summer 2018, we began analyzing TAs' comments from student papers.
+We extracted and tabulated all comments from approximately 500 pairs of reports (initial submissions and revisions) collected during a single semester in 3 different courses.
+Initial data cleaning involved removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent.
 In all, the final tabulated dataset included ~11,000 separate comments.
@@
 
-The resulting sorted contingency table provided a relative estimate of TA effort.
+The sorted contingency table provided us with a relative estimate of TA effort.
 
@@
 
-## The Central Problem
-While qualitative coding of TA comments is informative, it is unsustainable.
-A single investigator needed 2 months to develop the codebook then rate 11,000 comments.
+## The Central Problem of This Project
+Qualitative coding of TA comments is informative but unsustainable.
+Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort.
 Even with the codebook, individual coders will require training to achieve any accuracy.
@@
 # Literature Review
+##General Approach to Initial Dataset
+The following excerpt from *Text Mining with R: A Tidy Approach* (Julia Silge and David Robinson) summarizes how extended text (in this case, TA comments) can be organized as a tidy matrix.*All emphases are mine.*
+
+>We thus define the tidy text format as being a table with one-token-per-row.
+A token is a **meaningful unit of text**, such as a word, that we are interested in using for analysis, and tokenization is the process of splitting text into tokens.
+
+>This one-token-per-row structure is in contrast to the ways text is often stored in current analyses, perhaps as strings or in a document-term matrix.
+For tidy text mining, the token that is stored in each row is most often a single word, **but can also be an n-gram, sentence, or paragraph** .
+In the tidytext package, we provide functionality to tokenize by commonly used units of text like these and convert to a one-term-per-row format.
+
+Using tidy format (vs. other common text data formats) better maintains relationships between individual comments and metadata, which simplifies subsequent analysis.
+
+####\
 
 # Methods
-## Preliminary Proof of Concept
+Our initial working dataset is an anonymized Excel table with one column of "TA comments" extracted from student lab reports.
+Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation.
+Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.
+
+####\
+
+##Data Pre-Checks
+A CSV file named "coded_full_comments_dataset_Spring18anon.csv" should have correctly organized text and meta-data. Prior to import, review data table in Excel. Check:
+
+1.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.md"
+2.  Data have been de-identified and recoded using pre-assigned anonymous IDs.
+3.  Data table headers are in snake_case, and match tabulated list below. The same list is in the file "code_column_names.txt"
+
+Column Number|Column Name
+------------|----------------------
+1|unique.record
+2|report.id
+3|sort
+4|report.title (empty)
+5|student (Std\_nnn)
+6|course
+7|ta (TA\_nnn)
+8|lab
+9|tag
+10|type.TA
+11|grade.TA
+12|grading.time
+13|Rank
+14|hypothesis.ok
+15|data.ok
+16|citation.ok
+17|interpretation.ok
+18|organization.ok
+19|techflaws.ok
+20|writing.ok
+21|comments.incorporated
+22|ta.comment
+23|code.subject
+24|code.structure
+25|code.locus
+26|code.scope
+27|code.tone
+28|code.notes
+
+Once all data are in correct format, export Excel file to CSV with name formatted as **"coded_full_comments_dataset_SemYear.csv"**
+
+####\
+
+##Initial Intake and Post-Check
+```{r Libraries}
+# Call in tidyverse, other required libraries.
+library(tidyverse)
+library(tidytext)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(scales)
+```
+
+```{r}
+#This block reads in the full CSV into a starting dataframe named "base_data".
+#Subsequent analyses call in subsets of "base_data".
+base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
+```
+
+If the TA comments dataset is imported correctly, there will be a dataframe of ~11,000 rows and 28 columns named as listed in the **Data Pre-Checks** table above.
+
+Given some data points may be entered incorrectly by instructors or students, I included several post-import data checks to ensure data have been properly coded and columns contain valid entries only.
+
+```{r}
+#"Course" column should only have entries that correspond to 113, 114, 214, NA. 
+unique(base_data$course)
+```
+
+```{r}
+#Entries in "TAs" column should match the anonymous IDs that correspond to the TAs assigned to the relevant courses in the semester being analyzed. 
+#If there are incorrect or extra anonymous IDs in this list, the mostly likely source of the error is improper de-identification of the Excel file prior to import.
+unique(base_data$ta)
+```
+
+```{r}
+#"Lab" should only have entries that match the limited keywords in codebook.
+unique(base_data$lab)
+```
+
+```{r}
+#Look at the NUMBER of unique anonymous student IDs. Should be within 10% of (but not more than) the combined enrollment of the 3 target courses for the semester.
+unique(base_data$student)
+```
+
+```{r}
+#Check that the list of "code.subject" topics here matches allowed terms in codebook.
+unique(base_data$code.subject)
+```
+
+```{r}
+#Check that the list of "code.structure" terms extracted from the dataset and displayed here matches allowed terms in codebook.
+unique(base_data$code.structure)
+```
+
+####\
+
+## Preliminary Analysis
+The first question prior to deeper comparisons is **are the comments obviously different between groups?** Before conducting detailed analyses I tried several iterations of different visual comparison strategies.
+
+####\
+
+###Log-Odds Text Comparison
+The log-odds ratios method outlined below has been the most informative to date.
+The basic premise being tested is that certain words are more frequent in TA comments belonging to one sub-category, and less frequent in others.
+I first saw this approach as a way to compare word usage between separate Twitter accounts here: https://www.tidytextmining.com/twitter.html.
+
+For simplicity I reduced the analysis to just the four most common comment subjects.
+
+```{r}
+# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
+
+frequency_writing <- filter(base_data, code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
+frequency_writing.subcolumns <- frequency_writing %>% select(1, 22:23)
+```
+
+```{r}
+# Tokenize phrases, remove stop words listed in standard reference file.
+base_data_tokenized <- frequency_writing.subcolumns %>% 
+  unnest_tokens(word,ta.comment) %>% 
+  anti_join(stop_words)
+
+#At this point the dataframe should have "Unique.Record" in column 1, "code.subject" in column 2, and the unnested "word" in column 3.
+All columns are class = "character".
+Stop words listed in the "stop_words" R reference data file have been removed.
+
+```
+
+####Comparing Pearson Correlations of Word Frequencies
+If the words used by TAs in their comments differ between the categories, there should be fairly low correlations overall .
+```{r}
+#Block below re-organizes the data for comparisons. This pattern sets "2. Writing Quality" as the first dataset, then all other data in remaining columns.
+#Groups by code.subject, calculated proportional frequency in each group.
+#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
+#Stats need this format to be able to compare 2. WQ against other values.
+base_data_tokenized_sorted2 <- base_data_tokenized %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  mutate(proportion = n / sum(n)) %>% 
+  select(-n) %>% 
+  spread(code.subject, proportion) %>% 
+  gather(code.subject, proportion, `3. Technical and Scientific`:`4. Logic and Thinking`)
+
+#The dataframe called "base_data_tokenized_sorted2" is the final 4-column dataframe for statistical analysis.
+Individual words are in column 1, "2.
+Writing Quality" frequency is in Column 2, "code.subject" is in column 3, and "proportion" is in column 4.
+Columns 1 and 3 are class = "character", 2 and 4 are numeric ranges.
+
+#This structure is needed in order to compare word (in Column 1) frequencies for 2.
+Writing Quality (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes.
+WHICH subset of X values to use is coded in Column 3.
+```
+
+Calculating Pearson correlation between the frequency of words in Writing Quality versus the other two groups.
+
+```{r}
+cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3. Technical and Scientific",], 
+         ~ proportion + `2. Writing Quality`)
+```
+
+```{r}
+cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4. Logic and Thinking",], 
+         ~ proportion + `2. Writing Quality`)
+```
+
+####START OF BROKEN SECTION
+THERE IS AN ERROR HERE I CANNOT DEBUG!
+If I change "3.
+Technical and Scientific"  to "2.
+Writing Quality", the structure of the tibble breaks.
+It goes from the correctly sorted 4 columns to incorrectly sorted 3 columns.
+It is NOT A TYPO; I can literally copy/paste it in and out and break/correct the code.
+
+I cannot use a different order, or any other format. I am at a loss.
+
+```{}
+# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
+frequency_writing3 <- filter(base_data, code.subject=="3. Technical and Scientific"|code.subject=="2. Writing Quality"|code.subject=="4. Logic and Thinking")
+frequency_writing3.subcolumns <- frequency_writing3 %>% select(1, 22:23)
+```
+
+```{}
+# Tokenize phrases, remove stop words listed in standard reference file.
+base_data_tokenized3 <- frequency_writing3.subcolumns %>% 
+  unnest_tokens(word,ta.comment) %>% 
+  anti_join(stop_words)
+```
+
+```{}
+#Block below re-organizes the data for comparisons. This pattern sets "3. Writing Quality" as the first dataset, then all other data in remaining columns.
+#Groups by code.subject, calculated proportional frequency in each group.
+#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
+#Stats need this format to be able to compare 2. WQ against other values.
+base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  mutate(proportion = n / sum(n)) %>% 
+  select(-n) %>% 
+  spread(code.subject, proportion) %>% 
+  gather(code.subject, proportion, `2. Writing Quality`:`4. Logic and Thinking`)
+```
+
+```{}
+#Create groups by code.subject, calculates proportional frequency in each group.
+#Final step creates 5 columns: code.subject, n, word, total, frequency.
+base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  left_join(base_data_tokenized %>% 
+              group_by(code.subject) %>% 
+              summarise(total = n())) %>%
+  mutate(freq = n/total)
+```
+
+Re-organize the data for comparisons. This pattern sets 3. Technical and Scientific as the first dataset, then all other data in remaining columns.
+```{}
+#Groups by code.subject, calculated proportional frequency in each group.
+#Final step creates 4 columns: word, 3. Technical and Scientific, code.subject, proportion
+#Stats need this format to be able to compare 3. TS against other values.
+base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  mutate(proportion = n / sum(n)) %>% 
+  select(-n) %>% 
+  spread(code.subject, proportion) %>% 
+  gather(code.subject, proportion, `2. Writing Quality`:`4. Logic and Thinking`)
+
+#END OF BROKEN SECTION
+```
+
+***
+
+START OF SECTION I CANNOT COPY AND MODIFY. This section depends on the ability to calculate tables using the code above. So I cannot run this part either.
+
+```{r}
+#Create groups by code.subject, calculates proportional frequency in each group.
+#Final step creates 5 columns: code.subject, n, word, total, frequency.
+base_data_tokenized_sorted2 <- base_data_tokenized %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  left_join(base_data_tokenized %>% 
+  group_by(code.subject) %>% 
+  summarise(total = n())) %>%
+  mutate(freq = n/total)
+```
+
+```{r}
+base_data_tokenized_sorted2v3 <- base_data_tokenized_sorted2 %>% 
+  select(code.subject, word, freq) %>% 
+  spread(code.subject, freq) %>%
+  arrange(`2. Writing Quality`,`3. Technical and Scientific`)
+```
+
+```{r}
+ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`3. Technical and Scientific`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+  ggsave("Freqplot_3vs2.png", width = 10, height = 6)
+```
+
+```{r}
+#I can get part way around the problem by re-using the same data and calling a different subset for the Y axis.
+Want to add comparisons of 3 vs 4, but same tables cannot be created because dataframe keeps breaking when I try to change 2.
+WQ.
+
+base_data_tokenized_sorted2v4 <- base_data_tokenized_sorted2 %>% 
+  select(code.subject, word, freq) %>% 
+  spread(code.subject, freq) %>%
+  arrange(`2. Writing Quality`,`4. Logic and Thinking`)
+```
+
+```{r}
+ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`4. Logic and Thinking`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+  ggsave("Freqplot_4vs2.png", width = 10, height = 6)
+```
+
+END of broken / limping section
+
+####Calculating and Plotting Log-Odds Pairs
+**2. Writing versus 3. Technical**
+
+```{r}
+word_ratios2v3 <- base_data_tokenized %>%
+  count(word, code.subject) %>%
+  group_by(word) %>%
+  filter(sum(n) >= 10) %>%
+  ungroup() %>%
+  spread(code.subject, n, fill = 0) %>%
+  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
+  mutate(logratio = log(`2. Writing Quality`/`3. Technical and Scientific`)) %>%
+  arrange(desc(logratio))
+```
+
+```{r}
+word_ratios2v3 %>% 
+  arrange(abs(logratio))
+```
+
+```{r}
+word_ratios2v3 %>%
+  group_by(logratio < 0) %>%
+  top_n(20, abs(logratio)) %>%
+  ungroup() %>%
+  mutate(word = reorder(word, logratio)) %>%
+  ggplot(aes(word, logratio, fill = logratio < 0)) +
+  geom_col(show.legend = FALSE) +
+  coord_flip() +
+  ylab("log odds ratio (Writing-top/Technical-bottom)") +
+  scale_fill_discrete(name = "", labels = c("2. Writing Quality", "3. Technical and Scientific"))
+  #ggsave("Logplot_2vs3.png", width = 6, height = 10)
+
+```
+
+**3. Technical vs. 4. Logic**
+
+```{r}
+word_ratios3v4 <- base_data_tokenized %>%
+  count(word, code.subject) %>%
+  group_by(word) %>%
+  filter(sum(n) >= 10) %>%
+  ungroup() %>%
+  spread(code.subject, n, fill = 0) %>%
+  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
+  mutate(logratio = log(`3. Technical and Scientific`/`4. Logic and Thinking`)) %>%
+  arrange(desc(logratio))
+```
+
+```{r}
+word_ratios3v4 %>% 
+  arrange(abs(logratio))
+```
+
+```{r}
+word_ratios3v4 %>%
+  group_by(logratio < 0) %>%
+  top_n(20, abs(logratio)) %>%
+  ungroup() %>%
+  mutate(word = reorder(word, logratio)) %>%
+  ggplot(aes(word, logratio, fill = logratio < 0)) +
+  geom_col(show.legend = FALSE) +
+  coord_flip() +
+  ylab("log odds ratio (Technical-top/Logic-bottom)") +
+  scale_fill_discrete(name = "", labels = c("3. Technical and Scientific","4. Logic and Thinking"))
+  #ggsave("Logplot_3vs4.png", width = 6, height = 10)
+
+```
+
+**2. Writing versus 4. Logic**
+
+```{r}
+word_ratios2v4 <- base_data_tokenized %>%
+  count(word, code.subject) %>%
+  group_by(word) %>%
+  filter(sum(n) >= 10) %>%
+  ungroup() %>%
+  spread(code.subject, n, fill = 0) %>%
+  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
+  mutate(logratio = log(`2. Writing Quality`/`4. Logic and Thinking`)) %>%
+  arrange(desc(logratio))
+```
+
+```{r}
+word_ratios2v4 %>% 
+  arrange(abs(logratio))
+```
+
+```{r}
+word_ratios2v4 %>%
+  group_by(logratio < 0) %>%
+  top_n(20, abs(logratio)) %>%
+  ungroup() %>%
+  mutate(word = reorder(word, logratio)) %>%
+  ggplot(aes(word, logratio, fill = logratio < 0)) +
+  geom_col(show.legend = FALSE) +
+  coord_flip() +
+  ylab("log odds ratio (Writing-top/Logic-bottom)") +
+  scale_fill_discrete(name = "", labels = c("2. Writing Quality", "4. Logic and Thinking"))
+  #ggsave("Logplot_2vs4.png", width = 6, height = 10)
+
+```
+
+##Direct Comparison Using Only Words Appearing in All Subsets of Comments
+I think this is less informative, but it does let me see the data more clearly.Basic strategy is to take all comments from base_data, unnest tokens, filter stopwords, count and sort remaining terms, then write back to a new datatable called 'sortedwords.comments.all'.
+Calculate TOTAL words and store.
+Then calculate fractional frequencies for each word and store using mutate.
+
+This is an early version, so likely has some poorly developed code blocks that could be refactored.
+
+```{r}
+library(dplyr)
+library(tidytext)
+
+sortedwords.comments.all <- base_data_tokenized %>%
+  filter(!word %in% stop_words$word,
+         !word %in% str_remove_all(stop_words$word, "'"),
+         str_detect(word, "[a-z]")) %>%
+  count(word, word, sort = TRUE) %>%
+  ungroup()
+
+#Calculate the number of words in total in the subset
+total_words.all <- sortedwords.comments.all %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
+  all.fraction = (n*100)/total_words.all$total
+  )
+```
+
+##Summarize Word Frequencies in Subsets of Comments Based on 'Subject'
+Take subsets of comments from base_data_tokenized using code.subject as separating characteristic.
+
+###Other Analysis Approaches
+This approach summarizes the word frequencies in the FULL set of TA comments first, then breaks them down into sub-groups. It has examples of other code bits as well.
+
+The first block takes all comments from base_data, unnests tokens, filters stopwords, count and sort remaining terms, then writes results back to a new dataframe called 'sortedwords.comments.all'.
+
+TOTAL words is calculated and stored. Then fractional frequencies for each word are calculated and stored back to the dataframe using mutate.
+
+```{r}
+# Tokenize phrases, remove stop words listed in standard reference file.
+base_data_tokenized <- base_data %>% 
+  unnest_tokens(word,ta.comment)
+
+# An alternative approach to pulling out stopwords  
+sortedwords.comments.all <- base_data_tokenized %>%
+  filter(!word %in% stop_words$word,
+         !word %in% str_remove_all(stop_words$word, "'"),
+         str_detect(word, "[a-z]")) %>%
+  count(word, word, sort = TRUE) %>%
+  ungroup()
+
+#Calculate the number of words in total in the subset
+total_words.all <- sortedwords.comments.all %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
+  all.fraction = (n*100)/total_words.all$total
+  )
+```
+
+```{r}
+#Pull SUBSETs of tokenized data based on code.subject: 
+base_data_tokenized.basic <- subset(base_data_tokenized, code.subject == "1. Basic Criteria", select = (1:28))
+base_data_tokenized.writing <- subset(base_data_tokenized, code.subject == "2. Writing Quality", select = (1:28))
+base_data_tokenized.technical <- subset(base_data_tokenized, code.subject == "3. Technical and Scientific", select = (1:28))
+base_data_tokenized.logic <- subset(base_data_tokenized, code.subject == "4. Logic and Thinking", select = (1:28))
+base_data_tokenized.praise <- subset(base_data_tokenized, code.subject == "5. Praise or Concern", select = (1:28))
+base_data_tokenized.misconduct <- subset(base_data_tokenized, code.subject == "6. Misconduct", select = (1:28))
+base_data_tokenized.narrative <- subset(base_data_tokenized, code.subject == "12. Narrative Comments", select = (1:28))
+base_data_tokenized.sciname <- subset(base_data_tokenized, code.subject == "Scientific Name", select = (1:28))
+```
+
+For each subset, filter stopwords, count words and total, then calculate frequencies.
+
+```{r}
+sortedwords.basic <- base_data_tokenized.basic %>%
+  filter(!word %in% stop_words$word,
+         !word %in% str_remove_all(stop_words$word, "'"),
+         str_detect(word, "[a-z]")) %>%
+  count(word, word, sort = TRUE) %>%
+  ungroup()
+
+#Calculate the number of words in total in the subset
+total_words.basic <- sortedwords.basic %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.basic.fraction <- sortedwords.basic %>% as_tibble() %>% mutate(
+  basic.fraction = (n*100)/total_words.basic$total
+)
+```
+
+```{r}
+sortedwords.writing <- base_data_tokenized.writing %>%
+  filter(!word %in% stop_words$word,
+         !word %in% str_remove_all(stop_words$word, "'"),
+         str_detect(word, "[a-z]")) %>%
+  count(word, word, sort = TRUE) %>%
+  ungroup()
+
+#Calculate the number of words in total in the subset
+total_words.writing <- sortedwords.writing %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.writing.fraction <- sortedwords.writing %>% as_tibble() %>% mutate(
+  writing.fraction = (n*100)/total_words.writing$total
+)
+```
+
+```{r}
+sortedwords.technical <- base_data_tokenized.technical %>%
+  filter(!word %in% stop_words$word,
+         !word %in% str_remove_all(stop_words$word, "'"),
+         str_detect(word, "[a-z]")) %>%
+  count(word, word, sort = TRUE) %>%
+  ungroup()
+
+#Calculate the number of words in total in the subset
+total_words.technical <- sortedwords.technical %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.technical.fraction <- sortedwords.technical %>% as_tibble() %>% mutate(
+  technical.fraction = (n*100)/total_words.technical$total
+)
+```
+
+```{r}
+sortedwords.logic <- base_data_tokenized.logic %>%
+  filter(!word %in% stop_words$word,
+         !word %in% str_remove_all(stop_words$word, "'"),
+         str_detect(word, "[a-z]")) %>%
+  count(word, word, sort = TRUE) %>%
+  ungroup()
+
+#Calculate the number of words in total in the subset
+total_words.logic <- sortedwords.logic %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.logic.fraction <- sortedwords.logic %>% as_tibble() %>% mutate(
+  logic.fraction = (n*100)/total_words.logic$total
+)
+```
+
+Next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed.
+
+**WARNING**: Words that are MISSING from **any subset** cause the full set to be pruned. I am losing the unique words this way. **Need to figure out how to join without pruning.**
+
+```{r}
+total <- merge(sortedwords.comments.all.fraction,sortedwords.technical.fraction,by="word") 
+total2 <- merge(total,sortedwords.writing.fraction,by="word")
+total3 <- merge(total2,sortedwords.logic.fraction,by="word")
+#rename lines that were possibly confusing
+names(total3)[2] <- "all.count"
+names(total3)[4] <- "technical.count"
+names(total3)[6] <- "writing.count"
+names(total3)[8] <- "logic.count"
+
+#Mutate table to calculate then append fractional word frequency in the text.
+total4 <- total3 %>% as_tibble() %>% mutate(
+  writing.wt = (writing.count *100)/all.count
+)
+
+total5 <- total4 %>% as_tibble() %>% mutate(
+  technical.wt = (technical.count *100)/all.count
+)
+
+total6 <- total5 %>% as_tibble() %>% mutate(
+  logic.wt = (logic.count *100)/all.count
+)
+
+```
+
+The following six graphs only compare frequencies for ~770 words that appear in ALL THREE of the sub-categories.
+Graphs 1-3 compare frequencies in individual groups against word frequency overall.
+Generally, the frequency of a term in the whole set of comments is similar to frequency in an individual group (X vs Y positions similar.) However, some terms associated with Logic subgroup clearly fall below the trendline Graph 3, suggesting they are less frequent in that sub-category than overall.
+
+```{r}
+ggplot(total6, aes(`all.fraction`,`writing.fraction`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+```
+
+```{r}
+ggplot(total6, aes(`all.fraction`,`technical.fraction`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+```
+
+```{r}
+ggplot(total6, aes(`all.fraction`,`logic.fraction`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+```
+
+Graphs 4-6 are pair-wise comparisons of frequencies for sub-categories.
+Unlike the comparisons to overall frequency, most words are NOT clustered on the trendline, suggesting there are differences in word usage.
+
+```{r}
+ggplot(total6, aes(`technical.fraction`,`writing.fraction`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+```
+
+```{r}
+ggplot(total6, aes(`logic.fraction`,`writing.fraction`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+```
+
+```{r}
+ggplot(total6, aes(`logic.fraction`,`technical.fraction`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+```
 

### 2018-12-16 project.Rmd
update

--- 2018-11-29 project.Rmd
@@
 # Literature Review
+##General Approach to Initial Dataset
+The following excerpt from *Text Mining with R: A Tidy Approach* (Julia Silge and David Robinson) summarizes how extended text (in this case, TA comments) can be organized as a tidy matrix.*All emphases are mine.*
+
+>We thus define the tidy text format as being a table with one-token-per-row.
+A token is a **meaningful unit of text**, such as a word, that we are interested in using for analysis, and tokenization is the process of splitting text into tokens.
+
+>This one-token-per-row structure is in contrast to the ways text is often stored in current analyses, perhaps as strings or in a document-term matrix.
+For tidy text mining, the token that is stored in each row is most often a single word, **but can also be an n-gram, sentence, or paragraph** .
+In the tidytext package, we provide functionality to tokenize by commonly used units of text like these and convert to a one-term-per-row format.
+
+Using tidy format (vs. other common text data formats) better maintains relationships between individual comments and metadata, which simplifies subsequent analysis.
+
+####\
 
 # Methods
-## Preliminary Proof of Concept
 

### 2018-12-16 project_backup.Rmd
update

---
title: "Project"
author: "Dan Johnson"
date: "10/26/2018"
output: html_document
bibliography: library.bib
biblio-style: apalike
---
```{r, child="_setup.Rmd"}
```

# Abstract

####\

# Introduction
## Background
Undergraduate introductory STEM courses typically include technical writing training. Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed. Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills.

Technical writing training usually is part of a lab course associated with a lecture class. These lab sections may be be led by graduate teaching assistants or even advanced undergraduates. These instructors may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.

In 2014 we launched an automated data collection process we call SAWHET that collects systematic observations and sample data from student writers and graduate instructors. Our resarch goal is to better understand student and instructor behaviors within the technical writing instructional space. Our practical goal is to develop techniques for teaching technical writing that:

*  Can be implemented at large scale by less experienced teaching assistants;
*  Emphasize student development through holistic coaching;
*  Shift student and instructor attention away from simple copy-correction towards larger writing issues; and
*  Embody other best practices outlined in nearly 40 years of research literature on "writing in the disciplines."

We have a structured database of >5000 laboratory reports written by undergraduate students for introductory biology courses. In addition we have all comments that a TA made on the individual reports, plus associated metadata:

*  Context. Was it first or second report of the semester? Initial submission or revision? What course and topic in the course? What year in college was the author?
*  Instructor information. Which TA provided the feedback? What score was assigned, and what what the rationale behind that score?

We are using multiple text-mining methods to map key features of student writing, as well as connections and correlations between those features. These insights then inform how we teach technical writing and train graduate instructors.

####\

## Context for This Project & Prior Work
Students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead. To that end, we train graduate teaching assistants to use a **global first, specific later** approach to grading undergraduate technical writing. We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues. Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection. As students' writing improves, the TA can focus on smaller issues.

This strategy is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy. One of our ongoing questions has been, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**

In Summer 2018, we began analyzing TAs' comments from student papers. We extracted and tabulated all comments from approximately 500 pairs of reports (initial submissions and revisions) collected during a single semester in 3 different courses. Initial data cleaning involved removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent. In all, the final tabulated dataset included ~11,000 separate comments.

We next used a subset of 100 randomly selected comments to build an initial qualitative codebook for sorting comments into types. The initial codebook was re-tested using 1100 additional randomly selected comments. Using the final codebook, we sorted all 11,000 comments into separate categories (see LINK for complete codebook.)

Coded TA comments were sorted on two axes:

* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
* __Structure__: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?

The sorted contingency table provided us with a relative estimate of TA effort.

####\

## The Central Problem of This Project
Qualitative coding of TA comments is informative but unsustainable. Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort. Even with the codebook, individual coders will require training to achieve any accuracy. Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.

####\

# Literature Review
##General Approach to Initial Dataset
The following excerpt from *Text Mining with R: A Tidy Approach* (Julia Silge and David Robinson) summarizes how extended text (in this case, TA comments) can be organized as a tidy matrix.*All emphases are mine.*

>We thus define the tidy text format as being a table with one-token-per-row. A token is a **meaningful unit of text**, such as a word, that we are interested in using for analysis, and tokenization is the process of splitting text into tokens. 
>This one-token-per-row structure is in contrast to the ways text is often stored in current analyses, perhaps as strings or in a document-term matrix. For tidy text mining, the token that is stored in each row is most often a single word, **but can also be an n-gram, sentence, or paragraph** . In the tidytext package, we provide functionality to tokenize by commonly used units of text like these and convert to a one-term-per-row format.

Using tidy format (vs. other common text data formats) better maintains relationships between individual comments and metadata, which simplifies subsequent analysis.

####\

# Methods
Our initial working dataset is an anonymized Excel table with one column of "TA comments" extracted from student lab reports. Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation. Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.

####\

##Data Pre-Checks
A CSV file named "coded_full_comments_dataset_Spring18anon.csv" should have correctly organized text and meta-data. Prior to import, review data table in Excel. Check:

1.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.md"
2.  Data have been de-identified and recoded using pre-assigned anonymous IDs.
3.  Data table headers are in snake_case, and match tabulated list below. The same list is in the file "code_column_names.txt"

Column Number|Column Name
------------|----------------------
1|unique.record
2|report.id
3|sort
4|report.title (empty)
5|student (Std\_nnn)
6|course
7|ta (TA\_nnn)
8|lab
9|tag
10|type.TA
11|grade.TA
12|grading.time
13|Rank
14|hypothesis.ok
15|data.ok
16|citation.ok
17|interpretation.ok
18|organization.ok
19|techflaws.ok
20|writing.ok
21|comments.incorporated
22|ta.comment
23|code.subject
24|code.structure
25|code.locus
26|code.scope
27|code.tone
28|code.notes

Once all data are in correct format, export Excel file to CSV with name formatted as **"coded_full_comments_dataset_SemYear.csv"**

####\

##Initial Intake and Post-Check
```{r Libraries}
# Call in tidyverse, other required libraries.
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
```

```{r}
#This block reads in the full CSV into a starting dataframe named "base_data".
#Subsequent analyses call in subsets of "base_data".
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
```

If the TA comments dataset is imported correctly, there will be a dataframe of ~11,000 rows and 28 columns named as listed in the **Data Pre-Checks** table above.

Given some data points may be entered incorrectly by instructors or students, I included several post-import data checks to ensure data have been properly coded and columns contain valid entries only.

```{r}
#"Course" column should only have entries that correspond to 113, 114, 214, NA. 
unique(base_data$course)
```

```{r}
#Entries in "TAs" column should match the anonymous IDs that correspond to the TAs assigned to the relevant courses in the semester being analyzed. 
#If there are incorrect or extra anonymous IDs in this list, the mostly likely source of the error is improper de-identification of the Excel file prior to import.
unique(base_data$ta)
```

```{r}
#"Lab" should only have entries that match the limited keywords in codebook.
unique(base_data$lab)
```

```{r}
#Look at the NUMBER of unique anonymous student IDs. Should be within 10% of (but not more than) the combined enrollment of the 3 target courses for the semester.
unique(base_data$student)
```

```{r}
#Check that the list of "code.subject" topics here matches allowed terms in codebook.
unique(base_data$code.subject)
```

```{r}
#Check that the list of "code.structure" terms extracted from the dataset and displayed here matches allowed terms in codebook.
unique(base_data$code.structure)
```

####\

## Preliminary Analysis
The first question prior to deeper comparisons is **are the comments obviously different between groups?** Before conducting detailed analyses I tried several iterations of different visual comparison strategies.

####\

###Log-Odds Text Comparison
The log-odds ratios method outlined below has been the most informative to date. The basic premise being tested is that certain words are more frequent in TA comments belonging to one sub-category, and less frequent in others. I first saw this approach as a way to compare word usage between separate Twitter accounts here: https://www.tidytextmining.com/twitter.html.

For simplicity I reduced the analysis to just the four most common comment subjects.

```{r}
# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.

frequency_writing <- filter(base_data, code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
frequency_writing.subcolumns <- frequency_writing %>% select(1, 22:23)
```

```{r}
# Tokenize phrases, remove stop words listed in standard reference file.
base_data_tokenized <- frequency_writing.subcolumns %>% 
  unnest_tokens(word,ta.comment) %>% 
  anti_join(stop_words)

#At this point the dataframe should have "Unique.Record" in column 1, "code.subject" in column 2, and the unnested "word" in column 3. All columns are class = "character". Stop words listed in the "stop_words" R reference data file have been removed.

```

####Comparing Pearson Correlations of Word Frequencies
If the words used by TAs in their comments differ between the categories, there should be fairly low correlations overall .
```{r}
#Block below re-organizes the data for comparisons. This pattern sets "2. Writing Quality" as the first dataset, then all other data in remaining columns.
#Groups by code.subject, calculated proportional frequency in each group.
#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
#Stats need this format to be able to compare 2. WQ against other values.
base_data_tokenized_sorted2 <- base_data_tokenized %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  spread(code.subject, proportion) %>% 
  gather(code.subject, proportion, `3. Technical and Scientific`:`4. Logic and Thinking`)

#The dataframe called "base_data_tokenized_sorted2" is the final 4-column dataframe for statistical analysis. Individual words are in column 1, "2. Writing Quality" frequency is in Column 2, "code.subject" is in column 3, and "proportion" is in column 4. Columns 1 and 3 are class = "character", 2 and 4 are numeric ranges.

#This structure is needed in order to compare word (in Column 1) frequencies for 2. Writing Quality (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes. WHICH subset of X values to use is coded in Column 3.
```

Calculating Pearson correlation between the frequency of words in Writing Quality versus the other two groups.

```{r}
cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3. Technical and Scientific",], 
         ~ proportion + `2. Writing Quality`)
```

```{r}
cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4. Logic and Thinking",], 
         ~ proportion + `2. Writing Quality`)
```

####START OF BROKEN SECTION
THERE IS AN ERROR HERE I CANNOT DEBUG! If I change "3. Technical and Scientific"  to "2. Writing Quality", the structure of the tibble breaks. It goes from the correctly sorted 4 columns to incorrectly sorted 3 columns. It is NOT A TYPO; I can literally copy/paste it in and out and break/correct the code.

I cannot use a different order, or any other format. I am at a loss.

```{}
# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
frequency_writing3 <- filter(base_data, code.subject=="3. Technical and Scientific"|code.subject=="2. Writing Quality"|code.subject=="4. Logic and Thinking")
frequency_writing3.subcolumns <- frequency_writing3 %>% select(1, 22:23)
```

```{}
# Tokenize phrases, remove stop words listed in standard reference file.
base_data_tokenized3 <- frequency_writing3.subcolumns %>% 
  unnest_tokens(word,ta.comment) %>% 
  anti_join(stop_words)
```

```{}
#Block below re-organizes the data for comparisons. This pattern sets "3. Writing Quality" as the first dataset, then all other data in remaining columns.
#Groups by code.subject, calculated proportional frequency in each group.
#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
#Stats need this format to be able to compare 2. WQ against other values.
base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  spread(code.subject, proportion) %>% 
  gather(code.subject, proportion, `2. Writing Quality`:`4. Logic and Thinking`)
```

```{}
#Create groups by code.subject, calculates proportional frequency in each group.
#Final step creates 5 columns: code.subject, n, word, total, frequency.
base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  left_join(base_data_tokenized %>% 
              group_by(code.subject) %>% 
              summarise(total = n())) %>%
  mutate(freq = n/total)
```

Re-organize the data for comparisons. This pattern sets 3. Technical and Scientific as the first dataset, then all other data in remaining columns.
```{}
#Groups by code.subject, calculated proportional frequency in each group.
#Final step creates 4 columns: word, 3. Technical and Scientific, code.subject, proportion
#Stats need this format to be able to compare 3. TS against other values.
base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  spread(code.subject, proportion) %>% 
  gather(code.subject, proportion, `2. Writing Quality`:`4. Logic and Thinking`)

#END OF BROKEN SECTION
```

***

START OF SECTION I CANNOT COPY AND MODIFY. This section depends on the ability to calculate tables using the code above. So I cannot run this part either.

```{r}
#Create groups by code.subject, calculates proportional frequency in each group.
#Final step creates 5 columns: code.subject, n, word, total, frequency.
base_data_tokenized_sorted2 <- base_data_tokenized %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  left_join(base_data_tokenized %>% 
  group_by(code.subject) %>% 
  summarise(total = n())) %>%
  mutate(freq = n/total)
```

```{r}
base_data_tokenized_sorted2v3 <- base_data_tokenized_sorted2 %>% 
  select(code.subject, word, freq) %>% 
  spread(code.subject, freq) %>%
  arrange(`2. Writing Quality`,`3. Technical and Scientific`)
```

```{r}
ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`3. Technical and Scientific`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
  ggsave("Freqplot_3vs2.png", width = 10, height = 6)
```

```{r}
#I can get part way around the problem by re-using the same data and calling a different subset for the Y axis. Want to add comparisons of 3 vs 4, but same tables cannot be created because dataframe keeps breaking when I try to change 2. WQ.

base_data_tokenized_sorted2v4 <- base_data_tokenized_sorted2 %>% 
  select(code.subject, word, freq) %>% 
  spread(code.subject, freq) %>%
  arrange(`2. Writing Quality`,`4. Logic and Thinking`)
```

```{r}
ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`4. Logic and Thinking`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
  ggsave("Freqplot_4vs2.png", width = 10, height = 6)
```

END of broken / limping section

####Calculating and Plotting Log-Odds Pairs
**2. Writing versus 3. Technical**

```{r}
word_ratios2v3 <- base_data_tokenized %>%
  count(word, code.subject) %>%
  group_by(word) %>%
  filter(sum(n) >= 10) %>%
  ungroup() %>%
  spread(code.subject, n, fill = 0) %>%
  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
  mutate(logratio = log(`2. Writing Quality`/`3. Technical and Scientific`)) %>%
  arrange(desc(logratio))
```

```{r}
word_ratios2v3 %>% 
  arrange(abs(logratio))
```

```{r}
word_ratios2v3 %>%
  group_by(logratio < 0) %>%
  top_n(20, abs(logratio)) %>%
  ungroup() %>%
  mutate(word = reorder(word, logratio)) %>%
  ggplot(aes(word, logratio, fill = logratio < 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  ylab("log odds ratio (Writing-top/Technical-bottom)") +
  scale_fill_discrete(name = "", labels = c("2. Writing Quality", "3. Technical and Scientific"))
  #ggsave("Logplot_2vs3.png", width = 6, height = 10)

```

**3. Technical vs. 4. Logic**

```{r}
word_ratios3v4 <- base_data_tokenized %>%
  count(word, code.subject) %>%
  group_by(word) %>%
  filter(sum(n) >= 10) %>%
  ungroup() %>%
  spread(code.subject, n, fill = 0) %>%
  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
  mutate(logratio = log(`3. Technical and Scientific`/`4. Logic and Thinking`)) %>%
  arrange(desc(logratio))
```

```{r}
word_ratios3v4 %>% 
  arrange(abs(logratio))
```

```{r}
word_ratios3v4 %>%
  group_by(logratio < 0) %>%
  top_n(20, abs(logratio)) %>%
  ungroup() %>%
  mutate(word = reorder(word, logratio)) %>%
  ggplot(aes(word, logratio, fill = logratio < 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  ylab("log odds ratio (Technical-top/Logic-bottom)") +
  scale_fill_discrete(name = "", labels = c("3. Technical and Scientific","4. Logic and Thinking"))
  #ggsave("Logplot_3vs4.png", width = 6, height = 10)

```

**2. Writing versus 4. Logic**

```{r}
word_ratios2v4 <- base_data_tokenized %>%
  count(word, code.subject) %>%
  group_by(word) %>%
  filter(sum(n) >= 10) %>%
  ungroup() %>%
  spread(code.subject, n, fill = 0) %>%
  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
  mutate(logratio = log(`2. Writing Quality`/`4. Logic and Thinking`)) %>%
  arrange(desc(logratio))
```

```{r}
word_ratios2v4 %>% 
  arrange(abs(logratio))
```

```{r}
word_ratios2v4 %>%
  group_by(logratio < 0) %>%
  top_n(20, abs(logratio)) %>%
  ungroup() %>%
  mutate(word = reorder(word, logratio)) %>%
  ggplot(aes(word, logratio, fill = logratio < 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  ylab("log odds ratio (Writing-top/Logic-bottom)") +
  scale_fill_discrete(name = "", labels = c("2. Writing Quality", "4. Logic and Thinking"))
  #ggsave("Logplot_2vs4.png", width = 6, height = 10)

```

##Direct Comparison Using Only Words Appearing in All Subsets of Comments
I think this is less informative, but it does let me see the data more clearly.Basic strategy is to take all comments from base_data, unnest tokens, filter stopwords, count and sort remaining terms, then write back to a new datatable called 'sortedwords.comments.all'.  Calculate TOTAL words and store. Then calculate fractional frequencies for each word and store using mutate.

This is an early version, so likely has some poorly developed code blocks that could be refactored.

```{r}
library(dplyr)
library(tidytext)

sortedwords.comments.all <- base_data_tokenized %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.all <- sortedwords.comments.all %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
  all.fraction = (n*100)/total_words.all$total
  )
```

##Summarize Word Frequencies in Subsets of Comments Based on 'Subject'
Take subsets of comments from base_data_tokenized using code.subject as separating characteristic.

###Other Analysis Approaches
This approach summarizes the word frequencies in the FULL set of TA comments first, then breaks them down into sub-groups. It has examples of other code bits as well.

The first block takes all comments from base_data, unnests tokens, filters stopwords, count and sort remaining terms, then writes results back to a new dataframe called 'sortedwords.comments.all'.

TOTAL words is calculated and stored. Then fractional frequencies for each word are calculated and stored back to the dataframe using mutate.

```{r}
# Tokenize phrases, remove stop words listed in standard reference file.
base_data_tokenized <- base_data %>% 
  unnest_tokens(word,ta.comment)

# An alternative approach to pulling out stopwords  
sortedwords.comments.all <- base_data_tokenized %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.all <- sortedwords.comments.all %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
  all.fraction = (n*100)/total_words.all$total
  )
```

```{r}
#Pull SUBSETs of tokenized data based on code.subject: 
base_data_tokenized.basic <- subset(base_data_tokenized, code.subject == "1. Basic Criteria", select = (1:28))
base_data_tokenized.writing <- subset(base_data_tokenized, code.subject == "2. Writing Quality", select = (1:28))
base_data_tokenized.technical <- subset(base_data_tokenized, code.subject == "3. Technical and Scientific", select = (1:28))
base_data_tokenized.logic <- subset(base_data_tokenized, code.subject == "4. Logic and Thinking", select = (1:28))
base_data_tokenized.praise <- subset(base_data_tokenized, code.subject == "5. Praise or Concern", select = (1:28))
base_data_tokenized.misconduct <- subset(base_data_tokenized, code.subject == "6. Misconduct", select = (1:28))
base_data_tokenized.narrative <- subset(base_data_tokenized, code.subject == "12. Narrative Comments", select = (1:28))
base_data_tokenized.sciname <- subset(base_data_tokenized, code.subject == "Scientific Name", select = (1:28))
```

For each subset, filter stopwords, count words and total, then calculate frequencies.

```{r}
sortedwords.basic <- base_data_tokenized.basic %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.basic <- sortedwords.basic %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.basic.fraction <- sortedwords.basic %>% as_tibble() %>% mutate(
  basic.fraction = (n*100)/total_words.basic$total
)
```

```{r}
sortedwords.writing <- base_data_tokenized.writing %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.writing <- sortedwords.writing %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.writing.fraction <- sortedwords.writing %>% as_tibble() %>% mutate(
  writing.fraction = (n*100)/total_words.writing$total
)
```

```{r}
sortedwords.technical <- base_data_tokenized.technical %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.technical <- sortedwords.technical %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.technical.fraction <- sortedwords.technical %>% as_tibble() %>% mutate(
  technical.fraction = (n*100)/total_words.technical$total
)
```

```{r}
sortedwords.logic <- base_data_tokenized.logic %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.logic <- sortedwords.logic %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.logic.fraction <- sortedwords.logic %>% as_tibble() %>% mutate(
  logic.fraction = (n*100)/total_words.logic$total
)
```

Next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed.

**WARNING**: Words that are MISSING from **any subset** cause the full set to be pruned. I am losing the unique words this way. **Need to figure out how to join without pruning.**

```{r}
total <- merge(sortedwords.comments.all.fraction,sortedwords.technical.fraction,by="word") 
total2 <- merge(total,sortedwords.writing.fraction,by="word")
total3 <- merge(total2,sortedwords.logic.fraction,by="word")
#rename lines that were possibly confusing
names(total3)[2] <- "all.count"
names(total3)[4] <- "technical.count"
names(total3)[6] <- "writing.count"
names(total3)[8] <- "logic.count"

#Mutate table to calculate then append fractional word frequency in the text.
total4 <- total3 %>% as_tibble() %>% mutate(
  writing.wt = (writing.count *100)/all.count
)

total5 <- total4 %>% as_tibble() %>% mutate(
  technical.wt = (technical.count *100)/all.count
)

total6 <- total5 %>% as_tibble() %>% mutate(
  logic.wt = (logic.count *100)/all.count
)

```

The following six graphs only compare frequencies for ~770 words that appear in ALL THREE of the sub-categories. Graphs 1-3 compare frequencies in individual groups against word frequency overall. Generally, the frequency of a term in the whole set of comments is similar to frequency in an individual group (X vs Y positions similar.) However, some terms associated with Logic subgroup clearly fall below the trendline Graph 3, suggesting they are less frequent in that sub-category than overall.

```{r}
ggplot(total6, aes(`all.fraction`,`writing.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
```

```{r}
ggplot(total6, aes(`all.fraction`,`technical.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
```

```{r}
ggplot(total6, aes(`all.fraction`,`logic.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
```

Graphs 4-6 are pair-wise comparisons of frequencies for sub-categories. Unlike the comparisons to overall frequency, most words are NOT clustered on the trendline, suggesting there are differences in word usage.

```{r}
ggplot(total6, aes(`technical.fraction`,`writing.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
```

```{r}
ggplot(total6, aes(`logic.fraction`,`writing.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
```

```{r}
ggplot(total6, aes(`logic.fraction`,`technical.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
```

####\

## Data

####\

## Analysis

####\

# Results

####\

# Discussion

####\

# Conclusion

####\

# References

####\

### 2018-12-17 draft.Rmd
update

@@
 ---
-title: "Project"
+title: "Data Intake & Initial Exploration"
 author: "Dan Johnson"
-date: "10/26/2018"
+date: "12/16/2018"
 output: html_document
@@
 
-# Abstract
+The initial working dataset is an anonymized Excel table of 10,808 rows.
+One column has "TA comments" extracted from student lab reports.
+Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation.
+Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.
+
+####\  
+##Data Pre-Checks
+Prior to import, review data table in Excel. Check:
+
+1.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.md"
+2.  Data have been de-identified and recoded using pre-assigned anonymous IDs.
+3.  Data table headers are in snake_case, and match tabulated list below. The same list is in the file "code_column_names.txt"
+
+Column Number|Column Name|Column Number|Column Name
+------------|----------------------|------------|----------------------
+1|unique.record|15|data.ok
+2|report.id|16|citation.ok
+3|sort|17|interpretation.ok
+4|report.title (empty)|18|organization.ok
+5|student (Std\_nnn)|19|techflaws.ok
+6|course|20|writing.ok
+7|ta (TA\_nnn)|21|comments.incorporated
+8|lab|22|ta.comment
+9|tag|23|code.subject
+10|type.TA|24|code.structure
+11|grade.TA|25|code.locus
+12|grading.time|26|code.scope
+13|Rank|27|code.tone
+14|hypothesis.ok|28|code.notes
+
+Once all data are in correct format, export Excel file to CSV with name formatted as **"coded_full_comments_dataset_SemYear.csv"**
 
@@
 
-# Introduction
-## Background
-Undergraduate introductory STEM courses typically include technical writing training.
-Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed.
-Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills.
+##Initial Import and Post-Check
+```{r Libraries}
+# Call in tidyverse, other required libraries.
+library(tidyverse)
+library(tidytext)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(scales)
+```
 
-Technical writing training usually is part of a lab course associated with a lecture class.
-These lab sections may be be led by graduate teaching assistants or even advanced undergraduates.
-These instructors may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.
+```{r}
+#This block reads in the full CSV into a starting dataframe named "base_data".
+#Subsequent analyses call in subsets of "base_data".
+base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
 
-In 2014 we launched an automated data collection process we call SAWHET that collects systematic observations and sample data from student writers and graduate instructors.
-Our resarch goal is to better understand student and instructor behaviors within the technical writing instructional space.
-Our practical goal is to develop techniques for teaching technical writing that:
+#If the TA comments dataset is imported correctly, there will be a dataframe of ~11,000 rows and 28 columns named as listed in the **Data Pre-Checks** table above.
+```
 
-*  Can be implemented at large scale by less experienced teaching assistants;
-*  Emphasize student development through holistic coaching;
-*  Shift student and instructor attention away from simple copy-correction towards larger writing issues; and
-*  Embody other best practices outlined in nearly 40 years of research literature on "writing in the disciplines."
+Given some data points may be entered incorrectly by instructors or students, I included several post-import data checks to ensure data have been properly coded and columns contain valid entries only.
 
-We have a structured database of >5000 laboratory reports written by undergraduate students for introductory biology courses.
-In addition we have all comments that a TA made on the individual reports, plus associated metadata:
+```{r}
+#"Course" column should only have entries that correspond to 113, 114, 214, NA. 
+course<-unique(base_data$course)
 
-*  Context. Was it first or second report of the semester? Initial submission or revision? What course and topic in the course? What year in college was the author?
-*  Instructor information. Which TA provided the feedback? What score was assigned, and what what the rationale behind that score?
+#Entries in "TAs" column should match the anonymous IDs that correspond to the TAs assigned to the relevant courses in the semester being analyzed. 
+#If there are incorrect or extra anonymous IDs in this list, the mostly likely source of the error is improper de-identification of the Excel file prior to import.
+ta<-unique(base_data$ta)
 
-We are using multiple text-mining methods to map key features of student writing, as well as connections and correlations between those features.
-These insights then inform how we teach technical writing and train graduate instructors.
+#"Lab" should only have entries that match the limited keywords in codebook.
+lab<-unique(base_data$lab)
+
+#Look at the NUMBER of unique anonymous student IDs. Should be within 10% of (but not more than) the combined enrollment of the 3 target courses for the semester.
+student<-unique(base_data$student)
+
+#Check that the list of "code.subject" topics here matches allowed terms in codebook.
+subject<-unique(base_data$code.subject)
+
+#Check that the list of "code.structure" terms extracted from the dataset and displayed here matches allowed terms in codebook.
+structure<-unique(base_data$code.structure)
+```
 
@@
 
-## Context for This Project & Prior Work
-Students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead.
-To that end, we train graduate teaching assistants to use a **global first, specific later** approach to grading undergraduate technical writing.
-We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues.
-Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection.
-As students' writing improves, the TA can focus on smaller issues.
-
-This strategy is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy.
-One of our ongoing questions has been, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**
-
-In Summer 2018, we began analyzing TAs' comments from student papers.
-We extracted and tabulated all comments from approximately 500 pairs of reports (initial submissions and revisions) collected during a single semester in 3 different courses.
-Initial data cleaning involved removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent.
-In all, the final tabulated dataset included ~11,000 separate comments.
-
-We next used a subset of 100 randomly selected comments to build an initial qualitative codebook for sorting comments into types.
-The initial codebook was re-tested using 1100 additional randomly selected comments.
-Using the final codebook, we sorted all 11,000 comments into separate categories (see LINK for complete codebook.)
-
-Coded TA comments were sorted on two axes:
-
-* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
-* __Structure__: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?
-
-The sorted contingency table provided us with a relative estimate of TA effort.
+## Preliminary Exploration
+The first question to answer before making deeper comparisons is **are the comments obviously different between sub-categories?** Before conducting detailed analyses I tried several iterations of different visual comparison strategies.
 
@@
 
-## The Central Problem of This Project
-Qualitative coding of TA comments is informative but unsustainable.
-Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort.
-Even with the codebook, individual coders will require training to achieve any accuracy.
-Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.
+###Log-Odds Text Comparison
+The log-odds ratios method outlined below has been the most informative to date.
+The basic premise being tested is that certain words are more frequent in TA comments belonging to one sub-category, and less frequent in others.
+I first saw this approach as a way to compare word usage between separate Twitter accounts here: https://www.tidytextmining.com/twitter.html.
 
-####\
+For simplicity I reduced the analysis to just the four most common comment subjects.
 
-# Literature Review
-##General Approach to Initial Dataset
-The following excerpt from *Text Mining with R: A Tidy Approach* (Julia Silge and David Robinson) summarizes how extended text (in this case, TA comments) can be organized as a tidy matrix.*All emphases are mine.*
+```{r}
+# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
+frequency_writing <- filter(base_data, code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
+frequency_writing.subcolumns <- frequency_writing %>% select(1, 22:23)
 
->We thus define the tidy text format as being a table with one-token-per-row.
-A token is a **meaningful unit of text**, such as a word, that we are interested in using for analysis, and tokenization is the process of splitting text into tokens.
+# Tokenize phrases, remove stop words listed in standard reference file.
+base_data_tokenized <- frequency_writing.subcolumns %>% 
+  unnest_tokens(word,ta.comment) %>% 
+  anti_join(stop_words)
 
->This one-token-per-row structure is in contrast to the ways text is often stored in current analyses, perhaps as strings or in a document-term matrix.
-For tidy text mining, the token that is stored in each row is most often a single word, **but can also be an n-gram, sentence, or paragraph** .
-In the tidytext package, we provide functionality to tokenize by commonly used units of text like these and convert to a one-term-per-row format.
+#At this point the dataframe should have "Unique.Record" in column 1, "code.subject" in column 2, and the unnested "word" in column 3.
+All columns are class = "character".
+Stop words listed in the "stop_words" R reference data file have been removed.
 
-Using tidy format (vs. other common text data formats) better maintains relationships between individual comments and metadata, which simplifies subsequent analysis.
-
-####\
-
-# Methods
-
-####\
-
-## Preliminary Analysis
-###Log-Odds Text Comparison
+```
 
 ####Comparing Pearson Correlations of Word Frequencies
+If the words used by TAs in their comments differ between the categories, there should be fairly low correlations overall .
+```{r}
+#Block below re-organizes the data for comparisons. This pattern sets "2. Writing Quality" as the first dataset, then all other data in remaining columns.
+#Groups by code.subject, calculated proportional frequency in each group.
+#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
+#Stats need this format to be able to compare 2. WQ against other values.
+base_data_tokenized_sorted2 <- base_data_tokenized %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  mutate(proportion = n / sum(n)) %>% 
+  select(-n) %>% 
+  spread(code.subject, proportion) %>% 
+  gather(code.subject, proportion, `3. Technical and Scientific`:`4. Logic and Thinking`)
+
+#The dataframe called "base_data_tokenized_sorted2" is the final 4-column dataframe for statistical analysis.
+Individual words are in column 1, "2.
+Writing Quality" frequency is in Column 2, "code.subject" is in column 3, and "proportion" is in column 4.
+Columns 1 and 3 are class = "character", 2 and 4 are numeric ranges.
+
+#This structure is needed in order to compare word (in Column 1) frequencies for 2.
+Writing Quality (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes.
+WHICH subset of X values to use is coded in Column 3.
+```
+
+**Pearson correlation** between the frequency of words in Writing Quality versus Technical and Logic sub-categoris.
+
+```{r}
+cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3. Technical and Scientific",], 
+         ~ proportion + `2. Writing Quality`)
+```
+
+```{r}
+cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4. Logic and Thinking",], 
+         ~ proportion + `2. Writing Quality`)
+```
+
+***
+***
+####START OF BROKEN SECTION
+THERE IS AN ERROR HERE I CANNOT DEBUG!
+If I change "3.
+Technical and Scientific"  to "2.
+Writing Quality", the structure of the tibble breaks.
+It goes from the correctly sorted 4 columns to incorrectly sorted 3 columns.
+It is NOT A TYPO; I can literally copy/paste it in and out and break/correct the code.
+
+I cannot use a different order, or any other format. I am at a loss.
+
+```{}
+# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
+frequency_writing3 <- filter(base_data, code.subject=="3. Technical and Scientific"|code.subject=="2. Writing Quality"|code.subject=="4. Logic and Thinking")
+frequency_writing3.subcolumns <- frequency_writing3 %>% select(1, 22:23)
+```
+
+```{}
+# Tokenize phrases, remove stop words listed in standard reference file.
+base_data_tokenized3 <- frequency_writing3.subcolumns %>% 
+  unnest_tokens(word,ta.comment) %>% 
+  anti_join(stop_words)
+```
+
+```{}
+#Block below re-organizes the data for comparisons. This pattern sets "3. Writing Quality" as the first dataset, then all other data in remaining columns.
+#Groups by code.subject, calculated proportional frequency in each group.
+#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
+#Stats need this format to be able to compare 2. WQ against other values.
+base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  mutate(proportion = n / sum(n)) %>% 
+  select(-n) %>% 
+  spread(code.subject, proportion) %>% 
+  gather(code.subject, proportion, `2. Writing Quality`:`4. Logic and Thinking`)
+```
+
+```{}
+#Create groups by code.subject, calculates proportional frequency in each group.
+#Final step creates 5 columns: code.subject, n, word, total, frequency.
+base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  left_join(base_data_tokenized %>% 
+              group_by(code.subject) %>% 
+              summarise(total = n())) %>%
+  mutate(freq = n/total)
+```
+
+Re-organize the data for comparisons. This pattern sets 3. Technical and Scientific as the first dataset, then all other data in remaining columns.
+```{}
+#Groups by code.subject, calculated proportional frequency in each group.
+#Final step creates 4 columns: word, 3. Technical and Scientific, code.subject, proportion
+#Stats need this format to be able to compare 3. TS against other values.
+base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  mutate(proportion = n / sum(n)) %>% 
+  select(-n) %>% 
+  spread(code.subject, proportion) %>% 
+  gather(code.subject, proportion, `2. Writing Quality`:`4. Logic and Thinking`)
+
+#END OF BROKEN SECTION
+```
+
+***
+***
+
+I can run this block, but CANNOT COPY AND MODIFY THIS SECTION. It depends tables created using the broken code above. So I cannot run this part either.
+
+```{r}
+#Create groups by code.subject, calculates proportional frequency in each group.
+#Final step creates 5 columns: code.subject, n, word, total, frequency.
+base_data_tokenized_sorted2 <- base_data_tokenized %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  left_join(base_data_tokenized %>% 
+  group_by(code.subject) %>% 
+  summarise(total = n())) %>%
+  mutate(freq = n/total)
+```
+
+```{r}
+base_data_tokenized_sorted2v3 <- base_data_tokenized_sorted2 %>% 
+  select(code.subject, word, freq) %>% 
+  spread(code.subject, freq) %>%
+  arrange(`2. Writing Quality`,`3. Technical and Scientific`)
+```
+
+```{r}
+ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`3. Technical and Scientific`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+  ggsave("Freqplot_3vs2.png", width = 10, height = 6)
+```
+
+```{r}
+#I can get part way around the problem by re-using the same data and calling a different subset for the Y axis.
+Want to add comparisons of 3 vs 4, but same tables cannot be created because dataframe keeps breaking when I try to change 2.
+WQ.
+
+base_data_tokenized_sorted2v4 <- base_data_tokenized_sorted2 %>% 
+  select(code.subject, word, freq) %>% 
+  spread(code.subject, freq) %>%
+  arrange(`2. Writing Quality`,`4. Logic and Thinking`)
+```
+
+```{r}
+ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`4. Logic and Thinking`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+  ggsave("Freqplot_4vs2.png", width = 10, height = 6)
+```
+
+END of broken / limping section
+***
+***
 
 ####Calculating and Plotting Log-Odds Pairs
+**2. Writing versus 3. Technical**
 
-###Direct Comparison Using Only Words Appearing in All Subsets of Comments
+```{r}
+word_ratios2v3 <- base_data_tokenized %>%
+  count(word, code.subject) %>%
+  group_by(word) %>%
+  filter(sum(n) >= 10) %>%
+  ungroup() %>%
+  spread(code.subject, n, fill = 0) %>%
+  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
+  mutate(logratio = log(`2. Writing Quality`/`3. Technical and Scientific`)) %>%
+  arrange(desc(logratio))
+```
+
+```{r}
+word_ratios2v3 %>% 
+  arrange(abs(logratio))
+```
+
+```{r}
+word_ratios2v3 %>%
+  group_by(logratio < 0) %>%
+  top_n(20, abs(logratio)) %>%
+  ungroup() %>%
+  mutate(word = reorder(word, logratio)) %>%
+  ggplot(aes(word, logratio, fill = logratio < 0)) +
+  geom_col(show.legend = FALSE) +
+  coord_flip() +
+  ylab("log odds ratio (Writing-top/Technical-bottom)") +
+  scale_fill_discrete(name = "", labels = c("2. Writing Quality", "3. Technical and Scientific"))
+  #ggsave("Logplot_2vs3.png", width = 6, height = 10)
+
+```
+
+**3. Technical vs. 4. Logic**
+
+```{r}
+word_ratios3v4 <- base_data_tokenized %>%
+  count(word, code.subject) %>%
+  group_by(word) %>%
+  filter(sum(n) >= 10) %>%
+  ungroup() %>%
+  spread(code.subject, n, fill = 0) %>%
+  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
+  mutate(logratio = log(`3. Technical and Scientific`/`4. Logic and Thinking`)) %>%
+  arrange(desc(logratio))
+```
+
+```{r}
+word_ratios3v4 %>% 
+  arrange(abs(logratio))
+```
+
+```{r}
+word_ratios3v4 %>%
+  group_by(logratio < 0) %>%
+  top_n(20, abs(logratio)) %>%
+  ungroup() %>%
+  mutate(word = reorder(word, logratio)) %>%
+  ggplot(aes(word, logratio, fill = logratio < 0)) +
+  geom_col(show.legend = FALSE) +
+  coord_flip() +
+  ylab("log odds ratio (Technical-top/Logic-bottom)") +
+  scale_fill_discrete(name = "", labels = c("3. Technical and Scientific","4. Logic and Thinking"))
+  #ggsave("Logplot_3vs4.png", width = 6, height = 10)
+
+```
+
+**2. Writing versus 4. Logic**
+
+```{r}
+word_ratios2v4 <- base_data_tokenized %>%
+  count(word, code.subject) %>%
+  group_by(word) %>%
+  filter(sum(n) >= 10) %>%
+  ungroup() %>%
+  spread(code.subject, n, fill = 0) %>%
+  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
+  mutate(logratio = log(`2. Writing Quality`/`4. Logic and Thinking`)) %>%
+  arrange(desc(logratio))
+```
+
+```{r}
+word_ratios2v4 %>% 
+  arrange(abs(logratio))
+```
+
+```{r}
+word_ratios2v4 %>%
+  group_by(logratio < 0) %>%
+  top_n(20, abs(logratio)) %>%
+  ungroup() %>%
+  mutate(word = reorder(word, logratio)) %>%
+  ggplot(aes(word, logratio, fill = logratio < 0)) +
+  geom_col(show.legend = FALSE) +
+  coord_flip() +
+  ylab("log odds ratio (Writing-top/Logic-bottom)") +
+  scale_fill_discrete(name = "", labels = c("2. Writing Quality", "4. Logic and Thinking"))
+  #ggsave("Logplot_2vs4.png", width = 6, height = 10)
+
+```
+
+##Direct Comparison Using Only Words Appearing in All Subsets of Comments
+I think this is less informative, but it does let me see the data more clearly.Basic strategy is to take all comments from base_data, unnest tokens, filter stopwords, count and sort remaining terms, then write back to a new datatable called 'sortedwords.comments.all'.
+Calculate TOTAL words and store.
+Then calculate fractional frequencies for each word and store using mutate.
+
+This is an early version, so likely has some poorly developed code blocks that could be refactored.
+
+```{r}
+library(dplyr)
+library(tidytext)
+
+sortedwords.comments.all <- base_data_tokenized %>%
+  filter(!word %in% stop_words$word,
+         !word %in% str_remove_all(stop_words$word, "'"),
+         str_detect(word, "[a-z]")) %>%
+  count(word, word, sort = TRUE) %>%
+  ungroup()
+
+#Calculate the number of words in total in the subset
+total_words.all <- sortedwords.comments.all %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
+  all.fraction = (n*100)/total_words.all$total
+  )
+```
+
+##Summarize Word Frequencies in Subsets of Comments Based on 'Subject'
+Take subsets of comments from base_data_tokenized using code.subject as separating characteristic.
 
 ###Other Analysis Approaches
+This approach summarizes the word frequencies in the FULL set of TA comments first, then breaks them down into sub-groups. It has examples of other code bits as well.
 
-####\
+The first block takes all comments from base_data, unnests tokens, filters stopwords, count and sort remaining terms, then writes results back to a new dataframe called 'sortedwords.comments.all'.
 
-## Data
+TOTAL words is calculated and stored. Then fractional frequencies for each word are calculated and stored back to the dataframe using mutate.
 
-####\
+```{r}
+# Tokenize phrases, remove stop words listed in standard reference file.
+base_data_tokenized <- base_data %>% 
+  unnest_tokens(word,ta.comment)
 
-## Analysis
+# An alternative approach to pulling out stopwords  
+sortedwords.comments.all <- base_data_tokenized %>%
+  filter(!word %in% stop_words$word,
+         !word %in% str_remove_all(stop_words$word, "'"),
+         str_detect(word, "[a-z]")) %>%
+  count(word, word, sort = TRUE) %>%
+  ungroup()
 
-####\
+#Calculate the number of words in total in the subset
+total_words.all <- sortedwords.comments.all %>% 
+  summarize(total = sum(n))
 
-# Results
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
+  all.fraction = (n*100)/total_words.all$total
+  )
+```
 
-####\
+```{r}
+#Pull SUBSETs of tokenized data based on code.subject: 
+base_data_tokenized.basic <- subset(base_data_tokenized, code.subject == "1. Basic Criteria", select = (1:28))
+base_data_tokenized.writing <- subset(base_data_tokenized, code.subject == "2. Writing Quality", select = (1:28))
+base_data_tokenized.technical <- subset(base_data_tokenized, code.subject == "3. Technical and Scientific", select = (1:28))
+base_data_tokenized.logic <- subset(base_data_tokenized, code.subject == "4. Logic and Thinking", select = (1:28))
+base_data_tokenized.praise <- subset(base_data_tokenized, code.subject == "5. Praise or Concern", select = (1:28))
+base_data_tokenized.misconduct <- subset(base_data_tokenized, code.subject == "6. Misconduct", select = (1:28))
+base_data_tokenized.narrative <- subset(base_data_tokenized, code.subject == "12. Narrative Comments", select = (1:28))
+base_data_tokenized.sciname <- subset(base_data_tokenized, code.subject == "Scientific Name", select = (1:28))
+```
 
-# Discussion
+For each subset, filter stopwords, count words and total, then calculate frequencies.
 
-####\
+```{r}
+sortedwords.basic <- base_data_tokenized.basic %>%
+  filter(!word %in% stop_words$word,
+         !word %in% str_remove_all(stop_words$word, "'"),
+         str_detect(word, "[a-z]")) %>%
+  count(word, word, sort = TRUE) %>%
+  ungroup()
 
-# Conclusion
+#Calculate the number of words in total in the subset
+total_words.basic <- sortedwords.basic %>% 
+  summarize(total = sum(n))
 
-####\
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.basic.fraction <- sortedwords.basic %>% as_tibble() %>% mutate(
+  basic.fraction = (n*100)/total_words.basic$total
+)
+```
 
-# References
+```{r}
+sortedwords.writing <- base_data_tokenized.writing %>%
+  filter(!word %in% stop_words$word,
+         !word %in% str_remove_all(stop_words$word, "'"),
+         str_detect(word, "[a-z]")) %>%
+  count(word, word, sort = TRUE) %>%
+  ungroup()
 
-####\
+#Calculate the number of words in total in the subset
+total_words.writing <- sortedwords.writing %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.writing.fraction <- sortedwords.writing %>% as_tibble() %>% mutate(
+  writing.fraction = (n*100)/total_words.writing$total
+)
+```
+
+```{r}
+sortedwords.technical <- base_data_tokenized.technical %>%
+  filter(!word %in% stop_words$word,
+         !word %in% str_remove_all(stop_words$word, "'"),
+         str_detect(word, "[a-z]")) %>%
+  count(word, word, sort = TRUE) %>%
+  ungroup()
+
+#Calculate the number of words in total in the subset
+total_words.technical <- sortedwords.technical %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.technical.fraction <- sortedwords.technical %>% as_tibble() %>% mutate(
+  technical.fraction = (n*100)/total_words.technical$total
+)
+```
+
+```{r}
+sortedwords.logic <- base_data_tokenized.logic %>%
+  filter(!word %in% stop_words$word,
+         !word %in% str_remove_all(stop_words$word, "'"),
+         str_detect(word, "[a-z]")) %>%
+  count(word, word, sort = TRUE) %>%
+  ungroup()
+
+#Calculate the number of words in total in the subset
+total_words.logic <- sortedwords.logic %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.logic.fraction <- sortedwords.logic %>% as_tibble() %>% mutate(
+  logic.fraction = (n*100)/total_words.logic$total
+)
+```
+
+Next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed.
+
+**WARNING**: Words that are MISSING from **any subset** cause the full set to be pruned. I am losing the unique words this way. **Need to figure out how to join without pruning.**
+
+```{r}
+total <- merge(sortedwords.comments.all.fraction,sortedwords.technical.fraction,by="word") 
+total2 <- merge(total,sortedwords.writing.fraction,by="word")
+total3 <- merge(total2,sortedwords.logic.fraction,by="word")
+#rename lines that were possibly confusing
+names(total3)[2] <- "all.count"
+names(total3)[4] <- "technical.count"
+names(total3)[6] <- "writing.count"
+names(total3)[8] <- "logic.count"
+
+#Mutate table to calculate then append fractional word frequency in the text.
+total4 <- total3 %>% as_tibble() %>% mutate(
+  writing.wt = (writing.count *100)/all.count
+)
+
+total5 <- total4 %>% as_tibble() %>% mutate(
+  technical.wt = (technical.count *100)/all.count
+)
+
+total6 <- total5 %>% as_tibble() %>% mutate(
+  logic.wt = (logic.count *100)/all.count
+)
+
+```
+
+The following six graphs only compare frequencies for ~770 words that appear in ALL THREE of the sub-categories.
+Graphs 1-3 compare frequencies in individual groups against word frequency overall.
+Generally, the frequency of a term in the whole set of comments is similar to frequency in an individual group (X vs Y positions similar.) However, some terms associated with Logic subgroup clearly fall below the trendline Graph 3, suggesting they are less frequent in that sub-category than overall.
+
+```{r}
+ggplot(total6, aes(`all.fraction`,`writing.fraction`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+```
+
+```{r}
+ggplot(total6, aes(`all.fraction`,`technical.fraction`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+```
+
+```{r}
+ggplot(total6, aes(`all.fraction`,`logic.fraction`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+```
+
+Graphs 4-6 are pair-wise comparisons of frequencies for sub-categories.
+Unlike the comparisons to overall frequency, most words are NOT clustered on the trendline, suggesting there are differences in word usage.
+
+```{r}
+ggplot(total6, aes(`technical.fraction`,`writing.fraction`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+```
+
+```{r}
+ggplot(total6, aes(`logic.fraction`,`writing.fraction`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+```
+
+```{r}
+ggplot(total6, aes(`logic.fraction`,`technical.fraction`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+```

### 2018-12-18 Naive_bayes.Rmd
update

---
title: "Naive Bayes"
author: "Dan Johnson"
date: "12/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Demonstration of Applying Naive Bayes Method

The objective of this project is to classify SMS messages as spam or ham (not spam). A Naive Bayes classifier approach will be used. This example is taken from chapter 4 of Machine Learning with R, Second Edition"

An example of the conditional probability that will be computed is as follows:

P(Spam|Hospital) = P(Hospital|Spam)P(Spam)/P(Hospital)

which is the formula for determining the probability that a message is spam given that it contains the word "Hospital" in the message.

##Loading Data Into R.

```{r}
sms_raw <- read.table("~/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/data/Bayes/SMSSpamCollection.txt", header = FALSE, sep = "\t", quote = "", col.names = c("type","text"), stringsAsFactors = FALSE)
```

The data frame has 5574 observations with 2 observations of either spam or ham.

```{r}
str(sms_raw)
```

As seen above, the "type"" element is a character.Change it to a factor for the analysis.

```{r}
sms_raw$type <- factor(sms_raw$type)
str(sms_raw$type)
table(sms_raw$type)
```

Now the "type" variable is a factor with 2 levels. Of the 5574 messages, 747 are spam, which is about 13.4%.

##Text Mining

For Naive Bayes to run effectively, the test data needs to be transformed. This begins with using the "tm" package in R to create a volitile coprus that contains the "text" vector from our data frame.

```{r}
#install.packages("tm")
library(tm)
```

```{r}
sms_corpus <- VCorpus(VectorSource(sms_raw$text))
print(sms_corpus)
```

Check out the first few messages in the new corpus, which is basically a list that can be manipulated with list operations.

```{r}
inspect(sms_corpus[1:3])
```

Use "as.character" function to see what a message looks like.

```{r}
as.character(sms_corpus[[3]])
```

In order to standardize the messages, the data set must be tranformed to all lower case letters. The words "Free", "free", and "FREE" should all be treated as the same word. Use the "tm_map"" funtion in R, and use the "content_transformer" function to transform the text.

```{r}
sms_corpus_clean <- tm_map(sms_corpus, content_transformer(tolower))
```

Look at third message again to see if our data was transformed.

```{r}
as.character(sms_corpus[[3]])

as.character((sms_corpus_clean[[3]]))
```

The message was indeed tranformed to all lowercase letters. Now to remove the numbers using the "removeNumbers" function.

```{r}
sms_corpus_clean <- tm_map(sms_corpus_clean, removeNumbers)
```

Remome words that appear often but don't contribute to our objective. These words include "to", "and", "but" and "or".

```{r}
sms_corpus_clean <- tm_map(sms_corpus_clean, removeWords, stopwords())
```

Remove punctuation as well using the "removePunctuation" function.

```{r}
sms_corpus_clean <- tm_map(sms_corpus_clean, removePunctuation)

as.character((sms_corpus_clean[[3]]))
```

Perform "stemming" to the text data to strip the suffix from words like "jumping", so the words "jumping" "jumps" and "jumped" are all transformed into "jump". Stemming can be perfromed using the "tm" package with help from the "SnowballC" package.

```{r}
#install.packages("SnowballC")
library(SnowballC)

sms_corpus_clean <- tm_map(sms_corpus_clean, stemDocument)
```

And now the final step in text mining is to remove white space from the document.

```{r}
sms_corpus_clean <- tm_map(sms_corpus_clean, stripWhitespace)
```

```{r}
as.character(sms_corpus_clean[[3]])
```

Perform tokenization using the "DocumentTermMatrix" function. This creates a matrix in which the rows indicates documents (SMS messages in this case) and the columns indicate words. It should be noted that the "DocumentTermMaxtrix" function has the power to do all of the text mining above in one command.

```{r}
sms_dtm <- DocumentTermMatrix(sms_corpus_clean)
```

##Data Preparation
Split our data into training and testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data. Divide our data set into 75% training and 25% testing.

```{r}
.75 * 5574
.25 * 5574
```

Because the dataset is random, the first 4180 messages can be used for the training set - there’s no need to randomize the data first.

```{r}
sms_dtm_train <- sms_dtm[1:4180, ]
sms_dtm_test <- sms_dtm[4181:5559, ]
```

Save vectors labeling rows in the training and testing vectors

```{r}
sms_train_labels <- sms_raw[1:4180, ]$type
sms_test_labels <- sms_raw[4181:5559,]$type
```

Make sure that the proportion of spam is similar in the training and testing data set.

```{r}
prop.table(table(sms_train_labels))
```

```{r}
prop.table(table(sms_test_labels))
```

sms_test_labels
      ham      spam 
0.8694706 0.1305294

Each have approx. 13% spam.

##Visualization
Create a wordcloud of the frequency of the words in the dataset using the package “wordcloud”.

```{r}
library(wordcloud)
wordcloud(sms_corpus_clean, max.words = 50, random.order = FALSE)
```

Compare wordclouds between spam and ham.
```{r}
spam <- subset(sms_raw, type == "spam")
ham <- subset(sms_raw, type == "ham")
wordcloud(spam$text, max.words = 50, scale = c(3, 0.5))
```

```{r}
wordcloud(ham$text, max.words = 50, scale = c(3, 0.5))
```

##Preparation for Naive Bayes
Remove words from the matrix that appear less than 5 times.

```{r}
sms_freq_words <- findFreqTerms(sms_dtm_train, 5)
str(sms_freq_words)
```

Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.

```{r}
sms_dtm_freq_train <- sms_dtm_train[ , sms_freq_words]
sms_dtm_freq_test <- sms_dtm_test[ , sms_freq_words]
```

The naive bayes classifier works with categorical reatures, so we need to convert the matrix to “yes” and “no” categorical variables. To do this we’ll build a convert_counts function and apply it to our data.

```{r}
convert_counts <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}
```

This replaces values greater than 0 with yes, and values not greater than 0 with no. Let’s apply it to our data.

```{r}
sms_train <- apply(sms_dtm_freq_train, MARGIN = 2, convert_counts)
sms_test <- apply(sms_dtm_freq_test, MARGIN = 2, convert_counts)
```

The resulting matrixes will be character type, with cells indicating “yes” or “no” if the word represented by the column appears in the message represented by the row.

##Train Model on the Data.
Use the e1071 package to impliment the Naive Bayes algorithm on the data, and predict whether a message is likely to be spam or ham.

```{r}
library(e1071)

sms_classifier <- naiveBayes(sms_train, sms_train_labels)
```

##Predict and Evaluate the Model.

```{r}
sms_test_pred <- predict(sms_classifier, sms_test)
```

Evaluate the predition with the actual data using a crosstable from the gmodels package.

```{r}
library(gmodels)
CrossTable(sms_test_pred, sms_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

As shown in the table only 29/1379 messages were classified incorrectly. This means that the algorithm classifed the testing set as spam or ham with approx. 98% accuracy. That’s impressive. To improve the model, one might tamper with the Laplace value, colect more sms data, or try splitting the dataset randomly into training and testing.
I suspect the accuracy would increase as the dataset gets bigger. The more data there is to train the algorith, the more effective it would be in predicting Spam or Ham.

Show the 5 most frequent words in the sms data:
```{r}
sack <- TermDocumentMatrix(sms_corpus_clean)
pack <- as.matrix(sack)
snack <- sort(rowSums(pack), decreasing = TRUE)
hack <- data.frame(word = names(snack), freq=snack)
head(hack, 5)
```

And the 5 most frequent words from each class:

```{r}
wordcloud(spam$text, max.words = 5, scale = c(3, 0.5))

wordcloud(ham$text, max.words = 5, scale = c(3, 0.5))
```

As shown in the word clouds, the most frequent words from the spam messages are "call", "free", "now", "mobile", and "text". And in the Ham messages, the 5 most frequent words are "can", "get", "just", "will", and "now".

***
***

#My Test of Naive Bayes

##Loading Data Into R.
This block reads full CSV into a starting dataframe named "base_data". Next block isolates tables and rows into new dataframe "sms_raw2", and renames the columns.
```{r}
#CSV file named "coded_full_comments_dataset_Spring18anon.csv" already has correctly organized text and meta-data. First line reads it in.
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')

# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
frequency_writing <- filter(base_data, code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
sms_raw2 <- frequency_writing %>% select(23,22)

#Rename columns
names(sms_raw2)[1] <- "type"
names(sms_raw2)[2] <- "text"
```

```{r}
str(sms_raw2)
```

The "type"" element is a character. This changes it to a factor for the analysis.

```{r}
sms_raw2$type <- factor(sms_raw2$type)
str(sms_raw2$type)
table(sms_raw2$type)
```

##Text Mining
For Naive Bayes to run effectively, the test data needs to be transformed. This begins with using the "tm" package in R to create a volitile coprus that contains the "text" vector from our data frame.

```{r}
#install.packages("tm")
library(tm)
```

```{r}
sms_corpus2 <- VCorpus(VectorSource(sms_raw2$text))
print(sms_corpus2)
```

Check out the first few messages in the new corpus, which is basically a list that can be manipulated with list operations.

```{r}
inspect(sms_corpus2[1:3])
```

Use "as.character" function to see what a message looks like.

```{r}
as.character(sms_corpus2[[3]])
```

In order to standardize the messages, the data set must be tranformed to all lower case letters. The words "Free", "free", and "FREE" should all be treated as the same word. Use the "tm_map"" funtion in R, and use the "content_transformer" function to transform the text.

```{r}
sms_corpus_clean2 <- tm_map(sms_corpus2, content_transformer(tolower))
```

Look at third message again to see if our data was transformed.

```{r}
as.character(sms_corpus2[[3]])

as.character((sms_corpus_clean2[[3]]))
```

```{r}
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removeNumbers)
```

Remome words that appear often but don't contribute to our objective. These words include "to", "and", "but" and "or".

```{r}
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removeWords, stopwords())
```

Remove punctuation as well using the "removePunctuation" function.

```{r}
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removePunctuation)

as.character((sms_corpus_clean2[[3]]))
```

Perform "stemming" to the text data to strip the suffix from words like "jumping", so the words "jumping" "jumps" and "jumped" are all transformed into "jump". Stemming can be perfromed using the "tm" package with help from the "SnowballC" package.

```{r}
#install.packages("SnowballC")
#library(SnowballC)

sms_corpus_clean2 <- tm_map(sms_corpus_clean2, stemDocument)
```

And now the final step in text mining is to remove white space from the document.

```{r}
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, stripWhitespace)
```

```{r}
as.character(sms_corpus_clean2[[3]])
```

Perform tokenization using the "DocumentTermMatrix" function. This creates a matrix in which the rows indicate documents (SMS messages in this case) and the columns indicate words. It should be noted that the "DocumentTermMaxtrix" function has the power to do all of the text mining above in one command.

```{r}
sms_dtm2 <- DocumentTermMatrix(sms_corpus_clean2)
```

##Data Preparation
Split our data into training and testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data. Divide our data set into 75% training and 25% testing.

```{r}
.75 * 9129
.25 * 9129
```

Because the dataset is random, the first 4180 messages can be used for the training set - there's no need to randomize the data first.

```{r}
sms_dtm_train2 <- sms_dtm2[1:6846, ]
sms_dtm_test2 <- sms_dtm2[6847:9129, ]
```

Save vectors labeling rows in the training and testing vectors

```{r}
sms_train_labels2 <- sms_raw2[1:6846, ]$type
sms_test_labels2 <- sms_raw2[6847:9129,]$type
```

Make sure that the proportion of spam is similar in the training and testing data set.

```{r}
prop.table(table(sms_train_labels2))

prop.table(table(sms_test_labels2))
```

##Visualization
Create a wordcloud of the frequency of the words in the dataset using the package "wordcloud".

```{}
install.packages("wordcloud")
install.packages("RColorBrewer")
```

```{r}
#library(wordcloud)
wordcloud(sms_corpus_clean2, max.words = 50, random.order = FALSE)
```

Compare wordclouds between 3 groups.

```{r}
TWO <- subset(sms_raw2, type == "2. Writing Quality")
THREE <- subset(sms_raw2, type == "3. Technical and Scientific")
FOUR <- subset(sms_raw2, type == "4. Logic and Thinking")

wordcloud(TWO$text, max.words = 50, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 50, scale = c(3, 0.5))
wordcloud(FOUR$text, max.words = 50, scale = c(3, 0.5))
```

##Preparation for Naive Bayes
Remove words from the matrix that appear less than 5 times.

```{r}
sms_freq_words2 <- findFreqTerms(sms_dtm_train2, 5)
str(sms_freq_words2)
```

Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.

```{r}
sms_dtm_freq_train2 <- sms_dtm_train2[ , sms_freq_words2]
sms_dtm_freq_test2 <- sms_dtm_test2[ , sms_freq_words2]
```

The naive bayes classifier works with categorical reatures, so we need to convert the matrix to "yes" and "no" categorical variables. To do this we'll build a convert_counts function and apply it to our data.

```{r}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}
```

This replaces values greater than 0 with yes, and values not greater than 0 with no. Let's apply it to our data.

```{r}
sms_train2 <- apply(sms_dtm_freq_train2, MARGIN = 2, convert_counts2)

sms_test2 <- apply(sms_dtm_freq_test2, MARGIN = 2, convert_counts2)
```

The resulting matrixes will be character type, with cells indicating "yes" or "no" if the word represented by the column appears in the message represented by the row.

##Train Model on the Data.
Use the e1071 package to impliment the Naive Bayes algorithm on the data, and predict whether a message is likely to be spam or ham.

```{}
install.packages("e1071")
library(e1071)
```

```{r}
sms_classifier2 <- naiveBayes(sms_train2, sms_train_labels2)
```

##Predict and Evaluate the Model.

```{r}
sms_test_pred2 <- predict(sms_classifier2, sms_test2)
```

Evaluate the predition with the actual data using a crosstable from the gmodels package.

```{r}
#library(gmodels)
CrossTable(sms_test_pred2, sms_test_labels2, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

```{r}
1352/1379
```

As shown in the table only 27/1379 messages were classified incorrectly. This means that the algorithm classifed the testing set as spam or ham with approx. 98% accuracy. That's impressive. To improve the model, one might tamper with the Laplace value, colect more sms data, or try splitting the dataset randomly into training and testing.

I suspect the accuracy would increase as the dataset gets bigger. The more data there is to train the algorith, the more effective it would be in predicting Spam or Ham.

Show the 5 most frequent words in the sms data:

```{r}
sack2 <- TermDocumentMatrix(sms_corpus_clean2)
pack2 <- as.matrix(sack2)
snack2 <- sort(rowSums(pack2), decreasing = TRUE)
hack2 <- data.frame(word = names(snack2), freq=snack2)
head(hack2, 5)
```

And the 5 most frequent words from each class:

```{r}
wordcloud(TWO$text, max.words = 5, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 5, scale = c(3, 0.5))
wordcloud(FOUR$text, max.words = 5, scale = c(3, 0.5))
```

### 2018-12-18 draft.Rmd
update

@@
 ---
-title: "Data Intake & Initial Exploration"
+title: "Draft Description of Work in Progress"
 author: "Dan Johnson"
-date: "12/16/2018"
+date: "12/17/2018"
 output: html_document
@@
 ```
+#Overview
+This page provides a draft overview of my progress.
+A short narrative outlines the  discrete blocks of work and how they are connected.
+Links connect to longer pages that document intake, munging, analysis, and visualization in depth.
 
-The initial working dataset is an anonymized Excel table of 10,808 rows.
-One column has "TA comments" extracted from student lab reports.
+####\
+
+##Initial Dataset
+My working dataset is an Excel table of 10,808 rows.
+One column has "TA comments" extracted from student lab reports in Spring 2018.
 Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation.
@@
 
+Data were anonymized in Excel prior to conversion to CSV format. Names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings.
+
 ####\
+
+##Blocks of Work
+
+*  Data Pre-Checks and Import
+*  Initial Import and Post-Checks
+*  Preliminary Exploration: Log-Odds Text Comparison
+    +  Pearson Correlations of Word Frequencies: BROKEN SECTION
+    +  Mapping Analyses Visually: RUNS BUT CANNOT COPY AND MODIFY. It depends tables created using the broken code above. So I cannot run this part either.
+    +  Calculating and Plotting Log-Odds Pairs
+*  Preliminary Exploration: Direct Comparison of Word Lists
+*  Deeper Text Analysis Methods
+    +  Naive Bayes [/docs/Naive_bayes.html](link)
+    +  Random Forest (link)
+    +  Latent Dirichlet Allocation (link)
+    +  Etc. (link)
+    +  Etc. (link)
+
+####\
+
 ##Data Pre-Checks
@@
 
-1.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.md"
+1.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.rmd"
 2.  Data have been de-identified and recoded using pre-assigned anonymous IDs.
-3.  Data table headers are in snake_case, and match tabulated list below. The same list is in the file "code_column_names.txt"
+3.  Data table headers are in snake_case, and match tabulated list below. The same list is in the file "code_column_names.rmd"
 
@@
 ##Initial Import and Post-Check
-```{r Libraries}
+```{r Libraries_1}
 # Call in tidyverse, other required libraries.
@@
 
-#If the TA comments dataset is imported correctly, there will be a dataframe of ~11,000 rows and 28 columns named as listed in the **Data Pre-Checks** table above.
+#If dataset imports correctly, there will be a dataframe of ~11,000 rows and 28 columns named as listed in the **Data Pre-Checks** table above.
 ```
 
-Given some data points may be entered incorrectly by instructors or students, I included several post-import data checks to ensure data have been properly coded and columns contain valid entries only.
+Given some data points may be entered incorrectly by instructors or students, these post-import data checks help ensure data have been properly coded and columns contain valid entries only.
 
@@
 ###Log-Odds Text Comparison
-The log-odds ratios method outlined below has been the most informative to date.
-The basic premise being tested is that certain words are more frequent in TA comments belonging to one sub-category, and less frequent in others.
-I first saw this approach as a way to compare word usage between separate Twitter accounts here: https://www.tidytextmining.com/twitter.html.
+The log-odds ratios method outlined below finds and displays differences words that are more frequently found in TA comments belonging to one sub-category vs.
+another.
+I based this on a method for comparing word usage frequencies between two separate Twitter users [https://www.tidytextmining.com/twitter.html](that I found here.)
 
-For simplicity I reduced the analysis to just the four most common comment subjects.
+For simplicity I reduced my analysis to just the four most common comment subjects. This has been the most informative preliminary visualization I have found so far.
 
@@
 It goes from the correctly sorted 4 columns to incorrectly sorted 3 columns.
-It is NOT A TYPO; I can literally copy/paste it in and out and break/correct the code.
+It is NOT A TYPO; I can literally copy/paste the names in and out and break/correct the code.
 
-I cannot use a different order, or any other format. I am at a loss.
+I cannot fix it using a different order for variables. I am at a loss for why it breaks.
 
@@
 
-***
-***
+I can run this block, but CANNOT COPY AND MODIFY THIS SECTION. It needs the tables that should be created using the broken code above. So I cannot run this part either.
 
-I can run this block, but CANNOT COPY AND MODIFY THIS SECTION. It depends tables created using the broken code above. So I cannot run this part either.
-
-```{r}
+```{}
 #Create groups by code.subject, calculates proportional frequency in each group.
@@
 
-```{r}
+```{}
 base_data_tokenized_sorted2v3 <- base_data_tokenized_sorted2 %>% 
@@
 
-```{r}
+```{}
 ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`3. Technical and Scientific`)) +
@@
   geom_abline(color = "red")
-  ggsave("Freqplot_3vs2.png", width = 10, height = 6)
+  #ggsave("Freqplot_3vs2.png", width = 10, height = 6)
 ```
 
-```{r}
+```{}
 #I can get part way around the problem by re-using the same data and calling a different subset for the Y axis.
@@
 
-```{r}
+```{}
 ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`4. Logic and Thinking`)) +
@@
   geom_abline(color = "red")
-  ggsave("Freqplot_4vs2.png", width = 10, height = 6)
+  #ggsave("Freqplot_4vs2.png", width = 10, height = 6)
 ```
-
-END of broken / limping section
+END of broken / limping sections
 ***
@@
 
-####Calculating and Plotting Log-Odds Pairs
-**2. Writing versus 3. Technical**
+###Calculating and Plotting Log-Odds Pairs
 
@@
   arrange(desc(logratio))
+
+word_ratios3v4 <- base_data_tokenized %>%
+  count(word, code.subject) %>%
+  group_by(word) %>%
+  filter(sum(n) >= 10) %>%
+  ungroup() %>%
+  spread(code.subject, n, fill = 0) %>%
+  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
+  mutate(logratio = log(`3. Technical and Scientific`/`4. Logic and Thinking`)) %>%
+  arrange(desc(logratio))
+
+word_ratios2v4 <- base_data_tokenized %>%
+  count(word, code.subject) %>%
+  group_by(word) %>%
+  filter(sum(n) >= 10) %>%
+  ungroup() %>%
+  spread(code.subject, n, fill = 0) %>%
+  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
+  mutate(logratio = log(`2. Writing Quality`/`4. Logic and Thinking`)) %>%
+  arrange(desc(logratio))
 ```
@@
 word_ratios2v3 %>% 
+  arrange(abs(logratio))
+
+word_ratios3v4 %>% 
+  arrange(abs(logratio))
+
+word_ratios2v4 %>% 
   arrange(abs(logratio))
@@
 
-```
-
-**3. Technical vs. 4. Logic**
-
-```{r}
-word_ratios3v4 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`3. Technical and Scientific`/`4. Logic and Thinking`)) %>%
-  arrange(desc(logratio))
-```
-
-```{r}
-word_ratios3v4 %>% 
-  arrange(abs(logratio))
-```
-
-```{r}
 word_ratios3v4 %>%
@@
 
-```
-
-**2. Writing versus 4. Logic**
-
-```{r}
-word_ratios2v4 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`2. Writing Quality`/`4. Logic and Thinking`)) %>%
-  arrange(desc(logratio))
-```
-
-```{r}
-word_ratios2v4 %>% 
-  arrange(abs(logratio))
-```
-
-```{r}
 word_ratios2v4 %>%
@@
 
+####\
+
 ##Direct Comparison Using Only Words Appearing in All Subsets of Comments
-I think this is less informative, but it does let me see the data more clearly.Basic strategy is to take all comments from base_data, unnest tokens, filter stopwords, count and sort remaining terms, then write back to a new datatable called 'sortedwords.comments.all'.
+I think this is less informative, but it does let me see the data more clearly.
+
+Basic strategy is to take all comments from base_data, unnest tokens, filter stopwords, count and sort remaining terms, then write back to a new datatable called 'sortedwords.comments.all'.
 Calculate TOTAL words and store.
@@
 
-This is an early version, so likely has some poorly developed code blocks that could be refactored.
+I wrote this block so that, once CSV is imported, it does not require any of the code above. This is an early effort, so likely has some poorly developed code that could be refactored.
+
+```{}
+# Call in tidyverse, other required libraries.
+library(tidyverse)
+library(tidytext)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(scales)
+
+#Read in the full CSV into a starting dataframe named "base_data".
+#Dataframe should have ~11,000 rows and 28 columns .
+base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
+```
 
 ```{r}
-library(dplyr)
-library(tidytext)
-
-sortedwords.comments.all <- base_data_tokenized %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate the number of words in total in the subset
-total_words.all <- sortedwords.comments.all %>% 
-  summarize(total = sum(n))
-
-#Mutate table to calculate then append fractional word frequency in the text.
-sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
-  all.fraction = (n*100)/total_words.all$total
-  )
-```
-
-##Summarize Word Frequencies in Subsets of Comments Based on 'Subject'
-Take subsets of comments from base_data_tokenized using code.subject as separating characteristic.
-
-###Other Analysis Approaches
-This approach summarizes the word frequencies in the FULL set of TA comments first, then breaks them down into sub-groups. It has examples of other code bits as well.
-
-The first block takes all comments from base_data, unnests tokens, filters stopwords, count and sort remaining terms, then writes results back to a new dataframe called 'sortedwords.comments.all'.
-
-TOTAL words is calculated and stored. Then fractional frequencies for each word are calculated and stored back to the dataframe using mutate.
-
-```{r}
-# Tokenize phrases, remove stop words listed in standard reference file.
+# Tokenize phrases, remove the stop words listed in R's standard reference file.
 base_data_tokenized <- base_data %>% 
@@
 
-# An alternative approach to pulling out stopwords  
+# This is an ALTERNATIVE METHOD for pulling out stopwords.  
 sortedwords.comments.all <- base_data_tokenized %>%
@@
 )
-```
 
-```{r}
 sortedwords.technical <- base_data_tokenized.technical %>%
@@
 )
-```
 
-```{r}
 sortedwords.logic <- base_data_tokenized.logic %>%
@@
 
-Next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed.
+Next block joins the individual data tables for each subset in a larger file using column named 'word'.
+Columns get confusing names so they are renamed.
+**WARNING**: if a word is MISSING from **any subset** the data for that word is pruned from the full dataset.
+I am losing the unique words this way.
 
-**WARNING**: Words that are MISSING from **any subset** cause the full set to be pruned. I am losing the unique words this way. **Need to figure out how to join without pruning.**
+**Need to figure out how to join without pruning.**
 
@@
 
-The following six graphs only compare frequencies for ~770 words that appear in ALL THREE of the sub-categories.
+The following six graphs compare frequencies for ~770 words that appear in ALL THREE of the sub-categories.
+
 Graphs 1-3 compare frequencies in individual groups against word frequency overall.
@@
   geom_abline(color = "red")
-```
 
-```{r}
 ggplot(total6, aes(`all.fraction`,`technical.fraction`)) +
@@
   geom_abline(color = "red")
-```
 
-```{r}
 ggplot(total6, aes(`all.fraction`,`logic.fraction`)) +
@@
 Graphs 4-6 are pair-wise comparisons of frequencies for sub-categories.
-Unlike the comparisons to overall frequency, most words are NOT clustered on the trendline, suggesting there are differences in word usage.
+Unlike the comparisons to overall frequency, most words are NOT clustered along the trendline, suggesting there are greater differences in word usage frequencies.
 
@@
   geom_abline(color = "red")
-```
 
-```{r}
 ggplot(total6, aes(`logic.fraction`,`writing.fraction`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+
+ggplot(total6, aes(`logic.fraction`,`technical.fraction`)) +
   geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
@@
 
-```{r}
-ggplot(total6, aes(`logic.fraction`,`technical.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-```
+####\

### 2018-12-18 Naive_bayes.Rmd
update

@@
 
-#Demonstration of Applying Naive Bayes Method
+#Demo of Naive Bayes (NB) Method
+I copied the original demo web site [https://rpubs.com/Billyhansen6/318412](located here) to show how the process should work. I then duplicated the code and modified it to test with my TA comments.
 
+The initial trial was able to identify sub-categories of comments ~80% of the time.
+This is better than the inter-rater reliability of two human raters, but still low.
+For subsequent iterations of the NB method I plan to evaluate:
+
+*  Variations on data pre-processing: stemmed versus unstemmed, numbers removed vs. not, etc.
+*  Comparing n-grams rather than single words
+*  Pair-wise comparison versus comparing multiple groups simultaneously
+*  Modifying the variables within the NB analysis itself.
+
+####\
+
+##Hansen's Original Version
 The objective of this project is to classify SMS messages as spam or ham (not spam).
@@
 
-P(Spam|Hospital) = P(Hospital|Spam)P(Spam)/P(Hospital)
+$$P(Spam|Hospital) = P(Hospital|Spam)P(Spam)/P(Hospital)$$
 
@@
 
-##Loading Data Into R.
+####\
+
+##Loading Libraries and Data
 
 ```{r}
-sms_raw <- read.table("~/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/data/Bayes/SMSSpamCollection.txt", header = FALSE, sep = "\t", quote = "", col.names = c("type","text"), stringsAsFactors = FALSE)
+library(tidyverse)
+library(tidytext)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(scales)
+library(tm)
+library(SnowballC)
+library(wordcloud)
+library(e1071)
+library(gmodels)
+library(readr)
 ```
 
-The data frame has 5574 observations with 2 observations of either spam or ham.
+```{r}
+#Demo dataset of SMS messages
+sms_raw <- read.table("~/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/data/SMSSpamCollection.txt", header = FALSE, sep = "\t", quote = "", col.names = c("type","text"), stringsAsFactors = FALSE)
+```
+
+The data frame has 5574 observations marked as either spam or ham.
 
@@
 
-Now the "type" variable is a factor with 2 levels. Of the 5574 messages, 747 are spam, which is about 13.4%.
+Now the "type" variable is a factor with 2 levels. Of 5574 messages, 747 (about 13.4%) are spam.
+
+####\
 
 ##Text Mining
-
-For Naive Bayes to run effectively, the test data needs to be transformed. This begins with using the "tm" package in R to create a volitile coprus that contains the "text" vector from our data frame.
-
-```{r}
-#install.packages("tm")
-library(tm)
-```
+For Naive Bayes to run effectively, the test data needs to be transformed. This begins with using "tm" to create a volitile coprus that contains the "text" vector from our data frame.
 
@@
 as.character(sms_corpus[[3]])
-
 as.character((sms_corpus_clean[[3]]))
@@
 ```{r}
-#install.packages("SnowballC")
-library(SnowballC)
-
 sms_corpus_clean <- tm_map(sms_corpus_clean, stemDocument)
@@
 ```
+
+####\
 
@@
 
-sms_test_labels
-      ham      spam 
-0.8694706 0.1305294
+Each have approx. 13% spam.
 
-Each have approx. 13% spam.
+####\
 
@@
 ```{r}
-library(wordcloud)
 wordcloud(sms_corpus_clean, max.words = 50, random.order = FALSE)
@@
 wordcloud(spam$text, max.words = 50, scale = c(3, 0.5))
+wordcloud(ham$text, max.words = 50, scale = c(3, 0.5))
 ```
 
-```{r}
-wordcloud(ham$text, max.words = 50, scale = c(3, 0.5))
-```
+####\
 
@@
 
+####\
+
 ##Train Model on the Data.
@@
 ```{r}
-library(e1071)
-
 sms_classifier <- naiveBayes(sms_train, sms_train_labels)
 ```
+
+####\
 
@@
 ```{r}
-library(gmodels)
 CrossTable(sms_test_pred, sms_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
@@
 
-As shown in the table only 29/1379 messages were classified incorrectly.
+As shown in the table only 27/1379 messages were classified incorrectly.
 This means that the algorithm classifed the testing set as spam or ham with approx.
@@
 wordcloud(spam$text, max.words = 5, scale = c(3, 0.5))
-
 wordcloud(ham$text, max.words = 5, scale = c(3, 0.5))
@@
 #My Test of Naive Bayes
-
-##Loading Data Into R.
+##Loading Data
 This block reads full CSV into a starting dataframe named "base_data". Next block isolates tables and rows into new dataframe "sms_raw2", and renames the columns.
 ```{r}
-#CSV file named "coded_full_comments_dataset_Spring18anon.csv" already has correctly organized text and meta-data. First line reads it in.
-base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
-
-# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
-frequency_writing <- filter(base_data, code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
-sms_raw2 <- frequency_writing %>% select(23,22)
-
-#Rename columns
-names(sms_raw2)[1] <- "type"
-names(sms_raw2)[2] <- "text"
+#library(tm)
+#library(SnowballC)
+#library(wordcloud)
+#library(e1071)
+#library(gmodels)
+#library(readr)
 ```
@@
 ```{r}
-str(sms_raw2)
-```
+#Read in CSV file named "coded_full_comments_dataset_Spring18anon.csv".
+base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
 
-The "type"" element is a character. This changes it to a factor for the analysis.
+#Isolate rows to compare. 
+frequency_writing <- filter(base_data,code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
 
-```{r}
+#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
+sms_raw2 <- frequency_writing %>% select(23,22)
+
+#Rename the columns.
+names(sms_raw2)[1] <- "type"
+names(sms_raw2)[2] <- "text"
+
+#Simplify the coding terms
+sms_raw2[,1] <- ifelse(sms_raw2[,1] == "2.
+Writing Quality", "2_Writing", ifelse(sms_raw2[,1] == "3.
+Technical and Scientific", "3_Technical", ifelse(sms_raw2[,1] == "4.
+Logic and Thinking","4_Logic",99)))
+
+#Change "type" element from character to a factor for analysis.
 sms_raw2$type <- factor(sms_raw2$type)
@@
 
-##Text Mining
-For Naive Bayes to run effectively, the test data needs to be transformed. This begins with using the "tm" package in R to create a volitile coprus that contains the "text" vector from our data frame.
+####\
+
+##Text Transformation Steps
+This begins with using "tm" package in R to create a volitile coprus that contains the "text" vector from our data frame.
 
 ```{r}
-#install.packages("tm")
-library(tm)
-```
-
-```{r}
+#library(tm)
 sms_corpus2 <- VCorpus(VectorSource(sms_raw2$text))
@@
 
+This removes numbers. **May not want to do.**
 ```{r}
@@
 
-Remome words that appear often but don't contribute to our objective. These words include "to", "and", "but" and "or".
+Stopword removal takes out words that appear often but don't contribute to our objective. These words include "to", "and", "but" and "or". **One of variables to test out.**
 
@@
 
-Remove punctuation as well using the "removePunctuation" function.
+Remove punctuation as well using the "removePunctuation" function. **Try not doing this.**
 
@@
 
-Perform "stemming" to the text data to strip the suffix from words like "jumping", so the words "jumping" "jumps" and "jumped" are all transformed into "jump".
-Stemming can be perfromed using the "tm" package with help from the "SnowballC" package.
+Perform "stemming" to the text data to strip the suffix from words. **Try NOT doing this.**
 
 ```{r}
-#install.packages("SnowballC")
 #library(SnowballC)
-
 sms_corpus_clean2 <- tm_map(sms_corpus_clean2, stemDocument)
@@
 
+####\
+
 ##Data Preparation
@@
 
-Because the dataset is random, the first 4180 messages can be used for the training set - there's no need to randomize the data first.
+This code assumes comments are random. **Probably want to try randomizing them.**
 
@@
 
-Make sure that the proportion of spam is similar in the training and testing data set.
+Make sure that the proportion of each sub-category is similar in the training and testing data set.
 
@@
 prop.table(table(sms_train_labels2))
-
 prop.table(table(sms_test_labels2))
@@
 
+####\
+
 ##Visualization
 Create a wordcloud of the frequency of the words in the dataset using the package "wordcloud".
-
-```{}
-install.packages("wordcloud")
-install.packages("RColorBrewer")
-```
 
@@
 ```{r}
-TWO <- subset(sms_raw2, type == "2. Writing Quality")
-THREE <- subset(sms_raw2, type == "3. Technical and Scientific")
-FOUR <- subset(sms_raw2, type == "4. Logic and Thinking")
+TWO <- subset(sms_raw2, type == "2_Writing")
+THREE <- subset(sms_raw2, type == "3_Technical")
+FOUR <- subset(sms_raw2, type == "4_Logic")
 
@@
 
-The naive bayes classifier works with categorical reatures, so we need to convert the matrix to "yes" and "no" categorical variables.
+The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables.
 To do this we'll build a convert_counts function and apply it to our data.
@@
 sms_train2 <- apply(sms_dtm_freq_train2, MARGIN = 2, convert_counts2)
-
 sms_test2 <- apply(sms_dtm_freq_test2, MARGIN = 2, convert_counts2)
@@
 
-##Train Model on the Data.
-Use the e1071 package to impliment the Naive Bayes algorithm on the data, and predict whether a message is likely to be spam or ham.
+####\
 
-```{}
-install.packages("e1071")
-library(e1071)
-```
+##Train Model, Predict, Evaluate
+Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group TWO, THREE, or FOUR.
+Evaluate the prediction with the actual data using a crosstable from the gmodels package.
 
 ```{r}
+#library(e1071)
+#library(gmodels)
 sms_classifier2 <- naiveBayes(sms_train2, sms_train_labels2)
-```
-
-##Predict and Evaluate the Model.
-
-```{r}
 sms_test_pred2 <- predict(sms_classifier2, sms_test2)
-```
-
-Evaluate the predition with the actual data using a crosstable from the gmodels package.
-
-```{r}
-#library(gmodels)
 CrossTable(sms_test_pred2, sms_test_labels2, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
@@
 
-```{r}
-1352/1379
-```
-
-As shown in the table only 27/1379 messages were classified incorrectly.
-This means that the algorithm classifed the testing set as spam or ham with approx.
-98% accuracy.
-That's impressive.
-To improve the model, one might tamper with the Laplace value, colect more sms data, or try splitting the dataset randomly into training and testing.
-
-I suspect the accuracy would increase as the dataset gets bigger. The more data there is to train the algorith, the more effective it would be in predicting Spam or Ham.
-
-Show the 5 most frequent words in the sms data:
+Show the 5 most frequent words in the data:
 

### 2018-12-18 draft.Rmd
update

@@
 #Overview
-This page provides a draft overview of my progress.
-A short narrative outlines the  discrete blocks of work and how they are connected.
-Links connect to longer pages that document intake, munging, analysis, and visualization in depth.
+This page begins my draft outline of progress to date. Initial data import and exploration are described here. The other pages listed below document different pieces of the larger project.
+
+##Import and Initial Exploration
+This page describes to date on:
+
+*  Data pre-checks, import, and post-import validation process.
+*  Initial comparisons of sub-categories using log-odds ratios of word frequencies
+    +  Pearson Correlations of Word Frequencies: BROKEN SECTION
+    +  Mapping Analyses Visually: RUNS BUT CANNOT COPY AND MODIFY. It depends tables created using the broken code above. So I cannot run this part either.
+    +  Calculating and Plotting Log-Odds Pairs
+*  Direct comparison of word frequency lists
+
+##Deeper Text Analysis
+These linked pages summarize work to date on each method.
+
+*  Naive Bayes: https://adanieljohnson.github.io/default_website/Naive_bayes.html
+*  Random Forest (link)
+*  Latent Dirichlet Allocation (link)
+*  Etc. (link)
+*  Etc. (link)
 
@@
 
-##Initial Dataset
+#Initial Dataset
 My working dataset is an Excel table of 10,808 rows.
@@
 
-##Blocks of Work
-
-*  Data Pre-Checks and Import
-*  Initial Import and Post-Checks
-*  Preliminary Exploration: Log-Odds Text Comparison
-    +  Pearson Correlations of Word Frequencies: BROKEN SECTION
-    +  Mapping Analyses Visually: RUNS BUT CANNOT COPY AND MODIFY. It depends tables created using the broken code above. So I cannot run this part either.
-    +  Calculating and Plotting Log-Odds Pairs
-*  Preliminary Exploration: Direct Comparison of Word Lists
-*  Deeper Text Analysis Methods
-    +  Naive Bayes [/docs/Naive_bayes.html](link)
-    +  Random Forest (link)
-    +  Latent Dirichlet Allocation (link)
-    +  Etc. (link)
-    +  Etc. (link)
-
 ####\
 
-##Data Pre-Checks
+#Data Pre-Checks
 Prior to import, review data table in Excel. Check:

### 2018-12-21 Naive_bayes.Rmd
Update all

--- 2018-12-18 Naive_bayes.Rmd
@@
 ---
-title: "Naive Bayes"
+title: "Naive Bayes Workflow"
 author: "Dan Johnson"
-date: "12/16/2018"
+date: "12/21/2018"
 output: html_document
@@
 
-#Demonstration of Applying Naive Bayes Method
-
-The objective of this project is to classify SMS messages as spam or ham (not spam).
+##Summary
+This notebook documents a basic workflow for applying a naive Bayes model to terms in TA comments.
+I based it on a tutorial [https://rpubs.com/Billyhansen6/318412](located here).
+This is quoted from Hansen's original tutorial:
+>The objective of this project is to classify SMS messages as spam or ham (not spam).
 A Naive Bayes classifier approach will be used.
@@
 
-An example of the conditional probability that will be computed is as follows:
+>An example of the conditional probability that will be computed is as follows:
 
-P(Spam|Hospital) = P(Hospital|Spam)P(Spam)/P(Hospital)
+$$P(Spam|Hospital) = P(Hospital|Spam)P(Spam)/P(Hospital)$$
 
-which is the formula for determining the probability that a message is spam given that it contains the word "Hospital" in the message.
+>which is the formula for determining the probability that a message is spam given that it contains the word "Hospital" in the message.
 
-##Loading Data Into R.
+Lines ~30-205 are the baseline version of the workflow using single words.
+It correctly classifies comments in the top 4 sub-categories  ~79% of the time.
+While that is close to minimum acceptable inter-rater reliability of two human readers, it is well below my goal of >95% for use in routine automated analysis.
 
+Lines ~208-290 describe an alternative method using n-grams. When I tested this approach using 2-grams, classification was slightly worse, at 68.3% correct.
+
+Lines ~295-335 show how to use word clouds to visualize single word frequencies. I did not find this approach very useful.
+
+To keep this code clean for now I will run various permutations in a scratch version, then summarize findings starting on line #345. In these subsequent iterations I will evaluate:
+
+*  Different text pre-processing strategies: stemmed versus unstemmed, numbers removed vs. not, etc.
+*  Using n-grams rather than single words
+*  Making pair-wise comparison versus comparing multiple groups simultaneously
+*  Modifying the variables in the NB analysis itself.
+
+Finally, there are several things still missing or needed for this workflow:
+
+*  How to feed a text dataset in, and get a table of the predicted categories back out.
+*  How to store the values from the Bayes prediction table so I can use them in calculations rather than type them in by hand.
+*  How to randomize which comments are used for training, testing datasets.
+
+####\
+
+##Initial Setup
 ```{r}
-sms_raw <- read.table("~/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/data/Bayes/SMSSpamCollection.txt", header = FALSE, sep = "\t", quote = "", col.names = c("type","text"), stringsAsFactors = FALSE)
+library(tidyverse)
+library(tidytext)
+library(tidyr)
+library(tm)
+library(SnowballC)
+library(wordcloud)
+library(e1071)
+library(gmodels)
 ```
 
-The data frame has 5574 observations with 2 observations of either spam or ham.
+####\
 
+##Baseline Version 1.0 of Naive Bayes (NB) Protocol
+###Initial Data Input
 ```{r}
-str(sms_raw)
+#Read in TA comments from CSV file.
+base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
+
+#Select rows representing the sub-groups to compare. 
+comments_subset <- filter(base_data,code.subject=="1. Basic Criteria"|code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
+
+#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
+comments_raw <- comments_subset %>% select(23,22)
+
+#Rename the columns.
+names(comments_raw)[1] <- "type"
+names(comments_raw)[2] <- "text"
+
+#Simplify coding terms
+comments_raw[,1] <- ifelse(comments_raw[,1] == "1.
+Basic Criteria","1_Basic", ifelse(comments_raw[,1] == "2.
+Writing Quality","2_Writing", ifelse(comments_raw[,1] == "3.
+Technical and Scientific","3_Technical", ifelse(comments_raw[,1] == "4.
+Logic and Thinking","4_Logic",99))))
+
+#Change "type" element from character to a factor for analysis.
+comments_raw$type <- factor(comments_raw$type)
+str(comments_raw$type)
+table(comments_raw$type)
 ```
 
-As seen above, the "type"" element is a character.Change it to a factor for the analysis.
+Data set must be converted to a volative corpus using "tm" library then transformed.
+```{r}
+#Create the volatile coprus that contains the "text" vector from data frame.
+comments_corpus <- VCorpus(VectorSource(comments_raw$text))
+print(comments_corpus)
 
-```{r}
-sms_raw$type <- factor(sms_raw$type)
-str(sms_raw$type)
-table(sms_raw$type)
+#Check out the first few messages in the new corpus, which is basically a list that can be manipulated with list operations.
+inspect(comments_corpus[1:3])
+
+#Use "as.character" function to see what a message looks like.
+as.character(comments_corpus[[3]])
 ```
 
-Now the "type" variable is a factor with 2 levels. Of the 5574 messages, 747 are spam, which is about 13.4%.
-
-##Text Mining
-
-For Naive Bayes to run effectively, the test data needs to be transformed. This begins with using the "tm" package in R to create a volitile coprus that contains the "text" vector from our data frame.
-
+The OPTIONAL data transforms using "tm_map"" and "content_transformer" functions. Switch between "{r}" and "{}" to turn a transformation block off or on.
 ```{r}
-#install.packages("tm")
-library(tm)
+#Convert to all lower case letters.
+comments_corpus_clean <- tm_map(comments_corpus, content_transformer(tolower))
 ```
@@
 ```{r}
-sms_corpus <- VCorpus(VectorSource(sms_raw$text))
-print(sms_corpus)
-```
-
-Check out the first few messages in the new corpus, which is basically a list that can be manipulated with list operations.
-
-```{r}
-inspect(sms_corpus[1:3])
-```
-
-Use "as.character" function to see what a message looks like.
-
-```{r}
-as.character(sms_corpus[[3]])
-```
-
-In order to standardize the messages, the data set must be tranformed to all lower case letters.
-The words "Free", "free", and "FREE" should all be treated as the same word.
-Use the "tm_map"" funtion in R, and use the "content_transformer" function to transform the text.
-
-```{r}
-sms_corpus_clean <- tm_map(sms_corpus, content_transformer(tolower))
-```
-
-Look at third message again to see if our data was transformed.
-
-```{r}
-as.character(sms_corpus[[3]])
-
-as.character((sms_corpus_clean[[3]]))
-```
-
-The message was indeed tranformed to all lowercase letters. Now to remove the numbers using the "removeNumbers" function.
-
-```{r}
-sms_corpus_clean <- tm_map(sms_corpus_clean, removeNumbers)
-```
-
-Remome words that appear often but don't contribute to our objective. These words include "to", "and", "but" and "or".
-
-```{r}
-sms_corpus_clean <- tm_map(sms_corpus_clean, removeWords, stopwords())
-```
-
-Remove punctuation as well using the "removePunctuation" function.
-
-```{r}
-sms_corpus_clean <- tm_map(sms_corpus_clean, removePunctuation)
-
-as.character((sms_corpus_clean[[3]]))
-```
-
-Perform "stemming" to the text data to strip the suffix from words like "jumping", so the words "jumping" "jumps" and "jumped" are all transformed into "jump".
-Stemming can be perfromed using the "tm" package with help from the "SnowballC" package.
-
-```{r}
-#install.packages("SnowballC")
-library(SnowballC)
-
-sms_corpus_clean <- tm_map(sms_corpus_clean, stemDocument)
-```
-
-And now the final step in text mining is to remove white space from the document.
-
-```{r}
-sms_corpus_clean <- tm_map(sms_corpus_clean, stripWhitespace)
+#Remove numbers. 
+comments_corpus_clean <- tm_map(comments_corpus_clean, removeNumbers)
 ```
@@
 ```{r}
-as.character(sms_corpus_clean[[3]])
+#Stopword removal. 
+comments_corpus_clean <- tm_map(comments_corpus_clean, removeWords, stopwords())
 ```
 
-Perform tokenization using the "DocumentTermMatrix" function.
-This creates a matrix in which the rows indicates documents (SMS messages in this case) and the columns indicate words.
-It should be noted that the "DocumentTermMaxtrix" function has the power to do all of the text mining above in one command.
+```{r}
+#Remove punctuation using the "removePunctuation" function. 
+#This step removes evidence of questions, so may remove data.
+comments_corpus_clean <- tm_map(comments_corpus_clean, removePunctuation)
+```
 
 ```{r}
-sms_dtm <- DocumentTermMatrix(sms_corpus_clean)
+#Stemming the text data to strip the suffix from words.  
+comments_corpus_clean <- tm_map(comments_corpus_clean, stemDocument)
 ```
 
-##Data Preparation
-Split our data into training and testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data. Divide our data set into 75% training and 25% testing.
+Final prep and data check.
+```{r}
+#Final step removes white space from the document. This is NOT optional.
+comments_corpus_clean <- tm_map(comments_corpus_clean, stripWhitespace)
+
+#Look at an example of data.
+as.character((comments_corpus_clean[[3]]))
+```
+
+####\
+
+###Basic Tokenizer
+This version uses the "DocumentTermMatrix" function to creates a matrix in which the rows indicate individual documents (individual TA comments) and the columns indicate presence or absence of the word that is the header for the column.
+Be careful not to confuse it with "TermDocumentMatrix", which transposes the matrix.
+
+"DocumentTermMaxtrix" function also can perform all of the text prep above in one command.
 
 ```{r}
-.75 * 5574
-.25 * 5574
+comments_corpus_1gram <- DocumentTermMatrix(comments_corpus_clean)
+comments_corpus_1gram_tidy <- tidy(comments_corpus_1gram)
+inspect(comments_corpus_1gram)
 ```
 
-Because the dataset is random, the first 4180 messages can be used for the training set - there’s no need to randomize the data first.
+####\
 
+###Data Preparation
+Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data. 
 ```{r}
-sms_dtm_train <- sms_dtm[1:4180, ]
-sms_dtm_test <- sms_dtm[4181:5559, ]
+.75 * 9340 #Number of documents/rows in the dataset; product is #rows for training
+.25 * 9340 #Product is #rows for testing set
+```
+
+This code assumes comments are random. **Probably want to try randomizing them.**
+```{r}
+comments_dtm_train <- comments_corpus_1gram[1:7005, ]
+comments_dtm_test <- comments_corpus_1gram[7006:9340, ]
 ```
@@
 Save vectors labeling rows in the training and testing vectors
-
 ```{r}
-sms_train_labels <- sms_raw[1:4180, ]$type
-sms_test_labels <- sms_raw[4181:5559,]$type
+comments_train_labels <- comments_raw[1:7005, ]$type
+comments_test_labels <- comments_raw[7006:9340,]$type
 ```
 
-Make sure that the proportion of spam is similar in the training and testing data set.
-
+Make sure that the proportion of each sub-category is similar in the training and testing data set.
 ```{r}
-prop.table(table(sms_train_labels))
+prop.table(table(comments_train_labels))
+prop.table(table(comments_test_labels))
 ```
 
+####\
+
+###Preparation for Naive Bayes
+Remove words from the matrix that appear less than 5 times.
 ```{r}
-prop.table(table(sms_test_labels))
-```
-
-sms_test_labels
-      ham      spam 
-0.8694706 0.1305294
-
-Each have approx. 13% spam.
-
-##Visualization
-Create a wordcloud of the frequency of the words in the dataset using the package “wordcloud”.
-
-```{r}
-library(wordcloud)
-wordcloud(sms_corpus_clean, max.words = 50, random.order = FALSE)
-```
-
-Compare wordclouds between spam and ham.
-```{r}
-spam <- subset(sms_raw, type == "spam")
-ham <- subset(sms_raw, type == "ham")
-wordcloud(spam$text, max.words = 50, scale = c(3, 0.5))
-```
-
-```{r}
-wordcloud(ham$text, max.words = 50, scale = c(3, 0.5))
-```
-
-##Preparation for Naive Bayes
-Remove words from the matrix that appear less than 5 times.
-
-```{r}
-sms_freq_words <- findFreqTerms(sms_dtm_train, 5)
-str(sms_freq_words)
+comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
+str(comments_freq_words)
 ```
@@
 Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
-
 ```{r}
-sms_dtm_freq_train <- sms_dtm_train[ , sms_freq_words]
-sms_dtm_freq_test <- sms_dtm_test[ , sms_freq_words]
+comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
+comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
 ```
 
-The naive bayes classifier works with categorical reatures, so we need to convert the matrix to “yes” and “no” categorical variables.
-To do this we’ll build a convert_counts function and apply it to our data.
-
-```{r}
-convert_counts <- function(x) {
-  x <- ifelse(x > 0, "Yes", "No")
-}
-```
-
-This replaces values greater than 0 with yes, and values not greater than 0 with no. Let’s apply it to our data.
-
-```{r}
-sms_train <- apply(sms_dtm_freq_train, MARGIN = 2, convert_counts)
-sms_test <- apply(sms_dtm_freq_test, MARGIN = 2, convert_counts)
-```
-
-The resulting matrixes will be character type, with cells indicating “yes” or “no” if the word represented by the column appears in the message represented by the row.
-
-##Train Model on the Data.
-Use the e1071 package to impliment the Naive Bayes algorithm on the data, and predict whether a message is likely to be spam or ham.
-
-```{r}
-library(e1071)
-
-sms_classifier <- naiveBayes(sms_train, sms_train_labels)
-```
-
-##Predict and Evaluate the Model.
-
-```{r}
-sms_test_pred <- predict(sms_classifier, sms_test)
-```
-
-Evaluate the predition with the actual data using a crosstable from the gmodels package.
-
-```{r}
-library(gmodels)
-CrossTable(sms_test_pred, sms_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
-```
-
-As shown in the table only 29/1379 messages were classified incorrectly.
-This means that the algorithm classifed the testing set as spam or ham with approx.
-98% accuracy.
-That’s impressive.
-To improve the model, one might tamper with the Laplace value, colect more sms data, or try splitting the dataset randomly into training and testing.
-I suspect the accuracy would increase as the dataset gets bigger. The more data there is to train the algorith, the more effective it would be in predicting Spam or Ham.
-
-Show the 5 most frequent words in the sms data:
-```{r}
-sack <- TermDocumentMatrix(sms_corpus_clean)
-pack <- as.matrix(sack)
-snack <- sort(rowSums(pack), decreasing = TRUE)
-hack <- data.frame(word = names(snack), freq=snack)
-head(hack, 5)
-```
-
-And the 5 most frequent words from each class:
-
-```{r}
-wordcloud(spam$text, max.words = 5, scale = c(3, 0.5))
-
-wordcloud(ham$text, max.words = 5, scale = c(3, 0.5))
-```
-
-As shown in the word clouds, the most frequent words from the spam messages are "call", "free", "now", "mobile", and "text".
-And in the Ham messages, the 5 most frequent words are "can", "get", "just", "will", and "now".
-
-***
-***
-
-#My Test of Naive Bayes
-
-##Loading Data Into R.
-This block reads full CSV into a starting dataframe named "base_data". Next block isolates tables and rows into new dataframe "sms_raw2", and renames the columns.
-```{r}
-#CSV file named "coded_full_comments_dataset_Spring18anon.csv" already has correctly organized text and meta-data. First line reads it in.
-base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
-
-# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
-frequency_writing <- filter(base_data, code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
-sms_raw2 <- frequency_writing %>% select(23,22)
-
-#Rename columns
-names(sms_raw2)[1] <- "type"
-names(sms_raw2)[2] <- "text"
-```
-
-```{r}
-str(sms_raw2)
-```
-
-The "type"" element is a character. This changes it to a factor for the analysis.
-
-```{r}
-sms_raw2$type <- factor(sms_raw2$type)
-str(sms_raw2$type)
-table(sms_raw2$type)
-```
-
-##Text Mining
-For Naive Bayes to run effectively, the test data needs to be transformed. This begins with using the "tm" package in R to create a volitile coprus that contains the "text" vector from our data frame.
-
-```{r}
-#install.packages("tm")
-library(tm)
-```
-
-```{r}
-sms_corpus2 <- VCorpus(VectorSource(sms_raw2$text))
-print(sms_corpus2)
-```
-
-Check out the first few messages in the new corpus, which is basically a list that can be manipulated with list operations.
-
-```{r}
-inspect(sms_corpus2[1:3])
-```
-
-Use "as.character" function to see what a message looks like.
-
-```{r}
-as.character(sms_corpus2[[3]])
-```
-
-In order to standardize the messages, the data set must be tranformed to all lower case letters.
-The words "Free", "free", and "FREE" should all be treated as the same word.
-Use the "tm_map"" funtion in R, and use the "content_transformer" function to transform the text.
-
-```{r}
-sms_corpus_clean2 <- tm_map(sms_corpus2, content_transformer(tolower))
-```
-
-Look at third message again to see if our data was transformed.
-
-```{r}
-as.character(sms_corpus2[[3]])
-
-as.character((sms_corpus_clean2[[3]]))
-```
-
-```{r}
-sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removeNumbers)
-```
-
-Remome words that appear often but don't contribute to our objective. These words include "to", "and", "but" and "or".
-
-```{r}
-sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removeWords, stopwords())
-```
-
-Remove punctuation as well using the "removePunctuation" function.
-
-```{r}
-sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removePunctuation)
-
-as.character((sms_corpus_clean2[[3]]))
-```
-
-Perform "stemming" to the text data to strip the suffix from words like "jumping", so the words "jumping" "jumps" and "jumped" are all transformed into "jump".
-Stemming can be perfromed using the "tm" package with help from the "SnowballC" package.
-
-```{r}
-#install.packages("SnowballC")
-#library(SnowballC)
-
-sms_corpus_clean2 <- tm_map(sms_corpus_clean2, stemDocument)
-```
-
-And now the final step in text mining is to remove white space from the document.
-
-```{r}
-sms_corpus_clean2 <- tm_map(sms_corpus_clean2, stripWhitespace)
-```
-
-```{r}
-as.character(sms_corpus_clean2[[3]])
-```
-
-Perform tokenization using the "DocumentTermMatrix" function.
-This creates a matrix in which the rows indicate documents (SMS messages in this case) and the columns indicate words.
-It should be noted that the "DocumentTermMaxtrix" function has the power to do all of the text mining above in one command.
-
-```{r}
-sms_dtm2 <- DocumentTermMatrix(sms_corpus_clean2)
-```
-
-##Data Preparation
-Split our data into training and testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data. Divide our data set into 75% training and 25% testing.
-
-```{r}
-.75 * 9129
-.25 * 9129
-```
-
-Because the dataset is random, the first 4180 messages can be used for the training set - there's no need to randomize the data first.
-
-```{r}
-sms_dtm_train2 <- sms_dtm2[1:6846, ]
-sms_dtm_test2 <- sms_dtm2[6847:9129, ]
-```
-
-Save vectors labeling rows in the training and testing vectors
-
-```{r}
-sms_train_labels2 <- sms_raw2[1:6846, ]$type
-sms_test_labels2 <- sms_raw2[6847:9129,]$type
-```
-
-Make sure that the proportion of spam is similar in the training and testing data set.
-
-```{r}
-prop.table(table(sms_train_labels2))
-
-prop.table(table(sms_test_labels2))
-```
-
-##Visualization
-Create a wordcloud of the frequency of the words in the dataset using the package "wordcloud".
-
-```{}
-install.packages("wordcloud")
-install.packages("RColorBrewer")
-```
-
-```{r}
-#library(wordcloud)
-wordcloud(sms_corpus_clean2, max.words = 50, random.order = FALSE)
-```
-
-Compare wordclouds between 3 groups.
-
-```{r}
-TWO <- subset(sms_raw2, type == "2. Writing Quality")
-THREE <- subset(sms_raw2, type == "3. Technical and Scientific")
-FOUR <- subset(sms_raw2, type == "4. Logic and Thinking")
-
-wordcloud(TWO$text, max.words = 50, scale = c(3, 0.5))
-wordcloud(THREE$text, max.words = 50, scale = c(3, 0.5))
-wordcloud(FOUR$text, max.words = 50, scale = c(3, 0.5))
-```
-
-##Preparation for Naive Bayes
-Remove words from the matrix that appear less than 5 times.
-
-```{r}
-sms_freq_words2 <- findFreqTerms(sms_dtm_train2, 5)
-str(sms_freq_words2)
-```
-
-Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
-
-```{r}
-sms_dtm_freq_train2 <- sms_dtm_train2[ , sms_freq_words2]
-sms_dtm_freq_test2 <- sms_dtm_test2[ , sms_freq_words2]
-```
-
-The naive bayes classifier works with categorical reatures, so we need to convert the matrix to "yes" and "no" categorical variables.
+The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables.
 To do this we'll build a convert_counts function and apply it to our data.
-
 ```{r}
@@
 This replaces values greater than 0 with yes, and values not greater than 0 with no. Let's apply it to our data.
-
 ```{r}
-sms_train2 <- apply(sms_dtm_freq_train2, MARGIN = 2, convert_counts2)
-
-sms_test2 <- apply(sms_dtm_freq_test2, MARGIN = 2, convert_counts2)
+comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
+comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
 ```
@@
 
-##Train Model on the Data.
-Use the e1071 package to impliment the Naive Bayes algorithm on the data, and predict whether a message is likely to be spam or ham.
+####\
 
-```{}
-install.packages("e1071")
-library(e1071)
+###Train Model, Predict, Evaluate
+Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group ONE, TWO, THREE, or FOUR.
+Evaluate the prediction with the actual data using a crosstable from the gmodels package.
+```{r}
+comments_classifier <- naiveBayes(comments_train, comments_train_labels)
+comments_test_pred <- predict(comments_classifier, comments_test)
+CrossTable(comments_test_pred, comments_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
+```
+
+Calculate overall correct prediction rate.
+```{r}
+(42+454+1186+185)/23.35
+```
+
+***
+##Modified Naive Bayes (NB) Method Using N-grams
+The tokenizer command block above only works for single words.
+The block below uses tm's NLP commands to generate either 1-, 2-, or 3-grams with the same command.
+The only change needed is "#" in "(words(x), #)" below.
+For simplicity I left off all of the preparatory code up leading up to the final cleaned corpus.
+
+```{r}
+NLP_Tokenizer <- function(x) {
+      unlist(lapply(ngrams(words(x), 2), paste, collapse = " "), use.names = FALSE)
+}
+
+comments_dtm_2gram <- DocumentTermMatrix(comments_corpus_clean, control=list(tokenize = NLP_Tokenizer))
+comments_dtm_2gram_tidy <- tidy(comments_dtm_2gram)
+inspect(comments_dtm_2gram)
+```
+
+####\
+
+###Data Preparation
+Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data.
+
+```{r}
+.75 * 9340 #Number of rows in the dataset; product is #rows for training
+.25 * 9340 #Product is #rows for testing set
+```
+
+This code assumes comments are random. **Probably want to try randomizing them.**
+```{r}
+comments_dtm_train <- comments_dtm_2gram[1:7005, ]
+comments_dtm_test <- comments_dtm_2gram[7006:9340, ]
+```
+
+Save vectors labeling rows in the training and testing vectors
+```{r}
+comments_train_labels <- comments_raw[1:7005, ]$type
+comments_test_labels <- comments_raw[7006:9340,]$type
+```
+
+Make sure that the proportion of each sub-category is similar in the training and testing data set.
+```{r}
+prop.table(table(comments_train_labels))
+prop.table(table(comments_test_labels))
+```
+
+####\
+
+###Preparation for Naive Bayes
+Remove words from the matrix that appear less than 5 times.
+```{r}
+comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
+str(comments_freq_words)
+```
+
+Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
+```{r}
+comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
+comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
+```
+
+The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables.
+To do this we'll build a convert_counts function and apply it to our data.
+```{r}
+convert_counts2 <- function(x) {
+  x <- ifelse(x > 0, "Yes", "No")
+}
+```
+
+This replaces values greater than 0 with yes, and values not greater than 0 with no. Let's apply it to our data.
+```{r}
+comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
+comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
+```
+
+The resulting matrixes will be character type, with cells indicating "yes" or "no" if the word represented by the column appears in the message represented by the row.
+
+####\
+
+###Train Model, Predict, Evaluate
+Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group ONE, TWO, THREE, or FOUR.
+Evaluate the prediction with the actual data using a crosstable from the gmodels package.
+```{r}
+comments_classifier <- naiveBayes(comments_train, comments_train_labels)
+comments_test_pred <- predict(comments_classifier, comments_test)
+CrossTable(comments_test_pred, comments_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
 ```
@@
 ```{r}
-sms_classifier2 <- naiveBayes(sms_train2, sms_train_labels2)
+(34+214+1254+94)/23.35
 ```
 
-##Predict and Evaluate the Model.
+***
+***
+####\
 
+##Visualization
+This code generates a wordcloud of the frequency of single words in the dataset using the package "wordcloud".
 ```{r}
-sms_test_pred2 <- predict(sms_classifier2, sms_test2)
+wordcloud(comments_corpus_clean, max.words = 50, random.order = FALSE)
 ```
 
-Evaluate the predition with the actual data using a crosstable from the gmodels package.
+Compare wordclouds between 3 groups.
+```{r}
+ONE <- subset(comments_raw, type == "1_Basic")
+TWO <- subset(comments_raw, type == "2_Writing")
+THREE <- subset(comments_raw, type == "3_Technical")
+FOUR <- subset(comments_raw, type == "4_Logic")
 
-```{r}
-#library(gmodels)
-CrossTable(sms_test_pred2, sms_test_labels2, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
+wordcloud(ONE$text, max.words = 50, scale = c(3, 0.5))
+wordcloud(TWO$text, max.words = 50, scale = c(3, 0.5))
+wordcloud(THREE$text, max.words = 50, scale = c(3, 0.5))
+wordcloud(FOUR$text, max.words = 50, scale = c(3, 0.5))
 ```
 
+Show the 5 most frequent words in the data:
 ```{r}
-1352/1379
-```
-
-As shown in the table only 27/1379 messages were classified incorrectly.
-This means that the algorithm classifed the testing set as spam or ham with approx.
-98% accuracy.
-That's impressive.
-To improve the model, one might tamper with the Laplace value, colect more sms data, or try splitting the dataset randomly into training and testing.
-
-I suspect the accuracy would increase as the dataset gets bigger. The more data there is to train the algorith, the more effective it would be in predicting Spam or Ham.
-
-Show the 5 most frequent words in the sms data:
-
-```{r}
-sack2 <- TermDocumentMatrix(sms_corpus_clean2)
-pack2 <- as.matrix(sack2)
-snack2 <- sort(rowSums(pack2), decreasing = TRUE)
-hack2 <- data.frame(word = names(snack2), freq=snack2)
-head(hack2, 5)
+comment_sack <- TermDocumentMatrix(comments_corpus_clean)
+comment_pack <- as.matrix(comment_sack)
+comment_snack <- sort(rowSums(comment_pack), decreasing = TRUE)
+comment_hack <- data.frame(word = names(comment_snack), freq=comment_snack)
+head(comment_hack, 10)
 ```
@@
 And the 5 most frequent words from each class:
+```{r}
+wordcloud(ONE$text, max.words = 10, scale = c(3, 0.5))
+wordcloud(TWO$text, max.words = 10, scale = c(3, 0.5))
+wordcloud(THREE$text, max.words = 10, scale = c(3, 0.5))
+wordcloud(FOUR$text, max.words = 10, scale = c(3, 0.5))
+```
 
-```{r}
-wordcloud(TWO$text, max.words = 5, scale = c(3, 0.5))
-wordcloud(THREE$text, max.words = 5, scale = c(3, 0.5))
-wordcloud(FOUR$text, max.words = 5, scale = c(3, 0.5))
-```
+####\
+
+##Permutation Tests of Naive Bayes (NB) Method
+Version 1.0 correctly identifies comment subcategories ~79% of the time when using single words, and 68.3% of the time when using 2-grams. Permutations to test:
+
+*  Compare 2-grams, 3-grams rather than single words
+*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
+    +  Ex.: Basic vs. Not Basic (1_Basic vs. not 1_Basic)
+*  Modify the variables within the NB analysis itself. 
+    +  Laplace constant (Y/N), Cutoff value for words removed from list.
+*  Do/Do not stem the terms
+    +  This may remove evidence of tense
+*  Remove/Leave stopwords
+*  Remove/Leave capitalization
+*  Remove/Leave numbers
+    +  This may remove evidence of technical descriptors
+*  Remove/Leave punctuation
+    +  This may remove evidence of questions vs statements

### 2019-01-31 Naive_bayes_V2r.Rmd
Working text analysis forms

---
title: "Naive Bayes Workflow Ver 2.0.r"
author: "Dan Johnson"
date: "1/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Summary
This is **Version 2.0.r** of the basic workflow for classifying TA comments using a naive Bayes model. It pre-processes text using the same command set as Ver 2.0.b. The "r" designates this version has been modified to be able to randomize which rows are selected for training and test datasets.

####\

##Initial Setup
Calling the required libraries.

```{}
library(tidyverse)
library(tidytext)
library(tidyr)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
```

####\

##Version 2.0.r of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

###Extract, Organize Data Subset
Extract comments and classification data subset
```{r}
#Select rows representing the sub-groups to compare. 
comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
comments_rawVERB <- comments_subsetVERB %>% select(23,24,22)

#Rename the columns.
names(comments_rawVERB)[1] <- "subject"
names(comments_rawVERB)[2] <- "structure"
names(comments_rawVERB)[3] <- "text"

#Change "subject" element from character to a factor for analysis.
comments_rawVERB$subject <- factor(comments_rawVERB$subject)
str(comments_rawVERB$subject)
table(comments_rawVERB$subject)
```

```{r}
#Change "structure" element from character to a factor for analysis.
comments_rawVERB$structure <- factor(comments_rawVERB$structure)
str(comments_rawVERB$structure)
table(comments_rawVERB$structure)
```

```{r}
#Generate a vector of ~10,000 numbers (be sure it can be generated consistently)
#Adjust the number so it matches the number of rows in the original data frame
#Next, set the seed for random numbers, and randomize the order of the vector
vector <- 1:9340
set.seed(123) #change this seed value, and the data are randomized in a new sequence.
vector <- sample(vector)
glimpse(vector)
```

```{r}
#Append the vector as a new column to "comments_rawVERB" with name "randomizer"
#Sort rows of dataframe according to "randomizer" column using dplyer 'arrange' function. 
comments_rawVERB$randomizer <- c(vector)
comments_rawVERB_R <-comments_rawVERB %>% arrange(randomizer)
print(comments_rawVERB_R)

```

Convert data set to a volative corpus using "tm" library.
```{r}
#Create the volatile corpus that contains the "text" vector from data frame.
comments_corpusVERB_R <- VCorpus(VectorSource(comments_rawVERB_R$text))
print(comments_corpusVERB_R)

#Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
inspect(comments_corpusVERB_R[1:3])

#Use "as.character" function to see what a single text comment looks like.
as.character(comments_corpusVERB_R[[5]])
```

###Text Data Transformations
The text data transforms are separated out as individual "tm_map"" and "content_transformer" functions to simplify the process of testing various combinations. Conversion to lower case and removing punctuation are the baseline set. Switch between "{r}" and "{}" to turn the other transformation blocks off or on. Removing extra white spaces must be the last transformation, and is required.

```{r}
#Convert text to all lower case letters.
comments_corpus_cleanVERB_R <- tm_map(comments_corpusVERB_R, content_transformer(tolower))

#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removePunctuation)

#Remove numerals. 
#I am not sure this is beneficial; does it remove evidence indicating technical comments?
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeNumbers)

#Stopword removal. 
#This is a standard cleanup step, but again may remove useful terms.
#The "stopwords" file is a convention; other files can be substituted for it
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeWords, stopwords())

#The option to use stemming has been removed from this version.

#Final step removes white space from the document. This is NOT an optional step.
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, stripWhitespace)

#Look at an example of cleaned text comment to see if it fits what is expected.
as.character((comments_corpus_cleanVERB_R[[5]]))
```

####\

##Basic Naive Bayes (NB) Method Using N-grams
This code block uses tm's NLP commands. The same block can be used to assess 1-, 2-, or 3-grams. Only the "#" variable in "(words(x), #)" below needs to be changed.

```{r}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 3), paste, collapse = " "), use.names = FALSE)
}

comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB_R, control=list(tokenize = NLP_Tokenizer))
comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
inspect(comments_dtm_1gram)
```

####\

###Data Preparation
Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data.

```{}
.75 * 9340 #Number of rows in the dataset; product is #rows for training
.25 * 9340 #Product is #rows for testing set
```

This code uses the comments in the order they are listed in the original file. Block for randomizing them was done earlier.
```{r}
comments_dtm_train <- comments_dtm_1gram[1:7005, ]
comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
```

Save vectors labeling rows in the training and testing vectors
```{r}
comments_train_labels_subject <- comments_rawVERB_R[1:7005, ]$subject
comments_test_labels_subject <- comments_rawVERB_R[7006:9340,]$subject
```

```{r}
comments_train_labels_structure <- comments_rawVERB_R[1:7005, ]$structure
comments_test_labels_structure <- comments_rawVERB_R[7006:9340,]$structure
```

Make sure that the proportion of each sub-category is similar in the training and testing data sets.
```{r}
prop.table(table(comments_train_labels_subject))
prop.table(table(comments_test_labels_subject))
```

```{r}
prop.table(table(comments_train_labels_structure))
prop.table(table(comments_test_labels_structure))
```

####\

###Preparation for Naive Bayes
Remove words from the matrix that appear less than 5 times.
```{r}
comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
str(comments_freq_words)
```

Limit Document Term Matrix to only include words in the comments_freq_words vector. Using all the rows, but we want to limit the columns to these words in the frequency vector.
```{r}
comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
```

The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables. This is done using a **convert_counts function** and applying it to the data. This replaces values greater than 0 with yes, and values not greater than 0 with no. 
```{r}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}

comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
```

The resulting matrices have cells indicating "yes" or "no" if the word represented by the column appears in the text comment represented by the row.

####\

###Train Model, Predict, Evaluate for Subject
This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. It evaluates the prediction with the actual data using a crosstable from the gmodels package.
```{r}
comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject)
comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

The cross table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject.
```{r}
print("Overall subject accuracy")
(14+54+1318+28)/23.35  #Need to find a way to calculate this automatically. 
```

###Train Model, Predict, Evaluate for structure
This is the same block as above but uses the "structure" data column.

```{r}
comments_classifier_structure <- naiveBayes(comments_train, comments_train_labels_structure)
comments_test_pred_structure <- predict(comments_classifier_structure, comments_test)
CrossTable(comments_test_pred_structure, comments_test_labels_structure, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

Again, a cross table shows how many of the predicted categories match the correct categories for structure.

```{r}
print("Overall structure accuracy")
(66+64+949+5+1)/23.35  #Need to find a way to calculate this automatically. 
```

###OPTIONAL Output
To make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame. The full data frame will be too long to display; use "head" to check format, then output the full dataset to CSV for further review. By default the CSV file should be saved to the current working directory.

```{}
predicted_category_subject <- data.frame(comments_test_pred_subject,comments_test_labels_subject)
head(predicted_category_subject)
write.csv(predicted_category_subject, file = "PredictedCategorySubject_Data.csv")
```

```{}
predicted_category_structure <- data.frame(comments_test_pred_structure,comments_test_labels_structure)
head(predicted_category_structure)
write.csv(predicted_category_structure, file = "PredictedCategoryStructure_Data.csv")
```

***
START COMMENTS, OLD work notes
***

##Permutation Tests of Naive Bayes (NB) Method
Using single words, Version 2.0.b correctly identifies comment "subject" ~80% of the time and comment "structure" ~59% of the time.

Permutations to test with random subsets version (2.0.r):

*  Compare 1-grams, 2-grams, 3-grams
*  Compare outcomes for 3 independent randomized sets of 1-grams

The pseudo-code block below can pull random lines from a data table, but does not work correctly on a document term matrix. The second problem is how to pull the training and test data sets out and process them identically.

```{}
set.seed(1234) 		# ensures the same random set gets picked each time.

comments_raw_dtm_train <- sample_frac(comments_raw, .75)	# sample_frac pulls out 75% of "comments_raw" dataset randomly using the sample sequence identified by set.seed(1234)

comments_raw_dtm_test <- setdiff(comments_raw, comments_raw_dtm_train)	#setdiff creates a "test" dataset using anything NOT in the "comments_raw_dtm_train" subset of "comments_raw."
```

The solution was not to pull the data twice, but to randomly (but repeatably) shuffle the starting data table. How I solved this was:
1. Create a vector of sequential numbers the same length as the number of data frame rows.
2. Randomize the sequential numbers using set.seed to make pattern reproducible.
3. Add the randomized numbers as a new column to the data frame.
4. Re-order the data table to the randomizer column.

This makes it easy to generate a new sequence of randomized training and test samples to run replicates. Only one number needs to be changed.

```{}
#Creates a data frame that compares predicted and actual labels for subject and structure
tally <- data.frame("subject_label"=c(comments_test_labels_subject), "subject_predicted"=c(comments_test_pred_subject), "structure_label"=c(comments_test_labels_structure), "structure_predicted"=c(comments_test_pred_structure))
#tally <- as.numeric(as.character(tally))
```

```{}
if_else (tally(comments_test_labels_subject) != tally(comments_test_pred_subject),"YES","no")

#glimpse(tally)

#tally %>% mutate(match_subject=comments_test_labels_subject - comments_test_pred_subject)
#tally %>% mutate(match_structure=comments_test_labels_structure - comments_test_pred_structure)
```

***
END COMMENTS
***

### 2019-01-31 Naive_bayes_V2yn.Rmd
Working text analysis forms

--- 2019-01-31 Naive_bayes_V2r.Rmd
@@
 ---
-title: "Naive Bayes Workflow Ver 2.0.r"
+title: "Naive Bayes Workflow Ver 2.0.yn"
 author: "Dan Johnson"
@@
 ##Summary
-This is **Version 2.0.r** of the basic workflow for classifying TA comments using a naive Bayes model.
-It pre-processes text using the same command set as Ver 2.0.b.
-The "r" designates this version has been modified to be able to randomize which rows are selected for training and test datasets.
+This is **Version 2.0.yn** of the basic workflow for classifying TA comments using a naive Bayes model.
+It is optimized for testing pair-wise assignment to categories (Yes/No decision) beginning with the most frequently encountered.
+
+The main features and permutations to be evaluated are:
+
+*  Stepwise paired classification versus comparing multiple groups simultaneously
+
+Additional working notes are located at the end of this page.
 
@@
 
-```{}
+```{r}
 library(tidyverse)
@@
 
-##Version 2.0.r of Naive Bayes (NB) Protocol
+##Version 2.0.yn of Naive Bayes (NB) Protocol
 ###Initial Data Input
@@
 #Rename the columns.
-names(comments_rawVERB)[1] <- "subject"
-names(comments_rawVERB)[2] <- "structure"
+names(comments_rawVERB)[1] <- "type"
+names(comments_rawVERB)[2] <- "locus"
 names(comments_rawVERB)[3] <- "text"
 
-#Change "subject" element from character to a factor for analysis.
-comments_rawVERB$subject <- factor(comments_rawVERB$subject)
-str(comments_rawVERB$subject)
-table(comments_rawVERB$subject)
+#Change "type" element from character to a factor for analysis.
+comments_rawVERB$type <- factor(comments_rawVERB$type)
+str(comments_rawVERB$type)
+table(comments_rawVERB$type)
 ```
@@
 ```{r}
-#Change "structure" element from character to a factor for analysis.
-comments_rawVERB$structure <- factor(comments_rawVERB$structure)
-str(comments_rawVERB$structure)
-table(comments_rawVERB$structure)
+#Change "locus" element from character to a factor for analysis.
+comments_rawVERB$locus <- factor(comments_rawVERB$locus)
+str(comments_rawVERB$locus)
+table(comments_rawVERB$locus)
 ```
 
-```{r}
-#Generate a vector of ~10,000 numbers (be sure it can be generated consistently)
-#Adjust the number so it matches the number of rows in the original data frame
-#Next, set the seed for random numbers, and randomize the order of the vector
-vector <- 1:9340
-set.seed(123) #change this seed value, and the data are randomized in a new sequence.
-vector <- sample(vector)
-glimpse(vector)
-```
+***
+***
+Here is where I need to add the column used to randomly assign the data to training or testing sets.
 
-```{r}
-#Append the vector as a new column to "comments_rawVERB" with name "randomizer"
-#Sort rows of dataframe according to "randomizer" column using dplyer 'arrange' function. 
-comments_rawVERB$randomizer <- c(vector)
-comments_rawVERB_R <-comments_rawVERB %>% arrange(randomizer)
-print(comments_rawVERB_R)
+Use set.seed(1234) to create a stable randomized set.
 
-```
+***
+***
 
@@
 #Create the volatile corpus that contains the "text" vector from data frame.
-comments_corpusVERB_R <- VCorpus(VectorSource(comments_rawVERB_R$text))
-print(comments_corpusVERB_R)
+comments_corpusVERB <- VCorpus(VectorSource(comments_rawVERB$text))
+print(comments_corpusVERB)
 
 #Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
-inspect(comments_corpusVERB_R[1:3])
+inspect(comments_corpusVERB[1:3])
 
 #Use "as.character" function to see what a single text comment looks like.
-as.character(comments_corpusVERB_R[[5]])
+as.character(comments_corpusVERB[[5]])
 ```
@@
 #Convert text to all lower case letters.
-comments_corpus_cleanVERB_R <- tm_map(comments_corpusVERB_R, content_transformer(tolower))
+comments_corpus_cleanVERB <- tm_map(comments_corpusVERB, content_transformer(tolower))
+```
 
+```{r}
 #Remove punctuation using the "removePunctuation" function. 
 #This step removes evidence of questions, so may remove data.
-comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removePunctuation)
+comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removePunctuation)
+```
 
+```{r}
 #Remove numerals. 
 #I am not sure this is beneficial; does it remove evidence indicating technical comments?
-comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeNumbers)
 
+comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removeNumbers)
+```
+
+```{r}
 #Stopword removal. 
@@
 #The "stopwords" file is a convention; other files can be substituted for it
-comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeWords, stopwords())
 
-#The option to use stemming has been removed from this version.
+comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removeWords, stopwords())
+```
 
+Option to stem terms has been removed from this version.
+
+Final prep and data check.
+```{r}
 #Final step removes white space from the document. This is NOT an optional step.
-comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, stripWhitespace)
+comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, stripWhitespace)
 
 #Look at an example of cleaned text comment to see if it fits what is expected.
-as.character((comments_corpus_cleanVERB_R[[5]]))
+as.character((comments_corpus_cleanVERB[[5]]))
 ```
@@
 NLP_Tokenizer <- function(x) {
-      unlist(lapply(ngrams(words(x), 3), paste, collapse = " "), use.names = FALSE)
+      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
 }
 
-comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB_R, control=list(tokenize = NLP_Tokenizer))
+comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB, control=list(tokenize = NLP_Tokenizer))
 comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
@@
 
-This code uses the comments in the order they are listed in the original file. Block for randomizing them was done earlier.
-```{r}
+This code uses the comments in the order they are listed in the original file. **Need a block for  randomizing them.**
+```{}
 comments_dtm_train <- comments_dtm_1gram[1:7005, ]
@@
 Save vectors labeling rows in the training and testing vectors
-```{r}
-comments_train_labels_subject <- comments_rawVERB_R[1:7005, ]$subject
-comments_test_labels_subject <- comments_rawVERB_R[7006:9340,]$subject
+```{}
+comments_train_labels_type <- comments_rawVERB[1:7005, ]$type
+comments_test_labels_type <- comments_rawVERB[7006:9340,]$type
 ```
 
-```{r}
-comments_train_labels_structure <- comments_rawVERB_R[1:7005, ]$structure
-comments_test_labels_structure <- comments_rawVERB_R[7006:9340,]$structure
+```{}
+comments_train_labels_locus <- comments_rawVERB[1:7005, ]$locus
+comments_test_labels_locus <- comments_rawVERB[7006:9340,]$locus
 ```
 
-Make sure that the proportion of each sub-category is similar in the training and testing data sets.
-```{r}
-prop.table(table(comments_train_labels_subject))
-prop.table(table(comments_test_labels_subject))
+Make sure that the proportion of each sub-category is similar in the training and testing data set.
+```{}
+prop.table(table(comments_train_labels_type))
+prop.table(table(comments_test_labels_type))
 ```
 
-```{r}
-prop.table(table(comments_train_labels_structure))
-prop.table(table(comments_test_labels_structure))
+```{}
+prop.table(table(comments_train_labels_locus))
+prop.table(table(comments_test_labels_locus))
 ```
@@
 This replaces values greater than 0 with yes, and values not greater than 0 with no.
+
 ```{r}
@@
 
-###Train Model, Predict, Evaluate for Subject
+###Train Model, Predict, Evaluate for Type
 This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic.
@@
 ```{r}
-comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject)
-comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
-CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
+comments_classifier_type <- naiveBayes(comments_train, comments_train_labels_type)
+comments_test_pred_type <- predict(comments_classifier_type, comments_test)
+CrossTable(comments_test_pred_type, comments_test_labels_type, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
 ```
 
-The cross table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject.
+The cross table list how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting TYPE.
+
 ```{r}
-print("Overall subject accuracy")
-(14+54+1318+28)/23.35  #Need to find a way to calculate this automatically. 
+print("Overall TYPE accuracy")
+(14+45+1346+19)/23.35
 ```
 
-###Train Model, Predict, Evaluate for structure
-This is the same block as above but uses the "structure" data column.
+###Train Model, Predict, Evaluate for LOCUS
+This is the same block as above but uses the "Locus" data column.
 
 ```{r}
-comments_classifier_structure <- naiveBayes(comments_train, comments_train_labels_structure)
-comments_test_pred_structure <- predict(comments_classifier_structure, comments_test)
-CrossTable(comments_test_pred_structure, comments_test_labels_structure, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
+comments_classifier_locus <- naiveBayes(comments_train, comments_train_labels_locus)
+comments_test_pred_locus <- predict(comments_classifier_locus, comments_test)
+CrossTable(comments_test_pred_locus, comments_test_labels_locus, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
 ```
 
-Again, a cross table shows how many of the predicted categories match the correct categories for structure.
+Again, a cross table shows how many of the predicted categories match the correct categories for LOCUS.
 
 ```{r}
-print("Overall structure accuracy")
-(66+64+949+5+1)/23.35  #Need to find a way to calculate this automatically. 
+print("Overall LOCUS accuracy")
+(62+64+931+3)/23.35
 ```
@@
 ```{}
-predicted_category_subject <- data.frame(comments_test_pred_subject,comments_test_labels_subject)
-head(predicted_category_subject)
-write.csv(predicted_category_subject, file = "PredictedCategorySubject_Data.csv")
+predicted_category_type <- data.frame(comments_test_pred_type,comments_test_labels_type)
+head(predicted_category_type)
+write.csv(predicted_category_type, file = "PredictedCategoryType_Data.csv")
 ```
@@
 ```{}
-predicted_category_structure <- data.frame(comments_test_pred_structure,comments_test_labels_structure)
-head(predicted_category_structure)
-write.csv(predicted_category_structure, file = "PredictedCategoryStructure_Data.csv")
+predicted_category_locus <- data.frame(comments_test_pred_locus,comments_test_labels_locus)
+head(predicted_category_locus)
+write.csv(predicted_category_locus, file = "PredictedCategoryLocus_Data.csv")
 ```
@@
 ***
-START COMMENTS, OLD work notes
+START COMMENTS
 ***
@@
 ##Permutation Tests of Naive Bayes (NB) Method
-Using single words, Version 2.0.b correctly identifies comment "subject" ~80% of the time and comment "structure" ~59% of the time.
+Using single words, Version 2.0.b correctly identifies comment "type" ~80% of the time and comment "locus" ~59% of the time.
 
-Permutations to test with random subsets version (2.0.r):
+Permutations to Test:
 
 *  Compare 1-grams, 2-grams, 3-grams
-*  Compare outcomes for 3 independent randomized sets of 1-grams
+*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
+    +  Initial approach will be to split between largest group and balance in each category
+    +  Ex.: type is 3-technical vs. not 3_technical
 
-The pseudo-code block below can pull random lines from a data table, but does not work correctly on a document term matrix.
-The second problem is how to pull the training and test data sets out and process them identically.
+Known Issues:
 
-```{}
-set.seed(1234) 		# ensures the same random set gets picked each time.
-
-comments_raw_dtm_train <- sample_frac(comments_raw, .75)	# sample_frac pulls out 75% of "comments_raw" dataset randomly using the sample sequence identified by set.seed(1234)
-
-comments_raw_dtm_test <- setdiff(comments_raw, comments_raw_dtm_train)	#setdiff creates a "test" dataset using anything NOT in the "comments_raw_dtm_train" subset of "comments_raw."
-```
-
-The solution was not to pull the data twice, but to randomly (but repeatably) shuffle the starting data table. How I solved this was:
-1. Create a vector of sequential numbers the same length as the number of data frame rows.
-2. Randomize the sequential numbers using set.seed to make pattern reproducible.
-3. Add the randomized numbers as a new column to the data frame.
-4. Re-order the data table to the randomizer column.
-
-This makes it easy to generate a new sequence of randomized training and test samples to run replicates. Only one number needs to be changed.
-
-```{}
-#Creates a data frame that compares predicted and actual labels for subject and structure
-tally <- data.frame("subject_label"=c(comments_test_labels_subject), "subject_predicted"=c(comments_test_pred_subject), "structure_label"=c(comments_test_labels_structure), "structure_predicted"=c(comments_test_pred_structure))
-#tally <- as.numeric(as.character(tally))
-```
-
-```{}
-if_else (tally(comments_test_labels_subject) != tally(comments_test_pred_subject),"YES","no")
-
-#glimpse(tally)
-
-#tally %>% mutate(match_subject=comments_test_labels_subject - comments_test_pred_subject)
-#tally %>% mutate(match_structure=comments_test_labels_structure - comments_test_pred_structure)
-```
+1. The e1071 naive bayes classifier is binomial. Need to incorporate multinomial.
 

### 2019-01-31 VERB.Rmd
Working text analysis forms

--- 2018-12-18 draft.Rmd
@@
 ##Deeper Text Analysis
-These linked pages summarize work to date on each method.
+These linked pages summarize work to date on individual methods.
 
 *  Naive Bayes: https://adanieljohnson.github.io/default_website/Naive_bayes.html
-*  Random Forest (link)
-*  Latent Dirichlet Allocation (link)
-*  Etc. (link)
+*  To Be Decided #1. (link)
+*  To Be Decided #2. (link)
+*  To Be Decided #3. (link)
 *  Etc. (link)
@@
 #Initial Dataset
-My working dataset is an Excel table of 10,808 rows.
+The development dataset is an Excel table of 10,808 rows.
 One column has "TA comments" extracted from student lab reports in Spring 2018.
@@
 #Data Pre-Checks
-Prior to import, review data table in Excel. Check:
+Prior to import, review data table **in Excel**. Check:
 
@@
 2.  Data have been de-identified and recoded using pre-assigned anonymous IDs.
-3.  Data table headers are in snake_case, and match tabulated list below. The same list is in the file "code_column_names.rmd"
+3.  Data table headers match those listed below. The same list is in the file "code_column_names.rmd"
 
@@
 #Subsequent analyses call in subsets of "base_data".
-base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
+base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
 
@@
 
+For recoding comment labels, I am considering recoding **comment.subject** as:
+```
+0_null
+1_basic
+2_writing
+3_technical
+4_logic
+5_exposition
+6_misconduct
+7_administrative
+8_nobasis
+9_split
+10_context
+11_flag
+12_narrative
+13_sciname
+```
+Recoding **comment.structure** as:
+```
+0_null
+1_pointer
+2_copyedit
+3_general
+4_specific
+5_rationale
+6_holistic
+7_idiomatic
+8_nobasis
+```
+Any obvious problems with these?
+
 ####\
@@
 # Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
-frequency_writing <- filter(base_data, code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
+frequency_writing <- filter(base_data, code.subject=="2_Writing"|code.subject=="3_Technical"|code.subject=="4_Logic")
 frequency_writing.subcolumns <- frequency_writing %>% select(1, 22:23)
@@
 ####Comparing Pearson Correlations of Word Frequencies
-If the words used by TAs in their comments differ between the categories, there should be fairly low correlations overall .
+If the words used by TAs in their comments differ between the categories, there should be fairly low correlations overall.
 ```{r}
@@
 
-```{}
+```{r}
 # Call in tidyverse, other required libraries.

### 2019-01-31 index.Rmd
update

@@
 ## Background
-I am interested in understanding how students learn technical writing skills in biology, and how we can use data science methods to improve that process.
+I am interested in understanding how students develop technical writing skills in biology, and more specifically, how we can apply data science methods to improve that process.
 
-Our NSF-sponsored project asks if and how scripted instruction, automated support, and holistic feedback accelerates student development as writers.
+My current NSF-sponsored project asks if and how scripted instruction, automated support, and holistic feedback accelerates student development as writers.
 One of our specific aims is to see if changing the structure and focus of instructors' comments affects how students' writing develops over time.
-To do this, we need to extract instructor comments from student reports, classify those comments, and correlate the types of comments with student performance.
+To do this, we need to extract instructor comments from student reports, classify those comments, and correlate the types of comments TAs make with student performance.
+
+My project for the Faculty Learning Community for R is partly exploratory, in I am asking: what information is present and can be mined from the TA comments dataset?
+The project also is partly predictive in that I hope to build a text classification workflow that automates the process of categorizing TA comments on student writing.
 
@@
 In Summer 2018, we extracted and tabulated ~11,000 TA comments from ~500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses collected during a single semester.
-Using an iteratively generated codebook, all 11,000 comments were tagged in two categories:
+Using an iteratively generated codebook, all 11,000 comments were classified in two categories:
 
 * __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
-* __Structure__: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?
+* __Structure__: did the comment Point Out an error without providing other help, provide simple Copy Correction, provide declarative General or Specific Information only, or provide Holistic coaching that fosters student thinking and long-term improvement?
 
 #### \   
-Counts of the coded TA comments were sorted into a two-axis contingency table. This table provided two pieces of information:
+Frequency counts of unambiguously coded TA comments were sorted into a two-axis contingency table that provided two important insights:
 
@@
 ## Specific Project Goals
-TA comments and meta-data for each student report can be extracted and compiled into a CSV datafile using an existing Unix script.
+TA comments and meta-data for each student report can be extracted and compiled into a CSV datafile using existing Unix scripts.
 However routinely classifying TA comments by hand is impractical; 10-12,000 comments need to be scored EACH SEMESTER.
@@
 
-For my project I will be building an R-based text analysis workflow that:
+My specific project goals were/are to create an R-based text analysis workflow that:
 
 1.
-**Creates a de-identified dataset.** Required data will be extracted from the master CSV data set and organized in an anonymized data frame where all names, comments, and confidential metadata are re-keyed to unique random identifiers.
+**Converts TA comments and metadata into a de-identified "tidy" dataframe for R.** This includes converting the master CSV data set to an anonymized data frame with all names and confidential metadata re-keyed to unique random identifiers.
 2.
-**Classifies TA comments according to our existing codebook**.
+**Uses text features to assign TA comments to the categories defined in our existing codebook**.
 The goal is to be able to assign instructor comments into our pre-defined Subject and Structure codes with 90% accuracy or greater.
-Initially I will be evaluating whether terms frequencies can be used to classify comments.
-Other methods will be evaluated as they come to light.
-3.
-**Sorts TA comments using unsupervised grouping methods**.
-The codebook for classifying TA comments makes prior assumptions about most relevant text features.
-Unsupervised learning may uncover more robust or informative classification groups.
 

### 2019-01-31 Naive_bayes_V2yn.Rmd
Update

--- 2019-01-31 Naive_bayes_V2r.Rmd
@@
 ##Summary
-This is **Version 2.0.r** of the basic workflow for classifying TA comments using a naive Bayes model.
-It pre-processes text using the same command set as Ver 2.0.b.
-The "r" designates this version has been modified to be able to randomize which rows are selected for training and test datasets.
+This is **Version 2.0.yn** of a basic workflow for classifying TA comments using a naive Bayes model.
+It is optimized for testing pair-wise assignment to categories (Yes/No decision) beginning with the most frequently encountered:
+
+*  3_technical for the Subject category
+*  4_specific for the Structure category
+
+The main features and permutations to be evaluated are:
+
+*  Stepwise paired classification versus comparing multiple groups simultaneously
+
+Additional working notes are located at the end of this page.
 
@@
 ##Initial Setup
-Calling the required libraries.
+Calling required libraries.
 
-```{}
+```{r}
 library(tidyverse)
@@
 #Read in TA comments from CSV file.
-base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
+base_dataVerYN <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVerYN.csv')
 ```
@@
 #Select rows representing the sub-groups to compare. 
-comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")
+comments_subsetVerYN <- filter(base_dataVerYN,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")
 
 #Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
-comments_rawVERB <- comments_subsetVERB %>% select(23,24,22)
+comments_rawVerYN <- comments_subsetVerYN %>% select(23,24,22)
 
 #Rename the columns.
-names(comments_rawVERB)[1] <- "subject"
-names(comments_rawVERB)[2] <- "structure"
-names(comments_rawVERB)[3] <- "text"
+names(comments_rawVerYN)[1] <- "subject"
+names(comments_rawVerYN)[2] <- "structure"
+names(comments_rawVerYN)[3] <- "text"
 
 #Change "subject" element from character to a factor for analysis.
-comments_rawVERB$subject <- factor(comments_rawVERB$subject)
-str(comments_rawVERB$subject)
-table(comments_rawVERB$subject)
+comments_rawVerYN$subject <- factor(comments_rawVerYN$subject)
+str(comments_rawVerYN$subject)
+table(comments_rawVerYN$subject)
 ```
@@
 #Change "structure" element from character to a factor for analysis.
-comments_rawVERB$structure <- factor(comments_rawVERB$structure)
-str(comments_rawVERB$structure)
-table(comments_rawVERB$structure)
+comments_rawVerYN$structure <- factor(comments_rawVerYN$structure)
+str(comments_rawVerYN$structure)
+table(comments_rawVerYN$structure)
 ```
@@
 ```{r}
-#Append the vector as a new column to "comments_rawVERB" with name "randomizer"
+#Append the vector as a new column to "comments_rawVerYN" with name "randomizer"
 #Sort rows of dataframe according to "randomizer" column using dplyer 'arrange' function. 
-comments_rawVERB$randomizer <- c(vector)
-comments_rawVERB_R <-comments_rawVERB %>% arrange(randomizer)
-print(comments_rawVERB_R)
+comments_rawVerYN$randomizer <- c(vector)
+comments_rawVerYN_R <-comments_rawVerYN %>% arrange(randomizer)
+print(comments_rawVerYN_R)
 
@@
 #Create the volatile corpus that contains the "text" vector from data frame.
-comments_corpusVERB_R <- VCorpus(VectorSource(comments_rawVERB_R$text))
-print(comments_corpusVERB_R)
+comments_corpusVerYN_R <- VCorpus(VectorSource(comments_rawVerYN_R$text))
+print(comments_corpusVerYN_R)
 
 #Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
-inspect(comments_corpusVERB_R[1:3])
+inspect(comments_corpusVerYN_R[1:3])
 
 #Use "as.character" function to see what a single text comment looks like.
-as.character(comments_corpusVERB_R[[5]])
+as.character(comments_corpusVerYN_R[[5]])
 ```
@@
 #Convert text to all lower case letters.
-comments_corpus_cleanVERB_R <- tm_map(comments_corpusVERB_R, content_transformer(tolower))
+comments_corpus_cleanVerYN_R <- tm_map(comments_corpusVerYN_R, content_transformer(tolower))
 
@@
 #This step removes evidence of questions, so may remove data.
-comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removePunctuation)
+comments_corpus_cleanVerYN_R <- tm_map(comments_corpus_cleanVerYN_R, removePunctuation)
 
@@
 #I am not sure this is beneficial; does it remove evidence indicating technical comments?
-comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeNumbers)
+comments_corpus_cleanVerYN_R <- tm_map(comments_corpus_cleanVerYN_R, removeNumbers)
 
@@
 #The "stopwords" file is a convention; other files can be substituted for it
-comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeWords, stopwords())
+comments_corpus_cleanVerYN_R <- tm_map(comments_corpus_cleanVerYN_R, removeWords, stopwords())
 
@@
 #Final step removes white space from the document. This is NOT an optional step.
-comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, stripWhitespace)
+comments_corpus_cleanVerYN_R <- tm_map(comments_corpus_cleanVerYN_R, stripWhitespace)
 
 #Look at an example of cleaned text comment to see if it fits what is expected.
-as.character((comments_corpus_cleanVERB_R[[5]]))
+as.character((comments_corpus_cleanVerYN_R[[5]]))
 ```
@@
 NLP_Tokenizer <- function(x) {
-      unlist(lapply(ngrams(words(x), 3), paste, collapse = " "), use.names = FALSE)
+      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
 }
 
-comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB_R, control=list(tokenize = NLP_Tokenizer))
+comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVerYN_R, control=list(tokenize = NLP_Tokenizer))
 comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
@@
 ```{r}
-comments_train_labels_subject <- comments_rawVERB_R[1:7005, ]$subject
-comments_test_labels_subject <- comments_rawVERB_R[7006:9340,]$subject
+comments_train_labels_subject <- comments_rawVerYN_R[1:7005, ]$subject
+comments_test_labels_subject <- comments_rawVerYN_R[7006:9340,]$subject
 ```
@@
 ```{r}
-comments_train_labels_structure <- comments_rawVERB_R[1:7005, ]$structure
-comments_test_labels_structure <- comments_rawVERB_R[7006:9340,]$structure
+comments_train_labels_structure <- comments_rawVerYN_R[1:7005, ]$structure
+comments_test_labels_structure <- comments_rawVerYN_R[7006:9340,]$structure
 ```
@@
 ***
+
+####\
+
+##Initial Setup
+Calling the required libraries.
+
+```{}
+library(tidyverse)
+library(tidytext)
+library(tidyr)
+library(tm)
+library(SnowballC)
+library(wordcloud)
+library(e1071)
+library(gmodels)
+```
+
+####\
+
+##Version 2.0.yn of Naive Bayes (NB) Protocol
+###Initial Data Input
+```{}
+#Read in TA comments from CSV file.
+base_dataVerYN <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVerYN.csv')
+```
+
+###Extract, Organize Data Subset
+Extract comments and classification data subset
+```{}
+#Select rows representing the sub-groups to compare. 
+comments_subsetVerYN <- filter(base_dataVerYN,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")
+
+#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
+comments_rawVerYN <- comments_subsetVerYN %>% select(23,24,22)
+
+#Rename the columns.
+names(comments_rawVerYN)[1] <- "type"
+names(comments_rawVerYN)[2] <- "locus"
+names(comments_rawVerYN)[3] <- "text"
+
+#Change "type" element from character to a factor for analysis.
+comments_rawVerYN$type <- factor(comments_rawVerYN$type)
+str(comments_rawVerYN$type)
+table(comments_rawVerYN$type)
+```
+
+```{}
+#Change "locus" element from character to a factor for analysis.
+comments_rawVerYN$locus <- factor(comments_rawVerYN$locus)
+str(comments_rawVerYN$locus)
+table(comments_rawVerYN$locus)
+```
+
+***
+***
+Here is where I need to add the column used to randomly assign the data to training or testing sets.
+
+Use set.seed(1234) to create a stable randomized set.
+
+***
+***
+
+Convert data set to a volative corpus using "tm" library.
+```{}
+#Create the volatile corpus that contains the "text" vector from data frame.
+comments_corpusVerYN <- VCorpus(VectorSource(comments_rawVerYN$text))
+print(comments_corpusVerYN)
+
+#Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
+inspect(comments_corpusVerYN[1:3])
+
+#Use "as.character" function to see what a single text comment looks like.
+as.character(comments_corpusVerYN[[5]])
+```
+
+###Text Data Transformations
+The text data transforms are separated out as individual "tm_map"" and "content_transformer" functions to simplify the process of testing various combinations.
+Conversion to lower case and removing punctuation are the baseline set.
+Switch between "{r}" and "{}" to turn the other transformation blocks off or on.
+Removing extra white spaces must be the last transformation, and is required.
+
+```{}
+#Convert text to all lower case letters.
+comments_corpus_cleanVerYN <- tm_map(comments_corpusVerYN, content_transformer(tolower))
+```
+
+```{}
+#Remove punctuation using the "removePunctuation" function. 
+#This step removes evidence of questions, so may remove data.
+comments_corpus_cleanVerYN <- tm_map(comments_corpus_cleanVerYN, removePunctuation)
+```
+
+```{}
+#Remove numerals. 
+#I am not sure this is beneficial; does it remove evidence indicating technical comments?
+
+comments_corpus_cleanVerYN <- tm_map(comments_corpus_cleanVerYN, removeNumbers)
+```
+
+```{}
+#Stopword removal. 
+#This is a standard cleanup step, but again may remove useful terms.
+#The "stopwords" file is a convention; other files can be substituted for it
+
+comments_corpus_cleanVerYN <- tm_map(comments_corpus_cleanVerYN, removeWords, stopwords())
+```
+
+Option to stem terms has been removed from this version.
+
+Final prep and data check.
+```{}
+#Final step removes white space from the document. This is NOT an optional step.
+comments_corpus_cleanVerYN <- tm_map(comments_corpus_cleanVerYN, stripWhitespace)
+
+#Look at an example of cleaned text comment to see if it fits what is expected.
+as.character((comments_corpus_cleanVerYN[[5]]))
+```
+
+####\
+
+##Basic Naive Bayes (NB) Method Using N-grams
+This code block uses tm's NLP commands. The same block can be used to assess 1-, 2-, or 3-grams. Only the "#" variable in "(words(x), #)" below needs to be changed.
+
+```{}
+NLP_Tokenizer <- function(x) {
+      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
+}
+
+comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVerYN, control=list(tokenize = NLP_Tokenizer))
+comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
+inspect(comments_dtm_1gram)
+```
+
+####\
+
+###Data Preparation
+Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data.
+
+```{}
+.75 * 9340 #Number of rows in the dataset; product is #rows for training
+.25 * 9340 #Product is #rows for testing set
+```
+
+This code uses the comments in the order they are listed in the original file. **Need a block for  randomizing them.**
+```{}
+comments_dtm_train <- comments_dtm_1gram[1:7005, ]
+comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
+```
+
+Save vectors labeling rows in the training and testing vectors
+```{}
+comments_train_labels_type <- comments_rawVerYN[1:7005, ]$type
+comments_test_labels_type <- comments_rawVerYN[7006:9340,]$type
+```
+
+```{}
+comments_train_labels_locus <- comments_rawVerYN[1:7005, ]$locus
+comments_test_labels_locus <- comments_rawVerYN[7006:9340,]$locus
+```
+
+Make sure that the proportion of each sub-category is similar in the training and testing data set.
+```{}
+prop.table(table(comments_train_labels_type))
+prop.table(table(comments_test_labels_type))
+```
+
+```{}
+prop.table(table(comments_train_labels_locus))
+prop.table(table(comments_test_labels_locus))
+```
+
+####\
+
+###Preparation for Naive Bayes
+Remove words from the matrix that appear less than 5 times.
+```{}
+comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
+str(comments_freq_words)
+```
+
+Limit Document Term Matrix to only include words in the comments_freq_words vector. Using all the rows, but we want to limit the columns to these words in the frequency vector.
+```{}
+comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
+comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
+```
+
+The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables.
+This is done using a **convert_counts function** and applying it to the data.
+This replaces values greater than 0 with yes, and values not greater than 0 with no.
+
+```{}
+convert_counts2 <- function(x) {
+  x <- ifelse(x > 0, "Yes", "No")
+}
+
+comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
+comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
+```
+
+The resulting matrices have cells indicating "yes" or "no" if the word represented by the column appears in the text comment represented by the row.
+
+####\
+
+###Train Model, Predict, Evaluate for Type
+This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic.
+It evaluates the prediction with the actual data using a crosstable from the gmodels package.
+```{}
+comments_classifier_type <- naiveBayes(comments_train, comments_train_labels_type)
+comments_test_pred_type <- predict(comments_classifier_type, comments_test)
+CrossTable(comments_test_pred_type, comments_test_labels_type, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
+```
+
+The cross table list how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting TYPE.
+
+```{}
+print("Overall TYPE accuracy")
+(14+45+1346+19)/23.35
+```
+
+###Train Model, Predict, Evaluate for LOCUS
+This is the same block as above but uses the "Locus" data column.
+
+```{}
+comments_classifier_locus <- naiveBayes(comments_train, comments_train_labels_locus)
+comments_test_pred_locus <- predict(comments_classifier_locus, comments_test)
+CrossTable(comments_test_pred_locus, comments_test_labels_locus, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
+```
+
+Again, a cross table shows how many of the predicted categories match the correct categories for LOCUS.
+
+```{}
+print("Overall LOCUS accuracy")
+(62+64+931+3)/23.35
+```
+
+###OPTIONAL Output
+To make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame.
+The full data frame will be too long to display; use "head" to check format, then output the full dataset to CSV for further review.
+By default the CSV file should be saved to the current working directory.
+
+```{}
+predicted_category_type <- data.frame(comments_test_pred_type,comments_test_labels_type)
+head(predicted_category_type)
+write.csv(predicted_category_type, file = "PredictedCategoryType_Data.csv")
+```
+
+```{}
+predicted_category_locus <- data.frame(comments_test_pred_locus,comments_test_labels_locus)
+head(predicted_category_locus)
+write.csv(predicted_category_locus, file = "PredictedCategoryLocus_Data.csv")
+```
+
+***
+START COMMENTS
+***
+
+##Permutation Tests of Naive Bayes (NB) Method
+Using single words, Version 2.0.b correctly identifies comment "type" ~80% of the time and comment "locus" ~59% of the time.
+
+Permutations to Test:
+
+*  Compare 1-grams, 2-grams, 3-grams
+*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
+    +  Initial approach will be to split between largest group and balance in each category
+    +  Ex.: type is 3-technical vs. not 3_technical
+
+Known Issues:
+
+1. The e1071 naive bayes classifier is binomial. Need to incorporate multinomial.
+
+***
+END COMMENTS
+***

### 2019-02-05 Backups/Naive_bayes.Rmd
Update

--- 2018-12-21 Naive_bayes.Rmd
@@
 
-To keep this code clean for now I will run various permutations in a scratch version, then summarize findings starting on line #345. In these subsequent iterations I will evaluate:
+To keep this code clean for now I will run various permutations in a scratch version, then summarize findings starting on line #345. In the subsequent iterations I will evaluate:
 
@@
 *  How to randomize which comments are used for training, testing datasets.
+
+Other pages examine other classifiers besides Naive Bayes.
 
@@
 ```{r}
-#Remove numbers. 
+#Remove numerals. 
+#I am not sure this is beneficial; does it remove evidence indicating technical comments?
 comments_corpus_clean <- tm_map(comments_corpus_clean, removeNumbers)
@@
 #Stemming the text data to strip the suffix from words. 
+#My thinking is that stemming removes valuable tense data.I am unsure how valuable it is.
+
 comments_corpus_clean <- tm_map(comments_corpus_clean, stemDocument)
@@
 ```{r}
-#Final step removes white space from the document. This is NOT optional.
+#Final step removes white space from the document. This is NOT an optional step.
 comments_corpus_clean <- tm_map(comments_corpus_clean, stripWhitespace)
@@
 ```{r}
-(42+454+1186+185)/23.35
+(42+449+1185+186)/23.35
 ```
@@
 ##Modified Naive Bayes (NB) Method Using N-grams
-The tokenizer command block above only works for single words.
-The block below uses tm's NLP commands to generate either 1-, 2-, or 3-grams with the same command.
+The tokenizer command block above only works when using single words.
+The block below is an alternative method that uses tm's NLP commands to generate either 1-, 2-, or 3-grams with the same command.
 The only change needed is "#" in "(words(x), #)" below.

### 2019-02-05 Backups/Naive_bayes_V2a.Rmd
Update

--- 2019-02-05 Backups/Naive_bayes.Rmd
@@
 ---
-title: "Naive Bayes Workflow"
+title: "Naive Bayes Workflow Ver 2.0"
 author: "Dan Johnson"
-date: "12/21/2018"
+date: "1/29/2019"
 output: html_document
@@
 
-##Baseline Version 1.0 of Naive Bayes (NB) Protocol
+##Version 2.0 of Naive Bayes (NB) Protocol
 ###Initial Data Input
@@
 #Read in TA comments from CSV file.
-base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
+base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
+```
 
+###Extract, Organize Data Subset
+Extract comments and classification data subset
+```{r}
 #Select rows representing the sub-groups to compare. 
-comments_subset <- filter(base_data,code.subject=="1. Basic Criteria"|code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
+comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")
 
 #Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
-comments_raw <- comments_subset %>% select(23,22)
+comments_rawVERB <- comments_subsetVERB %>% select(23,24,22)
 
 #Rename the columns.
-names(comments_raw)[1] <- "type"
-names(comments_raw)[2] <- "text"
-
-#Simplify coding terms
-comments_raw[,1] <- ifelse(comments_raw[,1] == "1.
-Basic Criteria","1_Basic", ifelse(comments_raw[,1] == "2.
-Writing Quality","2_Writing", ifelse(comments_raw[,1] == "3.
-Technical and Scientific","3_Technical", ifelse(comments_raw[,1] == "4.
-Logic and Thinking","4_Logic",99))))
+names(comments_rawVERB)[1] <- "type"
+names(comments_rawVERB)[2] <- "locus"
+names(comments_rawVERB)[3] <- "text"
 
 #Change "type" element from character to a factor for analysis.
-comments_raw$type <- factor(comments_raw$type)
-str(comments_raw$type)
-table(comments_raw$type)
+comments_rawVERB$type <- factor(comments_rawVERB$type)
+str(comments_rawVERB$type)
+table(comments_rawVERB$type)
+```
+```{r}
+#Change "locus" element from character to a factor for analysis.
+comments_rawVERB$locus <- factor(comments_rawVERB$locus)
+str(comments_rawVERB$locus)
+table(comments_rawVERB$locus)
 ```
 
-Data set must be converted to a volative corpus using "tm" library then transformed.
+***
+***
+Here is where I need to add the column used to randomly assign the data to training or testing sets.
+
+Use set.seed(1234) to create a stable randomized set.
+
+***
+***
+
+Convert data set to a volative corpus using "tm" library.
 ```{r}
-#Create the volatile coprus that contains the "text" vector from data frame.
-comments_corpus <- VCorpus(VectorSource(comments_raw$text))
-print(comments_corpus)
+#Create the volatile corpus that contains the "text" vector from data frame.
+comments_corpusVERB <- VCorpus(VectorSource(comments_rawVERB$text))
+print(comments_corpusVERB)
 
-#Check out the first few messages in the new corpus, which is basically a list that can be manipulated with list operations.
-inspect(comments_corpus[1:3])
+#Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
+inspect(comments_corpusVERB[1:3])
 
-#Use "as.character" function to see what a message looks like.
-as.character(comments_corpus[[3]])
+#Use "as.character" function to see what a single text comment looks like.
+as.character(comments_corpusVERB[[5]])
 ```
 
-The OPTIONAL data transforms using "tm_map"" and "content_transformer" functions. Switch between "{r}" and "{}" to turn a transformation block off or on.
-```{r}
-#Convert to all lower case letters.
-comments_corpus_clean <- tm_map(comments_corpus, content_transformer(tolower))
-```
+###Text Data Transformations
+The text data transforms are separated out as individual "tm_map"" and "content_transformer" functions to simplify the process of testing various combinations.
+Conversion to lower case and removing punctuation are the baseline set.
+Switch between "{r}" and "{}" to turn the other transformation blocks off or on.
+Removing extra white spaces must be the last transformation, and is required.
 
 ```{r}
-#Remove numerals. 
-#I am not sure this is beneficial; does it remove evidence indicating technical comments?
-comments_corpus_clean <- tm_map(comments_corpus_clean, removeNumbers)
-```
-
-```{r}
-#Stopword removal. 
-comments_corpus_clean <- tm_map(comments_corpus_clean, removeWords, stopwords())
+#Convert text to all lower case letters.
+comments_corpus_cleanVERB <- tm_map(comments_corpusVERB, content_transformer(tolower))
 ```
@@
 #This step removes evidence of questions, so may remove data.
-comments_corpus_clean <- tm_map(comments_corpus_clean, removePunctuation)
+comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removePunctuation)
 ```
 
-```{r}
+```{}
+#Remove numerals. 
+#I am not sure this is beneficial; does it remove evidence indicating technical comments?
+
+comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removeNumbers)
+```
+
+```{}
+#Stopword removal. 
+#This is a standard cleanup step, but again may remove useful terms.
+#The "stopwords" file is a convention; other files can be substituted for it
+
+comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, removeWords, stopwords())
+```
+
+***
+START BREAK POINT - Stemming vs. lemmatization
+***
+
+```{}
 #Stemming the text data to strip the suffix from words. 
 #My thinking is that stemming removes valuable tense data.I am unsure how valuable it is.
+#Looking to later versions, part-of-speech tagging and lemmatization may be the better method.
 
-comments_corpus_clean <- tm_map(comments_corpus_clean, stemDocument)
+comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, stemDocument)
 ```
+
+***
+END BREAK POINT
+***
 
@@
 #Final step removes white space from the document. This is NOT an optional step.
-comments_corpus_clean <- tm_map(comments_corpus_clean, stripWhitespace)
+comments_corpus_cleanVERB <- tm_map(comments_corpus_cleanVERB, stripWhitespace)
 
-#Look at an example of data.
-as.character((comments_corpus_clean[[3]]))
+#Look at an example of cleaned text comment to see if it fits what is expected.
+as.character((comments_corpus_cleanVERB[[5]]))
 ```
@@
 
-###Basic Tokenizer
-This version uses the "DocumentTermMatrix" function to creates a matrix in which the rows indicate individual documents (individual TA comments) and the columns indicate presence or absence of the word that is the header for the column.
-Be careful not to confuse it with "TermDocumentMatrix", which transposes the matrix.
-
-"DocumentTermMaxtrix" function also can perform all of the text prep above in one command.
+##Basic Naive Bayes (NB) Method Using N-grams
+This code block uses tm's NLP commands. The same block can be used to assess 1-, 2-, or 3-grams. Only the "#" variable in "(words(x), #)" below needs to be changed.
 
 ```{r}
-comments_corpus_1gram <- DocumentTermMatrix(comments_corpus_clean)
-comments_corpus_1gram_tidy <- tidy(comments_corpus_1gram)
-inspect(comments_corpus_1gram)
+NLP_Tokenizer <- function(x) {
+      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
+}
+
+comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB, control=list(tokenize = NLP_Tokenizer))
+comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
+inspect(comments_dtm_1gram)
 ```
@@
 Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data.
+
 ```{r}
-.75 * 9340 #Number of documents/rows in the dataset; product is #rows for training
+.75 * 9340 #Number of rows in the dataset; product is #rows for training
 .25 * 9340 #Product is #rows for testing set
@@
 ```{r}
-comments_dtm_train <- comments_corpus_1gram[1:7005, ]
-comments_dtm_test <- comments_corpus_1gram[7006:9340, ]
+comments_dtm_train <- comments_dtm_1gram[1:7005, ]
+comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
 ```
@@
 ```{r}
-comments_train_labels <- comments_raw[1:7005, ]$type
-comments_test_labels <- comments_raw[7006:9340,]$type
+comments_train_labels <- comments_rawVERB[1:7005, ]$type
+comments_test_labels <- comments_rawVERB[7006:9340,]$type
+```
+
+```{r}
+comments_train_labels_locus <- comments_rawVERB[1:7005, ]$locus
+comments_test_labels_locus <- comments_rawVERB[7006:9340,]$locus
 ```
@@
 ```
+
+```{r}
+prop.table(table(comments_train_labels_locus))
+prop.table(table(comments_test_labels_locus))
+```
+
+***
+START BREAK POINT
+***
+The above code block incorrectly uses sequential roww, not random rows.  I need to find a way to incorporate randomization.
+
+The command block below can pull random lines from a data table, but does not work on a document term matrix.
+
+```{}
+set.seed(1234) 		# ensures the same random set gets picked each time.
+
+comments_raw_dtm_train <- sample_frac(comments_raw, .75)	# sample_frac pulls out 75% of "comments_raw" dataset randomly using the sample sequence identified by set.seed(1234)
+
+comments_raw_dtm_test <- setdiff(comments_raw, comments_raw_dtm_train)	#setdiff creates a "test" dataset using anything NOT in the "comments_raw_dtm_train" subset of "comments_raw."
+```
+
+Ways I think will work to solve this:
+1. Embed a randomized identifier as another column, then pull the later datasets out using that identifier.
+2. Generate a vector of random words (test, train, test, train, train, train, test, train...)
+
+***
+END BREAK POINT
+***
 
@@
 
-The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables.
-To do this we'll build a convert_counts function and apply it to our data.
+The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables.
+This is done using a **convert_counts function** and applying it to the data.
+This replaces values greater than 0 with yes, and values not greater than 0 with no.
+
 ```{r}
@@
 }
-```
 
-This replaces values greater than 0 with yes, and values not greater than 0 with no. Let's apply it to our data.
-```{r}
 comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
@@
 
-The resulting matrixes will be character type, with cells indicating "yes" or "no" if the word represented by the column appears in the message represented by the row.
+The resulting matrices have cells indicating "yes" or "no" if the word represented by the column appears in the text comment represented by the row.
 
@@
 ###Train Model, Predict, Evaluate
-Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group ONE, TWO, THREE, or FOUR.
+Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic.
 Evaluate the prediction with the actual data using a crosstable from the gmodels package.
@@
 
-Calculate overall correct prediction rate.
+The cross table will list how many of the predicted categories are in their correct categories. Adding diagonally and dividing will calculate the percent accuracy.
+
 ```{r}
-(42+449+1185+186)/23.35
+(44+473+1142+206)/23.35
+```
+
+To make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame.
+The full data frame will be too long to display; use "head" to check format, then output the full dataset to CSV for further review.
+By default the CSV file should be saved to the current working directory.
+
+```{r}
+predicted_category <- data.frame(comments_test_pred,comments_test_labels)
+head(predicted_category)
+write.csv(predicted_category, file = "PredictedCategory_Data.csv")
 ```
@@
 ***
-##Modified Naive Bayes (NB) Method Using N-grams
-The tokenizer command block above only works when using single words.
-The block below is an alternative method that uses tm's NLP commands to generate either 1-, 2-, or 3-grams with the same command.
-The only change needed is "#" in "(words(x), #)" below.
-For simplicity I left off all of the preparatory code up leading up to the final cleaned corpus.
+START BREAK POINT
+***
+1. The e1071 naive bayes classifier is binomial. Need to incorporate multinomial.
+2. Need to add prediction output.
+3. Is dropping numerical values down to categorical yes/no variables removing useful info?
 
-```{r}
-NLP_Tokenizer <- function(x) {
-      unlist(lapply(ngrams(words(x), 2), paste, collapse = " "), use.names = FALSE)
-}
-
-comments_dtm_2gram <- DocumentTermMatrix(comments_corpus_clean, control=list(tokenize = NLP_Tokenizer))
-comments_dtm_2gram_tidy <- tidy(comments_dtm_2gram)
-inspect(comments_dtm_2gram)
-```
-
-####\
-
-###Data Preparation
-Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data.
-
-```{r}
-.75 * 9340 #Number of rows in the dataset; product is #rows for training
-.25 * 9340 #Product is #rows for testing set
-```
-
-This code assumes comments are random. **Probably want to try randomizing them.**
-```{r}
-comments_dtm_train <- comments_dtm_2gram[1:7005, ]
-comments_dtm_test <- comments_dtm_2gram[7006:9340, ]
-```
-
-Save vectors labeling rows in the training and testing vectors
-```{r}
-comments_train_labels <- comments_raw[1:7005, ]$type
-comments_test_labels <- comments_raw[7006:9340,]$type
-```
-
-Make sure that the proportion of each sub-category is similar in the training and testing data set.
-```{r}
-prop.table(table(comments_train_labels))
-prop.table(table(comments_test_labels))
-```
-
-####\
-
-###Preparation for Naive Bayes
-Remove words from the matrix that appear less than 5 times.
-```{r}
-comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
-str(comments_freq_words)
-```
-
-Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
-```{r}
-comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
-comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
-```
-
-The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables.
-To do this we'll build a convert_counts function and apply it to our data.
-```{r}
-convert_counts2 <- function(x) {
-  x <- ifelse(x > 0, "Yes", "No")
-}
-```
-
-This replaces values greater than 0 with yes, and values not greater than 0 with no. Let's apply it to our data.
-```{r}
-comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
-comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
-```
-
-The resulting matrixes will be character type, with cells indicating "yes" or "no" if the word represented by the column appears in the message represented by the row.
-
-####\
-
-###Train Model, Predict, Evaluate
-Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group ONE, TWO, THREE, or FOUR.
-Evaluate the prediction with the actual data using a crosstable from the gmodels package.
-```{r}
-comments_classifier <- naiveBayes(comments_train, comments_train_labels)
-comments_test_pred <- predict(comments_classifier, comments_test)
-CrossTable(comments_test_pred, comments_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
-```
-
-```{r}
-(34+214+1254+94)/23.35
-```
+***
+END BREAK POINT
+***
 

### 2019-02-05 Backups/VERB.Rmd
Update

--- 2019-01-31 VERB.Rmd
@@
 
-*  Naive Bayes: https://adanieljohnson.github.io/default_website/Naive_bayes.html
+Naive Bayes
+
+*  Baseline Method: https://adanieljohnson.github.io/default_website/Naive_bayesV2b.html
+*  Randomization Method: https://adanieljohnson.github.io/default_website/Naive_bayesV2r.html
+*  In/Outgroup Binning: https://adanieljohnson.github.io/default_website/Naive_bayesV2yn.html
+*  In/Outgroup Binning: https://adanieljohnson.github.io/default_website/Naive_bayes_Smry.html
+
 *  To Be Decided #1. (link)

### 2019-02-05 draft.Rmd
Update

--- 2018-12-18 draft.Rmd
@@
 ---
-title: "Draft Description of Work in Progress"
+title: "Data Import and Initial Exploration"
 author: "Dan Johnson"
@@
 #Overview
-This page provides a draft overview of my progress.
-A short narrative outlines the  discrete blocks of work and how they are connected.
-Links connect to longer pages that document intake, munging, analysis, and visualization in depth.
+This page has three sections.
+**Part 1** describes the target dataset for this project.
+**Part 2** outlines the data pre-checks and import process.
+**Part 3** describes post-import validation developed for the target dataset, and how selected elements within the dataset are recoded for analysis.
+
+Initial visual exploration of the dataset is described here: https://adanieljohnson.github.io/default_website/initial_exploration.html
+
+These linked pages summarize work to date using categorical Naive Bayes as the text classification method. They provided a baseline from which I am evaluating other approaches.
+
+*  Baseline Method: https://adanieljohnson.github.io/default_website/Naive_bayesV2b.html
+*  Randomization Method: https://adanieljohnson.github.io/default_website/Naive_bayesV2r.html
+*  In/Outgroup Binning: https://adanieljohnson.github.io/default_website/Naive_bayesV2yn.html
+*  In/Outgroup Binning: https://adanieljohnson.github.io/default_website/Naive_bayes_Smry.html
+
+Other pages provide the theoretical background for the selected methods.
+Each deeper analysis method is described in detail on its own separate page.
+These are listed on the **[LINK] Project Outline** page.
 
@@
 
-##Initial Dataset
-My working dataset is an Excel table of 10,808 rows.
+#Part 1. Initial Dataset
+The development dataset is an Excel table of 10,808 rows.
 One column has "TA comments" extracted from student lab reports in Spring 2018.
@@
 
-Data were anonymized in Excel prior to conversion to CSV format. Names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings.
+Data were anonymized in Excel prior to conversion to CSV format. Names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings **[LINK]**.
 
@@
 
-##Blocks of Work
+#Part 2. Data Pre-Checks
+Prior to import, review the data table **in Excel**; check that:
 
-*  Data Pre-Checks and Import
-*  Initial Import and Post-Checks
-*  Preliminary Exploration: Log-Odds Text Comparison
-    +  Pearson Correlations of Word Frequencies: BROKEN SECTION
-    +  Mapping Analyses Visually: RUNS BUT CANNOT COPY AND MODIFY. It depends tables created using the broken code above. So I cannot run this part either.
-    +  Calculating and Plotting Log-Odds Pairs
-*  Preliminary Exploration: Direct Comparison of Word Lists
-*  Deeper Text Analysis Methods
-    +  Naive Bayes [/docs/Naive_bayes.html](link)
-    +  Random Forest (link)
-    +  Latent Dirichlet Allocation (link)
-    +  Etc. (link)
-    +  Etc. (link)
+1.  Student and TA names have been recoded using pre-assigned anonymous IDs.
+2.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.rmd"
+3.  Data table headers match those listed below. The same list is in the file "code_column_names.rmd"
 
-####\
-
-##Data Pre-Checks
-Prior to import, review data table in Excel. Check:
-
-1.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.rmd"
-2.  Data have been de-identified and recoded using pre-assigned anonymous IDs.
-3.  Data table headers are in snake_case, and match tabulated list below. The same list is in the file "code_column_names.rmd"
-
-Column Number|Column Name|Column Number|Column Name
-------------|----------------------|------------|----------------------
-1|unique.record|15|data.ok
-2|report.id|16|citation.ok
-3|sort|17|interpretation.ok
-4|report.title (empty)|18|organization.ok
-5|student (Std\_nnn)|19|techflaws.ok
-6|course|20|writing.ok
-7|ta (TA\_nnn)|21|comments.incorporated
-8|lab|22|ta.comment
-9|tag|23|code.subject
-10|type.TA|24|code.structure
-11|grade.TA|25|code.locus
-12|grading.time|26|code.scope
-13|Rank|27|code.tone
-14|hypothesis.ok|28|code.notes
+Column Number|Column Name|Accepted Values or Format
+------------|----------------------|-----------------------
+1|unique.record|Sp18.00001 (semester and year, then order in dataset)
+2|report.id|R_0fanqHFhaY2yLT7 or similar
+3|sort|remnant of coding
+4|report.title (empty)|string
+5|student|Std_drVZ2x1W (format is Std\_nnn)
+6|course|113, 114, or 214
+7|ta |TA_bc8EVX09 (format is TA\_nnn)
+8|lab|113: allocation, betta; 114: manduca, frog; 214: enzymes (old) photosynthesis, physarum, cyanophage (new)
+9|tag|first, second
+10|type.TA|submission, revision
+11|grade.TA|A, B, C, F
+12|grading.time|time in minutes
+13|Rank|remnant of coding
+14|hypothesis.ok|Yes, No
+15|data.ok|Yes, No
+16|citation.ok|Yes, No
+17|interpretation.ok|Yes, No
+18|organization.ok|Yes, No
+19|techflaws.ok|Yes, No
+20|writing.ok|Yes, No
+21|comments.incorporated|Yes, Somewhat, No
+22|ta.comment|text string
+23|code.subject|values in Ref. Tables
+24|code.structure|values in Ref. Tables
+25|code.locus|values in Ref. Tables
+26|code.scope|values in Ref. Tables
+27|code.tone|values in Ref. Tables
+28|code.notes|values in Ref. Tables
 
@@
 
-##Initial Import and Post-Check
+##Initial Library and Data Import
 ```{r Libraries_1}
@@
 ```{r}
-#This block reads in the full CSV into a starting dataframe named "base_data".
-#Subsequent analyses call in subsets of "base_data".
-base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
-
-#If dataset imports correctly, there will be a dataframe of ~11,000 rows and 28 columns named as listed in the **Data Pre-Checks** table above.
+#This block reads in full CSV into a starting dataframe named "base_data".
+#Subsequent analyses are performed on extracted subsets of "base_data".
+base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
 ```
 
+If dataset imports correctly, there will be a dataframe of ~11,000 rows and 28 columns named as listed in the **Data Pre-Checks** table above.
+
+####\
+
+#Part 3. Post-Import Checks
 Given some data points may be entered incorrectly by instructors or students, these post-import data checks help ensure data have been properly coded and columns contain valid entries only.
+(This process can be automated using the ```janitor``` data cleaning package.)
 
@@
 
-#"Lab" should only have entries that match the limited keywords in codebook.
+#"Lab" should only have entries that match the allowed limited keywords in codebook.
 lab<-unique(base_data$lab)
 
-#Look at the NUMBER of unique anonymous student IDs. Should be within 10% of (but not more than) the combined enrollment of the 3 target courses for the semester.
+#It is impractical to validate all randomized identifiers.
+Instead look at the NUMBER of unique anonymous student IDs.
+Quantity should be within 10% of (but not more than) the combined enrollment of the 3 target courses for the semester.
 student<-unique(base_data$student)
@@
 ```
-
 ####\
 
-## Preliminary Exploration
-The first question to answer before making deeper comparisons is **are the comments obviously different between sub-categories?** Before conducting detailed analyses I tried several iterations of different visual comparison strategies.
-
-####\
-
-###Log-Odds Text Comparison
-The log-odds ratios method outlined below finds and displays differences words that are more frequently found in TA comments belonging to one sub-category vs.
-another.
-I based this on a method for comparing word usage frequencies between two separate Twitter users [https://www.tidytextmining.com/twitter.html](that I found here.)
-
-For simplicity I reduced my analysis to just the four most common comment subjects. This has been the most informative preliminary visualization I have found so far.
+##Post-Import Processing
+Comment labels used in the project codebook contain punctuation and spacing that complicates subsequent analysis.
+Rather than modify the project codebook or external processes, it is simpler to recode the problematic labels after import.
+For example, the following code extracts a subset of rows from the full dataset, reduces those to 3 columns, and renames the **comment.subject** column as **subject**.
+Four problematic code labels in the subject column are replaced with simpler terms.
 
 ```{r}
-# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
-frequency_writing <- filter(base_data, code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
-frequency_writing.subcolumns <- frequency_writing %>% select(1, 22:23)
+#Read in TA comments from CSV file.
+base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
 
-# Tokenize phrases, remove stop words listed in standard reference file.
-base_data_tokenized <- frequency_writing.subcolumns %>% 
-  unnest_tokens(word,ta.comment) %>% 
-  anti_join(stop_words)
+#Select rows representing the sub-groups to compare. 
+comments_subset <- filter(base_data,code.subject=="1. Basic Criteria"|code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
 
-#At this point the dataframe should have "Unique.Record" in column 1, "code.subject" in column 2, and the unnested "word" in column 3.
-All columns are class = "character".
-Stop words listed in the "stop_words" R reference data file have been removed.
+#Reduce larger dataframe to 3 required columns of data (subject, structure, and text of comment), and put columns in order needed.
+comments_raw <- comments_subset %>% select(23,24,22)
 
+#Rename the columns.
+names(comments_raw)[1] <- "subject"
+names(comments_raw)[2] <- "structure"
+names(comments_raw)[3] <- "text"
+
+#Simplify coding terms used in "subject" column.
+comments_raw[,1] <- ifelse(comments_raw[,1] == "1.
+Basic Criteria","1_basic", ifelse(comments_raw[,1] == "2.
+Writing Quality","2_writing", ifelse(comments_raw[,1] == "3.
+Technical and Scientific","3_technical", ifelse(comments_raw[,1] == "4.
+Logic and Thinking","4_logic",99))))
+
+#Change "subject" element from character to a factor for analysis.
+comments_raw$subject <- factor(comments_raw$subject)
+str(comments_raw$subject)
+table(comments_raw$subject)
 ```
 
-####Comparing Pearson Correlations of Word Frequencies
-If the words used by TAs in their comments differ between the categories, there should be fairly low correlations overall .
-```{r}
-#Block below re-organizes the data for comparisons. This pattern sets "2. Writing Quality" as the first dataset, then all other data in remaining columns.
-#Groups by code.subject, calculated proportional frequency in each group.
-#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
-#Stats need this format to be able to compare 2. WQ against other values.
-base_data_tokenized_sorted2 <- base_data_tokenized %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  mutate(proportion = n / sum(n)) %>% 
-  select(-n) %>% 
-  spread(code.subject, proportion) %>% 
-  gather(code.subject, proportion, `3. Technical and Scientific`:`4. Logic and Thinking`)
-
-#The dataframe called "base_data_tokenized_sorted2" is the final 4-column dataframe for statistical analysis.
-Individual words are in column 1, "2.
-Writing Quality" frequency is in Column 2, "code.subject" is in column 3, and "proportion" is in column 4.
-Columns 1 and 3 are class = "character", 2 and 4 are numeric ranges.
-
-#This structure is needed in order to compare word (in Column 1) frequencies for 2.
-Writing Quality (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes.
-WHICH subset of X values to use is coded in Column 3.
-```
-
-**Pearson correlation** between the frequency of words in Writing Quality versus Technical and Logic sub-categoris.
-
-```{r}
-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3. Technical and Scientific",], 
-         ~ proportion + `2. Writing Quality`)
-```
-
-```{r}
-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4. Logic and Thinking",], 
-         ~ proportion + `2. Writing Quality`)
-```
-
-***
-***
-####START OF BROKEN SECTION
-THERE IS AN ERROR HERE I CANNOT DEBUG!
-If I change "3.
-Technical and Scientific"  to "2.
-Writing Quality", the structure of the tibble breaks.
-It goes from the correctly sorted 4 columns to incorrectly sorted 3 columns.
-It is NOT A TYPO; I can literally copy/paste the names in and out and break/correct the code.
-
-I cannot fix it using a different order for variables. I am at a loss for why it breaks.
-
-```{}
-# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
-frequency_writing3 <- filter(base_data, code.subject=="3. Technical and Scientific"|code.subject=="2. Writing Quality"|code.subject=="4. Logic and Thinking")
-frequency_writing3.subcolumns <- frequency_writing3 %>% select(1, 22:23)
-```
-
-```{}
-# Tokenize phrases, remove stop words listed in standard reference file.
-base_data_tokenized3 <- frequency_writing3.subcolumns %>% 
-  unnest_tokens(word,ta.comment) %>% 
-  anti_join(stop_words)
-```
-
-```{}
-#Block below re-organizes the data for comparisons. This pattern sets "3. Writing Quality" as the first dataset, then all other data in remaining columns.
-#Groups by code.subject, calculated proportional frequency in each group.
-#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
-#Stats need this format to be able to compare 2. WQ against other values.
-base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  mutate(proportion = n / sum(n)) %>% 
-  select(-n) %>% 
-  spread(code.subject, proportion) %>% 
-  gather(code.subject, proportion, `2. Writing Quality`:`4. Logic and Thinking`)
-```
-
-```{}
-#Create groups by code.subject, calculates proportional frequency in each group.
-#Final step creates 5 columns: code.subject, n, word, total, frequency.
-base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  left_join(base_data_tokenized %>% 
-              group_by(code.subject) %>% 
-              summarise(total = n())) %>%
-  mutate(freq = n/total)
-```
-
-Re-organize the data for comparisons. This pattern sets 3. Technical and Scientific as the first dataset, then all other data in remaining columns.
-```{}
-#Groups by code.subject, calculated proportional frequency in each group.
-#Final step creates 4 columns: word, 3. Technical and Scientific, code.subject, proportion
-#Stats need this format to be able to compare 3. TS against other values.
-base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  mutate(proportion = n / sum(n)) %>% 
-  select(-n) %>% 
-  spread(code.subject, proportion) %>% 
-  gather(code.subject, proportion, `2. Writing Quality`:`4. Logic and Thinking`)
-
-#END OF BROKEN SECTION
-```
-
-I can run this block, but CANNOT COPY AND MODIFY THIS SECTION. It needs the tables that should be created using the broken code above. So I cannot run this part either.
-
-```{}
-#Create groups by code.subject, calculates proportional frequency in each group.
-#Final step creates 5 columns: code.subject, n, word, total, frequency.
-base_data_tokenized_sorted2 <- base_data_tokenized %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  left_join(base_data_tokenized %>% 
-  group_by(code.subject) %>% 
-  summarise(total = n())) %>%
-  mutate(freq = n/total)
-```
-
-```{}
-base_data_tokenized_sorted2v3 <- base_data_tokenized_sorted2 %>% 
-  select(code.subject, word, freq) %>% 
-  spread(code.subject, freq) %>%
-  arrange(`2. Writing Quality`,`3. Technical and Scientific`)
-```
-
-```{}
-ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`3. Technical and Scientific`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-  #ggsave("Freqplot_3vs2.png", width = 10, height = 6)
-```
-
-```{}
-#I can get part way around the problem by re-using the same data and calling a different subset for the Y axis.
-Want to add comparisons of 3 vs 4, but same tables cannot be created because dataframe keeps breaking when I try to change 2.
-WQ.
-
-base_data_tokenized_sorted2v4 <- base_data_tokenized_sorted2 %>% 
-  select(code.subject, word, freq) %>% 
-  spread(code.subject, freq) %>%
-  arrange(`2. Writing Quality`,`4. Logic and Thinking`)
-```
-
-```{}
-ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`4. Logic and Thinking`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-  #ggsave("Freqplot_4vs2.png", width = 10, height = 6)
-```
-END of broken / limping sections
-***
-***
-
-###Calculating and Plotting Log-Odds Pairs
-
-```{r}
-word_ratios2v3 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`2. Writing Quality`/`3. Technical and Scientific`)) %>%
-  arrange(desc(logratio))
-
-word_ratios3v4 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`3. Technical and Scientific`/`4. Logic and Thinking`)) %>%
-  arrange(desc(logratio))
-
-word_ratios2v4 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`2. Writing Quality`/`4. Logic and Thinking`)) %>%
-  arrange(desc(logratio))
-```
-
-```{r}
-word_ratios2v3 %>% 
-  arrange(abs(logratio))
-
-word_ratios3v4 %>% 
-  arrange(abs(logratio))
-
-word_ratios2v4 %>% 
-  arrange(abs(logratio))
-```
-
-```{r}
-word_ratios2v3 %>%
-  group_by(logratio < 0) %>%
-  top_n(20, abs(logratio)) %>%
-  ungroup() %>%
-  mutate(word = reorder(word, logratio)) %>%
-  ggplot(aes(word, logratio, fill = logratio < 0)) +
-  geom_col(show.legend = FALSE) +
-  coord_flip() +
-  ylab("log odds ratio (Writing-top/Technical-bottom)") +
-  scale_fill_discrete(name = "", labels = c("2. Writing Quality", "3. Technical and Scientific"))
-  #ggsave("Logplot_2vs3.png", width = 6, height = 10)
-
-word_ratios3v4 %>%
-  group_by(logratio < 0) %>%
-  top_n(20, abs(logratio)) %>%
-  ungroup() %>%
-  mutate(word = reorder(word, logratio)) %>%
-  ggplot(aes(word, logratio, fill = logratio < 0)) +
-  geom_col(show.legend = FALSE) +
-  coord_flip() +
-  ylab("log odds ratio (Technical-top/Logic-bottom)") +
-  scale_fill_discrete(name = "", labels = c("3. Technical and Scientific","4. Logic and Thinking"))
-  #ggsave("Logplot_3vs4.png", width = 6, height = 10)
-
-word_ratios2v4 %>%
-  group_by(logratio < 0) %>%
-  top_n(20, abs(logratio)) %>%
-  ungroup() %>%
-  mutate(word = reorder(word, logratio)) %>%
-  ggplot(aes(word, logratio, fill = logratio < 0)) +
-  geom_col(show.legend = FALSE) +
-  coord_flip() +
-  ylab("log odds ratio (Writing-top/Logic-bottom)") +
-  scale_fill_discrete(name = "", labels = c("2. Writing Quality", "4. Logic and Thinking"))
-  #ggsave("Logplot_2vs4.png", width = 6, height = 10)
-
-```
-
-####\
-
-##Direct Comparison Using Only Words Appearing in All Subsets of Comments
-I think this is less informative, but it does let me see the data more clearly.
-
-Basic strategy is to take all comments from base_data, unnest tokens, filter stopwords, count and sort remaining terms, then write back to a new datatable called 'sortedwords.comments.all'.
-Calculate TOTAL words and store.
-Then calculate fractional frequencies for each word and store using mutate.
-
-I wrote this block so that, once CSV is imported, it does not require any of the code above. This is an early effort, so likely has some poorly developed code that could be refactored.
-
-```{}
-# Call in tidyverse, other required libraries.
-library(tidyverse)
-library(tidytext)
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-library(scales)
-
-#Read in the full CSV into a starting dataframe named "base_data".
-#Dataframe should have ~11,000 rows and 28 columns .
-base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
-```
-
-```{r}
-# Tokenize phrases, remove the stop words listed in R's standard reference file.
-base_data_tokenized <- base_data %>% 
-  unnest_tokens(word,ta.comment)
-
-# This is an ALTERNATIVE METHOD for pulling out stopwords.  
-sortedwords.comments.all <- base_data_tokenized %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate the number of words in total in the subset
-total_words.all <- sortedwords.comments.all %>% 
-  summarize(total = sum(n))
-
-#Mutate table to calculate then append fractional word frequency in the text.
-sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
-  all.fraction = (n*100)/total_words.all$total
-  )
-```
-
-```{r}
-#Pull SUBSETs of tokenized data based on code.subject: 
-base_data_tokenized.basic <- subset(base_data_tokenized, code.subject == "1. Basic Criteria", select = (1:28))
-base_data_tokenized.writing <- subset(base_data_tokenized, code.subject == "2. Writing Quality", select = (1:28))
-base_data_tokenized.technical <- subset(base_data_tokenized, code.subject == "3. Technical and Scientific", select = (1:28))
-base_data_tokenized.logic <- subset(base_data_tokenized, code.subject == "4. Logic and Thinking", select = (1:28))
-base_data_tokenized.praise <- subset(base_data_tokenized, code.subject == "5. Praise or Concern", select = (1:28))
-base_data_tokenized.misconduct <- subset(base_data_tokenized, code.subject == "6. Misconduct", select = (1:28))
-base_data_tokenized.narrative <- subset(base_data_tokenized, code.subject == "12. Narrative Comments", select = (1:28))
-base_data_tokenized.sciname <- subset(base_data_tokenized, code.subject == "Scientific Name", select = (1:28))
-```
-
-For each subset, filter stopwords, count words and total, then calculate frequencies.
-
-```{r}
-sortedwords.basic <- base_data_tokenized.basic %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate the number of words in total in the subset
-total_words.basic <- sortedwords.basic %>% 
-  summarize(total = sum(n))
-
-#Mutate table to calculate then append fractional word frequency in the text.
-sortedwords.basic.fraction <- sortedwords.basic %>% as_tibble() %>% mutate(
-  basic.fraction = (n*100)/total_words.basic$total
-)
-```
-
-```{r}
-sortedwords.writing <- base_data_tokenized.writing %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate the number of words in total in the subset
-total_words.writing <- sortedwords.writing %>% 
-  summarize(total = sum(n))
-
-#Mutate table to calculate then append fractional word frequency in the text.
-sortedwords.writing.fraction <- sortedwords.writing %>% as_tibble() %>% mutate(
-  writing.fraction = (n*100)/total_words.writing$total
-)
-
-sortedwords.technical <- base_data_tokenized.technical %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate the number of words in total in the subset
-total_words.technical <- sortedwords.technical %>% 
-  summarize(total = sum(n))
-
-#Mutate table to calculate then append fractional word frequency in the text.
-sortedwords.technical.fraction <- sortedwords.technical %>% as_tibble() %>% mutate(
-  technical.fraction = (n*100)/total_words.technical$total
-)
-
-sortedwords.logic <- base_data_tokenized.logic %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate the number of words in total in the subset
-total_words.logic <- sortedwords.logic %>% 
-  summarize(total = sum(n))
-
-#Mutate table to calculate then append fractional word frequency in the text.
-sortedwords.logic.fraction <- sortedwords.logic %>% as_tibble() %>% mutate(
-  logic.fraction = (n*100)/total_words.logic$total
-)
-```
-
-Next block joins the individual data tables for each subset in a larger file using column named 'word'.
-Columns get confusing names so they are renamed.
-**WARNING**: if a word is MISSING from **any subset** the data for that word is pruned from the full dataset.
-I am losing the unique words this way.
-
-**Need to figure out how to join without pruning.**
-
-```{r}
-total <- merge(sortedwords.comments.all.fraction,sortedwords.technical.fraction,by="word") 
-total2 <- merge(total,sortedwords.writing.fraction,by="word")
-total3 <- merge(total2,sortedwords.logic.fraction,by="word")
-#rename lines that were possibly confusing
-names(total3)[2] <- "all.count"
-names(total3)[4] <- "technical.count"
-names(total3)[6] <- "writing.count"
-names(total3)[8] <- "logic.count"
-
-#Mutate table to calculate then append fractional word frequency in the text.
-total4 <- total3 %>% as_tibble() %>% mutate(
-  writing.wt = (writing.count *100)/all.count
-)
-
-total5 <- total4 %>% as_tibble() %>% mutate(
-  technical.wt = (technical.count *100)/all.count
-)
-
-total6 <- total5 %>% as_tibble() %>% mutate(
-  logic.wt = (logic.count *100)/all.count
-)
-
-```
-
-The following six graphs compare frequencies for ~770 words that appear in ALL THREE of the sub-categories.
-
-Graphs 1-3 compare frequencies in individual groups against word frequency overall.
-Generally, the frequency of a term in the whole set of comments is similar to frequency in an individual group (X vs Y positions similar.) However, some terms associated with Logic subgroup clearly fall below the trendline Graph 3, suggesting they are less frequent in that sub-category than overall.
-
-```{r}
-ggplot(total6, aes(`all.fraction`,`writing.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-
-ggplot(total6, aes(`all.fraction`,`technical.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-
-ggplot(total6, aes(`all.fraction`,`logic.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-```
-
-Graphs 4-6 are pair-wise comparisons of frequencies for sub-categories.
-Unlike the comparisons to overall frequency, most words are NOT clustered along the trendline, suggesting there are greater differences in word usage frequencies.
-
-```{r}
-ggplot(total6, aes(`technical.fraction`,`writing.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-
-ggplot(total6, aes(`logic.fraction`,`writing.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-
-ggplot(total6, aes(`logic.fraction`,`technical.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-```
-
-####\
+Most in-depth analyses will have a similar extraction and post-processing code block.
+In practice, it has proven easier to modify the post-processing code than to maintain multiple separate copies of the original CSV dataset for the different analyses.

### 2019-02-05 index.Rmd
Update

@@
 ## Background
-I am interested in understanding how students develop technical writing skills in biology, and more specifically, how we can apply data science methods to improve that process.
+I am interested in understanding how students develop technical writing skills in biology, and more specifically, how to apply data science methods to improve that process.
 
 My current NSF-sponsored project asks if and how scripted instruction, automated support, and holistic feedback accelerates student development as writers.
-One of our specific aims is to see if changing the structure and focus of instructors' comments affects how students' writing develops over time.
-To do this, we need to extract instructor comments from student reports, classify those comments, and correlate the types of comments TAs make with student performance.
+One specific aim is to see if changing the structure and focus of instructors' comments affects how students' writing develops over time.
+To measure this we extract instructor comments from student reports, classify those comments, and correlate the types of comments TAs make with student performance.
 
@@
 
-* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
-* __Structure__: did the comment Point Out an error without providing other help, provide simple Copy Correction, provide declarative General or Specific Information only, or provide Holistic coaching that fosters student thinking and long-term improvement?
+* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic and thinking, or Other issues
+* __Structure__: did the comment Point to an error without providing other help, provide simple Copy Correction, provide declarative General or Specific Information only, or provide Holistic coaching that fosters student thinking and long-term improvement?
 
@@
 
+(Embed a copy of that table here.)
+
 #### \
+
 ## Specific Project Goals
-TA comments and meta-data for each student report can be extracted and compiled into a CSV datafile using existing Unix scripts.
+Unix scripts can extract and compile TA comments and meta-data for each student report into a CSV datafile.
 However routinely classifying TA comments by hand is impractical; 10-12,000 comments need to be scored EACH SEMESTER.
@@
 
-My specific project goals were/are to create an R-based text analysis workflow that:
+My specific project goals were/are to:
 
 1.
-**Converts TA comments and metadata into a de-identified "tidy" dataframe for R.** This includes converting the master CSV data set to an anonymized data frame with all names and confidential metadata re-keyed to unique random identifiers.
+**Create an R-based workflow that converts TA comments and metadata from student reports into a de-identified "tidy" dataframe for R.** This includes converting the master CSV data set to an anonymized data frame with all names and confidential metadata re-keyed to unique random identifiers.
 2.
-**Uses text features to assign TA comments to the categories defined in our existing codebook**.
+**Use text features to assign TA comments to the hand-coded categories defined in our existing codebook**.
 The goal is to be able to assign instructor comments into our pre-defined Subject and Structure codes with 90% accuracy or greater.
+3.
+**Assemble these methods into a semi-automated text classifier**.
+The final classifier model will be able to generate a two-axis contingency table similar to the one created using hand-coded data.
+It will be applied to newer datasets to determine if and how TA comments change over semesters.
 

### 2019-02-05 initial_exploration.Rmd
Update

--- 2018-12-18 draft.Rmd
@@
 ---
-title: "Draft Description of Work in Progress"
+title: "Initial Exploration"
 author: "Dan Johnson"
-date: "12/17/2018"
+date: "2/3/2019"
 output: html_document
-bibliography: library.bib
-biblio-style: apalike
 ---
-```{r, child="_setup.Rmd"}
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
-#Overview
-This page begins my draft outline of progress to date. Initial data import and exploration are described here. The other pages listed below document different pieces of the larger project.
 
-##Import and Initial Exploration
-This page describes to date on:
+#Preliminary Exploration of TA Comments Dataset
+The first question to answer before making deeper comparisons is **are TAs' comments obviously different between sub-categories?** This page describes several different visual comparisons I made of word frequencies before beginning in-depth text analyses.
 
-*  Data pre-checks, import, and post-import validation process.
-*  Initial comparisons of sub-categories using log-odds ratios of word frequencies
-    +  Pearson Correlations of Word Frequencies: BROKEN SECTION
-    +  Mapping Analyses Visually: RUNS BUT CANNOT COPY AND MODIFY. It depends tables created using the broken code above. So I cannot run this part either.
+*  Comparisons of word frequencies in subject sub-categories, using log-odds ratios
+    +  Pearson Correlations of Word Frequencies
+    +  Mapping Analyses Visually
     +  Calculating and Plotting Log-Odds Pairs
@@
 
-##Deeper Text Analysis
-These linked pages summarize work to date on each method.
-
-*  Naive Bayes: https://adanieljohnson.github.io/default_website/Naive_bayes.html
-*  Random Forest (link)
-*  Latent Dirichlet Allocation (link)
-*  Etc. (link)
-*  Etc. (link)
-
-####\
-
-#Initial Dataset
-My working dataset is an Excel table of 10,808 rows.
-One column has "TA comments" extracted from student lab reports in Spring 2018.
-Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation.
-Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.
-
-Data were anonymized in Excel prior to conversion to CSV format. Names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings.
-
-####\
-
-####\
-
-#Data Pre-Checks
-Prior to import, review data table in Excel. Check:
-
-1.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.rmd"
-2.  Data have been de-identified and recoded using pre-assigned anonymous IDs.
-3.  Data table headers are in snake_case, and match tabulated list below. The same list is in the file "code_column_names.rmd"
-
-Column Number|Column Name|Column Number|Column Name
-------------|----------------------|------------|----------------------
-1|unique.record|15|data.ok
-2|report.id|16|citation.ok
-3|sort|17|interpretation.ok
-4|report.title (empty)|18|organization.ok
-5|student (Std\_nnn)|19|techflaws.ok
-6|course|20|writing.ok
-7|ta (TA\_nnn)|21|comments.incorporated
-8|lab|22|ta.comment
-9|tag|23|code.subject
-10|type.TA|24|code.structure
-11|grade.TA|25|code.locus
-12|grading.time|26|code.scope
-13|Rank|27|code.tone
-14|hypothesis.ok|28|code.notes
-
-Once all data are in correct format, export Excel file to CSV with name formatted as **"coded_full_comments_dataset_SemYear.csv"**
-
-####\
-
-##Initial Import and Post-Check
+##Initial Library and Data Import
 ```{r Libraries_1}
@@
 ```{r}
-#This block reads in the full CSV into a starting dataframe named "base_data".
-#Subsequent analyses call in subsets of "base_data".
-base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
-
-#If dataset imports correctly, there will be a dataframe of ~11,000 rows and 28 columns named as listed in the **Data Pre-Checks** table above.
-```
-
-Given some data points may be entered incorrectly by instructors or students, these post-import data checks help ensure data have been properly coded and columns contain valid entries only.
-
-```{r}
-#"Course" column should only have entries that correspond to 113, 114, 214, NA. 
-course<-unique(base_data$course)
-
-#Entries in "TAs" column should match the anonymous IDs that correspond to the TAs assigned to the relevant courses in the semester being analyzed. 
-#If there are incorrect or extra anonymous IDs in this list, the mostly likely source of the error is improper de-identification of the Excel file prior to import.
-ta<-unique(base_data$ta)
-
-#"Lab" should only have entries that match the limited keywords in codebook.
-lab<-unique(base_data$lab)
-
-#Look at the NUMBER of unique anonymous student IDs. Should be within 10% of (but not more than) the combined enrollment of the 3 target courses for the semester.
-student<-unique(base_data$student)
-
-#Check that the list of "code.subject" topics here matches allowed terms in codebook.
-subject<-unique(base_data$code.subject)
-
-#Check that the list of "code.structure" terms extracted from the dataset and displayed here matches allowed terms in codebook.
-structure<-unique(base_data$code.structure)
-```
-
-####\
-
-## Preliminary Exploration
-The first question to answer before making deeper comparisons is **are the comments obviously different between sub-categories?** Before conducting detailed analyses I tried several iterations of different visual comparison strategies.
-
-####\
-
-###Log-Odds Text Comparison
-The log-odds ratios method outlined below finds and displays differences words that are more frequently found in TA comments belonging to one sub-category vs.
-another.
-I based this on a method for comparing word usage frequencies between two separate Twitter users [https://www.tidytextmining.com/twitter.html](that I found here.)
-
-For simplicity I reduced my analysis to just the four most common comment subjects. This has been the most informative preliminary visualization I have found so far.
-
-```{r}
-# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
-frequency_writing <- filter(base_data, code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
-frequency_writing.subcolumns <- frequency_writing %>% select(1, 22:23)
-
-# Tokenize phrases, remove stop words listed in standard reference file.
-base_data_tokenized <- frequency_writing.subcolumns %>% 
-  unnest_tokens(word,ta.comment) %>% 
-  anti_join(stop_words)
-
-#At this point the dataframe should have "Unique.Record" in column 1, "code.subject" in column 2, and the unnested "word" in column 3.
-All columns are class = "character".
-Stop words listed in the "stop_words" R reference data file have been removed.
-
-```
-
-####Comparing Pearson Correlations of Word Frequencies
-If the words used by TAs in their comments differ between the categories, there should be fairly low correlations overall .
-```{r}
-#Block below re-organizes the data for comparisons. This pattern sets "2. Writing Quality" as the first dataset, then all other data in remaining columns.
-#Groups by code.subject, calculated proportional frequency in each group.
-#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
-#Stats need this format to be able to compare 2. WQ against other values.
-base_data_tokenized_sorted2 <- base_data_tokenized %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  mutate(proportion = n / sum(n)) %>% 
-  select(-n) %>% 
-  spread(code.subject, proportion) %>% 
-  gather(code.subject, proportion, `3. Technical and Scientific`:`4. Logic and Thinking`)
-
-#The dataframe called "base_data_tokenized_sorted2" is the final 4-column dataframe for statistical analysis.
-Individual words are in column 1, "2.
-Writing Quality" frequency is in Column 2, "code.subject" is in column 3, and "proportion" is in column 4.
-Columns 1 and 3 are class = "character", 2 and 4 are numeric ranges.
-
-#This structure is needed in order to compare word (in Column 1) frequencies for 2.
-Writing Quality (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes.
-WHICH subset of X values to use is coded in Column 3.
-```
-
-**Pearson correlation** between the frequency of words in Writing Quality versus Technical and Logic sub-categoris.
-
-```{r}
-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3. Technical and Scientific",], 
-         ~ proportion + `2. Writing Quality`)
-```
-
-```{r}
-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4. Logic and Thinking",], 
-         ~ proportion + `2. Writing Quality`)
-```
-
-***
-***
-####START OF BROKEN SECTION
-THERE IS AN ERROR HERE I CANNOT DEBUG!
-If I change "3.
-Technical and Scientific"  to "2.
-Writing Quality", the structure of the tibble breaks.
-It goes from the correctly sorted 4 columns to incorrectly sorted 3 columns.
-It is NOT A TYPO; I can literally copy/paste the names in and out and break/correct the code.
-
-I cannot fix it using a different order for variables. I am at a loss for why it breaks.
-
-```{}
-# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
-frequency_writing3 <- filter(base_data, code.subject=="3. Technical and Scientific"|code.subject=="2. Writing Quality"|code.subject=="4. Logic and Thinking")
-frequency_writing3.subcolumns <- frequency_writing3 %>% select(1, 22:23)
-```
-
-```{}
-# Tokenize phrases, remove stop words listed in standard reference file.
-base_data_tokenized3 <- frequency_writing3.subcolumns %>% 
-  unnest_tokens(word,ta.comment) %>% 
-  anti_join(stop_words)
-```
-
-```{}
-#Block below re-organizes the data for comparisons. This pattern sets "3. Writing Quality" as the first dataset, then all other data in remaining columns.
-#Groups by code.subject, calculated proportional frequency in each group.
-#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
-#Stats need this format to be able to compare 2. WQ against other values.
-base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  mutate(proportion = n / sum(n)) %>% 
-  select(-n) %>% 
-  spread(code.subject, proportion) %>% 
-  gather(code.subject, proportion, `2. Writing Quality`:`4. Logic and Thinking`)
-```
-
-```{}
-#Create groups by code.subject, calculates proportional frequency in each group.
-#Final step creates 5 columns: code.subject, n, word, total, frequency.
-base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  left_join(base_data_tokenized %>% 
-              group_by(code.subject) %>% 
-              summarise(total = n())) %>%
-  mutate(freq = n/total)
-```
-
-Re-organize the data for comparisons. This pattern sets 3. Technical and Scientific as the first dataset, then all other data in remaining columns.
-```{}
-#Groups by code.subject, calculated proportional frequency in each group.
-#Final step creates 4 columns: word, 3. Technical and Scientific, code.subject, proportion
-#Stats need this format to be able to compare 3. TS against other values.
-base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  mutate(proportion = n / sum(n)) %>% 
-  select(-n) %>% 
-  spread(code.subject, proportion) %>% 
-  gather(code.subject, proportion, `2. Writing Quality`:`4. Logic and Thinking`)
-
-#END OF BROKEN SECTION
-```
-
-I can run this block, but CANNOT COPY AND MODIFY THIS SECTION. It needs the tables that should be created using the broken code above. So I cannot run this part either.
-
-```{}
-#Create groups by code.subject, calculates proportional frequency in each group.
-#Final step creates 5 columns: code.subject, n, word, total, frequency.
-base_data_tokenized_sorted2 <- base_data_tokenized %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  left_join(base_data_tokenized %>% 
-  group_by(code.subject) %>% 
-  summarise(total = n())) %>%
-  mutate(freq = n/total)
-```
-
-```{}
-base_data_tokenized_sorted2v3 <- base_data_tokenized_sorted2 %>% 
-  select(code.subject, word, freq) %>% 
-  spread(code.subject, freq) %>%
-  arrange(`2. Writing Quality`,`3. Technical and Scientific`)
-```
-
-```{}
-ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`3. Technical and Scientific`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-  #ggsave("Freqplot_3vs2.png", width = 10, height = 6)
-```
-
-```{}
-#I can get part way around the problem by re-using the same data and calling a different subset for the Y axis.
-Want to add comparisons of 3 vs 4, but same tables cannot be created because dataframe keeps breaking when I try to change 2.
-WQ.
-
-base_data_tokenized_sorted2v4 <- base_data_tokenized_sorted2 %>% 
-  select(code.subject, word, freq) %>% 
-  spread(code.subject, freq) %>%
-  arrange(`2. Writing Quality`,`4. Logic and Thinking`)
-```
-
-```{}
-ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`4. Logic and Thinking`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-  #ggsave("Freqplot_4vs2.png", width = 10, height = 6)
-```
-END of broken / limping sections
-***
-***
-
-###Calculating and Plotting Log-Odds Pairs
-
-```{r}
-word_ratios2v3 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`2. Writing Quality`/`3. Technical and Scientific`)) %>%
-  arrange(desc(logratio))
-
-word_ratios3v4 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`3. Technical and Scientific`/`4. Logic and Thinking`)) %>%
-  arrange(desc(logratio))
-
-word_ratios2v4 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`2. Writing Quality`/`4. Logic and Thinking`)) %>%
-  arrange(desc(logratio))
-```
-
-```{r}
-word_ratios2v3 %>% 
-  arrange(abs(logratio))
-
-word_ratios3v4 %>% 
-  arrange(abs(logratio))
-
-word_ratios2v4 %>% 
-  arrange(abs(logratio))
-```
-
-```{r}
-word_ratios2v3 %>%
-  group_by(logratio < 0) %>%
-  top_n(20, abs(logratio)) %>%
-  ungroup() %>%
-  mutate(word = reorder(word, logratio)) %>%
-  ggplot(aes(word, logratio, fill = logratio < 0)) +
-  geom_col(show.legend = FALSE) +
-  coord_flip() +
-  ylab("log odds ratio (Writing-top/Technical-bottom)") +
-  scale_fill_discrete(name = "", labels = c("2. Writing Quality", "3. Technical and Scientific"))
-  #ggsave("Logplot_2vs3.png", width = 6, height = 10)
-
-word_ratios3v4 %>%
-  group_by(logratio < 0) %>%
-  top_n(20, abs(logratio)) %>%
-  ungroup() %>%
-  mutate(word = reorder(word, logratio)) %>%
-  ggplot(aes(word, logratio, fill = logratio < 0)) +
-  geom_col(show.legend = FALSE) +
-  coord_flip() +
-  ylab("log odds ratio (Technical-top/Logic-bottom)") +
-  scale_fill_discrete(name = "", labels = c("3. Technical and Scientific","4. Logic and Thinking"))
-  #ggsave("Logplot_3vs4.png", width = 6, height = 10)
-
-word_ratios2v4 %>%
-  group_by(logratio < 0) %>%
-  top_n(20, abs(logratio)) %>%
-  ungroup() %>%
-  mutate(word = reorder(word, logratio)) %>%
-  ggplot(aes(word, logratio, fill = logratio < 0)) +
-  geom_col(show.legend = FALSE) +
-  coord_flip() +
-  ylab("log odds ratio (Writing-top/Logic-bottom)") +
-  scale_fill_discrete(name = "", labels = c("2. Writing Quality", "4. Logic and Thinking"))
-  #ggsave("Logplot_2vs4.png", width = 6, height = 10)
-
+#This block reads in full CSV into a starting dataframe named "base_data".
+#Subsequent analyses are performed on extracted subsets of "base_data".
+base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
 ```
@@
 I wrote this block so that, once CSV is imported, it does not require any of the code above. This is an early effort, so likely has some poorly developed code that could be refactored.
-
-```{}
-# Call in tidyverse, other required libraries.
-library(tidyverse)
-library(tidytext)
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-library(scales)
-
-#Read in the full CSV into a starting dataframe named "base_data".
-#Dataframe should have ~11,000 rows and 28 columns .
-base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
-```
 
@@
 #Pull SUBSETs of tokenized data based on code.subject: 
-base_data_tokenized.basic <- subset(base_data_tokenized, code.subject == "1. Basic Criteria", select = (1:28))
-base_data_tokenized.writing <- subset(base_data_tokenized, code.subject == "2. Writing Quality", select = (1:28))
-base_data_tokenized.technical <- subset(base_data_tokenized, code.subject == "3. Technical and Scientific", select = (1:28))
-base_data_tokenized.logic <- subset(base_data_tokenized, code.subject == "4. Logic and Thinking", select = (1:28))
-base_data_tokenized.praise <- subset(base_data_tokenized, code.subject == "5. Praise or Concern", select = (1:28))
-base_data_tokenized.misconduct <- subset(base_data_tokenized, code.subject == "6. Misconduct", select = (1:28))
-base_data_tokenized.narrative <- subset(base_data_tokenized, code.subject == "12. Narrative Comments", select = (1:28))
-base_data_tokenized.sciname <- subset(base_data_tokenized, code.subject == "Scientific Name", select = (1:28))
+base_data_tokenized.basic <- subset(base_data_tokenized, code.subject == "1_basic", select = (1:28))
+base_data_tokenized.writing <- subset(base_data_tokenized, code.subject == "2_writing", select = (1:28))
+base_data_tokenized.technical <- subset(base_data_tokenized, code.subject == "3_technical", select = (1:28))
+base_data_tokenized.logic <- subset(base_data_tokenized, code.subject == "4_logic", select = (1:28))
+#base_data_tokenized.praise <- subset(base_data_tokenized, code.subject == "5_exposition", select = (1:28))
+#base_data_tokenized.misconduct <- subset(base_data_tokenized, code.subject == "6_misconduct", select = (1:28))
+#base_data_tokenized.narrative <- subset(base_data_tokenized, code.subject == "12. Narrative Comments", select = (1:28))
+#base_data_tokenized.sciname <- subset(base_data_tokenized, code.subject == "Scientific Name", select = (1:28))
 ```
@@
 
-Next block joins the individual data tables for each subset in a larger file using column named 'word'.
-Columns get confusing names so they are renamed.
-**WARNING**: if a word is MISSING from **any subset** the data for that word is pruned from the full dataset.
-I am losing the unique words this way.
+Next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed.
 
-**Need to figure out how to join without pruning.**
+**WARNING**: if a word is MISSING from **any subset** the data for that word is pruned from the full dataset. I am losing the unique words this way. **Need to figure out how to join without pruning.**
 
@@
 ####\
+
+##Log-Odds Ratio for Text Comparison Between Text Categories
+The log-odds ratios method outlined below finds and displays differences words that are more frequently found in TA comments belonging to one sub-category vs.
+another.
+I based this on a method for comparing word usage frequencies between two separate Twitter users [https://www.tidytextmining.com/twitter.html](that I found here.) This page has two parts, one for the comparison for code.subject, and a second part for code.structure category.
+
+####\
+
+###Part 1. Log-Odds Ratio Comparisons Between Subject Categories
+For simplicity I reduced my analysis to just the four most common comment subjects. This has been the most informative preliminary visualization I have found so far.
+
+```{r}
+# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
+frequency_writing <- filter(base_data, code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")
+frequency_writing.subcolumns <- frequency_writing %>% select(1, 22:23)
+
+# Tokenize phrases, remove stop words listed in standard reference file.
+base_data_tokenized <- frequency_writing.subcolumns %>% 
+  unnest_tokens(word,ta.comment) %>% 
+  anti_join(stop_words)
+
+#At this point the dataframe should have "Unique.Record" in column 1, "code.subject" in column 2, and the unnested "word" in column 3.
+All columns are class = "character".
+Stop words listed in the "stop_words" R reference data file have been removed.
+
+```
+
+####Pearson Correlations of Log-Odds Word Frequencies
+If the words used by TAs in their comments differ between the categories, there should be fairly low correlations overall.
+
+```{r}
+#Block below re-organizes the data for comparisons. This pattern sets "2_writing" as the first dataset, then all other data in remaining columns.
+#Groups by code.subject, calculated proportional frequency in each group.
+#Final step creates 4 columns: word, 2_writing, code.subject, proportion
+#Stats need this format to be able to compare 2_writing against other values.
+base_data_tokenized_sorted2 <- base_data_tokenized %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  mutate(proportion = n / sum(n)) %>% 
+  select(-n) %>% 
+  spread(code.subject, proportion) %>% 
+  gather(code.subject, proportion, `3_technical`:`4_logic`)
+
+#The dataframe called "base_data_tokenized_sorted2" is the final 4-column dataframe for statistical analysis.
+Individual words are in column 1, "2_writing" frequency is in Column 2, "code.subject" is in column 3, and "proportion" is in column 4.
+Columns 1 and 3 are class = "character", 2 and 4 are numeric ranges.
+
+#This structure is needed in order to compare word (in Column 1) frequencies for 2.
+Writing Quality (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes.
+WHICH subset of X values to use is coded in Column 3.
+```
+
+**Pearson correlation** between frequency of words in 2_writing versus 3_technical and 2_writing versus 4_logic sub-categories.
+
+```{r}
+cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3_technical",], 
+         ~ proportion + `2_writing`)
+```
+
+```{r}
+cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4_logic",], 
+         ~ proportion + `2_writing`)
+```
+
+```{r}
+coroutput23 <-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3_technical",], 
+         ~ proportion + `2_writing`)
+coroutput24 <-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4_logic",], 
+         ~ proportion + `2_writing`)
+broom::tidy(coroutput23)
+broom::tidy(coroutput24)
+
+```
+
+```{r}
+knitr::kable(coroutput23)
+```
+
+Comparing sub-groups 3_technical and 4_logic requires recreating the dataframe in slightly different format that excludes the data rows for 2_writing.
+
+```{r}
+# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
+frequency_writing3 <- filter(base_data, code.subject=="3_technical"|code.subject=="4_logic")
+frequency_writing.subcolumns3 <- frequency_writing3 %>% select(1, 22:23)
+
+# Tokenize phrases, remove stop words listed in standard reference file.
+base_data_tokenized3 <- frequency_writing.subcolumns3 %>% 
+  unnest_tokens(word,ta.comment) %>% 
+  anti_join(stop_words)
+
+#At this point the dataframe should have "Unique.Record" in column 1, "code.subject" in column 2, and the unnested "word" in column 3.
+All columns are class = "character".
+Stop words listed in the "stop_words" R reference data file have been removed.
+
+```
+
+```{r}
+#Block below re-organizes the data for comparisons. This pattern sets "3_technical" as the first dataset, then all other data in remaining columns.
+#Groups by code.subject, calculated proportional frequency in each group.
+#Final step creates 4 columns: word, 3_technical, code.subject, proportion
+#Stats need this format to be able to compare 3_technical against other values.
+base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  mutate(proportion = n / sum(n)) %>% 
+  select(-n) %>% 
+  spread(code.subject, proportion) %>% 
+  gather(code.subject, proportion,`4_logic`)
+
+#The dataframe called "base_data_tokenized_sorted3" is the final 4-column dataframe for statistical analysis.
+Individual words are in column 1, "3_technical" frequency is in Column 2, "code.subject" is in column 3, and "proportion" is in column 4.
+Columns 1 and 3 are class = "character", 2 and 4 are numeric ranges.
+
+#This structure is needed in order to compare word (in Column 1) frequencies for 3_technical (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes.
+WHICH subset of X values to use is coded in Column 3.
+```
+
+**Pearson correlation** between the frequency of words in 3_technical versus 4_logic sub-categories.
+
+```{r}
+cor.test(data = base_data_tokenized_sorted3[base_data_tokenized_sorted3$code.subject == "4_logic",], 
+         ~ proportion + `3_technical`)
+```
+
+####Visualizing Word Frequency Differences Between Subject Sub-Categories
+
+```{r}
+#Create groups by code.subject, calculates proportional frequency in each group.
+#Final step creates 5 columns: code.subject, n, word, total, frequency.
+base_data_tokenized_sorted2 <- base_data_tokenized %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  left_join(base_data_tokenized %>% 
+  group_by(code.subject) %>% 
+  summarise(total = n())) %>%
+  mutate(freq = n/total)
+```
+
+```{r}
+base_data_tokenized_sorted2v3 <- base_data_tokenized_sorted2 %>% 
+  select(code.subject, word, freq) %>% 
+  spread(code.subject, freq) %>%
+  arrange(`2_writing`,`3_technical`)
+```
+
+```{r}
+ggplot(base_data_tokenized_sorted2v3, aes(`2_writing`,`3_technical`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+  #ggsave("Freqplot_3vs2.png", width = 10, height = 6)
+```
+
+```{r}
+#Comparing 2 vs. 4.
+base_data_tokenized_sorted2v4 <- base_data_tokenized_sorted2 %>% 
+  select(code.subject, word, freq) %>% 
+  spread(code.subject, freq) %>%
+  arrange(`2_writing`,`4_logic`)
+```
+
+```{r}
+ggplot(base_data_tokenized_sorted2v3, aes(`2_writing`,`4_logic`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+  #ggsave("Freqplot_4vs2.png", width = 10, height = 6)
+```
+
+```{r}
+#Create groups by code.subject, calculates proportional frequency in each group.
+#Final step creates 5 columns: code.subject, n, word, total, frequency.
+base_data_tokenized_sorted3 <- base_data_tokenized %>% 
+  group_by(code.subject) %>%
+  count(code.subject, word, sort=TRUE) %>% 
+  left_join(base_data_tokenized %>% 
+  group_by(code.subject) %>% 
+  summarise(total = n())) %>%
+  mutate(freq = n/total)
+```
+
+```{r}
+#Set up comparison for 3 vs. 4.
+base_data_tokenized_sorted3v4 <- base_data_tokenized_sorted3 %>% 
+  select(code.subject, word, freq) %>% 
+  spread(code.subject, freq) %>%
+  arrange(`2_writing`,`4_logic`)
+```
+
+```{r}
+ggplot(base_data_tokenized_sorted3v4, aes(`3_technical`,`4_logic`)) +
+  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
+  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
+  scale_x_log10(labels = percent_format()) +
+  scale_y_log10(labels = percent_format()) +
+  geom_abline(color = "red")
+  #ggsave("Freqplot_4vs3.png", width = 10, height = 6)
+```
+
+####Calculating and Plotting Log-Odds Pairs for Subject Sub-Categories
+
+```{r}
+word_ratios2v3 <- base_data_tokenized %>%
+  count(word, code.subject) %>%
+  group_by(word) %>%
+  filter(sum(n) >= 10) %>%
+  ungroup() %>%
+  spread(code.subject, n, fill = 0) %>%
+  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
+  mutate(logratio = log(`2_writing`/`3_technical`)) %>%
+  arrange(desc(logratio))
+
+word_ratios3v4 <- base_data_tokenized %>%
+  count(word, code.subject) %>%
+  group_by(word) %>%
+  filter(sum(n) >= 10) %>%
+  ungroup() %>%
+  spread(code.subject, n, fill = 0) %>%
+  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
+  mutate(logratio = log(`3_technical`/`4_logic`)) %>%
+  arrange(desc(logratio))
+
+word_ratios2v4 <- base_data_tokenized %>%
+  count(word, code.subject) %>%
+  group_by(word) %>%
+  filter(sum(n) >= 10) %>%
+  ungroup() %>%
+  spread(code.subject, n, fill = 0) %>%
+  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
+  mutate(logratio = log(`2_writing`/`4_logic`)) %>%
+  arrange(desc(logratio))
+```
+
+```{r}
+word_ratios2v3 %>% 
+  arrange(abs(logratio))
+
+word_ratios3v4 %>% 
+  arrange(abs(logratio))
+
+word_ratios2v4 %>% 
+  arrange(abs(logratio))
+```
+
+```{r}
+word_ratios2v3 %>%
+  group_by(logratio < 0) %>%
+  top_n(20, abs(logratio)) %>%
+  ungroup() %>%
+  mutate(word = reorder(word, logratio)) %>%
+  ggplot(aes(word, logratio, fill = logratio < 0)) +
+  geom_col(show.legend = FALSE) +
+  coord_flip() +
+  ylab("log odds ratio (Writing-top/Technical-bottom)") +
+  scale_fill_discrete(name = "", labels = c("2_writing", "3_technical"))
+  #ggsave("Logplot_2vs3.png", width = 6, height = 10)
+
+word_ratios3v4 %>%
+  group_by(logratio < 0) %>%
+  top_n(20, abs(logratio)) %>%
+  ungroup() %>%
+  mutate(word = reorder(word, logratio)) %>%
+  ggplot(aes(word, logratio, fill = logratio < 0)) +
+  geom_col(show.legend = FALSE) +
+  coord_flip() +
+  ylab("log odds ratio (Technical-top/Logic-bottom)") +
+  scale_fill_discrete(name = "", labels = c("3_technical","4_logic"))
+  #ggsave("Logplot_3vs4.png", width = 6, height = 10)
+
+word_ratios2v4 %>%
+  group_by(logratio < 0) %>%
+  top_n(20, abs(logratio)) %>%
+  ungroup() %>%
+  mutate(word = reorder(word, logratio)) %>%
+  ggplot(aes(word, logratio, fill = logratio < 0)) +
+  geom_col(show.legend = FALSE) +
+  coord_flip() +
+  ylab("log odds ratio (Writing-top/Logic-bottom)") +
+  scale_fill_discrete(name = "", labels = c("2_writing", "4_logic"))
+  #ggsave("Logplot_2vs4.png", width = 6, height = 10)
+
+```
+
+####\
+
+####\
+
+###Part 2. Log-Odds Ratio Comparisons Between Structure Categories
+
+####\

### 2019-02-21 background.Rmd
cleanup

--- 2018-12-16 project_backup.Rmd
@@
 ---
-title: "Project"
-author: "Dan Johnson"
-date: "10/26/2018"
-output: html_document
+title: "Background for the Project"
 bibliography: library.bib
@@
 ---
+
 ```{r, child="_setup.Rmd"}
@@
 
-# Abstract
-
-####\
-
-# Introduction
-## Background
 Undergraduate introductory STEM courses typically include technical writing training.
@@
 
-In 2014 we launched an automated data collection process we call SAWHET that collects systematic observations and sample data from student writers and graduate instructors.
+In 2014 we launched an automated data collection process we call **SAWHET** that systematically collects structured writing and other data from student writers and graduate instructors.
 Our resarch goal is to better understand student and instructor behaviors within the technical writing instructional space.
-Our practical goal is to develop techniques for teaching technical writing that:
+From this empirical basis our practical goal is to develop techniques for teaching technical writing that:
 
@@
 *  Emphasize student development through holistic coaching;
-*  Shift student and instructor attention away from simple copy-correction towards larger writing issues; and
-*  Embody other best practices outlined in nearly 40 years of research literature on "writing in the disciplines."
+*  Shift student and instructor attention away from copy-correction towards larger writing issues; and
+*  Embody other best practices from 40+ years of research literature on "writing in the disciplines."
 
@@
 ####\
-
-# Literature Review
-##General Approach to Initial Dataset
-The following excerpt from *Text Mining with R: A Tidy Approach* (Julia Silge and David Robinson) summarizes how extended text (in this case, TA comments) can be organized as a tidy matrix.*All emphases are mine.*
-
->We thus define the tidy text format as being a table with one-token-per-row.
-A token is a **meaningful unit of text**, such as a word, that we are interested in using for analysis, and tokenization is the process of splitting text into tokens.
->This one-token-per-row structure is in contrast to the ways text is often stored in current analyses, perhaps as strings or in a document-term matrix.
-For tidy text mining, the token that is stored in each row is most often a single word, **but can also be an n-gram, sentence, or paragraph** .
-In the tidytext package, we provide functionality to tokenize by commonly used units of text like these and convert to a one-term-per-row format.
-
-Using tidy format (vs. other common text data formats) better maintains relationships between individual comments and metadata, which simplifies subsequent analysis.
-
-####\
-
-# Methods
-Our initial working dataset is an anonymized Excel table with one column of "TA comments" extracted from student lab reports.
-Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation.
-Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.
-
-####\
-
-##Data Pre-Checks
-A CSV file named "coded_full_comments_dataset_Spring18anon.csv" should have correctly organized text and meta-data. Prior to import, review data table in Excel. Check:
-
-1.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.md"
-2.  Data have been de-identified and recoded using pre-assigned anonymous IDs.
-3.  Data table headers are in snake_case, and match tabulated list below. The same list is in the file "code_column_names.txt"
-
-Column Number|Column Name
-------------|----------------------
-1|unique.record
-2|report.id
-3|sort
-4|report.title (empty)
-5|student (Std\_nnn)
-6|course
-7|ta (TA\_nnn)
-8|lab
-9|tag
-10|type.TA
-11|grade.TA
-12|grading.time
-13|Rank
-14|hypothesis.ok
-15|data.ok
-16|citation.ok
-17|interpretation.ok
-18|organization.ok
-19|techflaws.ok
-20|writing.ok
-21|comments.incorporated
-22|ta.comment
-23|code.subject
-24|code.structure
-25|code.locus
-26|code.scope
-27|code.tone
-28|code.notes
-
-Once all data are in correct format, export Excel file to CSV with name formatted as **"coded_full_comments_dataset_SemYear.csv"**
-
-####\
-
-##Initial Intake and Post-Check
-```{r Libraries}
-# Call in tidyverse, other required libraries.
-library(tidyverse)
-library(tidytext)
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-library(scales)
-```
-
-```{r}
-#This block reads in the full CSV into a starting dataframe named "base_data".
-#Subsequent analyses call in subsets of "base_data".
-base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
-```
-
-If the TA comments dataset is imported correctly, there will be a dataframe of ~11,000 rows and 28 columns named as listed in the **Data Pre-Checks** table above.
-
-Given some data points may be entered incorrectly by instructors or students, I included several post-import data checks to ensure data have been properly coded and columns contain valid entries only.
-
-```{r}
-#"Course" column should only have entries that correspond to 113, 114, 214, NA. 
-unique(base_data$course)
-```
-
-```{r}
-#Entries in "TAs" column should match the anonymous IDs that correspond to the TAs assigned to the relevant courses in the semester being analyzed. 
-#If there are incorrect or extra anonymous IDs in this list, the mostly likely source of the error is improper de-identification of the Excel file prior to import.
-unique(base_data$ta)
-```
-
-```{r}
-#"Lab" should only have entries that match the limited keywords in codebook.
-unique(base_data$lab)
-```
-
-```{r}
-#Look at the NUMBER of unique anonymous student IDs. Should be within 10% of (but not more than) the combined enrollment of the 3 target courses for the semester.
-unique(base_data$student)
-```
-
-```{r}
-#Check that the list of "code.subject" topics here matches allowed terms in codebook.
-unique(base_data$code.subject)
-```
-
-```{r}
-#Check that the list of "code.structure" terms extracted from the dataset and displayed here matches allowed terms in codebook.
-unique(base_data$code.structure)
-```
-
-####\
-
-## Preliminary Analysis
-The first question prior to deeper comparisons is **are the comments obviously different between groups?** Before conducting detailed analyses I tried several iterations of different visual comparison strategies.
-
-####\
-
-###Log-Odds Text Comparison
-The log-odds ratios method outlined below has been the most informative to date.
-The basic premise being tested is that certain words are more frequent in TA comments belonging to one sub-category, and less frequent in others.
-I first saw this approach as a way to compare word usage between separate Twitter accounts here: https://www.tidytextmining.com/twitter.html.
-
-For simplicity I reduced the analysis to just the four most common comment subjects.
-
-```{r}
-# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
-
-frequency_writing <- filter(base_data, code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
-frequency_writing.subcolumns <- frequency_writing %>% select(1, 22:23)
-```
-
-```{r}
-# Tokenize phrases, remove stop words listed in standard reference file.
-base_data_tokenized <- frequency_writing.subcolumns %>% 
-  unnest_tokens(word,ta.comment) %>% 
-  anti_join(stop_words)
-
-#At this point the dataframe should have "Unique.Record" in column 1, "code.subject" in column 2, and the unnested "word" in column 3.
-All columns are class = "character".
-Stop words listed in the "stop_words" R reference data file have been removed.
-
-```
-
-####Comparing Pearson Correlations of Word Frequencies
-If the words used by TAs in their comments differ between the categories, there should be fairly low correlations overall .
-```{r}
-#Block below re-organizes the data for comparisons. This pattern sets "2. Writing Quality" as the first dataset, then all other data in remaining columns.
-#Groups by code.subject, calculated proportional frequency in each group.
-#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
-#Stats need this format to be able to compare 2. WQ against other values.
-base_data_tokenized_sorted2 <- base_data_tokenized %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  mutate(proportion = n / sum(n)) %>% 
-  select(-n) %>% 
-  spread(code.subject, proportion) %>% 
-  gather(code.subject, proportion, `3. Technical and Scientific`:`4. Logic and Thinking`)
-
-#The dataframe called "base_data_tokenized_sorted2" is the final 4-column dataframe for statistical analysis.
-Individual words are in column 1, "2.
-Writing Quality" frequency is in Column 2, "code.subject" is in column 3, and "proportion" is in column 4.
-Columns 1 and 3 are class = "character", 2 and 4 are numeric ranges.
-
-#This structure is needed in order to compare word (in Column 1) frequencies for 2.
-Writing Quality (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes.
-WHICH subset of X values to use is coded in Column 3.
-```
-
-Calculating Pearson correlation between the frequency of words in Writing Quality versus the other two groups.
-
-```{r}
-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3. Technical and Scientific",], 
-         ~ proportion + `2. Writing Quality`)
-```
-
-```{r}
-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4. Logic and Thinking",], 
-         ~ proportion + `2. Writing Quality`)
-```
-
-####START OF BROKEN SECTION
-THERE IS AN ERROR HERE I CANNOT DEBUG!
-If I change "3.
-Technical and Scientific"  to "2.
-Writing Quality", the structure of the tibble breaks.
-It goes from the correctly sorted 4 columns to incorrectly sorted 3 columns.
-It is NOT A TYPO; I can literally copy/paste it in and out and break/correct the code.
-
-I cannot use a different order, or any other format. I am at a loss.
-
-```{}
-# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
-frequency_writing3 <- filter(base_data, code.subject=="3. Technical and Scientific"|code.subject=="2. Writing Quality"|code.subject=="4. Logic and Thinking")
-frequency_writing3.subcolumns <- frequency_writing3 %>% select(1, 22:23)
-```
-
-```{}
-# Tokenize phrases, remove stop words listed in standard reference file.
-base_data_tokenized3 <- frequency_writing3.subcolumns %>% 
-  unnest_tokens(word,ta.comment) %>% 
-  anti_join(stop_words)
-```
-
-```{}
-#Block below re-organizes the data for comparisons. This pattern sets "3. Writing Quality" as the first dataset, then all other data in remaining columns.
-#Groups by code.subject, calculated proportional frequency in each group.
-#Final step creates 4 columns: word, 2. Writing Quality, code.subject, proportion
-#Stats need this format to be able to compare 2. WQ against other values.
-base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  mutate(proportion = n / sum(n)) %>% 
-  select(-n) %>% 
-  spread(code.subject, proportion) %>% 
-  gather(code.subject, proportion, `2. Writing Quality`:`4. Logic and Thinking`)
-```
-
-```{}
-#Create groups by code.subject, calculates proportional frequency in each group.
-#Final step creates 5 columns: code.subject, n, word, total, frequency.
-base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  left_join(base_data_tokenized %>% 
-              group_by(code.subject) %>% 
-              summarise(total = n())) %>%
-  mutate(freq = n/total)
-```
-
-Re-organize the data for comparisons. This pattern sets 3. Technical and Scientific as the first dataset, then all other data in remaining columns.
-```{}
-#Groups by code.subject, calculated proportional frequency in each group.
-#Final step creates 4 columns: word, 3. Technical and Scientific, code.subject, proportion
-#Stats need this format to be able to compare 3. TS against other values.
-base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  mutate(proportion = n / sum(n)) %>% 
-  select(-n) %>% 
-  spread(code.subject, proportion) %>% 
-  gather(code.subject, proportion, `2. Writing Quality`:`4. Logic and Thinking`)
-
-#END OF BROKEN SECTION
-```
-
-***
-
-START OF SECTION I CANNOT COPY AND MODIFY. This section depends on the ability to calculate tables using the code above. So I cannot run this part either.
-
-```{r}
-#Create groups by code.subject, calculates proportional frequency in each group.
-#Final step creates 5 columns: code.subject, n, word, total, frequency.
-base_data_tokenized_sorted2 <- base_data_tokenized %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  left_join(base_data_tokenized %>% 
-  group_by(code.subject) %>% 
-  summarise(total = n())) %>%
-  mutate(freq = n/total)
-```
-
-```{r}
-base_data_tokenized_sorted2v3 <- base_data_tokenized_sorted2 %>% 
-  select(code.subject, word, freq) %>% 
-  spread(code.subject, freq) %>%
-  arrange(`2. Writing Quality`,`3. Technical and Scientific`)
-```
-
-```{r}
-ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`3. Technical and Scientific`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-  ggsave("Freqplot_3vs2.png", width = 10, height = 6)
-```
-
-```{r}
-#I can get part way around the problem by re-using the same data and calling a different subset for the Y axis.
-Want to add comparisons of 3 vs 4, but same tables cannot be created because dataframe keeps breaking when I try to change 2.
-WQ.
-
-base_data_tokenized_sorted2v4 <- base_data_tokenized_sorted2 %>% 
-  select(code.subject, word, freq) %>% 
-  spread(code.subject, freq) %>%
-  arrange(`2. Writing Quality`,`4. Logic and Thinking`)
-```
-
-```{r}
-ggplot(base_data_tokenized_sorted2v3, aes(`2. Writing Quality`,`4. Logic and Thinking`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-  ggsave("Freqplot_4vs2.png", width = 10, height = 6)
-```
-
-END of broken / limping section
-
-####Calculating and Plotting Log-Odds Pairs
-**2. Writing versus 3. Technical**
-
-```{r}
-word_ratios2v3 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`2. Writing Quality`/`3. Technical and Scientific`)) %>%
-  arrange(desc(logratio))
-```
-
-```{r}
-word_ratios2v3 %>% 
-  arrange(abs(logratio))
-```
-
-```{r}
-word_ratios2v3 %>%
-  group_by(logratio < 0) %>%
-  top_n(20, abs(logratio)) %>%
-  ungroup() %>%
-  mutate(word = reorder(word, logratio)) %>%
-  ggplot(aes(word, logratio, fill = logratio < 0)) +
-  geom_col(show.legend = FALSE) +
-  coord_flip() +
-  ylab("log odds ratio (Writing-top/Technical-bottom)") +
-  scale_fill_discrete(name = "", labels = c("2. Writing Quality", "3. Technical and Scientific"))
-  #ggsave("Logplot_2vs3.png", width = 6, height = 10)
-
-```
-
-**3. Technical vs. 4. Logic**
-
-```{r}
-word_ratios3v4 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`3. Technical and Scientific`/`4. Logic and Thinking`)) %>%
-  arrange(desc(logratio))
-```
-
-```{r}
-word_ratios3v4 %>% 
-  arrange(abs(logratio))
-```
-
-```{r}
-word_ratios3v4 %>%
-  group_by(logratio < 0) %>%
-  top_n(20, abs(logratio)) %>%
-  ungroup() %>%
-  mutate(word = reorder(word, logratio)) %>%
-  ggplot(aes(word, logratio, fill = logratio < 0)) +
-  geom_col(show.legend = FALSE) +
-  coord_flip() +
-  ylab("log odds ratio (Technical-top/Logic-bottom)") +
-  scale_fill_discrete(name = "", labels = c("3. Technical and Scientific","4. Logic and Thinking"))
-  #ggsave("Logplot_3vs4.png", width = 6, height = 10)
-
-```
-
-**2. Writing versus 4. Logic**
-
-```{r}
-word_ratios2v4 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`2. Writing Quality`/`4. Logic and Thinking`)) %>%
-  arrange(desc(logratio))
-```
-
-```{r}
-word_ratios2v4 %>% 
-  arrange(abs(logratio))
-```
-
-```{r}
-word_ratios2v4 %>%
-  group_by(logratio < 0) %>%
-  top_n(20, abs(logratio)) %>%
-  ungroup() %>%
-  mutate(word = reorder(word, logratio)) %>%
-  ggplot(aes(word, logratio, fill = logratio < 0)) +
-  geom_col(show.legend = FALSE) +
-  coord_flip() +
-  ylab("log odds ratio (Writing-top/Logic-bottom)") +
-  scale_fill_discrete(name = "", labels = c("2. Writing Quality", "4. Logic and Thinking"))
-  #ggsave("Logplot_2vs4.png", width = 6, height = 10)
-
-```
-
-##Direct Comparison Using Only Words Appearing in All Subsets of Comments
-I think this is less informative, but it does let me see the data more clearly.Basic strategy is to take all comments from base_data, unnest tokens, filter stopwords, count and sort remaining terms, then write back to a new datatable called 'sortedwords.comments.all'.
-Calculate TOTAL words and store.
-Then calculate fractional frequencies for each word and store using mutate.
-
-This is an early version, so likely has some poorly developed code blocks that could be refactored.
-
-```{r}
-library(dplyr)
-library(tidytext)
-
-sortedwords.comments.all <- base_data_tokenized %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate the number of words in total in the subset
-total_words.all <- sortedwords.comments.all %>% 
-  summarize(total = sum(n))
-
-#Mutate table to calculate then append fractional word frequency in the text.
-sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
-  all.fraction = (n*100)/total_words.all$total
-  )
-```
-
-##Summarize Word Frequencies in Subsets of Comments Based on 'Subject'
-Take subsets of comments from base_data_tokenized using code.subject as separating characteristic.
-
-###Other Analysis Approaches
-This approach summarizes the word frequencies in the FULL set of TA comments first, then breaks them down into sub-groups. It has examples of other code bits as well.
-
-The first block takes all comments from base_data, unnests tokens, filters stopwords, count and sort remaining terms, then writes results back to a new dataframe called 'sortedwords.comments.all'.
-
-TOTAL words is calculated and stored. Then fractional frequencies for each word are calculated and stored back to the dataframe using mutate.
-
-```{r}
-# Tokenize phrases, remove stop words listed in standard reference file.
-base_data_tokenized <- base_data %>% 
-  unnest_tokens(word,ta.comment)
-
-# An alternative approach to pulling out stopwords  
-sortedwords.comments.all <- base_data_tokenized %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate the number of words in total in the subset
-total_words.all <- sortedwords.comments.all %>% 
-  summarize(total = sum(n))
-
-#Mutate table to calculate then append fractional word frequency in the text.
-sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
-  all.fraction = (n*100)/total_words.all$total
-  )
-```
-
-```{r}
-#Pull SUBSETs of tokenized data based on code.subject: 
-base_data_tokenized.basic <- subset(base_data_tokenized, code.subject == "1. Basic Criteria", select = (1:28))
-base_data_tokenized.writing <- subset(base_data_tokenized, code.subject == "2. Writing Quality", select = (1:28))
-base_data_tokenized.technical <- subset(base_data_tokenized, code.subject == "3. Technical and Scientific", select = (1:28))
-base_data_tokenized.logic <- subset(base_data_tokenized, code.subject == "4. Logic and Thinking", select = (1:28))
-base_data_tokenized.praise <- subset(base_data_tokenized, code.subject == "5. Praise or Concern", select = (1:28))
-base_data_tokenized.misconduct <- subset(base_data_tokenized, code.subject == "6. Misconduct", select = (1:28))
-base_data_tokenized.narrative <- subset(base_data_tokenized, code.subject == "12. Narrative Comments", select = (1:28))
-base_data_tokenized.sciname <- subset(base_data_tokenized, code.subject == "Scientific Name", select = (1:28))
-```
-
-For each subset, filter stopwords, count words and total, then calculate frequencies.
-
-```{r}
-sortedwords.basic <- base_data_tokenized.basic %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate the number of words in total in the subset
-total_words.basic <- sortedwords.basic %>% 
-  summarize(total = sum(n))
-
-#Mutate table to calculate then append fractional word frequency in the text.
-sortedwords.basic.fraction <- sortedwords.basic %>% as_tibble() %>% mutate(
-  basic.fraction = (n*100)/total_words.basic$total
-)
-```
-
-```{r}
-sortedwords.writing <- base_data_tokenized.writing %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate the number of words in total in the subset
-total_words.writing <- sortedwords.writing %>% 
-  summarize(total = sum(n))
-
-#Mutate table to calculate then append fractional word frequency in the text.
-sortedwords.writing.fraction <- sortedwords.writing %>% as_tibble() %>% mutate(
-  writing.fraction = (n*100)/total_words.writing$total
-)
-```
-
-```{r}
-sortedwords.technical <- base_data_tokenized.technical %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate the number of words in total in the subset
-total_words.technical <- sortedwords.technical %>% 
-  summarize(total = sum(n))
-
-#Mutate table to calculate then append fractional word frequency in the text.
-sortedwords.technical.fraction <- sortedwords.technical %>% as_tibble() %>% mutate(
-  technical.fraction = (n*100)/total_words.technical$total
-)
-```
-
-```{r}
-sortedwords.logic <- base_data_tokenized.logic %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate the number of words in total in the subset
-total_words.logic <- sortedwords.logic %>% 
-  summarize(total = sum(n))
-
-#Mutate table to calculate then append fractional word frequency in the text.
-sortedwords.logic.fraction <- sortedwords.logic %>% as_tibble() %>% mutate(
-  logic.fraction = (n*100)/total_words.logic$total
-)
-```
-
-Next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed.
-
-**WARNING**: Words that are MISSING from **any subset** cause the full set to be pruned. I am losing the unique words this way. **Need to figure out how to join without pruning.**
-
-```{r}
-total <- merge(sortedwords.comments.all.fraction,sortedwords.technical.fraction,by="word") 
-total2 <- merge(total,sortedwords.writing.fraction,by="word")
-total3 <- merge(total2,sortedwords.logic.fraction,by="word")
-#rename lines that were possibly confusing
-names(total3)[2] <- "all.count"
-names(total3)[4] <- "technical.count"
-names(total3)[6] <- "writing.count"
-names(total3)[8] <- "logic.count"
-
-#Mutate table to calculate then append fractional word frequency in the text.
-total4 <- total3 %>% as_tibble() %>% mutate(
-  writing.wt = (writing.count *100)/all.count
-)
-
-total5 <- total4 %>% as_tibble() %>% mutate(
-  technical.wt = (technical.count *100)/all.count
-)
-
-total6 <- total5 %>% as_tibble() %>% mutate(
-  logic.wt = (logic.count *100)/all.count
-)
-
-```
-
-The following six graphs only compare frequencies for ~770 words that appear in ALL THREE of the sub-categories.
-Graphs 1-3 compare frequencies in individual groups against word frequency overall.
-Generally, the frequency of a term in the whole set of comments is similar to frequency in an individual group (X vs Y positions similar.) However, some terms associated with Logic subgroup clearly fall below the trendline Graph 3, suggesting they are less frequent in that sub-category than overall.
-
-```{r}
-ggplot(total6, aes(`all.fraction`,`writing.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-```
-
-```{r}
-ggplot(total6, aes(`all.fraction`,`technical.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-```
-
-```{r}
-ggplot(total6, aes(`all.fraction`,`logic.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-```
-
-Graphs 4-6 are pair-wise comparisons of frequencies for sub-categories.
-Unlike the comparisons to overall frequency, most words are NOT clustered on the trendline, suggesting there are differences in word usage.
-
-```{r}
-ggplot(total6, aes(`technical.fraction`,`writing.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-```
-
-```{r}
-ggplot(total6, aes(`logic.fraction`,`writing.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-```
-
-```{r}
-ggplot(total6, aes(`logic.fraction`,`technical.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-```
-
-####\
-
-## Data
-
-####\
-
-## Analysis
-
-####\
-
-# Results
-
-####\
-
-# Discussion
-
-####\
-
-# Conclusion
-
-####\
-
-# References
-
-####\

### 2019-02-21 findings.Rmd
cleanup

--- 2018-12-16 project.Rmd
@@
 ---
-title: "Project"
+title: "Project Findings"
 author: "Dan Johnson"

### 2019-02-21 hiding/Ensemble_Notes.Rmd
cleanup

---
title: "Notes on Building a Final Ensemble Protocol"
author: "Dan Johnson"
date: "2/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Overview
This notebook summarizes my progress, thoughts while building a working Ensemble method for categorizing TA comments.

####\

##Replicate Cycles Test of Naive Bayes
One of my ideas was to see if I could refine Naive Bayes by running the same classifier with different randomized datasets, then putting individual outputs side by side and assigning final categories based on most frequent classification in the set.

To test this I ran 5 Subject classifier replicates using the code in *Naive_bayes_V2r_2b.rmd*, with 5 different ```set.seed``` values. The exported CSV files were combined to get a set of data with up to 5 classifier assignments for the same comment.

Using multiple randomized NB training runs does not seem to help. First, it was rare for the  randomized runs to assign different incorrect classes. Second, when I compared 5 independently randomized runs by eye, the incorrectly categorized items were repeatedly put into the SAME incorrect group each time. Based on this, an Ensemble method using H-stacked categorical NB is unlikely to reduce overall errors in classification.

Looking at the combined data DID provide other helpful insights into the overall classification procedure though. Naive Bayes does not seem to do well with comments that:

*  Contain many or mostly punctuation symbols, such as:
    +  : not ,
    +  ?
    +  ??
    +  ???
*  Are whole numbers
    +  1
    +  2
    +  10
*  Are one word questions or comments
    +  because?
    +  "and"
*	 Are sentences enclosed in punctuation
*  Are made up entirely of STOP/common words (I am not as sure on this one)

####\

##Overall Accuracy
I also looked at OVERALL accuracy of NB by recreating the 2-dimensional contingency table that I made when hand-coding, and comparing it side-by-side to NB classifications. From this I gained additional insights.

OVERALL, the 2-D Naive Bayes model performed very well for classifying comment Subject. When I re-ran the full original dataset (9340 comments) through the prediction model, NB estimates for Subject were within 1.5% of my hand-scored assignments. Structure classification was not as robust. Other observations:

*  NB over-classified comments as "copyediting of technical issues" by 8-10%, and under-classified comments that should be marked as "specific technical issues" by almost same amount.
*  Comments that belonged to groups making up less than 1-2% of all comments were more likely to be mis-classified. 
*  Longer text comments seemed to be mis-classified more often. Follow-up methods to evaluate: 
    +  Split longer comments into shorter phrases
    +  Evaluate only the first 5-10 words of a comment

It appears that Subject categories are more terms-centered, while Structure is based more on the sentiment and sentence structure of a comment. This suggests that I need to incorporate a sentiment classifier or PoS tagger step.

####\

##Mixed Ensemble Method
Based on the  seems like the most reasonable approach for maximum accuracy is going to be a mixed Ensemble method. The initial working model is to:

Add a pre-screening step prior to implementing NB:

*  Isolate "short-string" comments first. These are comments consisting of single words, two-word phrases, more punctuation than text, etc. These represent mostly pointers and copy-editing.
*  Look for 2-grams with high frequencies in one particular category. (Ex., "basic criteria" 2-gram identifies a Basic Criteria comment.)

####\

### 2019-02-21 hiding/Multinomial_bayes_V1b.Rmd
cleanup

--- 2019-01-31 Naive_bayes_V2r.Rmd
@@
 ---
-title: "Naive Bayes Workflow Ver 2.0.r"
+title: "Multinomial Naive Bayes Workflow Ver 3.0.b."
 author: "Dan Johnson"
-date: "1/29/2019"
+date: "2/13/2019"
 output: html_document
@@
 ##Summary
-This is **Version 2.0.r** of the basic workflow for classifying TA comments using a naive Bayes model.
-It pre-processes text using the same command set as Ver 2.0.b.
-The "r" designates this version has been modified to be able to randomize which rows are selected for training and test datasets.
+This is an alternative version of the basic NB workflow for classifying TA comments.
+The code is slightly different than what was used for Categorical NB testing, but takes in text using the same command set as NB-Ver 2.0.
+This alternative method lets me perform true multinomial analysis.
+I can switch between using binary yes/no categorization of words' presence or absence, and the actual frequencies.
+There is a trade-off though; when using non-categorical values, I am limited to two bins: in-group versus out-group.
+If I use multiple groups, the data fail to parse correctly.
+If I want to look at multiple bins for classification, I have to use categorical values.
+
+Overall categorical NB seems to perform better than true multinomial NB.
+I get 75-84% overall accuracy (in vs out) if I use yes/no categories for both Subject and Structure, but accuracy drops to ~65% when I use the raw numeric frequencies.
+Reducing the Laplace value does not affect overall performance.
 
@@
 
-```{}
-library(tidyverse)
-library(tidytext)
-library(tidyr)
+```{r}
+
 library(tm)
-library(SnowballC)
-library(wordcloud)
+library(RTextTools)
 library(e1071)
-library(gmodels)
+library(dplyr)
+library(caret)
+library(doMC)
+registerDoMC(core=detectCores())
+
+#library(tidyverse)
+#library(tidytext)
+#library(tidyr)
+#library(SnowballC)
+#library(wordcloud)
+#library(gmodels)
 ```
@@
 #Read in TA comments from CSV file.
-base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
+subset_M <- read.csv("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/comments_binary.csv", stringsAsFactors=FALSE)
+glimpse(subset_M)
 ```
+```{r}
+set.seed(1)
+subset_M <- subset_M[sample(nrow(subset_M)), ]
+subset_M <- subset_M[sample(nrow(subset_M)), ]
+glimpse(subset_M)
+```
+
+```{r}
+subset_M$subject <- as.factor(subset_M$subject)
+subset_M$structure <- as.factor(subset_M$structure)
+```
+
+```{r}
+corpus <- Corpus(VectorSource(subset_M$text))
+corpus
+inspect(corpus[1:3])
+```
+
+```{r}
+corpus.clean <- corpus %>%
+  tm_map(content_transformer(tolower)) %>% 
+  tm_map(removePunctuation) %>%
+  tm_map(removeNumbers) %>%
+  tm_map(removeWords, stopwords(kind="en")) %>%
+  tm_map(stripWhitespace)
+```
+
+```{r}
+dtm <- DocumentTermMatrix(corpus.clean)
+# Inspect the dtm
+inspect(dtm[40:50, 10:15])
+```
+
+```{r}
+subset_M.train <- subset_M[1:7005,]
+subset_M.test <- subset_M[7006:9340,]
+
+dtm.train <- dtm[1:7005,]
+dtm.test <- dtm[7006:9340,]
+
+corpus.clean.train <- corpus.clean[1:7005]
+corpus.clean.test <- corpus.clean[7006:9340]
+```
+
+```{r}
+dim(dtm.train)
+```
+
+```{r}
+fivefreq<- findFreqTerms(dtm.train, 5)
+length((fivefreq))
+
+dtm.train.nb <-DocumentTermMatrix(corpus.clean.train, control=list(dictionary=fivefreq))
+
+dim(dtm.train.nb)
+
+dtm.test.nb <- DocumentTermMatrix(corpus.clean.test, control=list(dictionary = fivefreq))
+
+dim(dtm.train.nb)
+```
+
+We use a variation of the multinomial Naive Bayes algorithm known as binarized (boolean feature) Naive Bayes due to Dan Jurafsky.
+In this method, the term frequencies are replaced by Boolean presence/absence features.
+The logic behind this being that for sentiment classification, word occurrence matters more than word frequency.
+```{r}
+# Function to convert the word frequencies to yes (presence) and no (absence) labels
+convert_count <- function(x) {
+  y <- ifelse(x > 0, 1,0)
+  y <- factor(y, levels=c(0,1), labels=c("No", "Yes"))
+  y
+}
+```
+
+```{r}
+# Function to convert the word frequencies to yes (presence) and no (absence) labels
+convert_count <- function(x) {
+  y <- x
+}
+```
+
+```{r}
+# Apply the convert_count function to get final training and testing DTMs
+trainNB <- apply(dtm.train.nb, 2, convert_count)
+testNB <- apply(dtm.test.nb, 2, convert_count)
+```
+
+```{r}
+# Train the classifier
+system.time( classifier <- naiveBayes(trainNB, subset_M.train$subject, laplace = 1) )
+system.time( classifier2 <- naiveBayes(trainNB, subset_M.train$structure, laplace = 1) )
+```
+
+Testing the Predictions
+```{r}
+# Use the NB classifier we built to make predictions on the test set.
+system.time( pred <- predict(classifier, newdata=testNB) )
+system.time( pred2 <- predict(classifier2, newdata=testNB) )
+```
+
+```{r}
+# Create a truth table by tabulating the predicted class labels with the actual class labels 
+table("Predictions"= pred,  "Actual" = subset_M.test$subject )
+table("Predictions2"= pred2,  "Actual" = subset_M.test$structure )
+```
+
+```{r}
+# Prepare the confusion matrix
+conf.mat <- confusionMatrix(pred, subset_M.test$subject)
+conf.mat
+print ("   ")
+
+conf.mat$byClass
+print ("   ")
+
+conf.mat$overall
+print ("   ")
+
+conf.mat$overall['Accuracy']
+
+# Prepare the confusion matrix
+conf.mat2 <- confusionMatrix(pred2, subset_M.test$structure)
+conf.mat2
+print ("   ")
+
+conf.mat2$byClass
+print ("   ")
+
+conf.mat2$overall
+print ("   ")
+
+conf.mat2$overall['Accuracy']
+
+```
+
+https://rpubs.com/cen0te/naivebayes-sentimentpolarity
 
@@
 Extract comments and classification data subset
-```{r}
+```{}
 #Select rows representing the sub-groups to compare. 
@@
 
-```{r}
+```{}
 #Change "structure" element from character to a factor for analysis.
@@
 
-```{r}
+```{}
 #Generate a vector of ~10,000 numbers (be sure it can be generated consistently)
@@
 
-```{r}
+```{}
 #Append the vector as a new column to "comments_rawVERB" with name "randomizer"
@@
 Convert data set to a volative corpus using "tm" library.
-```{r}
+```{}
 #Create the volatile corpus that contains the "text" vector from data frame.
@@
 
-```{r}
+```{}
 #Convert text to all lower case letters.
@@
 
-```{r}
+```{}
 NLP_Tokenizer <- function(x) {
-      unlist(lapply(ngrams(words(x), 3), paste, collapse = " "), use.names = FALSE)
+      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
 }
@@
 This code uses the comments in the order they are listed in the original file. Block for randomizing them was done earlier.
-```{r}
+```{}
 comments_dtm_train <- comments_dtm_1gram[1:7005, ]
@@
 Save vectors labeling rows in the training and testing vectors
-```{r}
+```{}
 comments_train_labels_subject <- comments_rawVERB_R[1:7005, ]$subject
@@
 
-```{r}
+```{}
 comments_train_labels_structure <- comments_rawVERB_R[1:7005, ]$structure
@@
 Make sure that the proportion of each sub-category is similar in the training and testing data sets.
-```{r}
+```{}
 prop.table(table(comments_train_labels_subject))
@@
 
-```{r}
+```{}
 prop.table(table(comments_train_labels_structure))
@@
 Remove words from the matrix that appear less than 5 times.
-```{r}
+```{}
 comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
@@
 Limit Document Term Matrix to only include words in the comments_freq_words vector. Using all the rows, but we want to limit the columns to these words in the frequency vector.
-```{r}
+```{}
 comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
@@
 This replaces values greater than 0 with yes, and values not greater than 0 with no.
-```{r}
+```{}
 convert_counts2 <- function(x) {
@@
 It evaluates the prediction with the actual data using a crosstable from the gmodels package.
-```{r}
+```{}
 comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject)
@@
 The cross table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject.
-```{r}
+```{}
 print("Overall subject accuracy")
-(14+54+1318+28)/23.35  #Need to find a way to calculate this automatically. 
+(43+473+1131+184)/23.35  #Need to find a way to calculate this automatically. 
 ```
@@
 
-```{r}
+```{}
 comments_classifier_structure <- naiveBayes(comments_train, comments_train_labels_structure)
@@
 
-```{r}
+```{}
 print("Overall structure accuracy")
-(66+64+949+5+1)/23.35  #Need to find a way to calculate this automatically. 
+(22+488+292+518+33+10)/23.35  #Need to find a way to calculate this automatically. 
 ```
@@
 
-##Permutation Tests of Naive Bayes (NB) Method
-Using single words, Version 2.0.b correctly identifies comment "subject" ~80% of the time and comment "structure" ~59% of the time.
-
-Permutations to test with random subsets version (2.0.r):
-
-*  Compare 1-grams, 2-grams, 3-grams
-*  Compare outcomes for 3 independent randomized sets of 1-grams
-
-The pseudo-code block below can pull random lines from a data table, but does not work correctly on a document term matrix.
-The second problem is how to pull the training and test data sets out and process them identically.
-
-```{}
-set.seed(1234) 		# ensures the same random set gets picked each time.
-
-comments_raw_dtm_train <- sample_frac(comments_raw, .75)	# sample_frac pulls out 75% of "comments_raw" dataset randomly using the sample sequence identified by set.seed(1234)
-
-comments_raw_dtm_test <- setdiff(comments_raw, comments_raw_dtm_train)	#setdiff creates a "test" dataset using anything NOT in the "comments_raw_dtm_train" subset of "comments_raw."
-```
-
-The solution was not to pull the data twice, but to randomly (but repeatably) shuffle the starting data table. How I solved this was:
-1. Create a vector of sequential numbers the same length as the number of data frame rows.
-2. Randomize the sequential numbers using set.seed to make pattern reproducible.
-3. Add the randomized numbers as a new column to the data frame.
-4. Re-order the data table to the randomizer column.
-
-This makes it easy to generate a new sequence of randomized training and test samples to run replicates. Only one number needs to be changed.
-
 ```{}

### 2019-02-21 hiding/Naive_bayesV2.Rmd
cleanup

--- 2019-02-05 Backups/Naive_bayes.Rmd
@@
 
+##PROBLEM
+The set.seed command can pull random lines from a data table, but does not work on a document term matrix.
+
+```{r}
+set.seed(1234) 		# ensures the same random set gets picked each time.
+
+comments_raw_dtm_train <- sample_frac(comments_raw, .75)	# sample_frac pulls out 75% of "comments_raw" dataset randomly using the sample sequence identified by set.seed(1234)
+
+comments_raw_dtm_test <- setdiff(comments_raw, comments_raw_dtm_train)	#setdiff creates a "test" dataset using anything NOT in the "comments_raw_dtm_train" subset of "comments_raw."
+```
+
 Data set must be converted to a volative corpus using "tm" library then transformed.
@@
 #Create the volatile coprus that contains the "text" vector from data frame.
-comments_corpus <- VCorpus(VectorSource(comments_raw$text))
-print(comments_corpus)
+comments_corpus_train <- VCorpus(VectorSource(comments_raw_dtm_train$text))
+print(comments_corpus_train)
 
 #Check out the first few messages in the new corpus, which is basically a list that can be manipulated with list operations.
-inspect(comments_corpus[1:3])
+inspect(comments_corpus_train[1:3])
 
 #Use "as.character" function to see what a message looks like.
-as.character(comments_corpus[[3]])
-```
-
-The OPTIONAL data transforms using "tm_map"" and "content_transformer" functions. Switch between "{r}" and "{}" to turn a transformation block off or on.
-```{r}
-#Convert to all lower case letters.
-comments_corpus_clean <- tm_map(comments_corpus, content_transformer(tolower))
+as.character(comments_corpus_train[[3]])
 ```
@@
 ```{r}
+#Create the volatile coprus that contains the "text" vector from data frame.
+comments_corpus_test <- VCorpus(VectorSource(comments_raw_dtm_test$text))
+print(comments_corpus_test)
+
+#Check out the first few messages in the new corpus, which is basically a list that can be manipulated with list operations.
+inspect(comments_corpus_test[1:3])
+
+#Use "as.character" function to see what a message looks like.
+as.character(comments_corpus_test[[3]])
+```
+
+***
+***
+The OPTIONAL data transforms start here. They use "tm_map"" and "content_transformer" functions. Switch between "{r}" and "{}" to turn a transformation block off or on.
+```{r}
+#Convert to all lower case letters.
+#Pretty standard conversion.
+comments_corpus_clean_train <- tm_map(comments_corpus_train, content_transformer(tolower))
+comments_corpus_clean_test <- tm_map(comments_corpus_test, content_transformer(tolower))
+```
+
+```{r}
+#Remove punctuation using the "removePunctuation" function. 
+#This step removes evidence of questions, so may remove data.
+comments_corpus_clean_train <- tm_map(comments_corpus_clean_train, removePunctuation)
+comments_corpus_clean_test <- tm_map(comments_corpus_clean_test, removePunctuation)
+```
+
+```{}
 #Remove numerals. 
@@
 
-```{r}
+```{}
 #Stopword removal. 
+#Can be refined by altering what reference file is used.
 comments_corpus_clean <- tm_map(comments_corpus_clean, removeWords, stopwords())
@@
 
-```{r}
-#Remove punctuation using the "removePunctuation" function. 
-#This step removes evidence of questions, so may remove data.
-comments_corpus_clean <- tm_map(comments_corpus_clean, removePunctuation)
-```
-
-```{r}
+```{}
 #Stemming the text data to strip the suffix from words. 
 #My thinking is that stemming removes valuable tense data.I am unsure how valuable it is.
+#Alternative is using part-of-speech tagging and lemmatization instead.
 
@@
 
-Final prep and data check.
+Final prep and data check for training set.
 ```{r}
 #Final step removes white space from the document. This is NOT an optional step.
-comments_corpus_clean <- tm_map(comments_corpus_clean, stripWhitespace)
+comments_corpus_clean_train <- tm_map(comments_corpus_clean_train, stripWhitespace)
 
 #Look at an example of data.
-as.character((comments_corpus_clean[[3]]))
+as.character((comments_corpus_clean_train[[3]]))
 ```
 
+Final prep and data check for testing set.
+```{r}
+#Final step removes white space from the document. This is NOT an optional step.
+comments_corpus_clean_test <- tm_map(comments_corpus_clean_test, stripWhitespace)
+
+#Look at an example of data.
+as.character((comments_corpus_clean_test[[3]]))
+```
 ####\
 
-###Basic Tokenizer
-This version uses the "DocumentTermMatrix" function to creates a matrix in which the rows indicate individual documents (individual TA comments) and the columns indicate presence or absence of the word that is the header for the column.
-Be careful not to confuse it with "TermDocumentMatrix", which transposes the matrix.
-
-"DocumentTermMaxtrix" function also can perform all of the text prep above in one command.
+##Modified Naive Bayes (NB) Method Using N-grams
+The tokenizer command block below uses tm's NLP commands to generate either 1-, 2-, or 3-grams with the same command. The only change needed is "#" in "(words(x), #)" below.
 
 ```{r}
-comments_corpus_1gram <- DocumentTermMatrix(comments_corpus_clean)
-comments_corpus_1gram_tidy <- tidy(comments_corpus_1gram)
-inspect(comments_corpus_1gram)
+NLP_Tokenizer <- function(x) {
+      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
+}
+
+comments_dtm_ngram_train <- DocumentTermMatrix(comments_corpus_clean_train, control=list(tokenize = NLP_Tokenizer))
+comments_dtm_ngram_train_tidy <- tidy(comments_dtm_ngram_train)
+inspect(comments_dtm_ngram_train)
+```
+
+```{r}
+NLP_Tokenizer <- function(x) {
+      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
+}
+
+comments_dtm_ngram_test <- DocumentTermMatrix(comments_corpus_clean_test, control=list(tokenize = NLP_Tokenizer))
+comments_dtm_ngram_test_tidy <- tidy(comments_dtm_ngram_test)
+inspect(comments_dtm_ngram_test)
 ```
@@
 ###Data Preparation
-Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data. 
+
 ```{r}
-.75 * 9340 #Number of documents/rows in the dataset; product is #rows for training
-.25 * 9340 #Product is #rows for testing set
-```
-
-This code assumes comments are random. **Probably want to try randomizing them.**
-```{r}
-comments_dtm_train <- comments_corpus_1gram[1:7005, ]
-comments_dtm_test <- comments_corpus_1gram[7006:9340, ]
+#comments_dtm_train <- comments_dtm_ngram_train
+#comments_dtm_test <- comments_dtm_ngram_test
+print("Train")
+glimpse(comments_dtm_ngram_train)
+print("Test")
+glimpse(comments_dtm_ngram_test)
 ```
@@
 ```{r}
-comments_train_labels <- comments_raw[1:7005, ]$type
-comments_test_labels <- comments_raw[7006:9340,]$type
+comments_train_labels <- comments_raw_dtm_train$type
+comments_test_labels <- comments_raw_dtm_test$type
+print("Train")
+glimpse(comments_train_labels)
+print("Test")
+glimpse(comments_test_labels)
 ```
@@
 ###Preparation for Naive Bayes
-Remove words from the matrix that appear less than 5 times.
+Remove words from the training matrix that appear less than 5 times.
 ```{r}
-comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
-str(comments_freq_words)
+comments_freq_words_train <- findFreqTerms(comments_dtm_ngram_train, 5)
+str(comments_freq_words_train)
+```
+
+Remove words from the test matrix that appear less than 5 times.
+```{r}
+comments_freq_words_test <- findFreqTerms(comments_dtm_ngram_test, 5)
+str(comments_freq_words_test)
 ```
@@
 ```{r}
-comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
-comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
+comments_dtm_freq_train <- comments_dtm_ngram_train[ , comments_freq_words_train]
+comments_dtm_freq_test <- comments_dtm_ngram_test[ , comments_freq_words_test]
 ```
@@
 comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
+print("Train")
+glimpse(comments_train)
+print("Test")
+glimpse(comments_test)
 ```
@@
 Evaluate the prediction with the actual data using a crosstable from the gmodels package.
-```{r}
-comments_classifier <- naiveBayes(comments_train, comments_train_labels)
-comments_test_pred <- predict(comments_classifier, comments_test)
-CrossTable(comments_test_pred, comments_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
-```
-
-Calculate overall correct prediction rate.
-```{r}
-(42+449+1185+186)/23.35
-```
 
 ***
-##Modified Naive Bayes (NB) Method Using N-grams
-The tokenizer command block above only works when using single words.
-The block below is an alternative method that uses tm's NLP commands to generate either 1-, 2-, or 3-grams with the same command.
-The only change needed is "#" in "(words(x), #)" below.
-For simplicity I left off all of the preparatory code up leading up to the final cleaned corpus.
+Here is where it breaks. It generates the comparison data (Step 1) and I can glimpse all the inputs and first output (see prior blocks). I just cannot run the prediction builder step. I keep getting:
 
-```{r}
-NLP_Tokenizer <- function(x) {
-      unlist(lapply(ngrams(words(x), 2), paste, collapse = " "), use.names = FALSE)
-}
-
-comments_dtm_2gram <- DocumentTermMatrix(comments_corpus_clean, control=list(tokenize = NLP_Tokenizer))
-comments_dtm_2gram_tidy <- tidy(comments_dtm_2gram)
-inspect(comments_dtm_2gram)
 ```
-
-####\
-
-###Data Preparation
-Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data.
-
-```{r}
-.75 * 9340 #Number of rows in the dataset; product is #rows for training
-.25 * 9340 #Product is #rows for testing set
-```
-
-This code assumes comments are random. **Probably want to try randomizing them.**
-```{r}
-comments_dtm_train <- comments_dtm_2gram[1:7005, ]
-comments_dtm_test <- comments_dtm_2gram[7006:9340, ]
-```
-
-Save vectors labeling rows in the training and testing vectors
-```{r}
-comments_train_labels <- comments_raw[1:7005, ]$type
-comments_test_labels <- comments_raw[7006:9340,]$type
-```
-
-Make sure that the proportion of each sub-category is similar in the training and testing data set.
-```{r}
-prop.table(table(comments_train_labels))
-prop.table(table(comments_test_labels))
-```
-
-####\
-
-###Preparation for Naive Bayes
-Remove words from the matrix that appear less than 5 times.
-```{r}
-comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
-str(comments_freq_words)
-```
-
-Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
-```{r}
-comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
-comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
-```
-
-The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables.
-To do this we'll build a convert_counts function and apply it to our data.
-```{r}
-convert_counts2 <- function(x) {
-  x <- ifelse(x > 0, "Yes", "No")
-}
-```
-
-This replaces values greater than 0 with yes, and values not greater than 0 with no. Let's apply it to our data.
-```{r}
-comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
-comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
-```
-
-The resulting matrixes will be character type, with cells indicating "yes" or "no" if the word represented by the column appears in the message represented by the row.
-
-####\
-
-###Train Model, Predict, Evaluate
-Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group ONE, TWO, THREE, or FOUR.
-Evaluate the prediction with the actual data using a crosstable from the gmodels package.
-```{r}
-comments_classifier <- naiveBayes(comments_train, comments_train_labels)
-comments_test_pred <- predict(comments_classifier, comments_test)
-CrossTable(comments_test_pred, comments_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
+Error in `[[<-.data.frame`(`*tmp*`, i, value = integer(0)) : 
+  replacement has 0 rows, data has 1946
 ```
@@
 ```{r}
-(34+214+1254+94)/23.35
+comments_classifier <- naiveBayes(comments_train, comments_train_labels)
+
+comments_test_pred <- predict(comments_classifier, comments_test)
+
+CrossTable(comments_test_pred, comments_train_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
 ```

### 2019-02-21 hiding/Naive_bayes_V2b.Rmd
cleanup

--- 2019-02-05 Backups/Naive_bayes_V2a.Rmd
@@
 ---
-title: "Naive Bayes Workflow Ver 2.0"
+title: "Naive Bayes Workflow Ver 2.0.b"
 author: "Dan Johnson"
@@
 ##Summary
-This notebook documents a basic workflow for applying a naive Bayes model to terms in TA comments.
-I based it on a tutorial [https://rpubs.com/Billyhansen6/318412](located here).
-This is quoted from Hansen's original tutorial:
->The objective of this project is to classify SMS messages as spam or ham (not spam).
+This is **Version 2.0.b** of the basic workflow for classifying TA comments using a naive Bayes model.
+It is optimized for text cleanup.
+It does not randomize the rows selected or make pair-wise assignments.
+
+The protocol is based on a published tutorial [https://rpubs.com/Billyhansen6/318412](located here).
+
+This quote from Hansen's original tutorial nicely explains the challenge of binary classification:
+>The objective of our sample project is to classify SMS messages as spam or ham (not spam).
 A Naive Bayes classifier approach will be used.
@@
 
-Lines ~30-205 are the baseline version of the workflow using single words.
-It correctly classifies comments in the top 4 sub-categories  ~79% of the time.
-While that is close to minimum acceptable inter-rater reliability of two human readers, it is well below my goal of >95% for use in routine automated analysis.
+***
 
-Lines ~208-290 describe an alternative method using n-grams. When I tested this approach using 2-grams, classification was slightly worse, at 68.3% correct.
-
-Lines ~295-335 show how to use word clouds to visualize single word frequencies. I did not find this approach very useful.
-
-To keep this code clean for now I will run various permutations in a scratch version, then summarize findings starting on line #345. In the subsequent iterations I will evaluate:
+The main features and permutations to be evaluated are:
 
 *  Different text pre-processing strategies: stemmed versus unstemmed, numbers removed vs. not, etc.
-*  Using n-grams rather than single words
-*  Making pair-wise comparison versus comparing multiple groups simultaneously
-*  Modifying the variables in the NB analysis itself.
+*  Using 1- vs. 2- vs. 3-grams
+*  Stepwise paired classification versus comparing multiple groups simultaneously
 
-Finally, there are several things still missing or needed for this workflow:
-
-*  How to feed a text dataset in, and get a table of the predicted categories back out.
-*  How to store the values from the Bayes prediction table so I can use them in calculations rather than type them in by hand.
-*  How to randomize which comments are used for training, testing datasets.
-
-Other pages examine other classifiers besides Naive Bayes.
+Additional working notes are located at the end of this page.
 
@@
 ##Initial Setup
+Calling the required libraries.
+
 ```{r}
@@
 
-##Version 2.0 of Naive Bayes (NB) Protocol
+##Version 2.0.b of Naive Bayes (NB) Protocol
 ###Initial Data Input
@@
 
-```{}
+```{r}
 #Remove numerals. 
@@
 
-```{}
+```{r}
 #Stopword removal. 
@@
 ***
-START BREAK POINT - Stemming vs. lemmatization
+START BRANCH POINT - Stemming vs. lemmatization
 ***
@@
 ***
-END BREAK POINT
+END BRANCH POINT
 ***
@@
 NLP_Tokenizer <- function(x) {
-      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
+      unlist(lapply(ngrams(words(x), 3), paste, collapse = " "), use.names = FALSE)
 }
@@
 
-This code assumes comments are random. **Probably want to try randomizing them.**
+This code uses the comments in the order they are listed in the original file. **Need a block for  randomizing them.**
 ```{r}
@@
 ```{r}
-comments_train_labels <- comments_rawVERB[1:7005, ]$type
-comments_test_labels <- comments_rawVERB[7006:9340,]$type
+comments_train_labels_type <- comments_rawVERB[1:7005, ]$type
+comments_test_labels_type <- comments_rawVERB[7006:9340,]$type
 ```
@@
 ```{r}
-prop.table(table(comments_train_labels))
-prop.table(table(comments_test_labels))
+prop.table(table(comments_train_labels_type))
+prop.table(table(comments_test_labels_type))
 ```
@@
 ***
-The above code block incorrectly uses sequential roww, not random rows.  I need to find a way to incorporate randomization.
+The above code block incorrectly uses sequential rows, not random rows.  I need to find a way to incorporate randomization.
 
@@
 
-Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
+Limit Document Term Matrix to only include words in the comments_freq_words vector. Using all the rows, but we want to limit the columns to these words in the frequency vector.
 ```{r}
@@
 
-###Train Model, Predict, Evaluate
-Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic.
-Evaluate the prediction with the actual data using a crosstable from the gmodels package.
+###Train Model, Predict, Evaluate for Type
+This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic.
+It evaluates the prediction with the actual data using a crosstable from the gmodels package.
 ```{r}
-comments_classifier <- naiveBayes(comments_train, comments_train_labels)
-comments_test_pred <- predict(comments_classifier, comments_test)
-CrossTable(comments_test_pred, comments_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
+comments_classifier_type <- naiveBayes(comments_train, comments_train_labels_type)
+comments_test_pred_type <- predict(comments_classifier_type, comments_test)
+CrossTable(comments_test_pred_type, comments_test_labels_type, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
 ```
 
-The cross table will list how many of the predicted categories are in their correct categories. Adding diagonally and dividing will calculate the percent accuracy.
+The cross table list how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting TYPE.
 
 ```{r}
-(44+473+1142+206)/23.35
+print("Overall TYPE accuracy")
+(14+45+1346+19)/23.35
 ```
 
+###Train Model, Predict, Evaluate for LOCUS
+This is the same block as above but uses the "Locus" data column.
+
+```{r}
+comments_classifier_locus <- naiveBayes(comments_train, comments_train_labels_locus)
+comments_test_pred_locus <- predict(comments_classifier_locus, comments_test)
+CrossTable(comments_test_pred_locus, comments_test_labels_locus, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
+```
+
+Again, a cross table shows how many of the predicted categories match the correct categories for LOCUS.
+
+```{r}
+print("Overall LOCUS accuracy")
+(62+64+931+3)/23.35
+```
+
+###OPTIONAL Output
 To make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame.
@@
 
-```{r}
-predicted_category <- data.frame(comments_test_pred,comments_test_labels)
-head(predicted_category)
-write.csv(predicted_category, file = "PredictedCategory_Data.csv")
+```{}
+predicted_category_type <- data.frame(comments_test_pred_type,comments_test_labels_type)
+head(predicted_category_type)
+write.csv(predicted_category_type, file = "PredictedCategoryType_Data.csv")
+```
+
+```{}
+predicted_category_locus <- data.frame(comments_test_pred_locus,comments_test_labels_locus)
+head(predicted_category_locus)
+write.csv(predicted_category_locus, file = "PredictedCategoryLocus_Data.csv")
 ```
@@
 ***
-START BREAK POINT
-***
-1. The e1071 naive bayes classifier is binomial. Need to incorporate multinomial.
-2. Need to add prediction output.
-3. Is dropping numerical values down to categorical yes/no variables removing useful info?
-
-***
-END BREAK POINT
+START COMMENTS
 ***
 
-***
-***
-####\
+##Permutation Tests of Naive Bayes (NB) Method
+Using single words, Version 2.0 correctly identifies comment "type" ~80% of the time and comment "locus" ~59% of the time.
 
-##Visualization
-This code generates a wordcloud of the frequency of single words in the dataset using the package "wordcloud".
-```{r}
-wordcloud(comments_corpus_clean, max.words = 50, random.order = FALSE)
-```
+Permutations to Test:
 
-Compare wordclouds between 3 groups.
-```{r}
-ONE <- subset(comments_raw, type == "1_Basic")
-TWO <- subset(comments_raw, type == "2_Writing")
-THREE <- subset(comments_raw, type == "3_Technical")
-FOUR <- subset(comments_raw, type == "4_Logic")
-
-wordcloud(ONE$text, max.words = 50, scale = c(3, 0.5))
-wordcloud(TWO$text, max.words = 50, scale = c(3, 0.5))
-wordcloud(THREE$text, max.words = 50, scale = c(3, 0.5))
-wordcloud(FOUR$text, max.words = 50, scale = c(3, 0.5))
-```
-
-Show the 5 most frequent words in the data:
-```{r}
-comment_sack <- TermDocumentMatrix(comments_corpus_clean)
-comment_pack <- as.matrix(comment_sack)
-comment_snack <- sort(rowSums(comment_pack), decreasing = TRUE)
-comment_hack <- data.frame(word = names(comment_snack), freq=comment_snack)
-head(comment_hack, 10)
-```
-
-And the 5 most frequent words from each class:
-```{r}
-wordcloud(ONE$text, max.words = 10, scale = c(3, 0.5))
-wordcloud(TWO$text, max.words = 10, scale = c(3, 0.5))
-wordcloud(THREE$text, max.words = 10, scale = c(3, 0.5))
-wordcloud(FOUR$text, max.words = 10, scale = c(3, 0.5))
-```
-
-####\
-
-##Permutation Tests of Naive Bayes (NB) Method
-Version 1.0 correctly identifies comment subcategories ~79% of the time when using single words, and 68.3% of the time when using 2-grams. Permutations to test:
-
-*  Compare 2-grams, 3-grams rather than single words
+*  Compare 1-grams, 2-grams, 3-grams
 *  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
-    +  Ex.: Basic vs. Not Basic (1_Basic vs. not 1_Basic)
+    +  Initial approach will be to split between largest group and balance in each category
+    +  Ex.: type is 3-technical vs. not 3_technical
 *  Modify the variables within the NB analysis itself. 
@@
     +  This may remove evidence of tense
+    +  Better strategy is to lemmatize
 *  Remove/Leave stopwords
+    +  Look into alternative terms lists to remove
 *  Remove/Leave capitalization
@@
     +  This may remove evidence of questions vs statements
+
+Known Issues:
+
+1. The e1071 naive bayes classifier is binomial. Need to incorporate multinomial.
+2. Is dropping numerical values down to categorical yes/no variables removing useful info?
+3. How to lemmatize instead of stemming?
+
+Keep the code blocks organized so that testing various permutations is mostly on/off.
+
+Create a SEPARATE version for changing larger blocks.
+
+Finally, there are several things still missing or needed for this workflow:
+
+*  How to feed a text dataset in, and get a table of the predicted categories back out.
+*  How to store the values from the Bayes prediction table so I can use them in calculations rather than type them in by hand.
+*  How to randomize which comments are used for training, testing datasets.
+
+Other pages examine other classifiers besides Naive Bayes.
+
+***
+END COMMENTS
+***

### 2019-02-21 hiding/Naive_bayes_V2r.Rmd
cleanup

--- 2019-01-31 Naive_bayes_V2r.Rmd
@@
 NLP_Tokenizer <- function(x) {
-      unlist(lapply(ngrams(words(x), 3), paste, collapse = " "), use.names = FALSE)
+      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
 }
@@
 print("Overall subject accuracy")
-(14+54+1318+28)/23.35  #Need to find a way to calculate this automatically. 
+(43+473+1131+184)/23.35  #Need to find a way to calculate this automatically. 
 ```
@@
 print("Overall structure accuracy")
-(66+64+949+5+1)/23.35  #Need to find a way to calculate this automatically. 
+(22+488+292+518+33+10)/23.35  #Need to find a way to calculate this automatically. 
 ```
@@
 
-```{}
+```{r}
 predicted_category_subject <- data.frame(comments_test_pred_subject,comments_test_labels_subject)
@@
 
-```{}
+```{r}
 predicted_category_structure <- data.frame(comments_test_pred_structure,comments_test_labels_structure)
@@
 
-##Permutation Tests of Naive Bayes (NB) Method
-Using single words, Version 2.0.b correctly identifies comment "subject" ~80% of the time and comment "structure" ~59% of the time.
-
-Permutations to test with random subsets version (2.0.r):
-
-*  Compare 1-grams, 2-grams, 3-grams
-*  Compare outcomes for 3 independent randomized sets of 1-grams
-
-The pseudo-code block below can pull random lines from a data table, but does not work correctly on a document term matrix.
-The second problem is how to pull the training and test data sets out and process them identically.
-
-```{}
-set.seed(1234) 		# ensures the same random set gets picked each time.
-
-comments_raw_dtm_train <- sample_frac(comments_raw, .75)	# sample_frac pulls out 75% of "comments_raw" dataset randomly using the sample sequence identified by set.seed(1234)
-
-comments_raw_dtm_test <- setdiff(comments_raw, comments_raw_dtm_train)	#setdiff creates a "test" dataset using anything NOT in the "comments_raw_dtm_train" subset of "comments_raw."
-```
-
-The solution was not to pull the data twice, but to randomly (but repeatably) shuffle the starting data table. How I solved this was:
-1. Create a vector of sequential numbers the same length as the number of data frame rows.
-2. Randomize the sequential numbers using set.seed to make pattern reproducible.
-3. Add the randomized numbers as a new column to the data frame.
-4. Re-order the data table to the randomizer column.
-
-This makes it easy to generate a new sequence of randomized training and test samples to run replicates. Only one number needs to be changed.
-
-```{}
+```{r}
 #Creates a data frame that compares predicted and actual labels for subject and structure
@@
 
-```{}
+```{r}
 if_else (tally(comments_test_labels_subject) != tally(comments_test_pred_subject),"YES","no")

### 2019-02-21 hiding/project.Rmd
cleanup

--- 2018-12-16 project.Rmd
@@
 ---
-title: "Project"
+title: "Project Summary"
 author: "Dan Johnson"

### 2019-02-21 hiding/questions.rmd
cleanup

# R Code
#### I cannot get numbers on CSV extracted data frames.

Detailed notes are in "Data_intake_1.rmd", starting Line 102. Any ideas?

####\

#### Why can I not append the read.csv command?
This works fine:
```
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18.csv')
head(base_data)
```

This fails:
```
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18.csv', header=FALSE)
head(base_data)
```
I get:
```	
	Error in read_csv(file = "data/coded_full_comments_dataset_Spring18.csv",  : 
  unused argument (header = FALSE)
```

-----

# Project

-----

# Github
#### Question about fork and branch. 
When I fork another repo, I get comments saying I am "#" steps ahead of it. Should I clone instead?

-----

# Archived
#### How to create files in R, then push to GitHub as a complete repo
Instructions for doing it on this website. [https://cfss.uchicago.edu/git05.html]()

-----

####\

### 2019-02-21 index.Rmd
cleanup

--- 2019-01-31 index.Rmd
@@
 ---
-title: "Analyzing TA Comments on Student Writing in R"
+title: "Automated Classification of TA Comments on Student Writing"
 ---
@@
 
-## Background
-I am interested in understanding how students develop technical writing skills in biology, and more specifically, how we can apply data science methods to improve that process.
+## Overview
+I am interested in understanding how students develop technical writing skills in biology, and how we might improve that process by applying data science methods.
 
 My current NSF-sponsored project asks if and how scripted instruction, automated support, and holistic feedback accelerates student development as writers.
-One of our specific aims is to see if changing the structure and focus of instructors' comments affects how students' writing develops over time.
-To do this, we need to extract instructor comments from student reports, classify those comments, and correlate the types of comments TAs make with student performance.
-
-My project for the Faculty Learning Community for R is partly exploratory, in I am asking: what information is present and can be mined from the TA comments dataset?
-The project also is partly predictive in that I hope to build a text classification workflow that automates the process of categorizing TA comments on student writing.
+One specific aim is to see if changing the structure and focus of **instructors' comments** affects how students' writing develops over time.
+To assess this we must extract instructor comments from student reports, categorize those comments according to their subject and structure, then correlate the types of comments TAs make with student performance.
 
@@
 ## Prior Work
-In Summer 2018, we extracted and tabulated ~11,000 TA comments from ~500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses collected during a single semester.
-Using an iteratively generated codebook, all 11,000 comments were classified in two categories:
+In Summer 2018, we extracted and tabulated ~11,000 TA comments from >500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses during a single semester.
+Using an iteratively generated codebook, the 11,000 TA comments were sepparately classified in two categories:
 
-* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
-* __Structure__: did the comment Point Out an error without providing other help, provide simple Copy Correction, provide declarative General or Specific Information only, or provide Holistic coaching that fosters student thinking and long-term improvement?
+* __Subject__: did the comment focus mainly on **Basic Criteria, Technical Issues, Writing Quality, Logic and Thinking,** or **Other** issues
+* __Structure__: did the comment **Point** to an error without providing other help, provide simple **Copy Correction**, provide declarative **General or Specific Information** only, or provide **Holistic Coaching** that fosters student thinking and long-term improvement?
 
 #### \   
-Frequency counts of unambiguously coded TA comments were sorted into a two-axis contingency table that provided two important insights:
+Counts of coded TA comments were tabulated into a 2-way contingency table that provided us with a direct measure of the most common types of comments TAs made on student work, and a relative estimate of which elements of writing TAs emphasize most often.
 
-* A snapshot of the most common types of comments TAs made, and 
-* A relative estimate of where TAs are placing their greatest effort.
+While classifying TA comments this way is informative, doing so by hand is impractical: 10-12,000 comments need to be scored EACH semester.
 
 #### \
-## Specific Project Goals
-TA comments and meta-data for each student report can be extracted and compiled into a CSV datafile using existing Unix scripts.
-However routinely classifying TA comments by hand is impractical; 10-12,000 comments need to be scored EACH SEMESTER.
-An automated classification method is needed both for speed and consistency.
 
-My specific project goals were/are to create an R-based text analysis workflow that:
+##Specific Project Goals
+My project for the **Faculty Learning Community for R** was to create an automated comment classifier that could generate two-axis contingency tables comparable to the one generated using hand-coded data.
+This automated classifier will be used as part of the larger NSF-funded research project to:
 
-1.
-**Converts TA comments and metadata into a de-identified "tidy" dataframe for R.** This includes converting the master CSV data set to an anonymized data frame with all names and confidential metadata re-keyed to unique random identifiers.
-2.
-**Uses text features to assign TA comments to the categories defined in our existing codebook**.
-The goal is to be able to assign instructor comments into our pre-defined Subject and Structure codes with 90% accuracy or greater.
+*  Improve both speed and consistency when classifying TA comments across multiple semesters. 
+*  Monitor how the types and focus of TAs comments on student writing change over time.
 
 #### \   
+## Site Map
+Insert a text-based site map of the major pages and branch points here.
+
+#### \

### 2019-02-21 nextsteps.Rmd
cleanup

--- 2019-02-21 background.Rmd
@@
 ---
-title: "Background for the Project"
+title: "Next Steps"
 bibliography: library.bib
@@
 
-Undergraduate introductory STEM courses typically include technical writing training.
-Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed.
-Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills.
-
-Technical writing training usually is part of a lab course associated with a lecture class.
-These lab sections may be be led by graduate teaching assistants or even advanced undergraduates.
-These instructors may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.
-
-In 2014 we launched an automated data collection process we call **SAWHET** that systematically collects structured writing and other data from student writers and graduate instructors.
-Our resarch goal is to better understand student and instructor behaviors within the technical writing instructional space.
-From this empirical basis our practical goal is to develop techniques for teaching technical writing that:
-
-*  Can be implemented at large scale by less experienced teaching assistants;
-*  Emphasize student development through holistic coaching;
-*  Shift student and instructor attention away from copy-correction towards larger writing issues; and
-*  Embody other best practices from 40+ years of research literature on "writing in the disciplines."
-
-We have a structured database of >5000 laboratory reports written by undergraduate students for introductory biology courses.
-In addition we have all comments that a TA made on the individual reports, plus associated metadata:
-
-*  Context. Was it first or second report of the semester? Initial submission or revision? What course and topic in the course? What year in college was the author?
-*  Instructor information. Which TA provided the feedback? What score was assigned, and what what the rationale behind that score?
-
-We are using multiple text-mining methods to map key features of student writing, as well as connections and correlations between those features.
-These insights then inform how we teach technical writing and train graduate instructors.
-
-####\
-
-## Context for This Project & Prior Work
-Students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead.
-To that end, we train graduate teaching assistants to use a **global first, specific later** approach to grading undergraduate technical writing.
-We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues.
-Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection.
-As students' writing improves, the TA can focus on smaller issues.
-
-This strategy is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy.
-One of our ongoing questions has been, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**
-
-In Summer 2018, we began analyzing TAs' comments from student papers.
-We extracted and tabulated all comments from approximately 500 pairs of reports (initial submissions and revisions) collected during a single semester in 3 different courses.
-Initial data cleaning involved removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent.
-In all, the final tabulated dataset included ~11,000 separate comments.
-
-We next used a subset of 100 randomly selected comments to build an initial qualitative codebook for sorting comments into types.
-The initial codebook was re-tested using 1100 additional randomly selected comments.
-Using the final codebook, we sorted all 11,000 comments into separate categories (see LINK for complete codebook.)
-
-Coded TA comments were sorted on two axes:
-
-* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
-* __Structure__: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?
-
-The sorted contingency table provided us with a relative estimate of TA effort.
-
-####\
-
-## The Central Problem of This Project
 Qualitative coding of TA comments is informative but unsustainable.

### 2019-02-21 docs/hiding/Multinomial_bayes_V1.Rmd
update

--- 2019-02-21 hiding/Naive_bayes_V2r.Rmd
@@
 ---
-title: "Naive Bayes Workflow Ver 2.0.r"
+title: "Multinomial Naive Bayes Workflow Ver 1.0."
 author: "Dan Johnson"
-date: "1/29/2019"
+date: "2/13/2019"
 output: html_document
@@
 ##Summary
-This is **Version 2.0.r** of the basic workflow for classifying TA comments using a naive Bayes model.
-It pre-processes text using the same command set as Ver 2.0.b.
-The "r" designates this version has been modified to be able to randomize which rows are selected for training and test datasets.
+This is **Version 1.0** of a basic workflow for classifying TA comments using a multinomial naive Bayes model.
+It pre-processes text using the same command set as NB-Ver 2.0.
+This version replaces categorical naive Bayes with a multinomial classifier.
 
@@
 ```{r}
-library(tidyverse)
-library(tidytext)
-library(tidyr)
+
 library(tm)
-library(SnowballC)
-library(wordcloud)
+library(RTextTools)
 library(e1071)
-library(gmodels)
+library(dplyr)
+library(caret)
+library(doMC)
+registerDoMC(core=detectCores())
+
+#library(tidyverse)
+#library(tidytext)
+#library(tidyr)
+#library(SnowballC)
+#library(wordcloud)
+#library(gmodels)
 ```
@@
 #Read in TA comments from CSV file.
-base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
+subset_M <- read.csv("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/comments_subset.csv", stringsAsFactors=FALSE)
+glimpse(subset_M)
 ```
+```{r}
+set.seed(1)
+subset_M <- subset_M[sample(nrow(subset_M)), ]
+subset_M <- subset_M[sample(nrow(subset_M)), ]
+glimpse(subset_M)
+```
+
+```{r}
+subset_M$subject <- as.factor(subset_M$subject)
+subset_M$structure <- as.factor(subset_M$structure)
+```
+
+```{r}
+corpus <- Corpus(VectorSource(subset_M$text))
+corpus
+inspect(corpus[1:3])
+```
+
+```{r}
+corpus.clean <- corpus %>%
+  tm_map(content_transformer(tolower)) %>% 
+  tm_map(removePunctuation) %>%
+  tm_map(removeNumbers) %>%
+  tm_map(removeWords, stopwords()) %>%
+  tm_map(stripWhitespace)
+```
+
+```{r}
+dtm <- DocumentTermMatrix(corpus.clean)
+# Inspect the dtm
+inspect(dtm[40:50, 10:15])
+```
+
+```{r}
+subset_M.train <- subset_M[1:7005,]
+subset_M.test <- subset_M[7006:9340,]
+
+dtm.train <- dtm[1:7005,]
+dtm.test <- dtm[7006:9340,]
+
+corpus.clean.train <- corpus.clean[1:7005]
+corpus.clean.test <- corpus.clean[7006:9340]
+```
+
+```{r}
+dim(dtm.train)
+```
+
+```{r}
+fivefreq<- findFreqTerms(dtm.train, 5)
+length((fivefreq))
+
+dtm.train.nb <-DocumentTermMatrix(corpus.clean.train, control=list(dictionary=fivefreq))
+
+dim(dtm.train.nb)
+
+dtm.test.nb <- DocumentTermMatrix(corpus.clean.test, control=list(dictionary = fivefreq))
+
+dim(dtm.train.nb)
+```
+
+We use a variation of the multinomial Naive Bayes algorithm known as binarized (boolean feature) Naive Bayes due to Dan Jurafsky.
+In this method, the term frequencies are replaced by Boolean presence/absence features.
+The logic behind this being that for sentiment classification, word occurrence matters more than word frequency.
+```{r}
+# Function to convert the word frequencies to yes (presence) and no (absence) labels
+convert_count <- function(x) {
+  y <- ifelse(x > 0, 1,0)
+  y <- factor(y, levels=c(0,1), labels=c("No", "Yes"))
+  y
+}
+```
+
+```{r}
+# Function to avoid converting word frequencies to yes (presence) and no (absence) labels
+convert_count <- function(x) {
+  y <- x
+}
+```
+
+# Apply the convert_count function to get final training and testing DTMs
+```{r}
+trainNB <- apply(dtm.train.nb, 2, convert_count)
+testNB <- apply(dtm.test.nb, 2, convert_count)
+```
+
+```{r}
+# Train the classifier
+system.time( classifier <- naiveBayes(trainNB, subset_M.train$subject, laplace = 1) )
+#system.time( classifier <- naiveBayes(trainNB, subset_M.train$structure, laplace = 1) )
+```
+
+Testing the Predictions
+```{r}
+# Use the NB classifier we built to make predictions on the test set.
+system.time( pred <- predict(classifier, newdata=testNB) )
+```
+
+```{r}
+# Create a truth table by tabulating the predicted class labels with the actual class labels 
+table("Predictions"= pred,  "Actual" = subset_M.test$subject )
+```
+
+```{r}
+# Prepare the confusion matrix
+conf.mat <- confusionMatrix(pred, subset_M.test$subject)
+conf.mat
+print ("   ")
+
+conf.mat$byClass
+print ("   ")
+
+conf.mat$overall
+print ("   ")
+
+conf.mat$overall['Accuracy']
+
+```
+
+https://rpubs.com/cen0te/naivebayes-sentimentpolarity
 

### 2019-02-21 docs/hiding/Multinomial_bayes_V3_0_b.Rmd
update

---
title: "Multinomial Naive Bayes Workflow Ver 3.0.b."
author: "Dan Johnson"
date: "2/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Summary
This is an alternative version of my basic NB workflow for classifying TA comments. The code is slightly different than what was used for Categorical NB testing, but takes in text using the same command set as NB-Ver 2.0. It is based on a tutorial located here:
[https://rpubs.com/cen0te/naivebayes-sentimentpolarity]

This alternative code allows for true multinomial analysis, by switching between using binary yes/no categorization of words' presence or absence, and the actual frequencies. There is a trade-off though; to use non-categorical values, I can only have 2 assignment bins: in-group versus out-group. If I use multiple groups, the data fail to parse correctly. If I want to look at multiple classification bins at the same time, I have to use the categorical values method.

The point seems moot overall, since categorical NB seems to perform better than true multinomial NB. For both Subject and Structure, I reach 75-84% overall accuracy if I sort comments into ingroup/outgroup categories using binary (yes/no) word frequencies. Accuracy drops to ~65% if I use the raw numeric frequencies. Changing the Laplace value does not affect overall performance.

Naive Bayes permutations seem to plateau at 80-85%, regardless of permutations.

####\

##Initial Setup
Calling the required libraries.

```{r}
library(tm)
library(RTextTools)
library(e1071)
library(dplyr)
library(caret)
library(doMC)
registerDoMC(core=detectCores())

library(quanteda)
library(RColorBrewer)
library(ggplot2)

#library(tidyverse) 
#library(tidytext)
#library(tidyr)
```

####\

##Version 3.0.b of Naive Bayes (NB) Protocol
###Initial Data Input
```{r}
#Read in TA comments from CSV file. 
#For convenience, I am reading in a shorter CSV file created previously using ver.2.0 code. This CSV file only lists 
#whether a comment is in or out of the 2 main groups, "Subject:3_technical" and "Structure:4_specific"

subset_M <- read.csv("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/comments_binary.csv", stringsAsFactors=FALSE)
glimpse(subset_M)
```

```{r}
set.seed(2012) #Randomizer
subset_M <- subset_M[sample(nrow(subset_M)), ]
glimpse(subset_M)
```

```{r}
text.corpus<-corpus(subset_M$text)
docvars(text.corpus)<-subset_M$subject
glimpse(text.corpus)
```

```{r}
ingroup.plot<-corpus_subset(text.corpus,docvar1=="ingroup")
ingroup.plot<-dfm(ingroup.plot, tolower=TRUE,remove_punct=TRUE,remove_twitter=TRUE,remove_numbers=TRUE,remove=stopwords(source="smart"))
ingroup.col <-brewer.pal(10, "BrBG")
textplot_wordcloud(ingroup.plot, min_count=16, color=ingroup.col)
title("Ingroup Wordcloud", col.main="grey14")

outgroup.plot<-corpus_subset(text.corpus,docvar1=="outgroup")
outgroup.plot<-dfm(outgroup.plot, tolower=TRUE,remove_punct=TRUE,remove_twitter=TRUE,remove_numbers=TRUE,remove=stopwords(source="smart"))
outgroup.col <-brewer.pal(10, "BrBG")
textplot_wordcloud(outgroup.plot, min_count=16, color=outgroup.col)
title("Outgroup Wordcloud", col.main="grey14")
```

Let’s separate the training and test data-
```{r}
#separating Train and test data
text.train<-subset_M[1:7005,]
text.test<-subset_M[7006:nrow(subset_M),]

msg.dfm <- dfm(text.corpus, tolower = TRUE)  #generating document freq matrix
msg.dfm <- dfm_trim(msg.dfm, min_termfreqError = 5, min_docfreq = 3)  
msg.dfm <- dfm_weight(msg.dfm)

head(msg.dfm)
```

#trining and testing data of dfm 
```{r}
msg.dfm.train<-msg.dfm[1:7005,]

msg.dfm.test<-msg.dfm[7006:nrow(subset_M),]

```

Training the Naive Bayes classifier-

nb.classifier<-textmodel_NB(msg.dfm.train,spam.train[,1])
nb.classifier
## Fitted Naive Bayes model:
## Call:
##  textmodel_NB.dfm(x = msg.dfm.train, y = spam.train[, 1])
## 
## 
## Training classes and priors:
## spam  ham 
##  0.5  0.5 
## 
##        Likelihoods:      Class Posteriors:
## 30 x 4 Matrix of class "dgeMatrix"
##                 spam          ham       spam       ham
## you     5.001507e-03 0.0096798156 0.34067144 0.6593286
## have    4.322289e-03 0.0042303673 0.50537386 0.4946261
## 1       2.695748e-03 0.0009529526 0.73882413 0.2611759
## new     3.492485e-03 0.0010753934 0.76457487 0.2354251
## .       6.965338e-03 0.0168302131 0.29271598 0.7072840
## please  2.339097e-03 0.0011593603 0.66860811 0.3313919
## call    1.058603e-02 0.0021859571 0.82884759 0.1711524
## i       8.439760e-04 0.0112106647 0.07001254 0.9299875
## wait    1.860817e-04 0.0011538316 0.13887596 0.8611240
## for     5.699340e-03 0.0045025239 0.55865674 0.4413433
## hope    2.334040e-04 0.0017258550 0.11912872 0.8808713
## tonight 1.137075e-04 0.0011106417 0.09287182 0.9071282
## too     3.802754e-05 0.0017024748 0.02184860 0.9781514
## bad     1.232420e-04 0.0006045270 0.16934219 0.8306578
## as      1.339518e-03 0.0020699791 0.39287852 0.6071215
## well    2.938089e-04 0.0017334850 0.14492664 0.8550734
## but     2.528948e-04 0.0043933716 0.05442968 0.9455703
## rock    3.802754e-05 0.0002684845 0.12406542 0.8759346
## night   3.003905e-04 0.0017976398 0.14317739 0.8568226
## anyway  3.802754e-05 0.0005405216 0.06572915 0.9342709
## going   1.538819e-04 0.0023951976 0.06036762 0.9396324
## a       7.856726e-03 0.0064918622 0.54756091 0.4524391
## now     6.254232e-03 0.0028758075 0.68501697 0.3149830
## good    6.723203e-04 0.0030342352 0.18138681 0.8186132
## speak   8.003838e-04 0.0004728416 0.62862694 0.3713731
## to      1.113210e-02 0.0075761991 0.59503541 0.4049646
## soon    2.642059e-04 0.0010608467 0.19939274 0.8006073
## today   1.120666e-03 0.0019041688 0.37048833 0.6295117
## is      4.451802e-03 0.0058153446 0.43359683 0.5664032
## accept  3.802754e-05 0.0003188419 0.10655871 0.8934413

The model outputs the Probabilities of the message being Spam or ham.
Let’s Test the Model

pred<-predict(nb.classifier,msg.dfm.test)

#generating a confusion matrix

# use pred$nb.predicted to extract the class labels
table(predicted=pred$nb.predicted,actual=spam.test[,1])
##       actual
## predicted ham spam
##      ham  952    7
##      spam  16  140

16 text examples wrongly classified for ham and 7 examples wrongly classified for spam.
In the Confusion matrix,the diagonals are the correctly classified examples and the off-diagonals are the incorrectly classified text instances.

Now let’s calculate the accuracy of the model –

#acccuracy of the classifier on Test data
mean(pred$nb.predicted==spam.test[,1])*100
## [1] 97.93722

Now 98% accuracy is a good amount of accuracy on unseen random test data.

```{r}
subset_M$subject <- as.factor(subset_M$subject)
subset_M$structure <- as.factor(subset_M$structure)
```

```{r}
corpus <- Corpus(VectorSource(subset_M$text))
corpus
inspect(corpus[4:6])
```

```{r}
#In my code this block is split into substeps to make A/B testing easier.
corpus.clean <- corpus %>%
  tm_map(content_transformer(tolower)) %>% 
  tm_map(removePunctuation) %>%
  tm_map(removeNumbers) %>%
  tm_map(removeWords, stopwords(kind="en")) %>%
  tm_map(stripWhitespace)
```
```{r}
corpus.dfm <- dfm(corpus.clean, tolower=TRUE)

# Inspect the dtm
#inspect(dtm[40:50, 10:15])
```

```{r}
dtm <- DocumentTermMatrix(corpus.clean)

# Inspect the dtm
inspect(dtm[40:50, 10:15])
```

```{r}
subset_M.train <- subset_M[1:7005,]
subset_M.test <- subset_M[7006:9340,]

dtm.train <- dtm[1:7005,]
dtm.test <- dtm[7006:9340,]

corpus.clean.train <- corpus.clean[1:7005]
corpus.clean.test <- corpus.clean[7006:9340]
```

```{r}
dim(dtm.train)
```

```{r}
fivefreq<- findFreqTerms(dtm.train, 5)
length((fivefreq))

dtm.train.nb <-DocumentTermMatrix(corpus.clean.train, control=list(dictionary=fivefreq))

dim(dtm.train.nb)

dtm.test.nb <- DocumentTermMatrix(corpus.clean.test, control=list(dictionary = fivefreq))

dim(dtm.train.nb)
```

Here are the two blocks of code that switch between using a variation of the multinomial Naive Bayes algorithm known as binarized (boolean feature) Naive Bayes (from Dan Jurafsky) versus non-binarized. In this method, the term frequencies are replaced by Boolean presence/absence features. The logic behind this is that for sentiment classification, word occurrence matters more than word frequency. For my purposes, either might work, so both must be tested; put "r" in the opening brackets for whichever is needed.

```{r}
# Function to convert the word frequencies to yes (presence) and no (absence) labels
convert_count <- function(x) {
  y <- ifelse(x > 0, 1,0)
  y <- factor(y, levels=c(0,1), labels=c("No", "Yes"))
  y
}
```

```{}
# Function to KEEP word frequencies while still using the same variable names and code structure.
convert_count <- function(x) {
  y <- x
}
```

```{r}
# Apply the convert_count function to get final training and testing DTMs (whichever was chosen earlier)
trainNB <- apply(dtm.train.nb, 2, convert_count)
testNB <- apply(dtm.test.nb, 2, convert_count)
```

```{r}
# Train the classifier. Not sure what system.time is needed for, except to show when block finishes.
system.time( classifier <- naiveBayes(trainNB, subset_M.train$subject, laplace = 1) )
system.time( classifier2 <- naiveBayes(trainNB, subset_M.train$structure, laplace = 1) )
```

##Testing the Predictions

```{r}
# Use the NB classifier we built to make predictions on the test set. Again, why use system.time?
system.time( pred <- predict(classifier, newdata=testNB) )
system.time( pred2 <- predict(classifier2, newdata=testNB) )
```

```{r}
# Create a truth table by tabulating the predicted class labels with the actual class labels 
line<-c("******************")
print("Subject")
table("Predictions"= pred,  "Actual" = subset_M.test$subject )
print(line)
print("Structure")
table("Predictions2"= pred2,  "Actual" = subset_M.test$structure )
```

```{r}
# Prepare the confusion matrix
conf.mat <- confusionMatrix(pred, subset_M.test$subject)
conf.mat
glimpse(line)

conf.mat$byClass
glimpse(line)

conf.mat$overall
glimpse(line)

conf.mat$overall['Accuracy']

glimpse(line)
glimpse(line)

# Prepare the confusion matrix
conf.mat2 <- confusionMatrix(pred2, subset_M.test$structure)
conf.mat2
glimpse(line)

conf.mat2$byClass
glimpse(line)

conf.mat2$overall
glimpse(line)

conf.mat2$overall['Accuracy']

```

###OPTIONAL Output
To make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame. The full data frame will be too long to display; use "head" to check format, then output the full dataset to CSV for further review. By default the CSV file should be saved to the current working directory.

table("Predictions"= pred,  "Actual" = subset_M.test$subject )
print(line)
print("Structure")
table("Predictions2"= pred2,  "Actual" = subset_M.test$structure )

```{r}
predicted_category_subject <- data.frame(pred,subset_M.test$subject)
head(predicted_category_subject)
write.csv(predicted_category_subject, file = "PredictedCategorySubject_Data.csv")
```

```{r}
predicted_category_structure <- data.frame(pred,subset_M.test$structure)
head(predicted_category_structure)
write.csv(predicted_category_structure, file = "PredictedCategoryStructure_Data.csv")
```

***
START COMMENTS, OLD work notes
***

***
END COMMENTS
***

### 2019-02-21 docs/hiding/Multinomial_bayes_V3_1.Rmd
update

--- 2019-02-21 docs/hiding/Multinomial_bayes_V3_0_b.Rmd
@@
 ```
-
 ```{r}
@@
 msg.dfm <- dfm(text.corpus, tolower = TRUE)  #generating document freq matrix
-msg.dfm <- dfm_trim(msg.dfm, min_termfreqError = 5, min_docfreq = 3)  
+msg.dfm <- dfm_trim(msg.dfm, min_termfreqError = 5, min_docfreq = 1)  
 msg.dfm <- dfm_weight(msg.dfm)
@@
 
+head(msg.dfm.train)
+head(msg.dfm.test)
 ```
 
-Training the Naive Bayes classifier-
+```{r}
+#library(quanteda)
+nb.classifier<-textmodel_nb(msg.dfm.train, text.train[,1])
+nb.classifier
+```
 
-nb.classifier<-textmodel_NB(msg.dfm.train,spam.train[,1])
-nb.classifier
-## Fitted Naive Bayes model:
-## Call:
-##  textmodel_NB.dfm(x = msg.dfm.train, y = spam.train[, 1])
-## 
-## 
-## Training classes and priors:
-## spam  ham 
-##  0.5  0.5 
-## 
-##        Likelihoods:      Class Posteriors:
-## 30 x 4 Matrix of class "dgeMatrix"
-##                 spam          ham       spam       ham
-## you     5.001507e-03 0.0096798156 0.34067144 0.6593286
-## have    4.322289e-03 0.0042303673 0.50537386 0.4946261
-## 1       2.695748e-03 0.0009529526 0.73882413 0.2611759
-## new     3.492485e-03 0.0010753934 0.76457487 0.2354251
-## .       6.965338e-03 0.0168302131 0.29271598 0.7072840
-## please  2.339097e-03 0.0011593603 0.66860811 0.3313919
-## call    1.058603e-02 0.0021859571 0.82884759 0.1711524
-## i       8.439760e-04 0.0112106647 0.07001254 0.9299875
-## wait    1.860817e-04 0.0011538316 0.13887596 0.8611240
-## for     5.699340e-03 0.0045025239 0.55865674 0.4413433
-## hope    2.334040e-04 0.0017258550 0.11912872 0.8808713
-## tonight 1.137075e-04 0.0011106417 0.09287182 0.9071282
-## too     3.802754e-05 0.0017024748 0.02184860 0.9781514
-## bad     1.232420e-04 0.0006045270 0.16934219 0.8306578
-## as      1.339518e-03 0.0020699791 0.39287852 0.6071215
-## well    2.938089e-04 0.0017334850 0.14492664 0.8550734
-## but     2.528948e-04 0.0043933716 0.05442968 0.9455703
-## rock    3.802754e-05 0.0002684845 0.12406542 0.8759346
-## night   3.003905e-04 0.0017976398 0.14317739 0.8568226
-## anyway  3.802754e-05 0.0005405216 0.06572915 0.9342709
-## going   1.538819e-04 0.0023951976 0.06036762 0.9396324
-## a       7.856726e-03 0.0064918622 0.54756091 0.4524391
-## now     6.254232e-03 0.0028758075 0.68501697 0.3149830
-## good    6.723203e-04 0.0030342352 0.18138681 0.8186132
-## speak   8.003838e-04 0.0004728416 0.62862694 0.3713731
-## to      1.113210e-02 0.0075761991 0.59503541 0.4049646
-## soon    2.642059e-04 0.0010608467 0.19939274 0.8006073
-## today   1.120666e-03 0.0019041688 0.37048833 0.6295117
-## is      4.451802e-03 0.0058153446 0.43359683 0.5664032
-## accept  3.802754e-05 0.0003188419 0.10655871 0.8934413
-
-The model outputs the Probabilities of the message being Spam or ham.
-Let’s Test the Model
-
-pred<-predict(nb.classifier,msg.dfm.test)
+```{r}
+predGroup <- predict(nb.classifier,msg.dfm.test)
+```
 
@@
 # use pred$nb.predicted to extract the class labels
-table(predicted=pred$nb.predicted,actual=spam.test[,1])
+```{r}
+table(predicted=predGroup[["names"]],actual=text.test[,1])
+mean(predGroup$names==text.test[,1])*100
+
+```
+
 ##       actual
@@
 #acccuracy of the classifier on Test data
-mean(pred$nb.predicted==spam.test[,1])*100
 ## [1] 97.93722

### 2019-02-21 docs/hiding/Naive_bayes_V2r_2a.Rmd
update

--- 2019-02-21 hiding/Naive_bayes_V2r.Rmd
@@
 author: "Dan Johnson"
-date: "1/29/2019"
+date: "2/14/2019"
 output: html_document
@@
 ##Summary
-This is **Version 2.0.r** of the basic workflow for classifying TA comments using a naive Bayes model.
+This is **Version 2r.2a** of the basic workflow for classifying TA comments using a naive Bayes model.
 It pre-processes text using the same command set as Ver 2.0.b.
 The "r" designates this version has been modified to be able to randomize which rows are selected for training and test datasets.
+
+This updated version refines the final output tables so they are more informative.
 
@@
 library(tidyverse)
+library(tm)
 library(tidytext)
-library(tidyr)
-library(tm)
-library(SnowballC)
-library(wordcloud)
 library(e1071)
-library(gmodels)
+
+#library(RTextTools)
+#library(caret)
+#library(doMC)
+#registerDoMC(core=detectCores())
+#library(tidyr)
+#library(dplyr)
+#library(gmodels)
 ```
@@
 vector <- 1:9340
-set.seed(123) #change this seed value, and the data are randomized in a new sequence.
+set.seed(567) #change this seed value, and the data are randomized in a new sequence.
 vector <- sample(vector)
@@
 #This step removes evidence of questions, so may remove data.
+#THIS ALSO REMOVES EVIDENCE OF POINTER ITEMS
 comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removePunctuation)
@@
 comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
+comments_dtm_all <- comments_dtm_1gram[1:9340, ]
 ```
@@
 comments_test_labels_subject <- comments_rawVERB_R[7006:9340,]$subject
+comments_all_labels_subject <- comments_rawVERB_R[1:9340,]$subject
 ```
@@
 comments_test_labels_structure <- comments_rawVERB_R[7006:9340,]$structure
+comments_all_labels_structure <- comments_rawVERB_R[1:9340,]$structure
 ```
@@
 prop.table(table(comments_test_labels_subject))
+prop.table(table(comments_all_labels_subject))
 ```
@@
 prop.table(table(comments_test_labels_structure))
+prop.table(table(comments_all_labels_structure))
 ```
@@
 comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
+comments_dtm_freq_all <- comments_dtm_all[ , comments_freq_words]
 ```
@@
 }
-
-comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
-comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
 ```
@@
 The resulting matrices have cells indicating "yes" or "no" if the word represented by the column appears in the text comment represented by the row.
+
+```{}
+# Function to KEEP word frequencies while still using the same variable names and code structure. The resulting matrices have cells with word frequencies instead.
+convert_counts2 <- function(x) {
+  y <- x
+}
+```
+
+```{r}
+comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
+comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
+comments_all <- apply(comments_dtm_freq_all, MARGIN = 2, convert_counts2)
+```
 
@@
 ```{r}
-comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject)
+comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject, laplace=1)
 comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
-CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
+comments_all_pred_subject <- predict(comments_classifier_subject, comments_all)
+
+# Create a truth table by tabulating the predicted class labels with the actual class labels 
+table("Predictions"= comments_test_pred_subject,  "Actual" = comments_test_labels_subject )
+
+table("Predictions"= comments_all_pred_subject,  "Actual" = comments_all_labels_subject )
+
+#table("Predictions2"= comments_test_pred_structure,  "Actual" = comments_test_labels_structure )
+
+#The truth table lists how many of the predicted categories are in their correct categories.
+Adding diagonally and dividing by total count gives the percent accuracy at predicting subject.
+this is what the confusion matrix calculates.
+The faster way is to calculate a confusion matrix, which includes BOTH the truth table, and the final overall accuracy.
+
+#The Cross Table command below (from "gmodels") generates a more detailed output table with frequency values for each individual cell.
+#CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
 ```
 
-The cross table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject.
 ```{r}
-print("Overall subject accuracy")
-(43+473+1131+184)/23.35  #Need to find a way to calculate this automatically. 
+#"line" is just a spacer text for readability.
+line<-("**************************************************************")
+
+# Prepare the confusion matrix. 
+conf.mat <- confusionMatrix(comments_test_pred_subject, comments_test_labels_subject)
+conf.mat2 <- confusionMatrix(comments_all_pred_subject, comments_all_labels_subject)
+
+# This print-out has both the overall and by-group accuracy values.
+conf.mat2
+print (line)
+
+# This print-out has by-group accuracy values in a single column. 
+conf.mat2$byClass
+print (line)
+
+# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
+conf.mat2$overall
+print (line)
+
+# This outputs JUST the overall accuracy value. 
+conf.mat2$overall['Accuracy']
+
+# This print-out has both the overall and by-group accuracy values.
+conf.mat
+print (line)
+
+# This print-out has by-group accuracy values in a single column. 
+conf.mat$byClass
+print (line)
+
+# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
+conf.mat$overall
+print (line)
+
+# This outputs JUST the overall accuracy value. 
+conf.mat$overall['Accuracy']
+
 ```
@@
 
-```{r}
-comments_classifier_structure <- naiveBayes(comments_train, comments_train_labels_structure)
+```{}
+comments_classifier_structure <- naiveBayes(comments_train, comments_train_labels_structure, laplace=1)
 comments_test_pred_structure <- predict(comments_classifier_structure, comments_test)
@@
 
-Again, a cross table shows how many of the predicted categories match the correct categories for structure.
+###OPTIONAL Output
+Make a table showing original comments, predicted category for comment based on the Naive Bayes test, and the actual label entered by a human observer.
+Export to a CSV file, with the randomizer value in the name of the file.
+By default the CSV file should be saved to the current working directory.
 
 ```{r}
-print("Overall structure accuracy")
-(22+488+292+518+33+10)/23.35  #Need to find a way to calculate this automatically. 
+#Creates a data frame that compares predicted and actual labels for subject and structure
+tallyall <- data.frame("text"=c(comments_rawVERB_R[1:9340,]$text),"subject_label"=c(comments_all_labels_subject), "subject_predicted"=c(comments_all_pred_subject))
+
+#Write a CSV file out with data from this run.
+write.csv(tallyall, file = "SubjectsAll_Randomizer_567.csv")
+
 ```
 
-###OPTIONAL Output
-To make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame.
+```{r}
+#Creates a data frame that compares predicted and actual labels for subject and structure
+tally <- data.frame("text"=c(comments_rawVERB_R[7006:9340,]$text),"subject_label"=c(comments_test_labels_subject), "subject_predicted"=c(comments_test_pred_subject))
+
+#Write a CSV file out with data from this run.
+write.csv(tally, file = "Subjects_Randomizer_567.csv")
+
+```
+
+***
+START COMMENTS, OLD work notes
+***
+
+Make a table showing predicted categories for comments based on the Naive Bayes test, versus the actual label entered by a human observer, merge the two datasets into a single data frame.
 The full data frame will be too long to display; use "head" to check format, then output the full dataset to CSV for further review.
@@
 
-***
-START COMMENTS, OLD work notes
-***
-
-```{r}
-#Creates a data frame that compares predicted and actual labels for subject and structure
-tally <- data.frame("subject_label"=c(comments_test_labels_subject), "subject_predicted"=c(comments_test_pred_subject), "structure_label"=c(comments_test_labels_structure), "structure_predicted"=c(comments_test_pred_structure))
-#tally <- as.numeric(as.character(tally))
-```
-
 ```{r}

### 2019-02-21 docs/hiding/Naive_bayes_V2r_2b.Rmd
update

--- 2019-02-21 docs/hiding/Naive_bayes_V2r_2a.Rmd
@@
 ---
-title: "Naive Bayes Workflow Ver 2.0.r"
+title: "Naive Bayes Workflow Ver 2r.2b"
 author: "Dan Johnson"
@@
 ##Summary
-This is **Version 2r.2a** of the basic workflow for classifying TA comments using a naive Bayes model.
+This is **Version 2r.2b** of the basic workflow for classifying TA comments using a naive Bayes model.
 It pre-processes text using the same command set as Ver 2.0.b.
 The "r" designates this version has been modified to be able to randomize which rows are selected for training and test datasets.
+This updated version refines the final output tables so they are more informative.
 
-This updated version refines the final output tables so they are more informative.
+Changes since 2a: this time I am trying out what happens if I go to ingroup/outgroup only, then run the ENTIRE dataset. I want to know how far off of my hand-coded table the RATIOS of effort are.
 
@@
 
-##Version 2.0.r of Naive Bayes (NB) Protocol
+##Version 2.r.2b of Naive Bayes (NB) Protocol
+
 ###Initial Data Input
@@
 #Read in TA comments from CSV file.
-base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
+base_dataVerYN <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
 ```
@@
 #Select rows representing the sub-groups to compare. 
-comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")
+comments_subsetVerYN <- filter(base_dataVerYN,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")
 
-#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
-comments_rawVERB <- comments_subsetVERB %>% select(23,24,22)
+#Reduce larger dataframe to 3 required columns of data, and put columns in order needed.
+comments_rawVerYN <- comments_subsetVerYN %>% select(23,24,22)
 
 #Rename the columns.
-names(comments_rawVERB)[1] <- "subject"
-names(comments_rawVERB)[2] <- "structure"
-names(comments_rawVERB)[3] <- "text"
+names(comments_rawVerYN)[1] <- "subject"
+names(comments_rawVerYN)[2] <- "structure"
+names(comments_rawVerYN)[3] <- "text"
 
-#Change "subject" element from character to a factor for analysis.
-comments_rawVERB$subject <- factor(comments_rawVERB$subject)
-str(comments_rawVERB$subject)
-table(comments_rawVERB$subject)
+#This block uses 'forcats' to recode the Subject and Structure factors as either ingroup or outgroup.
+#Changing this block is how I change and use other Subject and Structure factors as the ingroup.
+#Looking ahead, could use this block and a variable name to simplify coding (NO, does not work).
+
+comments_rawVerYN_V <- comments_rawVerYN %>% 
+  mutate(subject = fct_recode(subject, ingroup = "3_technical", outgroup="1_basic", outgroup="2_writing", outgroup="4_logic"))
+
+#%>% mutate(structure = fct_recode(structure, ingroup = "4_specific", outgroup="1_pointer",outgroup="2_copyedit",outgroup="3_general",outgroup="6_holistic",outgroup="7_idiomatic",outgroup="8_nobasis" ))
 ```
@@
 ```{r}
+#Change "subject" element from character to a factor for analysis.
+comments_rawVerYN_V$subject <- factor(comments_rawVerYN_V$subject)
+str(comments_rawVerYN_V$subject)
+table(comments_rawVerYN_V$subject)
+```
+
+```{}
 #Change "structure" element from character to a factor for analysis.
-comments_rawVERB$structure <- factor(comments_rawVERB$structure)
-str(comments_rawVERB$structure)
-table(comments_rawVERB$structure)
+comments_rawVerYN_V$structure <- factor(comments_rawVerYN_V$structure)
+str(comments_rawVerYN_V$structure)
+table(comments_rawVerYN_V$structure)
 ```
@@
 vector <- 1:9340
-set.seed(567) #change this seed value, and the data are randomized in a new sequence.
+set.seed(123) #change this seed value, and the data are randomized in a new sequence.
 vector <- sample(vector)
@@
 ```{r}
-#Append the vector as a new column to "comments_rawVERB" with name "randomizer"
+#Append the vector as a new column to "comments_rawVerYN" with name "randomizer"
 #Sort rows of dataframe according to "randomizer" column using dplyer 'arrange' function. 
-comments_rawVERB$randomizer <- c(vector)
-comments_rawVERB_R <-comments_rawVERB %>% arrange(randomizer)
-print(comments_rawVERB_R)
+comments_rawVerYN_V$randomizer <- c(vector)
+comments_rawVerYN_R <-comments_rawVerYN_V %>% arrange(randomizer)
+print(comments_rawVerYN_R)
 
@@
 #This step removes evidence of questions, so may remove data.
-#THIS ALSO REMOVES EVIDENCE OF POINTER ITEMS
-comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removePunctuation)
+comments_corpus_cleanVerYN_R <- tm_map(comments_corpus_cleanVerYN_R, removePunctuation)
 
@@
 #I am not sure this is beneficial; does it remove evidence indicating technical comments?
-comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeNumbers)
+#comments_corpus_cleanVerYN_R <- tm_map(comments_corpus_cleanVerYN_R, removeNumbers)
 
@@
 
-comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB_R, control=list(tokenize = NLP_Tokenizer))
+comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVerYN_R, control=list(tokenize = NLP_Tokenizer))
 comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
@@
 comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
-comments_dtm_all <- comments_dtm_1gram[1:9340, ]
 ```
@@
 ```{r}
-comments_train_labels_subject <- comments_rawVERB_R[1:7005, ]$subject
-comments_test_labels_subject <- comments_rawVERB_R[7006:9340,]$subject
-comments_all_labels_subject <- comments_rawVERB_R[1:9340,]$subject
+comments_train_labels_subject <- comments_rawVerYN_R[1:7005, ]$subject
+comments_test_labels_subject <- comments_rawVerYN_R[7006:9340,]$subject
 ```
 
-```{r}
-comments_train_labels_structure <- comments_rawVERB_R[1:7005, ]$structure
-comments_test_labels_structure <- comments_rawVERB_R[7006:9340,]$structure
-comments_all_labels_structure <- comments_rawVERB_R[1:9340,]$structure
+```{}}
+comments_train_labels_structure <- comments_rawVerYN_R[1:7005, ]$structure
+comments_test_labels_structure <- comments_rawVerYN_R[7006:9340,]$structure
 ```
@@
 prop.table(table(comments_test_labels_subject))
-prop.table(table(comments_all_labels_subject))
 ```
 
-```{r}
+```{}
 prop.table(table(comments_train_labels_structure))
 prop.table(table(comments_test_labels_structure))
-prop.table(table(comments_all_labels_structure))
 ```
@@
 comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
-comments_dtm_freq_all <- comments_dtm_all[ , comments_freq_words]
 ```
@@
 This replaces values greater than 0 with yes, and values not greater than 0 with no.
+The resulting matrices have cells indicating "yes" or "no" if the word represented by the column appears in the text comment represented by the row.
+
 ```{r}
@@
 }
+
+comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
+comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
 ```
 
-The resulting matrices have cells indicating "yes" or "no" if the word represented by the column appears in the text comment represented by the row.
+The alternate version below uses the original word frequencies without binarization.
 
@@
 }
-```
 
-```{r}
 comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
 comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
-comments_all <- apply(comments_dtm_freq_all, MARGIN = 2, convert_counts2)
 ```
@@
 comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
-comments_all_pred_subject <- predict(comments_classifier_subject, comments_all)
 
@@
 table("Predictions"= comments_test_pred_subject,  "Actual" = comments_test_labels_subject )
-
-table("Predictions"= comments_all_pred_subject,  "Actual" = comments_all_labels_subject )
-
 #table("Predictions2"= comments_test_pred_structure,  "Actual" = comments_test_labels_structure )
@@
 conf.mat <- confusionMatrix(comments_test_pred_subject, comments_test_labels_subject)
-conf.mat2 <- confusionMatrix(comments_all_pred_subject, comments_all_labels_subject)
-
-# This print-out has both the overall and by-group accuracy values.
-conf.mat2
-print (line)
-
-# This print-out has by-group accuracy values in a single column. 
-conf.mat2$byClass
-print (line)
-
-# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
-conf.mat2$overall
-print (line)
-
-# This outputs JUST the overall accuracy value. 
-conf.mat2$overall['Accuracy']
 
@@
 #Creates a data frame that compares predicted and actual labels for subject and structure
-tallyall <- data.frame("text"=c(comments_rawVERB_R[1:9340,]$text),"subject_label"=c(comments_all_labels_subject), "subject_predicted"=c(comments_all_pred_subject))
+tally <- data.frame("text"=c(comments_rawVerYN_R[7006:9340,]$text),"subject_label"=c(comments_test_labels_subject), "subject_predicted"=c(comments_test_pred_subject))
 
 #Write a CSV file out with data from this run.
-write.csv(tallyall, file = "SubjectsAll_Randomizer_567.csv")
-
-```
-
-```{r}
-#Creates a data frame that compares predicted and actual labels for subject and structure
-tally <- data.frame("text"=c(comments_rawVERB_R[7006:9340,]$text),"subject_label"=c(comments_test_labels_subject), "subject_predicted"=c(comments_test_pred_subject))
-
-#Write a CSV file out with data from this run.
-write.csv(tally, file = "Subjects_Randomizer_567.csv")
+write.csv(tally, file = "Subjects_InOut_Randomizer_123.csv")
 

### 2019-02-21 docs/hiding/Naive_bayes_V2r_2c.Rmd
update

---
title: "Naive Bayes Analysis Ver 2r.2c"
author: "Dan Johnson"
date: "2/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Summary
This workflow uses Naive Bayes to classify TA comments based on Subject and Structure. The major additions and revisions in **Version 2r.2c** are:

*  Randomly selected data rows used for training and test datasets. 
*  Final output matrix tables are more informative. 
*  Addition of a back-analysis of entire dataset, not just testing sub-set.
*  Final output CSV file is a tidy data table of the original comments, the hand-scored and predicted subject, and hand-scored and predicted structure.

####\

##Initial Setup
Calling the required libraries.

```{r libraries}
library(tidyverse)
library(tm)
library(tidytext)
library(e1071)
library(caret)

#library(RTextTools)
#library(doMC)
#registerDoMC(core=detectCores())
#library(tidyr)
#library(dplyr)
#library(gmodels)
```

####\

##Version 2.0.r of Naive Bayes (NB) Protocol
###Initial Data Input
```{r data_input}
#Read in TA comments from CSV file.
base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

###Extract, Organize Data Subset
Extract comments and classification data as subsets of CSV.
```{r data_subsetting}
#Select rows representing the sub-groups to compare. 
comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
comments_rawVERB <- comments_subsetVERB %>% select(23,24,22)

#Rename the columns.
names(comments_rawVERB)[1] <- "subject"
names(comments_rawVERB)[2] <- "structure"
names(comments_rawVERB)[3] <- "text"

#Change "subject" element from character to a factor for analysis.
comments_rawVERB$subject <- factor(comments_rawVERB$subject)
str(comments_rawVERB$subject)
table(comments_rawVERB$subject)

#Change "structure" element from character to a factor for analysis.
comments_rawVERB$structure <- factor(comments_rawVERB$structure)
str(comments_rawVERB$structure)
table(comments_rawVERB$structure)
```

```{r randomizer}
#Generates a reproducible vector of ~10,000 randomly assorted numbers using set.seed 
#Adjust the number so it matches the number of rows in the original data frame
#Next, set the seed for random numbers, and randomize the order of the vector
vector <- 1:9340 #second value must match number of rows in working dataset
set.seed(123) #change this seed value, and the data are randomized in a new sequence.
vector <- sample(vector)
glimpse(vector)

#Append the vector as a new column to "comments_rawVERB" with name "randomizer"
#Sort rows of dataframe according to "randomizer" column using dplyer 'arrange' function. 
comments_rawVERB$randomizer <- c(vector)
comments_rawVERB_R <-comments_rawVERB %>% arrange(randomizer)
print(comments_rawVERB_R)
```

```{r corpusmaker}
#Convert data set to a volative corpus using "tm" library.
#Create the volatile corpus that contains the "text" vector from data frame.
comments_corpusVERB_R <- VCorpus(VectorSource(comments_rawVERB_R$text))
print(comments_corpusVERB_R)

#Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
inspect(comments_corpusVERB_R[1:3])

#Use "as.character" function to see what a single text comment looks like.
as.character(comments_corpusVERB_R[[5]])
```

###Text Data Transformations
The text data transforms are separated out as individual "tm_map"" and "content_transformer" functions to simplify the process of testing various combinations. Conversion to lower case and removing punctuation are the baseline set. Switch between "{r}" and "{}" to turn the other transformation blocks off or on. Removing extra white spaces must be the last transformation, and is required.

```{r textcleanup}
#Convert text to all lower case letters.
comments_corpus_cleanVERB_R <- tm_map(comments_corpusVERB_R, content_transformer(tolower))

#Remove punctuation using the "removePunctuation" function. 
#This step removes evidence of questions, so may remove data.
#THIS ALSO REMOVES EVIDENCE OF POINTER ITEMS
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removePunctuation)

#Remove numerals. 
#This step may remove evidence indicating technical comments
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeNumbers)

#Stopword removal. 
#This is a standard cleanup step, but again may remove useful terms.
#The "stopwords" file is a convention; other files can be substituted for it
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeWords, stopwords())

#The option to apply stemming has been removed from this version.

#Final step removes white space from the document. This is NOT an optional step.
comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, stripWhitespace)

#Look at an example of cleaned text comment to see if it fits what is expected.
as.character((comments_corpus_cleanVERB_R[[5]]))
```

####\

##Basic Naive Bayes (NB) Method Using N-grams
This code block uses tm's NLP commands. The same block can be used to assess 1-, 2-, or 3-grams by modifying the "#" variable in "(words(x), #)".

```{r tokenizer}
NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
}

comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB_R, control=list(tokenize = NLP_Tokenizer))
comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
inspect(comments_dtm_1gram)
```

####\

###Data Splitting
Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data.

```{}
#This is for convenience. It is not required if the number of training, test rows is known.
.75 * 9340 #Number of rows in the dataset; product is #rows for training
.25 * 9340 #Product is #rows for testing set
```

This code divides the previously randomized comments into training, testing, and re-analysis subsets of data.
```{r divider}
comments_dtm_all <- comments_dtm_1gram[1:9340, ]
comments_dtm_train <- comments_dtm_1gram[1:7005, ]
comments_dtm_test <- comments_dtm_1gram[7006:9340, ]
```

Saves 2 vectors that label the rows in the training and testing vectors
```{r labeler}
comments_all_labels_subject <- comments_rawVERB_R[1:9340,]$subject
comments_train_labels_subject <- comments_rawVERB_R[1:7005, ]$subject
comments_test_labels_subject <- comments_rawVERB_R[7006:9340,]$subject

comments_all_labels_structure <- comments_rawVERB_R[1:9340,]$structure
comments_train_labels_structure <- comments_rawVERB_R[1:7005, ]$structure
comments_test_labels_structure <- comments_rawVERB_R[7006:9340,]$structure
```

The proportion of each sub-category should be similar (<1% difference) between the full dataset, and training and testing data subsets.
```{r proportion_check}
prop.table(table(comments_all_labels_subject))
prop.table(table(comments_train_labels_subject))
prop.table(table(comments_test_labels_subject))

prop.table(table(comments_all_labels_structure))
prop.table(table(comments_train_labels_structure))
prop.table(table(comments_test_labels_structure))
```

####\

###Preparation for Naive Bayes
Remove words from the matrix that appear less than 5 times.
```{r common_words}
comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
str(comments_freq_words)
```

Limit Document Term Matrix to only include words in the comments_freq_words vector. Using all the rows, but we want to limit the columns to these words in the frequency vector.
```{r limit_dtm}
comments_dtm_freq_all <- comments_dtm_all[ , comments_freq_words]
comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
```

The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables. This is done using a **convert_counts function** and applying it to the data. This replaces values greater than 0 with yes, and values not greater than 0 with no. 
```{r convert_function}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}
```

The resulting matrices have cells indicating "yes" or "no" if the word represented by the column appears in the text comment represented by the row.

```{}
# Function to KEEP word frequencies while still using the same variable names and code structure. The resulting matrices have cells with word frequencies instead.
convert_counts2 <- function(x) {
  y <- x
}
```

```{r make_binary}
comments_all <- apply(comments_dtm_freq_all, MARGIN = 2, convert_counts2)
comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
```

####\

###Train Model, Predict, Evaluate for Subject
This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. It evaluates the prediction with the actual data using a crosstable from the gmodels package.
```{r bayes_training}
comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject, laplace=1)
comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
comments_all_pred_subject <- predict(comments_classifier_subject, comments_all)

comments_classifier_structure <- naiveBayes(comments_train, comments_train_labels_structure, laplace=1)
comments_test_pred_structure <- predict(comments_classifier_structure, comments_test)
comments_all_pred_structure <- predict(comments_classifier_structure, comments_all)

# Create a truth table by tabulating the predicted class labels with the actual class labels 
table("Predictions"= comments_test_pred_subject,  "Actual Subject Test" = comments_test_labels_subject )
table("Predictions"= comments_all_pred_subject,  "Actual Subject All" = comments_all_labels_subject )
table("Predictions"= comments_test_pred_structure,  "Actual Structure Test" = comments_test_labels_structure )
table("Predictions"= comments_all_pred_structure,  "Actual Structure All" = comments_all_labels_structure )

#The truth table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting subject. this is what the confusion matrix calculates. The faster way is to calculate a confusion matrix, which includes BOTH the truth table, and the final overall accuracy.

#The Cross Table command below (from "gmodels") generates a more detailed output table with frequency values for each individual cell.
#CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

```{r reportOut}
#"line" is just a spacer text for readability.
line<-("**************************************************************")

# Prepare the confusion matrix. 
conf.mat1 <- confusionMatrix(comments_test_pred_subject, comments_test_labels_subject)
conf.mat2 <- confusionMatrix(comments_all_pred_subject, comments_all_labels_subject)
conf.mat3 <- confusionMatrix(comments_test_pred_structure, comments_test_labels_structure)
conf.mat4 <- confusionMatrix(comments_all_pred_structure, comments_all_labels_structure)

# This print-out has both the overall and by-group accuracy values.
conf.mat1
print (line)
print ("Test, Subject")
# This print-out has by-group accuracy values in a single column. 
conf.mat1$byClass
print (line)
print ("Test, Subject")

# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
conf.mat1$overall
print (line)
print ("Test, Subject")

# This outputs JUST the overall accuracy value. 
conf.mat1$overall['Accuracy']
print ("Test, Subject")

# This print-out has both the overall and by-group accuracy values.
conf.mat2
print (line)
print ("All, Subject")

# This print-out has by-group accuracy values in a single column. 
conf.mat2$byClass
print (line)
print ("All, Subject")

# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
conf.mat2$overall
print (line)
print ("All, Subject")

# This outputs JUST the overall accuracy value. 
conf.mat2$overall['Accuracy']
print ("All, Subject")

# This print-out has both the overall and by-group accuracy values.
conf.mat3
print (line)
print ("Test, Structure")

# This print-out has by-group accuracy values in a single column. 
conf.mat3$byClass
print (line)
print ("Test, Structure")

# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
conf.mat3$overall
print (line)
print ("Test, Structure")

# This outputs JUST the overall accuracy value. 
conf.mat3$overall['Accuracy']
print ("Test, Structure")

# This print-out has both the overall and by-group accuracy values.
conf.mat4
print (line)
print ("All, Structure")

# This print-out has by-group accuracy values in a single column. 
conf.mat4$byClass
print (line)
print ("All, Structure")

# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
conf.mat4$overall
print (line)
print ("All, Structure")

# This outputs JUST the overall accuracy value. 
print ("All, Structure")
conf.mat4$overall['Accuracy']

```

####\

###Output for Review, Analysis
The code block below creates a CSV file containing the original TA comments, the subject and structure labels assigned by a human observer, and the corresponding predicted categories for each comment based on the Naive Bayes estimator. Results are exported to a CSV file. By default the CSV file should be saved to the current working directory.

```{r writeCSV}
#Creates a data frame that compares predicted and actual labels for subject and structure
tallymatrix <- data.frame("text"=c(comments_rawVERB_R[1:9340,]$text),"subject_label"=c(comments_all_labels_subject), "subject_predicted"=c(comments_all_pred_subject),"structure_label"=c(comments_all_labels_structure), "structure_predicted"=c(comments_all_pred_structure))

#Creates a summary counts table that compares hand-coded counts to Bayes-estimated counts.  
final.tally <- table("Hand Coded Subject"=tallymatrix$subject_label, "Hand Coded Structure"=tallymatrix$structure_label)
final.tally2 <- table("Bayes Predicted Subject"=tallymatrix$subject_predicted, "Bayes Predicted Structure"=tallymatrix$structure_predicted)

#Write out the final data from the run to a CSV file. For reference, include the randomizer value used in the name of the file.
write.csv(tallymatrix, file = "SandS_Matrix_Randomizer_123.csv")

#Print to console the hand coded versus estimated counts as a 2-way contingency table of subject vs. structure.
final.tally
prop_tally<-prop.table(final.tally)
print(prop_tally)
print(line)
print(line)
final.tally2
prop_tally2<-prop.table(final.tally2)
print(prop_tally2)
```

***
***

This is an alternate method to get the same data tables out using ```tidy``` format. It seems much longer to me, so likely is over-written.
```{r tidy_prediction_table}
#Use broom::tidy to reorganize into a data frame.
tidy_tally1 <-tidy(final.tally)
tidy_tally2 <-tidy(final.tally2)
tidy_tally3 <-tidy(final.tally)
tidy_tally4 <-tidy(final.tally2)

#Use mutate to add a proportion of total column. And yes, it calculates based on subgroups rather than whole column.
prop_tally1<-mutate(tidy_tally1, prop = n/sum(n))
prop_tally1<-select(prop_tally1, 1,2,4)
prop_tally2<-mutate(tidy_tally2, prop = n/sum(n))
prop_tally2<-select(prop_tally2, 1,2,4)
prop_tally3<-mutate(tidy_tally3, prop = n/sum(n))
prop_tally3<-select(prop_tally3, 1,2,4)
prop_tally4<-mutate(tidy_tally4, prop = n/sum(n))
prop_tally4<-select(prop_tally4, 1,2,4)

#Spread tidy data into a readable table
spread(prop_tally1,key=Hand.Coded.Structure, value=prop)
spread(prop_tally2,key=Bayes.Predicted.Structure, value=prop)
spread(prop_tally3,key=Hand.Coded.Subject, value=prop)
spread(prop_tally4,key=Bayes.Predicted.Subject, value=prop)
```

```{r}
#Calculate the sum of values for each group.
proptallysum1<-prop_tally1 %>% group_by(Hand.Coded.Subject) %>% summarize(total=sum(prop))
proptallysum1
proptallysum2<-prop_tally2 %>% group_by(Bayes.Predicted.Subject) %>% summarize(total=sum(prop))
proptallysum2
proptallysum3<-prop_tally3 %>% group_by(Hand.Coded.Structure) %>% summarize(total=sum(prop))
proptallysum3
proptallysum4<-prop_tally4 %>% group_by(Bayes.Predicted.Structure) %>% summarize(total=sum(prop))
proptallysum4

```

##Graph Outputs
Plotting Hand-coding vs. Bayes

Need to combine two datasets into one, and create subsets based on the

```{r}
tidy(proptallysum1)
ggplot(data=proptallysum1, aes(x=Hand.Coded.Subject, y=total, group=1))+
  geom_line() +
  geom_point()

ggplot(data=proptallysum2, aes(x=Bayes.Predicted.Subject, y=total, group=1))+
  geom_line() +
  geom_point()
 ```

### 2019-02-21 docs/hiding/Naive_bayes_V2yn.Rmd
update

--- 2019-01-31 Naive_bayes_V2yn.Rmd
@@
 library(tm)
-library(SnowballC)
-library(wordcloud)
 library(e1071)
@@
 
-##Version 2.0.r of Naive Bayes (NB) Protocol
+##Version 2.0.YN of Naive Bayes (NB) Protocol
 ###Initial Data Input
@@
 #Read in TA comments from CSV file.
-base_dataVerYN <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVerYN.csv')
+base_dataVerYN <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
 ```
@@
 
-#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
+#Reduce larger dataframe to 3 required columns of data, and put columns in order needed.
 comments_rawVerYN <- comments_subsetVerYN %>% select(23,24,22)
@@
 
+#This block uses 'forcats' to recode the Subject and Structure factors as either ingroup or outgroup.
+#Changing this block is how I change and use other Subject and Structure factors as the ingroup.
+#Looking ahead, could use this block and a variable name to simplify coding (NO, does not work).
+
+comments_rawVerYN_V <- comments_rawVerYN %>% 
+  mutate(subject = fct_recode(subject, ingroup = "3_technical", outgroup="1_basic", outgroup="2_writing", outgroup="4_logic")) %>% 
+  mutate(structure = fct_recode(structure, ingroup = "4_specific", outgroup="1_pointer",outgroup="2_copyedit",outgroup="3_general",outgroup="6_holistic",outgroup="7_idiomatic",outgroup="8_nobasis" ))
+```
+
+***
+POINT OF CONCERN ABOUT CODING
+***
+Doing multiple comments for a larger group like Structure is tedious. This block of code:
+```
+comments_rawVerYN_V <- comments_rawVerYN %>% 
+    mutate(structure = fct_recode(structure, ingroup = "4_specific", outgroup="1_pointer",outgroup="2_copyedit",outgroup="3_general",outgroup="6_holistic",outgroup="7_idiomatic",outgroup="8_nobasis" ))
+```
+Has two issues. 
+1.
+Each item being recoded as outgroup has to be typed in individually.
+If I try to enter multiple items in a list (either by embedding the list in a variable, or by concatenating the items to all be recoded as 'outgroup'), I get an error message saying you cannot apply fct_recode to more than one pair of items.
+Seems to be a lot of code to write to recode several variables to the same name.
+
+2. This block was suggested, but is for writing the summary list of who gets COUNTED HOW; it does not affect the actual coding. 
+```
+partyid2 <- fct_collapse(gss_cat$partyid,
+  missing = c("No answer", "Don't know"),
+  other = "Other party",
+  rep = c("Strong republican", "Not str republican"),
+  ind = c("Ind,near rep", "Independent", "Ind,near dem"),
+  dem = c("Not str democrat", "Strong democrat")
+)
+```
+I do not want to collapse the COUNTS values together, I want to recode multiple variables simultaneously.
+I want to choose ALL of the NON-INGROUP items with one command and not add them individually.
+This is pseudo-code describing what I want to do:
+
+```{}
+comments_rawVerYN %>%
+  mutate(subject = fct_recode(subject, ingroup = "3_technical", outgroup= ANYTHING_BUT "3_technical"))
+```
+
+3. Other suggestions I found:
+
+```
+https://community.rstudio.com/t/quickly-recoding-large-ish-vectors-in-a-tibble/2447
+
+# object to decode and the key
+
+tbl <- tibble(CODE = sample(letters[1:3], 1e+06, replace = TRUE))
+
+key <- tribble(~CODE,   ~FRUIT, 
+                 "a",  "apple", 
+                 "b", "banana", 
+                 "c",  "cherry"
+               )
+
+# Alternative One: dplyr::recode (what I used above basically)
+
+mutate(tbl, FRUIT = recode(CODE, a = "apple", b = "banana", c = "cherry")) 
+#> # A tibble: 1,000,000 x 2
+#>     CODE  FRUIT
+#>    <chr>  <chr>
+#>  1     c cherry
+#>  2     b banana
+#>  3     a  apple
+#>  4     b banana
+#>  5     c cherry
+#>  6     c cherry
+#>  7     b banana
+#>  8     c cherry
+#>  9     a  apple
+#> 10     a  apple
+#> # ... with 999,990 more rows
+
+# Alternative Two: named vector
+
+# create a named vector from the key object and use it to recode
+named_vector <- 
+    key %>% 
+    pmap(~set_names(..2, ..1)) %>% 
+    unlist()
+
+mutate(tbl, FRUIT = named_vector[CODE])
+#> # A tibble: 1,000,000 x 2
+#>     CODE  FRUIT
+#>    <chr>  <chr>
+#>  1     c cherry
+#>  2     b banana
+#>  3     a  apple
+#>  4     b banana
+#>  5     c cherry
+#>  6     c cherry
+#>  7     b banana
+#>  8     c cherry
+#>  9     a  apple
+#> 10     a  apple
+#> # ... with 999,990 more rows
+
+left_join(tibblename, key)
+
+benchmark(left_join(tbl, key, by = "CODE"), 
+          columns = c("replications", "elapsed", "relative", "user.self", "sys.self"))
+
+tbl <- tibble(CODE = sample(letters[1:3], 1e+06, replace = TRUE))
+
+rev <- tbl %>%
+  mutate(type =
+           case_when(
+             CODE == "a" ~ "apple",
+             CODE == "b" ~ "banana",
+             CODE == "c" ~ "cherry",
+             FALSE ~ "no fruit"
+           )
+)
+
+tbl <- tibble(CODE = sample(letters[1:3], 1e+06, replace = TRUE))
+
+key <- tribble(~CODE,   ~FRUIT, 
+               "a",  "apple", 
+               "b", "banana", 
+               "c",  "cherry"
+)
+
+tbl_with_factor <- mutate(tbl, CODE = as.factor(CODE))
+key_with_factor <- mutate(key, CODE = as.factor(CODE))
+
+```
+
+***
+
+```{r}
 #Change "subject" element from character to a factor for analysis.
-comments_rawVerYN$subject <- factor(comments_rawVerYN$subject)
-str(comments_rawVerYN$subject)
-table(comments_rawVerYN$subject)
+comments_rawVerYN_V$subject <- factor(comments_rawVerYN_V$subject)
+str(comments_rawVerYN_V$subject)
+table(comments_rawVerYN_V$subject)
 ```
@@
 #Change "structure" element from character to a factor for analysis.
-comments_rawVerYN$structure <- factor(comments_rawVerYN$structure)
-str(comments_rawVerYN$structure)
-table(comments_rawVerYN$structure)
+comments_rawVerYN_V$structure <- factor(comments_rawVerYN_V$structure)
+str(comments_rawVerYN_V$structure)
+table(comments_rawVerYN_V$structure)
 ```
@@
 #Sort rows of dataframe according to "randomizer" column using dplyer 'arrange' function. 
-comments_rawVerYN$randomizer <- c(vector)
-comments_rawVerYN_R <-comments_rawVerYN %>% arrange(randomizer)
+comments_rawVerYN_V$randomizer <- c(vector)
+comments_rawVerYN_R <-comments_rawVerYN_V %>% arrange(randomizer)
 print(comments_rawVerYN_R)
@@
 print("Overall subject accuracy")
-(14+54+1318+28)/23.35  #Need to find a way to calculate this automatically. 
+(772+1174)/23.35  #Need to find a way to calculate this automatically. 
 ```
@@
 print("Overall structure accuracy")
-(66+64+949+5+1)/23.35  #Need to find a way to calculate this automatically. 
+(1117+592)/23.35  #Need to find a way to calculate this automatically. 
 ```
@@
 ***
-
-##Permutation Tests of Naive Bayes (NB) Method
-Using single words, Version 2.0.b correctly identifies comment "subject" ~80% of the time and comment "structure" ~59% of the time.
-
-Permutations to test with random subsets version (2.0.r):
-
-*  Compare 1-grams, 2-grams, 3-grams
-*  Compare outcomes for 3 independent randomized sets of 1-grams
-
-The pseudo-code block below can pull random lines from a data table, but does not work correctly on a document term matrix.
-The second problem is how to pull the training and test data sets out and process them identically.
-
-```{}
-set.seed(1234) 		# ensures the same random set gets picked each time.
-
-comments_raw_dtm_train <- sample_frac(comments_raw, .75)	# sample_frac pulls out 75% of "comments_raw" dataset randomly using the sample sequence identified by set.seed(1234)
-
-comments_raw_dtm_test <- setdiff(comments_raw, comments_raw_dtm_train)	#setdiff creates a "test" dataset using anything NOT in the "comments_raw_dtm_train" subset of "comments_raw."
-```
-
-The solution was not to pull the data twice, but to randomly (but repeatably) shuffle the starting data table. How I solved this was:
-1. Create a vector of sequential numbers the same length as the number of data frame rows.
-2. Randomize the sequential numbers using set.seed to make pattern reproducible.
-3. Add the randomized numbers as a new column to the data frame.
-4. Re-order the data table to the randomizer column.
-
-This makes it easy to generate a new sequence of randomized training and test samples to run replicates. Only one number needs to be changed.
 

### 2019-02-21 docs/hiding/Random_forest_demo.Rmd
update

--- 2019-02-21 docs/hiding/Multinomial_bayes_V3_1.Rmd
@@
 ---
-title: "Multinomial Naive Bayes Workflow Ver 3.0.b."
+title: "Demo of Random Forest Method Ver 1.0.a."
 author: "Dan Johnson"
@@
 ##Summary
-This is an alternative version of my basic NB workflow for classifying TA comments.
-The code is slightly different than what was used for Categorical NB testing, but takes in text using the same command set as NB-Ver 2.0.
-It is based on a tutorial located here:
-[https://rpubs.com/cen0te/naivebayes-sentimentpolarity]
-
-This alternative code allows for true multinomial analysis, by switching between using binary yes/no categorization of words' presence or absence, and the actual frequencies.
-There is a trade-off though; to use non-categorical values, I can only have 2 assignment bins: in-group versus out-group.
-If I use multiple groups, the data fail to parse correctly.
-If I want to look at multiple classification bins at the same time, I have to use the categorical values method.
-
-The point seems moot overall, since categorical NB seems to perform better than true multinomial NB.
-For both Subject and Structure, I reach 75-84% overall accuracy if I sort comments into ingroup/outgroup categories using binary (yes/no) word frequencies.
-Accuracy drops to ~65% if I use the raw numeric frequencies.
-Changing the Laplace value does not affect overall performance.
-
-Naive Bayes permutations seem to plateau at 80-85%, regardless of permutations.
+This is testing the random forest method described here:
+[https://datascienceplus.com/random-forests-in-r/]
 
@@
 ```{r}
-library(tm)
-library(RTextTools)
-library(e1071)
-library(dplyr)
-library(caret)
-library(doMC)
-registerDoMC(core=detectCores())
+#library(randomForest) #tools
+#library(RTextTools) #demo data. Package which contains the Boston housing dataset
+#library(MASS)
+attach(Boston)
+set.seed(101)
 
-library(quanteda)
-library(RColorBrewer)
-library(ggplot2)
-
-#library(tidyverse) 
-#library(tidytext)
-#library(tidyr)
-```
-
-####\
-
-##Version 3.0.b of Naive Bayes (NB) Protocol
-###Initial Data Input
-```{r}
-#Read in TA comments from CSV file. 
-#For convenience, I am reading in a shorter CSV file created previously using ver.2.0 code. This CSV file only lists 
-#whether a comment is in or out of the 2 main groups, "Subject:3_technical" and "Structure:4_specific"
-
-subset_M <- read.csv("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/comments_binary.csv", stringsAsFactors=FALSE)
-glimpse(subset_M)
-```
-```{r}
-set.seed(2012) #Randomizer
-subset_M <- subset_M[sample(nrow(subset_M)), ]
-glimpse(subset_M)
+dim(Boston)
 ```
@@
 ```{r}
-text.corpus<-corpus(subset_M$text)
-docvars(text.corpus)<-subset_M$subject
-glimpse(text.corpus)
+#training Sample with 300 observations
+train=sample(1:nrow(Boston),300)
+#?Boston  #to search on the dataset
+
+Boston.rf=randomForest(medv ~ . , data = Boston , subset = train)
+Boston.rf
+```
+
+```{r}
+plot(Boston.rf)
+```
+
+```{r}
+oob.err=double(13)
+test.err=double(13)
+
+#mtry is no of Variables randomly chosen at each split
+for(mtry in 1:13) 
+{
+  rf=randomForest(medv ~ . , data = Boston , subset = train,mtry=mtry,ntree=400) 
+  oob.err[mtry] = rf$mse[400] #Error of all Trees fitted
+
+  pred<-predict(rf,Boston[-train,]) #Predictions on Test Set for each Tree
+  test.err[mtry]= with(Boston[-train,], mean( (medv - pred)^2)) #Mean Squared Test Error
+
+  cat(mtry," ") #printing the output to the console
+
+}
+```
+
+```{r}
+test.err
+oob.err
+matplot(1:mtry , cbind(oob.err,test.err), pch=19 , col=c("red","blue"),type="b",ylab="Mean Squared Error",xlab="Number of Predictors Considered at each Split")
+legend("topright",legend=c("Out of Bag Error","Test Error"),pch=19, col=c("red","blue"))
 ```

### 2019-02-21 docs/hiding/reference_tables.Rmd
update

---
title: "reference_tables"
author: "Dan Johnson"
date: "2/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reference Tables

Labels in **comment.subject** are routinely recoded as:

0_null
1_basic
2_writing
3_technical
4_logic
5_exposition
6_misconduct
7_administrative
8_nobasis
9_split
10_context
11_flag
12_narrative
13_sciname

Labels in **comment.structure** are routinely recoded as:

0_null
1_pointer
2_copyedit
3_general
4_specific
5_rationale
6_holistic
7_idiomatic
8_nobasis

### 2019-02-21 background.Rmd
updates

@@
 Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed.
+
 Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills.
+Technical writing training usually is part of a lab course associated with a lecture class.
+Many times lab sections are led by graduate teaching assistants or even advanced undergraduates.
+TAs may have under-developed writing skills themselves, lack pedagogical content knowledge needed to guide students effectively, or not have been explicitly trained to teach writing in a particular way.
 
-Technical writing training usually is part of a lab course associated with a lecture class.
-These lab sections may be be led by graduate teaching assistants or even advanced undergraduates.
-These instructors may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.
+In 2014 we launched a report data intake process we call **SAWHET** that systematically collects writing and other data from student writers and graduate instructors.
+Using these data we are mapping key features of student writing, as well as connections and correlations between those features.
+These insights then inform how we teach technical writing and train graduate instructors.
 
-In 2014 we launched an automated data collection process we call **SAWHET** that systematically collects structured writing and other data from student writers and graduate instructors.
-Our resarch goal is to better understand student and instructor behaviors within the technical writing instructional space.
-From this empirical basis our practical goal is to develop techniques for teaching technical writing that:
+Currently our structured database contains >5000 laboratory reports written by undergraduate students for introductory biology courses, all in standardized .docx and .txt formats.
+Additional data and metadata available for each report includes:
 
-*  Can be implemented at large scale by less experienced teaching assistants;
+*  All comments that a TA made on each report
+*  Context. Was it first or second report of the semester? Initial submission or revision? Course and topic prompt? Year in college of the author?
+*  Instructor information. Which TA provided feedback? What score was assigned? What was rationale for score?
+
+Our first goal is to better understand student and instructor behaviors within the technical writing instructional space.
+Our second and more practical goal is to develop techniques for teaching technical writing that:
+
+*  Embody best practices from 40+ years of research on "writing in the disciplines;"
 *  Emphasize student development through holistic coaching;
 *  Shift student and instructor attention away from copy-correction towards larger writing issues; and
-*  Embody other best practices from 40+ years of research literature on "writing in the disciplines."
-
-We have a structured database of >5000 laboratory reports written by undergraduate students for introductory biology courses.
-In addition we have all comments that a TA made on the individual reports, plus associated metadata:
-
-*  Context. Was it first or second report of the semester? Initial submission or revision? What course and topic in the course? What year in college was the author?
-*  Instructor information. Which TA provided the feedback? What score was assigned, and what what the rationale behind that score?
-
-We are using multiple text-mining methods to map key features of student writing, as well as connections and correlations between those features.
-These insights then inform how we teach technical writing and train graduate instructors.
+*  Can be implemented at scale by less experienced teaching assistants.
 
@@
 
-## Context for This Project & Prior Work
-Students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead.
-To that end, we train graduate teaching assistants to use a **global first, specific later** approach to grading undergraduate technical writing.
+## Context for This Project
+From prior studies (REFs), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead.
+To that end, we train graduate teaching assistants to use a **global issues first, specific issues later** approach to correcting undergraduate technical writing.
 We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues.
@@
 As students' writing improves, the TA can focus on smaller issues.
+Our bins-based report grading protocol reinforces this strategy.
 
-This strategy is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy.
-One of our ongoing questions has been, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**
+The coaching-oriented approach is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy.
+Effort self-reporting by TAs is likely to be unreliable.
+We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**
 
-In Summer 2018, we began analyzing TAs' comments from student papers.
-We extracted and tabulated all comments from approximately 500 pairs of reports (initial submissions and revisions) collected during a single semester in 3 different courses.
-Initial data cleaning involved removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent.
-In all, the final tabulated dataset included ~11,000 separate comments.
+####\
+
+## Prior Work
+In 2018, we began trying to answer this question by analyzing TAs' comments from student papers directly.
+We extracted and tabulated all comments from approximately 500 pairs of reports.
+After initial data cleaning (removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent), we had a working dataset of ~11,000 separate comments.
 
@@
 
-* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
-* __Structure__: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?
+* **Subject**: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
+* **Structure**: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?
 
-The sorted contingency table provided us with a relative estimate of TA effort.
+Raw counts for each subject/structure pair were converted to fractional frequencies. The final table (excerpt is shown below) provided us with a direct estimate of TA effort.
+
+![*Excerpted data from frequency table for hand-coded TA comments on student reports*](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/images/Hand_frequency_table.png)
 
@@
 ## The Central Problem of This Project
-Qualitative coding of TA comments is informative but unsustainable.
-Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort.
-Even with the codebook, individual coders will require training to achieve any accuracy.
-Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.
+Coding TA comments this way is informative but doing so by hand is unsustainable.
+Developing the codebook then rating the initial set of 11,000 comments required nearly 2 months of investigator effort.
+Even with the codebook in place, individual coders need training to achieve sufficient accuracy, and there is significant risk of "coding drift" over time.
+
+Given 10-12,000 comments are generated EACH SEMESTER, an automated classification method is needed.
 

### 2019-02-21 index.Rmd
updates

@@
 ## Overview
-I am interested in understanding how students develop technical writing skills in biology, and how we might improve that process by applying data science methods.
+I am using classroom research and data science to better understand how students develop technical writing skills in large-enrollment biology courses.
 
-My current NSF-sponsored project asks if and how scripted instruction, automated support, and holistic feedback accelerates student development as writers.
-One specific aim is to see if changing the structure and focus of **instructors' comments** affects how students' writing develops over time.
-To assess this we must extract instructor comments from student reports, categorize those comments according to their subject and structure, then correlate the types of comments TAs make with student performance.
+My current NSF-sponsored grant asks whether we can accelerate student development as writers using a mix of scripted active instruction, automated support of writing in progress, and holistic feedback.
+One specific question we have is how **changing the structure and main focus of instructors' comments** affects students' writing development over time.
+To answer this we extract instructor comments from student reports, categorize the comments according to their subject and structure, then correlate the types of comments instructors make with student performance.
 
@@
 ## Prior Work
-In Summer 2018, we extracted and tabulated ~11,000 TA comments from >500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses during a single semester.
-Using an iteratively generated codebook, the 11,000 TA comments were sepparately classified in two categories:
+In Summer 2018, we extracted and tabulated ~11,000 comments made by teaching assistants (TA) from >500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses during a single semester.
+Using an iteratively generated codebook, each of the comments was separately classified in two categories:
 
@@
 #### \   
-Counts of coded TA comments were tabulated into a 2-way contingency table that provided us with a direct measure of the most common types of comments TAs made on student work, and a relative estimate of which elements of writing TAs emphasize most often.
+Counts of coded TA comments were tabulated according to subject and structure, and frequencies for each subject/structure pair calculated as a fraction of all comments.
+This 2-way contingency table provided us with a direct estimate of the most common types of comments TAs made on student work, and a relative estimate of which elements of writing TAs emphasize most often.
 
-While classifying TA comments this way is informative, doing so by hand is impractical: 10-12,000 comments need to be scored EACH semester.
+Classifying TA comments this way is informative, but doing so by hand is impractical: 10-12,000 comments need to be scored EACH semester.
+There also is a significant risk of "coding drift" over time, reducing accuracy.
 
@@
 My project for the **Faculty Learning Community for R** was to create an automated comment classifier that could generate two-axis contingency tables comparable to the one generated using hand-coded data.
-This automated classifier will be used as part of the larger NSF-funded research project to:
+This automated classifier will used within the larger NSF-funded research project to:
 
@@
 ## Site Map
-Insert a text-based site map of the major pages and branch points here.
+Insert a text-based overview and site map of the major pages and branch points here.
 

### 2019-02-21 background.Rmd
update

@@
 ---
-title: "Background for the Project"
+title: "Background"
 bibliography: library.bib
@@
 ####\  
-
-## Context for This Project
+##Context for This Project
 From prior studies (REFs), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead.
@@
 
-## Prior Work
+###Prior Work
 In 2018, we began trying to answer this question by analyzing TAs' comments from student papers directly.
@@
 
-## The Central Problem of This Project
+##Central Problem
 Coding TA comments this way is informative but doing so by hand is unsustainable.
@@
 ####\
+
+##Work Plan
+
+1.  Compile initial dataset, and de-identify both TAs and students by replacing names with anonymous IDs.
+2.  Export TA comments, metadata from student reports as a "tidy" dataframe.
+3.  Explore comment data structure independently of classification to identify useful elements for feature engineering.
+4.  Write the initial Naive Bayes workflow for supervised text classification.
+5.  Assign TA comments to categories defined in our existing codebook, then compare outcomes from hand-coding versus automated classification.
+6.  Using features (#3) and comparison data from initial classification protocol (#4) as a guide, test iterations of the classifier method and optimize. 
+7.  Apply the optimized classifier to the original comment dataset, identify all comments that are being classified incorrectly, and look for patterns.
+8.  Write a small set of rule or REGEX pattern-based searches that can identify comments which are more likely to be classified incorrectly. 
+9.  Append rules/pattern-based pre-screening to the optimized Naive Bayes classifier.
+10.  Validate this vertical Ensemble Classifier on a subset of ~2000 TA comments extracted from a non-overlapping set of student lab reports.
+
+####\

### 2019-02-21 index.Rmd
update

@@
 
-## Overview
+##Overview
 I am using classroom research and data science to better understand how students develop technical writing skills in large-enrollment biology courses.
@@
 
-#### \   
-## Prior Work
 In Summer 2018, we extracted and tabulated ~11,000 comments made by teaching assistants (TA) from >500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses during a single semester.
@@
 
+#### \   
+##Specific Goals
 Classifying TA comments this way is informative, but doing so by hand is impractical: 10-12,000 comments need to be scored EACH semester.
@@
 
-#### \
-
-##Specific Project Goals
 My project for the **Faculty Learning Community for R** was to create an automated comment classifier that could generate two-axis contingency tables comparable to the one generated using hand-coded data.
@@
 #### \   
-## Site Map
+##Project Site Map
 Insert a text-based overview and site map of the major pages and branch points here.
 
+When appropriate, anchor pages branch to extended supporting documentation.
+
 #### \

### 2019-02-21 litreview.Rmd
update

---
title: "Literature Review"
author: "Dan Johnson"
date: "2/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Literature base

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

### 2019-02-21 background.Rmd
update

--- 2019-02-21 background.Rmd
@@
 
-Undergraduate introductory STEM courses typically include technical writing training.
-Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed.
+Undergraduate introductory STEM courses often include technical writing training.
+Too often it is treated as an incidental skill that each student develops "along the way." At other times the instructional strategy is  haphazardly designed, implemented, and executed.
 
-Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills.
-Technical writing training usually is part of a lab course associated with a lecture class.
-Many times lab sections are led by graduate teaching assistants or even advanced undergraduates.
+Adding to the learning challenge, faculty may not be the ones helping students develop writing skills.
+If technical writing training is part of a lab course, graduate teaching assistants or even advanced undergraduates may be who teaches technical writing.
 TAs may have under-developed writing skills themselves, lack pedagogical content knowledge needed to guide students effectively, or not have been explicitly trained to teach writing in a particular way.
@@
 ####\  
-
-## Context for This Project
+##Context & Prior Work
 From prior studies (REFs), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead.
-To that end, we train graduate teaching assistants to use a **global issues first, specific issues later** approach to correcting undergraduate technical writing.
-We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues.
-Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection.
-As students' writing improves, the TA can focus on smaller issues.
+To that end, our program trains graduate TAs to use a **global issues first, specific issues later** approach to correcting undergraduate technical writing.
+We ask them to spend a shorter amount of time writing fewer comments that focus on broader issues.
+Ideally, TAs provide no more than 3-5 comments per text page, and the comments given should encourage student reflection.
+As students' writing improves, TAs can focus on smaller issues.
 Our bins-based report grading protocol reinforces this strategy.
 
-The coaching-oriented approach is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy.
-Effort self-reporting by TAs is likely to be unreliable.
-We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**
+Our coaching-oriented approach is well-supported by prior research, but we are uncertain how accurately TAs implement the recommended strategy.
+We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"** Self-reported TA effort is likely to be unreliable.
+What we need is a more direct measure of effort.
 
-####\
-
-## Prior Work
-In 2018, we began trying to answer this question by analyzing TAs' comments from student papers directly.
+In 2018, we began trying to analyze TAs' comments from student papers directly.
 We extracted and tabulated all comments from approximately 500 pairs of reports.
@@
 
-We next used a subset of 100 randomly selected comments to build an initial qualitative codebook for sorting comments into types.
-The initial codebook was re-tested using 1100 additional randomly selected comments.
-Using the final codebook, we sorted all 11,000 comments into separate categories (see LINK for complete codebook.)
+We next used a subset of 110 randomly selected comments to build an initial qualitative codebook for classifying comments based on subject and structure.
+This initial codebook was tested and revised using blocks of 1100 additional randomly selected comments until the categories and criteria stabilized.
+Using the final codebook, we then sorted all 11,000 comments into separate categories (see PAGE_LINK for complete codebook.)
 
-Coded TA comments were sorted on two axes:
+TA comments were coded and sorted according to:
 
@@
 
-Raw counts for each subject/structure pair were converted to fractional frequencies. The final table (excerpt is shown below) provided us with a direct estimate of TA effort.
+Raw counts for each subject/structure pair were then converted to fractional frequencies. The final frequency table (excerpted below) provided us with a direct estimate of TA effort and emphasis.
 
@@
 
-## The Central Problem of This Project
-Coding TA comments this way is informative but doing so by hand is unsustainable.
-Developing the codebook then rating the initial set of 11,000 comments required nearly 2 months of investigator effort.
-Even with the codebook in place, individual coders need training to achieve sufficient accuracy, and there is significant risk of "coding drift" over time.
+##Central Problem
+Analyzing TA comments this way is informative but coding by hand is unsustainable.
+Developing the codebook then rating the initial set of 11,000 comments required nearly 80 hours of investigator effort.
+Even with the codebook in hand, individual coders need training to achieve sufficient accuracy, and there is significant risk of "coding drift" over time.
 
@@
 ####\
+
+##Work Plan
+The work plan for this project is:
+
+1.  Compile initial dataset and de-identify both TAs and students by replacing names with anonymous IDs.
+2.  Import TA comments and relevant metadata from student reports into R as a "tidy" dataframe.
+3.  Explore comment data structure, n-gram frequencies, etc., to identify potentially useful elements for feature engineering.
+4.  Write an initial Naive Bayes workflow for supervised text classification.
+5.  Assign TA comments to categories using the NB classifier, then compare outcomes from automated classification to hand-coded results.
+6.  Using features (from Step #3) and comparison data from initial classification protocol (from Step #4) as guides, test iterations and optimize the classifier method. 
+7.  Apply the optimized classifier to the original comment dataset, identify all comments that are being classified incorrectly, and look for potential patterns.
+8.  Write a small set of rule- or REGEX pattern-based searches that can identify comments which are more likely to be classified incorrectly. 
+9.  Append rules/pattern-based pre-screening to the optimized Naive Bayes classifier.
+10.
+Validate this vertical Ensemble Classifier by classifying a subset of ~2000 new TA comments extracted from a non-overlapping set of student lab reports.
+Compare automated results for new data set to hand-coded categories.
+
+####\

### 2019-02-21 findings.Rmd
update

--- 2018-12-16 project.Rmd
@@
 ---
-title: "Project"
+title: "Findings"
 author: "Dan Johnson"
-date: "10/26/2018"
+date: "2/22/2018"
 output: html_document
@@
 
-# Abstract
+##Classifier Accuracy
+The tables below show observed frequencies for 9,340 TA comments coded by hand, and frequencies for the same TA comments categorized using the optimized Naive Bayes classifier **without** pre-screening.
+The far right columns are overall subject frequencies, the bottom rows, overall structure frequencies.
 
@@
 
-# Introduction
-## Background
-Undergraduate introductory STEM courses typically include technical writing training.
-Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed.
-Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills.
-
-Technical writing training usually is part of a lab course associated with a lecture class.
-These lab sections may be be led by graduate teaching assistants or even advanced undergraduates.
-These instructors may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.
-
-In 2014 we launched an automated data collection process we call SAWHET that collects systematic observations and sample data from student writers and graduate instructors.
-Our resarch goal is to better understand student and instructor behaviors within the technical writing instructional space.
-Our practical goal is to develop techniques for teaching technical writing that:
-
-*  Can be implemented at large scale by less experienced teaching assistants;
-*  Emphasize student development through holistic coaching;
-*  Shift student and instructor attention away from simple copy-correction towards larger writing issues; and
-*  Embody other best practices outlined in nearly 40 years of research literature on "writing in the disciplines."
-
-We have a structured database of >5000 laboratory reports written by undergraduate students for introductory biology courses.
-In addition we have all comments that a TA made on the individual reports, plus associated metadata:
-
-*  Context. Was it first or second report of the semester? Initial submission or revision? What course and topic in the course? What year in college was the author?
-*  Instructor information. Which TA provided the feedback? What score was assigned, and what what the rationale behind that score?
-
-We are using multiple text-mining methods to map key features of student writing, as well as connections and correlations between those features.
-These insights then inform how we teach technical writing and train graduate instructors.
+![*Comparison of frequency tables generated by hand-coding TA comments (top) versus predicted categories using optimized NB classifier (bottom).
+*](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/images/Result_of_NB_2D_mapping_Feb2019.png)
 
@@
 
-## Context for This Project & Prior Work
-Students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead.
-To that end, we train graduate teaching assistants to use a **global first, specific later** approach to grading undergraduate technical writing.
-We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues.
-Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection.
-As students' writing improves, the TA can focus on smaller issues.
+On average, the classifier is only 80-90% accurate for one subject sub-category or one specific comment.
+OVERALL though, the optimized NB classifier estimated relative frequencies of Subject subcategories very well; there was <1.5% difference in fractional frequencies between observed (hand-coded) and computationally assigned Subject sub-categories.
 
-This strategy is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy.
-One of our ongoing questions has been, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**
-
-In Summer 2018, we began analyzing TAs' comments from student papers.
-We extracted and tabulated all comments from approximately 500 pairs of reports (initial submissions and revisions) collected during a single semester in 3 different courses.
-Initial data cleaning involved removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent.
-In all, the final tabulated dataset included ~11,000 separate comments.
-
-We next used a subset of 100 randomly selected comments to build an initial qualitative codebook for sorting comments into types.
-The initial codebook was re-tested using 1100 additional randomly selected comments.
-Using the final codebook, we sorted all 11,000 comments into separate categories (see LINK for complete codebook.)
-
-Coded TA comments were sorted on two axes:
-
-* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
-* __Structure__: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?
-
-The sorted contingency table provided us with a relative estimate of TA effort.
+Overall accuracy was considerably lower for Structure sub-categories.
+Specifically, the NB classifier over-assigned comments to the "Copy-editing" sub-category, and under-assigned comments to the "Specifics" sub-category.
+Most incorrectly classified comments belonged to the "Technical" Subject sub-category.
+This **suggest** there may be specific terms present in comments within this sub-category that are problematic, and warrant further exploration.
 
@@
 
-## The Central Problem of This Project
-Qualitative coding of TA comments is informative but unsustainable.
-Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort.
-Even with the codebook, individual coders will require training to achieve any accuracy.
-Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.
+##Data Story 2
 
@@
 
-# Literature Review
-##General Approach to Initial Dataset
-The following excerpt from *Text Mining with R: A Tidy Approach* (Julia Silge and David Robinson) summarizes how extended text (in this case, TA comments) can be organized as a tidy matrix.*All emphases are mine.*
-
->We thus define the tidy text format as being a table with one-token-per-row.
-A token is a **meaningful unit of text**, such as a word, that we are interested in using for analysis, and tokenization is the process of splitting text into tokens.
-
->This one-token-per-row structure is in contrast to the ways text is often stored in current analyses, perhaps as strings or in a document-term matrix.
-For tidy text mining, the token that is stored in each row is most often a single word, **but can also be an n-gram, sentence, or paragraph** .
-In the tidytext package, we provide functionality to tokenize by commonly used units of text like these and convert to a one-term-per-row format.
-
-Using tidy format (vs. other common text data formats) better maintains relationships between individual comments and metadata, which simplifies subsequent analysis.
+#Tool or Analysis 2
 
@@
 
-# Methods
+##Data Story 3
 
@@
 
-## Data
+##Data Story 4
 
@@
 
-## Analysis
-
 ####\
-
-# Results
-
-####\
-
-# Discussion
-
-####\
-
-# Conclusion
-
-####\
-
-# References
-
-####\

### 2019-02-21 litreview.Rmd
update

@@
 This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+
+I found this summary, and a description of the approach at
+[https://www.tidytextmining.com/topicmodeling.html]():
+
+>"In text mining, we often have collections of documents, such as blog posts or news articles, that we’d like to divide into natural groups so that we can understand them separately.
+Topic modeling is a method for unsupervised classification of such documents, similar to clustering on numeric data, which finds natural groups of items even when we’re not sure what we’re looking for.
+
+>"Latent Dirichlet allocation (LDA) is a particularly popular method for fitting a topic model.
+It treats each document as a mixture of topics, and each topic as a mixture of words.
+This allows documents to “overlap” each other in terms of content, rather than being separated into discrete groups, in a way that mirrors typical use of natural language."
+
+My challenge is that LDA clusters items based on its own analysis. My groups are already sorted based on the **existing coding book.** The individual groups are already labeled.
+
+####\
+
+**Option: Comparison Modeling Based on Twitter Strategies**
+
+What if I treated each of the pre-defined sub-sets of comments I created using codebook analysis as if they were tweets from separate users? That approach is described here: 
+	https://www.tidytextmining.com/twitter.html
+
+Strategy would be:
+
+*  Un-nest tokens to create a tidy frame and remove stop words using anti-join
+*  Calculate word frequencies for each sub-type
+*  Compare frequencies by calculating log-odds ratios
+
+####\
+
+**Option: Comparison Modeling Via Text Classification**
+
+What if I treated each of the comments as text blocks where pre-defined words are more or less likely to be associated with the category. Jerid developed a Shiny app that does something similar:
+https://francojc.github.io/2015/08/15/generating-annotated-text-in-shiny/
+
+####\
+
+**Option: Latent Dirichlet Allocation**
+
+# Session 4 (Dec. 3, 2018)
+Looking at Jerid's *Intro to Statistical Thinking* clarified some points in my approach.
+I definitely need to be thinking about this problem more in terms of unsupervised exploration.
+It did raise some follow-up questions:
+
+*  A training set with 80% of the data seems excessive.
+If there is a large dataset, is there any cost or benefit to splitting to 33/33/34%, or even 25% x 4?
+The 80/20 and 60/20/20% also seems out of balance, but there is probably logic I do not see.
+*  In the exploration phase, is re-sampling such a major concern?
+
+**Separate note:** thank you many times over for the Regex101 link! Writing regex correctly has long been a pain for me! This greatly simplifies the process.
+
+####\

### 2019-02-22 NB_2r2c.Rmd
update

--- 2019-02-21 docs/hiding/Naive_bayes_V2r_2c.Rmd
@@
 author: "Dan Johnson"
-date: "2/14/2019"
+date: "2/21/2019"
 output: html_document
@@
 ##Summary
-This workflow uses Naive Bayes to classify TA comments based on Subject and Structure. The major additions and revisions in **Version 2r.2c** are:
+This it the current Naive Bayes workflow used to classify TA comments based on Subject and Structure. Additions and revisions made to **Version 2r.2c** are:
 
-*  Randomly selected data rows used for training and test datasets. 
+*  Random selection of data rows used for training and test datasets. 
 *  Final output matrix tables are more informative. 
-*  Addition of a back-analysis of entire dataset, not just testing sub-set.
-*  Final output CSV file is a tidy data table of the original comments, the hand-scored and predicted subject, and hand-scored and predicted structure.
+*  Addition of a back-analysis step that reviews entire dataset, not just testing sub-set.
+*  Outputs a tidy data table in CSV format that contains the original comments (pre-clean-up) plus their hand-scored and predicted subject and structure.
 
@@
 library(caret)
-
-#library(RTextTools)
-#library(doMC)
-#registerDoMC(core=detectCores())
-#library(tidyr)
-#library(dplyr)
-#library(gmodels)
 ```
@@
 ###Initial Data Input
-```{r data_input}
+```{r input data}
 #Read in TA comments from CSV file.
@@
 Extract comments and classification data as subsets of CSV.
-```{r data_subsetting}
+```{r subset data}
 #Select rows representing the sub-groups to compare. 
@@
 ```{}
-# Function to KEEP word frequencies while still using the same variable names and code structure. The resulting matrices have cells with word frequencies instead.
+# Alternate version of convert_function code block.
+This version KEEPs word frequencies while still using the same variable names and code structure.
+The resulting matrices have cells with word frequencies instead.
+#Only works when the NB classifier is looking at ingroup/outgroup. Breaks on multiple categories.
+#
 convert_counts2 <- function(x) {
@@
 
-```{r make_binary}
+```{r apply_conversion}
 comments_all <- apply(comments_dtm_freq_all, MARGIN = 2, convert_counts2)
@@
 ```{r bayes_training}
+# Train the classifier on training data, then test on test subset and full dataset.
 comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject, laplace=1)
@@
 this is what the confusion matrix calculates.
-The faster way is to calculate a confusion matrix, which includes BOTH the truth table, and the final overall accuracy.
 
-#The Cross Table command below (from "gmodels") generates a more detailed output table with frequency values for each individual cell.
+#The faster way is to calculate a confusion matrix, which includes BOTH the truth table, and the final overall accuracy.
+
+#Alternatively, the Cross Table command below (from "gmodels" package) generates a more detailed output table with frequency values for each individual cell.
 #CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
@@
 ##Graph Outputs
-Plotting Hand-coding vs. Bayes
+Comparing outcome from Hand-coding vs. Bayes
 
-Need to combine two datasets into one, and create subsets based on the
+Need to combine two datasets into one.
 
 ```{r}
-tidy(proptallysum1)
-ggplot(data=proptallysum1, aes(x=Hand.Coded.Subject, y=total, group=1))+
-  geom_line() +
-  geom_point()
-
+ggplot(data=proptallysum1, aes(`Hand.Coded.Subject`, `total`,group=1))+
+  geom_line()
 ggplot(data=proptallysum2, aes(x=Bayes.Predicted.Subject, y=total, group=1))+
-  geom_line() +
-  geom_point()
+  geom_line()
+ggplot(data=proptallysum3, aes(x=Hand.Coded.Structure, y=total, group=1))+
+  geom_line()
+ggplot(data=proptallysum4, aes(x=Bayes.Predicted.Structure, y=total, group=1))+
+  geom_line()
 ```

### 2019-02-22 bayes_history.Rmd
update

---
title: "Bayes Design"
author: "Dan Johnson"
date: "2/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Naive Bayes Design History

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2019-02-22 findings.Rmd
update

@@
 title: "Findings"
-author: "Dan Johnson"
-date: "2/22/2018"
 output: html_document
@@
 
-##Classifier Accuracy
+#Classifier Accuracy
+*Updated 2/22/19*
+
 The tables below show observed frequencies for 9,340 TA comments coded by hand, and frequencies for the same TA comments categorized using the optimized Naive Bayes classifier **without** pre-screening.
@@
 
+#Overview
+This notebook summarizes my progress, thoughts while building a working Ensemble method for categorizing TA comments.
+
+####\
+
+##Replicate Cycles Test of Naive Bayes
+One of my ideas was to see if I could refine Naive Bayes by running the same classifier with different randomized datasets, then putting individual outputs side by side and assigning final categories based on most frequent classification in the set.
+
+To test this I ran 5 Subject classifier replicates using the code in *Naive_bayes_V2r_2b.rmd*, with 5 different ```set.seed``` values.
+The exported CSV files were combined to get a set of data with up to 5 classifier assignments for the same comment.
+
+Using multiple randomized NB training runs does not seem to help.
+First, it was rare for the  randomized runs to assign different incorrect classes.
+Second, when I compared 5 independently randomized runs by eye, the incorrectly categorized items were repeatedly put into the SAME incorrect group each time.
+Based on this, an Ensemble method using H-stacked categorical NB is unlikely to reduce overall errors in classification.
+
+Looking at the combined data DID provide other helpful insights into the overall classification procedure though. Naive Bayes does not seem to do well with comments that:
+
+*  Contain many or mostly punctuation symbols, such as:
+    +  : not ,
+    +  ?
+    +  ??
+    +  ???
+*  Are whole numbers
+    +  1
+    +  2
+    +  10
+*  Are one word questions or comments
+    +  because?
+    +  "and"
+*	 Are sentences enclosed in punctuation
+*  Are made up entirely of STOP/common words (I am not as sure on this one)
+
+####\
+
+##Overall Accuracy
+I also looked at OVERALL accuracy of NB by recreating the 2-dimensional contingency table that I made when hand-coding, and comparing it side-by-side to NB classifications.
+From this I gained additional insights.
+
+OVERALL, the 2-D Naive Bayes model performed very well for classifying comment Subject.
+When I re-ran the full original dataset (9340 comments) through the prediction model, NB estimates for Subject were within 1.5% of my hand-scored assignments.
+Structure classification was not as robust.
+Other observations:
+
+*  NB over-classified comments as "copyediting of technical issues" by 8-10%, and under-classified comments that should be marked as "specific technical issues" by almost same amount.
+*  Comments that belonged to groups making up less than 1-2% of all comments were more likely to be mis-classified. 
+*  Longer text comments seemed to be mis-classified more often. Follow-up methods to evaluate: 
+    +  Split longer comments into shorter phrases
+    +  Evaluate only the first 5-10 words of a comment
+
+It appears that Subject categories are more terms-centered, while Structure is based more on the sentiment and sentence structure of a comment.
+This suggests that I need to incorporate a sentiment classifier or PoS tagger step.
+
+####\
+
+##Mixed Ensemble Method
+Based on the  seems like the most reasonable approach for maximum accuracy is going to be a mixed Ensemble method. The initial working model is to:
+
+Add a pre-screening step prior to implementing NB:
+
+*  Isolate "short-string" comments first. These are comments consisting of single words, two-word phrases, more punctuation than text, etc. These represent mostly pointers and copy-editing.
+*  Look for 2-grams with high frequencies in one particular category. (Ex., "basic criteria" 2-gram identifies a Basic Criteria comment.)
+
+####\
+
 ####\

### 2019-02-22 initial_exploration.Rmd
update

@@
 ---
-title: "Initial Exploration"
-author: "Dan Johnson"
-date: "2/3/2019"
+title: "Initial Explorations"
 output: html_document
@@
 
-#Preliminary Exploration of TA Comments Dataset
-The first question to answer before making deeper comparisons is **are TAs' comments obviously different between sub-categories?** This page describes several different visual comparisons I made of word frequencies before beginning in-depth text analyses.
+The first question to answer before making deeper comparisons is **are TAs' comments obviously different between sub-categories?** To that end, I ran three different analyses of word usage in TA comments.
 
-*  Comparisons of word frequencies in subject sub-categories, using log-odds ratios
-    +  Pearson Correlations of Word Frequencies
-    +  Mapping Analyses Visually
-    +  Calculating and Plotting Log-Odds Pairs
-*  Direct comparison of word frequency lists
+*  Pearson correlations for word use frequencies. This compares words found in all Subject sub-categories
+*  Word frequency differences based on log-odds ratios. This highlights terms that are the most different between sub-categories.
+*  Unique n-grams mapping. This identifies text strings, phrases, and rare terms that are unique (or nearly unique) to one sub-category.
 
-##Initial Library and Data Import
+#Setup
 ```{r Libraries_1}
@@
 library(scales)
-```
 
-```{r}
-#This block reads in full CSV into a starting dataframe named "base_data".
-#Subsequent analyses are performed on extracted subsets of "base_data".
+#Read in CSV as a starting dataframe named "base_data". 
 base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
@@
 
-##Direct Comparison Using Only Words Appearing in All Subsets of Comments
-I think this is less informative, but it does let me see the data more clearly.
-
-Basic strategy is to take all comments from base_data, unnest tokens, filter stopwords, count and sort remaining terms, then write back to a new datatable called 'sortedwords.comments.all'.
-Calculate TOTAL words and store.
-Then calculate fractional frequencies for each word and store using mutate.
-
-I wrote this block so that, once CSV is imported, it does not require any of the code above. This is an early effort, so likely has some poorly developed code that could be refactored.
+#Pearson Correlations of Word Frequencies
+##Are Shared Words Unequally Distributed?
+This analysis takes all comments from base_data, tokenizes and filters stopwords, then counts and sorts fractional frequencies for each word in each category.
+**NOTE**: this analysis only compares word use for terms present in ALL sub-categories.
+If a word is MISSING from **any subcategory**, the data row for that word is pruned from the dataset.
+This eliminates all unique words in favor of words found in all sub-categories.
 
 ```{r}
-# Tokenize phrases, remove the stop words listed in R's standard reference file.
+#Tokenize phrases, remove the stop words listed in R's standard reference file.
 base_data_tokenized <- base_data %>% 
@@
 
-# This is an ALTERNATIVE METHOD for pulling out stopwords.  
+#ALTERNATIVE METHOD for pulling out stopwords.  
 sortedwords.comments.all <- base_data_tokenized %>%
@@
 
-#Calculate the number of words in total in the subset
+#Calculate total # of words
 total_words.all <- sortedwords.comments.all %>% 
@@
 
-#Mutate table to calculate then append fractional word frequency in the text.
+#Calculate then append fractional word frequencies.
 sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
@@
   )
-```
 
-```{r}
-#Pull SUBSETs of tokenized data based on code.subject: 
+#Select SUBSETs of tokenized data based on code.subject: 
 base_data_tokenized.basic <- subset(base_data_tokenized, code.subject == "1_basic", select = (1:28))
@@
 base_data_tokenized.logic <- subset(base_data_tokenized, code.subject == "4_logic", select = (1:28))
-#base_data_tokenized.praise <- subset(base_data_tokenized, code.subject == "5_exposition", select = (1:28))
-#base_data_tokenized.misconduct <- subset(base_data_tokenized, code.subject == "6_misconduct", select = (1:28))
-#base_data_tokenized.narrative <- subset(base_data_tokenized, code.subject == "12. Narrative Comments", select = (1:28))
-#base_data_tokenized.sciname <- subset(base_data_tokenized, code.subject == "Scientific Name", select = (1:28))
-```
 
-For each subset, filter stopwords, count words and total, then calculate frequencies.
-
-```{r}
+#For each subset, filter stopwords, count words and total, then calculate frequencies.
 sortedwords.basic <- base_data_tokenized.basic %>%
@@
 )
-```
-
-```{r}
+#Repeats for subset 2
 sortedwords.writing <- base_data_tokenized.writing %>%
@@
 
-#Calculate the number of words in total in the subset
 total_words.writing <- sortedwords.writing %>% 
@@
 
-#Mutate table to calculate then append fractional word frequency in the text.
 sortedwords.writing.fraction <- sortedwords.writing %>% as_tibble() %>% mutate(
@@
 
+#Repeats for subset 3
 sortedwords.technical <- base_data_tokenized.technical %>%
@@
 
-#Calculate the number of words in total in the subset
 total_words.technical <- sortedwords.technical %>% 
@@
 
-#Mutate table to calculate then append fractional word frequency in the text.
 sortedwords.technical.fraction <- sortedwords.technical %>% as_tibble() %>% mutate(
@@
 
+#Repeats for subset 4
 sortedwords.logic <- base_data_tokenized.logic %>%
@@
 
-#Calculate the number of words in total in the subset
 total_words.logic <- sortedwords.logic %>% 
@@
 
-#Mutate table to calculate then append fractional word frequency in the text.
 sortedwords.logic.fraction <- sortedwords.logic %>% as_tibble() %>% mutate(
@@
 
-Next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed.
-
-**WARNING**: if a word is MISSING from **any subset** the data for that word is pruned from the full dataset. I am losing the unique words this way. **Need to figure out how to join without pruning.**
-
+The next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed. 
 ```{r}
@@
 
-The following six graphs compare frequencies for ~770 words that appear in ALL THREE of the sub-categories.
-
-Graphs 1-3 compare frequencies in individual groups against word frequency overall.
-Generally, the frequency of a term in the whole set of comments is similar to frequency in an individual group (X vs Y positions similar.) However, some terms associated with Logic subgroup clearly fall below the trendline Graph 3, suggesting they are less frequent in that sub-category than overall.
-
+The following six graphs compare frequencies for ~770 words that appear in ALL three sub-categories: Writing, Technical, and Logic.
 ```{r}
@@
   geom_abline(color = "red")
-```
 
-Graphs 4-6 are pair-wise comparisons of frequencies for sub-categories.
-Unlike the comparisons to overall frequency, most words are NOT clustered along the trendline, suggesting there are greater differences in word usage frequencies.
-
-```{r}
 ggplot(total6, aes(`technical.fraction`,`writing.fraction`)) +
@@
 
-####\
+Graphs 1-3 show frequencies in individual groups versus word frequency overall.
+Generally, frequency of terms in the whole set of comments should be similar to frequency in an individual group (X vs Y positions similar.) However, some terms associated with Logic subgroup (Graph 3) are far from the best-fit line, suggesting they are under-used for comments in that sub-category.
 
-##Log-Odds Ratio for Text Comparison Between Text Categories
-The log-odds ratios method outlined below finds and displays differences words that are more frequently found in TA comments belonging to one sub-category vs.
-another.
-I based this on a method for comparing word usage frequencies between two separate Twitter users [https://www.tidytextmining.com/twitter.html](that I found here.) This page has two parts, one for the comparison for code.subject, and a second part for code.structure category.
+Graphs 4-6 show pair-wise comparisons between sub-categories.
+Unlike overall frequency, most words are NOT clustered along the best-fit line, suggesting there is still more variation in word use between sub-categories.
 
@@
 
-###Part 1. Log-Odds Ratio Comparisons Between Subject Categories
-For simplicity I reduced my analysis to just the four most common comment subjects. This has been the most informative preliminary visualization I have found so far.
+#Log-Odds Ratios
+##Which Words Are Shared LEAST Between Categories?
+Unlike the preceding analysis, log-odds ratio comparison does not eliminate unshared terms.
+Instead, it identifies which words are MOST likely to be used in one sub-category, but not another.
+For simplicity I reduced the analysis to just the four most common comment subjects.
+This has been the most informative preliminary visualization I have found so far.
+
+This code block is based on a method for comparing word usage frequencies between two separate Twitter users [https://www.tidytextmining.com/twitter.html](described here.) If the words used by TAs in their comments differ between sub-categories, there should be fairly low Pearson correlations between word frequencies.
 
@@
 
-```
-
-####Pearson Correlations of Log-Odds Word Frequencies
-If the words used by TAs in their comments differ between the categories, there should be fairly low correlations overall.
-
-```{r}
 #Block below re-organizes the data for comparisons. This pattern sets "2_writing" as the first dataset, then all other data in remaining columns.
@@
 WHICH subset of X values to use is coded in Column 3.
-```
 
-**Pearson correlation** between frequency of words in 2_writing versus 3_technical and 2_writing versus 4_logic sub-categories.
-
-```{r}
+#Pearson correlation between frequency of words in 2_writing versus 3_technical and 2_writing versus 4_logic sub-categories.
 cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3_technical",], 
          ~ proportion + `2_writing`)
-```
 
-```{r}
 cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4_logic",], 
          ~ proportion + `2_writing`)
-```
 
-```{r}
 coroutput23 <-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3_technical",], 
@@
          ~ proportion + `2_writing`)
-broom::tidy(coroutput23)
-broom::tidy(coroutput24)
 
-```
+#Comparing sub-groups 3_technical and 4_logic requires recreating the dataframe in a slightly different format that excludes the data rows for 2_writing.
 
-```{r}
-knitr::kable(coroutput23)
-```
-
-Comparing sub-groups 3_technical and 4_logic requires recreating the dataframe in slightly different format that excludes the data rows for 2_writing.
-
-```{r}
 # Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
@@
 
-```
-
-```{r}
 #Block below re-organizes the data for comparisons. This pattern sets "3_technical" as the first dataset, then all other data in remaining columns.
@@
 WHICH subset of X values to use is coded in Column 3.
+
+#Pearson correlation** between the frequency of words in 3_technical versus 4_logic sub-categories.
+coroutput34 <-cor.test(data = base_data_tokenized_sorted3[base_data_tokenized_sorted3$code.subject == "4_logic",], 
+         ~ proportion + `3_technical`)
+
+broom::tidy(coroutput23)
+broom::tidy(coroutput24)
+broom::tidy(coroutput34)
 ```
 
-**Pearson correlation** between the frequency of words in 3_technical versus 4_logic sub-categories.
+####\
 
-```{r}
-cor.test(data = base_data_tokenized_sorted3[base_data_tokenized_sorted3$code.subject == "4_logic",], 
-         ~ proportion + `3_technical`)
-```
-
-####Visualizing Word Frequency Differences Between Subject Sub-Categories
+###Visualizing Frequency Differences Between Sub-Categories
 
@@
   mutate(freq = n/total)
-```
 
-```{r}
 base_data_tokenized_sorted2v3 <- base_data_tokenized_sorted2 %>% 
@@
   arrange(`2_writing`,`3_technical`)
-```
 
-```{r}
 ggplot(base_data_tokenized_sorted2v3, aes(`2_writing`,`3_technical`)) +
@@
   #ggsave("Freqplot_3vs2.png", width = 10, height = 6)
-```
 
-```{r}
 #Comparing 2 vs. 4.
@@
   arrange(`2_writing`,`4_logic`)
-```
 
-```{r}
 ggplot(base_data_tokenized_sorted2v3, aes(`2_writing`,`4_logic`)) +
@@
   #ggsave("Freqplot_4vs2.png", width = 10, height = 6)
-```
 
-```{r}
 #Create groups by code.subject, calculates proportional frequency in each group.
@@
   mutate(freq = n/total)
-```
 
-```{r}
 #Set up comparison for 3 vs. 4.
@@
   arrange(`2_writing`,`4_logic`)
-```
 
-```{r}
 ggplot(base_data_tokenized_sorted3v4, aes(`3_technical`,`4_logic`)) +
@@
 
-####Calculating and Plotting Log-Odds Pairs for Subject Sub-Categories
+###Calculating and Plotting Log-Odds Pairs for Subject Sub-Categories
 
@@
   arrange(desc(logratio))
-```
 
-```{r}
 word_ratios2v3 %>% 
   arrange(abs(logratio))
-
 word_ratios3v4 %>% 
   arrange(abs(logratio))
-
 word_ratios2v4 %>% 
@@
   group_by(logratio < 0) %>%
-  top_n(20, abs(logratio)) %>%
+  top_n(25, abs(logratio)) %>%
   ungroup() %>%
@@
 
-####\
+#N-Grams Mapping 
+##Which Terms, Phrases Are Unique to a Sub-Category?
 
-###Part 2. Log-Odds Ratio Comparisons Between Structure Categories
+```{}
+#Pseudo-code
+Import comments
+Tokenize the comments as 1-, 2, 3-grams
+Add counts within subset, calculate frequency
+Split by subcategory
+Align 4 subcategories, highest to lowest
+
+```
 
 ####\  
+```{r}
+# Call in tidyverse, other required libraries.
+library(tidyverse)
+library(tidytext)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(scales)
+
+#Read in CSV as a starting dataframe named "base_data". 
+base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
+```
+
+```{r}
+#Tokenize phrases as N-grams.
+
+base_data_tokenized2n <- base_data %>% 
+  unnest_tokens(ngram, ta.comment, token="ngrams", n=2)
+
+sortedwords.comments.all2n <- base_data_tokenized2n %>%
+  count(ngram, sort = TRUE) %>%
+  ungroup()
+
+total_words.all2n <- sortedwords.comments.all2n %>% 
+  summarize(total = sum(n))
+
+sortedwords.comments.all.fraction2n <- sortedwords.comments.all2n %>% as_tibble() %>% mutate(
+  all.fraction = (n*100)/total_words.all2n$total
+  )
+
+#Select SUBSETs of tokenized data based on code.subject: 
+base_data_tokenized.basic2n <- subset(base_data_tokenized2n, code.subject == "1_basic", select = (1:28))
+base_data_tokenized.writing2n <- subset(base_data_tokenized2n, code.subject == "2_writing", select = (1:28))
+base_data_tokenized.technical2n <- subset(base_data_tokenized2n, code.subject == "3_technical", select = (1:28))
+base_data_tokenized.logic2n <- subset(base_data_tokenized2n, code.subject == "4_logic", select = (1:28))
+
+#For each subset, filter stopwords, count words and total, then calculate frequencies.
+sortedwords.basic2n <- base_data_tokenized.basic2n %>%
+  count(ngram, sort = TRUE) %>%
+  ungroup()
+
+#Calculate the number of words in total in the subset
+total_words.basic2n <- sortedwords.basic2n %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.basic.fraction2n <- sortedwords.basic2n %>% as_tibble() %>% mutate(
+  basic.fraction2n = (n*100)/total_words.basic2n$total
+)
+
+#Repeats for subset 2
+#For each subset, filter stopwords, count words and total, then calculate frequencies.
+sortedwords.writing2n <- base_data_tokenized.writing2n %>%
+  count(ngram, sort = TRUE) %>%
+  ungroup()
+
+#Calculate the number of words in total in the subset
+total_words.writing2n <- sortedwords.writing2n %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.writing.fraction2n <- sortedwords.writing2n %>% as_tibble() %>% mutate(
+  writing.fraction2n = (n*100)/total_words.writing2n$total
+)
+
+#Repeats for subset 3
+#For each subset, filter stopwords, count words and total, then calculate frequencies.
+sortedwords.technical2n <- base_data_tokenized.technical2n %>%
+  count(ngram, sort = TRUE) %>%
+  ungroup()
+
+#Calculate the number of words in total in the subset
+total_words.technical2n <- sortedwords.technical2n %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.technical.fraction2n <- sortedwords.technical2n %>% as_tibble() %>% mutate(
+  technical.fraction2n = (n*100)/total_words.technical2n$total
+)
+
+#Repeats for subset 4
+#For each subset, filter stopwords, count words and total, then calculate frequencies.
+sortedwords.logic2n <- base_data_tokenized.logic2n %>%
+  count(ngram, sort = TRUE) %>%
+  ungroup()
+
+#Calculate the number of words in total in the subset
+total_words.logic2n <- sortedwords.logic2n %>% 
+  summarize(total = sum(n))
+
+#Mutate table to calculate then append fractional word frequency in the text.
+sortedwords.logic.fraction2n <- sortedwords.logic2n %>% as_tibble() %>% mutate(
+  logic.fraction2n = (n*100)/total_words.logic2n$total
+)
+```

### 2019-02-22 background.Rmd
updates

@@
 ##Context & Prior Work
-From prior studies (REFs), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead.
-To that end, our program trains graduate TAs to use a **global issues first, specific issues later** approach to correcting undergraduate technical writing.
-We ask them to spend a shorter amount of time writing fewer comments that focus on broader issues.
-Ideally, TAs provide no more than 3-5 comments per text page, and the comments given should encourage student reflection.
-As students' writing improves, TAs can focus on smaller issues.
-Our bins-based report grading protocol reinforces this strategy.
+From prior studies (*see the Literature Reveiw for details*), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead.
+To that end, our program trains graduate TAs to use a **global issues / biggest issues first** approach to correcting undergraduate technical writing.
+Ideally:
 
-Our coaching-oriented approach is well-supported by prior research, but we are uncertain how accurately TAs implement the recommended strategy.
+*  TAs spend a shorter amount of time writing fewer comments that focus on broader issues.
+*  TAs provide no more than 3-5 comments per text page.
+*  Comments encourage student reflection and self-correction rather than prescribe specific changes. 
+*  As students' writing improves, TAs shift their focus to the next most-significant errors or issues.
+
+Our bins-based report grading protocol reinforces this feedback strategy.
+
+This holistic, coaching-oriented approach is well-supported by prior research, but we are uncertain how accurately TAs implement the recommended strategy.
 We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"** Self-reported TA effort is likely to be unreliable.
-What we need is a more direct measure of effort.
+What is needed is a more direct measure of TA effort.
 
-In 2018, we began trying to analyze TAs' comments from student papers directly.
-We extracted and tabulated all comments from approximately 500 pairs of reports.
+In 2018, we began trying to analyze TAs' comments on student papers directly.
+We extracted and tabulated all comments from ~500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses during a single semester.
 After initial data cleaning (removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent), we had a working dataset of ~11,000 separate comments.
@@
 
-* **Subject**: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
-* **Structure**: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?
+* __Subject__: did the comment focus mainly on **Basic Criteria** (minimum pass/fail requirements), **Technical Issues, Writing Quality, Logic and Thinking,** or **Other** issues
+* __Structure__: did the comment **Point** to an error without providing other help, provide simple **Copy Correction**, provide declarative **General or Specific Information** only, or provide **Holistic Coaching** that fosters student thinking and long-term improvement?
 
-Raw counts for each subject/structure pair were then converted to fractional frequencies. The final frequency table (excerpted below) provided us with a direct estimate of TA effort and emphasis.
+Raw counts of comments for each Subject/Structure pair were then converted to fractional frequencies.
+The final frequency table (excerpted below) provided us with direct estimates of the most common types of comments TAs made on student work, and of which elements of writing TAs emphasize most often.
 
@@
 
-##Central Problem
-Analyzing TA comments this way is informative but coding by hand is unsustainable.
+##The Challenge
+First, analyzing TA comments this way is informative but coding by hand is unsustainable.
 Developing the codebook then rating the initial set of 11,000 comments required nearly 80 hours of investigator effort.
-Even with the codebook in hand, individual coders need training to achieve sufficient accuracy, and there is significant risk of "coding drift" over time.
+Given 10-12,000 comments are generated EACH SEMESTER, a faster classification method is essential for this NSF project to succeed.
 
-Given 10-12,000 comments are generated EACH SEMESTER, an automated classification method is needed.
+Second, even with the codebook in hand, individual coders need training to achieve sufficient accuracy, and there is significant risk of "coding drift" (changes in how code features are interpreted by a single rater over time).
+Automated classification would eliminate coder training and minimize potential for coding drift.
 
 ####\
-
-##Work Plan
-The work plan for this project is:
-
-1.  Compile initial dataset and de-identify both TAs and students by replacing names with anonymous IDs.
-2.  Import TA comments and relevant metadata from student reports into R as a "tidy" dataframe.
-3.  Explore comment data structure, n-gram frequencies, etc., to identify potentially useful elements for feature engineering.
-4.  Write an initial Naive Bayes workflow for supervised text classification.
-5.  Assign TA comments to categories using the NB classifier, then compare outcomes from automated classification to hand-coded results.
-6.  Using features (from Step #3) and comparison data from initial classification protocol (from Step #4) as guides, test iterations and optimize the classifier method. 
-7.  Apply the optimized classifier to the original comment dataset, identify all comments that are being classified incorrectly, and look for potential patterns.
-8.  Write a small set of rule- or REGEX pattern-based searches that can identify comments which are more likely to be classified incorrectly. 
-9.  Append rules/pattern-based pre-screening to the optimized Naive Bayes classifier.
-10.
-Validate this vertical Ensemble Classifier by classifying a subset of ~2000 new TA comments extracted from a non-overlapping set of student lab reports.
-Compare automated results for new data set to hand-coded categories.
-
-####\

### 2019-02-22 build.Rmd
updates

---
title: "The Build"
author: "Dan Johnson"
date: "2/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Overview
This page outlines the process for building various components used in the project. Details of each  component are provided in these sub-pages.

Structure of the initial dataset: https://adanieljohnson.github.io/default_website/datastructure.html

Import and post-import data validation: https://adanieljohnson.github.io/default_website/dataimport.html

####\

#Describe an Element

####\

#Describe an Element

####\

### 2019-02-22 dataimport.Rmd
updates

--- 2019-02-05 draft.Rmd
@@
 ---
-title: "Data Import and Initial Exploration"
+title: "Data Import & Checking"
 author: "Dan Johnson"
-date: "12/17/2018"
+date: "2/22/2019"
 output: html_document
@@
 ```
+
 #Overview
-This page has three sections.
-**Part 1** describes the target dataset for this project.
-**Part 2** outlines the data pre-checks and import process.
-**Part 3** describes post-import validation developed for the target dataset, and how selected elements within the dataset are recoded for analysis.
+This page outlines the process for building various components used in the project. Details of each  component are provided in these sub-pages.
 
-Initial visual exploration of the dataset is described here: https://adanieljohnson.github.io/default_website/initial_exploration.html
+Structure of the initial dataset: https://adanieljohnson.github.io/default_website/datastructure.html
 
-These linked pages summarize work to date using categorical Naive Bayes as the text classification method. They provided a baseline from which I am evaluating other approaches.
-
-*  Baseline Method: https://adanieljohnson.github.io/default_website/Naive_bayesV2b.html
-*  Randomization Method: https://adanieljohnson.github.io/default_website/Naive_bayesV2r.html
-*  In/Outgroup Binning: https://adanieljohnson.github.io/default_website/Naive_bayesV2yn.html
-*  In/Outgroup Binning: https://adanieljohnson.github.io/default_website/Naive_bayes_Smry.html
-
-Other pages provide the theoretical background for the selected methods.
-Each deeper analysis method is described in detail on its own separate page.
-These are listed on the **[LINK] Project Outline** page.
+Import and post-import data validation: https://adanieljohnson.github.io/default_website/dataimport.html
 
@@
 
-#Part 1. Initial Dataset
+#Describe an Element
+
+####\
+
+#Describe an Element
+
+####\
+
+#Overview
+This page describes the target dataset, data pre-checks and import process, post-import validation, and how selected elements are recoded for analysis.
+
+Preliminary exploration of the dataset is described here: https://adanieljohnson.github.io/default_website/initial_exploration.html
+
+Rationale and initial steps in development of the Classifier are described here: https://adanieljohnson.github.io/default_website/bayes_history.html
+
+Code for the current version of the Naive Bayes Classifier is located here:
+https://adanieljohnson.github.io/default_website/NB_2r2c.html
+
+####\
+
+#Target Dataset
 The development dataset is an Excel table of 10,808 rows.
@@
 
-#Part 2. Data Pre-Checks
-Prior to import, review the data table **in Excel**; check that:
+#Pre-Check and Import
+Prior to import, review the data table **in Excel** and check that:
 
@@
 
-Most in-depth analyses will have a similar extraction and post-processing code block.
-In practice, it has proven easier to modify the post-processing code than to maintain multiple separate copies of the original CSV dataset for the different analyses.
+Most analyses will use a similar extraction and post-processing code block.
+In practice, it is easier to modify the post-processing code than to maintain multiple separate copies of the original CSV dataset for different analyses.
+
+####\

### 2019-02-22 datastructure.Rmd
updates

--- 2019-02-22 dataimport.Rmd
@@
 ---
-title: "Data Import & Checking"
+title: "Data Structure"
 author: "Dan Johnson"
@@
 output: html_document
-bibliography: library.bib
-biblio-style: apalike
 ---
-```{r, child="_setup.Rmd"}
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
@@
 #Overview
-This page describes the target dataset, data pre-checks and import process, post-import validation, and how selected elements are recoded for analysis.
+This page describes how the target dataset is structured.
 
-Preliminary exploration of the dataset is described here: https://adanieljohnson.github.io/default_website/initial_exploration.html
+How data are imported then validated for analysis is described here: https://adanieljohnson.github.io/default_website/dataimport.html
 
@@
 #Target Dataset
-The development dataset is an Excel table of 10,808 rows.
-One column has "TA comments" extracted from student lab reports in Spring 2018.
+The dataset being used to develop this project is an Excel table containing 10,808 TA comments extracted from student lab reports submitted during Spring 2018.
 Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation.
 Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.
-
-Data were anonymized in Excel prior to conversion to CSV format. Names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings **[LINK]**.
 
@@
 
+Data were anonymized in Excel, replacing names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings **[LINK]**.
+
 Once all data are in correct format, export Excel file to CSV with name formatted as **"coded_full_comments_dataset_SemYear.csv"**
-
-####\
-
-##Initial Library and Data Import
-```{r Libraries_1}
-# Call in tidyverse, other required libraries.
-library(tidyverse)
-library(tidytext)
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-library(scales)
-```
-
-```{r}
-#This block reads in full CSV into a starting dataframe named "base_data".
-#Subsequent analyses are performed on extracted subsets of "base_data".
-base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
-```
-
-If dataset imports correctly, there will be a dataframe of ~11,000 rows and 28 columns named as listed in the **Data Pre-Checks** table above.
-
-####\
-
-#Part 3. Post-Import Checks
-Given some data points may be entered incorrectly by instructors or students, these post-import data checks help ensure data have been properly coded and columns contain valid entries only.
-(This process can be automated using the ```janitor``` data cleaning package.)
-
-```{r}
-#"Course" column should only have entries that correspond to 113, 114, 214, NA. 
-course<-unique(base_data$course)
-
-#Entries in "TAs" column should match the anonymous IDs that correspond to the TAs assigned to the relevant courses in the semester being analyzed. 
-#If there are incorrect or extra anonymous IDs in this list, the mostly likely source of the error is improper de-identification of the Excel file prior to import.
-ta<-unique(base_data$ta)
-
-#"Lab" should only have entries that match the allowed limited keywords in codebook.
-lab<-unique(base_data$lab)
-
-#It is impractical to validate all randomized identifiers.
-Instead look at the NUMBER of unique anonymous student IDs.
-Quantity should be within 10% of (but not more than) the combined enrollment of the 3 target courses for the semester.
-student<-unique(base_data$student)
-
-#Check that the list of "code.subject" topics here matches allowed terms in codebook.
-subject<-unique(base_data$code.subject)
-
-#Check that the list of "code.structure" terms extracted from the dataset and displayed here matches allowed terms in codebook.
-structure<-unique(base_data$code.structure)
-```
-####\
-
-##Post-Import Processing
-Comment labels used in the project codebook contain punctuation and spacing that complicates subsequent analysis.
-Rather than modify the project codebook or external processes, it is simpler to recode the problematic labels after import.
-For example, the following code extracts a subset of rows from the full dataset, reduces those to 3 columns, and renames the **comment.subject** column as **subject**.
-Four problematic code labels in the subject column are replaced with simpler terms.
-
-```{r}
-#Read in TA comments from CSV file.
-base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
-
-#Select rows representing the sub-groups to compare. 
-comments_subset <- filter(base_data,code.subject=="1. Basic Criteria"|code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
-
-#Reduce larger dataframe to 3 required columns of data (subject, structure, and text of comment), and put columns in order needed.
-comments_raw <- comments_subset %>% select(23,24,22)
-
-#Rename the columns.
-names(comments_raw)[1] <- "subject"
-names(comments_raw)[2] <- "structure"
-names(comments_raw)[3] <- "text"
-
-#Simplify coding terms used in "subject" column.
-comments_raw[,1] <- ifelse(comments_raw[,1] == "1.
-Basic Criteria","1_basic", ifelse(comments_raw[,1] == "2.
-Writing Quality","2_writing", ifelse(comments_raw[,1] == "3.
-Technical and Scientific","3_technical", ifelse(comments_raw[,1] == "4.
-Logic and Thinking","4_logic",99))))
-
-#Change "subject" element from character to a factor for analysis.
-comments_raw$subject <- factor(comments_raw$subject)
-str(comments_raw$subject)
-table(comments_raw$subject)
-```
-
-Most analyses will use a similar extraction and post-processing code block.
-In practice, it is easier to modify the post-processing code than to maintain multiple separate copies of the original CSV dataset for different analyses.
-
-####\

### 2019-02-22 index.Rmd
updates

@@
 
-In Summer 2018, we extracted and tabulated ~11,000 comments made by teaching assistants (TA) from >500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses during a single semester.
-Using an iteratively generated codebook, each of the comments was separately classified in two categories:
-
-* __Subject__: did the comment focus mainly on **Basic Criteria, Technical Issues, Writing Quality, Logic and Thinking,** or **Other** issues
-* __Structure__: did the comment **Point** to an error without providing other help, provide simple **Copy Correction**, provide declarative **General or Specific Information** only, or provide **Holistic Coaching** that fosters student thinking and long-term improvement?
-
-#### \   
-Counts of coded TA comments were tabulated according to subject and structure, and frequencies for each subject/structure pair calculated as a fraction of all comments.
-This 2-way contingency table provided us with a direct estimate of the most common types of comments TAs made on student work, and a relative estimate of which elements of writing TAs emphasize most often.
-
-#### \   
-##Specific Goals
 Classifying TA comments this way is informative, but doing so by hand is impractical: 10-12,000 comments need to be scored EACH semester.
@@
 #### \   
+##Work Plan
+The work plan for this project is:
+
+1.  Compile an initial dataset for testing, and de-identify both TAs and students by replacing names with anonymous IDs.
+2.  Import TA comments and relevant metadata from student reports into R as a "tidy" dataframe.
+3.  Use basic text analysis methods to explore comment data structure, n-gram frequencies, etc., and identify potentially useful elements for feature engineering.
+4.  Write a baseline workflow using Naive Bayes for supervised text classification.
+5.
+Assign TA comments to categories using the NB classifier, then establish baseline accuracy values using the optimized NB classifier by comparing outcomes of automated classification to hand-coded results.
+6.  Using features (from Step #3) and comparison data from NB classification protocol (from Step #5) as guides, further optimize the NB classifier method. 
+7.  When accuracy stabilizes, apply the optimized NB classifier to the original FULL comment dataset, identify all comments that are being classified incorrectly, and look for potential patterns.
+8.  Write a small set of rule- or REGEX pattern-based searches that can identify comments which are more likely to be classified incorrectly. 
+9.  Append rules/pattern-based pre-screening to the optimized Naive Bayes classifier to create a vertical Ensemble Classifier.
+10.
+Validate the Ensemble Classifier using a subset of ~2000 new hand-coded TA comments extracted from a non-overlapping set of student lab reports.
+Compare automated results for new data set to hand-coded categories.
+
+Looking beyond the current project, I will be using the NB-based Ensemble Classifier as the control method for evaluating other text classifier strategies.
+
+#### \   
 ##Project Site Map

### 2019-02-22 initial_exploration.Rmd
updates

@@
 ```
+
+#Overview
+This page outlines the process for building various components used in the project. Details of each  component are provided in these sub-pages.
+
+Structure of the initial dataset: https://adanieljohnson.github.io/default_website/datastructure.html
+
+Import and post-import data validation: https://adanieljohnson.github.io/default_website/dataimport.html
+
+####\
+
+#Describe an Element
+
+####\
+
+#Describe an Element
+
+####\
 

### 2019-02-22 build.Rmd
update

@@
 #Overview
-This page outlines the process for building various components used in the project. Details of each  component are provided in these sub-pages.
+This page links to sub-pages that describe how various components used in the project were built. Follow the links for details.
 
-Structure of the initial dataset: https://adanieljohnson.github.io/default_website/datastructure.html
+Extracting TAs' comments: https://adanieljohnson.github.io/default_website/docxcommentextraction.html
 
-Import and post-import data validation: https://adanieljohnson.github.io/default_website/dataimport.html
+Structure of development dataset; data import & validation: https://adanieljohnson.github.io/default_website/dataimport.html
+
+Current Naive Bayes classifier: https://adanieljohnson.github.io/default_website/NB_2r2c.html
 

### 2019-02-22 dataimport.Rmd
update

@@
 #Overview
-This page outlines the process for building various components used in the project. Details of each  component are provided in these sub-pages.
-
-Structure of the initial dataset: https://adanieljohnson.github.io/default_website/datastructure.html
-
-Import and post-import data validation: https://adanieljohnson.github.io/default_website/dataimport.html
-
-####\
-
-#Describe an Element
-
-####\
-
-#Describe an Element
-
-####\
-
-#Overview
 This page describes the target dataset, data pre-checks and import process, post-import validation, and how selected elements are recoded for analysis.
 
-Preliminary exploration of the dataset is described here: https://adanieljohnson.github.io/default_website/initial_exploration.html
-
-Rationale and initial steps in development of the Classifier are described here: https://adanieljohnson.github.io/default_website/bayes_history.html
-
-Code for the current version of the Naive Bayes Classifier is located here:
-https://adanieljohnson.github.io/default_website/NB_2r2c.html
-
-####\
-
-#Target Dataset
-The development dataset is an Excel table of 10,808 rows.
-One column has "TA comments" extracted from student lab reports in Spring 2018.
+The dataset being used to develop this project is an Excel table containing 10,808 TA comments extracted from student lab reports submitted during Spring 2018.
 Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation.
 Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.
-
-Data were anonymized in Excel prior to conversion to CSV format. Names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings **[LINK]**.
 
@@
 
+Data were anonymized in Excel, replacing names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings **[LINK]**.
+
 Once all data are in correct format, export Excel file to CSV with name formatted as **"coded_full_comments_dataset_SemYear.csv"**
@@
 
-##Initial Library and Data Import
+#Target Dataset
+The development dataset is an Excel table of 10,808 rows.
+One column has "TA comments" extracted from student lab reports in Spring 2018.
+Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation.
+Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.
+
+Data were anonymized in Excel prior to conversion to CSV format. Names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings **[LINK]**.
+
+####\
+
+#Pre-Checks Before Import
+Prior to import, review the data table **in Excel** and check that:
+
+1.  Student and TA names have been recoded using pre-assigned anonymous IDs.
+2.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.rmd" [LINK]
+3.  Data table headers match those listed below. The same list is in the file "code_column_names.rmd" [LINK]
+
+Column Number|Column Name|Accepted Values or Format
+------------|----------------------|-----------------------
+1|unique.record|Sp18.00001 (semester and year, then order in dataset)
+2|report.id|R_0fanqHFhaY2yLT7 or similar
+3|sort|remnant of coding
+4|report.title (empty)|string
+5|student|Std_drVZ2x1W (format is Std\_nnn)
+6|course|113, 114, or 214
+7|ta |TA_bc8EVX09 (format is TA\_nnn)
+8|lab|113: allocation, betta; 114: manduca, frog; 214: enzymes (old) photosynthesis, physarum, cyanophage (new)
+9|tag|first, second
+10|type.TA|submission, revision
+11|grade.TA|A, B, C, F
+12|grading.time|time in minutes
+13|Rank|remnant of coding
+14|hypothesis.ok|Yes, No
+15|data.ok|Yes, No
+16|citation.ok|Yes, No
+17|interpretation.ok|Yes, No
+18|organization.ok|Yes, No
+19|techflaws.ok|Yes, No
+20|writing.ok|Yes, No
+21|comments.incorporated|Yes, Somewhat, No
+22|ta.comment|text string
+23|code.subject|values in Ref. Tables
+24|code.structure|values in Ref. Tables
+25|code.locus|values in Ref. Tables
+26|code.scope|values in Ref. Tables
+27|code.tone|values in Ref. Tables
+28|code.notes|values in Ref. Tables
+
+Once all data are in correct format, export Excel file to CSV with name formatted as **"coded_full_comments_dataset_SemYear.csv"**
+
+####\
+
+#Initial Data Import
 ```{r Libraries_1}
@@
 
-#Part 3. Post-Import Checks
-Given some data points may be entered incorrectly by instructors or students, these post-import data checks help ensure data have been properly coded and columns contain valid entries only.
+#Post-Import Checks
+Given some data points may be entered incorrectly by instructors or students, these post-import data checks ensure data are properly coded and columns contain valid entries only.
 (This process can be automated using the ```janitor``` data cleaning package.)
@@
 course<-unique(base_data$course)
+```
 
+```{r}
 #Entries in "TAs" column should match the anonymous IDs that correspond to the TAs assigned to the relevant courses in the semester being analyzed. 
@@
 ta<-unique(base_data$ta)
+```
 
+```{r}
 #"Lab" should only have entries that match the allowed limited keywords in codebook.
 lab<-unique(base_data$lab)
+```
 
+```{r}
 #It is impractical to validate all randomized identifiers.
@@
 student<-unique(base_data$student)
+```
 
+```{r}
 #Check that the list of "code.subject" topics here matches allowed terms in codebook.
 subject<-unique(base_data$code.subject)
+```
 
+```{r}
 #Check that the list of "code.structure" terms extracted from the dataset and displayed here matches allowed terms in codebook.

### 2019-02-22 docxcommentextraction.Rmd
update

---
title: "DOCX Comment Extraction"
author: "Dan Johnson"
date: "2/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Pseudo-Code and Sources
Create a list of files to cycle through OR put all files to process in ONE folder.
For each file, use 'read_docx', then 'docx_extract_all_cmnts'
Use 'do.call("rbind", list(dataframes to merge))' to cycle through and append the extracted comments to ONE dataframe

https://stackoverflow.com/questions/50439684/how-to-extract-plain-text-from-docx-file-using-r

https://stackoverflow.com/questions/5965698/merge-unequal-dataframes-and-replace-missing-rows-with-0

https://stat.ethz.ch/R-manual/R-devel/library/base/html/merge.html

https://www.dummies.com/programming/r/how-to-use-the-merge-function-with-data-sets-in-r/

https://datascienceplus.com/combining-data-in-r-the-skill-of-merging-joining-and-stacking/

https://www.rdocumentation.org/packages/analogue/versions/0.17-1/topics/join
Load required Libraries
```{r}
library(docxtractr)
```

```{r}
Commentframe1<-read_docx("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/DOCX_comment_extraction/TestingReport.docx", track_changes = NULL)

Commentframe2<-read_docx("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/DOCX_comment_extraction/TestingCommentExtraction.docx", track_changes = NULL)

comments1<-docx_extract_all_cmnts(Commentframe1, include_text = FALSE)
comments2<-docx_extract_all_cmnts(Commentframe2, include_text = FALSE)

#Merging more than two dataframes
#To merge more than two dataframes within R, we can combine our rbind call with do.call.
# > do.call("rbind", list(dataframes to merge))
#
#How this works:
#  The 'do.call' function allows us to use any existing R functions, such as 'rbind', but instead of passing just two dataframes in this instance, we are able to pass multiple.
#
#Therefore, when used alongside rbind, we can use do.call to merge more than two dataframes at once.
#
#Putting this all together, we can merge our dataframes back together as follows:
# > ufoMerged <- do.call("rbind", list(ufo1, ufo2, ufo3, ufo4))
#
#We can confirm our data has been successfully merged by checking the number of rows:
#
#> NROW(ufoMerged)
#[1] 80332

#comments1And2 <- rbind(comments1,comments2)
do.call("rbind", list(comments1,comments2))

write.csv2(comments1And2, file = "TAcomments.csv")
```

### 2019-02-22 hiding/datastructure.Rmd
update

--- 2019-02-22 datastructure.Rmd
@@
 ---
-title: "Data Structure"
+title: "DROPPED"
 author: "Dan Johnson"
@@
 #Overview
-This page outlines the process for building various components used in the project. Details of each  component are provided in these sub-pages.
-
-Structure of the initial dataset: https://adanieljohnson.github.io/default_website/datastructure.html
-
-Import and post-import data validation: https://adanieljohnson.github.io/default_website/dataimport.html
-
-####\
-
-#Describe an Element
-
-####\
-
-#Describe an Element
-
-####\
-
-#Overview
-This page describes how the target dataset is structured.
-
-How data are imported then validated for analysis is described here: https://adanieljohnson.github.io/default_website/dataimport.html
-
-Rationale and initial steps in development of the Classifier are described here: https://adanieljohnson.github.io/default_website/bayes_history.html
-
-Code for the current version of the Naive Bayes Classifier is located here:
-https://adanieljohnson.github.io/default_website/NB_2r2c.html
-
-####\
-
-#Target Dataset
 The dataset being used to develop this project is an Excel table containing 10,808 TA comments extracted from student lab reports submitted during Spring 2018.

### 2019-02-22 initial_exploration.Rmd
update

--- 2019-02-22 initial_exploration.Rmd
@@
 ---
-title: "Initial Explorations"
+title: "Initial Exploration"
 output: html_document
@@
 
-The first question to answer before making deeper comparisons is **are TAs' comments obviously different between sub-categories?** To that end, I ran three different analyses of word usage in TA comments.
+#Overview
+The first question to answer before making deeper comparisons is, **are word frequencies different between TAs' comments from different sub-categories?** To that end, I ran three different analyses of word usage frequency.
 
-*  Pearson correlations for word use frequencies. This compares words found in all Subject sub-categories
-*  Word frequency differences based on log-odds ratios. This highlights terms that are the most different between sub-categories.
-*  Unique n-grams mapping. This identifies text strings, phrases, and rare terms that are unique (or nearly unique) to one sub-category.
+*  __Pearson correlations for word use frequencies__. This only compares words found in all of the Subject sub-categories
+*  __Word frequency differences based on log-odds ratios__. This highlights terms that are the most different between sub-categories.
+*  __N-grams mapping__. This identifies text strings, phrases, and terms that are unique (or nearly unique) to one sub-category.
+
+####\
 
@@
 #Pearson Correlations of Word Frequencies
-##Are Shared Words Unequally Distributed?
+##Are Shared Words Used At Different Frequencies?
 This analysis takes all comments from base_data, tokenizes and filters stopwords, then counts and sorts fractional frequencies for each word in each category.
@@
 
-#Log-Odds Ratios
+#Log-Odds Ratios of Word Frequencies
 ##Which Words Are Shared LEAST Between Categories?
@@
 #N-Grams Mapping 
-##Which Terms, Phrases Are Unique to a Sub-Category?
+##Are Certain Terms and Phrases Unique to One Sub-Category?
+One of our earliest observations was that the n-gram "basic criteria" only appeared in TA comments from three different Subject sub-categories.
+This suggests certain terms and n-grams might be eliminated
 
+However it w  the relative frequency was much higher
 ```{}

### 2019-02-23 Naive_Bayes_Pkg_test.Rmd
update

---
title: "Naive Bayes Package Test"
author: "Dan Johnson"
date: "1/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Trying out Naive Bayes Package

Source of package:
CRAN repository

```{r}
library(tidyverse)
```

```{r}
glimpse(iris)

```

```{r}
data(iris)
nb <- naive_bayes(Species ~ ., data = iris)
plot(nb)

```

```{r}
nb_kernel <- naive_bayes(x = iris[-5], y = iris[ ,5], usekernel = TRUE)
plot(nb_kernel)

```

***
BRANCH
***

```{r}
library(naivebayes)
```

```{r}
#Select rows representing the sub-groups to compare. 
comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
comments_rawVERB <- comments_subsetVERB %>% select(23,14,15,16,17,18,19,20)

#Rename the columns.
names(comments_rawVERB)[1] <- "type"

glimpse(comments_rawVERB)

```

```{r}
data(comments_rawVERB)
nb <- naive_bayes(type ~ ., data = comments_rawVERB)
plot(nb)

```

```{r}
nb_kernel <- naive_bayes(x = iris[-5], y = iris[ ,5], usekernel = TRUE)
plot(nb_kernel)

```

***
END BRANCH
***

```{r}
X1 <- as.data.frame(matrix(sample(letters[5:9], vars * rows, TRUE),
                           ncol = vars))
nb_cat <- naive_bayes(x = X1, y = y)
nb_cat
system.time(pred2 <- predict(nb_cat, X1))

```

## Not run: 
vars <- 10
rows <- 500000
y <- sample(c("a", "b"), rows, TRUE)

# Only categorical variables
X1 <- as.data.frame(matrix(sample(letters[5:9], vars * rows, TRUE),
                           ncol = vars))
nb_cat <- naive_bayes(x = X1, y = y)
nb_cat
system.time(pred2 <- predict(nb_cat, X1))

# Only numeric variables
X2 <- as.data.frame(matrix(rnorm(vars * rows), ncol = vars))
nb_num <- naive_bayes(x = X2, y = y)
nb_num
system.time(pred2 <- predict(nb_num, X2))

## End(Not run)

### 2019-02-23 Naive_bayes_notes_5.Rmd
update

--- 2018-12-21 Naive_bayes.Rmd
@@
 ---
-title: "Naive Bayes Workflow"
+title: "Permutations of Naive Bayes"
 author: "Dan Johnson"
@@
 
-##Summary
-This notebook documents a basic workflow for applying a naive Bayes model to terms in TA comments.
-I based it on a tutorial [https://rpubs.com/Billyhansen6/318412](located here).
-This is quoted from Hansen's original tutorial:
->The objective of this project is to classify SMS messages as spam or ham (not spam).
-A Naive Bayes classifier approach will be used.
-This example is taken from chapter 4 of Machine Learning with R, Second Edition"
+##Permutation Tests of Naive Bayes (NB) Method
+Version 1.0 correctly identifies comment subcategories ~79% of the time when using single words, and 68.3% of the time when using 2-grams. Permutations to test:
 
->An example of the conditional probability that will be computed is as follows:
+Uni-grams, four categories
+Lowercase=YES, remove_numbers=no, remove_stopwords=no, stemming=no, remove_whitespace=YES
+Freq_cutoff=5
+  Overall Discrimination: 78.6%
+  Highest Discrimination: 91.4%, 3_Technical
 
-$$P(Spam|Hospital) = P(Hospital|Spam)P(Spam)/P(Hospital)$$
+Bi-grams, four categories
+Lowercase=YES, remove_numbers=no, remove_stopwords=no, stemming=no, remove_whitespace=YES
+Freq_cutoff=5
+  Overall Discrimination: 73.6%
+  Highest Discrimination: 82.7%, 3_Technical
 
->which is the formula for determining the probability that a message is spam given that it contains the word "Hospital" in the message.
+Uni-grams, four categories
+Lowercase=YES, remove_numbers=no, remove_stopwords=YES, stemming=no, remove_whitespace=YES
+Freq_cutoff=5
+  Overall Discrimination: 
+  Highest Discrimination:
 
-Lines ~30-205 are the baseline version of the workflow using single words.
-It correctly classifies comments in the top 4 sub-categories  ~79% of the time.
-While that is close to minimum acceptable inter-rater reliability of two human readers, it is well below my goal of >95% for use in routine automated analysis.
+Bi-grams, four categories
+Lowercase=YES, remove_numbers=no, remove_stopwords=YES, stemming=no, remove_whitespace=YES
+Freq_cutoff=5
+  Overall Discrimination: 
+  Highest Discrimination:
 
-Lines ~208-290 describe an alternative method using n-grams. When I tested this approach using 2-grams, classification was slightly worse, at 68.3% correct.
-
-Lines ~295-335 show how to use word clouds to visualize single word frequencies. I did not find this approach very useful.
-
-To keep this code clean for now I will run various permutations in a scratch version, then summarize findings starting on line #345. In these subsequent iterations I will evaluate:
-
-*  Different text pre-processing strategies: stemmed versus unstemmed, numbers removed vs. not, etc.
-*  Using n-grams rather than single words
-*  Making pair-wise comparison versus comparing multiple groups simultaneously
-*  Modifying the variables in the NB analysis itself.
-
-Finally, there are several things still missing or needed for this workflow:
-
-*  How to feed a text dataset in, and get a table of the predicted categories back out.
-*  How to store the values from the Bayes prediction table so I can use them in calculations rather than type them in by hand.
-*  How to randomize which comments are used for training, testing datasets.
-
-####\
+*  Compare 2-grams, 3-grams rather than single words
+*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
+    +  Ex.: Basic vs. Not Basic (1_Basic vs. not 1_Basic)
+*  Modify the variables within the NB analysis itself. 
+    +  Laplace constant (Y/N), Cutoff value for words removed from list.
+*  Do/Do not stem the terms
+    +  This may remove evidence of tense
+*  Remove/Leave stopwords
+*  Remove/Leave capitalization
+*  Remove/Leave numbers
+    +  This may remove evidence of technical descriptors
+*  Remove/Leave punctuation
+    +  This may remove evidence of questions vs statements
 
@@
 
-##Baseline Version 1.0 of Naive Bayes (NB) Protocol
 ###Initial Data Input
@@
 
-```{r}
+```{}
 #Remove numbers. 
@@
 
-```{r}
+```{}
 #Remove punctuation using the "removePunctuation" function. 
@@
 
-```{r}
+```{}
 #Stemming the text data to strip the suffix from words.  
@@
 
-###Basic Tokenizer
-This version uses the "DocumentTermMatrix" function to creates a matrix in which the rows indicate individual documents (individual TA comments) and the columns indicate presence or absence of the word that is the header for the column.
-Be careful not to confuse it with "TermDocumentMatrix", which transposes the matrix.
-
-"DocumentTermMaxtrix" function also can perform all of the text prep above in one command.
-
-```{r}
-comments_corpus_1gram <- DocumentTermMatrix(comments_corpus_clean)
-comments_corpus_1gram_tidy <- tidy(comments_corpus_1gram)
-inspect(comments_corpus_1gram)
-```
-
-####\
-
-###Data Preparation
-Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data. 
-```{r}
-.75 * 9340 #Number of documents/rows in the dataset; product is #rows for training
-.25 * 9340 #Product is #rows for testing set
-```
-
-This code assumes comments are random. **Probably want to try randomizing them.**
-```{r}
-comments_dtm_train <- comments_corpus_1gram[1:7005, ]
-comments_dtm_test <- comments_corpus_1gram[7006:9340, ]
-```
-
-Save vectors labeling rows in the training and testing vectors
-```{r}
-comments_train_labels <- comments_raw[1:7005, ]$type
-comments_test_labels <- comments_raw[7006:9340,]$type
-```
-
-Make sure that the proportion of each sub-category is similar in the training and testing data set.
-```{r}
-prop.table(table(comments_train_labels))
-prop.table(table(comments_test_labels))
-```
-
-####\
-
-###Preparation for Naive Bayes
-Remove words from the matrix that appear less than 5 times.
-```{r}
-comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
-str(comments_freq_words)
-```
-
-Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
-```{r}
-comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
-comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
-```
-
-The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables.
-To do this we'll build a convert_counts function and apply it to our data.
-```{r}
-convert_counts2 <- function(x) {
-  x <- ifelse(x > 0, "Yes", "No")
-}
-```
-
-This replaces values greater than 0 with yes, and values not greater than 0 with no. Let's apply it to our data.
-```{r}
-comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
-comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
-```
-
-The resulting matrixes will be character type, with cells indicating "yes" or "no" if the word represented by the column appears in the message represented by the row.
-
-####\
-
-###Train Model, Predict, Evaluate
-Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group ONE, TWO, THREE, or FOUR.
-Evaluate the prediction with the actual data using a crosstable from the gmodels package.
-```{r}
-comments_classifier <- naiveBayes(comments_train, comments_train_labels)
-comments_test_pred <- predict(comments_classifier, comments_test)
-CrossTable(comments_test_pred, comments_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
-```
-
-Calculate overall correct prediction rate.
-```{r}
-(42+454+1186+185)/23.35
-```
-
-***
 ##Modified Naive Bayes (NB) Method Using N-grams
-The tokenizer command block above only works for single words.
-The block below uses tm's NLP commands to generate either 1-, 2-, or 3-grams with the same command.
-The only change needed is "#" in "(words(x), #)" below.
-For simplicity I left off all of the preparatory code up leading up to the final cleaned corpus.
-
+Uses tm's NLP commands to generate either 1-, 2-, or 3-grams with the same command. The only change needed is "#" in "(words(x), #)" below.
 ```{r}
 NLP_Tokenizer <- function(x) {
-      unlist(lapply(ngrams(words(x), 2), paste, collapse = " "), use.names = FALSE)
+      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
 }
@@
 ```{r}
-(34+214+1254+94)/23.35
+(44+375+1116+183)/23.35
 ```
@@
 ***
-####\
-
-##Visualization
-This code generates a wordcloud of the frequency of single words in the dataset using the package "wordcloud".
-```{r}
-wordcloud(comments_corpus_clean, max.words = 50, random.order = FALSE)
-```
-
-Compare wordclouds between 3 groups.
-```{r}
-ONE <- subset(comments_raw, type == "1_Basic")
-TWO <- subset(comments_raw, type == "2_Writing")
-THREE <- subset(comments_raw, type == "3_Technical")
-FOUR <- subset(comments_raw, type == "4_Logic")
-
-wordcloud(ONE$text, max.words = 50, scale = c(3, 0.5))
-wordcloud(TWO$text, max.words = 50, scale = c(3, 0.5))
-wordcloud(THREE$text, max.words = 50, scale = c(3, 0.5))
-wordcloud(FOUR$text, max.words = 50, scale = c(3, 0.5))
-```
-
-Show the 5 most frequent words in the data:
-```{r}
-comment_sack <- TermDocumentMatrix(comments_corpus_clean)
-comment_pack <- as.matrix(comment_sack)
-comment_snack <- sort(rowSums(comment_pack), decreasing = TRUE)
-comment_hack <- data.frame(word = names(comment_snack), freq=comment_snack)
-head(comment_hack, 10)
-```
-
-And the 5 most frequent words from each class:
-```{r}
-wordcloud(ONE$text, max.words = 10, scale = c(3, 0.5))
-wordcloud(TWO$text, max.words = 10, scale = c(3, 0.5))
-wordcloud(THREE$text, max.words = 10, scale = c(3, 0.5))
-wordcloud(FOUR$text, max.words = 10, scale = c(3, 0.5))
-```
-
-####\
-
-##Permutation Tests of Naive Bayes (NB) Method
-Version 1.0 correctly identifies comment subcategories ~79% of the time when using single words, and 68.3% of the time when using 2-grams. Permutations to test:
-
-*  Compare 2-grams, 3-grams rather than single words
-*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
-    +  Ex.: Basic vs. Not Basic (1_Basic vs. not 1_Basic)
-*  Modify the variables within the NB analysis itself. 
-    +  Laplace constant (Y/N), Cutoff value for words removed from list.
-*  Do/Do not stem the terms
-    +  This may remove evidence of tense
-*  Remove/Leave stopwords
-*  Remove/Leave capitalization
-*  Remove/Leave numbers
-    +  This may remove evidence of technical descriptors
-*  Remove/Leave punctuation
-    +  This may remove evidence of questions vs statements

### 2019-02-23 Text Classifier Models.Rmd
update

---
title: "Text Classification Primer"
author: "Dan Johnson"
date: "1/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this text classification project I am using an initial codebook and corpus of hand-coded TA comments to test various approaches to semi- and fully automated classification methods and, if appropriate, revise the underlying classification scheme.

This page provides background on general mechanics of text classification and various classifier methods. Other pages describe the dataset structure [LINK] (TA comments corpus), and [LINK] development of the initial codebook.

#Workflow
Text classification problems can be broken down into four overlapping sub-problems. Each piece may be more or less important to the overall problem, but all need to be considered:

1.  Feature selection and engineering
2.  Topic categorization and modeling
3.  Selecting an appropriate text classifier method
4.  Evaluating the outcomes

##Feature Selection and Engineering
This is the process of selecting, tabulating, calculating, or creating text-based features for subsequent evaluation. This includes:

*  Identifying existing, directly observable text features within the dataset
*  Engineering additional features using established NLP methods
*	 Reverse engineering relevant features based on an empirically defined feature set.

The features listed below are used regularly for natural language processing. Not all will be appropriate for all situations though; selecting, evaluating, comparing, and validating the appropriate features for the classification task is an iterative process, not a one-time decision.

*  Count-based Vectors
	+  Word Count – total number of words in the documents
	+  Character Count – total number of characters in the documents
	+  Punctuation Count – total number of punctuation marks in the documents
	+  Upper Case Count – total number of upper case words in the documents
	+  Title Word Count – total number of consecutive upper case (title) words in the documents
*  Distribution and Frequency Calculations
	+  Average word density of sentences 
	+  Average length of the words used in the documents
	+  TF-IDF
	+  Reading level estimates (described below)
*  Part of Speech Tagging and Frequency Distribution of Tags 
	+ Noun Count
	+ Verb Count
	+ Adjective Count
	+ Adverb Count
	+ Pronoun Count
*  Word Embedding (described below)

Additional features are available for longer texts such as multi-sentence paragraphs.

*  Word choices
*  Lexical diversity (described below)
*  Stance patterns - hedging, booster, generalization words or phrasing 
*  Concision - comparisons of noun phrase density versus comment clause density
*  Cohesion - use of words that are additive (also), countering (however), specify or emphasize (even, both, especially), connecting (former, latter)
*  Civility and uncertainty - language that acknowledges limits of statement, critiques ideas not people, and makes fewer generalizations

###Lexical Diversity
This is not the same as word choices. This explanation is an excerpt from "quanteda: Quantitative Analysis of Textual Data", which describes how to use ```textstat_lexdiv``` to calculate lexical diversity. 
https://rdrr.io/cran/quanteda/man/textstat_lexdiv.html
http://quanteda.io/reference/textstat_lexdiv.html

Consider a speaker, who uses the term "allow" multiple times throughout the speech, compared to an another speaker who uses terms "allow, concur, acquiesce, accede, and avow"" for the same word. The latter speech has more lexical diversity than the former. Lexical diversity is widely believed to be an important parameter to rate a document in terms of textual richness and effectiveness.

Lexical diversity, in simple terms, is a measurement of the breadth and variety of vocabulary used in a document. The different measures of lexical diversity are TTR, MSTTR, MATTR, C, R, CTTR, U, S, K, Maas, HD-D, MTLD, and MTLD-MA.

koRpus package in R also provides functions to estimate lexical diversity or complexity.
https://subscription.packtpub.com/book/big_data_and_business_intelligence/9781783551811/2/ch02lvl1sec16/lexical-diversity

###Readability metrics
These estimate the relative grade level or age of a typical reader who could comprehend 80+% (typically) of the text. Most readability metrics require longer blocks of text (100+ words) to achieve a stable score. They use some combination of word count, average syllables per word, and sentence length to estimate reading level.

The metrics differ in how appropriate they are for specific situations. Some were developed for general writing, others for specific school age groups, and still others for technical documents; the US Navy in particular developed a suite of metrics for evaluating their training documents.

The ```readability library``` (v0.1.1) is a collection of readability tools that utilize the ```syllable``` package for fast calculation of readability scores by grouping variables. It calculates Flesch Kincaid, Gunning Fog Index, Coleman Liau, SMOG, Automated Readability Index and an average of the 5 readability scores. These all are general scales; other scores may not be available in R packages.

###Word Embedding
**Word embedding** is a set of language modeling and feature learning techniques in natural language processing (NLP) where words or phrases from the vocabulary are mapped to vectors of real numbers. Conceptually it involves a mathematical embedding from a space with one dimension per word to a continuous vector space with a much lower dimension.

Methods to generate the mapping include neural networks, dimensionality reduction, probabilistic models, knowledge base methods, and explicit representation in terms of the context in which words appear. The technique of representing words as vectors has roots in the 1960s with the development of the vector space model for information retrieval.

Word embeddings come in two different styles, one in which words are expressed as vectors of co-occurring words, and another in which words are expressed as vectors of linguistic contexts in which the words occur. Most new word embedding techniques rely on a neural network architecture instead of more traditional n-gram models and unsupervised learning.

One of the main limitations of word embeddings (word vector space models in general) is that possible meanings of a word are conflated into a single representation (a single vector in the semantic space). Sense embeddings are a solution to this problem: individual meanings of words are represented as distinct vectors in the space.

Tools like **Principal Component Analysis (PCA)** and **T-Distributed Stochastic Neighbour Embedding (t-SNE)** are used to reduce the dimensionality of word vector spaces and visualize word embeddings and clusters.

Software for training and using word embeddings includes Tomas Mikolov's Word2vec, Stanford University's GloVe, fastText, Gensim, AllenNLP's Elmo, Indra and Deeplearning4j. Three of these tools are described further below.

####Word2vec
Word2vec is a group of shallow, two-layer neural network models that produce word embeddings. The toolkit can train vector space models to reconstruct linguistic contexts of words. Word2vec takes as its input a large corpus of text and produces a vector space, typically of several hundred dimensions, with each unique word in the corpus being assigned a corresponding vector in the space. Word vectors are positioned in the vector space so words that share common contexts in the corpus are located in close proximity to one another in the space.

The use of different model parameters and different corpus sizes can greatly affect the quality of a word2vec model. Accuracy can be improved in a number of ways, including the choice of model architecture (CBOW or Skip-Gram), increasing the training data set, increasing the number of vector dimensions, and increasing the window size of words considered by the algorithm. Each of these improvements comes with the cost of increased computational complexity and therefore increased model generation time.

In models using large corpora and a high number of dimensions, the skip-gram model yields the highest overall accuracy, and consistently produces the highest accuracy on semantic relationships, as well as yielding the highest syntactic accuracy in most cases. However, the CBOW is less computationally expensive and yields similar accuracy results.

Accuracy increases overall as the number of words used increases, and as the number of dimensions increases. Altszyler et al. (2017) studied Word2vec performance in two semantic tests for different corpus size. They found that Word2vec has a steep learning curve, outperforming another word-embedding technique (LSA) when it is trained with medium to large corpus size (more than 10 million words). However, with a small training corpus LSA showed better performance. Additionally they show that the best parameter setting depends on the task and the training corpus. Nevertheless, for skip-gram models trained in medium size corpora, with 50 dimensions, a window size of 15 and 10 negative samples seems to be a good parameter setting.

####fastText
fastText is a Python-based library created by Facebook's AI Research (FAIR) lab for learning of word embeddings and text classification. The model allows to create an unsupervised learning or supervised learning algorithm for obtaining vector representations for words. fastText uses a neural network for word embedding.

https://fasttext.cc/

####GloVe
GloVe, short for Global Vectors, is a model for distributed word representation. The model is an unsupervised learning algorithm for obtaining vector representations for words. Training is performed on aggregated global word-word co-occurrence statistics from a corpus, and the resulting representations showcase interesting linear substructures of the word vector space. It is developed as an open-source project at Stanford. Pre-trained word vectors data are made available under the Public Domain Dedication and License v1.0.

GloVe uses Euclidean distance (or cosine similarity) between two word vector for measuring the linguistic or semantic similarity of the corresponding words. Sometimes, the nearest neighbors according to this metric reveal rare but relevant words that lie outside an average human's vocabulary. For example, here are the closest words to the target word **frog**:

*  frog
*  frogs
*  toad
*  litoria
*  leptodactylidae
*  rana
*  lizard
*  eleutherodactylus

The similarity metrics used for nearest neighbor evaluations produce a single scalar that quantifies the relatedness of two words. This simplicity can be problematic since two given words almost always exhibit more intricate relationships than can be captured by a single number. For example, man may be regarded as similar to woman in that both words describe human beings; on the other hand, the two words are often considered opposites since they highlight a primary axis along which humans differ from one another.

In order to capture in a quantitative way the nuance necessary to distinguish man from woman, it is necessary for a model to associate more than a single number to the word pair. A natural and simple candidate for an enlarged set of discriminative numbers is the vector difference between the two word vectors. GloVe is designed in order that such vector differences capture as much as possible the meaning specified by the juxtaposition of two words.

https://nlp.stanford.edu/projects/glove/

##Sub-Problem Two: Topic Categorization and Modeling Methods
Establishing meaningful text classification categories depends on the goals of the project or the question being asked. Are relevant categories known already, or do they need to be established? What are the main topics

**Curated Categories** (my provisional term until I find the appropriate equivalent in current use) are established by subject matter experts. The established categories can be constructed from:

*  A limited keywords vocabulary (National Library of Medicine MeSH categories for example)
*  Disciplinary conventions or shared best practices (the Dublin Core is an example)
*  Prior empirical studies
*  The dataset itself, using *de novo* iterative coding. This process is described further here [LINK] and here [LINK].

**Topic Modeling** is a group of techniques for identifying groups of words (the "topic")  that best summarize the information in a collection of documents. Topic modeling is an iterative/ generative process that does not require any prior knowledge of the documents or relationships.

###Topic Modeling Methods to Identify Features
####Latent Dirichlet Allocation
(LDA) is a common method for generating topic model features. LDA is an iterative model which starts from a fixed number of topics. Each topic is represented as a distribution over words, and each document is then represented as a distribution over topics. Although the tokens themselves are meaningless, the probability distributions over words provided by the topics provide a sense of the different ideas contained in the documents.

####Brown Clustering
A hierarchical clustering algorithm that is applied to text, grouping words into clusters that are assumed to be semantically related by virtue of their having been embedded in similar contexts.

In natural language processing, Brown clustering or IBM clustering is a form of hierarchical clustering of words based on the contexts in which they occur. The intuition behind the method is that a class-based language model (also called cluster n-gram model), i.e. one where probabilities of words are based on the classes (clusters) of previous words, is used to address the data sparsity problem inherent in language modeling.

Brown groups items (i.e., types) into classes, using a binary merging criterion based on the log-probability of a text under a class-based language model, i.e. a probability model that takes the clustering into account.

####Latent semantic analysis
LSA is a technique in natural language processing of analyzing relationships between a set of documents and the terms they contain by producing a set of concepts related to the documents and terms. LSA assumes that words that are close in meaning will occur in similar pieces of text (the distributional hypothesis). A matrix containing word counts per paragraph (rows represent unique words and columns represent each paragraph) is constructed from a large piece of text and a mathematical technique called singular value decomposition (SVD) is used to reduce the number of rows while preserving the similarity structure among columns. Words are then compared by taking the cosine of the angle between the two vectors (or the dot product between the normalizations of the two vectors) formed by any two rows. Values close to 1 represent very similar words while values close to 0 represent very dissimilar words.

In the context of its application to information retrieval, it is sometimes called latent semantic indexing (LSI).

LSA can use a term-document matrix which describes the occurrences of terms in documents; it is a sparse matrix whose rows correspond to terms and whose columns correspond to documents. A typical example of the weighting of the elements of the matrix is tf-idf (term frequency–inverse document frequency): the weight of an element of the matrix is proportional to the number of times the terms appear in each document, where rare terms are upweighted to reflect their relative importance.

Applications of LSA include:

*  Comparing documents in low-dimensional space (data clustering, document classification).
*  Finding relations between terms (synonymy and polysemy). Synonymy is the phenomenon where different words describe the same idea. Polysemy is the phenomenon where the same word has multiple meanings.
*  Given a query of terms, translating it into the low-dimensional space, and find matching documents (information retrieval).
*  Analyzing word association in text corpus.

LSA overcomes two of the most problematic constraints of Boolean keyword queries: multiple words that have similar meanings (synonymy) and words that have more than one meaning (polysemy).

LSA is also used to perform automated document categorization. Document categorization is the assignment of documents to one or more predefined categories based on their similarity to the conceptual content of the categories. Several experiments have demonstrated that there are a number of correlations between the way LSI and humans process and categorize text.

LSA uses example documents to establish the conceptual basis for each category. During categorization processing, the concepts contained in the documents being categorized are compared to the concepts contained in the example items, and a category (or categories) is assigned to the documents based on the similarities between the concepts they contain and the concepts that are contained in the example documents.

LSA automatically adapts to new and changing terminology, and has been shown to be very tolerant of noise (i.e., misspelled words, typographical errors, unreadable characters, etc.).[24] This is especially important for applications using text derived from Optical Character Recognition (OCR) and speech-to-text conversion. LSA also deals effectively with sparse, ambiguous, and contradictory data.

Text does not need to be in sentence form for LSA to be effective. It can work with lists, free-form notes, email, Web-based content, etc. As long as a collection of text contains multiple terms, LSA can be used to identify patterns in the relationships between the important terms and concepts contained in the text.

##Sub-Problem Three: Choosing a Text Classifier Method
The many methods for classifying texts fall into three general categories:

*  Pattern matching strategies (does a particular phrase appear?)
*  Algorithms (how closely does the query text meet pre-set criteria?)
*  Neural networks (the network decides its own criteria, aka machine learning)

This edited excerpt nicely describes the overall approach and problems of text classification. It is from: **xxxxx**

Understanding how chatbots work can help us understand text classifiers, because the classifier is the fundamental piece that makes the chatbot work. Let’s look at the inner workings of a Multinomial Naive Bayes algorithm for text classification and natural language processing (NLP).

This classifier is “naive” because each word is treated as having no connection with other words in the sentence being classified. In other words it assumes independence between "features," in this case, words.

For example, in the sentence “the fox jumped over the log”, there is no relationship between “jumped” and “fox” or “log”. In NLP, this collection of words is referred to as “a bag of words”. While this is incredibly naive, the classifier isn’t attempting to understand the meaning of a sentence, it’s trying to classify it.

###Pattern Matchers
Early chatbots used **pattern matching** to classify text and produce a response. This is often referred to as “brute force” as the author of the system needs to describe every pattern for which there is a response.

A standard structure for these patterns is “AIML” (artificial intelligence markup language). Its use of the term “artificial intelligence” is quite an embellishment, but that’s another story.

A simple pattern matching definition:
```
<aiml version = "1.0.1" encoding = "UTF-8"?>
   <category>
      <pattern> WHO IS ALBERT EINSTEIN </pattern>
      <template>Albert Einstein was a German physicist.</template>
   </category>

   <category>
      <pattern> WHO IS Isaac NEWTON </pattern>
      <template>Isaac Newton was a English physicist and mathematician.</template>
   </category>

   <category>
      <pattern>DO YOU KNOW WHO * IS</pattern>
      <template>
         <srai>WHO IS <star/></srai>
      </template>
   </category>
</aiml>
```

The machine then produces:

> Human: Do you know who Albert Einstein is
> Robot: Albert Einstein was a German physicist.

It knows who a physicist is only because his or her name has an associated pattern. Likewise it responds to anything solely because of an authored pattern. Given hundreds or thousands of patterns you might see a chatbot “persona” emerge.

In 2000 a chatbot built by John Denning and colleagues using this approach was in the news for passing the "Turing test." It emulated the replies of a 13 year old boy from Ukraine (broken English and all). I met with John in 2015 and he made no false pretenses about the internal workings of this automaton. It may have been “brute force” but it proved a point: parts of a conversation can be made to appear “natural” using a sufficiently large definition of patterns. It proved Alan Turing’s assertion, that this question of a machine fooling humans was “meaningless”.

An example of this approach used in building chatbots is PandoraBots, who claim over 285k chatbots have been constructed using their framework.

Automated phone voice response systems that recognize patterns of spoken words and direct calls accordingly use a similar approach.

###Algorithm-Based Classification
Creating a brute-force pattern matcher is daunting: for each unique input a pattern must be available to specify a response. An algorithm-based method approaches the problem mathematically.

One text classification algorithm is "Multinomial Naive Bayes." Given a training set of sentences, each belonging to a class (that is, a particular group), we can count the occurrence of each word in each class, account for its commonality and assign each word a score for each class. Factoring for commonality is important: matching the word “it” is considerably less meaningful than a match for the word “cheese”. When we input a new sentence, scores for each word in the new sentence are compared to scores for those same words in the training set. The class with the highest OVERALL score for the input sentence is the class to which the input sentence most likely belongs.

Look at this sample training set:
```
class: weather
    "is it nice outside?"
    "how is it outside?"
    "is the weather nice?"

class: greeting
    "how are you?"
    "hello there"
    "how is it going?"
```

Let’s classify a few sample input sentences:
```
input: "Hi there"
 term: "hi" (no matches in either class)
 term: "there" (matches a word in class: greeting)
 classification: greeting (score=1)

input: "What’s it like outside?"
 term: "it" (matches words in class: weather (2), greeting(1))
 term: "outside (class: weather (2) )
 classification: weather (score=4)
```

Notice that the classification for "What’s it like outside" found a term in the "greeting" class but the term similarities to the "weather" class produced a higher score. By using an equation we are looking for word matches given some sample sentences for each class, and we avoid having to identify every pattern.

The classification score produced identifies the class with the highest term matches (accounting for commonality of words) but this has limitations. A score is not the same as a probability, a score tells us which intent is most like the sentence but not the likelihood of it being a match. Thus it is difficult to apply a threshold for which classification scores to accept or not. Having the highest score from this type of algorithm only provides a relative basis, it may still be an inherently weak classification. Also the algorithm doesn’t account for what a sentence is not, it only counts what it is like. You might say this approach doesn’t consider what makes a sentence not a given class.

Many text and spoken word frameworks use similar algorithms such as this to classify intent. Most of what’s taking place is word counting against training datasets. It’s "naive" but surprisingly effective.

I will be using an algorithm-based classification strategy for this project.

###Neural Networks
These methods are beyond the scope of this project. Anyone interested in knowing more can search online for information about:

*  Shallow Neural Networks
*  Deep Neural Networks
	+  Convolutional Neural Network (CNN)
	+  Long Short Term Modelr (LSTM)
	+  Gated Recurrent Unit (GRU)
	+  Bidirectional RNN
	+  Recurrent Convolutional Neural Network (RCNN)
	+  Other Variants of Deep Neural Networks

###Designing an Evaluation Workflow
The workflow for comparing classification algorithms starts with identifying potentially usable classifier methods. Several approaches are described later. For each potential classifier method, the next steps are:

*  Establish randomized training and test datasets
*  Train the classifier algorithm, then evaluate that model on a test dataset
*  Calculate an accuracy score using Gini index or similar method
*  Select best performing classifiers and either refine, or operationalize for new datasets

###Picking the Algorithm(s)
Algorithms approach text classification several ways.

The **bag of keywords approach** requires compiling a list of "key terms" that qualifies the type of content in question to a certain topic. If one or more qualifying keywords is present in a document, the document is assigned to the topic. One of the many problems with this approach is that identifying and organizing a list of terms (preferred, alternate, etc.) is labor intensive (much like simple pattern matching), and the natural ambiguity of language (one keyword can have different meanings) causes lots of false positives. This makes the bag of keywords approach not only not scalable, but also inaccurate.

**Rules-based approaches** use linguistic rules that capture all of the elements and attributes of a document to assign it to a category. Rules can be written manually or generated by automatic analysis and then validated manually (reducing manual effort up to 90%). Rules can be understood and improved (unlike the black box system used with statistics-based algorithms). A rules-based approach is flexible, powerful (much more than a bag of keywords) and easy to express. This approach works best when combined with semantic technology that can uncover deeper connections within text (meaning, relevancy, relationship between concepts, etc.)

**Statistical approaches** use a “training set” of documents that covers the same topic to identify key features (usually words). The algorithm (Bayesian, LSA, etc.) calculates term frequencies, then builds implicit rules in order to classify other content. This approach has no understanding of meaning. In addition, systems using these types of text classification algorithms are essentially a “black box”; no one can explain why specific terms are selected by the algorithm or how they are being weighted. In the event the classification is incorrect, there is no accurate way to modify a rule for better results.

Algorithms also can be separated into un-supervised and supervised methods.

**Unsupervised learning** uses test data that has not been labeled, classified or categorized. Unsupervised learning identifies commonalities in the training data, then classifies new texts based on the presence or absence of such commonalities.

Unsupervised learning algorithms appearing frequently in literature are:

*  Clustering
	+  Hierarchical clustering
	+  k-means
	+  Mixture models
*  Principal component analysis (PCA)

**Supervised learning** uses labeled training data to defined important text features. In supervised learning, each example is a pair consisting of an input object (typically a vector) and a desired output value. A supervised learning algorithm analyzes the training data and produces an inferred function, which can be used for mapping new examples. .

Supervised text classification algorithms appearing regularly in the literature are:

*  Naive Bayes
	+  Binary independent NB
	+  Multinomial NB
*  Linear Classifier
*  Logistic Regression
*  Support vector machines
	+  Basic SVM
	+  SVM multiple classes
	+  SVM multiple kernels
*  Decision trees
*  Linear Discriminant Analysis
*  Nearest neighbor
	+  k-Nearest Neighbor
	+  Weighted k-Nearest Neighbor
*  Similarity Learning
*  Relevance feedback
	+  Rocchio
	+  Rocchio in a query zone
*  Ensemble methods
	+  Bagging based Models
	+  Stacking based Models
	+  Boosting based Models
	+  Random Forest Models

For reference I assembled short excerpts from online resources or published articles describing most of these methods.

###Naive Bayes
A classification technique based on Bayes’ Theorem with an assumption of independence among predictors. A Naive Bayes classifier assumes that the presence of a particular feature in a class is unrelated to the presence of any other feature here. NB can be run on:

*  Word count or n-gram count vectors
*  Word or n-gram level TF-IDF vectors
*  Character Level TF-IDF vectors
	+  Would be useful for evaluating use of "?" and "="

Data are organized into 2 dictionaries: **corpus words** (each stemmed word and the # of occurances) and **class words** (each class and the list of stemmed words within it). The algorithm then uses these data structures to classify new text.

###Linear Classifier
The linear regression algorithm is one of the fundamental supervised machine-learning algorithms due to its relative simplicity and well-known properties.

The goal of statistical classification is to use an object's characteristics to identify which class (or group) it belongs to. A linear classifier achieves this by making a classification decision based on the value of a linear combination of the characteristics. An object's characteristics are also known as feature values and are typically presented to the machine in a vector called a feature vector. Such classifiers work well for practical problems such as document classification, and more generally for problems with many variables (features), reaching accuracy levels comparable to non-linear classifiers while taking less time to train and use.

###Logistic Regression
A logistic regression measures the relationship between the categorical dependent variable and one or more independent variables by estimating probabilities using a logistic/sigmoid function. LR can be run on:

*  Word count or n-gram count vectors
*  Word or n-gram level TF-IDF vectors
*  Character Level TF-IDF vectors

Resources for Logistic Regression 
https://www.analyticsvidhya.com/blog/2015/10/basics-logistic-regression/

R for logistic regression
https://www.saedsayad.com/logistic_regression.htm

###Support Vector Machine (SVM) Model
This is a supervised machine learning algorithm which can be used for both classification or regression challenges. The model extracts a best possible hyper-plane / line that segregates the two classes. The only example I found of SVM was on Ngram Level TF-IDF Vectors. Support vector machines can be:

*  Basic SVM
*  SVM multiple classes
*  SVM multiple kernels

###Decision Trees
Flowchart-like structures in which each internal node represents a "test" on an attribute (e.g. whether a coin flip comes up heads or tails), each branch represents the outcome of the test, and each leaf node represents a class label (decision taken after computing all attributes). The paths from root to leaf represent classification rules.

In decision analysis, a decision tree and the closely related influence diagram are used as a visual and analytical decision support tool, where the expected values (or expected utility) of competing alternatives are calculated.

A decision tree consists of three types of nodes:

*  Decision nodes – typically represented by squares
*  Chance nodes – typically represented by circles
*  End nodes – typically represented by triangles

Decision tree learning is the construction of a decision tree from class-labeled training tuples. A decision tree is a flow-chart-like structure, where each internal (non-leaf) node denotes a test on an attribute, each branch represents the outcome of a test, and each leaf (or terminal) node holds a class label. The topmost node in a tree is the root node.

The goal is to create a model that predicts the value of a target variable based on several input variables. An example is shown in the diagram at right. Each interior node corresponds to one of the input variables; there are edges to children for each of the possible values of that input variable. Each leaf represents a value of the target variable given the values of the input variables represented by the path from the root to the leaf.

A decision tree is a simple representation for classifying examples. For this section, assume that all of the input features have finite discrete domains, and there is a single target feature called the "classification". Each element of the domain of the classification is called a class. A decision tree or a classification tree is a tree in which each internal (non-leaf) node is labeled with an input feature. The arcs coming from a node labeled with an input feature are labeled with each of the possible values of the target or output feature or the arc leads to a subordinate decision node on a different input feature. Each leaf of the tree is labeled with a class or a probability distribution over the classes.

A tree can be "learned" by splitting the source set into subsets based on an attribute value test. This process is repeated on each derived subset in a recursive manner called recursive partitioning. The recursion is completed when the subset at a node has all the same value of the target variable, or when splitting no longer adds value to the predictions. This process of top-down induction of decision trees (TDIDT) is an example of a greedy algorithm, and it is by far the most common strategy for learning decision trees from data.

In data mining, decision trees can be described also as the combination of mathematical and computational techniques to aid the description, categorization and generalization of a given set of data.

###Linear Discriminant Analysis
LDA works when the measurements made on independent variables for each observation are continuous quantities. When dealing with categorical independent variables, the equivalent technique is discriminant correspondence analysis. The statistical system R includes the packages: MASS, ade4, ca, vegan, ExPosition, andFactoMineR which perform correspondence analysis and multiple correspondence analysis.

As the name indicates, discriminant correspondence analysis (DCA) is an extension of discriminant analysis (DA) and correspondence analysis (CA). Like discriminant analysis, the goal of DCA is to categorize observations in pre-defined groups, and like  correspondence analysis, it is used with nominal variables.The main idea behind DCA is to represent each group by thesum of its observations and to perform a simple CA on the groupsby variables matrix. The original observations are then projectedas  supplementary  elements and each observation is assigned to the closest group. The comparison between the a priori and the a posteriori classifications can be used to assess the quality of thendiscrimination. A similar procedure can be used to assign new observations to categories. The stability of the analysis can be evaluated using cross-validation techniques such as jackknifing or boot-strapping.

For the R model: https://pbil.univ-lyon1.fr/ade4/ade4-html/discrimin.coa.html

###k-Nearest Neighbors
k-NN is a non-parametric method used for classification and regression. In both cases, the input consists of the k closest training examples in the feature space. The output depends on whether k-NN is used for classification or regression. In k-NN classification, the output is a class membership. An object is classified by a majority vote of its neighbors, with the object being assigned to the class most common among its k nearest neighbors (k is a positive integer, typically small). If k = 1, then the object is simply assigned to the class of that single nearest neighbor.

###Similarity Learning
Similarity learning is an area of supervised machine learning in artificial intelligence. It is closely related to regression and classification, but the goal is to learn from examples a similarity function that measures how similar or related two objects are. It has applications in ranking, in recommendation systems, visual identity tracking, face verification, and speaker verification.

###Relevance Feedback/Rocchio Algorithm
Relevance feedback is a feature of some information retrieval systems. The idea behind relevance feedback is to take the results that are initially returned from a given query, to gather user feedback, and to use information about whether or not those results are relevant to perform a new query. We can usefully distinguish between three types of feedback: explicit feedback, implicit feedback, and blind or "pseudo" feedback.

The Rocchio algorithm is based on a method of relevance feedback. Like many other retrieval systems, the Rocchio feedback approach was developed using the Vector Space Model. The algorithm is based on the assumption that most users have a general conception of which documents should be denoted as relevant or non-relevant. Therefore, the user's search query is revised to include an arbitrary percentage of relevant and non-relevant documents as a means of increasing the search engine's recall, and possibly the precision as well.

For more:
https://nlp.stanford.edu/IR-book/html/htmledition/rocchio-classification-1.html

###Ensemble Methods
The excerpt below nice describes how ensemble methods work. It is reprinted from: *Ensemble Methods Explained in Simple English*, by Tavish Srivastava, posted August 2, 2015.

Ensemble modeling is a powerful way to improve the performance of your model. It usually pays off to apply ensemble learning over and above various models you might be building.

Example: I want to invest in a company XYZ. I am not sure about its performance though. So, I look for advice on whether the stock price will increase more than 6% per annum or not? I decide to approach various experts having diverse domain experience:

>1.  Employee of Company XYZ: This person knows the internal functionality of the company and have the insider information about the functionality of the firm. But he lacks a broader perspective on how are competitors innovating, how is the technology evolving and what will be the impact of this evolution on Company XYZ’s product. In the past, he has been right 70% times.

>1.  Financial Advisor of Company XYZ: This person has a broader perspective on how companies strategy will fair of in this competitive environment. However, he lacks a view on how the company’s internal policies are fairing off. In the past, he has been right 75% times.

>1.  Stock Market Trader: This person has observed the company’s stock price over past 3 years. He knows the seasonality trends and how the overall market is performing. He also has developed a strong intuition on how stocks might vary over time. In the past, he has been right 70% times.

>1.  Employee of a competitor: This person knows the internal functionality of the competitor firms and is aware of certain changes which are yet to be brought. He lacks a sight of company in focus and the external factors which can relate the growth of competitor with the company of subject. In the past, he has been right 60% of times.

>1.  Market Research team in same segment: This team analyzes the customer preference of company XYZ’s product over others and how is this changing with time. Because he deals with customer side, he is unaware of the changes company XYZ will bring because of alignment to its own goals. In the past, they have been right 75% of times.

>1.  Social Media Expert: This person can help us understand how has company XYZ has positioned its products in the market. And how are the sentiment of customers changing over time towards company. He is unaware of any kind of details beyond digital marketing. In the past, he has been right 65% of times.

Given the broad spectrum of access we have, we can probably combine all the information and make an informed decision. In a scenario when all the 6 experts/teams verify that it’s a good decision(assuming all the predictions are independent of each other), we will get a combined accuracy rate of:

```
		  1 - 30%*25%*30%*40%*25%*35%

		= 1 - 0.07875 = 99.92125%
```

Assumption: The assumption used here that all the predictions are completely independent is slightly extreme as they are expected to be correlated. However, we see how we can be so sure by combining various predictions together.

Let us now change the scenario slightly. This time we have 6 experts, all of them are employee of company XYZ working in the same division. Everyone has a propensity of 70% to advocate correctly.

>*What if we combine all this advice together, can we still raise up our confidence to >99%?*

Obviously not, as all the predictions are based on very similar set of information. They are certain to be influenced by similar set of information and the only variation in their advice would be due to their personal opinions & collected facts about the firm.

Ensemble learning is the art of combining individual models together to improve the stability and predictive power of the model.

####Error Sources in Ensemble Learning (Variance vs. Bias)
The error emerging from any model can be broken down into three components mathematically.

**Error of a model** Why is this important in the current context? To understand what really goes behind an ensemble model, we need to first understand what causes error in the model.

**Bias error** quantifies how much on an average the predicted values differ from the actual value. A high bias error means we have a under-performing model which keeps on missing important trends.

**Variance** quantifies how much the predictions made on same observations differ from each other. A high variance model will over-fit on your training population and perform badly on any observation beyond training.

Normally, as you increase the complexity of your model, you will see a reduction in error due to lower bias in the model. However, this only happens till a particular point. As you continue to make your model more complex, you end up **over-fitting your model** and hence your model will start suffering from high variance.

A champion model should maintain a balance between these two types of errors. This is known as the **trade-off management of bias-variance errors.** Ensemble learning is one way to execute this trade off analysis.

####Commonly used Ensemble learning techniques
**Bagging** tries to implement similar learners on small sample populations and then takes a mean of all the predictions. In generalized bagging, you can use different learners on different population.  As you can expect this helps us to reduce the variance error.

**Boosting** is an iterative technique which adjust the weight of an observation based on the last classification. If an observation was classified incorrectly, it tries to increase the weight of this observation and vice versa. Boosting in general decreases the bias error and builds strong predictive models. However, they may sometimes over fit on the training data.

**Stacking** is a very interesting way of combining models. Here we use a learner to combine output from different learners. This can lead to decrease in either bias or variance error depending on the combining learner we use.

####Random Forest Models
Random Forest models are a type of ensemble models, particularly bagging models. They are part of the tree based model family. RF can be run on:

*  Word count or n-gram count vectors
*  Word level TF-IDF vectors

Resources for Random Forest: 
https://www.analyticsvidhya.com/blog/2014/06/introduction-random-forest-simplified/

####Extreme Gradient Boosting Model
Boosting models are another type of ensemble models part of tree based models. Boosting is a machine learning ensemble meta-algorithm for primarily reducing bias, and also variance in supervised learning, and a family of machine learning algorithms that convert weak learners to strong ones. A weak learner is defined to be a classifier that is only slightly correlated with the true classification (it can label examples better than random guessing). XGB can be run on:

*  Word count or n-gram count vectors
*  Word level TF-IDF vectors
*  Character Level TF-IDF vectors

###Hierarchical Clustering
In data mining and statistics, hierarchical clustering (also called hierarchical cluster analysis or HCA) is a method of cluster analysis which seeks to build a hierarchy of clusters.

###k-Means Clustering
k-means clustering is a method of vector quantization, originally from signal processing, that is popular for cluster analysis in data mining. k-means clustering aims to partition n observations into k clusters in which each observation belongs to the cluster with the nearest mean, serving as a prototype of the cluster. This results in a partitioning of the data space into Voronoi cells.

The problem is computationally difficult (NP-hard); however, efficient heuristic algorithms converge quickly to a local optimum. These are usually similar to the expectation-maximization algorithm for mixtures of Gaussian distributions via an iterative refinement approach employed by both k-means and Gaussian mixture modeling. They both use cluster centers to model the data; however, k-means clustering tends to find clusters of comparable spatial extent, while the expectation-maximization mechanism allows clusters to have different shapes.

The algorithm has a loose relationship to the k-nearest neighbor classifier, a popular machine learning technique for classification that is often confused with k-means due to the name. Applying the 1-nearest neighbor classifier the cluster centers obtained by k-means classifies new data into the existing clusters. This is known as nearest centroid classifier or Rocchio algorithm.

###Mixture Model
In statistics, a mixture model is a probabilistic model for representing the presence of subpopulations within an overall population, without requiring that an observed data set should identify the sub-population to which an individual observation belongs. Formally a mixture model corresponds to the mixture distribution that represents the probability distribution of observations in the overall population. However, while problems associated with "mixture distributions" relate to deriving the properties of the overall population from those of the sub-populations, "mixture models" are used to make statistical inferences about the properties of the sub-populations given only observations on the pooled population, without sub-population identity information.

Topics in a document: Assume that a document is composed of N different words from a total vocabulary of size V, where each word corresponds to one of K possible topics. The distribution of such words could be modelled as a mixture of K different V-dimensional categorical distributions. A model of this sort is commonly termed a topic model. Note that expectation maximization applied to such a model will typically fail to produce realistic results, due (among other things) to the excessive number of parameters. Some sorts of additional assumptions are typically necessary to get good results.

A prior distribution is placed over the parameters describing the topic distributions, using a Dirichlet distribution with a concentration parameter that is set significantly below 1, so as to encourage sparse distributions (where only a small number of words have significantly non-zero probabilities).

Some sort of additional constraint is placed over the topic identities of words, to take advantage of natural clustering.

For example, a Markov chain could be placed on the topic identities (i.e., the latent variables specifying the mixture component of each observation), corresponding to the fact that nearby words belong to similar topics. (This results in a hidden Markov model, specifically one where a prior distribution is placed over state transitions that favors transitions that stay in the same state.)

Another possibility is the latent Dirichlet allocation model, which divides up the words into D different documents and assumes that in each document only a small number of topics occur with any frequency.

###Principal Component Analysis 
PCA is a statistical procedure that uses an orthogonal transformation to convert a set of observations of possibly correlated variables (entities each of which takes on various numerical values) into a set of values of linearly uncorrelated variables called principal components. This transformation is defined in such a way that the first principal component has the largest possible variance (that is, accounts for as much of the variability in the data as possible), and each succeeding component in turn has the highest variance possible under the constraint that it is orthogonal to the preceding components. The resulting vectors (each being a linear combination of the variables and containing n observations) are an uncorrelated orthogonal basis set. PCA is sensitive to the relative scaling of the original variables.

PCA is mostly used as a tool in exploratory data analysis and for making predictive models. It is often used to visualize genetic distance and relatedness between populations. PCA can be done by eigenvalue decomposition of a data covariance (or correlation) matrix or singular value decomposition of a data matrix, usually after a normalization step of the initial data. The normalization of each attribute consists of mean centering – subtracting each data value from its variable's measured mean so that its empirical mean (average) is zero – and, possibly, normalizing each variable's variance to make it equal to 1. The results of a PCA are usually discussed in terms of component scores, sometimes called factor scores (the transformed variable values corresponding to a particular data point), and loadings (the weight by which each standardized original variable should be multiplied to get the component score). If component scores are standardized to unit variance loadings must contain the data variance in them (and that is the magnitude of eigenvalues). If component scores are not standardized (therefore they contain the data variance) then loadings must be unit-scaled, ("normalized") and these weights are called eigenvectors; they are the cosines of orthogonal rotation of variables into principal components or back.

PCA is the simplest of the true eigenvector-based multivariate analyses. Often, its operation can be thought of as revealing the internal structure of the data in a way that best explains the variance in the data. If a multivariate dataset is visualised as a set of coordinates in a high-dimensional data space (1 axis per variable), PCA can supply the user with a lower-dimensional picture, a projection of this object when viewed from its most informative viewpoint. This is done by using only the first few principal components so that the dimensionality of the transformed data is reduced.

PCA is closely related to factor analysis. Factor analysis typically incorporates more domain specific assumptions about the underlying structure and solves eigenvectors of a slightly different matrix.

PCA is also related to canonical correlation analysis (CCA). CCA defines coordinate systems that optimally describe the cross-covariance between two datasets while PCA defines a new orthogonal coordinate system that optimally describes variance in a single dataset.

##Strategies for Improving Text Classification Models
Accuracy can be improved using one or more of these strategies.

1. Text Cleaning. Removing stopwords, punctuations marks, suffix variations etc. can  reduce the noise present in text data.

2. Hstacking arrays of text / NLP features with text feature vectors in sequence horizontally (column wise). After extracting features initially, try combining different feature vectors to improve the accuracy of the classifier.

3. Tuning the Classifier. Adjusting the parameters of the classifier is an important step towards a best fit model.

4. Testing Different Ensemble Models. Stacking different models and blending their outputs can help to further improve the results.

5.  Weighting words: we can significantly improve our algorithm by accounting for the commonality of each word. The word “is” should carry a lower weigh than the word “sandwich” in most cases, because it is more common.

##Evaluation
How to evaluate the different components of the overall text classification problem.

###Gini Index
One of the most common methods for quantifying the discrimination level of a feature is the use of a measure known as the gini-index.

###Cohen's Kappa Score
Cohen's kappa is a measure of interrater reliability used in social sciences. This excerpt from an R blog explains it.

**Using kappa score for comparisons of groups and methods**
https://www.personality-project.org/r/html/kappa.html
https://www.r-bloggers.com/k-is-for-cohens-kappa/

Last April, during the A to Z of Statistics, I blogged about Cohen’s kappa, a measure of interrater reliability. Cohen’s kappa is a way to assess whether two raters or judges are rating something the same way. And thanks to an R package called **irr**, it’s very easy to compute. But first, let’s talk about why you would use Cohen’s kappa and why it’s superior to a more simple measure of interrater reliability, interrater agreement.

***
***

###Neural Networks
A neural network is a mathematical model that is designed to behave similar to biological neurons and nervous system. These models are used to recognize complex patterns and relationships that exists within a labelled data.

Artificial neural networks, invented in the 1940’s, are a way of calculating an output from an input (a classification) using weighted connections (“synapses”) that are calculated from repeated iterations through training data. Each pass through the training data alters the weights such that the neural network produces the output with greater “accuracy” (lower error rate).

There’s not much new about these structures, except today’s software is using much faster processors and can work with a lot more memory. The combination of working memory and speed is crucial when you’re doing hundreds of thousands of matrix multiplications (the neural network’s essential math operation).

As in the prior method, each class is given with some number of example sentences. Once again each sentence is broken down by word (stemmed) and each word becomes an input for the neural network. The synaptic weights are then calculated by iterating through the training data thousands of times, each time adjusting the weights slightly to greater accuracy. By recalculating back across multiple layers (“back-propagation”) the weights of all synapses are calibrated while the results are compared to the training data output. These weights are like a ‘strength’ measure, in a neuron the synaptic weight is what causes something to be more memorable than not. You remember a thing more because you’ve seen it more times: each time the ‘weight’ increases slightly.

At some point the adjustment reaches a point of diminishing returns, this is called “over-fitting” and going beyond this is counter-productive.

The trained neural network is less code than an comparable algorithm but it requires a potentially large matrix of “weights”. In a relatively small sample, where the training sentences have 150 unique words and 30 classes this would be a matrix of 150x30. Imagine multiplying a matrix of this size 100,000 times to establish a sufficiently low error rate. This is where processing speed comes in.

If the neural network sounds magnificently sophisticated, relax, it boils down to matrix multiplication and a formula for reducing values between -1 and 1 or some other minimal range. A middle-school math student could learn this in a few hours. The hard work is achieving clean training data.

Just as there are variations in pattern matching code and in algorithms, there are variations in neural networks, some more complex than others. The basic machinery is the same. The essential work is that of classification.

Networks are looking for patterns in collections of terms, each term is reduced to a token. In this machine words have no meaning except for their patterned existence within training data. The label “artificial intelligence” applied to such machinery is mostly BS.

###Types of Neural Networks
A shallow neural network contains mainly three types of layers – input layer, hidden layer, and output layer.

Deep Neural Networks are more complex neural networks in which the hidden layers performs much more complex operations than simple sigmoid or relu activations. Different types of deep learning models can be applied in text classification problems.

####Convolutional Neural Network
In Convolutional neural networks, convolutions over the input layer are used to compute the output. This results in local connections, where each region of the input is connected to a neuron in the output. Each layer applies different filters and combines their results.

####Recurrent Neural Network: LSTM
Unlike Feed-forward neural networks in which activation outputs are propagated only in one direction, the activation outputs from neurons propagate in both directions (from inputs to outputs and from outputs to inputs) in Recurrent Neural Networks. This creates loops in the neural network architecture which acts as a ‘memory state’ of the neurons. This state allows the neurons an ability to remember what have been learned so far.

The memory state in RNNs gives an advantage over traditional neural networks but a problem called Vanishing Gradient is associated with them. In this problem, while learning with a large number of layers, it becomes really hard for the network to learn and tune the parameters of the earlier layers. To address this problem, A new type of RNNs called LSTMs (Long Short Term Memory) Models have been developed.

####Recurrent Neural Network: GRU
Gated Recurrent Units are another form of recurrent neural networks.

####Bidirectional RNN
RNN layers can be wrapped in Bidirectional layers as well.

####Recurrent Convolutional Neural Network
Once the essential architectures have been tried out, one can try different variants of these layers such as recurrent convolutional neural network. Another variants can be:

*  Hierarchical Attention Networks
*  Sequence to Sequence Models with Attention
*  Bidirectional Recurrent Convolutional Neural Networks
*  CNNs and RNNs with more number of layers

### 2019-02-23 docs/hiding/classifier_oldversion.Rmd
update

--- 2019-02-05 draft.Rmd
@@
 ---
-title: "Data Import and Initial Exploration"
+title: "Classifier Development"
 author: "Dan Johnson"

### 2019-02-23 docs/hiding/reference_tables.Rmd
update

@@
 ---
-title: "reference_tables"
+title: "Reference Tables"
 author: "Dan Johnson"
@@
 
-## Reference Tables
-
+#TA Comment Categories
 Labels in **comment.subject** are routinely recoded as:

### 2019-02-23 hiding/Xclassifier_oldversion.Rmd
update

--- 2019-02-23 docs/hiding/classifier_oldversion.Rmd
@@
 date: "12/17/2018"
-output: html_document
 bibliography: library.bib
@@
 These are listed on the **[LINK] Project Outline** page.
-
-####\
-
-#Part 1. Initial Dataset
-The development dataset is an Excel table of 10,808 rows.
-One column has "TA comments" extracted from student lab reports in Spring 2018.
-Each TA comment is stored as a separate row of the table, and is a mixed alphanumeric string of one or more words, numbers, and punctuation.
-Other columns record unique report, student, and TA IDs; grade assigned to the report; other standardized information about the original report from which the comment was extracted; and the hand-coded subject and structure of each comment.
-
-Data were anonymized in Excel prior to conversion to CSV format. Names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings **[LINK]**.
-
-####\
-
-#Part 2. Data Pre-Checks
-Prior to import, review the data table **in Excel**; check that:
-
-1.  Student and TA names have been recoded using pre-assigned anonymous IDs.
-2.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.rmd"
-3.  Data table headers match those listed below. The same list is in the file "code_column_names.rmd"
-
-Column Number|Column Name|Accepted Values or Format
-------------|----------------------|-----------------------
-1|unique.record|Sp18.00001 (semester and year, then order in dataset)
-2|report.id|R_0fanqHFhaY2yLT7 or similar
-3|sort|remnant of coding
-4|report.title (empty)|string
-5|student|Std_drVZ2x1W (format is Std\_nnn)
-6|course|113, 114, or 214
-7|ta |TA_bc8EVX09 (format is TA\_nnn)
-8|lab|113: allocation, betta; 114: manduca, frog; 214: enzymes (old) photosynthesis, physarum, cyanophage (new)
-9|tag|first, second
-10|type.TA|submission, revision
-11|grade.TA|A, B, C, F
-12|grading.time|time in minutes
-13|Rank|remnant of coding
-14|hypothesis.ok|Yes, No
-15|data.ok|Yes, No
-16|citation.ok|Yes, No
-17|interpretation.ok|Yes, No
-18|organization.ok|Yes, No
-19|techflaws.ok|Yes, No
-20|writing.ok|Yes, No
-21|comments.incorporated|Yes, Somewhat, No
-22|ta.comment|text string
-23|code.subject|values in Ref. Tables
-24|code.structure|values in Ref. Tables
-25|code.locus|values in Ref. Tables
-26|code.scope|values in Ref. Tables
-27|code.tone|values in Ref. Tables
-28|code.notes|values in Ref. Tables
-
-Once all data are in correct format, export Excel file to CSV with name formatted as **"coded_full_comments_dataset_SemYear.csv"**
 

### 2019-02-23 hiding/projectnotes.Rmd
update

--- 2018-12-16 project.Rmd
@@
 ---
-title: "Project"
+title: "Project Notes"
 author: "Dan Johnson"
@@
 
-# Abstract
-
-####\
-
-# Introduction
-## Background
-Undergraduate introductory STEM courses typically include technical writing training.
-Often it is treated as an incidental skill that each student develops "along the way." Or, the instructional strategy is  haphazardly designed, implemented, and executed.
-Adding to the challenge, faculty may not be directly responsible for helping students develop writing skills.
-
-Technical writing training usually is part of a lab course associated with a lecture class.
-These lab sections may be be led by graduate teaching assistants or even advanced undergraduates.
-These instructors may have under-developed writing skills themselves, or lack pedagogical content knowledge needed to guide students effectively.
-
-In 2014 we launched an automated data collection process we call SAWHET that collects systematic observations and sample data from student writers and graduate instructors.
-Our resarch goal is to better understand student and instructor behaviors within the technical writing instructional space.
-Our practical goal is to develop techniques for teaching technical writing that:
-
-*  Can be implemented at large scale by less experienced teaching assistants;
-*  Emphasize student development through holistic coaching;
-*  Shift student and instructor attention away from simple copy-correction towards larger writing issues; and
-*  Embody other best practices outlined in nearly 40 years of research literature on "writing in the disciplines."
-
-We have a structured database of >5000 laboratory reports written by undergraduate students for introductory biology courses.
-In addition we have all comments that a TA made on the individual reports, plus associated metadata:
-
-*  Context. Was it first or second report of the semester? Initial submission or revision? What course and topic in the course? What year in college was the author?
-*  Instructor information. Which TA provided the feedback? What score was assigned, and what what the rationale behind that score?
-
-We are using multiple text-mining methods to map key features of student writing, as well as connections and correlations between those features.
-These insights then inform how we teach technical writing and train graduate instructors.
-
-####\
-
-## Context for This Project & Prior Work
-Students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead.
-To that end, we train graduate teaching assistants to use a **global first, specific later** approach to grading undergraduate technical writing.
-We ask TAs to spend a limited block of time writing fewer comments that focus on broader issues.
-Ideally, TAs provide no more than 3-5 comments per text page, and comments should encourage student reflection.
-As students' writing improves, the TA can focus on smaller issues.
-
-This strategy is well-supported by prior research, but we are uncertain how well TAs can implement the recommended strategy.
-One of our ongoing questions has been, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**
-
-In Summer 2018, we began analyzing TAs' comments from student papers.
-We extracted and tabulated all comments from approximately 500 pairs of reports (initial submissions and revisions) collected during a single semester in 3 different courses.
-Initial data cleaning involved removing duplicates, splitting comments addressing multiple independent writing issues, and removing comments lacking any context for judging intent.
-In all, the final tabulated dataset included ~11,000 separate comments.
-
-We next used a subset of 100 randomly selected comments to build an initial qualitative codebook for sorting comments into types.
-The initial codebook was re-tested using 1100 additional randomly selected comments.
-Using the final codebook, we sorted all 11,000 comments into separate categories (see LINK for complete codebook.)
-
-Coded TA comments were sorted on two axes:
-
-* __Subject__: did the comment focus mainly on Basic criteria, Technical issues, Writing quality, Logic, or Other issues
-* __Structure__: was the comment mainly Copy correction, General or Specific information, Pointing to another resource, or Holistic coaching for improvement?
-
-The sorted contingency table provided us with a relative estimate of TA effort.
-
-####\
-
-## The Central Problem of This Project
-Qualitative coding of TA comments is informative but unsustainable.
-Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort.
-Even with the codebook, individual coders will require training to achieve any accuracy.
-Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.
-
-####\
-
-# Literature Review
 ##General Approach to Initial Dataset
-The following excerpt from *Text Mining with R: A Tidy Approach* (Julia Silge and David Robinson) summarizes how extended text (in this case, TA comments) can be organized as a tidy matrix.*All emphases are mine.*
+This excerpt from *Text Mining with R: A Tidy Approach* (Julia Silge and David Robinson) summarizes how extended text (in this case, TA comments) can be organized as a tidy matrix.*All emphases are mine.*
 
@@
 
-# Methods
+## How to create files in R, then push to GitHub as a complete repo
+Instructions for doing it on this website. [https://cfss.uchicago.edu/git05.html]()
 
@@
 
-## Data
+####\
 
@@
 
-## Analysis
+####\
 
 ####\
-
-# Results
-
-####\
-
-# Discussion
-
-####\
-
-# Conclusion
-
-####\
-
-# References
-
-####\

### 2019-02-23 litreview.Rmd
update

--- 2019-02-21 litreview.Rmd
@@
 title: "Literature Review"
-author: "Dan Johnson"
 date: "2/21/2019"
-output: html_document
+bibliography: library.bib
+biblio-style: apalike
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, child="_setup.Rmd"}
 ```
 
-##Literature base
+#Overview
+This page links out to sub-pages that describe various background topics. Follow the links for details.
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+Description: https://adanieljohnson.github.io/default_website/pagename.html
+
+Technical Writing Training
+
+Text Classification Strategies
+
+####\

### 2019-02-23 reference_tables.Rmd
update

---
title: "Automated Classification of TA Comments on Student Writing"
author: "Dan Johnson"
date: "2/3/2019"
output: html_document
---

```{r, child="_setup.Rmd"}
```

#TA Comment Categories
Labels in **comment.subject** are routinely recoded as:

*  0_null
*  1_basic
*  2_writing
*  3_technical
*  4_logic
*  5_exposition
*  6_misconduct
*  7_administrative
*  8_nobasis
*  9_split
*  10_context
*  11_flag
*  12_narrative
*  13_sciname

Labels in **comment.structure** are routinely recoded as:

*  0_null
*  1_pointer
*  2_copyedit
*  3_general
*  4_specific
*  5_rationale
*  6_holistic
*  7_idiomatic
*  8_nobasis

### 2019-02-24 codebook.rmd
update

---
title: "Comments Codebook"
date: "2/23/2019"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

We evaluating TAs’ comments on student reports to see if and how their comments change as students get more feedback from automated system.

To build the data table, commented reports from a single semester (MS Word .docx format) were processed using BaSH scripts to extract relevant file data (author, ID#, etc.) and individual TA comments from the raw XML, and write the data to a single CSV file. TA comments containing multiple points or suggestions were replicated and edited to produce a working data table with only one comment per entry row.

Inclusion and exclusion criteria for assigning TA comments to categories and sub-categories was established iteratively using two training sets of 120 and 1200 comments selected randomly from the full dataset of ~11,000. Each main category (Subject, Structure, etc.) was divided into 4-8 loosely defined provisional sub-categories.Each comment from the 120-item training set was assigned to a provisional sub-category within each category. Sub-categories from different main categories were not mutually exclusive; for example, if a comment was assigned to the “Writing Quality” sub-category under the “Subject” category, that comment could be assigned to any of the provisional sub-categories in “Structure,” any of the sub-categories in “Locus,” etc.

When comments could not be assigned with high confidence to a provisional sub-category, defining features of existing provisional sub-categories were refined further and, if necessary, additional provisional sub-categories established. Then the previously assigned comments were re-evaluated using the updated criteria, until all 120 comments in the training set could be assigned reliably.

Once stable sub-categories were established, examples of TA comments belonging to each sub-category were extracted from the main dataset and combined with finalized criteria to form a working code scheme. The working code scheme was tested and refined again by categorizing the second training set of 1200 comments, to produce the final coding scheme.

To score the full data table, the 1200 scored training comments were rejoined with the unscored comments, and all comments alphabetized. This distributed scored comments randomly within the larger data table, and identified comments with similar wording. This final integrated dataset was used to score the full set of TA comments.

####\

#Subject Category
Is primary emphasis of the comment on:

*  Fundamental pieces
*  How to communicate generally?
*  How to collect, analyze and present data correctly?
*  The underlying rationale for the report/experiment?
*  Other issues?

Below are the choices for identifying the subject of the comment. Criteria for basic, technical, and writing choices are the same as used in the report grading rubric. To simplify sorting, administrative tags used to mark comments for further processing are included in this category.

**Code.Subject** Term|Description
--------------------|-------------------------------------
0. Null | Used to reset a previously tagged comment
1. Basic Criteria | Relates to 1 or more of 5 basic criteria
2. Writing Quality | Relates to writing mechanics, quality
3. Technical and Scientific | Relates to a technical element
4. Logic and Thinking | Focuses on logic and argument
5. Praise or Concern | Statements directly to student, either positive or negative. Focused on student mindset more than text.
6. Misconduct | Focuses on plagiarism, academic misconduct issue
7. Policy, Administrative | Reiterates a standing class or course policy
8. No basis to judge | Cannot tell from text or context what comment means.
9. Split | If 2+ comments are together, duplicate line, split it, then annotate
10. Check Context | Go back and look at original document, then re-code
11. Flag | Comment is not appropriate for some reason. Includes factual errors, non-standard policies, deviations from norms.
12. Narrative Comments | Comment broadly addresses report
13. Scientific Name | Comment specifically about use of scientific names
14. Other | Comment does not meet any of above criteria

####\

**0. Null** is an administrative marker used to reset a comment.

**1. Basic Criteria** establish whether report is ready for grading, or unacceptably flawed. Comments in this category focus on the 5 basic requirements that ALL scientific writing must meet, plus the academic honesty requirement. Includes all comments about:

*  Written work represents the student’s own scientific research or work; not copied.
*  There is a clear rationale, hypothesis, or purpose of the study PRESENT (quality is scored later.)
*  Written work is complete, appropriately organized for audience, and all essential elements are present.
*  Summarized data are PRESENT. Data have been analyzed in some way.
*  There is an interpretation of the results.
*  Discussion states clearly whether or not hypothesis is supported and why.
*  Study is based within a larger body of literature; citations in Introduction, Discussion.

**2. Writing Quality** comments focus on written communication issues, including general conventions of scientific writing as a genre. Includes all comments on:

*  Writing style is appropriate for the target audience.
*  Style, format, organization of writing matches the expectations of a scientific audience.
*  Clear, concise wording.
*  Precise language appropriate to scientific audience.
*  Technically presented; avoids casual, imprecise, emotional language.
*  Written work is free of major writing errors (spelling, grammar, punctuation.)
*  Citations are presented consistently and professionally throughout the text and in the list of works cited section.
*  Citations are presented consistently, professionally in text and Works Cited section.

Comments about the following belong here.

*  Repeated info 
*  Errors in referencing figures, tables
*  Word choice generally (not tech. vocab.)
*  Extraneous information
*  “Include a sentence about…”
*  What goes in which sections
*  HTML code-based errors are writing quality pointers (should catch in general editing)
*  General to specific
*  “Include a sentence about…” can be categorized many ways. Most often will be general writing, but could be logic or technical.

**3. Technical and Scientific Issues** comments focus on whether: methods are communicated clearly and correctly; data and analyses are clear, appropriate, accurate, and unbiased; and mechanics specific to technical writing and data presentation are executed correctly. Includes all comments such as:

*  Methods are appropriately defined and explained, given the student’s research question.
*  The data analysis is clear, appropriate, accurate and unbiased.
*  Statistical summaries are executed, interpreted, and reported correctly.
*  Data are summarized, reported, displayed correctly.
*  Tables and figures are clear, effective, and informative.
*  Errors in interpreting data belong here, not in Logic.
*  Numbers and units
*  Reporting raw data
*  Why the particular study organism goes here.
*  Corrections to:
*  Use of technical terms
*  Factual information
*  Prove vs. support goes here.

**4. Logic and Thinking** comments are not part of our lab report grading model currently, but still important to assess. Comments tagged with this subject focus on: 1) quality, accuracy of claim-evidence-reasoning chain; 2) use of hypothetico-deductive reasoning model; 3) consistency of arguments with internal, external evidence. Includes all comments about:

*  Written work clearly articulates the student’s thinking regarding research goals.
*  Research question is clear.
*  Rationale, background for research question is clear.
*  Work is placed in context of RELEVANT scientific literature.
*  Cited literature support, develop a larger argument.
*  Statement of hypothesis is clear, appropriately organized.
*  Written work is logically consistent.
*  Student’s thought process about how parts connect is clear.
*  Logic that connects the arguments and conclusions is sound, clear.
*  All parts contribute to understanding outcomes. No distracting arguments.
*  Written work interprets the results skillfully, conservatively.
*  Implications of the findings are placed in a reasonable context.
*  Strong, integrated ending goes here, BUT if comment is about language rather than thinking, goes under Writing.

####\

#Structure Category
Is the comment idiomatic or informative? Is there general or specific information contained in comment? Is it instructional only, or does the comment foster broader thinking?

**Code.Structure** Term|Description
--------------------|-------------------------------------
0. Null | Used to reset a previously tagged comment
1. Pointer | Simple notation indicating an error
2. Copy Correction | Simple instructions for single correction
3. General info | Provides general instructions
4. Specific info | Provides specific instructions
5. Rationale | Comment contains process info only
6. Holistic | Comment has both content and process info
7. Idiomatic | Comment does not match standard requirements
8. No basis to judge | Not enough information to categorize

**1. Pointer** is simple notation indicating there is an error, or that a correction is needed. Includes comments like:

*  ??
*  Misspelling, typo, format, check, etc.
*  *Repeated phrase. (Only points to immediate error; compare to below)

**2. Copy correction** comments provide specific instructions for correcting or modifying copy in ONE SPECIFIC location, but do not address the logic or reason. Includes comments like:

*  Stylistic corrections, word choices
*  An apostrophe makes this a possessive noun, not a plural one. 
*  Colloquial (word choice) (but no explanation)
*  No direct quotes. Paraphrase.
*  Always capitalize “P”
*  Use this particular wording
*  Instructions to add or delete a word, phrase, image, element
*  Omit/delete/remove \_\_\_.
*  Add \_\_\_(description)\_\_\_. 
*  Unnecessary/ redundant
*  Provide specific data or information this way (p value, statistical tests, etc.)
*  Add/remove error bars
*  Citations and cross-references
*  Include a citation here. OR, This must be cited.
*  Citation format OR Citation format is [last name for first author : year]
*  Instructions on correcting a single SPECIFIC citation directly
*  Refer to this figure, table, etc. 
*  Not true (corrects error of fact without further information.)
*  Delete this repeated phrase. (Specific change to make.)

**3. General info** comments share GENERAL instructions or knowledge, but have no guidance on where specific changes are needed or how to apply, and offer no underlying rationale. Includes comments like:

*  For your lab report you should be concise and straight forward. 
*  Just include the relevant information.
*  Some sections are missing (not clear which ones)
*  Ambiguous, awkward
*  Rewrite this/ condense this/ avoid this word or term generally
*  *Avoid repeating phrases. (More instructive than above but not specific)
*  Tense, voice rules in general go in this group
*  Methods should be past tense
*  General corrections to citations, references without providing specifics
*  Use citation style in lab resource guide, & like was on the writing quiz.
*  See p.42 in the BioCore Resource Guide for correct in-text citation format 
*  Need citations IN THIS SECTION (specific when points to spot IN A section)
*  Reference your figures and tables.
*  Raw data (is specific only if provides additional context.)
*  Avoid recipe style (with no further explanation)
*  Numbers need units
*  What about statistics?
*  Comment on using loaded words: prove, significant,

**4. Specific info** comments provide SPECIFIC corrective info/context, but lack a starting point for broader thinking. Includes comments like:

*  Provides clear organization for whole section.
*  Put x, then y, then z.
*  Defines larger specific changes in text.
*  At the end of the discussion indicate possible implications of your experiment at a broader scale
*  Did not say if hypothesis was supported (specific b/c of content, context)
*  Add this missing section; you are missing the \_\_\_ section. 
*  \_\_\_Section\_\_\_ should contain \_\_\_x\_\_\_.
*  Move/add \_\_\_ to BEFORE/AFTER \_\_\_. 
*  What does “X” (error bars) mean? (simple interpretation)
*  Clarify this step in procedure or analysis
*  Move a specific table of data to text, or vice versa.
*  Do not interpret your results in Results section (specific b/c of guidance)
*  You need to explain \_\_\_ better. (More specific instructions that just “awkward” or “revise”.
*  *Avoid repeating phrases. Change the order, make it nicer to read. (Clear way to implement instructions, relative to two examples above, but does not invite deeper thought.)
*  You should try to paraphrase anything more than 2-3 words, although if it is an incredibly common phrase, you can use it without fear of plagiarism. 
*  This specific citation is not relevant to your argument/ point.
*  Avoid recipe style. Look at other articles to see how they did it.
*  Corrections to data presentation
*  Raw data. Need to summarize.
*  These numbers need units or additional information.
*  *Not true. Look in \_\_\_x\_\_\_ to learn more. (Provides a source for details)

**5. Rationale** comments are a question or statement that prompts for, points to, or guides thinking, but lacks any context. Uncommon. Includes comments like:

*  Why, when, how? What is your reasoning?

**6. Holistic** comments have BOTH specific knowledge/info and guidance/rationale that invites thinking. Specific comments provide both sources for information (what) plus invitation to think (why) in combination. General comments have BOTH general knowledge/info and guidance/rationale. More open ended, will require deeper thought. (General information to guide thought (what) plus invitation to think (why)). Expect these to be most challenging to write and to respond to. Includes comments like:

*  Did you mean for each leg before and after the injection?
*  What is this in the moth life cycle?
*  Are you sure it is the correct tense for this section? Check with other primary lit.
*  Any primary literature sources ( articles) that deal with interspecific interactions in betta fish? It would be very useful to cite and talk about those here, if there are.
*  Why is this a helpful species to study?

####\

####Contingency Table For Holistic Scores

Scores need to have BOTH information and process (what and why) to be holistic

\ | Process ("why" or “thinking” part) |
---- | ---- | ----
**Amt. Info Provided**| *Absent* | *Present*
*Limited or None* | Idiomatic, Pointer, Null | Rationale
*General* | General info | Holistic - general
*Specific* | Specific info | Holistic - specific

####\

**7. Idiomatic** comments point out discipline-dependent or personally idiomatic issues we do not grade on. Includes personal preferences, outlier issues, comments about italics, underlines, etc. Includes comments like:

*  Always use correct format for scientific names, i.e., italicize or underline. 
(Idiomatic because not required in our grading scheme. Incidentally, this is not accurate form either, which is one answer to ongoing arguments.)
*  You provided both scientific and common names for your plant species, so can use common names throughout the rest of the report including the title and abstract. Revise accordingly. (Idiomatic because this is not standard to all fields)

####\

##Example of How to Parse a Comment
**Original TA comment:**

Ah, so it didn’t move directly towards the Physarum. Here it’s important to make that distinction in how you took measurements in your methods section. Did you go based on leading edge? If so, this data point would probably be inconclusive in the results section. Did you go based on a central axis upon which you took measurements? Instead, that’d be different.

There are two questions in the comment, which suggests that it is a holistic comment to pull knowledge out through thinking. However, the TA comment provides some interpretation in the next sentence. So this comment is seeking or providing specific information, while also directly fostering student’s deeper thinking. It would be tagged as “Holistic."

### 2019-02-24 Text_classifier_models.Rmd
update

--- 2019-02-23 Text Classifier Models.Rmd
@@
 title: "Text Classification Primer"
-author: "Dan Johnson"
-date: "1/9/2019"
-output: html_document
+date: "2/21/2019"
+bibliography: library.bib
+biblio-style: apalike
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, child="_setup.Rmd"}
 ```
 
-For this text classification project I am using an initial codebook and corpus of hand-coded TA comments to test various approaches to semi- and fully automated classification methods and, if appropriate, revise the underlying classification scheme.
+#Overview
+This page provides general background on theory and approaches to text classification, and descriptions of several classification methods.
 
-This page provides background on general mechanics of text classification and various classifier methods.
-Other pages describe the dataset structure [LINK] (TA comments corpus), and [LINK] development of the initial codebook.
+####\
 
-#Workflow
+#Basic Workflow
 Text classification problems can be broken down into four overlapping sub-problems. Each piece may be more or less important to the overall problem, but all need to be considered:
@@
 4.  Evaluating the outcomes
+
+####\
 
@@
 
+####\
+
 ###Lexical Diversity
@@
 
-Consider a speaker, who uses the term "allow" multiple times throughout the speech, compared to an another speaker who uses terms "allow, concur, acquiesce, accede, and avow"" for the same word.
+>Consider a speaker, who uses the term "allow" multiple times throughout the speech, compared to an another speaker who uses terms "allow, concur, acquiesce, accede, and avow"" for the same word.
 The latter speech has more lexical diversity than the former.
@@
 
-Lexical diversity, in simple terms, is a measurement of the breadth and variety of vocabulary used in a document.
+>Lexical diversity, in simple terms, is a measurement of the breadth and variety of vocabulary used in a document.
 The different measures of lexical diversity are TTR, MSTTR, MATTR, C, R, CTTR, U, S, K, Maas, HD-D, MTLD, and MTLD-MA.
@@
 https://subscription.packtpub.com/book/big_data_and_business_intelligence/9781783551811/2/ch02lvl1sec16/lexical-diversity
+
+####\
 
@@
 These all are general scales; other scores may not be available in R packages.
+
+####\
 
@@
 Three of these tools are described further below.
+
+####\
 
@@
 
+####\
+
 ####fastText
@@
 https://fasttext.cc/
+
+####\
 
@@
 
+####\
+
 ##Sub-Problem Two: Topic Categorization and Modeling Methods
@@
 Topic modeling is an iterative/ generative process that does not require any prior knowledge of the documents or relationships.
+
+####\
 
@@
 a probability model that takes the clustering into account.
+
+####\
 
@@
 
+####\
+
 ##Sub-Problem Three: Choosing a Text Classifier Method
@@
 
+```{}
 Understanding how chatbots work can help us understand text classifiers, because the classifier is the fundamental piece that makes the chatbot work.
@@
 
-###Pattern Matchers
+Pattern Matchers
 Early chatbots used **pattern matching** to classify text and produce a response.
@@
 A simple pattern matching definition:
-```
+
 <aiml version = "1.0.1" encoding = "UTF-8"?>
@@
 </aiml>
-```
 
@@
 
-###Algorithm-Based Classification
+Algorithm-Based Classification
+
 Creating a brute-force pattern matcher is daunting: for each unique input a pattern must be available to specify a response. An algorithm-based method approaches the problem mathematically.
@@
 Look at this sample training set:
-```
+
 class: weather
@@
     "how is it going?"
-```
 
 Let’s classify a few sample input sentences:
-```
+
 input: "Hi there"
@@
  classification: weather (score=4)
-```
 
@@
 It’s "naive" but surprisingly effective.
+```
 
 I will be using an algorithm-based classification strategy for this project.
+
+####\
 
@@
 
+####\
+
 ###Designing an Evaluation Workflow
@@
 *  Select best performing classifiers and either refine, or operationalize for new datasets
+
+####\
 
@@
 Algorithms also can be separated into un-supervised and supervised methods. 
-
 **Unsupervised learning** uses test data that has not been labeled, classified or categorized.
@@
 
+####\
+
 ###Naive Bayes
@@
 
+####\
+
 ###Linear Classifier
@@
 Such classifiers work well for practical problems such as document classification, and more generally for problems with many variables (features), reaching accuracy levels comparable to non-linear classifiers while taking less time to train and use.
+
+####\
 
@@
 
+####\
+
 ###Support Vector Machine (SVM) Model
@@
 *  SVM multiple kernels
+
+####\
 
@@
 
+####\
+
 ###Linear Discriminant Analysis
@@
 
+####\
+
 ###k-Nearest Neighbors
@@
 
+####\
+
 ###Similarity Learning
@@
 It has applications in ranking, in recommendation systems, visual identity tracking, face verification, and speaker verification.
+
+####\
 
@@
 https://nlp.stanford.edu/IR-book/html/htmledition/rocchio-classification-1.html
+
+####\
 
@@
 
+####\
+
 ####Error Sources in Ensemble Learning (Variance vs. Bias)
@@
 
+####\
+
 ####Commonly used Ensemble learning techniques
@@
 
+####\
+
 ####Random Forest Models
@@
 https://www.analyticsvidhya.com/blog/2014/06/introduction-random-forest-simplified/
+
+####\
 
@@
 
+####\
+
 ###Hierarchical Clustering
 In data mining and statistics, hierarchical clustering (also called hierarchical cluster analysis or HCA) is a method of cluster analysis which seeks to build a hierarchy of clusters.
+
+####\
 
@@
 This is known as nearest centroid classifier or Rocchio algorithm.
+
+####\
 
@@
 Another possibility is the latent Dirichlet allocation model, which divides up the words into D different documents and assumes that in each document only a small number of topics occur with any frequency.
+
+####\
 
@@
 
+####\
+
 ##Strategies for Improving Text Classification Models
@@
 
+####\
+
 ##Evaluation
@@
 
+####\
+
 ###Gini Index
 One of the most common methods for quantifying the discrimination level of a feature is the use of a measure known as the gini-index.
+
+####\
 
@@
 
-***
-***
+####\
 
@@
 
+####\
+
 ###Types of Neural Networks
@@
 
+####\
+
 ####Convolutional Neural Network
@@
 Each layer applies different filters and combines their results.
+
+####\
 
@@
 
+####\
+
 ####Recurrent Neural Network: GRU
@@
 
+####\
+
 ####Bidirectional RNN
 RNN layers can be wrapped in Bidirectional layers as well.
+
+####\
 
@@
 *  CNNs and RNNs with more number of layers
+
+####\

### 2019-02-24 build.Rmd
update

@@
 title: "The Build"
-author: "Dan Johnson"
 date: "2/22/2019"
-output: html_document
+bibliography: library.bib
+biblio-style: apalike
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, child="_setup.Rmd"}
 ```
@@
 #Overview
-This page links to sub-pages that describe how various components used in the project were built. Follow the links for details.
+This page links to sub-pages that describe the build process for various components used in the overall project. Follow the links for details.
 
-Extracting TAs' comments: https://adanieljohnson.github.io/default_website/docxcommentextraction.html
+Extracting TAs' comments (*draft*): https://adanieljohnson.github.io/default_website/docxcommentextraction.html
+
+Codebook for classifying TAs' comments: https://adanieljohnson.github.io/default_website/codebook.html
 
@@
 
-Current Naive Bayes classifier: https://adanieljohnson.github.io/default_website/NB_2r2c.html
+Current version of the Naive Bayes classifier: https://adanieljohnson.github.io/default_website/NB_2r2c.html
+
+Expanding the Naive Bayes classifier to an Ensemble Model (*draft*): https://adanieljohnson.github.io/default_website/Ensemble_Notes.html
 
 ####\
-
-#Describe an Element
-
-####\
-
-#Describe an Element
-
-####\

### 2019-02-24 codebook.rmd
update

@@
 ---
-title: "Comments Codebook"
+title: "The TA Comments Codebook"
 date: "2/23/2019"
@@
 
-We evaluating TAs’ comments on student reports to see if and how their comments change as students get more feedback from automated system.
+We classify TAs’ comments on student reports to see if and how their comments change as students get more feedback from automated system.
 
-To build the data table, commented reports from a single semester (MS Word .docx format) were processed using BaSH scripts to extract relevant file data (author, ID#, etc.) and individual TA comments from the raw XML, and write the data to a single CSV file.
+To build a TA comments data table, commented reports from a single semester (MS Word .docx format) are processed using R or BaSH scripts to extract relevant file data (author, ID#, etc.) and individual TA comments from the raw XML, and write the data to a single CSV file.
 TA comments containing multiple points or suggestions were replicated and edited to produce a working data table with only one comment per entry row.
 
-Inclusion and exclusion criteria for assigning TA comments to categories and sub-categories was established iteratively using two training sets of 120 and 1200 comments selected randomly from the full dataset of ~11,000.
+Inclusion and exclusion criteria for assigning TA comments to categories and sub-categories were established iteratively using two training sets of 120 and 1200 comments selected randomly from the full dataset of ~11,000.
 Each main category (Subject, Structure, etc.) was divided into 4-8 loosely defined provisional sub-categories.Each comment from the 120-item training set was assigned to a provisional sub-category within each category.
-Sub-categories from different main categories were not mutually exclusive; for example, if a comment was assigned to the “Writing Quality” sub-category under the “Subject” category, that comment could be assigned to any of the provisional sub-categories in “Structure,” any of the sub-categories in “Locus,” etc.
+
+Note: sub-categories from different main categories are not mutually exclusive; for example, if a comment is assigned to the “Writing Quality” sub-category under the “Subject” category, that comment can be assigned to any of the provisional sub-categories in “Structure,” any of the sub-categories in “Locus,” etc.
 
@@
 
-Ah, so it didn’t move directly towards the Physarum.
+>Ah, so it didn’t move directly towards the Physarum.
 Here it’s important to make that distinction in how you took measurements in your methods section.

### 2019-02-24 docxcommentextraction.Rmd
update

@@
 
-#Pseudo-Code and Sources
+The testing dataset used in this project was created using BaSH scripting and one-off tools. This page documents progress on a workflow to replicate that process using R tools.
+
+####\
+
+#General Approach
 Create a list of files to cycle through OR put all files to process in ONE folder.
-For each file, use 'read_docx', then 'docx_extract_all_cmnts'
+For each file, use 'read_docx', then 'docx_extract_all_cmnts'.
 Use 'do.call("rbind", list(dataframes to merge))' to cycle through and append the extracted comments to ONE dataframe
+
+####\
+
+#Code Sources, Examples
 
@@
 Load required Libraries
+
+####\
+
+#Sample Working Code
 ```{r}
@@
 
-```{r}
+Call in Word files using ```read_docx```. Use ```docx_extract_all_cmts``` to pull out comments and ```rbind``` to combine them.
+
+To merge more than two dataframes within R, combine an ```rbind``` call with ```do.call```.
+The 'do.call' function uses any existing R functions, such as 'rbind', but instead of passing just two dataframes, it passes multiple data frames.
+Then use ```write.csv2``` to write the data to a CSV format.
+
+```{}
 Commentframe1<-read_docx("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/DOCX_comment_extraction/TestingReport.docx", track_changes = NULL)
+comments1<-docx_extract_all_cmnts(Commentframe1, include_text = FALSE)
 
 Commentframe2<-read_docx("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/DOCX_comment_extraction/TestingCommentExtraction.docx", track_changes = NULL)
-
-comments1<-docx_extract_all_cmnts(Commentframe1, include_text = FALSE)
 comments2<-docx_extract_all_cmnts(Commentframe2, include_text = FALSE)
 
-#Merging more than two dataframes
-#To merge more than two dataframes within R, we can combine our rbind call with do.call.
-# > do.call("rbind", list(dataframes to merge))
-#
-#How this works:
-#  The 'do.call' function allows us to use any existing R functions, such as 'rbind', but instead of passing just two dataframes in this instance, we are able to pass multiple.
-#
-#Therefore, when used alongside rbind, we can use do.call to merge more than two dataframes at once.
-#
-#Putting this all together, we can merge our dataframes back together as follows:
-# > ufoMerged <- do.call("rbind", list(ufo1, ufo2, ufo3, ufo4))
-#
-#We can confirm our data has been successfully merged by checking the number of rows:
-#
-#> NROW(ufoMerged)
-#[1] 80332
+# comments1And2 <- rbind(comments1,comments2)
+# OR
+# comments1And2 <- do.call("rbind", list(comments1,comments2))
 
-#comments1And2 <- rbind(comments1,comments2)
-do.call("rbind", list(comments1,comments2))
+#Check data have been successfully merged by checking the number of rows
+NROW(comments1And2)
+# [1] 32
 
+#When all lines are appended, write the CSV file 
 write.csv2(comments1And2, file = "TAcomments.csv")
+
 ```

### 2019-02-24 initial_exploration.Rmd
update

@@
 title: "Initial Exploration"
-output: html_document
+date: "2/21/2019"
+bibliography: library.bib
+biblio-style: apalike
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, child="_setup.Rmd"}
 ```
@@
 #Overview
-The first question to answer before making deeper comparisons is, **are word frequencies different between TAs' comments from different sub-categories?** To that end, I ran three different analyses of word usage frequency.
+The first question to answer before making deeper comparisons is, **are word frequencies different between TAs' comments from different sub-categories?** This page describes the methods used for this initial exploration.
+
+Results of these analyses are discussed on this page: https://adanieljohnson.github.io/default_website/findings.html
+
+####\
+
+#The Analyses
+I ran three different analyses of word usage frequency.
 
@@
 *  __Word frequency differences based on log-odds ratios__. This highlights terms that are the most different between sub-categories.
-*  __N-grams mapping__. This identifies text strings, phrases, and terms that are unique (or nearly unique) to one sub-category.
+*  __N-grams mapping__ (*draft*). This identifies text strings, phrases, and terms that are unique (or nearly unique) to one sub-category.
 
@@
 One of our earliest observations was that the n-gram "basic criteria" only appeared in TA comments from three different Subject sub-categories.
-This suggests certain terms and n-grams might be eliminated
+Of these, the phrase made up <0.5% of comments in two categories, but in the third category, "basic criteria" made up >5% of all comments.
+This suggests that the "basic criteria" n-gram is essentially diagnostic for that sub-category.
+This suggests other terms and n-grams that are eliminated in the previous two methods may be equally diagnostic for one sub-category.
 
-However it w  the relative frequency was much higher
+The block below describes a pseudo-code approach.
 ```{}
-#Pseudo-code
 Import comments

### 2019-02-25 Text_classifier_models.Rmd
updates

@@
 #Overview
-This page provides general background on theory and approaches to text classification, and descriptions of several classification methods.
+This page provides general background on theory and approaches to text classification. Links below connect to sub-pages that describe more specific topics. Follow the links for details.
+
+Topic Categorization and Modeling Methods: https://adanieljohnson.github.io/default_website/Topic_categorization.html
+
+Choosing a Text Classifier Method: https://adanieljohnson.github.io/default_website/Choosing_classifier.html
+
+Designing a Classifier Evaluation Workflow: https://adanieljohnson.github.io/default_website/Classifier_comparison.html
+
+Strategies for Improving, Evaluating Text Classifiers: https://adanieljohnson.github.io/default_website/Improving_classifier.html
 
@@
 ####\
-
-##Sub-Problem Two: Topic Categorization and Modeling Methods
-Establishing meaningful text classification categories depends on the goals of the project or the question being asked.
-Are relevant categories known already, or do they need to be established?
-What are the main topics
-
-**Curated Categories** (my provisional term until I find the appropriate equivalent in current use) are established by subject matter experts. The established categories can be constructed from:
-
-*  A limited keywords vocabulary (National Library of Medicine MeSH categories for example)
-*  Disciplinary conventions or shared best practices (the Dublin Core is an example)
-*  Prior empirical studies
-*  The dataset itself, using *de novo* iterative coding. This process is described further here [LINK] and here [LINK].
-
-**Topic Modeling** is a group of techniques for identifying groups of words (the "topic")  that best summarize the information in a collection of documents.
-Topic modeling is an iterative/ generative process that does not require any prior knowledge of the documents or relationships.
-
-####\
-
-###Topic Modeling Methods to Identify Features
-####Latent Dirichlet Allocation
-(LDA) is a common method for generating topic model features.
-LDA is an iterative model which starts from a fixed number of topics.
-Each topic is represented as a distribution over words, and each document is then represented as a distribution over topics.
-Although the tokens themselves are meaningless, the probability distributions over words provided by the topics provide a sense of the different ideas contained in the documents.
-
-####Brown Clustering
-A hierarchical clustering algorithm that is applied to text, grouping words into clusters that are assumed to be semantically related by virtue of their having been embedded in similar contexts.
-
-In natural language processing, Brown clustering or IBM clustering is a form of hierarchical clustering of words based on the contexts in which they occur.
-The intuition behind the method is that a class-based language model (also called cluster n-gram model), i.e.
-one where probabilities of words are based on the classes (clusters) of previous words, is used to address the data sparsity problem inherent in language modeling.
-
-Brown groups items (i.e., types) into classes, using a binary merging criterion based on the log-probability of a text under a class-based language model, i.e.
-a probability model that takes the clustering into account.
-
-####\
-
-####Latent semantic analysis
-LSA is a technique in natural language processing of analyzing relationships between a set of documents and the terms they contain by producing a set of concepts related to the documents and terms.
-LSA assumes that words that are close in meaning will occur in similar pieces of text (the distributional hypothesis).
-A matrix containing word counts per paragraph (rows represent unique words and columns represent each paragraph) is constructed from a large piece of text and a mathematical technique called singular value decomposition (SVD) is used to reduce the number of rows while preserving the similarity structure among columns.
-Words are then compared by taking the cosine of the angle between the two vectors (or the dot product between the normalizations of the two vectors) formed by any two rows.
-Values close to 1 represent very similar words while values close to 0 represent very dissimilar words.
-
-In the context of its application to information retrieval, it is sometimes called latent semantic indexing (LSI).
-
-LSA can use a term-document matrix which describes the occurrences of terms in documents; it is a sparse matrix whose rows correspond to terms and whose columns correspond to documents.
-A typical example of the weighting of the elements of the matrix is tf-idf (term frequency–inverse document frequency): the weight of an element of the matrix is proportional to the number of times the terms appear in each document, where rare terms are upweighted to reflect their relative importance.
-
-Applications of LSA include:
-
-*  Comparing documents in low-dimensional space (data clustering, document classification).
-*  Finding relations between terms (synonymy and polysemy).
-Synonymy is the phenomenon where different words describe the same idea.
-Polysemy is the phenomenon where the same word has multiple meanings.
-*  Given a query of terms, translating it into the low-dimensional space, and find matching documents (information retrieval).
-*  Analyzing word association in text corpus.
-
-LSA overcomes two of the most problematic constraints of Boolean keyword queries: multiple words that have similar meanings (synonymy) and words that have more than one meaning (polysemy).
-
-LSA is also used to perform automated document categorization.
-Document categorization is the assignment of documents to one or more predefined categories based on their similarity to the conceptual content of the categories.
-Several experiments have demonstrated that there are a number of correlations between the way LSI and humans process and categorize text.
-
-LSA uses example documents to establish the conceptual basis for each category.
-During categorization processing, the concepts contained in the documents being categorized are compared to the concepts contained in the example items, and a category (or categories) is assigned to the documents based on the similarities between the concepts they contain and the concepts that are contained in the example documents.
-
-LSA automatically adapts to new and changing terminology, and has been shown to be very tolerant of noise (i.e., misspelled words, typographical errors, unreadable characters, etc.).[24] This is especially important for applications using text derived from Optical Character Recognition (OCR) and speech-to-text conversion.
-LSA also deals effectively with sparse, ambiguous, and contradictory data.
-
-Text does not need to be in sentence form for LSA to be effective.
-It can work with lists, free-form notes, email, Web-based content, etc.
-As long as a collection of text contains multiple terms, LSA can be used to identify patterns in the relationships between the important terms and concepts contained in the text.
-
-####\
-
-##Sub-Problem Three: Choosing a Text Classifier Method
-The many methods for classifying texts fall into three general categories:
-
-*  Pattern matching strategies (does a particular phrase appear?)
-*  Algorithms (how closely does the query text meet pre-set criteria?)
-*  Neural networks (the network decides its own criteria, aka machine learning)
-
-This edited excerpt nicely describes the overall approach and problems of text classification. It is from: **xxxxx**
-
-```{}
-Understanding how chatbots work can help us understand text classifiers, because the classifier is the fundamental piece that makes the chatbot work.
-Let’s look at the inner workings of a Multinomial Naive Bayes algorithm for text classification and natural language processing (NLP).
-
-This classifier is “naive” because each word is treated as having no connection with other words in the sentence being classified.
-In other words it assumes independence between "features," in this case, words.
-
-For example, in the sentence “the fox jumped over the log”, there is no relationship between “jumped” and “fox” or “log”.
-In NLP, this collection of words is referred to as “a bag of words”.
-While this is incredibly naive, the classifier isn’t attempting to understand the meaning of a sentence, it’s trying to classify it.
-
-Pattern Matchers
-Early chatbots used **pattern matching** to classify text and produce a response.
-This is often referred to as “brute force” as the author of the system needs to describe every pattern for which there is a response.
-
-A standard structure for these patterns is “AIML” (artificial intelligence markup language). Its use of the term “artificial intelligence” is quite an embellishment, but that’s another story.
-
-A simple pattern matching definition:
-
-<aiml version = "1.0.1" encoding = "UTF-8"?>
-   <category>
-      <pattern> WHO IS ALBERT EINSTEIN </pattern>
-      <template>Albert Einstein was a German physicist.</template>
-   </category>
-
-   <category>
-      <pattern> WHO IS Isaac NEWTON </pattern>
-      <template>Isaac Newton was a English physicist and mathematician.</template>
-   </category>
-
-   <category>
-      <pattern>DO YOU KNOW WHO * IS</pattern>
-      <template>
-         <srai>WHO IS <star/></srai>
-      </template>
-   </category>
-</aiml>
-
-The machine then produces:
-
-> Human: Do you know who Albert Einstein is
-> Robot: Albert Einstein was a German physicist.
-
-It knows who a physicist is only because his or her name has an associated pattern.
-Likewise it responds to anything solely because of an authored pattern.
-Given hundreds or thousands of patterns you might see a chatbot “persona” emerge.
-
-In 2000 a chatbot built by John Denning and colleagues using this approach was in the news for passing the "Turing test." It emulated the replies of a 13 year old boy from Ukraine (broken English and all).
-I met with John in 2015 and he made no false pretenses about the internal workings of this automaton.
-It may have been “brute force” but it proved a point: parts of a conversation can be made to appear “natural” using a sufficiently large definition of patterns.
-It proved Alan Turing’s assertion, that this question of a machine fooling humans was “meaningless”.
-
-An example of this approach used in building chatbots is PandoraBots, who claim over 285k chatbots have been constructed using their framework.
-
-Automated phone voice response systems that recognize patterns of spoken words and direct calls accordingly use a similar approach.
-
-Algorithm-Based Classification
-
-Creating a brute-force pattern matcher is daunting: for each unique input a pattern must be available to specify a response. An algorithm-based method approaches the problem mathematically.
-
-One text classification algorithm is "Multinomial Naive Bayes." Given a training set of sentences, each belonging to a class (that is, a particular group), we can count the occurrence of each word in each class, account for its commonality and assign each word a score for each class.
-Factoring for commonality is important: matching the word “it” is considerably less meaningful than a match for the word “cheese”.
-When we input a new sentence, scores for each word in the new sentence are compared to scores for those same words in the training set.
-The class with the highest OVERALL score for the input sentence is the class to which the input sentence most likely belongs.
-
-Look at this sample training set:
-
-class: weather
-    "is it nice outside?"
-    "how is it outside?"
-    "is the weather nice?"
-
-class: greeting
-    "how are you?"
-    "hello there"
-    "how is it going?"
-
-Let’s classify a few sample input sentences:
-
-input: "Hi there"
- term: "hi" (no matches in either class)
- term: "there" (matches a word in class: greeting)
- classification: greeting (score=1)
-
-input: "What’s it like outside?"
- term: "it" (matches words in class: weather (2), greeting(1))
- term: "outside (class: weather (2) )
- classification: weather (score=4)
-
-Notice that the classification for "What’s it like outside" found a term in the "greeting" class but the term similarities to the "weather" class produced a higher score.
-By using an equation we are looking for word matches given some sample sentences for each class, and we avoid having to identify every pattern.
-
-The classification score produced identifies the class with the highest term matches (accounting for commonality of words) but this has limitations.
-A score is not the same as a probability, a score tells us which intent is most like the sentence but not the likelihood of it being a match.
-Thus it is difficult to apply a threshold for which classification scores to accept or not.
-Having the highest score from this type of algorithm only provides a relative basis, it may still be an inherently weak classification.
-Also the algorithm doesn’t account for what a sentence is not, it only counts what it is like.
-You might say this approach doesn’t consider what makes a sentence not a given class.
-
-Many text and spoken word frameworks use similar algorithms such as this to classify intent.
-Most of what’s taking place is word counting against training datasets.
-It’s "naive" but surprisingly effective.
-```
-
-I will be using an algorithm-based classification strategy for this project.
-
-####\
-
-###Neural Networks
-These methods are beyond the scope of this project. Anyone interested in knowing more can search online for information about:
-
-*  Shallow Neural Networks
-*  Deep Neural Networks
-	+  Convolutional Neural Network (CNN)
-	+  Long Short Term Modelr (LSTM)
-	+  Gated Recurrent Unit (GRU)
-	+  Bidirectional RNN
-	+  Recurrent Convolutional Neural Network (RCNN)
-	+  Other Variants of Deep Neural Networks
-
-####\
-
-###Designing an Evaluation Workflow
-The workflow for comparing classification algorithms starts with identifying potentially usable classifier methods.
-Several approaches are described later.
-For each potential classifier method, the next steps are:
-
-*  Establish randomized training and test datasets
-*  Train the classifier algorithm, then evaluate that model on a test dataset
-*  Calculate an accuracy score using Gini index or similar method
-*  Select best performing classifiers and either refine, or operationalize for new datasets
-
-####\
-
-###Picking the Algorithm(s)
-Algorithms approach text classification several ways.
-
-The **bag of keywords approach** requires compiling a list of "key terms" that qualifies the type of content in question to a certain topic.
-If one or more qualifying keywords is present in a document, the document is assigned to the topic.
-One of the many problems with this approach is that identifying and organizing a list of terms (preferred, alternate, etc.) is labor intensive (much like simple pattern matching), and the natural ambiguity of language (one keyword can have different meanings) causes lots of false positives.
-This makes the bag of keywords approach not only not scalable, but also inaccurate.
-
-**Rules-based approaches** use linguistic rules that capture all of the elements and attributes of a document to assign it to a category.
-Rules can be written manually or generated by automatic analysis and then validated manually (reducing manual effort up to 90%).
-Rules can be understood and improved (unlike the black box system used with statistics-based algorithms).
-A rules-based approach is flexible, powerful (much more than a bag of keywords) and easy to express.
-This approach works best when combined with semantic technology that can uncover deeper connections within text (meaning, relevancy, relationship between concepts, etc.)
-
-**Statistical approaches** use a “training set” of documents that covers the same topic to identify key features (usually words).
-The algorithm (Bayesian, LSA, etc.) calculates term frequencies, then builds implicit rules in order to classify other content.
-This approach has no understanding of meaning.
-In addition, systems using these types of text classification algorithms are essentially a “black box”; no one can explain why specific terms are selected by the algorithm or how they are being weighted.
-In the event the classification is incorrect, there is no accurate way to modify a rule for better results.
-
-Algorithms also can be separated into un-supervised and supervised methods. 
-**Unsupervised learning** uses test data that has not been labeled, classified or categorized.
-Unsupervised learning identifies commonalities in the training data, then classifies new texts based on the presence or absence of such commonalities.
-
-Unsupervised learning algorithms appearing frequently in literature are:
-
-*  Clustering
-	+  Hierarchical clustering
-	+  k-means
-	+  Mixture models
-*  Principal component analysis (PCA)
-
-**Supervised learning** uses labeled training data to defined important text features.
-In supervised learning, each example is a pair consisting of an input object (typically a vector) and a desired output value.
-A supervised learning algorithm analyzes the training data and produces an inferred function, which can be used for mapping new examples.
-.
-
-Supervised text classification algorithms appearing regularly in the literature are:
-
-*  Naive Bayes
-	+  Binary independent NB
-	+  Multinomial NB
-*  Linear Classifier
-*  Logistic Regression
-*  Support vector machines
-	+  Basic SVM
-	+  SVM multiple classes
-	+  SVM multiple kernels
-*  Decision trees
-*  Linear Discriminant Analysis
-*  Nearest neighbor
-	+  k-Nearest Neighbor
-	+  Weighted k-Nearest Neighbor
-*  Similarity Learning
-*  Relevance feedback
-	+  Rocchio
-	+  Rocchio in a query zone
-*  Ensemble methods
-	+  Bagging based Models
-	+  Stacking based Models
-	+  Boosting based Models
-	+  Random Forest Models
-
-For reference I assembled short excerpts from online resources or published articles describing most of these methods.
-
-####\
-
-###Naive Bayes
-A classification technique based on Bayes’ Theorem with an assumption of independence among predictors.
-A Naive Bayes classifier assumes that the presence of a particular feature in a class is unrelated to the presence of any other feature here.
-NB can be run on:
-
-*  Word count or n-gram count vectors
-*  Word or n-gram level TF-IDF vectors
-*  Character Level TF-IDF vectors
-	+  Would be useful for evaluating use of "?" and "="
-
-Data are organized into 2 dictionaries: **corpus words** (each stemmed word and the # of occurances) and **class words** (each class and the list of stemmed words within it).
-The algorithm then uses these data structures to classify new text.
-
-####\
-
-###Linear Classifier
-The linear regression algorithm is one of the fundamental supervised machine-learning algorithms due to its relative simplicity and well-known properties.
-
-The goal of statistical classification is to use an object's characteristics to identify which class (or group) it belongs to.
-A linear classifier achieves this by making a classification decision based on the value of a linear combination of the characteristics.
-An object's characteristics are also known as feature values and are typically presented to the machine in a vector called a feature vector.
-Such classifiers work well for practical problems such as document classification, and more generally for problems with many variables (features), reaching accuracy levels comparable to non-linear classifiers while taking less time to train and use.
-
-####\
-
-###Logistic Regression
-A logistic regression measures the relationship between the categorical dependent variable and one or more independent variables by estimating probabilities using a logistic/sigmoid function.
-LR can be run on:
-
-*  Word count or n-gram count vectors
-*  Word or n-gram level TF-IDF vectors
-*  Character Level TF-IDF vectors
-
-Resources for Logistic Regression 
-https://www.analyticsvidhya.com/blog/2015/10/basics-logistic-regression/
-
-R for logistic regression
-https://www.saedsayad.com/logistic_regression.htm
-
-####\
-
-###Support Vector Machine (SVM) Model
-This is a supervised machine learning algorithm which can be used for both classification or regression challenges.
-The model extracts a best possible hyper-plane / line that segregates the two classes.
-The only example I found of SVM was on Ngram Level TF-IDF Vectors.
-Support vector machines can be:
-
-*  Basic SVM
-*  SVM multiple classes
-*  SVM multiple kernels
-
-####\
-
-###Decision Trees
-Flowchart-like structures in which each internal node represents a "test" on an attribute (e.g.
-whether a coin flip comes up heads or tails), each branch represents the outcome of the test, and each leaf node represents a class label (decision taken after computing all attributes).
-The paths from root to leaf represent classification rules.
-
-In decision analysis, a decision tree and the closely related influence diagram are used as a visual and analytical decision support tool, where the expected values (or expected utility) of competing alternatives are calculated.
-
-A decision tree consists of three types of nodes:
-
-*  Decision nodes – typically represented by squares
-*  Chance nodes – typically represented by circles
-*  End nodes – typically represented by triangles
-
-Decision tree learning is the construction of a decision tree from class-labeled training tuples.
-A decision tree is a flow-chart-like structure, where each internal (non-leaf) node denotes a test on an attribute, each branch represents the outcome of a test, and each leaf (or terminal) node holds a class label.
-The topmost node in a tree is the root node.
-
-The goal is to create a model that predicts the value of a target variable based on several input variables.
-An example is shown in the diagram at right.
-Each interior node corresponds to one of the input variables; there are edges to children for each of the possible values of that input variable.
-Each leaf represents a value of the target variable given the values of the input variables represented by the path from the root to the leaf.
-
-A decision tree is a simple representation for classifying examples.
-For this section, assume that all of the input features have finite discrete domains, and there is a single target feature called the "classification".
-Each element of the domain of the classification is called a class.
-A decision tree or a classification tree is a tree in which each internal (non-leaf) node is labeled with an input feature.
-The arcs coming from a node labeled with an input feature are labeled with each of the possible values of the target or output feature or the arc leads to a subordinate decision node on a different input feature.
-Each leaf of the tree is labeled with a class or a probability distribution over the classes.
-
-A tree can be "learned" by splitting the source set into subsets based on an attribute value test.
-This process is repeated on each derived subset in a recursive manner called recursive partitioning.
-The recursion is completed when the subset at a node has all the same value of the target variable, or when splitting no longer adds value to the predictions.
-This process of top-down induction of decision trees (TDIDT) is an example of a greedy algorithm, and it is by far the most common strategy for learning decision trees from data.
-
-In data mining, decision trees can be described also as the combination of mathematical and computational techniques to aid the description, categorization and generalization of a given set of data.
-
-####\
-
-###Linear Discriminant Analysis
-LDA works when the measurements made on independent variables for each observation are continuous quantities.
-When dealing with categorical independent variables, the equivalent technique is discriminant correspondence analysis.
-The statistical system R includes the packages: MASS, ade4, ca, vegan, ExPosition, andFactoMineR which perform correspondence analysis and multiple correspondence analysis.
-
-As the name indicates, discriminant correspondence analysis (DCA) is an extension of discriminant analysis (DA) and correspondence analysis (CA).
-Like discriminant analysis, the goal of DCA is to categorize observations in pre-defined groups, and like  correspondence analysis, it is used with nominal variables.The main idea behind DCA is to represent each group by thesum of its observations and to perform a simple CA on the groupsby variables matrix.
-The original observations are then projectedas  supplementary  elements and each observation is assigned to the closest group.
-The comparison between the a priori and the a posteriori classifications can be used to assess the quality of thendiscrimination.
-A similar procedure can be used to assign new observations to categories.
-The stability of the analysis can be evaluated using cross-validation techniques such as jackknifing or boot-strapping.
-
-For the R model: https://pbil.univ-lyon1.fr/ade4/ade4-html/discrimin.coa.html
-
-####\
-
-###k-Nearest Neighbors
-k-NN is a non-parametric method used for classification and regression.
-In both cases, the input consists of the k closest training examples in the feature space.
-The output depends on whether k-NN is used for classification or regression.
-In k-NN classification, the output is a class membership.
-An object is classified by a majority vote of its neighbors, with the object being assigned to the class most common among its k nearest neighbors (k is a positive integer, typically small).
-If k = 1, then the object is simply assigned to the class of that single nearest neighbor.
-
-####\
-
-###Similarity Learning
-Similarity learning is an area of supervised machine learning in artificial intelligence.
-It is closely related to regression and classification, but the goal is to learn from examples a similarity function that measures how similar or related two objects are.
-It has applications in ranking, in recommendation systems, visual identity tracking, face verification, and speaker verification.
-
-####\
-
-###Relevance Feedback/Rocchio Algorithm
-Relevance feedback is a feature of some information retrieval systems.
-The idea behind relevance feedback is to take the results that are initially returned from a given query, to gather user feedback, and to use information about whether or not those results are relevant to perform a new query.
-We can usefully distinguish between three types of feedback: explicit feedback, implicit feedback, and blind or "pseudo" feedback.
-
-The Rocchio algorithm is based on a method of relevance feedback.
-Like many other retrieval systems, the Rocchio feedback approach was developed using the Vector Space Model.
-The algorithm is based on the assumption that most users have a general conception of which documents should be denoted as relevant or non-relevant.
-Therefore, the user's search query is revised to include an arbitrary percentage of relevant and non-relevant documents as a means of increasing the search engine's recall, and possibly the precision as well.
-
-For more:
-https://nlp.stanford.edu/IR-book/html/htmledition/rocchio-classification-1.html
-
-####\
-
-###Ensemble Methods
-The excerpt below nice describes how ensemble methods work. It is reprinted from: *Ensemble Methods Explained in Simple English*, by Tavish Srivastava, posted August 2, 2015.
-
-Ensemble modeling is a powerful way to improve the performance of your model. It usually pays off to apply ensemble learning over and above various models you might be building.
-
-Example: I want to invest in a company XYZ.
-I am not sure about its performance though.
-So, I look for advice on whether the stock price will increase more than 6% per annum or not?
-I decide to approach various experts having diverse domain experience:
-
->1.
-Employee of Company XYZ: This person knows the internal functionality of the company and have the insider information about the functionality of the firm.
-But he lacks a broader perspective on how are competitors innovating, how is the technology evolving and what will be the impact of this evolution on Company XYZ’s product.
-In the past, he has been right 70% times.
-
->1.
-Financial Advisor of Company XYZ: This person has a broader perspective on how companies strategy will fair of in this competitive environment.
-However, he lacks a view on how the company’s internal policies are fairing off.
-In the past, he has been right 75% times.
-
->1.
-Stock Market Trader: This person has observed the company’s stock price over past 3 years.
-He knows the seasonality trends and how the overall market is performing.
-He also has developed a strong intuition on how stocks might vary over time.
-In the past, he has been right 70% times.
-
->1.
-Employee of a competitor: This person knows the internal functionality of the competitor firms and is aware of certain changes which are yet to be brought.
-He lacks a sight of company in focus and the external factors which can relate the growth of competitor with the company of subject.
-In the past, he has been right 60% of times.
-
->1.
-Market Research team in same segment: This team analyzes the customer preference of company XYZ’s product over others and how is this changing with time.
-Because he deals with customer side, he is unaware of the changes company XYZ will bring because of alignment to its own goals.
-In the past, they have been right 75% of times.
-
->1.
-Social Media Expert: This person can help us understand how has company XYZ has positioned its products in the market.
-And how are the sentiment of customers changing over time towards company.
-He is unaware of any kind of details beyond digital marketing.
-In the past, he has been right 65% of times.
-
-Given the broad spectrum of access we have, we can probably combine all the information and make an informed decision.
-In a scenario when all the 6 experts/teams verify that it’s a good decision(assuming all the predictions are independent of each other), we will get a combined accuracy rate of:
-
-```
-		  1 - 30%*25%*30%*40%*25%*35%
-
-		= 1 - 0.07875 = 99.92125%
-```
-
-Assumption: The assumption used here that all the predictions are completely independent is slightly extreme as they are expected to be correlated.
-However, we see how we can be so sure by combining various predictions together.
-
-Let us now change the scenario slightly. This time we have 6 experts, all of them are employee of company XYZ working in the same division. Everyone has a propensity of 70% to advocate correctly.
-
->*What if we combine all this advice together, can we still raise up our confidence to >99%?*
-
-Obviously not, as all the predictions are based on very similar set of information.
-They are certain to be influenced by similar set of information and the only variation in their advice would be due to their personal opinions & collected facts about the firm.
-
-Ensemble learning is the art of combining individual models together to improve the stability and predictive power of the model.
-
-####\
-
-####Error Sources in Ensemble Learning (Variance vs. Bias)
-The error emerging from any model can be broken down into three components mathematically.
-
-**Error of a model** Why is this important in the current context? To understand what really goes behind an ensemble model, we need to first understand what causes error in the model.
-
-**Bias error** quantifies how much on an average the predicted values differ from the actual value. A high bias error means we have a under-performing model which keeps on missing important trends.
-
-**Variance** quantifies how much the predictions made on same observations differ from each other.
-A high variance model will over-fit on your training population and perform badly on any observation beyond training.
-
-Normally, as you increase the complexity of your model, you will see a reduction in error due to lower bias in the model.
-However, this only happens till a particular point.
-As you continue to make your model more complex, you end up **over-fitting your model** and hence your model will start suffering from high variance.
-
-A champion model should maintain a balance between these two types of errors.
-This is known as the **trade-off management of bias-variance errors.** Ensemble learning is one way to execute this trade off analysis.
-
-####\
-
-####Commonly used Ensemble learning techniques
-**Bagging** tries to implement similar learners on small sample populations and then takes a mean of all the predictions.
-In generalized bagging, you can use different learners on different population.
-As you can expect this helps us to reduce the variance error.
-
-**Boosting** is an iterative technique which adjust the weight of an observation based on the last classification.
-If an observation was classified incorrectly, it tries to increase the weight of this observation and vice versa.
-Boosting in general decreases the bias error and builds strong predictive models.
-However, they may sometimes over fit on the training data.
-
-**Stacking** is a very interesting way of combining models.
-Here we use a learner to combine output from different learners.
-This can lead to decrease in either bias or variance error depending on the combining learner we use.
-
-####\
-
-####Random Forest Models
-Random Forest models are a type of ensemble models, particularly bagging models. They are part of the tree based model family. RF can be run on:
-
-*  Word count or n-gram count vectors
-*  Word level TF-IDF vectors
-
-Resources for Random Forest: 
-https://www.analyticsvidhya.com/blog/2014/06/introduction-random-forest-simplified/
-
-####\
-
-####Extreme Gradient Boosting Model
-Boosting models are another type of ensemble models part of tree based models.
-Boosting is a machine learning ensemble meta-algorithm for primarily reducing bias, and also variance in supervised learning, and a family of machine learning algorithms that convert weak learners to strong ones.
-A weak learner is defined to be a classifier that is only slightly correlated with the true classification (it can label examples better than random guessing).
-XGB can be run on:
-
-*  Word count or n-gram count vectors
-*  Word level TF-IDF vectors
-*  Character Level TF-IDF vectors
-
-####\
-
-###Hierarchical Clustering
-In data mining and statistics, hierarchical clustering (also called hierarchical cluster analysis or HCA) is a method of cluster analysis which seeks to build a hierarchy of clusters.
-
-####\
-
-###k-Means Clustering
-k-means clustering is a method of vector quantization, originally from signal processing, that is popular for cluster analysis in data mining.
-k-means clustering aims to partition n observations into k clusters in which each observation belongs to the cluster with the nearest mean, serving as a prototype of the cluster.
-This results in a partitioning of the data space into Voronoi cells.
-
-The problem is computationally difficult (NP-hard); however, efficient heuristic algorithms converge quickly to a local optimum.
-These are usually similar to the expectation-maximization algorithm for mixtures of Gaussian distributions via an iterative refinement approach employed by both k-means and Gaussian mixture modeling.
-They both use cluster centers to model the data; however, k-means clustering tends to find clusters of comparable spatial extent, while the expectation-maximization mechanism allows clusters to have different shapes.
-
-The algorithm has a loose relationship to the k-nearest neighbor classifier, a popular machine learning technique for classification that is often confused with k-means due to the name.
-Applying the 1-nearest neighbor classifier the cluster centers obtained by k-means classifies new data into the existing clusters.
-This is known as nearest centroid classifier or Rocchio algorithm.
-
-####\
-
-###Mixture Model
-In statistics, a mixture model is a probabilistic model for representing the presence of subpopulations within an overall population, without requiring that an observed data set should identify the sub-population to which an individual observation belongs.
-Formally a mixture model corresponds to the mixture distribution that represents the probability distribution of observations in the overall population.
-However, while problems associated with "mixture distributions" relate to deriving the properties of the overall population from those of the sub-populations, "mixture models" are used to make statistical inferences about the properties of the sub-populations given only observations on the pooled population, without sub-population identity information.
-
-Topics in a document: Assume that a document is composed of N different words from a total vocabulary of size V, where each word corresponds to one of K possible topics.
-The distribution of such words could be modelled as a mixture of K different V-dimensional categorical distributions.
-A model of this sort is commonly termed a topic model.
-Note that expectation maximization applied to such a model will typically fail to produce realistic results, due (among other things) to the excessive number of parameters.
-Some sorts of additional assumptions are typically necessary to get good results.
-
-A prior distribution is placed over the parameters describing the topic distributions, using a Dirichlet distribution with a concentration parameter that is set significantly below 1, so as to encourage sparse distributions (where only a small number of words have significantly non-zero probabilities).
-
-Some sort of additional constraint is placed over the topic identities of words, to take advantage of natural clustering.
-
-For example, a Markov chain could be placed on the topic identities (i.e., the latent variables specifying the mixture component of each observation), corresponding to the fact that nearby words belong to similar topics.
-(This results in a hidden Markov model, specifically one where a prior distribution is placed over state transitions that favors transitions that stay in the same state.)
-
-Another possibility is the latent Dirichlet allocation model, which divides up the words into D different documents and assumes that in each document only a small number of topics occur with any frequency.
-
-####\
-
-###Principal Component Analysis 
-PCA is a statistical procedure that uses an orthogonal transformation to convert a set of observations of possibly correlated variables (entities each of which takes on various numerical values) into a set of values of linearly uncorrelated variables called principal components.
-This transformation is defined in such a way that the first principal component has the largest possible variance (that is, accounts for as much of the variability in the data as possible), and each succeeding component in turn has the highest variance possible under the constraint that it is orthogonal to the preceding components.
-The resulting vectors (each being a linear combination of the variables and containing n observations) are an uncorrelated orthogonal basis set.
-PCA is sensitive to the relative scaling of the original variables.
-
-PCA is mostly used as a tool in exploratory data analysis and for making predictive models.
-It is often used to visualize genetic distance and relatedness between populations.
-PCA can be done by eigenvalue decomposition of a data covariance (or correlation) matrix or singular value decomposition of a data matrix, usually after a normalization step of the initial data.
-The normalization of each attribute consists of mean centering – subtracting each data value from its variable's measured mean so that its empirical mean (average) is zero – and, possibly, normalizing each variable's variance to make it equal to 1.
-The results of a PCA are usually discussed in terms of component scores, sometimes called factor scores (the transformed variable values corresponding to a particular data point), and loadings (the weight by which each standardized original variable should be multiplied to get the component score).
-If component scores are standardized to unit variance loadings must contain the data variance in them (and that is the magnitude of eigenvalues).
-If component scores are not standardized (therefore they contain the data variance) then loadings must be unit-scaled, ("normalized") and these weights are called eigenvectors; they are the cosines of orthogonal rotation of variables into principal components or back.
-
-PCA is the simplest of the true eigenvector-based multivariate analyses.
-Often, its operation can be thought of as revealing the internal structure of the data in a way that best explains the variance in the data.
-If a multivariate dataset is visualised as a set of coordinates in a high-dimensional data space (1 axis per variable), PCA can supply the user with a lower-dimensional picture, a projection of this object when viewed from its most informative viewpoint.
-This is done by using only the first few principal components so that the dimensionality of the transformed data is reduced.
-
-PCA is closely related to factor analysis.
-Factor analysis typically incorporates more domain specific assumptions about the underlying structure and solves eigenvectors of a slightly different matrix.
-
-PCA is also related to canonical correlation analysis (CCA).
-CCA defines coordinate systems that optimally describe the cross-covariance between two datasets while PCA defines a new orthogonal coordinate system that optimally describes variance in a single dataset.
-
-####\
-
-##Strategies for Improving Text Classification Models
-Accuracy can be improved using one or more of these strategies.
-
-1. Text Cleaning. Removing stopwords, punctuations marks, suffix variations etc. can  reduce the noise present in text data.
-
-2.
-Hstacking arrays of text / NLP features with text feature vectors in sequence horizontally (column wise).
-After extracting features initially, try combining different feature vectors to improve the accuracy of the classifier.
-
-3. Tuning the Classifier. Adjusting the parameters of the classifier is an important step towards a best fit model.
-
-4. Testing Different Ensemble Models. Stacking different models and blending their outputs can help to further improve the results.
-
-5.
-Weighting words: we can significantly improve our algorithm by accounting for the commonality of each word.
-The word “is” should carry a lower weigh than the word “sandwich” in most cases, because it is more common.
-
-####\
-
-##Evaluation
-How to evaluate the different components of the overall text classification problem.
-
-####\
-
-###Gini Index
-One of the most common methods for quantifying the discrimination level of a feature is the use of a measure known as the gini-index.
-
-####\
-
-###Cohen's Kappa Score
-Cohen's kappa is a measure of interrater reliability used in social sciences. This excerpt from an R blog explains it.
-
-**Using kappa score for comparisons of groups and methods**
-https://www.personality-project.org/r/html/kappa.html
-https://www.r-bloggers.com/k-is-for-cohens-kappa/
-
-Last April, during the A to Z of Statistics, I blogged about Cohen’s kappa, a measure of interrater reliability.
-Cohen’s kappa is a way to assess whether two raters or judges are rating something the same way.
-And thanks to an R package called **irr**, it’s very easy to compute.
-But first, let’s talk about why you would use Cohen’s kappa and why it’s superior to a more simple measure of interrater reliability, interrater agreement.
-
-####\
-
-###Neural Networks
-A neural network is a mathematical model that is designed to behave similar to biological neurons and nervous system.
-These models are used to recognize complex patterns and relationships that exists within a labelled data.
-
-Artificial neural networks, invented in the 1940’s, are a way of calculating an output from an input (a classification) using weighted connections (“synapses”) that are calculated from repeated iterations through training data.
-Each pass through the training data alters the weights such that the neural network produces the output with greater “accuracy” (lower error rate).
-
-There’s not much new about these structures, except today’s software is using much faster processors and can work with a lot more memory.
-The combination of working memory and speed is crucial when you’re doing hundreds of thousands of matrix multiplications (the neural network’s essential math operation).
-
-As in the prior method, each class is given with some number of example sentences.
-Once again each sentence is broken down by word (stemmed) and each word becomes an input for the neural network.
-The synaptic weights are then calculated by iterating through the training data thousands of times, each time adjusting the weights slightly to greater accuracy.
-By recalculating back across multiple layers (“back-propagation”) the weights of all synapses are calibrated while the results are compared to the training data output.
-These weights are like a ‘strength’ measure, in a neuron the synaptic weight is what causes something to be more memorable than not.
-You remember a thing more because you’ve seen it more times: each time the ‘weight’ increases slightly.
-
-At some point the adjustment reaches a point of diminishing returns, this is called “over-fitting” and going beyond this is counter-productive.
-
-The trained neural network is less code than an comparable algorithm but it requires a potentially large matrix of “weights”.
-In a relatively small sample, where the training sentences have 150 unique words and 30 classes this would be a matrix of 150x30.
-Imagine multiplying a matrix of this size 100,000 times to establish a sufficiently low error rate.
-This is where processing speed comes in.
-
-If the neural network sounds magnificently sophisticated, relax, it boils down to matrix multiplication and a formula for reducing values between -1 and 1 or some other minimal range.
-A middle-school math student could learn this in a few hours.
-The hard work is achieving clean training data.
-
-Just as there are variations in pattern matching code and in algorithms, there are variations in neural networks, some more complex than others.
-The basic machinery is the same.
-The essential work is that of classification.
-
-Networks are looking for patterns in collections of terms, each term is reduced to a token.
-In this machine words have no meaning except for their patterned existence within training data.
-The label “artificial intelligence” applied to such machinery is mostly BS.
-
-####\
-
-###Types of Neural Networks
-A shallow neural network contains mainly three types of layers – input layer, hidden layer, and output layer.
-
-Deep Neural Networks are more complex neural networks in which the hidden layers performs much more complex operations than simple sigmoid or relu activations.
-Different types of deep learning models can be applied in text classification problems.
-
-####\
-
-####Convolutional Neural Network
-In Convolutional neural networks, convolutions over the input layer are used to compute the output.
-This results in local connections, where each region of the input is connected to a neuron in the output.
-Each layer applies different filters and combines their results.
-
-####\
-
-####Recurrent Neural Network: LSTM
-Unlike Feed-forward neural networks in which activation outputs are propagated only in one direction, the activation outputs from neurons propagate in both directions (from inputs to outputs and from outputs to inputs) in Recurrent Neural Networks.
-This creates loops in the neural network architecture which acts as a ‘memory state’ of the neurons.
-This state allows the neurons an ability to remember what have been learned so far.
-
-The memory state in RNNs gives an advantage over traditional neural networks but a problem called Vanishing Gradient is associated with them.
-In this problem, while learning with a large number of layers, it becomes really hard for the network to learn and tune the parameters of the earlier layers.
-To address this problem, A new type of RNNs called LSTMs (Long Short Term Memory) Models have been developed.
-
-####\
-
-####Recurrent Neural Network: GRU
-Gated Recurrent Units are another form of recurrent neural networks.
-
-####\
-
-####Bidirectional RNN
-RNN layers can be wrapped in Bidirectional layers as well.
-
-####\
-
-####Recurrent Convolutional Neural Network
-Once the essential architectures have been tried out, one can try different variants of these layers such as recurrent convolutional neural network. Another variants can be:
-
-*  Hierarchical Attention Networks
-*  Sequence to Sequence Models with Attention
-*  Bidirectional Recurrent Convolutional Neural Networks
-*  CNNs and RNNs with more number of layers
-
-####\

### 2019-02-25 Choosing_classifier.Rmd
updates

--- 2019-02-23 Text Classifier Models.Rmd
@@
 ---
-title: "Text Classification Primer"
-author: "Dan Johnson"
-date: "1/9/2019"
-output: html_document
+title: "Choosing a Text Classifier Method"
+date: "2/21/2019"
+bibliography: library.bib
+biblio-style: apalike
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, child="_setup.Rmd"}
 ```
 
-For this text classification project I am using an initial codebook and corpus of hand-coded TA comments to test various approaches to semi- and fully automated classification methods and, if appropriate, revise the underlying classification scheme.
-
-This page provides background on general mechanics of text classification and various classifier methods.
-Other pages describe the dataset structure [LINK] (TA comments corpus), and [LINK] development of the initial codebook.
-
-#Workflow
-Text classification problems can be broken down into four overlapping sub-problems. Each piece may be more or less important to the overall problem, but all need to be considered:
-
-1.  Feature selection and engineering
-2.  Topic categorization and modeling
-3.  Selecting an appropriate text classifier method
-4.  Evaluating the outcomes
-
-##Feature Selection and Engineering
-This is the process of selecting, tabulating, calculating, or creating text-based features for subsequent evaluation. This includes:
-
-*  Identifying existing, directly observable text features within the dataset
-*  Engineering additional features using established NLP methods
-*	 Reverse engineering relevant features based on an empirically defined feature set.
-
-The features listed below are used regularly for natural language processing.
-Not all will be appropriate for all situations though; selecting, evaluating, comparing, and validating the appropriate features for the classification task is an iterative process, not a one-time decision.
-
-*  Count-based Vectors
-	+  Word Count – total number of words in the documents
-	+  Character Count – total number of characters in the documents
-	+  Punctuation Count – total number of punctuation marks in the documents
-	+  Upper Case Count – total number of upper case words in the documents
-	+  Title Word Count – total number of consecutive upper case (title) words in the documents
-*  Distribution and Frequency Calculations
-	+  Average word density of sentences 
-	+  Average length of the words used in the documents
-	+  TF-IDF
-	+  Reading level estimates (described below)
-*  Part of Speech Tagging and Frequency Distribution of Tags 
-	+ Noun Count
-	+ Verb Count
-	+ Adjective Count
-	+ Adverb Count
-	+ Pronoun Count
-*  Word Embedding (described below)
-
-Additional features are available for longer texts such as multi-sentence paragraphs.
-
-*  Word choices
-*  Lexical diversity (described below)
-*  Stance patterns - hedging, booster, generalization words or phrasing 
-*  Concision - comparisons of noun phrase density versus comment clause density
-*  Cohesion - use of words that are additive (also), countering (however), specify or emphasize (even, both, especially), connecting (former, latter)
-*  Civility and uncertainty - language that acknowledges limits of statement, critiques ideas not people, and makes fewer generalizations
-
-###Lexical Diversity
-This is not the same as word choices.
-This explanation is an excerpt from "quanteda: Quantitative Analysis of Textual Data", which describes how to use ```textstat_lexdiv``` to calculate lexical diversity.
-https://rdrr.io/cran/quanteda/man/textstat_lexdiv.html
-http://quanteda.io/reference/textstat_lexdiv.html
-
-Consider a speaker, who uses the term "allow" multiple times throughout the speech, compared to an another speaker who uses terms "allow, concur, acquiesce, accede, and avow"" for the same word.
-The latter speech has more lexical diversity than the former.
-Lexical diversity is widely believed to be an important parameter to rate a document in terms of textual richness and effectiveness.
-
-Lexical diversity, in simple terms, is a measurement of the breadth and variety of vocabulary used in a document.
-The different measures of lexical diversity are TTR, MSTTR, MATTR, C, R, CTTR, U, S, K, Maas, HD-D, MTLD, and MTLD-MA.
-
-koRpus package in R also provides functions to estimate lexical diversity or complexity.
-https://subscription.packtpub.com/book/big_data_and_business_intelligence/9781783551811/2/ch02lvl1sec16/lexical-diversity
-
-###Readability metrics
-These estimate the relative grade level or age of a typical reader who could comprehend 80+% (typically) of the text.
-Most readability metrics require longer blocks of text (100+ words) to achieve a stable score.
-They use some combination of word count, average syllables per word, and sentence length to estimate reading level.
-
-The metrics differ in how appropriate they are for specific situations.
-Some were developed for general writing, others for specific school age groups, and still others for technical documents; the US Navy in particular developed a suite of metrics for evaluating their training documents.
-
-The ```readability library``` (v0.1.1) is a collection of readability tools that utilize the ```syllable``` package for fast calculation of readability scores by grouping variables.
-It calculates Flesch Kincaid, Gunning Fog Index, Coleman Liau, SMOG, Automated Readability Index and an average of the 5 readability scores.
-These all are general scales; other scores may not be available in R packages.
-
-###Word Embedding
-**Word embedding** is a set of language modeling and feature learning techniques in natural language processing (NLP) where words or phrases from the vocabulary are mapped to vectors of real numbers.
-Conceptually it involves a mathematical embedding from a space with one dimension per word to a continuous vector space with a much lower dimension.
-
-Methods to generate the mapping include neural networks, dimensionality reduction, probabilistic models, knowledge base methods, and explicit representation in terms of the context in which words appear.
-The technique of representing words as vectors has roots in the 1960s with the development of the vector space model for information retrieval.
-
-Word embeddings come in two different styles, one in which words are expressed as vectors of co-occurring words, and another in which words are expressed as vectors of linguistic contexts in which the words occur.
-Most new word embedding techniques rely on a neural network architecture instead of more traditional n-gram models and unsupervised learning.
-
-One of the main limitations of word embeddings (word vector space models in general) is that possible meanings of a word are conflated into a single representation (a single vector in the semantic space).
-Sense embeddings are a solution to this problem: individual meanings of words are represented as distinct vectors in the space.
-
-Tools like **Principal Component Analysis (PCA)** and **T-Distributed Stochastic Neighbour Embedding (t-SNE)** are used to reduce the dimensionality of word vector spaces and visualize word embeddings and clusters.
-
-Software for training and using word embeddings includes Tomas Mikolov's Word2vec, Stanford University's GloVe, fastText, Gensim, AllenNLP's Elmo, Indra and Deeplearning4j.
-Three of these tools are described further below.
-
-####Word2vec
-Word2vec is a group of shallow, two-layer neural network models that produce word embeddings.
-The toolkit can train vector space models to reconstruct linguistic contexts of words.
-Word2vec takes as its input a large corpus of text and produces a vector space, typically of several hundred dimensions, with each unique word in the corpus being assigned a corresponding vector in the space.
-Word vectors are positioned in the vector space so words that share common contexts in the corpus are located in close proximity to one another in the space.
-
-The use of different model parameters and different corpus sizes can greatly affect the quality of a word2vec model.
-Accuracy can be improved in a number of ways, including the choice of model architecture (CBOW or Skip-Gram), increasing the training data set, increasing the number of vector dimensions, and increasing the window size of words considered by the algorithm.
-Each of these improvements comes with the cost of increased computational complexity and therefore increased model generation time.
-
-In models using large corpora and a high number of dimensions, the skip-gram model yields the highest overall accuracy, and consistently produces the highest accuracy on semantic relationships, as well as yielding the highest syntactic accuracy in most cases.
-However, the CBOW is less computationally expensive and yields similar accuracy results.
-
-Accuracy increases overall as the number of words used increases, and as the number of dimensions increases.
-Altszyler et al.
-(2017) studied Word2vec performance in two semantic tests for different corpus size.
-They found that Word2vec has a steep learning curve, outperforming another word-embedding technique (LSA) when it is trained with medium to large corpus size (more than 10 million words).
-However, with a small training corpus LSA showed better performance.
-Additionally they show that the best parameter setting depends on the task and the training corpus.
-Nevertheless, for skip-gram models trained in medium size corpora, with 50 dimensions, a window size of 15 and 10 negative samples seems to be a good parameter setting.
-
-####fastText
-fastText is a Python-based library created by Facebook's AI Research (FAIR) lab for learning of word embeddings and text classification.
-The model allows to create an unsupervised learning or supervised learning algorithm for obtaining vector representations for words.
-fastText uses a neural network for word embedding.
-
-https://fasttext.cc/
-
-####GloVe
-GloVe, short for Global Vectors, is a model for distributed word representation.
-The model is an unsupervised learning algorithm for obtaining vector representations for words.
-Training is performed on aggregated global word-word co-occurrence statistics from a corpus, and the resulting representations showcase interesting linear substructures of the word vector space.
-It is developed as an open-source project at Stanford.
-Pre-trained word vectors data are made available under the Public Domain Dedication and License v1.0.
-
-GloVe uses Euclidean distance (or cosine similarity) between two word vector for measuring the linguistic or semantic similarity of the corresponding words.
-Sometimes, the nearest neighbors according to this metric reveal rare but relevant words that lie outside an average human's vocabulary.
-For example, here are the closest words to the target word **frog**:
-
-*  frog
-*  frogs
-*  toad
-*  litoria
-*  leptodactylidae
-*  rana
-*  lizard
-*  eleutherodactylus
-
-The similarity metrics used for nearest neighbor evaluations produce a single scalar that quantifies the relatedness of two words.
-This simplicity can be problematic since two given words almost always exhibit more intricate relationships than can be captured by a single number.
-For example, man may be regarded as similar to woman in that both words describe human beings; on the other hand, the two words are often considered opposites since they highlight a primary axis along which humans differ from one another.
-
-In order to capture in a quantitative way the nuance necessary to distinguish man from woman, it is necessary for a model to associate more than a single number to the word pair.
-A natural and simple candidate for an enlarged set of discriminative numbers is the vector difference between the two word vectors.
-GloVe is designed in order that such vector differences capture as much as possible the meaning specified by the juxtaposition of two words.
-
-https://nlp.stanford.edu/projects/glove/
-
-##Sub-Problem Two: Topic Categorization and Modeling Methods
-Establishing meaningful text classification categories depends on the goals of the project or the question being asked.
-Are relevant categories known already, or do they need to be established?
-What are the main topics
-
-**Curated Categories** (my provisional term until I find the appropriate equivalent in current use) are established by subject matter experts. The established categories can be constructed from:
-
-*  A limited keywords vocabulary (National Library of Medicine MeSH categories for example)
-*  Disciplinary conventions or shared best practices (the Dublin Core is an example)
-*  Prior empirical studies
-*  The dataset itself, using *de novo* iterative coding. This process is described further here [LINK] and here [LINK].
-
-**Topic Modeling** is a group of techniques for identifying groups of words (the "topic")  that best summarize the information in a collection of documents.
-Topic modeling is an iterative/ generative process that does not require any prior knowledge of the documents or relationships.
-
-###Topic Modeling Methods to Identify Features
-####Latent Dirichlet Allocation
-(LDA) is a common method for generating topic model features.
-LDA is an iterative model which starts from a fixed number of topics.
-Each topic is represented as a distribution over words, and each document is then represented as a distribution over topics.
-Although the tokens themselves are meaningless, the probability distributions over words provided by the topics provide a sense of the different ideas contained in the documents.
-
-####Brown Clustering
-A hierarchical clustering algorithm that is applied to text, grouping words into clusters that are assumed to be semantically related by virtue of their having been embedded in similar contexts.
-
-In natural language processing, Brown clustering or IBM clustering is a form of hierarchical clustering of words based on the contexts in which they occur.
-The intuition behind the method is that a class-based language model (also called cluster n-gram model), i.e.
-one where probabilities of words are based on the classes (clusters) of previous words, is used to address the data sparsity problem inherent in language modeling.
-
-Brown groups items (i.e., types) into classes, using a binary merging criterion based on the log-probability of a text under a class-based language model, i.e.
-a probability model that takes the clustering into account.
-
-####Latent semantic analysis
-LSA is a technique in natural language processing of analyzing relationships between a set of documents and the terms they contain by producing a set of concepts related to the documents and terms.
-LSA assumes that words that are close in meaning will occur in similar pieces of text (the distributional hypothesis).
-A matrix containing word counts per paragraph (rows represent unique words and columns represent each paragraph) is constructed from a large piece of text and a mathematical technique called singular value decomposition (SVD) is used to reduce the number of rows while preserving the similarity structure among columns.
-Words are then compared by taking the cosine of the angle between the two vectors (or the dot product between the normalizations of the two vectors) formed by any two rows.
-Values close to 1 represent very similar words while values close to 0 represent very dissimilar words.
-
-In the context of its application to information retrieval, it is sometimes called latent semantic indexing (LSI).
-
-LSA can use a term-document matrix which describes the occurrences of terms in documents; it is a sparse matrix whose rows correspond to terms and whose columns correspond to documents.
-A typical example of the weighting of the elements of the matrix is tf-idf (term frequency–inverse document frequency): the weight of an element of the matrix is proportional to the number of times the terms appear in each document, where rare terms are upweighted to reflect their relative importance.
-
-Applications of LSA include:
-
-*  Comparing documents in low-dimensional space (data clustering, document classification).
-*  Finding relations between terms (synonymy and polysemy).
-Synonymy is the phenomenon where different words describe the same idea.
-Polysemy is the phenomenon where the same word has multiple meanings.
-*  Given a query of terms, translating it into the low-dimensional space, and find matching documents (information retrieval).
-*  Analyzing word association in text corpus.
-
-LSA overcomes two of the most problematic constraints of Boolean keyword queries: multiple words that have similar meanings (synonymy) and words that have more than one meaning (polysemy).
-
-LSA is also used to perform automated document categorization.
-Document categorization is the assignment of documents to one or more predefined categories based on their similarity to the conceptual content of the categories.
-Several experiments have demonstrated that there are a number of correlations between the way LSI and humans process and categorize text.
-
-LSA uses example documents to establish the conceptual basis for each category.
-During categorization processing, the concepts contained in the documents being categorized are compared to the concepts contained in the example items, and a category (or categories) is assigned to the documents based on the similarities between the concepts they contain and the concepts that are contained in the example documents.
-
-LSA automatically adapts to new and changing terminology, and has been shown to be very tolerant of noise (i.e., misspelled words, typographical errors, unreadable characters, etc.).[24] This is especially important for applications using text derived from Optical Character Recognition (OCR) and speech-to-text conversion.
-LSA also deals effectively with sparse, ambiguous, and contradictory data.
-
-Text does not need to be in sentence form for LSA to be effective.
-It can work with lists, free-form notes, email, Web-based content, etc.
-As long as a collection of text contains multiple terms, LSA can be used to identify patterns in the relationships between the important terms and concepts contained in the text.
-
-##Sub-Problem Three: Choosing a Text Classifier Method
+#Choosing a Text Classifier Method
 The many methods for classifying texts fall into three general categories:
@@
 It’s "naive" but surprisingly effective.
+```
 
@@
 
-###Neural Networks
-These methods are beyond the scope of this project. Anyone interested in knowing more can search online for information about:
-
-*  Shallow Neural Networks
-*  Deep Neural Networks
-	+  Convolutional Neural Network (CNN)
-	+  Long Short Term Modelr (LSTM)
-	+  Gated Recurrent Unit (GRU)
-	+  Bidirectional RNN
-	+  Recurrent Convolutional Neural Network (RCNN)
-	+  Other Variants of Deep Neural Networks
-
-###Designing an Evaluation Workflow
-The workflow for comparing classification algorithms starts with identifying potentially usable classifier methods.
-Several approaches are described later.
-For each potential classifier method, the next steps are:
-
-*  Establish randomized training and test datasets
-*  Train the classifier algorithm, then evaluate that model on a test dataset
-*  Calculate an accuracy score using Gini index or similar method
-*  Select best performing classifiers and either refine, or operationalize for new datasets
-
-###Picking the Algorithm(s)
-Algorithms approach text classification several ways.
-
-The **bag of keywords approach** requires compiling a list of "key terms" that qualifies the type of content in question to a certain topic.
-If one or more qualifying keywords is present in a document, the document is assigned to the topic.
-One of the many problems with this approach is that identifying and organizing a list of terms (preferred, alternate, etc.) is labor intensive (much like simple pattern matching), and the natural ambiguity of language (one keyword can have different meanings) causes lots of false positives.
-This makes the bag of keywords approach not only not scalable, but also inaccurate.
-
-**Rules-based approaches** use linguistic rules that capture all of the elements and attributes of a document to assign it to a category.
-Rules can be written manually or generated by automatic analysis and then validated manually (reducing manual effort up to 90%).
-Rules can be understood and improved (unlike the black box system used with statistics-based algorithms).
-A rules-based approach is flexible, powerful (much more than a bag of keywords) and easy to express.
-This approach works best when combined with semantic technology that can uncover deeper connections within text (meaning, relevancy, relationship between concepts, etc.)
-
-**Statistical approaches** use a “training set” of documents that covers the same topic to identify key features (usually words).
-The algorithm (Bayesian, LSA, etc.) calculates term frequencies, then builds implicit rules in order to classify other content.
-This approach has no understanding of meaning.
-In addition, systems using these types of text classification algorithms are essentially a “black box”; no one can explain why specific terms are selected by the algorithm or how they are being weighted.
-In the event the classification is incorrect, there is no accurate way to modify a rule for better results.
-
-Algorithms also can be separated into un-supervised and supervised methods.
-
-**Unsupervised learning** uses test data that has not been labeled, classified or categorized.
-Unsupervised learning identifies commonalities in the training data, then classifies new texts based on the presence or absence of such commonalities.
-
-Unsupervised learning algorithms appearing frequently in literature are:
-
-*  Clustering
-	+  Hierarchical clustering
-	+  k-means
-	+  Mixture models
-*  Principal component analysis (PCA)
-
-**Supervised learning** uses labeled training data to defined important text features.
-In supervised learning, each example is a pair consisting of an input object (typically a vector) and a desired output value.
-A supervised learning algorithm analyzes the training data and produces an inferred function, which can be used for mapping new examples.
-.
-
-Supervised text classification algorithms appearing regularly in the literature are:
-
-*  Naive Bayes
-	+  Binary independent NB
-	+  Multinomial NB
-*  Linear Classifier
-*  Logistic Regression
-*  Support vector machines
-	+  Basic SVM
-	+  SVM multiple classes
-	+  SVM multiple kernels
-*  Decision trees
-*  Linear Discriminant Analysis
-*  Nearest neighbor
-	+  k-Nearest Neighbor
-	+  Weighted k-Nearest Neighbor
-*  Similarity Learning
-*  Relevance feedback
-	+  Rocchio
-	+  Rocchio in a query zone
-*  Ensemble methods
-	+  Bagging based Models
-	+  Stacking based Models
-	+  Boosting based Models
-	+  Random Forest Models
-
-For reference I assembled short excerpts from online resources or published articles describing most of these methods.
-
-###Naive Bayes
-A classification technique based on Bayes’ Theorem with an assumption of independence among predictors.
-A Naive Bayes classifier assumes that the presence of a particular feature in a class is unrelated to the presence of any other feature here.
-NB can be run on:
-
-*  Word count or n-gram count vectors
-*  Word or n-gram level TF-IDF vectors
-*  Character Level TF-IDF vectors
-	+  Would be useful for evaluating use of "?" and "="
-
-Data are organized into 2 dictionaries: **corpus words** (each stemmed word and the # of occurances) and **class words** (each class and the list of stemmed words within it).
-The algorithm then uses these data structures to classify new text.
-
-###Linear Classifier
-The linear regression algorithm is one of the fundamental supervised machine-learning algorithms due to its relative simplicity and well-known properties.
-
-The goal of statistical classification is to use an object's characteristics to identify which class (or group) it belongs to.
-A linear classifier achieves this by making a classification decision based on the value of a linear combination of the characteristics.
-An object's characteristics are also known as feature values and are typically presented to the machine in a vector called a feature vector.
-Such classifiers work well for practical problems such as document classification, and more generally for problems with many variables (features), reaching accuracy levels comparable to non-linear classifiers while taking less time to train and use.
-
-###Logistic Regression
-A logistic regression measures the relationship between the categorical dependent variable and one or more independent variables by estimating probabilities using a logistic/sigmoid function.
-LR can be run on:
-
-*  Word count or n-gram count vectors
-*  Word or n-gram level TF-IDF vectors
-*  Character Level TF-IDF vectors
-
-Resources for Logistic Regression 
-https://www.analyticsvidhya.com/blog/2015/10/basics-logistic-regression/
-
-R for logistic regression
-https://www.saedsayad.com/logistic_regression.htm
-
-###Support Vector Machine (SVM) Model
-This is a supervised machine learning algorithm which can be used for both classification or regression challenges.
-The model extracts a best possible hyper-plane / line that segregates the two classes.
-The only example I found of SVM was on Ngram Level TF-IDF Vectors.
-Support vector machines can be:
-
-*  Basic SVM
-*  SVM multiple classes
-*  SVM multiple kernels
-
-###Decision Trees
-Flowchart-like structures in which each internal node represents a "test" on an attribute (e.g.
-whether a coin flip comes up heads or tails), each branch represents the outcome of the test, and each leaf node represents a class label (decision taken after computing all attributes).
-The paths from root to leaf represent classification rules.
-
-In decision analysis, a decision tree and the closely related influence diagram are used as a visual and analytical decision support tool, where the expected values (or expected utility) of competing alternatives are calculated.
-
-A decision tree consists of three types of nodes:
-
-*  Decision nodes – typically represented by squares
-*  Chance nodes – typically represented by circles
-*  End nodes – typically represented by triangles
-
-Decision tree learning is the construction of a decision tree from class-labeled training tuples.
-A decision tree is a flow-chart-like structure, where each internal (non-leaf) node denotes a test on an attribute, each branch represents the outcome of a test, and each leaf (or terminal) node holds a class label.
-The topmost node in a tree is the root node.
-
-The goal is to create a model that predicts the value of a target variable based on several input variables.
-An example is shown in the diagram at right.
-Each interior node corresponds to one of the input variables; there are edges to children for each of the possible values of that input variable.
-Each leaf represents a value of the target variable given the values of the input variables represented by the path from the root to the leaf.
-
-A decision tree is a simple representation for classifying examples.
-For this section, assume that all of the input features have finite discrete domains, and there is a single target feature called the "classification".
-Each element of the domain of the classification is called a class.
-A decision tree or a classification tree is a tree in which each internal (non-leaf) node is labeled with an input feature.
-The arcs coming from a node labeled with an input feature are labeled with each of the possible values of the target or output feature or the arc leads to a subordinate decision node on a different input feature.
-Each leaf of the tree is labeled with a class or a probability distribution over the classes.
-
-A tree can be "learned" by splitting the source set into subsets based on an attribute value test.
-This process is repeated on each derived subset in a recursive manner called recursive partitioning.
-The recursion is completed when the subset at a node has all the same value of the target variable, or when splitting no longer adds value to the predictions.
-This process of top-down induction of decision trees (TDIDT) is an example of a greedy algorithm, and it is by far the most common strategy for learning decision trees from data.
-
-In data mining, decision trees can be described also as the combination of mathematical and computational techniques to aid the description, categorization and generalization of a given set of data.
-
-###Linear Discriminant Analysis
-LDA works when the measurements made on independent variables for each observation are continuous quantities.
-When dealing with categorical independent variables, the equivalent technique is discriminant correspondence analysis.
-The statistical system R includes the packages: MASS, ade4, ca, vegan, ExPosition, andFactoMineR which perform correspondence analysis and multiple correspondence analysis.
-
-As the name indicates, discriminant correspondence analysis (DCA) is an extension of discriminant analysis (DA) and correspondence analysis (CA).
-Like discriminant analysis, the goal of DCA is to categorize observations in pre-defined groups, and like  correspondence analysis, it is used with nominal variables.The main idea behind DCA is to represent each group by thesum of its observations and to perform a simple CA on the groupsby variables matrix.
-The original observations are then projectedas  supplementary  elements and each observation is assigned to the closest group.
-The comparison between the a priori and the a posteriori classifications can be used to assess the quality of thendiscrimination.
-A similar procedure can be used to assign new observations to categories.
-The stability of the analysis can be evaluated using cross-validation techniques such as jackknifing or boot-strapping.
-
-For the R model: https://pbil.univ-lyon1.fr/ade4/ade4-html/discrimin.coa.html
-
-###k-Nearest Neighbors
-k-NN is a non-parametric method used for classification and regression.
-In both cases, the input consists of the k closest training examples in the feature space.
-The output depends on whether k-NN is used for classification or regression.
-In k-NN classification, the output is a class membership.
-An object is classified by a majority vote of its neighbors, with the object being assigned to the class most common among its k nearest neighbors (k is a positive integer, typically small).
-If k = 1, then the object is simply assigned to the class of that single nearest neighbor.
-
-###Similarity Learning
-Similarity learning is an area of supervised machine learning in artificial intelligence.
-It is closely related to regression and classification, but the goal is to learn from examples a similarity function that measures how similar or related two objects are.
-It has applications in ranking, in recommendation systems, visual identity tracking, face verification, and speaker verification.
-
-###Relevance Feedback/Rocchio Algorithm
-Relevance feedback is a feature of some information retrieval systems.
-The idea behind relevance feedback is to take the results that are initially returned from a given query, to gather user feedback, and to use information about whether or not those results are relevant to perform a new query.
-We can usefully distinguish between three types of feedback: explicit feedback, implicit feedback, and blind or "pseudo" feedback.
-
-The Rocchio algorithm is based on a method of relevance feedback.
-Like many other retrieval systems, the Rocchio feedback approach was developed using the Vector Space Model.
-The algorithm is based on the assumption that most users have a general conception of which documents should be denoted as relevant or non-relevant.
-Therefore, the user's search query is revised to include an arbitrary percentage of relevant and non-relevant documents as a means of increasing the search engine's recall, and possibly the precision as well.
-
-For more:
-https://nlp.stanford.edu/IR-book/html/htmledition/rocchio-classification-1.html
-
-###Ensemble Methods
-The excerpt below nice describes how ensemble methods work. It is reprinted from: *Ensemble Methods Explained in Simple English*, by Tavish Srivastava, posted August 2, 2015.
-
-Ensemble modeling is a powerful way to improve the performance of your model. It usually pays off to apply ensemble learning over and above various models you might be building.
-
-Example: I want to invest in a company XYZ.
-I am not sure about its performance though.
-So, I look for advice on whether the stock price will increase more than 6% per annum or not?
-I decide to approach various experts having diverse domain experience:
-
->1.
-Employee of Company XYZ: This person knows the internal functionality of the company and have the insider information about the functionality of the firm.
-But he lacks a broader perspective on how are competitors innovating, how is the technology evolving and what will be the impact of this evolution on Company XYZ’s product.
-In the past, he has been right 70% times.
-
->1.
-Financial Advisor of Company XYZ: This person has a broader perspective on how companies strategy will fair of in this competitive environment.
-However, he lacks a view on how the company’s internal policies are fairing off.
-In the past, he has been right 75% times.
-
->1.
-Stock Market Trader: This person has observed the company’s stock price over past 3 years.
-He knows the seasonality trends and how the overall market is performing.
-He also has developed a strong intuition on how stocks might vary over time.
-In the past, he has been right 70% times.
-
->1.
-Employee of a competitor: This person knows the internal functionality of the competitor firms and is aware of certain changes which are yet to be brought.
-He lacks a sight of company in focus and the external factors which can relate the growth of competitor with the company of subject.
-In the past, he has been right 60% of times.
-
->1.
-Market Research team in same segment: This team analyzes the customer preference of company XYZ’s product over others and how is this changing with time.
-Because he deals with customer side, he is unaware of the changes company XYZ will bring because of alignment to its own goals.
-In the past, they have been right 75% of times.
-
->1.
-Social Media Expert: This person can help us understand how has company XYZ has positioned its products in the market.
-And how are the sentiment of customers changing over time towards company.
-He is unaware of any kind of details beyond digital marketing.
-In the past, he has been right 65% of times.
-
-Given the broad spectrum of access we have, we can probably combine all the information and make an informed decision.
-In a scenario when all the 6 experts/teams verify that it’s a good decision(assuming all the predictions are independent of each other), we will get a combined accuracy rate of:
-
-```
-		  1 - 30%*25%*30%*40%*25%*35%
-
-		= 1 - 0.07875 = 99.92125%
-```
-
-Assumption: The assumption used here that all the predictions are completely independent is slightly extreme as they are expected to be correlated.
-However, we see how we can be so sure by combining various predictions together.
-
-Let us now change the scenario slightly. This time we have 6 experts, all of them are employee of company XYZ working in the same division. Everyone has a propensity of 70% to advocate correctly.
-
->*What if we combine all this advice together, can we still raise up our confidence to >99%?*
-
-Obviously not, as all the predictions are based on very similar set of information.
-They are certain to be influenced by similar set of information and the only variation in their advice would be due to their personal opinions & collected facts about the firm.
-
-Ensemble learning is the art of combining individual models together to improve the stability and predictive power of the model.
-
-####Error Sources in Ensemble Learning (Variance vs. Bias)
-The error emerging from any model can be broken down into three components mathematically.
-
-**Error of a model** Why is this important in the current context? To understand what really goes behind an ensemble model, we need to first understand what causes error in the model.
-
-**Bias error** quantifies how much on an average the predicted values differ from the actual value. A high bias error means we have a under-performing model which keeps on missing important trends.
-
-**Variance** quantifies how much the predictions made on same observations differ from each other.
-A high variance model will over-fit on your training population and perform badly on any observation beyond training.
-
-Normally, as you increase the complexity of your model, you will see a reduction in error due to lower bias in the model.
-However, this only happens till a particular point.
-As you continue to make your model more complex, you end up **over-fitting your model** and hence your model will start suffering from high variance.
-
-A champion model should maintain a balance between these two types of errors.
-This is known as the **trade-off management of bias-variance errors.** Ensemble learning is one way to execute this trade off analysis.
-
-####Commonly used Ensemble learning techniques
-**Bagging** tries to implement similar learners on small sample populations and then takes a mean of all the predictions.
-In generalized bagging, you can use different learners on different population.
-As you can expect this helps us to reduce the variance error.
-
-**Boosting** is an iterative technique which adjust the weight of an observation based on the last classification.
-If an observation was classified incorrectly, it tries to increase the weight of this observation and vice versa.
-Boosting in general decreases the bias error and builds strong predictive models.
-However, they may sometimes over fit on the training data.
-
-**Stacking** is a very interesting way of combining models.
-Here we use a learner to combine output from different learners.
-This can lead to decrease in either bias or variance error depending on the combining learner we use.
-
-####Random Forest Models
-Random Forest models are a type of ensemble models, particularly bagging models. They are part of the tree based model family. RF can be run on:
-
-*  Word count or n-gram count vectors
-*  Word level TF-IDF vectors
-
-Resources for Random Forest: 
-https://www.analyticsvidhya.com/blog/2014/06/introduction-random-forest-simplified/
-
-####Extreme Gradient Boosting Model
-Boosting models are another type of ensemble models part of tree based models.
-Boosting is a machine learning ensemble meta-algorithm for primarily reducing bias, and also variance in supervised learning, and a family of machine learning algorithms that convert weak learners to strong ones.
-A weak learner is defined to be a classifier that is only slightly correlated with the true classification (it can label examples better than random guessing).
-XGB can be run on:
-
-*  Word count or n-gram count vectors
-*  Word level TF-IDF vectors
-*  Character Level TF-IDF vectors
-
-###Hierarchical Clustering
-In data mining and statistics, hierarchical clustering (also called hierarchical cluster analysis or HCA) is a method of cluster analysis which seeks to build a hierarchy of clusters.
-
-###k-Means Clustering
-k-means clustering is a method of vector quantization, originally from signal processing, that is popular for cluster analysis in data mining.
-k-means clustering aims to partition n observations into k clusters in which each observation belongs to the cluster with the nearest mean, serving as a prototype of the cluster.
-This results in a partitioning of the data space into Voronoi cells.
-
-The problem is computationally difficult (NP-hard); however, efficient heuristic algorithms converge quickly to a local optimum.
-These are usually similar to the expectation-maximization algorithm for mixtures of Gaussian distributions via an iterative refinement approach employed by both k-means and Gaussian mixture modeling.
-They both use cluster centers to model the data; however, k-means clustering tends to find clusters of comparable spatial extent, while the expectation-maximization mechanism allows clusters to have different shapes.
-
-The algorithm has a loose relationship to the k-nearest neighbor classifier, a popular machine learning technique for classification that is often confused with k-means due to the name.
-Applying the 1-nearest neighbor classifier the cluster centers obtained by k-means classifies new data into the existing clusters.
-This is known as nearest centroid classifier or Rocchio algorithm.
-
-###Mixture Model
-In statistics, a mixture model is a probabilistic model for representing the presence of subpopulations within an overall population, without requiring that an observed data set should identify the sub-population to which an individual observation belongs.
-Formally a mixture model corresponds to the mixture distribution that represents the probability distribution of observations in the overall population.
-However, while problems associated with "mixture distributions" relate to deriving the properties of the overall population from those of the sub-populations, "mixture models" are used to make statistical inferences about the properties of the sub-populations given only observations on the pooled population, without sub-population identity information.
-
-Topics in a document: Assume that a document is composed of N different words from a total vocabulary of size V, where each word corresponds to one of K possible topics.
-The distribution of such words could be modelled as a mixture of K different V-dimensional categorical distributions.
-A model of this sort is commonly termed a topic model.
-Note that expectation maximization applied to such a model will typically fail to produce realistic results, due (among other things) to the excessive number of parameters.
-Some sorts of additional assumptions are typically necessary to get good results.
-
-A prior distribution is placed over the parameters describing the topic distributions, using a Dirichlet distribution with a concentration parameter that is set significantly below 1, so as to encourage sparse distributions (where only a small number of words have significantly non-zero probabilities).
-
-Some sort of additional constraint is placed over the topic identities of words, to take advantage of natural clustering.
-
-For example, a Markov chain could be placed on the topic identities (i.e., the latent variables specifying the mixture component of each observation), corresponding to the fact that nearby words belong to similar topics.
-(This results in a hidden Markov model, specifically one where a prior distribution is placed over state transitions that favors transitions that stay in the same state.)
-
-Another possibility is the latent Dirichlet allocation model, which divides up the words into D different documents and assumes that in each document only a small number of topics occur with any frequency.
-
-###Principal Component Analysis 
-PCA is a statistical procedure that uses an orthogonal transformation to convert a set of observations of possibly correlated variables (entities each of which takes on various numerical values) into a set of values of linearly uncorrelated variables called principal components.
-This transformation is defined in such a way that the first principal component has the largest possible variance (that is, accounts for as much of the variability in the data as possible), and each succeeding component in turn has the highest variance possible under the constraint that it is orthogonal to the preceding components.
-The resulting vectors (each being a linear combination of the variables and containing n observations) are an uncorrelated orthogonal basis set.
-PCA is sensitive to the relative scaling of the original variables.
-
-PCA is mostly used as a tool in exploratory data analysis and for making predictive models.
-It is often used to visualize genetic distance and relatedness between populations.
-PCA can be done by eigenvalue decomposition of a data covariance (or correlation) matrix or singular value decomposition of a data matrix, usually after a normalization step of the initial data.
-The normalization of each attribute consists of mean centering – subtracting each data value from its variable's measured mean so that its empirical mean (average) is zero – and, possibly, normalizing each variable's variance to make it equal to 1.
-The results of a PCA are usually discussed in terms of component scores, sometimes called factor scores (the transformed variable values corresponding to a particular data point), and loadings (the weight by which each standardized original variable should be multiplied to get the component score).
-If component scores are standardized to unit variance loadings must contain the data variance in them (and that is the magnitude of eigenvalues).
-If component scores are not standardized (therefore they contain the data variance) then loadings must be unit-scaled, ("normalized") and these weights are called eigenvectors; they are the cosines of orthogonal rotation of variables into principal components or back.
-
-PCA is the simplest of the true eigenvector-based multivariate analyses.
-Often, its operation can be thought of as revealing the internal structure of the data in a way that best explains the variance in the data.
-If a multivariate dataset is visualised as a set of coordinates in a high-dimensional data space (1 axis per variable), PCA can supply the user with a lower-dimensional picture, a projection of this object when viewed from its most informative viewpoint.
-This is done by using only the first few principal components so that the dimensionality of the transformed data is reduced.
-
-PCA is closely related to factor analysis.
-Factor analysis typically incorporates more domain specific assumptions about the underlying structure and solves eigenvectors of a slightly different matrix.
-
-PCA is also related to canonical correlation analysis (CCA).
-CCA defines coordinate systems that optimally describe the cross-covariance between two datasets while PCA defines a new orthogonal coordinate system that optimally describes variance in a single dataset.
-
-##Strategies for Improving Text Classification Models
-Accuracy can be improved using one or more of these strategies.
-
-1. Text Cleaning. Removing stopwords, punctuations marks, suffix variations etc. can  reduce the noise present in text data.
-
-2.
-Hstacking arrays of text / NLP features with text feature vectors in sequence horizontally (column wise).
-After extracting features initially, try combining different feature vectors to improve the accuracy of the classifier.
-
-3. Tuning the Classifier. Adjusting the parameters of the classifier is an important step towards a best fit model.
-
-4. Testing Different Ensemble Models. Stacking different models and blending their outputs can help to further improve the results.
-
-5.
-Weighting words: we can significantly improve our algorithm by accounting for the commonality of each word.
-The word “is” should carry a lower weigh than the word “sandwich” in most cases, because it is more common.
-
-##Evaluation
-How to evaluate the different components of the overall text classification problem.
-
-###Gini Index
-One of the most common methods for quantifying the discrimination level of a feature is the use of a measure known as the gini-index.
-
-###Cohen's Kappa Score
-Cohen's kappa is a measure of interrater reliability used in social sciences. This excerpt from an R blog explains it.
-
-**Using kappa score for comparisons of groups and methods**
-https://www.personality-project.org/r/html/kappa.html
-https://www.r-bloggers.com/k-is-for-cohens-kappa/
-
-Last April, during the A to Z of Statistics, I blogged about Cohen’s kappa, a measure of interrater reliability.
-Cohen’s kappa is a way to assess whether two raters or judges are rating something the same way.
-And thanks to an R package called **irr**, it’s very easy to compute.
-But first, let’s talk about why you would use Cohen’s kappa and why it’s superior to a more simple measure of interrater reliability, interrater agreement.
-
-***
-***
+####\
 
@@
 
-As in the prior method, each class is given with some number of example sentences.
+Each class is given with some number of example sentences.
 Once again each sentence is broken down by word (stemmed) and each word becomes an input for the neural network.
@@
 
-###Types of Neural Networks
+These methods are beyond the scope of this project. As a reference I have provided some information about these sub-types:
+
+*  Shallow Neural Networks
+*  Deep Neural Networks
+	+  Convolutional Neural Network (CNN)
+	+  Long Short Term Modelr (LSTM)
+	+  Gated Recurrent Unit (GRU)
+	+  Bidirectional RNN
+	+  Recurrent Convolutional Neural Network (RCNN)
+	+  Other Variants of Deep Neural Networks
+
 A shallow neural network contains mainly three types of layers – input layer, hidden layer, and output layer.
-
 Deep Neural Networks are more complex neural networks in which the hidden layers performs much more complex operations than simple sigmoid or relu activations.
@@
 
-####Convolutional Neural Network
 In Convolutional neural networks, convolutions over the input layer are used to compute the output.
@@
 
-####Recurrent Neural Network: LSTM
 Unlike Feed-forward neural networks in which activation outputs are propagated only in one direction, the activation outputs from neurons propagate in both directions (from inputs to outputs and from outputs to inputs) in Recurrent Neural Networks.
@@
 
-####Recurrent Neural Network: GRU
-Gated Recurrent Units are another form of recurrent neural networks.
+Gated Recurrent Units are another form of recurrent neural networks. RNN layers can be wrapped in Bidirectional layers as well.
 
-####Bidirectional RNN
-RNN layers can be wrapped in Bidirectional layers as well.
-
-####Recurrent Convolutional Neural Network
 Once the essential architectures have been tried out, one can try different variants of these layers such as recurrent convolutional neural network. Another variants can be:

### 2019-02-25 Classifier_comparison.Rmd
updates

--- 2019-02-23 Text Classifier Models.Rmd
@@
 ---
-title: "Text Classification Primer"
-author: "Dan Johnson"
-date: "1/9/2019"
-output: html_document
+title: "Designing a Classifier Evaluation Workflow"
+date: "2/21/2019"
+bibliography: library.bib
+biblio-style: apalike
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, child="_setup.Rmd"}
 ```
 
-For this text classification project I am using an initial codebook and corpus of hand-coded TA comments to test various approaches to semi- and fully automated classification methods and, if appropriate, revise the underlying classification scheme.
-
-This page provides background on general mechanics of text classification and various classifier methods.
-Other pages describe the dataset structure [LINK] (TA comments corpus), and [LINK] development of the initial codebook.
-
-#Workflow
-Text classification problems can be broken down into four overlapping sub-problems. Each piece may be more or less important to the overall problem, but all need to be considered:
-
-1.  Feature selection and engineering
-2.  Topic categorization and modeling
-3.  Selecting an appropriate text classifier method
-4.  Evaluating the outcomes
-
-##Feature Selection and Engineering
-This is the process of selecting, tabulating, calculating, or creating text-based features for subsequent evaluation. This includes:
-
-*  Identifying existing, directly observable text features within the dataset
-*  Engineering additional features using established NLP methods
-*	 Reverse engineering relevant features based on an empirically defined feature set.
-
-The features listed below are used regularly for natural language processing.
-Not all will be appropriate for all situations though; selecting, evaluating, comparing, and validating the appropriate features for the classification task is an iterative process, not a one-time decision.
-
-*  Count-based Vectors
-	+  Word Count – total number of words in the documents
-	+  Character Count – total number of characters in the documents
-	+  Punctuation Count – total number of punctuation marks in the documents
-	+  Upper Case Count – total number of upper case words in the documents
-	+  Title Word Count – total number of consecutive upper case (title) words in the documents
-*  Distribution and Frequency Calculations
-	+  Average word density of sentences 
-	+  Average length of the words used in the documents
-	+  TF-IDF
-	+  Reading level estimates (described below)
-*  Part of Speech Tagging and Frequency Distribution of Tags 
-	+ Noun Count
-	+ Verb Count
-	+ Adjective Count
-	+ Adverb Count
-	+ Pronoun Count
-*  Word Embedding (described below)
-
-Additional features are available for longer texts such as multi-sentence paragraphs.
-
-*  Word choices
-*  Lexical diversity (described below)
-*  Stance patterns - hedging, booster, generalization words or phrasing 
-*  Concision - comparisons of noun phrase density versus comment clause density
-*  Cohesion - use of words that are additive (also), countering (however), specify or emphasize (even, both, especially), connecting (former, latter)
-*  Civility and uncertainty - language that acknowledges limits of statement, critiques ideas not people, and makes fewer generalizations
-
-###Lexical Diversity
-This is not the same as word choices.
-This explanation is an excerpt from "quanteda: Quantitative Analysis of Textual Data", which describes how to use ```textstat_lexdiv``` to calculate lexical diversity.
-https://rdrr.io/cran/quanteda/man/textstat_lexdiv.html
-http://quanteda.io/reference/textstat_lexdiv.html
-
-Consider a speaker, who uses the term "allow" multiple times throughout the speech, compared to an another speaker who uses terms "allow, concur, acquiesce, accede, and avow"" for the same word.
-The latter speech has more lexical diversity than the former.
-Lexical diversity is widely believed to be an important parameter to rate a document in terms of textual richness and effectiveness.
-
-Lexical diversity, in simple terms, is a measurement of the breadth and variety of vocabulary used in a document.
-The different measures of lexical diversity are TTR, MSTTR, MATTR, C, R, CTTR, U, S, K, Maas, HD-D, MTLD, and MTLD-MA.
-
-koRpus package in R also provides functions to estimate lexical diversity or complexity.
-https://subscription.packtpub.com/book/big_data_and_business_intelligence/9781783551811/2/ch02lvl1sec16/lexical-diversity
-
-###Readability metrics
-These estimate the relative grade level or age of a typical reader who could comprehend 80+% (typically) of the text.
-Most readability metrics require longer blocks of text (100+ words) to achieve a stable score.
-They use some combination of word count, average syllables per word, and sentence length to estimate reading level.
-
-The metrics differ in how appropriate they are for specific situations.
-Some were developed for general writing, others for specific school age groups, and still others for technical documents; the US Navy in particular developed a suite of metrics for evaluating their training documents.
-
-The ```readability library``` (v0.1.1) is a collection of readability tools that utilize the ```syllable``` package for fast calculation of readability scores by grouping variables.
-It calculates Flesch Kincaid, Gunning Fog Index, Coleman Liau, SMOG, Automated Readability Index and an average of the 5 readability scores.
-These all are general scales; other scores may not be available in R packages.
-
-###Word Embedding
-**Word embedding** is a set of language modeling and feature learning techniques in natural language processing (NLP) where words or phrases from the vocabulary are mapped to vectors of real numbers.
-Conceptually it involves a mathematical embedding from a space with one dimension per word to a continuous vector space with a much lower dimension.
-
-Methods to generate the mapping include neural networks, dimensionality reduction, probabilistic models, knowledge base methods, and explicit representation in terms of the context in which words appear.
-The technique of representing words as vectors has roots in the 1960s with the development of the vector space model for information retrieval.
-
-Word embeddings come in two different styles, one in which words are expressed as vectors of co-occurring words, and another in which words are expressed as vectors of linguistic contexts in which the words occur.
-Most new word embedding techniques rely on a neural network architecture instead of more traditional n-gram models and unsupervised learning.
-
-One of the main limitations of word embeddings (word vector space models in general) is that possible meanings of a word are conflated into a single representation (a single vector in the semantic space).
-Sense embeddings are a solution to this problem: individual meanings of words are represented as distinct vectors in the space.
-
-Tools like **Principal Component Analysis (PCA)** and **T-Distributed Stochastic Neighbour Embedding (t-SNE)** are used to reduce the dimensionality of word vector spaces and visualize word embeddings and clusters.
-
-Software for training and using word embeddings includes Tomas Mikolov's Word2vec, Stanford University's GloVe, fastText, Gensim, AllenNLP's Elmo, Indra and Deeplearning4j.
-Three of these tools are described further below.
-
-####Word2vec
-Word2vec is a group of shallow, two-layer neural network models that produce word embeddings.
-The toolkit can train vector space models to reconstruct linguistic contexts of words.
-Word2vec takes as its input a large corpus of text and produces a vector space, typically of several hundred dimensions, with each unique word in the corpus being assigned a corresponding vector in the space.
-Word vectors are positioned in the vector space so words that share common contexts in the corpus are located in close proximity to one another in the space.
-
-The use of different model parameters and different corpus sizes can greatly affect the quality of a word2vec model.
-Accuracy can be improved in a number of ways, including the choice of model architecture (CBOW or Skip-Gram), increasing the training data set, increasing the number of vector dimensions, and increasing the window size of words considered by the algorithm.
-Each of these improvements comes with the cost of increased computational complexity and therefore increased model generation time.
-
-In models using large corpora and a high number of dimensions, the skip-gram model yields the highest overall accuracy, and consistently produces the highest accuracy on semantic relationships, as well as yielding the highest syntactic accuracy in most cases.
-However, the CBOW is less computationally expensive and yields similar accuracy results.
-
-Accuracy increases overall as the number of words used increases, and as the number of dimensions increases.
-Altszyler et al.
-(2017) studied Word2vec performance in two semantic tests for different corpus size.
-They found that Word2vec has a steep learning curve, outperforming another word-embedding technique (LSA) when it is trained with medium to large corpus size (more than 10 million words).
-However, with a small training corpus LSA showed better performance.
-Additionally they show that the best parameter setting depends on the task and the training corpus.
-Nevertheless, for skip-gram models trained in medium size corpora, with 50 dimensions, a window size of 15 and 10 negative samples seems to be a good parameter setting.
-
-####fastText
-fastText is a Python-based library created by Facebook's AI Research (FAIR) lab for learning of word embeddings and text classification.
-The model allows to create an unsupervised learning or supervised learning algorithm for obtaining vector representations for words.
-fastText uses a neural network for word embedding.
-
-https://fasttext.cc/
-
-####GloVe
-GloVe, short for Global Vectors, is a model for distributed word representation.
-The model is an unsupervised learning algorithm for obtaining vector representations for words.
-Training is performed on aggregated global word-word co-occurrence statistics from a corpus, and the resulting representations showcase interesting linear substructures of the word vector space.
-It is developed as an open-source project at Stanford.
-Pre-trained word vectors data are made available under the Public Domain Dedication and License v1.0.
-
-GloVe uses Euclidean distance (or cosine similarity) between two word vector for measuring the linguistic or semantic similarity of the corresponding words.
-Sometimes, the nearest neighbors according to this metric reveal rare but relevant words that lie outside an average human's vocabulary.
-For example, here are the closest words to the target word **frog**:
-
-*  frog
-*  frogs
-*  toad
-*  litoria
-*  leptodactylidae
-*  rana
-*  lizard
-*  eleutherodactylus
-
-The similarity metrics used for nearest neighbor evaluations produce a single scalar that quantifies the relatedness of two words.
-This simplicity can be problematic since two given words almost always exhibit more intricate relationships than can be captured by a single number.
-For example, man may be regarded as similar to woman in that both words describe human beings; on the other hand, the two words are often considered opposites since they highlight a primary axis along which humans differ from one another.
-
-In order to capture in a quantitative way the nuance necessary to distinguish man from woman, it is necessary for a model to associate more than a single number to the word pair.
-A natural and simple candidate for an enlarged set of discriminative numbers is the vector difference between the two word vectors.
-GloVe is designed in order that such vector differences capture as much as possible the meaning specified by the juxtaposition of two words.
-
-https://nlp.stanford.edu/projects/glove/
-
-##Sub-Problem Two: Topic Categorization and Modeling Methods
-Establishing meaningful text classification categories depends on the goals of the project or the question being asked.
-Are relevant categories known already, or do they need to be established?
-What are the main topics
-
-**Curated Categories** (my provisional term until I find the appropriate equivalent in current use) are established by subject matter experts. The established categories can be constructed from:
-
-*  A limited keywords vocabulary (National Library of Medicine MeSH categories for example)
-*  Disciplinary conventions or shared best practices (the Dublin Core is an example)
-*  Prior empirical studies
-*  The dataset itself, using *de novo* iterative coding. This process is described further here [LINK] and here [LINK].
-
-**Topic Modeling** is a group of techniques for identifying groups of words (the "topic")  that best summarize the information in a collection of documents.
-Topic modeling is an iterative/ generative process that does not require any prior knowledge of the documents or relationships.
-
-###Topic Modeling Methods to Identify Features
-####Latent Dirichlet Allocation
-(LDA) is a common method for generating topic model features.
-LDA is an iterative model which starts from a fixed number of topics.
-Each topic is represented as a distribution over words, and each document is then represented as a distribution over topics.
-Although the tokens themselves are meaningless, the probability distributions over words provided by the topics provide a sense of the different ideas contained in the documents.
-
-####Brown Clustering
-A hierarchical clustering algorithm that is applied to text, grouping words into clusters that are assumed to be semantically related by virtue of their having been embedded in similar contexts.
-
-In natural language processing, Brown clustering or IBM clustering is a form of hierarchical clustering of words based on the contexts in which they occur.
-The intuition behind the method is that a class-based language model (also called cluster n-gram model), i.e.
-one where probabilities of words are based on the classes (clusters) of previous words, is used to address the data sparsity problem inherent in language modeling.
-
-Brown groups items (i.e., types) into classes, using a binary merging criterion based on the log-probability of a text under a class-based language model, i.e.
-a probability model that takes the clustering into account.
-
-####Latent semantic analysis
-LSA is a technique in natural language processing of analyzing relationships between a set of documents and the terms they contain by producing a set of concepts related to the documents and terms.
-LSA assumes that words that are close in meaning will occur in similar pieces of text (the distributional hypothesis).
-A matrix containing word counts per paragraph (rows represent unique words and columns represent each paragraph) is constructed from a large piece of text and a mathematical technique called singular value decomposition (SVD) is used to reduce the number of rows while preserving the similarity structure among columns.
-Words are then compared by taking the cosine of the angle between the two vectors (or the dot product between the normalizations of the two vectors) formed by any two rows.
-Values close to 1 represent very similar words while values close to 0 represent very dissimilar words.
-
-In the context of its application to information retrieval, it is sometimes called latent semantic indexing (LSI).
-
-LSA can use a term-document matrix which describes the occurrences of terms in documents; it is a sparse matrix whose rows correspond to terms and whose columns correspond to documents.
-A typical example of the weighting of the elements of the matrix is tf-idf (term frequency–inverse document frequency): the weight of an element of the matrix is proportional to the number of times the terms appear in each document, where rare terms are upweighted to reflect their relative importance.
-
-Applications of LSA include:
-
-*  Comparing documents in low-dimensional space (data clustering, document classification).
-*  Finding relations between terms (synonymy and polysemy).
-Synonymy is the phenomenon where different words describe the same idea.
-Polysemy is the phenomenon where the same word has multiple meanings.
-*  Given a query of terms, translating it into the low-dimensional space, and find matching documents (information retrieval).
-*  Analyzing word association in text corpus.
-
-LSA overcomes two of the most problematic constraints of Boolean keyword queries: multiple words that have similar meanings (synonymy) and words that have more than one meaning (polysemy).
-
-LSA is also used to perform automated document categorization.
-Document categorization is the assignment of documents to one or more predefined categories based on their similarity to the conceptual content of the categories.
-Several experiments have demonstrated that there are a number of correlations between the way LSI and humans process and categorize text.
-
-LSA uses example documents to establish the conceptual basis for each category.
-During categorization processing, the concepts contained in the documents being categorized are compared to the concepts contained in the example items, and a category (or categories) is assigned to the documents based on the similarities between the concepts they contain and the concepts that are contained in the example documents.
-
-LSA automatically adapts to new and changing terminology, and has been shown to be very tolerant of noise (i.e., misspelled words, typographical errors, unreadable characters, etc.).[24] This is especially important for applications using text derived from Optical Character Recognition (OCR) and speech-to-text conversion.
-LSA also deals effectively with sparse, ambiguous, and contradictory data.
-
-Text does not need to be in sentence form for LSA to be effective.
-It can work with lists, free-form notes, email, Web-based content, etc.
-As long as a collection of text contains multiple terms, LSA can be used to identify patterns in the relationships between the important terms and concepts contained in the text.
-
-##Sub-Problem Three: Choosing a Text Classifier Method
-The many methods for classifying texts fall into three general categories:
-
-*  Pattern matching strategies (does a particular phrase appear?)
-*  Algorithms (how closely does the query text meet pre-set criteria?)
-*  Neural networks (the network decides its own criteria, aka machine learning)
-
-This edited excerpt nicely describes the overall approach and problems of text classification. It is from: **xxxxx**
-
-Understanding how chatbots work can help us understand text classifiers, because the classifier is the fundamental piece that makes the chatbot work.
-Let’s look at the inner workings of a Multinomial Naive Bayes algorithm for text classification and natural language processing (NLP).
-
-This classifier is “naive” because each word is treated as having no connection with other words in the sentence being classified.
-In other words it assumes independence between "features," in this case, words.
-
-For example, in the sentence “the fox jumped over the log”, there is no relationship between “jumped” and “fox” or “log”.
-In NLP, this collection of words is referred to as “a bag of words”.
-While this is incredibly naive, the classifier isn’t attempting to understand the meaning of a sentence, it’s trying to classify it.
-
-###Pattern Matchers
-Early chatbots used **pattern matching** to classify text and produce a response.
-This is often referred to as “brute force” as the author of the system needs to describe every pattern for which there is a response.
-
-A standard structure for these patterns is “AIML” (artificial intelligence markup language). Its use of the term “artificial intelligence” is quite an embellishment, but that’s another story.
-
-A simple pattern matching definition:
-```
-<aiml version = "1.0.1" encoding = "UTF-8"?>
-   <category>
-      <pattern> WHO IS ALBERT EINSTEIN </pattern>
-      <template>Albert Einstein was a German physicist.</template>
-   </category>
-
-   <category>
-      <pattern> WHO IS Isaac NEWTON </pattern>
-      <template>Isaac Newton was a English physicist and mathematician.</template>
-   </category>
-
-   <category>
-      <pattern>DO YOU KNOW WHO * IS</pattern>
-      <template>
-         <srai>WHO IS <star/></srai>
-      </template>
-   </category>
-</aiml>
-```
-
-The machine then produces:
-
-> Human: Do you know who Albert Einstein is
-> Robot: Albert Einstein was a German physicist.
-
-It knows who a physicist is only because his or her name has an associated pattern.
-Likewise it responds to anything solely because of an authored pattern.
-Given hundreds or thousands of patterns you might see a chatbot “persona” emerge.
-
-In 2000 a chatbot built by John Denning and colleagues using this approach was in the news for passing the "Turing test." It emulated the replies of a 13 year old boy from Ukraine (broken English and all).
-I met with John in 2015 and he made no false pretenses about the internal workings of this automaton.
-It may have been “brute force” but it proved a point: parts of a conversation can be made to appear “natural” using a sufficiently large definition of patterns.
-It proved Alan Turing’s assertion, that this question of a machine fooling humans was “meaningless”.
-
-An example of this approach used in building chatbots is PandoraBots, who claim over 285k chatbots have been constructed using their framework.
-
-Automated phone voice response systems that recognize patterns of spoken words and direct calls accordingly use a similar approach.
-
-###Algorithm-Based Classification
-Creating a brute-force pattern matcher is daunting: for each unique input a pattern must be available to specify a response. An algorithm-based method approaches the problem mathematically.
-
-One text classification algorithm is "Multinomial Naive Bayes." Given a training set of sentences, each belonging to a class (that is, a particular group), we can count the occurrence of each word in each class, account for its commonality and assign each word a score for each class.
-Factoring for commonality is important: matching the word “it” is considerably less meaningful than a match for the word “cheese”.
-When we input a new sentence, scores for each word in the new sentence are compared to scores for those same words in the training set.
-The class with the highest OVERALL score for the input sentence is the class to which the input sentence most likely belongs.
-
-Look at this sample training set:
-```
-class: weather
-    "is it nice outside?"
-    "how is it outside?"
-    "is the weather nice?"
-
-class: greeting
-    "how are you?"
-    "hello there"
-    "how is it going?"
-```
-
-Let’s classify a few sample input sentences:
-```
-input: "Hi there"
- term: "hi" (no matches in either class)
- term: "there" (matches a word in class: greeting)
- classification: greeting (score=1)
-
-input: "What’s it like outside?"
- term: "it" (matches words in class: weather (2), greeting(1))
- term: "outside (class: weather (2) )
- classification: weather (score=4)
-```
-
-Notice that the classification for "What’s it like outside" found a term in the "greeting" class but the term similarities to the "weather" class produced a higher score.
-By using an equation we are looking for word matches given some sample sentences for each class, and we avoid having to identify every pattern.
-
-The classification score produced identifies the class with the highest term matches (accounting for commonality of words) but this has limitations.
-A score is not the same as a probability, a score tells us which intent is most like the sentence but not the likelihood of it being a match.
-Thus it is difficult to apply a threshold for which classification scores to accept or not.
-Having the highest score from this type of algorithm only provides a relative basis, it may still be an inherently weak classification.
-Also the algorithm doesn’t account for what a sentence is not, it only counts what it is like.
-You might say this approach doesn’t consider what makes a sentence not a given class.
-
-Many text and spoken word frameworks use similar algorithms such as this to classify intent.
-Most of what’s taking place is word counting against training datasets.
-It’s "naive" but surprisingly effective.
-
-I will be using an algorithm-based classification strategy for this project.
-
-###Neural Networks
-These methods are beyond the scope of this project. Anyone interested in knowing more can search online for information about:
-
-*  Shallow Neural Networks
-*  Deep Neural Networks
-	+  Convolutional Neural Network (CNN)
-	+  Long Short Term Modelr (LSTM)
-	+  Gated Recurrent Unit (GRU)
-	+  Bidirectional RNN
-	+  Recurrent Convolutional Neural Network (RCNN)
-	+  Other Variants of Deep Neural Networks
-
-###Designing an Evaluation Workflow
 The workflow for comparing classification algorithms starts with identifying potentially usable classifier methods.
@@
 
-###Picking the Algorithm(s)
+####\
+
+#Picking the Algorithm(s)
 Algorithms approach text classification several ways.
@@
 
-###Naive Bayes
+####\
+
+##Naive Bayes
 A classification technique based on Bayes’ Theorem with an assumption of independence among predictors.
@@
 
-###Linear Classifier
+####\
+
+##Linear Classifier
 The linear regression algorithm is one of the fundamental supervised machine-learning algorithms due to its relative simplicity and well-known properties.
@@
 
-###Logistic Regression
+####\
+
+##Logistic Regression
 A logistic regression measures the relationship between the categorical dependent variable and one or more independent variables by estimating probabilities using a logistic/sigmoid function.
@@
 
-###Support Vector Machine (SVM) Model
+####\
+
+##Support Vector Machine (SVM) Model
 This is a supervised machine learning algorithm which can be used for both classification or regression challenges.
@@
 
-###Decision Trees
+####\
+
+##Decision Trees
 Flowchart-like structures in which each internal node represents a "test" on an attribute (e.g.
@@
 
-###Linear Discriminant Analysis
+####\
+
+##Linear Discriminant Analysis
 LDA works when the measurements made on independent variables for each observation are continuous quantities.
@@
 
-###k-Nearest Neighbors
+####\
+
+##k-Nearest Neighbors
 k-NN is a non-parametric method used for classification and regression.
@@
 
-###Similarity Learning
+####\
+
+##Similarity Learning
 Similarity learning is an area of supervised machine learning in artificial intelligence.
@@
 
-###Relevance Feedback/Rocchio Algorithm
+####\
+
+##Relevance Feedback/Rocchio Algorithm
 Relevance feedback is a feature of some information retrieval systems.
@@
 
-###Ensemble Methods
+####\
+
+##Ensemble Methods
 The excerpt below nice describes how ensemble methods work. It is reprinted from: *Ensemble Methods Explained in Simple English*, by Tavish Srivastava, posted August 2, 2015.
@@
 
->1.
+>2.
 Financial Advisor of Company XYZ: This person has a broader perspective on how companies strategy will fair of in this competitive environment.
@@
 
->1.
+>3.
 Stock Market Trader: This person has observed the company’s stock price over past 3 years.
@@
 
->1.
+>4.
 Employee of a competitor: This person knows the internal functionality of the competitor firms and is aware of certain changes which are yet to be brought.
@@
 
->1.
+>5.
 Market Research team in same segment: This team analyzes the customer preference of company XYZ’s product over others and how is this changing with time.
@@
 
->1.
+>6.
 Social Media Expert: This person can help us understand how has company XYZ has positioned its products in the market.
@@
 
-####Error Sources in Ensemble Learning (Variance vs. Bias)
+####\
+
+###Error Sources in Ensemble Learning (Variance vs. Bias)
 The error emerging from any model can be broken down into three components mathematically.
@@
 
-####Commonly used Ensemble learning techniques
+####\
+
+###Commonly used Ensemble learning techniques
 **Bagging** tries to implement similar learners on small sample populations and then takes a mean of all the predictions.
@@
 
-####Random Forest Models
+####\
+
+###Random Forest Models
 Random Forest models are a type of ensemble models, particularly bagging models. They are part of the tree based model family. RF can be run on:
@@
 
-####Extreme Gradient Boosting Model
+####\
+
+###Extreme Gradient Boosting Model
 Boosting models are another type of ensemble models part of tree based models.
@@
 
-###Hierarchical Clustering
+####\
+
+##Hierarchical Clustering
 In data mining and statistics, hierarchical clustering (also called hierarchical cluster analysis or HCA) is a method of cluster analysis which seeks to build a hierarchy of clusters.
 
-###k-Means Clustering
+####\
+
+##k-Means Clustering
 k-means clustering is a method of vector quantization, originally from signal processing, that is popular for cluster analysis in data mining.
@@
 
-###Mixture Model
+####\
+
+##Mixture Model
 In statistics, a mixture model is a probabilistic model for representing the presence of subpopulations within an overall population, without requiring that an observed data set should identify the sub-population to which an individual observation belongs.
@@
 
-###Principal Component Analysis 
+####\
+
+##Principal Component Analysis 
 PCA is a statistical procedure that uses an orthogonal transformation to convert a set of observations of possibly correlated variables (entities each of which takes on various numerical values) into a set of values of linearly uncorrelated variables called principal components.
@@
 
-##Strategies for Improving Text Classification Models
-Accuracy can be improved using one or more of these strategies.
-
-1. Text Cleaning. Removing stopwords, punctuations marks, suffix variations etc. can  reduce the noise present in text data.
-
-2.
-Hstacking arrays of text / NLP features with text feature vectors in sequence horizontally (column wise).
-After extracting features initially, try combining different feature vectors to improve the accuracy of the classifier.
-
-3. Tuning the Classifier. Adjusting the parameters of the classifier is an important step towards a best fit model.
-
-4. Testing Different Ensemble Models. Stacking different models and blending their outputs can help to further improve the results.
-
-5.
-Weighting words: we can significantly improve our algorithm by accounting for the commonality of each word.
-The word “is” should carry a lower weigh than the word “sandwich” in most cases, because it is more common.
-
-##Evaluation
-How to evaluate the different components of the overall text classification problem.
-
-###Gini Index
-One of the most common methods for quantifying the discrimination level of a feature is the use of a measure known as the gini-index.
-
-###Cohen's Kappa Score
-Cohen's kappa is a measure of interrater reliability used in social sciences. This excerpt from an R blog explains it.
-
-**Using kappa score for comparisons of groups and methods**
-https://www.personality-project.org/r/html/kappa.html
-https://www.r-bloggers.com/k-is-for-cohens-kappa/
-
-Last April, during the A to Z of Statistics, I blogged about Cohen’s kappa, a measure of interrater reliability.
-Cohen’s kappa is a way to assess whether two raters or judges are rating something the same way.
-And thanks to an R package called **irr**, it’s very easy to compute.
-But first, let’s talk about why you would use Cohen’s kappa and why it’s superior to a more simple measure of interrater reliability, interrater agreement.
-
-***
-***
-
-###Neural Networks
-A neural network is a mathematical model that is designed to behave similar to biological neurons and nervous system.
-These models are used to recognize complex patterns and relationships that exists within a labelled data.
-
-Artificial neural networks, invented in the 1940’s, are a way of calculating an output from an input (a classification) using weighted connections (“synapses”) that are calculated from repeated iterations through training data.
-Each pass through the training data alters the weights such that the neural network produces the output with greater “accuracy” (lower error rate).
-
-There’s not much new about these structures, except today’s software is using much faster processors and can work with a lot more memory.
-The combination of working memory and speed is crucial when you’re doing hundreds of thousands of matrix multiplications (the neural network’s essential math operation).
-
-As in the prior method, each class is given with some number of example sentences.
-Once again each sentence is broken down by word (stemmed) and each word becomes an input for the neural network.
-The synaptic weights are then calculated by iterating through the training data thousands of times, each time adjusting the weights slightly to greater accuracy.
-By recalculating back across multiple layers (“back-propagation”) the weights of all synapses are calibrated while the results are compared to the training data output.
-These weights are like a ‘strength’ measure, in a neuron the synaptic weight is what causes something to be more memorable than not.
-You remember a thing more because you’ve seen it more times: each time the ‘weight’ increases slightly.
-
-At some point the adjustment reaches a point of diminishing returns, this is called “over-fitting” and going beyond this is counter-productive.
-
-The trained neural network is less code than an comparable algorithm but it requires a potentially large matrix of “weights”.
-In a relatively small sample, where the training sentences have 150 unique words and 30 classes this would be a matrix of 150x30.
-Imagine multiplying a matrix of this size 100,000 times to establish a sufficiently low error rate.
-This is where processing speed comes in.
-
-If the neural network sounds magnificently sophisticated, relax, it boils down to matrix multiplication and a formula for reducing values between -1 and 1 or some other minimal range.
-A middle-school math student could learn this in a few hours.
-The hard work is achieving clean training data.
-
-Just as there are variations in pattern matching code and in algorithms, there are variations in neural networks, some more complex than others.
-The basic machinery is the same.
-The essential work is that of classification.
-
-Networks are looking for patterns in collections of terms, each term is reduced to a token.
-In this machine words have no meaning except for their patterned existence within training data.
-The label “artificial intelligence” applied to such machinery is mostly BS.
-
-###Types of Neural Networks
-A shallow neural network contains mainly three types of layers – input layer, hidden layer, and output layer.
-
-Deep Neural Networks are more complex neural networks in which the hidden layers performs much more complex operations than simple sigmoid or relu activations.
-Different types of deep learning models can be applied in text classification problems.
-
-####Convolutional Neural Network
-In Convolutional neural networks, convolutions over the input layer are used to compute the output.
-This results in local connections, where each region of the input is connected to a neuron in the output.
-Each layer applies different filters and combines their results.
-
-####Recurrent Neural Network: LSTM
-Unlike Feed-forward neural networks in which activation outputs are propagated only in one direction, the activation outputs from neurons propagate in both directions (from inputs to outputs and from outputs to inputs) in Recurrent Neural Networks.
-This creates loops in the neural network architecture which acts as a ‘memory state’ of the neurons.
-This state allows the neurons an ability to remember what have been learned so far.
-
-The memory state in RNNs gives an advantage over traditional neural networks but a problem called Vanishing Gradient is associated with them.
-In this problem, while learning with a large number of layers, it becomes really hard for the network to learn and tune the parameters of the earlier layers.
-To address this problem, A new type of RNNs called LSTMs (Long Short Term Memory) Models have been developed.
-
-####Recurrent Neural Network: GRU
-Gated Recurrent Units are another form of recurrent neural networks.
-
-####Bidirectional RNN
-RNN layers can be wrapped in Bidirectional layers as well.
-
-####Recurrent Convolutional Neural Network
-Once the essential architectures have been tried out, one can try different variants of these layers such as recurrent convolutional neural network. Another variants can be:
-
-*  Hierarchical Attention Networks
-*  Sequence to Sequence Models with Attention
-*  Bidirectional Recurrent Convolutional Neural Networks
-*  CNNs and RNNs with more number of layers
+####\

### 2019-02-25 Improving_classifier.Rmd
updates

--- 2019-02-23 Text Classifier Models.Rmd
@@
 ---
-title: "Text Classification Primer"
-author: "Dan Johnson"
-date: "1/9/2019"
-output: html_document
+title: "Improving & Evaluating Text Classifiers"
+date: "2/21/2019"
+bibliography: library.bib
+biblio-style: apalike
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, child="_setup.Rmd"}
 ```
 
-For this text classification project I am using an initial codebook and corpus of hand-coded TA comments to test various approaches to semi- and fully automated classification methods and, if appropriate, revise the underlying classification scheme.
-
-This page provides background on general mechanics of text classification and various classifier methods.
-Other pages describe the dataset structure [LINK] (TA comments corpus), and [LINK] development of the initial codebook.
-
-#Workflow
-Text classification problems can be broken down into four overlapping sub-problems. Each piece may be more or less important to the overall problem, but all need to be considered:
-
-1.  Feature selection and engineering
-2.  Topic categorization and modeling
-3.  Selecting an appropriate text classifier method
-4.  Evaluating the outcomes
-
-##Feature Selection and Engineering
-This is the process of selecting, tabulating, calculating, or creating text-based features for subsequent evaluation. This includes:
-
-*  Identifying existing, directly observable text features within the dataset
-*  Engineering additional features using established NLP methods
-*	 Reverse engineering relevant features based on an empirically defined feature set.
-
-The features listed below are used regularly for natural language processing.
-Not all will be appropriate for all situations though; selecting, evaluating, comparing, and validating the appropriate features for the classification task is an iterative process, not a one-time decision.
-
-*  Count-based Vectors
-	+  Word Count – total number of words in the documents
-	+  Character Count – total number of characters in the documents
-	+  Punctuation Count – total number of punctuation marks in the documents
-	+  Upper Case Count – total number of upper case words in the documents
-	+  Title Word Count – total number of consecutive upper case (title) words in the documents
-*  Distribution and Frequency Calculations
-	+  Average word density of sentences 
-	+  Average length of the words used in the documents
-	+  TF-IDF
-	+  Reading level estimates (described below)
-*  Part of Speech Tagging and Frequency Distribution of Tags 
-	+ Noun Count
-	+ Verb Count
-	+ Adjective Count
-	+ Adverb Count
-	+ Pronoun Count
-*  Word Embedding (described below)
-
-Additional features are available for longer texts such as multi-sentence paragraphs.
-
-*  Word choices
-*  Lexical diversity (described below)
-*  Stance patterns - hedging, booster, generalization words or phrasing 
-*  Concision - comparisons of noun phrase density versus comment clause density
-*  Cohesion - use of words that are additive (also), countering (however), specify or emphasize (even, both, especially), connecting (former, latter)
-*  Civility and uncertainty - language that acknowledges limits of statement, critiques ideas not people, and makes fewer generalizations
-
-###Lexical Diversity
-This is not the same as word choices.
-This explanation is an excerpt from "quanteda: Quantitative Analysis of Textual Data", which describes how to use ```textstat_lexdiv``` to calculate lexical diversity.
-https://rdrr.io/cran/quanteda/man/textstat_lexdiv.html
-http://quanteda.io/reference/textstat_lexdiv.html
-
-Consider a speaker, who uses the term "allow" multiple times throughout the speech, compared to an another speaker who uses terms "allow, concur, acquiesce, accede, and avow"" for the same word.
-The latter speech has more lexical diversity than the former.
-Lexical diversity is widely believed to be an important parameter to rate a document in terms of textual richness and effectiveness.
-
-Lexical diversity, in simple terms, is a measurement of the breadth and variety of vocabulary used in a document.
-The different measures of lexical diversity are TTR, MSTTR, MATTR, C, R, CTTR, U, S, K, Maas, HD-D, MTLD, and MTLD-MA.
-
-koRpus package in R also provides functions to estimate lexical diversity or complexity.
-https://subscription.packtpub.com/book/big_data_and_business_intelligence/9781783551811/2/ch02lvl1sec16/lexical-diversity
-
-###Readability metrics
-These estimate the relative grade level or age of a typical reader who could comprehend 80+% (typically) of the text.
-Most readability metrics require longer blocks of text (100+ words) to achieve a stable score.
-They use some combination of word count, average syllables per word, and sentence length to estimate reading level.
-
-The metrics differ in how appropriate they are for specific situations.
-Some were developed for general writing, others for specific school age groups, and still others for technical documents; the US Navy in particular developed a suite of metrics for evaluating their training documents.
-
-The ```readability library``` (v0.1.1) is a collection of readability tools that utilize the ```syllable``` package for fast calculation of readability scores by grouping variables.
-It calculates Flesch Kincaid, Gunning Fog Index, Coleman Liau, SMOG, Automated Readability Index and an average of the 5 readability scores.
-These all are general scales; other scores may not be available in R packages.
-
-###Word Embedding
-**Word embedding** is a set of language modeling and feature learning techniques in natural language processing (NLP) where words or phrases from the vocabulary are mapped to vectors of real numbers.
-Conceptually it involves a mathematical embedding from a space with one dimension per word to a continuous vector space with a much lower dimension.
-
-Methods to generate the mapping include neural networks, dimensionality reduction, probabilistic models, knowledge base methods, and explicit representation in terms of the context in which words appear.
-The technique of representing words as vectors has roots in the 1960s with the development of the vector space model for information retrieval.
-
-Word embeddings come in two different styles, one in which words are expressed as vectors of co-occurring words, and another in which words are expressed as vectors of linguistic contexts in which the words occur.
-Most new word embedding techniques rely on a neural network architecture instead of more traditional n-gram models and unsupervised learning.
-
-One of the main limitations of word embeddings (word vector space models in general) is that possible meanings of a word are conflated into a single representation (a single vector in the semantic space).
-Sense embeddings are a solution to this problem: individual meanings of words are represented as distinct vectors in the space.
-
-Tools like **Principal Component Analysis (PCA)** and **T-Distributed Stochastic Neighbour Embedding (t-SNE)** are used to reduce the dimensionality of word vector spaces and visualize word embeddings and clusters.
-
-Software for training and using word embeddings includes Tomas Mikolov's Word2vec, Stanford University's GloVe, fastText, Gensim, AllenNLP's Elmo, Indra and Deeplearning4j.
-Three of these tools are described further below.
-
-####Word2vec
-Word2vec is a group of shallow, two-layer neural network models that produce word embeddings.
-The toolkit can train vector space models to reconstruct linguistic contexts of words.
-Word2vec takes as its input a large corpus of text and produces a vector space, typically of several hundred dimensions, with each unique word in the corpus being assigned a corresponding vector in the space.
-Word vectors are positioned in the vector space so words that share common contexts in the corpus are located in close proximity to one another in the space.
-
-The use of different model parameters and different corpus sizes can greatly affect the quality of a word2vec model.
-Accuracy can be improved in a number of ways, including the choice of model architecture (CBOW or Skip-Gram), increasing the training data set, increasing the number of vector dimensions, and increasing the window size of words considered by the algorithm.
-Each of these improvements comes with the cost of increased computational complexity and therefore increased model generation time.
-
-In models using large corpora and a high number of dimensions, the skip-gram model yields the highest overall accuracy, and consistently produces the highest accuracy on semantic relationships, as well as yielding the highest syntactic accuracy in most cases.
-However, the CBOW is less computationally expensive and yields similar accuracy results.
-
-Accuracy increases overall as the number of words used increases, and as the number of dimensions increases.
-Altszyler et al.
-(2017) studied Word2vec performance in two semantic tests for different corpus size.
-They found that Word2vec has a steep learning curve, outperforming another word-embedding technique (LSA) when it is trained with medium to large corpus size (more than 10 million words).
-However, with a small training corpus LSA showed better performance.
-Additionally they show that the best parameter setting depends on the task and the training corpus.
-Nevertheless, for skip-gram models trained in medium size corpora, with 50 dimensions, a window size of 15 and 10 negative samples seems to be a good parameter setting.
-
-####fastText
-fastText is a Python-based library created by Facebook's AI Research (FAIR) lab for learning of word embeddings and text classification.
-The model allows to create an unsupervised learning or supervised learning algorithm for obtaining vector representations for words.
-fastText uses a neural network for word embedding.
-
-https://fasttext.cc/
-
-####GloVe
-GloVe, short for Global Vectors, is a model for distributed word representation.
-The model is an unsupervised learning algorithm for obtaining vector representations for words.
-Training is performed on aggregated global word-word co-occurrence statistics from a corpus, and the resulting representations showcase interesting linear substructures of the word vector space.
-It is developed as an open-source project at Stanford.
-Pre-trained word vectors data are made available under the Public Domain Dedication and License v1.0.
-
-GloVe uses Euclidean distance (or cosine similarity) between two word vector for measuring the linguistic or semantic similarity of the corresponding words.
-Sometimes, the nearest neighbors according to this metric reveal rare but relevant words that lie outside an average human's vocabulary.
-For example, here are the closest words to the target word **frog**:
-
-*  frog
-*  frogs
-*  toad
-*  litoria
-*  leptodactylidae
-*  rana
-*  lizard
-*  eleutherodactylus
-
-The similarity metrics used for nearest neighbor evaluations produce a single scalar that quantifies the relatedness of two words.
-This simplicity can be problematic since two given words almost always exhibit more intricate relationships than can be captured by a single number.
-For example, man may be regarded as similar to woman in that both words describe human beings; on the other hand, the two words are often considered opposites since they highlight a primary axis along which humans differ from one another.
-
-In order to capture in a quantitative way the nuance necessary to distinguish man from woman, it is necessary for a model to associate more than a single number to the word pair.
-A natural and simple candidate for an enlarged set of discriminative numbers is the vector difference between the two word vectors.
-GloVe is designed in order that such vector differences capture as much as possible the meaning specified by the juxtaposition of two words.
-
-https://nlp.stanford.edu/projects/glove/
-
-##Sub-Problem Two: Topic Categorization and Modeling Methods
-Establishing meaningful text classification categories depends on the goals of the project or the question being asked.
-Are relevant categories known already, or do they need to be established?
-What are the main topics
-
-**Curated Categories** (my provisional term until I find the appropriate equivalent in current use) are established by subject matter experts. The established categories can be constructed from:
-
-*  A limited keywords vocabulary (National Library of Medicine MeSH categories for example)
-*  Disciplinary conventions or shared best practices (the Dublin Core is an example)
-*  Prior empirical studies
-*  The dataset itself, using *de novo* iterative coding. This process is described further here [LINK] and here [LINK].
-
-**Topic Modeling** is a group of techniques for identifying groups of words (the "topic")  that best summarize the information in a collection of documents.
-Topic modeling is an iterative/ generative process that does not require any prior knowledge of the documents or relationships.
-
-###Topic Modeling Methods to Identify Features
-####Latent Dirichlet Allocation
-(LDA) is a common method for generating topic model features.
-LDA is an iterative model which starts from a fixed number of topics.
-Each topic is represented as a distribution over words, and each document is then represented as a distribution over topics.
-Although the tokens themselves are meaningless, the probability distributions over words provided by the topics provide a sense of the different ideas contained in the documents.
-
-####Brown Clustering
-A hierarchical clustering algorithm that is applied to text, grouping words into clusters that are assumed to be semantically related by virtue of their having been embedded in similar contexts.
-
-In natural language processing, Brown clustering or IBM clustering is a form of hierarchical clustering of words based on the contexts in which they occur.
-The intuition behind the method is that a class-based language model (also called cluster n-gram model), i.e.
-one where probabilities of words are based on the classes (clusters) of previous words, is used to address the data sparsity problem inherent in language modeling.
-
-Brown groups items (i.e., types) into classes, using a binary merging criterion based on the log-probability of a text under a class-based language model, i.e.
-a probability model that takes the clustering into account.
-
-####Latent semantic analysis
-LSA is a technique in natural language processing of analyzing relationships between a set of documents and the terms they contain by producing a set of concepts related to the documents and terms.
-LSA assumes that words that are close in meaning will occur in similar pieces of text (the distributional hypothesis).
-A matrix containing word counts per paragraph (rows represent unique words and columns represent each paragraph) is constructed from a large piece of text and a mathematical technique called singular value decomposition (SVD) is used to reduce the number of rows while preserving the similarity structure among columns.
-Words are then compared by taking the cosine of the angle between the two vectors (or the dot product between the normalizations of the two vectors) formed by any two rows.
-Values close to 1 represent very similar words while values close to 0 represent very dissimilar words.
-
-In the context of its application to information retrieval, it is sometimes called latent semantic indexing (LSI).
-
-LSA can use a term-document matrix which describes the occurrences of terms in documents; it is a sparse matrix whose rows correspond to terms and whose columns correspond to documents.
-A typical example of the weighting of the elements of the matrix is tf-idf (term frequency–inverse document frequency): the weight of an element of the matrix is proportional to the number of times the terms appear in each document, where rare terms are upweighted to reflect their relative importance.
-
-Applications of LSA include:
-
-*  Comparing documents in low-dimensional space (data clustering, document classification).
-*  Finding relations between terms (synonymy and polysemy).
-Synonymy is the phenomenon where different words describe the same idea.
-Polysemy is the phenomenon where the same word has multiple meanings.
-*  Given a query of terms, translating it into the low-dimensional space, and find matching documents (information retrieval).
-*  Analyzing word association in text corpus.
-
-LSA overcomes two of the most problematic constraints of Boolean keyword queries: multiple words that have similar meanings (synonymy) and words that have more than one meaning (polysemy).
-
-LSA is also used to perform automated document categorization.
-Document categorization is the assignment of documents to one or more predefined categories based on their similarity to the conceptual content of the categories.
-Several experiments have demonstrated that there are a number of correlations between the way LSI and humans process and categorize text.
-
-LSA uses example documents to establish the conceptual basis for each category.
-During categorization processing, the concepts contained in the documents being categorized are compared to the concepts contained in the example items, and a category (or categories) is assigned to the documents based on the similarities between the concepts they contain and the concepts that are contained in the example documents.
-
-LSA automatically adapts to new and changing terminology, and has been shown to be very tolerant of noise (i.e., misspelled words, typographical errors, unreadable characters, etc.).[24] This is especially important for applications using text derived from Optical Character Recognition (OCR) and speech-to-text conversion.
-LSA also deals effectively with sparse, ambiguous, and contradictory data.
-
-Text does not need to be in sentence form for LSA to be effective.
-It can work with lists, free-form notes, email, Web-based content, etc.
-As long as a collection of text contains multiple terms, LSA can be used to identify patterns in the relationships between the important terms and concepts contained in the text.
-
-##Sub-Problem Three: Choosing a Text Classifier Method
-The many methods for classifying texts fall into three general categories:
-
-*  Pattern matching strategies (does a particular phrase appear?)
-*  Algorithms (how closely does the query text meet pre-set criteria?)
-*  Neural networks (the network decides its own criteria, aka machine learning)
-
-This edited excerpt nicely describes the overall approach and problems of text classification. It is from: **xxxxx**
-
-Understanding how chatbots work can help us understand text classifiers, because the classifier is the fundamental piece that makes the chatbot work.
-Let’s look at the inner workings of a Multinomial Naive Bayes algorithm for text classification and natural language processing (NLP).
-
-This classifier is “naive” because each word is treated as having no connection with other words in the sentence being classified.
-In other words it assumes independence between "features," in this case, words.
-
-For example, in the sentence “the fox jumped over the log”, there is no relationship between “jumped” and “fox” or “log”.
-In NLP, this collection of words is referred to as “a bag of words”.
-While this is incredibly naive, the classifier isn’t attempting to understand the meaning of a sentence, it’s trying to classify it.
-
-###Pattern Matchers
-Early chatbots used **pattern matching** to classify text and produce a response.
-This is often referred to as “brute force” as the author of the system needs to describe every pattern for which there is a response.
-
-A standard structure for these patterns is “AIML” (artificial intelligence markup language). Its use of the term “artificial intelligence” is quite an embellishment, but that’s another story.
-
-A simple pattern matching definition:
-```
-<aiml version = "1.0.1" encoding = "UTF-8"?>
-   <category>
-      <pattern> WHO IS ALBERT EINSTEIN </pattern>
-      <template>Albert Einstein was a German physicist.</template>
-   </category>
-
-   <category>
-      <pattern> WHO IS Isaac NEWTON </pattern>
-      <template>Isaac Newton was a English physicist and mathematician.</template>
-   </category>
-
-   <category>
-      <pattern>DO YOU KNOW WHO * IS</pattern>
-      <template>
-         <srai>WHO IS <star/></srai>
-      </template>
-   </category>
-</aiml>
-```
-
-The machine then produces:
-
-> Human: Do you know who Albert Einstein is
-> Robot: Albert Einstein was a German physicist.
-
-It knows who a physicist is only because his or her name has an associated pattern.
-Likewise it responds to anything solely because of an authored pattern.
-Given hundreds or thousands of patterns you might see a chatbot “persona” emerge.
-
-In 2000 a chatbot built by John Denning and colleagues using this approach was in the news for passing the "Turing test." It emulated the replies of a 13 year old boy from Ukraine (broken English and all).
-I met with John in 2015 and he made no false pretenses about the internal workings of this automaton.
-It may have been “brute force” but it proved a point: parts of a conversation can be made to appear “natural” using a sufficiently large definition of patterns.
-It proved Alan Turing’s assertion, that this question of a machine fooling humans was “meaningless”.
-
-An example of this approach used in building chatbots is PandoraBots, who claim over 285k chatbots have been constructed using their framework.
-
-Automated phone voice response systems that recognize patterns of spoken words and direct calls accordingly use a similar approach.
-
-###Algorithm-Based Classification
-Creating a brute-force pattern matcher is daunting: for each unique input a pattern must be available to specify a response. An algorithm-based method approaches the problem mathematically.
-
-One text classification algorithm is "Multinomial Naive Bayes." Given a training set of sentences, each belonging to a class (that is, a particular group), we can count the occurrence of each word in each class, account for its commonality and assign each word a score for each class.
-Factoring for commonality is important: matching the word “it” is considerably less meaningful than a match for the word “cheese”.
-When we input a new sentence, scores for each word in the new sentence are compared to scores for those same words in the training set.
-The class with the highest OVERALL score for the input sentence is the class to which the input sentence most likely belongs.
-
-Look at this sample training set:
-```
-class: weather
-    "is it nice outside?"
-    "how is it outside?"
-    "is the weather nice?"
-
-class: greeting
-    "how are you?"
-    "hello there"
-    "how is it going?"
-```
-
-Let’s classify a few sample input sentences:
-```
-input: "Hi there"
- term: "hi" (no matches in either class)
- term: "there" (matches a word in class: greeting)
- classification: greeting (score=1)
-
-input: "What’s it like outside?"
- term: "it" (matches words in class: weather (2), greeting(1))
- term: "outside (class: weather (2) )
- classification: weather (score=4)
-```
-
-Notice that the classification for "What’s it like outside" found a term in the "greeting" class but the term similarities to the "weather" class produced a higher score.
-By using an equation we are looking for word matches given some sample sentences for each class, and we avoid having to identify every pattern.
-
-The classification score produced identifies the class with the highest term matches (accounting for commonality of words) but this has limitations.
-A score is not the same as a probability, a score tells us which intent is most like the sentence but not the likelihood of it being a match.
-Thus it is difficult to apply a threshold for which classification scores to accept or not.
-Having the highest score from this type of algorithm only provides a relative basis, it may still be an inherently weak classification.
-Also the algorithm doesn’t account for what a sentence is not, it only counts what it is like.
-You might say this approach doesn’t consider what makes a sentence not a given class.
-
-Many text and spoken word frameworks use similar algorithms such as this to classify intent.
-Most of what’s taking place is word counting against training datasets.
-It’s "naive" but surprisingly effective.
-
-I will be using an algorithm-based classification strategy for this project.
-
-###Neural Networks
-These methods are beyond the scope of this project. Anyone interested in knowing more can search online for information about:
-
-*  Shallow Neural Networks
-*  Deep Neural Networks
-	+  Convolutional Neural Network (CNN)
-	+  Long Short Term Modelr (LSTM)
-	+  Gated Recurrent Unit (GRU)
-	+  Bidirectional RNN
-	+  Recurrent Convolutional Neural Network (RCNN)
-	+  Other Variants of Deep Neural Networks
-
-###Designing an Evaluation Workflow
-The workflow for comparing classification algorithms starts with identifying potentially usable classifier methods.
-Several approaches are described later.
-For each potential classifier method, the next steps are:
-
-*  Establish randomized training and test datasets
-*  Train the classifier algorithm, then evaluate that model on a test dataset
-*  Calculate an accuracy score using Gini index or similar method
-*  Select best performing classifiers and either refine, or operationalize for new datasets
-
-###Picking the Algorithm(s)
-Algorithms approach text classification several ways.
-
-The **bag of keywords approach** requires compiling a list of "key terms" that qualifies the type of content in question to a certain topic.
-If one or more qualifying keywords is present in a document, the document is assigned to the topic.
-One of the many problems with this approach is that identifying and organizing a list of terms (preferred, alternate, etc.) is labor intensive (much like simple pattern matching), and the natural ambiguity of language (one keyword can have different meanings) causes lots of false positives.
-This makes the bag of keywords approach not only not scalable, but also inaccurate.
-
-**Rules-based approaches** use linguistic rules that capture all of the elements and attributes of a document to assign it to a category.
-Rules can be written manually or generated by automatic analysis and then validated manually (reducing manual effort up to 90%).
-Rules can be understood and improved (unlike the black box system used with statistics-based algorithms).
-A rules-based approach is flexible, powerful (much more than a bag of keywords) and easy to express.
-This approach works best when combined with semantic technology that can uncover deeper connections within text (meaning, relevancy, relationship between concepts, etc.)
-
-**Statistical approaches** use a “training set” of documents that covers the same topic to identify key features (usually words).
-The algorithm (Bayesian, LSA, etc.) calculates term frequencies, then builds implicit rules in order to classify other content.
-This approach has no understanding of meaning.
-In addition, systems using these types of text classification algorithms are essentially a “black box”; no one can explain why specific terms are selected by the algorithm or how they are being weighted.
-In the event the classification is incorrect, there is no accurate way to modify a rule for better results.
-
-Algorithms also can be separated into un-supervised and supervised methods.
-
-**Unsupervised learning** uses test data that has not been labeled, classified or categorized.
-Unsupervised learning identifies commonalities in the training data, then classifies new texts based on the presence or absence of such commonalities.
-
-Unsupervised learning algorithms appearing frequently in literature are:
-
-*  Clustering
-	+  Hierarchical clustering
-	+  k-means
-	+  Mixture models
-*  Principal component analysis (PCA)
-
-**Supervised learning** uses labeled training data to defined important text features.
-In supervised learning, each example is a pair consisting of an input object (typically a vector) and a desired output value.
-A supervised learning algorithm analyzes the training data and produces an inferred function, which can be used for mapping new examples.
-.
-
-Supervised text classification algorithms appearing regularly in the literature are:
-
-*  Naive Bayes
-	+  Binary independent NB
-	+  Multinomial NB
-*  Linear Classifier
-*  Logistic Regression
-*  Support vector machines
-	+  Basic SVM
-	+  SVM multiple classes
-	+  SVM multiple kernels
-*  Decision trees
-*  Linear Discriminant Analysis
-*  Nearest neighbor
-	+  k-Nearest Neighbor
-	+  Weighted k-Nearest Neighbor
-*  Similarity Learning
-*  Relevance feedback
-	+  Rocchio
-	+  Rocchio in a query zone
-*  Ensemble methods
-	+  Bagging based Models
-	+  Stacking based Models
-	+  Boosting based Models
-	+  Random Forest Models
-
-For reference I assembled short excerpts from online resources or published articles describing most of these methods.
-
-###Naive Bayes
-A classification technique based on Bayes’ Theorem with an assumption of independence among predictors.
-A Naive Bayes classifier assumes that the presence of a particular feature in a class is unrelated to the presence of any other feature here.
-NB can be run on:
-
-*  Word count or n-gram count vectors
-*  Word or n-gram level TF-IDF vectors
-*  Character Level TF-IDF vectors
-	+  Would be useful for evaluating use of "?" and "="
-
-Data are organized into 2 dictionaries: **corpus words** (each stemmed word and the # of occurances) and **class words** (each class and the list of stemmed words within it).
-The algorithm then uses these data structures to classify new text.
-
-###Linear Classifier
-The linear regression algorithm is one of the fundamental supervised machine-learning algorithms due to its relative simplicity and well-known properties.
-
-The goal of statistical classification is to use an object's characteristics to identify which class (or group) it belongs to.
-A linear classifier achieves this by making a classification decision based on the value of a linear combination of the characteristics.
-An object's characteristics are also known as feature values and are typically presented to the machine in a vector called a feature vector.
-Such classifiers work well for practical problems such as document classification, and more generally for problems with many variables (features), reaching accuracy levels comparable to non-linear classifiers while taking less time to train and use.
-
-###Logistic Regression
-A logistic regression measures the relationship between the categorical dependent variable and one or more independent variables by estimating probabilities using a logistic/sigmoid function.
-LR can be run on:
-
-*  Word count or n-gram count vectors
-*  Word or n-gram level TF-IDF vectors
-*  Character Level TF-IDF vectors
-
-Resources for Logistic Regression 
-https://www.analyticsvidhya.com/blog/2015/10/basics-logistic-regression/
-
-R for logistic regression
-https://www.saedsayad.com/logistic_regression.htm
-
-###Support Vector Machine (SVM) Model
-This is a supervised machine learning algorithm which can be used for both classification or regression challenges.
-The model extracts a best possible hyper-plane / line that segregates the two classes.
-The only example I found of SVM was on Ngram Level TF-IDF Vectors.
-Support vector machines can be:
-
-*  Basic SVM
-*  SVM multiple classes
-*  SVM multiple kernels
-
-###Decision Trees
-Flowchart-like structures in which each internal node represents a "test" on an attribute (e.g.
-whether a coin flip comes up heads or tails), each branch represents the outcome of the test, and each leaf node represents a class label (decision taken after computing all attributes).
-The paths from root to leaf represent classification rules.
-
-In decision analysis, a decision tree and the closely related influence diagram are used as a visual and analytical decision support tool, where the expected values (or expected utility) of competing alternatives are calculated.
-
-A decision tree consists of three types of nodes:
-
-*  Decision nodes – typically represented by squares
-*  Chance nodes – typically represented by circles
-*  End nodes – typically represented by triangles
-
-Decision tree learning is the construction of a decision tree from class-labeled training tuples.
-A decision tree is a flow-chart-like structure, where each internal (non-leaf) node denotes a test on an attribute, each branch represents the outcome of a test, and each leaf (or terminal) node holds a class label.
-The topmost node in a tree is the root node.
-
-The goal is to create a model that predicts the value of a target variable based on several input variables.
-An example is shown in the diagram at right.
-Each interior node corresponds to one of the input variables; there are edges to children for each of the possible values of that input variable.
-Each leaf represents a value of the target variable given the values of the input variables represented by the path from the root to the leaf.
-
-A decision tree is a simple representation for classifying examples.
-For this section, assume that all of the input features have finite discrete domains, and there is a single target feature called the "classification".
-Each element of the domain of the classification is called a class.
-A decision tree or a classification tree is a tree in which each internal (non-leaf) node is labeled with an input feature.
-The arcs coming from a node labeled with an input feature are labeled with each of the possible values of the target or output feature or the arc leads to a subordinate decision node on a different input feature.
-Each leaf of the tree is labeled with a class or a probability distribution over the classes.
-
-A tree can be "learned" by splitting the source set into subsets based on an attribute value test.
-This process is repeated on each derived subset in a recursive manner called recursive partitioning.
-The recursion is completed when the subset at a node has all the same value of the target variable, or when splitting no longer adds value to the predictions.
-This process of top-down induction of decision trees (TDIDT) is an example of a greedy algorithm, and it is by far the most common strategy for learning decision trees from data.
-
-In data mining, decision trees can be described also as the combination of mathematical and computational techniques to aid the description, categorization and generalization of a given set of data.
-
-###Linear Discriminant Analysis
-LDA works when the measurements made on independent variables for each observation are continuous quantities.
-When dealing with categorical independent variables, the equivalent technique is discriminant correspondence analysis.
-The statistical system R includes the packages: MASS, ade4, ca, vegan, ExPosition, andFactoMineR which perform correspondence analysis and multiple correspondence analysis.
-
-As the name indicates, discriminant correspondence analysis (DCA) is an extension of discriminant analysis (DA) and correspondence analysis (CA).
-Like discriminant analysis, the goal of DCA is to categorize observations in pre-defined groups, and like  correspondence analysis, it is used with nominal variables.The main idea behind DCA is to represent each group by thesum of its observations and to perform a simple CA on the groupsby variables matrix.
-The original observations are then projectedas  supplementary  elements and each observation is assigned to the closest group.
-The comparison between the a priori and the a posteriori classifications can be used to assess the quality of thendiscrimination.
-A similar procedure can be used to assign new observations to categories.
-The stability of the analysis can be evaluated using cross-validation techniques such as jackknifing or boot-strapping.
-
-For the R model: https://pbil.univ-lyon1.fr/ade4/ade4-html/discrimin.coa.html
-
-###k-Nearest Neighbors
-k-NN is a non-parametric method used for classification and regression.
-In both cases, the input consists of the k closest training examples in the feature space.
-The output depends on whether k-NN is used for classification or regression.
-In k-NN classification, the output is a class membership.
-An object is classified by a majority vote of its neighbors, with the object being assigned to the class most common among its k nearest neighbors (k is a positive integer, typically small).
-If k = 1, then the object is simply assigned to the class of that single nearest neighbor.
-
-###Similarity Learning
-Similarity learning is an area of supervised machine learning in artificial intelligence.
-It is closely related to regression and classification, but the goal is to learn from examples a similarity function that measures how similar or related two objects are.
-It has applications in ranking, in recommendation systems, visual identity tracking, face verification, and speaker verification.
-
-###Relevance Feedback/Rocchio Algorithm
-Relevance feedback is a feature of some information retrieval systems.
-The idea behind relevance feedback is to take the results that are initially returned from a given query, to gather user feedback, and to use information about whether or not those results are relevant to perform a new query.
-We can usefully distinguish between three types of feedback: explicit feedback, implicit feedback, and blind or "pseudo" feedback.
-
-The Rocchio algorithm is based on a method of relevance feedback.
-Like many other retrieval systems, the Rocchio feedback approach was developed using the Vector Space Model.
-The algorithm is based on the assumption that most users have a general conception of which documents should be denoted as relevant or non-relevant.
-Therefore, the user's search query is revised to include an arbitrary percentage of relevant and non-relevant documents as a means of increasing the search engine's recall, and possibly the precision as well.
-
-For more:
-https://nlp.stanford.edu/IR-book/html/htmledition/rocchio-classification-1.html
-
-###Ensemble Methods
-The excerpt below nice describes how ensemble methods work. It is reprinted from: *Ensemble Methods Explained in Simple English*, by Tavish Srivastava, posted August 2, 2015.
-
-Ensemble modeling is a powerful way to improve the performance of your model. It usually pays off to apply ensemble learning over and above various models you might be building.
-
-Example: I want to invest in a company XYZ.
-I am not sure about its performance though.
-So, I look for advice on whether the stock price will increase more than 6% per annum or not?
-I decide to approach various experts having diverse domain experience:
-
->1.
-Employee of Company XYZ: This person knows the internal functionality of the company and have the insider information about the functionality of the firm.
-But he lacks a broader perspective on how are competitors innovating, how is the technology evolving and what will be the impact of this evolution on Company XYZ’s product.
-In the past, he has been right 70% times.
-
->1.
-Financial Advisor of Company XYZ: This person has a broader perspective on how companies strategy will fair of in this competitive environment.
-However, he lacks a view on how the company’s internal policies are fairing off.
-In the past, he has been right 75% times.
-
->1.
-Stock Market Trader: This person has observed the company’s stock price over past 3 years.
-He knows the seasonality trends and how the overall market is performing.
-He also has developed a strong intuition on how stocks might vary over time.
-In the past, he has been right 70% times.
-
->1.
-Employee of a competitor: This person knows the internal functionality of the competitor firms and is aware of certain changes which are yet to be brought.
-He lacks a sight of company in focus and the external factors which can relate the growth of competitor with the company of subject.
-In the past, he has been right 60% of times.
-
->1.
-Market Research team in same segment: This team analyzes the customer preference of company XYZ’s product over others and how is this changing with time.
-Because he deals with customer side, he is unaware of the changes company XYZ will bring because of alignment to its own goals.
-In the past, they have been right 75% of times.
-
->1.
-Social Media Expert: This person can help us understand how has company XYZ has positioned its products in the market.
-And how are the sentiment of customers changing over time towards company.
-He is unaware of any kind of details beyond digital marketing.
-In the past, he has been right 65% of times.
-
-Given the broad spectrum of access we have, we can probably combine all the information and make an informed decision.
-In a scenario when all the 6 experts/teams verify that it’s a good decision(assuming all the predictions are independent of each other), we will get a combined accuracy rate of:
-
-```
-		  1 - 30%*25%*30%*40%*25%*35%
-
-		= 1 - 0.07875 = 99.92125%
-```
-
-Assumption: The assumption used here that all the predictions are completely independent is slightly extreme as they are expected to be correlated.
-However, we see how we can be so sure by combining various predictions together.
-
-Let us now change the scenario slightly. This time we have 6 experts, all of them are employee of company XYZ working in the same division. Everyone has a propensity of 70% to advocate correctly.
-
->*What if we combine all this advice together, can we still raise up our confidence to >99%?*
-
-Obviously not, as all the predictions are based on very similar set of information.
-They are certain to be influenced by similar set of information and the only variation in their advice would be due to their personal opinions & collected facts about the firm.
-
-Ensemble learning is the art of combining individual models together to improve the stability and predictive power of the model.
-
-####Error Sources in Ensemble Learning (Variance vs. Bias)
-The error emerging from any model can be broken down into three components mathematically.
-
-**Error of a model** Why is this important in the current context? To understand what really goes behind an ensemble model, we need to first understand what causes error in the model.
-
-**Bias error** quantifies how much on an average the predicted values differ from the actual value. A high bias error means we have a under-performing model which keeps on missing important trends.
-
-**Variance** quantifies how much the predictions made on same observations differ from each other.
-A high variance model will over-fit on your training population and perform badly on any observation beyond training.
-
-Normally, as you increase the complexity of your model, you will see a reduction in error due to lower bias in the model.
-However, this only happens till a particular point.
-As you continue to make your model more complex, you end up **over-fitting your model** and hence your model will start suffering from high variance.
-
-A champion model should maintain a balance between these two types of errors.
-This is known as the **trade-off management of bias-variance errors.** Ensemble learning is one way to execute this trade off analysis.
-
-####Commonly used Ensemble learning techniques
-**Bagging** tries to implement similar learners on small sample populations and then takes a mean of all the predictions.
-In generalized bagging, you can use different learners on different population.
-As you can expect this helps us to reduce the variance error.
-
-**Boosting** is an iterative technique which adjust the weight of an observation based on the last classification.
-If an observation was classified incorrectly, it tries to increase the weight of this observation and vice versa.
-Boosting in general decreases the bias error and builds strong predictive models.
-However, they may sometimes over fit on the training data.
-
-**Stacking** is a very interesting way of combining models.
-Here we use a learner to combine output from different learners.
-This can lead to decrease in either bias or variance error depending on the combining learner we use.
-
-####Random Forest Models
-Random Forest models are a type of ensemble models, particularly bagging models. They are part of the tree based model family. RF can be run on:
-
-*  Word count or n-gram count vectors
-*  Word level TF-IDF vectors
-
-Resources for Random Forest: 
-https://www.analyticsvidhya.com/blog/2014/06/introduction-random-forest-simplified/
-
-####Extreme Gradient Boosting Model
-Boosting models are another type of ensemble models part of tree based models.
-Boosting is a machine learning ensemble meta-algorithm for primarily reducing bias, and also variance in supervised learning, and a family of machine learning algorithms that convert weak learners to strong ones.
-A weak learner is defined to be a classifier that is only slightly correlated with the true classification (it can label examples better than random guessing).
-XGB can be run on:
-
-*  Word count or n-gram count vectors
-*  Word level TF-IDF vectors
-*  Character Level TF-IDF vectors
-
-###Hierarchical Clustering
-In data mining and statistics, hierarchical clustering (also called hierarchical cluster analysis or HCA) is a method of cluster analysis which seeks to build a hierarchy of clusters.
-
-###k-Means Clustering
-k-means clustering is a method of vector quantization, originally from signal processing, that is popular for cluster analysis in data mining.
-k-means clustering aims to partition n observations into k clusters in which each observation belongs to the cluster with the nearest mean, serving as a prototype of the cluster.
-This results in a partitioning of the data space into Voronoi cells.
-
-The problem is computationally difficult (NP-hard); however, efficient heuristic algorithms converge quickly to a local optimum.
-These are usually similar to the expectation-maximization algorithm for mixtures of Gaussian distributions via an iterative refinement approach employed by both k-means and Gaussian mixture modeling.
-They both use cluster centers to model the data; however, k-means clustering tends to find clusters of comparable spatial extent, while the expectation-maximization mechanism allows clusters to have different shapes.
-
-The algorithm has a loose relationship to the k-nearest neighbor classifier, a popular machine learning technique for classification that is often confused with k-means due to the name.
-Applying the 1-nearest neighbor classifier the cluster centers obtained by k-means classifies new data into the existing clusters.
-This is known as nearest centroid classifier or Rocchio algorithm.
-
-###Mixture Model
-In statistics, a mixture model is a probabilistic model for representing the presence of subpopulations within an overall population, without requiring that an observed data set should identify the sub-population to which an individual observation belongs.
-Formally a mixture model corresponds to the mixture distribution that represents the probability distribution of observations in the overall population.
-However, while problems associated with "mixture distributions" relate to deriving the properties of the overall population from those of the sub-populations, "mixture models" are used to make statistical inferences about the properties of the sub-populations given only observations on the pooled population, without sub-population identity information.
-
-Topics in a document: Assume that a document is composed of N different words from a total vocabulary of size V, where each word corresponds to one of K possible topics.
-The distribution of such words could be modelled as a mixture of K different V-dimensional categorical distributions.
-A model of this sort is commonly termed a topic model.
-Note that expectation maximization applied to such a model will typically fail to produce realistic results, due (among other things) to the excessive number of parameters.
-Some sorts of additional assumptions are typically necessary to get good results.
-
-A prior distribution is placed over the parameters describing the topic distributions, using a Dirichlet distribution with a concentration parameter that is set significantly below 1, so as to encourage sparse distributions (where only a small number of words have significantly non-zero probabilities).
-
-Some sort of additional constraint is placed over the topic identities of words, to take advantage of natural clustering.
-
-For example, a Markov chain could be placed on the topic identities (i.e., the latent variables specifying the mixture component of each observation), corresponding to the fact that nearby words belong to similar topics.
-(This results in a hidden Markov model, specifically one where a prior distribution is placed over state transitions that favors transitions that stay in the same state.)
-
-Another possibility is the latent Dirichlet allocation model, which divides up the words into D different documents and assumes that in each document only a small number of topics occur with any frequency.
-
-###Principal Component Analysis 
-PCA is a statistical procedure that uses an orthogonal transformation to convert a set of observations of possibly correlated variables (entities each of which takes on various numerical values) into a set of values of linearly uncorrelated variables called principal components.
-This transformation is defined in such a way that the first principal component has the largest possible variance (that is, accounts for as much of the variability in the data as possible), and each succeeding component in turn has the highest variance possible under the constraint that it is orthogonal to the preceding components.
-The resulting vectors (each being a linear combination of the variables and containing n observations) are an uncorrelated orthogonal basis set.
-PCA is sensitive to the relative scaling of the original variables.
-
-PCA is mostly used as a tool in exploratory data analysis and for making predictive models.
-It is often used to visualize genetic distance and relatedness between populations.
-PCA can be done by eigenvalue decomposition of a data covariance (or correlation) matrix or singular value decomposition of a data matrix, usually after a normalization step of the initial data.
-The normalization of each attribute consists of mean centering – subtracting each data value from its variable's measured mean so that its empirical mean (average) is zero – and, possibly, normalizing each variable's variance to make it equal to 1.
-The results of a PCA are usually discussed in terms of component scores, sometimes called factor scores (the transformed variable values corresponding to a particular data point), and loadings (the weight by which each standardized original variable should be multiplied to get the component score).
-If component scores are standardized to unit variance loadings must contain the data variance in them (and that is the magnitude of eigenvalues).
-If component scores are not standardized (therefore they contain the data variance) then loadings must be unit-scaled, ("normalized") and these weights are called eigenvectors; they are the cosines of orthogonal rotation of variables into principal components or back.
-
-PCA is the simplest of the true eigenvector-based multivariate analyses.
-Often, its operation can be thought of as revealing the internal structure of the data in a way that best explains the variance in the data.
-If a multivariate dataset is visualised as a set of coordinates in a high-dimensional data space (1 axis per variable), PCA can supply the user with a lower-dimensional picture, a projection of this object when viewed from its most informative viewpoint.
-This is done by using only the first few principal components so that the dimensionality of the transformed data is reduced.
-
-PCA is closely related to factor analysis.
-Factor analysis typically incorporates more domain specific assumptions about the underlying structure and solves eigenvectors of a slightly different matrix.
-
-PCA is also related to canonical correlation analysis (CCA).
-CCA defines coordinate systems that optimally describe the cross-covariance between two datasets while PCA defines a new orthogonal coordinate system that optimally describes variance in a single dataset.
-
-##Strategies for Improving Text Classification Models
+#Improving Text Classification
 Accuracy can be improved using one or more of these strategies.
@@
 
-***
-***
-
-###Neural Networks
-A neural network is a mathematical model that is designed to behave similar to biological neurons and nervous system.
-These models are used to recognize complex patterns and relationships that exists within a labelled data.
-
-Artificial neural networks, invented in the 1940’s, are a way of calculating an output from an input (a classification) using weighted connections (“synapses”) that are calculated from repeated iterations through training data.
-Each pass through the training data alters the weights such that the neural network produces the output with greater “accuracy” (lower error rate).
-
-There’s not much new about these structures, except today’s software is using much faster processors and can work with a lot more memory.
-The combination of working memory and speed is crucial when you’re doing hundreds of thousands of matrix multiplications (the neural network’s essential math operation).
-
-As in the prior method, each class is given with some number of example sentences.
-Once again each sentence is broken down by word (stemmed) and each word becomes an input for the neural network.
-The synaptic weights are then calculated by iterating through the training data thousands of times, each time adjusting the weights slightly to greater accuracy.
-By recalculating back across multiple layers (“back-propagation”) the weights of all synapses are calibrated while the results are compared to the training data output.
-These weights are like a ‘strength’ measure, in a neuron the synaptic weight is what causes something to be more memorable than not.
-You remember a thing more because you’ve seen it more times: each time the ‘weight’ increases slightly.
-
-At some point the adjustment reaches a point of diminishing returns, this is called “over-fitting” and going beyond this is counter-productive.
-
-The trained neural network is less code than an comparable algorithm but it requires a potentially large matrix of “weights”.
-In a relatively small sample, where the training sentences have 150 unique words and 30 classes this would be a matrix of 150x30.
-Imagine multiplying a matrix of this size 100,000 times to establish a sufficiently low error rate.
-This is where processing speed comes in.
-
-If the neural network sounds magnificently sophisticated, relax, it boils down to matrix multiplication and a formula for reducing values between -1 and 1 or some other minimal range.
-A middle-school math student could learn this in a few hours.
-The hard work is achieving clean training data.
-
-Just as there are variations in pattern matching code and in algorithms, there are variations in neural networks, some more complex than others.
-The basic machinery is the same.
-The essential work is that of classification.
-
-Networks are looking for patterns in collections of terms, each term is reduced to a token.
-In this machine words have no meaning except for their patterned existence within training data.
-The label “artificial intelligence” applied to such machinery is mostly BS.
-
-###Types of Neural Networks
-A shallow neural network contains mainly three types of layers – input layer, hidden layer, and output layer.
-
-Deep Neural Networks are more complex neural networks in which the hidden layers performs much more complex operations than simple sigmoid or relu activations.
-Different types of deep learning models can be applied in text classification problems.
-
-####Convolutional Neural Network
-In Convolutional neural networks, convolutions over the input layer are used to compute the output.
-This results in local connections, where each region of the input is connected to a neuron in the output.
-Each layer applies different filters and combines their results.
-
-####Recurrent Neural Network: LSTM
-Unlike Feed-forward neural networks in which activation outputs are propagated only in one direction, the activation outputs from neurons propagate in both directions (from inputs to outputs and from outputs to inputs) in Recurrent Neural Networks.
-This creates loops in the neural network architecture which acts as a ‘memory state’ of the neurons.
-This state allows the neurons an ability to remember what have been learned so far.
-
-The memory state in RNNs gives an advantage over traditional neural networks but a problem called Vanishing Gradient is associated with them.
-In this problem, while learning with a large number of layers, it becomes really hard for the network to learn and tune the parameters of the earlier layers.
-To address this problem, A new type of RNNs called LSTMs (Long Short Term Memory) Models have been developed.
-
-####Recurrent Neural Network: GRU
-Gated Recurrent Units are another form of recurrent neural networks.
-
-####Bidirectional RNN
-RNN layers can be wrapped in Bidirectional layers as well.
-
-####Recurrent Convolutional Neural Network
-Once the essential architectures have been tried out, one can try different variants of these layers such as recurrent convolutional neural network. Another variants can be:
-
-*  Hierarchical Attention Networks
-*  Sequence to Sequence Models with Attention
-*  Bidirectional Recurrent Convolutional Neural Networks
-*  CNNs and RNNs with more number of layers
+####\

### 2019-02-25 Topic_categorization.Rmd
updates

--- 2019-02-23 Text Classifier Models.Rmd
@@
 ---
-title: "Text Classification Primer"
-author: "Dan Johnson"
-date: "1/9/2019"
-output: html_document
+title: "Topic Categorization and Modeling"
+date: "2/21/2019"
+bibliography: library.bib
+biblio-style: apalike
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, child="_setup.Rmd"}
 ```
 
-For this text classification project I am using an initial codebook and corpus of hand-coded TA comments to test various approaches to semi- and fully automated classification methods and, if appropriate, revise the underlying classification scheme.
-
-This page provides background on general mechanics of text classification and various classifier methods.
-Other pages describe the dataset structure [LINK] (TA comments corpus), and [LINK] development of the initial codebook.
-
-#Workflow
-Text classification problems can be broken down into four overlapping sub-problems. Each piece may be more or less important to the overall problem, but all need to be considered:
-
-1.  Feature selection and engineering
-2.  Topic categorization and modeling
-3.  Selecting an appropriate text classifier method
-4.  Evaluating the outcomes
-
-##Feature Selection and Engineering
-This is the process of selecting, tabulating, calculating, or creating text-based features for subsequent evaluation. This includes:
-
-*  Identifying existing, directly observable text features within the dataset
-*  Engineering additional features using established NLP methods
-*	 Reverse engineering relevant features based on an empirically defined feature set.
-
-The features listed below are used regularly for natural language processing.
-Not all will be appropriate for all situations though; selecting, evaluating, comparing, and validating the appropriate features for the classification task is an iterative process, not a one-time decision.
-
-*  Count-based Vectors
-	+  Word Count – total number of words in the documents
-	+  Character Count – total number of characters in the documents
-	+  Punctuation Count – total number of punctuation marks in the documents
-	+  Upper Case Count – total number of upper case words in the documents
-	+  Title Word Count – total number of consecutive upper case (title) words in the documents
-*  Distribution and Frequency Calculations
-	+  Average word density of sentences 
-	+  Average length of the words used in the documents
-	+  TF-IDF
-	+  Reading level estimates (described below)
-*  Part of Speech Tagging and Frequency Distribution of Tags 
-	+ Noun Count
-	+ Verb Count
-	+ Adjective Count
-	+ Adverb Count
-	+ Pronoun Count
-*  Word Embedding (described below)
-
-Additional features are available for longer texts such as multi-sentence paragraphs.
-
-*  Word choices
-*  Lexical diversity (described below)
-*  Stance patterns - hedging, booster, generalization words or phrasing 
-*  Concision - comparisons of noun phrase density versus comment clause density
-*  Cohesion - use of words that are additive (also), countering (however), specify or emphasize (even, both, especially), connecting (former, latter)
-*  Civility and uncertainty - language that acknowledges limits of statement, critiques ideas not people, and makes fewer generalizations
-
-###Lexical Diversity
-This is not the same as word choices.
-This explanation is an excerpt from "quanteda: Quantitative Analysis of Textual Data", which describes how to use ```textstat_lexdiv``` to calculate lexical diversity.
-https://rdrr.io/cran/quanteda/man/textstat_lexdiv.html
-http://quanteda.io/reference/textstat_lexdiv.html
-
-Consider a speaker, who uses the term "allow" multiple times throughout the speech, compared to an another speaker who uses terms "allow, concur, acquiesce, accede, and avow"" for the same word.
-The latter speech has more lexical diversity than the former.
-Lexical diversity is widely believed to be an important parameter to rate a document in terms of textual richness and effectiveness.
-
-Lexical diversity, in simple terms, is a measurement of the breadth and variety of vocabulary used in a document.
-The different measures of lexical diversity are TTR, MSTTR, MATTR, C, R, CTTR, U, S, K, Maas, HD-D, MTLD, and MTLD-MA.
-
-koRpus package in R also provides functions to estimate lexical diversity or complexity.
-https://subscription.packtpub.com/book/big_data_and_business_intelligence/9781783551811/2/ch02lvl1sec16/lexical-diversity
-
-###Readability metrics
-These estimate the relative grade level or age of a typical reader who could comprehend 80+% (typically) of the text.
-Most readability metrics require longer blocks of text (100+ words) to achieve a stable score.
-They use some combination of word count, average syllables per word, and sentence length to estimate reading level.
-
-The metrics differ in how appropriate they are for specific situations.
-Some were developed for general writing, others for specific school age groups, and still others for technical documents; the US Navy in particular developed a suite of metrics for evaluating their training documents.
-
-The ```readability library``` (v0.1.1) is a collection of readability tools that utilize the ```syllable``` package for fast calculation of readability scores by grouping variables.
-It calculates Flesch Kincaid, Gunning Fog Index, Coleman Liau, SMOG, Automated Readability Index and an average of the 5 readability scores.
-These all are general scales; other scores may not be available in R packages.
-
-###Word Embedding
-**Word embedding** is a set of language modeling and feature learning techniques in natural language processing (NLP) where words or phrases from the vocabulary are mapped to vectors of real numbers.
-Conceptually it involves a mathematical embedding from a space with one dimension per word to a continuous vector space with a much lower dimension.
-
-Methods to generate the mapping include neural networks, dimensionality reduction, probabilistic models, knowledge base methods, and explicit representation in terms of the context in which words appear.
-The technique of representing words as vectors has roots in the 1960s with the development of the vector space model for information retrieval.
-
-Word embeddings come in two different styles, one in which words are expressed as vectors of co-occurring words, and another in which words are expressed as vectors of linguistic contexts in which the words occur.
-Most new word embedding techniques rely on a neural network architecture instead of more traditional n-gram models and unsupervised learning.
-
-One of the main limitations of word embeddings (word vector space models in general) is that possible meanings of a word are conflated into a single representation (a single vector in the semantic space).
-Sense embeddings are a solution to this problem: individual meanings of words are represented as distinct vectors in the space.
-
-Tools like **Principal Component Analysis (PCA)** and **T-Distributed Stochastic Neighbour Embedding (t-SNE)** are used to reduce the dimensionality of word vector spaces and visualize word embeddings and clusters.
-
-Software for training and using word embeddings includes Tomas Mikolov's Word2vec, Stanford University's GloVe, fastText, Gensim, AllenNLP's Elmo, Indra and Deeplearning4j.
-Three of these tools are described further below.
-
-####Word2vec
-Word2vec is a group of shallow, two-layer neural network models that produce word embeddings.
-The toolkit can train vector space models to reconstruct linguistic contexts of words.
-Word2vec takes as its input a large corpus of text and produces a vector space, typically of several hundred dimensions, with each unique word in the corpus being assigned a corresponding vector in the space.
-Word vectors are positioned in the vector space so words that share common contexts in the corpus are located in close proximity to one another in the space.
-
-The use of different model parameters and different corpus sizes can greatly affect the quality of a word2vec model.
-Accuracy can be improved in a number of ways, including the choice of model architecture (CBOW or Skip-Gram), increasing the training data set, increasing the number of vector dimensions, and increasing the window size of words considered by the algorithm.
-Each of these improvements comes with the cost of increased computational complexity and therefore increased model generation time.
-
-In models using large corpora and a high number of dimensions, the skip-gram model yields the highest overall accuracy, and consistently produces the highest accuracy on semantic relationships, as well as yielding the highest syntactic accuracy in most cases.
-However, the CBOW is less computationally expensive and yields similar accuracy results.
-
-Accuracy increases overall as the number of words used increases, and as the number of dimensions increases.
-Altszyler et al.
-(2017) studied Word2vec performance in two semantic tests for different corpus size.
-They found that Word2vec has a steep learning curve, outperforming another word-embedding technique (LSA) when it is trained with medium to large corpus size (more than 10 million words).
-However, with a small training corpus LSA showed better performance.
-Additionally they show that the best parameter setting depends on the task and the training corpus.
-Nevertheless, for skip-gram models trained in medium size corpora, with 50 dimensions, a window size of 15 and 10 negative samples seems to be a good parameter setting.
-
-####fastText
-fastText is a Python-based library created by Facebook's AI Research (FAIR) lab for learning of word embeddings and text classification.
-The model allows to create an unsupervised learning or supervised learning algorithm for obtaining vector representations for words.
-fastText uses a neural network for word embedding.
-
-https://fasttext.cc/
-
-####GloVe
-GloVe, short for Global Vectors, is a model for distributed word representation.
-The model is an unsupervised learning algorithm for obtaining vector representations for words.
-Training is performed on aggregated global word-word co-occurrence statistics from a corpus, and the resulting representations showcase interesting linear substructures of the word vector space.
-It is developed as an open-source project at Stanford.
-Pre-trained word vectors data are made available under the Public Domain Dedication and License v1.0.
-
-GloVe uses Euclidean distance (or cosine similarity) between two word vector for measuring the linguistic or semantic similarity of the corresponding words.
-Sometimes, the nearest neighbors according to this metric reveal rare but relevant words that lie outside an average human's vocabulary.
-For example, here are the closest words to the target word **frog**:
-
-*  frog
-*  frogs
-*  toad
-*  litoria
-*  leptodactylidae
-*  rana
-*  lizard
-*  eleutherodactylus
-
-The similarity metrics used for nearest neighbor evaluations produce a single scalar that quantifies the relatedness of two words.
-This simplicity can be problematic since two given words almost always exhibit more intricate relationships than can be captured by a single number.
-For example, man may be regarded as similar to woman in that both words describe human beings; on the other hand, the two words are often considered opposites since they highlight a primary axis along which humans differ from one another.
-
-In order to capture in a quantitative way the nuance necessary to distinguish man from woman, it is necessary for a model to associate more than a single number to the word pair.
-A natural and simple candidate for an enlarged set of discriminative numbers is the vector difference between the two word vectors.
-GloVe is designed in order that such vector differences capture as much as possible the meaning specified by the juxtaposition of two words.
-
-https://nlp.stanford.edu/projects/glove/
-
-##Sub-Problem Two: Topic Categorization and Modeling Methods
+#Topic Categorization and Modeling Methods
 Establishing meaningful text classification categories depends on the goals of the project or the question being asked.
@@
 
-##Sub-Problem Three: Choosing a Text Classifier Method
-The many methods for classifying texts fall into three general categories:
-
-*  Pattern matching strategies (does a particular phrase appear?)
-*  Algorithms (how closely does the query text meet pre-set criteria?)
-*  Neural networks (the network decides its own criteria, aka machine learning)
-
-This edited excerpt nicely describes the overall approach and problems of text classification. It is from: **xxxxx**
-
-Understanding how chatbots work can help us understand text classifiers, because the classifier is the fundamental piece that makes the chatbot work.
-Let’s look at the inner workings of a Multinomial Naive Bayes algorithm for text classification and natural language processing (NLP).
-
-This classifier is “naive” because each word is treated as having no connection with other words in the sentence being classified.
-In other words it assumes independence between "features," in this case, words.
-
-For example, in the sentence “the fox jumped over the log”, there is no relationship between “jumped” and “fox” or “log”.
-In NLP, this collection of words is referred to as “a bag of words”.
-While this is incredibly naive, the classifier isn’t attempting to understand the meaning of a sentence, it’s trying to classify it.
-
-###Pattern Matchers
-Early chatbots used **pattern matching** to classify text and produce a response.
-This is often referred to as “brute force” as the author of the system needs to describe every pattern for which there is a response.
-
-A standard structure for these patterns is “AIML” (artificial intelligence markup language). Its use of the term “artificial intelligence” is quite an embellishment, but that’s another story.
-
-A simple pattern matching definition:
-```
-<aiml version = "1.0.1" encoding = "UTF-8"?>
-   <category>
-      <pattern> WHO IS ALBERT EINSTEIN </pattern>
-      <template>Albert Einstein was a German physicist.</template>
-   </category>
-
-   <category>
-      <pattern> WHO IS Isaac NEWTON </pattern>
-      <template>Isaac Newton was a English physicist and mathematician.</template>
-   </category>
-
-   <category>
-      <pattern>DO YOU KNOW WHO * IS</pattern>
-      <template>
-         <srai>WHO IS <star/></srai>
-      </template>
-   </category>
-</aiml>
-```
-
-The machine then produces:
-
-> Human: Do you know who Albert Einstein is
-> Robot: Albert Einstein was a German physicist.
-
-It knows who a physicist is only because his or her name has an associated pattern.
-Likewise it responds to anything solely because of an authored pattern.
-Given hundreds or thousands of patterns you might see a chatbot “persona” emerge.
-
-In 2000 a chatbot built by John Denning and colleagues using this approach was in the news for passing the "Turing test." It emulated the replies of a 13 year old boy from Ukraine (broken English and all).
-I met with John in 2015 and he made no false pretenses about the internal workings of this automaton.
-It may have been “brute force” but it proved a point: parts of a conversation can be made to appear “natural” using a sufficiently large definition of patterns.
-It proved Alan Turing’s assertion, that this question of a machine fooling humans was “meaningless”.
-
-An example of this approach used in building chatbots is PandoraBots, who claim over 285k chatbots have been constructed using their framework.
-
-Automated phone voice response systems that recognize patterns of spoken words and direct calls accordingly use a similar approach.
-
-###Algorithm-Based Classification
-Creating a brute-force pattern matcher is daunting: for each unique input a pattern must be available to specify a response. An algorithm-based method approaches the problem mathematically.
-
-One text classification algorithm is "Multinomial Naive Bayes." Given a training set of sentences, each belonging to a class (that is, a particular group), we can count the occurrence of each word in each class, account for its commonality and assign each word a score for each class.
-Factoring for commonality is important: matching the word “it” is considerably less meaningful than a match for the word “cheese”.
-When we input a new sentence, scores for each word in the new sentence are compared to scores for those same words in the training set.
-The class with the highest OVERALL score for the input sentence is the class to which the input sentence most likely belongs.
-
-Look at this sample training set:
-```
-class: weather
-    "is it nice outside?"
-    "how is it outside?"
-    "is the weather nice?"
-
-class: greeting
-    "how are you?"
-    "hello there"
-    "how is it going?"
-```
-
-Let’s classify a few sample input sentences:
-```
-input: "Hi there"
- term: "hi" (no matches in either class)
- term: "there" (matches a word in class: greeting)
- classification: greeting (score=1)
-
-input: "What’s it like outside?"
- term: "it" (matches words in class: weather (2), greeting(1))
- term: "outside (class: weather (2) )
- classification: weather (score=4)
-```
-
-Notice that the classification for "What’s it like outside" found a term in the "greeting" class but the term similarities to the "weather" class produced a higher score.
-By using an equation we are looking for word matches given some sample sentences for each class, and we avoid having to identify every pattern.
-
-The classification score produced identifies the class with the highest term matches (accounting for commonality of words) but this has limitations.
-A score is not the same as a probability, a score tells us which intent is most like the sentence but not the likelihood of it being a match.
-Thus it is difficult to apply a threshold for which classification scores to accept or not.
-Having the highest score from this type of algorithm only provides a relative basis, it may still be an inherently weak classification.
-Also the algorithm doesn’t account for what a sentence is not, it only counts what it is like.
-You might say this approach doesn’t consider what makes a sentence not a given class.
-
-Many text and spoken word frameworks use similar algorithms such as this to classify intent.
-Most of what’s taking place is word counting against training datasets.
-It’s "naive" but surprisingly effective.
-
-I will be using an algorithm-based classification strategy for this project.
-
-###Neural Networks
-These methods are beyond the scope of this project. Anyone interested in knowing more can search online for information about:
-
-*  Shallow Neural Networks
-*  Deep Neural Networks
-	+  Convolutional Neural Network (CNN)
-	+  Long Short Term Modelr (LSTM)
-	+  Gated Recurrent Unit (GRU)
-	+  Bidirectional RNN
-	+  Recurrent Convolutional Neural Network (RCNN)
-	+  Other Variants of Deep Neural Networks
-
-###Designing an Evaluation Workflow
-The workflow for comparing classification algorithms starts with identifying potentially usable classifier methods.
-Several approaches are described later.
-For each potential classifier method, the next steps are:
-
-*  Establish randomized training and test datasets
-*  Train the classifier algorithm, then evaluate that model on a test dataset
-*  Calculate an accuracy score using Gini index or similar method
-*  Select best performing classifiers and either refine, or operationalize for new datasets
-
-###Picking the Algorithm(s)
-Algorithms approach text classification several ways.
-
-The **bag of keywords approach** requires compiling a list of "key terms" that qualifies the type of content in question to a certain topic.
-If one or more qualifying keywords is present in a document, the document is assigned to the topic.
-One of the many problems with this approach is that identifying and organizing a list of terms (preferred, alternate, etc.) is labor intensive (much like simple pattern matching), and the natural ambiguity of language (one keyword can have different meanings) causes lots of false positives.
-This makes the bag of keywords approach not only not scalable, but also inaccurate.
-
-**Rules-based approaches** use linguistic rules that capture all of the elements and attributes of a document to assign it to a category.
-Rules can be written manually or generated by automatic analysis and then validated manually (reducing manual effort up to 90%).
-Rules can be understood and improved (unlike the black box system used with statistics-based algorithms).
-A rules-based approach is flexible, powerful (much more than a bag of keywords) and easy to express.
-This approach works best when combined with semantic technology that can uncover deeper connections within text (meaning, relevancy, relationship between concepts, etc.)
-
-**Statistical approaches** use a “training set” of documents that covers the same topic to identify key features (usually words).
-The algorithm (Bayesian, LSA, etc.) calculates term frequencies, then builds implicit rules in order to classify other content.
-This approach has no understanding of meaning.
-In addition, systems using these types of text classification algorithms are essentially a “black box”; no one can explain why specific terms are selected by the algorithm or how they are being weighted.
-In the event the classification is incorrect, there is no accurate way to modify a rule for better results.
-
-Algorithms also can be separated into un-supervised and supervised methods.
-
-**Unsupervised learning** uses test data that has not been labeled, classified or categorized.
-Unsupervised learning identifies commonalities in the training data, then classifies new texts based on the presence or absence of such commonalities.
-
-Unsupervised learning algorithms appearing frequently in literature are:
-
-*  Clustering
-	+  Hierarchical clustering
-	+  k-means
-	+  Mixture models
-*  Principal component analysis (PCA)
-
-**Supervised learning** uses labeled training data to defined important text features.
-In supervised learning, each example is a pair consisting of an input object (typically a vector) and a desired output value.
-A supervised learning algorithm analyzes the training data and produces an inferred function, which can be used for mapping new examples.
-.
-
-Supervised text classification algorithms appearing regularly in the literature are:
-
-*  Naive Bayes
-	+  Binary independent NB
-	+  Multinomial NB
-*  Linear Classifier
-*  Logistic Regression
-*  Support vector machines
-	+  Basic SVM
-	+  SVM multiple classes
-	+  SVM multiple kernels
-*  Decision trees
-*  Linear Discriminant Analysis
-*  Nearest neighbor
-	+  k-Nearest Neighbor
-	+  Weighted k-Nearest Neighbor
-*  Similarity Learning
-*  Relevance feedback
-	+  Rocchio
-	+  Rocchio in a query zone
-*  Ensemble methods
-	+  Bagging based Models
-	+  Stacking based Models
-	+  Boosting based Models
-	+  Random Forest Models
-
-For reference I assembled short excerpts from online resources or published articles describing most of these methods.
-
-###Naive Bayes
-A classification technique based on Bayes’ Theorem with an assumption of independence among predictors.
-A Naive Bayes classifier assumes that the presence of a particular feature in a class is unrelated to the presence of any other feature here.
-NB can be run on:
-
-*  Word count or n-gram count vectors
-*  Word or n-gram level TF-IDF vectors
-*  Character Level TF-IDF vectors
-	+  Would be useful for evaluating use of "?" and "="
-
-Data are organized into 2 dictionaries: **corpus words** (each stemmed word and the # of occurances) and **class words** (each class and the list of stemmed words within it).
-The algorithm then uses these data structures to classify new text.
-
-###Linear Classifier
-The linear regression algorithm is one of the fundamental supervised machine-learning algorithms due to its relative simplicity and well-known properties.
-
-The goal of statistical classification is to use an object's characteristics to identify which class (or group) it belongs to.
-A linear classifier achieves this by making a classification decision based on the value of a linear combination of the characteristics.
-An object's characteristics are also known as feature values and are typically presented to the machine in a vector called a feature vector.
-Such classifiers work well for practical problems such as document classification, and more generally for problems with many variables (features), reaching accuracy levels comparable to non-linear classifiers while taking less time to train and use.
-
-###Logistic Regression
-A logistic regression measures the relationship between the categorical dependent variable and one or more independent variables by estimating probabilities using a logistic/sigmoid function.
-LR can be run on:
-
-*  Word count or n-gram count vectors
-*  Word or n-gram level TF-IDF vectors
-*  Character Level TF-IDF vectors
-
-Resources for Logistic Regression 
-https://www.analyticsvidhya.com/blog/2015/10/basics-logistic-regression/
-
-R for logistic regression
-https://www.saedsayad.com/logistic_regression.htm
-
-###Support Vector Machine (SVM) Model
-This is a supervised machine learning algorithm which can be used for both classification or regression challenges.
-The model extracts a best possible hyper-plane / line that segregates the two classes.
-The only example I found of SVM was on Ngram Level TF-IDF Vectors.
-Support vector machines can be:
-
-*  Basic SVM
-*  SVM multiple classes
-*  SVM multiple kernels
-
-###Decision Trees
-Flowchart-like structures in which each internal node represents a "test" on an attribute (e.g.
-whether a coin flip comes up heads or tails), each branch represents the outcome of the test, and each leaf node represents a class label (decision taken after computing all attributes).
-The paths from root to leaf represent classification rules.
-
-In decision analysis, a decision tree and the closely related influence diagram are used as a visual and analytical decision support tool, where the expected values (or expected utility) of competing alternatives are calculated.
-
-A decision tree consists of three types of nodes:
-
-*  Decision nodes – typically represented by squares
-*  Chance nodes – typically represented by circles
-*  End nodes – typically represented by triangles
-
-Decision tree learning is the construction of a decision tree from class-labeled training tuples.
-A decision tree is a flow-chart-like structure, where each internal (non-leaf) node denotes a test on an attribute, each branch represents the outcome of a test, and each leaf (or terminal) node holds a class label.
-The topmost node in a tree is the root node.
-
-The goal is to create a model that predicts the value of a target variable based on several input variables.
-An example is shown in the diagram at right.
-Each interior node corresponds to one of the input variables; there are edges to children for each of the possible values of that input variable.
-Each leaf represents a value of the target variable given the values of the input variables represented by the path from the root to the leaf.
-
-A decision tree is a simple representation for classifying examples.
-For this section, assume that all of the input features have finite discrete domains, and there is a single target feature called the "classification".
-Each element of the domain of the classification is called a class.
-A decision tree or a classification tree is a tree in which each internal (non-leaf) node is labeled with an input feature.
-The arcs coming from a node labeled with an input feature are labeled with each of the possible values of the target or output feature or the arc leads to a subordinate decision node on a different input feature.
-Each leaf of the tree is labeled with a class or a probability distribution over the classes.
-
-A tree can be "learned" by splitting the source set into subsets based on an attribute value test.
-This process is repeated on each derived subset in a recursive manner called recursive partitioning.
-The recursion is completed when the subset at a node has all the same value of the target variable, or when splitting no longer adds value to the predictions.
-This process of top-down induction of decision trees (TDIDT) is an example of a greedy algorithm, and it is by far the most common strategy for learning decision trees from data.
-
-In data mining, decision trees can be described also as the combination of mathematical and computational techniques to aid the description, categorization and generalization of a given set of data.
-
-###Linear Discriminant Analysis
-LDA works when the measurements made on independent variables for each observation are continuous quantities.
-When dealing with categorical independent variables, the equivalent technique is discriminant correspondence analysis.
-The statistical system R includes the packages: MASS, ade4, ca, vegan, ExPosition, andFactoMineR which perform correspondence analysis and multiple correspondence analysis.
-
-As the name indicates, discriminant correspondence analysis (DCA) is an extension of discriminant analysis (DA) and correspondence analysis (CA).
-Like discriminant analysis, the goal of DCA is to categorize observations in pre-defined groups, and like  correspondence analysis, it is used with nominal variables.The main idea behind DCA is to represent each group by thesum of its observations and to perform a simple CA on the groupsby variables matrix.
-The original observations are then projectedas  supplementary  elements and each observation is assigned to the closest group.
-The comparison between the a priori and the a posteriori classifications can be used to assess the quality of thendiscrimination.
-A similar procedure can be used to assign new observations to categories.
-The stability of the analysis can be evaluated using cross-validation techniques such as jackknifing or boot-strapping.
-
-For the R model: https://pbil.univ-lyon1.fr/ade4/ade4-html/discrimin.coa.html
-
-###k-Nearest Neighbors
-k-NN is a non-parametric method used for classification and regression.
-In both cases, the input consists of the k closest training examples in the feature space.
-The output depends on whether k-NN is used for classification or regression.
-In k-NN classification, the output is a class membership.
-An object is classified by a majority vote of its neighbors, with the object being assigned to the class most common among its k nearest neighbors (k is a positive integer, typically small).
-If k = 1, then the object is simply assigned to the class of that single nearest neighbor.
-
-###Similarity Learning
-Similarity learning is an area of supervised machine learning in artificial intelligence.
-It is closely related to regression and classification, but the goal is to learn from examples a similarity function that measures how similar or related two objects are.
-It has applications in ranking, in recommendation systems, visual identity tracking, face verification, and speaker verification.
-
-###Relevance Feedback/Rocchio Algorithm
-Relevance feedback is a feature of some information retrieval systems.
-The idea behind relevance feedback is to take the results that are initially returned from a given query, to gather user feedback, and to use information about whether or not those results are relevant to perform a new query.
-We can usefully distinguish between three types of feedback: explicit feedback, implicit feedback, and blind or "pseudo" feedback.
-
-The Rocchio algorithm is based on a method of relevance feedback.
-Like many other retrieval systems, the Rocchio feedback approach was developed using the Vector Space Model.
-The algorithm is based on the assumption that most users have a general conception of which documents should be denoted as relevant or non-relevant.
-Therefore, the user's search query is revised to include an arbitrary percentage of relevant and non-relevant documents as a means of increasing the search engine's recall, and possibly the precision as well.
-
-For more:
-https://nlp.stanford.edu/IR-book/html/htmledition/rocchio-classification-1.html
-
-###Ensemble Methods
-The excerpt below nice describes how ensemble methods work. It is reprinted from: *Ensemble Methods Explained in Simple English*, by Tavish Srivastava, posted August 2, 2015.
-
-Ensemble modeling is a powerful way to improve the performance of your model. It usually pays off to apply ensemble learning over and above various models you might be building.
-
-Example: I want to invest in a company XYZ.
-I am not sure about its performance though.
-So, I look for advice on whether the stock price will increase more than 6% per annum or not?
-I decide to approach various experts having diverse domain experience:
-
->1.
-Employee of Company XYZ: This person knows the internal functionality of the company and have the insider information about the functionality of the firm.
-But he lacks a broader perspective on how are competitors innovating, how is the technology evolving and what will be the impact of this evolution on Company XYZ’s product.
-In the past, he has been right 70% times.
-
->1.
-Financial Advisor of Company XYZ: This person has a broader perspective on how companies strategy will fair of in this competitive environment.
-However, he lacks a view on how the company’s internal policies are fairing off.
-In the past, he has been right 75% times.
-
->1.
-Stock Market Trader: This person has observed the company’s stock price over past 3 years.
-He knows the seasonality trends and how the overall market is performing.
-He also has developed a strong intuition on how stocks might vary over time.
-In the past, he has been right 70% times.
-
->1.
-Employee of a competitor: This person knows the internal functionality of the competitor firms and is aware of certain changes which are yet to be brought.
-He lacks a sight of company in focus and the external factors which can relate the growth of competitor with the company of subject.
-In the past, he has been right 60% of times.
-
->1.
-Market Research team in same segment: This team analyzes the customer preference of company XYZ’s product over others and how is this changing with time.
-Because he deals with customer side, he is unaware of the changes company XYZ will bring because of alignment to its own goals.
-In the past, they have been right 75% of times.
-
->1.
-Social Media Expert: This person can help us understand how has company XYZ has positioned its products in the market.
-And how are the sentiment of customers changing over time towards company.
-He is unaware of any kind of details beyond digital marketing.
-In the past, he has been right 65% of times.
-
-Given the broad spectrum of access we have, we can probably combine all the information and make an informed decision.
-In a scenario when all the 6 experts/teams verify that it’s a good decision(assuming all the predictions are independent of each other), we will get a combined accuracy rate of:
-
-```
-		  1 - 30%*25%*30%*40%*25%*35%
-
-		= 1 - 0.07875 = 99.92125%
-```
-
-Assumption: The assumption used here that all the predictions are completely independent is slightly extreme as they are expected to be correlated.
-However, we see how we can be so sure by combining various predictions together.
-
-Let us now change the scenario slightly. This time we have 6 experts, all of them are employee of company XYZ working in the same division. Everyone has a propensity of 70% to advocate correctly.
-
->*What if we combine all this advice together, can we still raise up our confidence to >99%?*
-
-Obviously not, as all the predictions are based on very similar set of information.
-They are certain to be influenced by similar set of information and the only variation in their advice would be due to their personal opinions & collected facts about the firm.
-
-Ensemble learning is the art of combining individual models together to improve the stability and predictive power of the model.
-
-####Error Sources in Ensemble Learning (Variance vs. Bias)
-The error emerging from any model can be broken down into three components mathematically.
-
-**Error of a model** Why is this important in the current context? To understand what really goes behind an ensemble model, we need to first understand what causes error in the model.
-
-**Bias error** quantifies how much on an average the predicted values differ from the actual value. A high bias error means we have a under-performing model which keeps on missing important trends.
-
-**Variance** quantifies how much the predictions made on same observations differ from each other.
-A high variance model will over-fit on your training population and perform badly on any observation beyond training.
-
-Normally, as you increase the complexity of your model, you will see a reduction in error due to lower bias in the model.
-However, this only happens till a particular point.
-As you continue to make your model more complex, you end up **over-fitting your model** and hence your model will start suffering from high variance.
-
-A champion model should maintain a balance between these two types of errors.
-This is known as the **trade-off management of bias-variance errors.** Ensemble learning is one way to execute this trade off analysis.
-
-####Commonly used Ensemble learning techniques
-**Bagging** tries to implement similar learners on small sample populations and then takes a mean of all the predictions.
-In generalized bagging, you can use different learners on different population.
-As you can expect this helps us to reduce the variance error.
-
-**Boosting** is an iterative technique which adjust the weight of an observation based on the last classification.
-If an observation was classified incorrectly, it tries to increase the weight of this observation and vice versa.
-Boosting in general decreases the bias error and builds strong predictive models.
-However, they may sometimes over fit on the training data.
-
-**Stacking** is a very interesting way of combining models.
-Here we use a learner to combine output from different learners.
-This can lead to decrease in either bias or variance error depending on the combining learner we use.
-
-####Random Forest Models
-Random Forest models are a type of ensemble models, particularly bagging models. They are part of the tree based model family. RF can be run on:
-
-*  Word count or n-gram count vectors
-*  Word level TF-IDF vectors
-
-Resources for Random Forest: 
-https://www.analyticsvidhya.com/blog/2014/06/introduction-random-forest-simplified/
-
-####Extreme Gradient Boosting Model
-Boosting models are another type of ensemble models part of tree based models.
-Boosting is a machine learning ensemble meta-algorithm for primarily reducing bias, and also variance in supervised learning, and a family of machine learning algorithms that convert weak learners to strong ones.
-A weak learner is defined to be a classifier that is only slightly correlated with the true classification (it can label examples better than random guessing).
-XGB can be run on:
-
-*  Word count or n-gram count vectors
-*  Word level TF-IDF vectors
-*  Character Level TF-IDF vectors
-
-###Hierarchical Clustering
-In data mining and statistics, hierarchical clustering (also called hierarchical cluster analysis or HCA) is a method of cluster analysis which seeks to build a hierarchy of clusters.
-
-###k-Means Clustering
-k-means clustering is a method of vector quantization, originally from signal processing, that is popular for cluster analysis in data mining.
-k-means clustering aims to partition n observations into k clusters in which each observation belongs to the cluster with the nearest mean, serving as a prototype of the cluster.
-This results in a partitioning of the data space into Voronoi cells.
-
-The problem is computationally difficult (NP-hard); however, efficient heuristic algorithms converge quickly to a local optimum.
-These are usually similar to the expectation-maximization algorithm for mixtures of Gaussian distributions via an iterative refinement approach employed by both k-means and Gaussian mixture modeling.
-They both use cluster centers to model the data; however, k-means clustering tends to find clusters of comparable spatial extent, while the expectation-maximization mechanism allows clusters to have different shapes.
-
-The algorithm has a loose relationship to the k-nearest neighbor classifier, a popular machine learning technique for classification that is often confused with k-means due to the name.
-Applying the 1-nearest neighbor classifier the cluster centers obtained by k-means classifies new data into the existing clusters.
-This is known as nearest centroid classifier or Rocchio algorithm.
-
-###Mixture Model
-In statistics, a mixture model is a probabilistic model for representing the presence of subpopulations within an overall population, without requiring that an observed data set should identify the sub-population to which an individual observation belongs.
-Formally a mixture model corresponds to the mixture distribution that represents the probability distribution of observations in the overall population.
-However, while problems associated with "mixture distributions" relate to deriving the properties of the overall population from those of the sub-populations, "mixture models" are used to make statistical inferences about the properties of the sub-populations given only observations on the pooled population, without sub-population identity information.
-
-Topics in a document: Assume that a document is composed of N different words from a total vocabulary of size V, where each word corresponds to one of K possible topics.
-The distribution of such words could be modelled as a mixture of K different V-dimensional categorical distributions.
-A model of this sort is commonly termed a topic model.
-Note that expectation maximization applied to such a model will typically fail to produce realistic results, due (among other things) to the excessive number of parameters.
-Some sorts of additional assumptions are typically necessary to get good results.
-
-A prior distribution is placed over the parameters describing the topic distributions, using a Dirichlet distribution with a concentration parameter that is set significantly below 1, so as to encourage sparse distributions (where only a small number of words have significantly non-zero probabilities).
-
-Some sort of additional constraint is placed over the topic identities of words, to take advantage of natural clustering.
-
-For example, a Markov chain could be placed on the topic identities (i.e., the latent variables specifying the mixture component of each observation), corresponding to the fact that nearby words belong to similar topics.
-(This results in a hidden Markov model, specifically one where a prior distribution is placed over state transitions that favors transitions that stay in the same state.)
-
-Another possibility is the latent Dirichlet allocation model, which divides up the words into D different documents and assumes that in each document only a small number of topics occur with any frequency.
-
-###Principal Component Analysis 
-PCA is a statistical procedure that uses an orthogonal transformation to convert a set of observations of possibly correlated variables (entities each of which takes on various numerical values) into a set of values of linearly uncorrelated variables called principal components.
-This transformation is defined in such a way that the first principal component has the largest possible variance (that is, accounts for as much of the variability in the data as possible), and each succeeding component in turn has the highest variance possible under the constraint that it is orthogonal to the preceding components.
-The resulting vectors (each being a linear combination of the variables and containing n observations) are an uncorrelated orthogonal basis set.
-PCA is sensitive to the relative scaling of the original variables.
-
-PCA is mostly used as a tool in exploratory data analysis and for making predictive models.
-It is often used to visualize genetic distance and relatedness between populations.
-PCA can be done by eigenvalue decomposition of a data covariance (or correlation) matrix or singular value decomposition of a data matrix, usually after a normalization step of the initial data.
-The normalization of each attribute consists of mean centering – subtracting each data value from its variable's measured mean so that its empirical mean (average) is zero – and, possibly, normalizing each variable's variance to make it equal to 1.
-The results of a PCA are usually discussed in terms of component scores, sometimes called factor scores (the transformed variable values corresponding to a particular data point), and loadings (the weight by which each standardized original variable should be multiplied to get the component score).
-If component scores are standardized to unit variance loadings must contain the data variance in them (and that is the magnitude of eigenvalues).
-If component scores are not standardized (therefore they contain the data variance) then loadings must be unit-scaled, ("normalized") and these weights are called eigenvectors; they are the cosines of orthogonal rotation of variables into principal components or back.
-
-PCA is the simplest of the true eigenvector-based multivariate analyses.
-Often, its operation can be thought of as revealing the internal structure of the data in a way that best explains the variance in the data.
-If a multivariate dataset is visualised as a set of coordinates in a high-dimensional data space (1 axis per variable), PCA can supply the user with a lower-dimensional picture, a projection of this object when viewed from its most informative viewpoint.
-This is done by using only the first few principal components so that the dimensionality of the transformed data is reduced.
-
-PCA is closely related to factor analysis.
-Factor analysis typically incorporates more domain specific assumptions about the underlying structure and solves eigenvectors of a slightly different matrix.
-
-PCA is also related to canonical correlation analysis (CCA).
-CCA defines coordinate systems that optimally describe the cross-covariance between two datasets while PCA defines a new orthogonal coordinate system that optimally describes variance in a single dataset.
-
-##Strategies for Improving Text Classification Models
-Accuracy can be improved using one or more of these strategies.
-
-1. Text Cleaning. Removing stopwords, punctuations marks, suffix variations etc. can  reduce the noise present in text data.
-
-2.
-Hstacking arrays of text / NLP features with text feature vectors in sequence horizontally (column wise).
-After extracting features initially, try combining different feature vectors to improve the accuracy of the classifier.
-
-3. Tuning the Classifier. Adjusting the parameters of the classifier is an important step towards a best fit model.
-
-4. Testing Different Ensemble Models. Stacking different models and blending their outputs can help to further improve the results.
-
-5.
-Weighting words: we can significantly improve our algorithm by accounting for the commonality of each word.
-The word “is” should carry a lower weigh than the word “sandwich” in most cases, because it is more common.
-
-##Evaluation
-How to evaluate the different components of the overall text classification problem.
-
-###Gini Index
-One of the most common methods for quantifying the discrimination level of a feature is the use of a measure known as the gini-index.
-
-###Cohen's Kappa Score
-Cohen's kappa is a measure of interrater reliability used in social sciences. This excerpt from an R blog explains it.
-
-**Using kappa score for comparisons of groups and methods**
-https://www.personality-project.org/r/html/kappa.html
-https://www.r-bloggers.com/k-is-for-cohens-kappa/
-
-Last April, during the A to Z of Statistics, I blogged about Cohen’s kappa, a measure of interrater reliability.
-Cohen’s kappa is a way to assess whether two raters or judges are rating something the same way.
-And thanks to an R package called **irr**, it’s very easy to compute.
-But first, let’s talk about why you would use Cohen’s kappa and why it’s superior to a more simple measure of interrater reliability, interrater agreement.
-
-***
-***
-
-###Neural Networks
-A neural network is a mathematical model that is designed to behave similar to biological neurons and nervous system.
-These models are used to recognize complex patterns and relationships that exists within a labelled data.
-
-Artificial neural networks, invented in the 1940’s, are a way of calculating an output from an input (a classification) using weighted connections (“synapses”) that are calculated from repeated iterations through training data.
-Each pass through the training data alters the weights such that the neural network produces the output with greater “accuracy” (lower error rate).
-
-There’s not much new about these structures, except today’s software is using much faster processors and can work with a lot more memory.
-The combination of working memory and speed is crucial when you’re doing hundreds of thousands of matrix multiplications (the neural network’s essential math operation).
-
-As in the prior method, each class is given with some number of example sentences.
-Once again each sentence is broken down by word (stemmed) and each word becomes an input for the neural network.
-The synaptic weights are then calculated by iterating through the training data thousands of times, each time adjusting the weights slightly to greater accuracy.
-By recalculating back across multiple layers (“back-propagation”) the weights of all synapses are calibrated while the results are compared to the training data output.
-These weights are like a ‘strength’ measure, in a neuron the synaptic weight is what causes something to be more memorable than not.
-You remember a thing more because you’ve seen it more times: each time the ‘weight’ increases slightly.
-
-At some point the adjustment reaches a point of diminishing returns, this is called “over-fitting” and going beyond this is counter-productive.
-
-The trained neural network is less code than an comparable algorithm but it requires a potentially large matrix of “weights”.
-In a relatively small sample, where the training sentences have 150 unique words and 30 classes this would be a matrix of 150x30.
-Imagine multiplying a matrix of this size 100,000 times to establish a sufficiently low error rate.
-This is where processing speed comes in.
-
-If the neural network sounds magnificently sophisticated, relax, it boils down to matrix multiplication and a formula for reducing values between -1 and 1 or some other minimal range.
-A middle-school math student could learn this in a few hours.
-The hard work is achieving clean training data.
-
-Just as there are variations in pattern matching code and in algorithms, there are variations in neural networks, some more complex than others.
-The basic machinery is the same.
-The essential work is that of classification.
-
-Networks are looking for patterns in collections of terms, each term is reduced to a token.
-In this machine words have no meaning except for their patterned existence within training data.
-The label “artificial intelligence” applied to such machinery is mostly BS.
-
-###Types of Neural Networks
-A shallow neural network contains mainly three types of layers – input layer, hidden layer, and output layer.
-
-Deep Neural Networks are more complex neural networks in which the hidden layers performs much more complex operations than simple sigmoid or relu activations.
-Different types of deep learning models can be applied in text classification problems.
-
-####Convolutional Neural Network
-In Convolutional neural networks, convolutions over the input layer are used to compute the output.
-This results in local connections, where each region of the input is connected to a neuron in the output.
-Each layer applies different filters and combines their results.
-
-####Recurrent Neural Network: LSTM
-Unlike Feed-forward neural networks in which activation outputs are propagated only in one direction, the activation outputs from neurons propagate in both directions (from inputs to outputs and from outputs to inputs) in Recurrent Neural Networks.
-This creates loops in the neural network architecture which acts as a ‘memory state’ of the neurons.
-This state allows the neurons an ability to remember what have been learned so far.
-
-The memory state in RNNs gives an advantage over traditional neural networks but a problem called Vanishing Gradient is associated with them.
-In this problem, while learning with a large number of layers, it becomes really hard for the network to learn and tune the parameters of the earlier layers.
-To address this problem, A new type of RNNs called LSTMs (Long Short Term Memory) Models have been developed.
-
-####Recurrent Neural Network: GRU
-Gated Recurrent Units are another form of recurrent neural networks.
-
-####Bidirectional RNN
-RNN layers can be wrapped in Bidirectional layers as well.
-
-####Recurrent Convolutional Neural Network
-Once the essential architectures have been tried out, one can try different variants of these layers such as recurrent convolutional neural network. Another variants can be:
-
-*  Hierarchical Attention Networks
-*  Sequence to Sequence Models with Attention
-*  Bidirectional Recurrent Convolutional Neural Networks
-*  CNNs and RNNs with more number of layers
+####\

### 2019-02-25 hiding/Base_Bayes_classifier.Rmd
updates

---
title: "Base Model Classifier"
author: "Dan Johnson"
date: "2/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(caret)
library(tm)
library(arm)
```

```{r}
# Training data.
#data2 <- c('Cats like to chase mice.', 'Dogs like to eat big bones.')

data_train <- c('A filter excludes all light, but the red wavelength. Not the other way around.','A good control here would have been to soak the plates in water.','A myogram is a representation of the temporal curse of a contraction. It is not something that you use to record the data.','A one-tailed paired t-test is more appropriate here, please re-analyze the data.','A one-way ANOVA test should’ve been used first. Then if there was significance, use the Tukey-Kramer post hoc test to see which treatments different significantly from each other or not. Since excel doesn’t give you p-values in Tukey-Kramer, then you can use a two-sample t test to determine what the p-values are from the groups that are significant','A paired t-test is more appropriate for your data, as it takes into account the fact that you used the same fish for both treatments.','A photo of the plants would help support this!','A sentence about why you are using this organism would be good.','A separate chart for standard deviation is not needed as you have it already as a bar on your graphs.','A t-test is not necessary to calculate the SD. The SD is a descriptor of the sample, it indicates the amount of dispersion of the set of data values.','Above, you say that you measured masses. But you only talk about the results for length here. Would be good to show the data for all of the measurements you talk about in the methods section','Absorbance of what? What’s being absorbed? (This needs more context.)','According to statistics they are similar, not slightly different','According to this, the eight caterpillars were treated with Martin’s IGR','According to your formula (before - after treatment) this value should be positive because the contraction was stronger before the injection. So correct this.','ACh is not involved in the propagation of the action potential. It is involved in the synapse and it is required for the generation of an AP in the muscle.','Adapt really isn’t the right word here; it refers to a specific evolutionary process as opposed to the plasticity of a trait','Add a sentence about why you used Rana pipiens as a subject','Add a sentence or more about Rana pipiens and why you are using it.','Adding an x-axis label does not mean you should remove the labels clarifying which is control and which is experimental (“treated”). If you don’t wish to have too much text, you could also make control and experimental different colors and add a key showing what the different colors mean.','Additionally, the way is written could be improved. Because you didn’t get statistical significant, then you could conclude that the results did not support the hypothesis.','After the class explanation, try doing one t-test for all the data.  Let me know if you need help.','Again, detail for the methods section.','Again, error bars? What do they mean?','Again, I’d revisit this, since we’re just talking about absorption via spec here—not the chloroplasts’ absorption of light.','Again, no need to mention the null','Again, this can all go in your methods. The independent and dependent variable are made clear already by saying that you will be testing the effects of Martins IGR.  I understand that you are just trying to state them explicitly, as the resource guide tells you to, but this is unnecessary.','Again, this should be percent change—not absorbance—plotted if you want to report percent change values in your results section. You don’t need to include the blank, but all of these data points should have individual error bars since you had replicates.','Again, what do your error bars represent?','Again, what we labeled the tubes isn’t important','Agar and Water amounts remain constant throughout control and test plates')
          #,'Agar and Water amounts remain constant throughout your control and treatment plates. For treatment group, you added 4ml of 20x test compound solution on top of it.','Ah, but it is a single cell! (Isn’t Physarum cool?!)','Ah, so a good control here would have been covering the control PDA plates with water and allowing them to soak for 10 minutes too.','Ah, so it didn’t move directly towards the Physarum. Here it’s important to make that distinction in how you took measurements in your methods section. Did you go based on leading edge? If so, this data point would probably be inconclusive in the results section. Did you go based on a central axis upon which you took measurements? Instead, that’d be different.','All information that is not common knowledge needs to be cited.','All of them or are you referring to Martin’s IGR?','All of these tables are not needed. The first two are summed up in your figures and the second two should be in paragraph form in the results.','All of this should be saved for the discussion.  It talks to much about why you did things rather than the cold results.','All the legs received the lidocaine treatment. Write this in accordance to that.','All these are the results of the experiment.')

data_test <- c('All these caterpillar pictures are nice, but if you’re going to include them, you need to reference them somewhere in your report, generally in methods','All these numbers are confusing and make your result section hard to follow.','All these should be your results. That is written exactly like it should be in the results.','All this should be in the methods section.','All three amongst each other or the three treatments against the control?','All your error bars seem incorrect, as you don’t need error bars when each bar is one individual','Almost perfect','Also check that your standard deviation bars are reflecting the correct values, they look like they are the same size and they shouldn’t be.','Also citation','Also clean the image up a bit and delete the “Chart area” box','Also include the ANOVA and post hoc test results as a figure so you can cite your table in your results to show where you got your p-value from','Also include the proportions of water and agar in these plates','Also include your t-statistic and degrees of freedom','Also introduce statistical tests you performed to analyze the data','Also introduce the stat test you performed to analyze your data as final sentence in methods data.','Also mention briefly what you will actually be doing in the experiment','Also refence your figures when you use them as leverage in the discussion.','Also the error bars are not placed correctly','Also there’s some problems with your error bars. See me if you need help making the graphs','Also These graphs both say Experimental Group on the x axis.  This data is not a good choice to show in a figure because you don’t have any data for the tetany.  Instead of showing a graph, show a figure which can be a photo, a diagram, things like that.  It will need a caption but the in depth description of the figure could be put in the intro/discussion based on what you choose.For example:  you could show a diagram of a myogram of normal contraction and of tetany and use it to explain what happened to the leg.  This would count as a figure and would be a strong addition to the paper.','Also your stats say there is no difference between the two groups','Also- oat plants do flower.')
```

```{r}
corpus <- VCorpus(VectorSource(data_train))

# Create a document term matrix.
tdm <- DocumentTermMatrix(corpus, list(removePunctuation = TRUE, stopwords = TRUE, stemming = TRUE, removeNumbers = TRUE))

# Convert to a data.frame for training and assign a classification (factor) to each document.
train <- as.matrix(tdm)
train <- cbind(train, c(0, 1))
colnames(train)[ncol(train)] <- 'y'
train <- as.data.frame(train)
train$y <- as.factor(train$y)
```

```{r}
# Train.
fit <- train(y ~ ., data = train, method = 'bayesglm')
```

```{r}
# Check accuracy on training.
predict(fit, newdata = train)
```

```{r}
# Test data.
corpus2 <- VCorpus(VectorSource(data_test))
tdm2 <- DocumentTermMatrix(corpus2, control = list(dictionary = Terms(tdm), removePunctuation = TRUE, stopwords = TRUE, stemming = TRUE, removeNumbers = TRUE))
test2 <- as.matrix(tdm2)

# Check accuracy on test.
predict(fit, newdata = test2)
```

### 2019-02-25 hiding/Best_practices.Rmd
updates

---
title: "Best Practices for Writing R Code"
author: "downloaded"
date: "2/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Question:
How can I write R code that other people can understand and use?

Your objectives:

*  Describe changes made to code for future reference."
*  Understand the importance about requirements and dependencies for your code."
*  Know when to use setwd()."
*  To identify and segregate distinct components in your code using # or #-."

Additional best practice recommendations

- "Start each program with a description of what it does."
- "Then load all required packages."
- "Consider what working directory you are in when sourcing a script."
- "Use comments to mark off sections of code."
- "Put function definitions at the top of your file, or in a separate file if there are many."
- "Name and style code consistently."
- "Break code into small, discrete pieces."
- "Factor out common operations rather than repeating them."
- "Keep all of the source files for a project in one directory and use relative paths to access them."
- "Keep track of the memory used by your program."
- "Always start with a clean environment instead of saving the workspace."
- "Keep track of session information in your project folder."
- "Have someone else review your code."
- "Use version control."

```{r source, include = FALSE}
source("../bin/chunk-options.R")
knitr_fig_path("06-best-practices-R-")
```

Originating URL: https://swcarpentry.github.io/r-novice-inflammation/06-best-practices-R/

### Keep track of who wrote your code and its intended purpose

Starting your code with an annotated description of what the code does when it is run will help you when you have to look at or change it in the future. Just one or two lines at the beginning of the file can save you or someone else a lot of time and effort when trying to understand what a particular script does.

```{r comment1}
# This is code to replicate the analyses and figures from my 2014 Science
# paper. Code developed by Sarah Supp, Tracy Teal, and Jon Borelli
```

### Be explicit about the requirements and dependencies of your code

Loading all of the packages that will be necessary to run your code (using `library`) is a nice way of indicating which packages are necessary to run your code. It can be frustrating to make it two-thirds of the way through a long-running script only to find out that a dependency hasn't been installed.

```{r loadpkgs, eval=FALSE}
library(ggplot2)
library(reshape)
library(vegan)
```
Another way you can be explicit about the requirements of your code and improve it's reproducibility is to limit the "hard-coding" of the input and output files for your script. If your code will read in data from a file, define a variable early in your code that stores the path to that file. For example

```{r, eval=FALSE}
input_file <- "data/data.csv" 
output_file <- "data/results.csv"
# read input
input_data <- read.csv(input_file)
# get number of samples in data
sample_number <- nrow(input_data)
# generate results
results <- some_other_function(input_file, sample_number)
# write results
write.table(results, results_file)
```

is preferable to

```{r, eval=FALSE}
# check
input_data <- read.csv("data/data.csv")
# get number of samples in data
sample_number <- nrow(input_data)
# generate results
results <- some_other_function("data/data.csv", sample_number)
# write results
write.table("data/results.csv", results_file)
```

It is also worth considering what the working directory is. If the working directory must change, it is best to do that at the beginning of the script.

## Be careful when using `setwd()`
One should exercise caution when using `setwd()`. Changing directories in a script file can limit reproducibility:

* `setwd()` will return an error if the directory to which you're trying to change doesn't exist or if the user doesn't have the correct permissions to access that directory. This becomes a problem when sharing scripts between users who have organized their directories differently.
* If/when your script terminates with an error, you might leave the user in a different directory than the one they started in, and if they then call the script again, this will cause further problems. If you must use `setwd()`, it is best to put it at the top of the script to avoid these problems.

The following error message indicates that R has failed to set the working directory you specified:
 ```
Error in setwd("~/path/to/working/directory") : cannot change working directory
```

It is best practice to have the user running the script begin in a consistent directory on their machine and then use relative file paths from that directory to access files (see below).

### Identify and segregate distinct components in your code
It's easy to annotate and mark your code using `#` or `#-` to set off sections of your code and to make finding specific parts of your code easier. For example, it's often helpful when writing code to separate the function definitions. If you create only one or a few custom functions in your script, put them toward the top of your code. If you have written many functions, put them all in their own .R file and then `source` those files. `source` will define all of these functions so that your code can make use of them as needed.

```{r source_ex, eval=FALSE}
source("my_genius_fxns.R")
```

### Other Ideas
1. Use a consistent style within your code. For example, name all matrices something ending in `_mat`. Consistency makes code easier to read and problems easier to spot.

2. Keep your code in bite-sized chunks. If a single function or loop gets too long, consider looking for ways to break it into smaller pieces.

3. Don't repeat yourself--automate! If you are repeating the same code over and over, use a loop or a function to repeat that code for you. Needless repetition doesn't just waste time--it also increases the likelihood you'll make a costly mistake!

4. Keep all of your source files for a project in the same directory, then use relative paths as necessary to access them. For example, use

```{r relpath, eval=FALSE}
dat <- read.csv(file = "files/dataset-2013-01.csv", header = TRUE)
```

rather than:

```{r abspath, eval=FALSE}
dat <- read.csv(file = "/Users/Karthik/Documents/sannic-project/files/dataset-2013-01.csv", header = TRUE)
```

5. R can run into memory issues. It is a common problem to run out of memory after running R scripts for a long time. To inspect the objects in your current R environment, you can list the objects, search current packages, and remove objects that are currently not in use. A good practice when running long lines of computationally intensive code is to remove temporary objects after they have served their purpose. However, sometimes, R will not clean up unused memory for a while after you delete objects. You can force R to tidy up its memory by using `gc()`.

```{r gc_ex, eval=FALSE}
# Sample dataset of 1000 rows
interim_object <- data.frame(rep(1:100, 10),
                             rep(101:200, 10),
                             rep(201:300, 10))
object.size(interim_object) # Reports the memory size allocated to the object
rm("interim_object") # Removes only the object itself and not necessarily the memory allotted to it
gc() # Force R to release memory it is no longer using
ls() # Lists all the objects in your current workspace
rm(list = ls()) # If you want to delete all the objects in the workspace and start with a clean slate
```

6. Don't save a session history (the default option in R, when it asks if you want an `RData` file). Instead, start in a clean environment so that older objects don't remain in your environment any longer than they need to. If that happens, it can lead to unexpected results.

7. Wherever possible, keep track of `sessionInfo()` somewhere in your project folder. Session information is invaluable because it captures all of the packages used in the current project. If a newer version of a package changes the way a function behaves, you can always go back and reinstall the version that worked (Note: At least on CRAN, all older versions of packages are permanently archived).

8. Collaborate. Grab a buddy and practice "code review". Review is used for preparing experiments and manuscripts; why not use it for code as well? Our code is also a major scientific achievement and the product of lots of hard work!

9. Develop your code using version control and frequent updates! You can find lessons about version control on [software-carpentry.org/lessons][swc-lessons].

### 2019-02-25 hiding/Check_in_Feb_4.Rmd
updates

---
title: "Feb 4 Check-In"
author: "Dan Johnson"
date: "2/3/19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#New ideas From FLC Meeting #6

Need help with:

Coding issues

*  How to ID, prune unneeded packages. NO SIMPLE SOLUTION. Cut and test seems best.

*  When I join columns in correlation analysis, if a word is MISSING from **any subset** the data for that word is pruned from the full dataset. I am losing the unique words this way. 
    +  Assume the NA would blow up correlations, which is why it is done this way.
    +  Need to figure out how to join without pruning, then find the UNIQUE words.

Post-Processing issues. Use broom package

*  How to calculate the value automatically in the Naive Bayes final step. Data post-processing. I have stable code for my binomial naive Bayes classifier, and I am screening different permutations to see what the best approach is to refining the model. There is one step I cannot figure out how to code though.My primary output is a cross-comparison table showing what fraction of the test set comments is correctly identified. Adding the counts diagonally and dividing by the total number of test cases gives me the overall accuracy.The table has all of the data that I want to use for other calculations, but I cannot figure out how to either: 1) call values from the table, or 2) pull

 Output issues

*  How to take the Excel/csv tables I have for permutations of naive Bayes, Pearson's r, and bring them in for display in summary page. broom

Add to My Code:

*  shell(cmd, shell, flag = "/c", intern = FALSE, wait = TRUE,
      translate = FALSE, mustWork = FALSE, ...) runs UNIX commands
https://www.rdocumentation.org/packages/base/versions/3.4.0/topics/shell

*  Using csvy format. R supports .csvy using csvy package and read_csvy() and write_csvy()
Reference Sites:

*  Using broom for cleanup
https://cran.r-project.org/web/packages/broom/vignettes/broom.html

Packages to Explore:

*  fs. https://fs.r-lib.org/index.html

fs provides a cross-platform, uniform interface to file system operations. It shares the same back-end component as nodejs, the libuv C library, which brings the benefit of extensive real-world use and rigorous cross-platform testing. fs functions are divided into four main categories:

    path_ for manipulating and constructing paths
    file_ for files
    dir_ for directories
    link_ for links
fs is designed to work well with the pipe, though because it is a minimal-dependency infrastructure package it doesn’t provide the pipe itself. You will need to attach magrittr or similar. fs functions also work well in conjunction with other tidyverse packages, like dplyr and purrr.

To Do List:

*  Best practices when writing R code
https://swcarpentry.github.io/r-novice-inflammation/06-best-practices-R/

#File Management

```{r}
headlines2 <- read_docx("R_0eqZDDQ0nG2obuh_text.docx") %>% head(9)
View(headlines2)
```

R can manipulate files and folders. This is a primer:
http://theautomatic.net/2018/07/11/manipulate-files-r/

Get, change working directory
Create files of any type based on extension
Copy a file or folder
List files in a directory, including by sub-type
Use rowbind to join file contents
Get snapshot data
Remove files, folders
Check for files, folders
Get the basenames for files, folders
Get file extensions
Call shell to open the file
Move a file

# read in all the CSV files
all_data_frames <- lapply(list.files(pattern = ".csv"), read.csv)

# stack all data frames together
single_data_frame <- Reduce(rbind, all_data_frames)

Default
draft
General_workflow
initial_exploration
Naive_bayes_V2b, r, yn
pre_work
project
questions
reference_tables
course_notes

ta_comments_FLC
Text Classifier Models
Random_forest_notes
Latent_dirichlet_allocation_notes
  folder=data
    codebook_for_comments.md
    code_column_names.md
    TA_and_Student_IDs.csv
    ID_lookup.csv
  folder=docs_and_notes
    Grading Criteria.rmd
    Raw_data_structure.rmd

###Planned Project Pages
It may seem excessive, but my thinking is that:
1. Our group is the first so should establish some guidelines for creating tools.
2. Work I put in here translates over to the final, shared product for NSF project.

*  toc = Table of Contents
*	 project = Project Summary

*	 Background
    +  Theoretical background on text classifiers
    +  Practical background on project challenges

*	 Project Methods and Guides
    +  Anonymizing Data
    +  Standard Workflow
    +  Reference Tables
        -  Recoding lists
    +  Project Codebook

*  Pages template:
    +  Page Goals
    +  Page Overview
    +  Page Content

*  Initial Exploration
    +  Data Processing
    +  Preliminary Visualization

*  Deeper Analyses
    +  Categorical Naive Bayes
        -  Baseline cNB
        -  Randomized cNB
        -  Binary cNB
        -  Summary of cNB Findings

***
***

#Jan 14 Check-In, and New ideas From FLC Meeting #5

Check my data for correlations.
	"cor" blows up on missing data
	Try "use" pairwise complete

*  Use R profile for storing API tokens

```{r}
file.edit(file.path("~", ".Rprofile")) # edit .Rprofile in HOME
file.edit(".Rprofile") # edit project specific .Rprofile

# There is information about the options you can set via
# help("Rprofile")
```

Tools to consider:

*  UDPipe for syntactic parsing
*  Stylometry library
*  QualtRics package for connecting to API (no longer supported by CRAN)
*  By hand exploration followed up with computational exploration

Reference sources:

*  R Open Sci

To Do List:

*  Need to add some data intermediates to my work to help others visualize.
*  Taskviews on CRAN
*  Look at lemmatization, not stemming

Look into Bayes model commands:

*  prob
*  predict

Packages to evaluate

*  caret
*  skimR
*  parsnip
*  recipe
*  tidymodels
*  UDPipe
*  package = usethis

##What Is Working

Suggested: 
Check my data for small errors that break correlations.
	"cor" blows up on missing data
	Try "use" pairwise complete

Tokenizer and terms cleanup seems to work well. (Still struggling with understanding how data get transformed in certain copied code blocks.)
  Suggested: Look into lemmatization, not stemming

## Coding & Interpretation Questions

1.  Randomization. Does the straight split I used bias the data? 
  Comment: YES, so instead use set.seed to pick more randomized dataset.

Suggestions:
  Try: 
    Leave one out cross-validation
    K-folds cross-validation

I do have a few things for you to consider before we do get together. The first thing I suggest you do, which shouldn't be too much of a problem to create, is to provide the output of some of your datasets (particularly thinking about the contrast sets that you use in the correlation analysis). A quick `glimpse(data)` printed out would help me get a better sense of the structure you are using as input and output at each stage of your analysis.

Third thing that comes to mind has to do with the Naive Bayes analysis. As you've seen, naive bayes function from the `e1071` package only works with categorical features. For many NLP tasks this oversimplifies features making the appearance of a word a 'yes' or 'no' decision, which reduces information concerning the frequency of occurrence of words. It's been a while since I've used naive bayes in my own work, but last time I did, I ended up writing my own algorithm to deal with non-categorical features. I'm not altogether happy with the algorithm I coded, but when we meet we can talk about whether there are other package options out there now, or whether my function might be of some help to you.

Anyway, things are looking good. Keep up the good work and great documentation!

## Bigger Questions
Python/Scython porting

After posting an initial naive Bayes trial, Jerid wrote:
>Naive bayes function from `e1071` package only works with categorical features. For many NLP tasks this oversimplifies features making the appearance of a word a 'yes' or 'no' decision.

I spent time looking at various methods as part of background research. The summary is posted elsewhere. Question is whether I should focus on:

*  Continued refinement of NB, or exploring other supervised models? 
    +  How deep to go into different methods?
    +  When to drop one model, move to another?
*  Using my existing code-book to reverse engineer a pure rules based method instead?
    +  Faster for me, but what about others?
*  Extending previous item, create a HYBRID of statistical and rules-based?
    +  Words to evaluate are pre-selected limited vocabulary (see codebook for lists)
    +  Words are scored present/absent in a comment, then comment sorted based on most likely.
*  Step back and rethink the categories using LDA, etc.? 
    +  Seems counter-productive given I have a model built already.

### 2019-02-25 hiding/Latent_dirichlet_allocation_notes.Rmd
updates

---
title: "Latent Dirichlet Analysis v1"
author: "Dan Johnson"
date: "12/14/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Sources:
Primary
https://eight2late.wordpress.com/2015/09/29/a-gentle-introduction-to-topic-modeling-using-r/

```{r}
#load text mining library
library(tm)

#set working directory (modify path as needed)
setwd("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/data/textmining/")

#load files into corpus
#get listing of .txt files in directory
filenames <- list.files(getwd(),pattern="*.txt")

#read files into a character vector
files <- lapply(filenames,readLines)

#create corpus from vector
docs <- Corpus(VectorSource(files))

#inspect a particular document in corpus
writeLines(as.character(docs[[30]]))

#start preprocessing
#Transform to lower case
docs <-tm_map(docs,content_transformer(tolower))

#remove potentially problematic symbols
toSpace <- content_transformer(function(x, pattern) { return (gsub(pattern, " ", x))})
docs <- tm_map(docs, toSpace, "-")
docs <- tm_map(docs, toSpace, "’")
docs <- tm_map(docs, toSpace, "‘")
docs <- tm_map(docs, toSpace, "•")
docs <- tm_map(docs, toSpace, "“")
docs <- tm_map(docs, toSpace, "”")

#remove punctuation
docs <- tm_map(docs, removePunctuation)
#Strip digits
docs <- tm_map(docs, removeNumbers)
#remove stopwords
docs <- tm_map(docs, removeWords, stopwords("english"))
#remove whitespace
docs <- tm_map(docs, stripWhitespace)
#Good practice to check every now and then
writeLines(as.character(docs[[30]]))
#Stem document
docs <- tm_map(docs,stemDocument)

#fix up 1) differences between us and aussie english 2) general errors
docs <- tm_map(docs, content_transformer(gsub),
pattern = "organiz", replacement = "organ")
docs <- tm_map(docs, content_transformer(gsub),
pattern = "organis", replacement = "organ")
docs <- tm_map(docs, content_transformer(gsub),
pattern = "andgovern", replacement = "govern")
docs <- tm_map(docs, content_transformer(gsub),
pattern = "inenterpris", replacement = "enterpris")
docs <- tm_map(docs, content_transformer(gsub),
pattern = "team-", replacement = "team")

#define and eliminate all custom stopwords
myStopwords <- c("can", "say","one","way","use",
"also","howev","tell","will",
"much","need","take","tend","even",
"like","particular","rather","said",
"get","well","make","ask","come","end",
"first","two","help","often","may",
"might","see","someth","thing","point",
"post","look","right","now","think","‘ve ",
"‘re ","anoth","put","set","new","good",
"want","sure","kind","larg","yes,","day","etc",
"quit","sinc","attempt","lack","seen","awar",
"littl","ever","moreov","though","found","abl",
"enough","far","earli","away","achiev","draw",
"last","never","brief","bit","entir","brief",
"great","lot")
docs <- tm_map(docs, removeWords, myStopwords)

#inspect a document as a check
writeLines(as.character(docs[[30]]))

#Create document-term matrix
dtm <- DocumentTermMatrix(docs)

#convert rownames to filenames
rownames(dtm) <- filenames

#collapse matrix by summing over columns
freq <- colSums(as.matrix(dtm))

#length should be total number of terms
length(freq)

#create sort order (descending)
ord <- order(freq,decreasing=TRUE)

#List all terms in decreasing order of freq and write to disk
freq[ord]
write.csv(freq[ord],"word_freq.csv")
```

##Part 2

```{r}
#load topic models library
library(topicmodels)

#set working directory (modify path as needed)
setwd("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/data/textmining/")

#Set parameters for Gibbs sampling
burnin <- 4000
iter <- 2000
thin <- 500
seed <-list(2003,5,63,100001,765)
nstart <- 5
best <- TRUE

#Number of topics
k <- 5

#That done, we can now do the actual work – run the topic modelling algorithm on our corpus. Here is the code:

#Run LDA using Gibbs sampling
ldaOut <-LDA(dtm,k, method="Gibbs", control=list(nstart=nstart, seed = seed, best=best, burnin = burnin, iter = iter, thin=thin))

#write out results
#docs to topics
ldaOut.topics <- as.matrix(topics(ldaOut))
write.csv(ldaOut.topics,file=paste("LDAGibbs",k,"DocsToTopics.csv"))

#top 6 terms in each topic
ldaOut.terms <- as.matrix(terms(ldaOut,6))
write.csv(ldaOut.terms,file=paste("LDAGibbs",k,"TopicsToTerms.csv"))

#probabilities associated with each topic assignment
topicProbabilities <- as.data.frame(ldaOut@gamma)
write.csv(topicProbabilities,file=paste("LDAGibbs",k,"TopicProbabilities.csv"))

#Find relative importance of top 2 topics
topic1ToTopic2 <- lapply(1:nrow(dtm),function(x)
sort(topicProbabilities[x,])[k]/sort(topicProbabilities[x,])[k-1])

#Find relative importance of second and third most important topics
topic2ToTopic3 <- lapply(1:nrow(dtm),function(x)
sort(topicProbabilities[x,])[k-1]/sort(topicProbabilities[x,])[k-2])

#write to file
write.csv(topic1ToTopic2,file=paste("LDAGibbs",k,"Topic1ToTopic2.csv"))
write.csv(topic2ToTopic3,file=paste("LDAGibbs",k,"Topic2ToTopic3.csv"))
```

***
***

##Other Examples of LDA Procedures
https://rpubs.com/Junny31/318845
https://rstudio-pubs-static.s3.amazonaws.com/266565_171416f6c4be464fb11f7d8200c0b8f7.html

***
***

##Another Approach Comparing LDA (unsupervised) and tf-idf (supervised)

https://www.kaggle.com/rtatman/nlp-in-r-topic-modelling
>Latent Dirichlet (said like deer-ih-CLAY) allocation, more commonly shortened to LDA, is one unsupervised way of doing topic modeling. In machine learning, a problem is unsupervised if you don't have labeled examples of the thing you're trying to figure out.

>LDA rests on the assumption that each document is going to have some subset of topics in it, and that each topic is associated with a certain subset of words. Imagine, for example, that you have three text ads:

    One is for Chow-Chow Crunch, a new dog food
    One is for Danny's Downhome Diner, a new restaurant
    One is for Fifi's Chic Puppy Boutique, a store that sells dog accessories

>Both the diner and the dog food ad will probably have words associated with the topic of food in them, like "tasty", "filling" or "delicious". Both the dog food and the pet store will probably have words associated with the topic of pets, like "canine", "pet" or "owner". And, because they're all ads, it likely that all three documents will have words associated with advertising in them, like "sale", "price" or "special".

>The tricky part is that words can also belong to more than one topic. I just said that "special" could be associated with advertising... but it can also be associated with food, especially in terms like "blue plate special" or "chef's special".

>What LDA outputs, then, is two estimates:

    An estimate of how much each topic contributes to each document
    An estimate of how much each word contributes to each topic

>For exploratory analysis, especially if you have a large number of documents, the second estimate is usually more interesting. In order to make our lives easier (since we're going to ran an LDA several times) let's write a function that takes a text column from a data frame and returns a plot of the most informative words for a given number of topics.

```{r}
# function to get & plot the most informative terms by a specificed number
# of topics, using LDA
top_terms_by_topic_LDA <- function(input_text, # should be a columm from a dataframe
                                   plot = T, # return a plot? TRUE by defult
                                   number_of_topics = 4) # number of topics (4 by default)
{    
    # create a corpus (type of object expected by tm) and document term matrix
    Corpus <- Corpus(VectorSource(input_text)) # make a corpus object
    DTM <- DocumentTermMatrix(Corpus) # get the count of words/document

    # remove any empty rows in our document term matrix (if there are any 
    # we'll get an error when we try to run our LDA)
    unique_indexes <- unique(DTM$i) # get the index of each unique value
    DTM <- DTM[unique_indexes,] # get a subset of only those indexes

    # preform LDA & get the words/topic in a tidy text format
    lda <- LDA(DTM, k = number_of_topics, control = list(seed = 1234))
    topics <- tidy(lda, matrix = "beta")

    # get the top ten terms for each topic
    top_terms <- topics  %>% # take the topics data frame and..
      group_by(topic) %>% # treat each topic as a different group
      top_n(10, beta) %>% # get the top 10 most informative words
      ungroup() %>% # ungroup
      arrange(topic, -beta) # arrange words in descending informativeness

    # if the user asks for a plot (TRUE by default)
    if(plot == T){
        # plot the top ten terms for each topic in order
        top_terms %>% # take the top terms
          mutate(term = reorder(term, beta)) %>% # sort terms by beta value 
          ggplot(aes(term, beta, fill = factor(topic))) + # plot beta by theme
          geom_col(show.legend = FALSE) + # as a bar plot
          facet_wrap(~ topic, scales = "free") + # which each topic in a seperate plot
          labs(x = NULL, y = "Beta") + # no x label, change y label 
          coord_flip() # turn bars sideways
    }else{ 
        # if the user does not request a plot
        # return a list of sorted terms instead
        return(top_terms)
    }
}
```

### 2019-02-25 hiding/Naive_bayes_notes4 (original).Rmd
updates

---
title: "Naive_bayesV3"
author: "Dan Johnson"
date: "12/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Permutations of Naive Bayes (NB) Method
I based the original workflow on this tutorial [https://rpubs.com/Billyhansen6/318412]. Version 1.0 correctly identify comment subcategories ~80% of the time. This is better than the inter-rater reliability of two human raters, but still low. This notebook documents tests of various permutations of the NB method including:

*  Compare 2-grams, 3-grams rather than single words
*  Make pair-wise comparisons versus comparing multiple groups simultaneously
*  Modify the variables within the NB analysis itself.
*  Do/Do not stem the terms
*  Remove/Leave stopwords
*  Remove/Leave capitalization
*  Remove/Leave numbers
*  Remove/Leave punctuation

####\

##Starting Version for Bigrams, Trigrams
The original code is in rows 22 to ~235 of of Version 2 of this document. Works as of 12/18/2018. This is modified for n-grams instead of single words.

```{r}
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
library(readr)
```

```{r}
#Read in CSV file named "coded_full_comments_dataset_Spring18anon.csv".
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')

#Isolate rows to compare. 
frequency_writing <- filter(base_data,code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
sms_raw2 <- frequency_writing %>% select(23,22)

#Rename the columns.
names(sms_raw2)[1] <- "type"
names(sms_raw2)[2] <- "text"

#Simplify coding terms
sms_raw2[,1] <- ifelse(sms_raw2[,1] == "2. Writing Quality", "2_Writing", ifelse(sms_raw2[,1] == "3. Technical and Scientific", "3_Technical", ifelse(sms_raw2[,1] == "4. Logic and Thinking","4_Logic",99)))

#Change "type" element from character to a factor for analysis.
sms_raw2$type <- factor(sms_raw2$type)
str(sms_raw2$type)
table(sms_raw2$type)
```

```{r}
##Text Transformation using "tm" package in R to create a volitile coprus that contains the "text" vector from our data frame.
sms_corpus2 <- VCorpus(VectorSource(sms_raw2$text))

#Convert to all lower case letters with "tm_map"" funtion in R, then use the "content_transformer" function to transform the text.
sms_corpus_clean2 <- tm_map(sms_corpus2, content_transformer(tolower))

#This removes numbers. **May not want to do.**
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removeNumbers)

#Stopword removal. **One of variables to test out.**
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removeWords, stopwords())

#Remove punctuation as well using the "removePunctuation" function. Try not doing this. Removes evidence of questions.
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removePunctuation)
```

Perform "stemming" to the text data to strip the suffix from words. **Try NOT doing this.** 
```{}
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, stemDocument)
```

Final step in text mining is to remove white spaces. 
```{r}
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, stripWhitespace)
```

Code to inspect a line of text.
```{r}
as.character(sms_corpus_clean2[[3]])
```

#Single word tokenization
Block tokenizes using the "DocumentTermMatrix" function. This creates a matrix in which the rows indicate documents (SMS messages in this case) and the columns indicate words. It should be noted that the "DocumentTermMaxtrix" function has the power to do all of the text mining above in one command. The "tidy" command is a hack that makes it easier to inspect the document terms matrix.
```{r}
sms_dtm2 <- DocumentTermMatrix(sms_corpus_clean2)
sms_dtm2_tidy <- tidy(sms_dtm2)
```

#N-gram tokenization
This code block uses tm's NLP functions to generate n-grams from the full corpus of TA comments

```{r}

NLP_Tokenizer <- function(x) {
      unlist(lapply(ngrams(words(x), 2), paste, collapse = " "), use.names = FALSE)
}

sms_dtm2_bigrams <- TermDocumentMatrix(sms_corpus_clean2, control=list(tokenize = NLP_Tokenizer))

inspect(sms_dtm2_bigrams)

```

This should build an inspectable doc-terms matrix using n-grams described in "sms_dtm2_bigrams"
```{r}
sms_dtm2_bigrams_tidy <- tidy(sms_dtm2_bigrams)

#sms_dtm_trigram <- DocumentTermMatrix(sms_corpus_clean2, control = list(tokenize=tdm_NLP))
#sms_dtm_trigram_tidy <- tidy(sms_dtm_trigram)
```

Data Tables for Analysis
sms_dtm2
sms_dtm2_bigrams
sms_corpus_clean2

**Along the way I picked up this quick code block for generating 2-grams and 3-grams from full corpus of TA comments using the "corpus"" library**

```{}
library(corpus)
corpus <- sms_corpus_clean2 
text_filter(corpus)$drop_punct==TRUE
term_stats(corpus, ngrams = 2:3)
```

Data Tables for Analysis
sms_dtm2
sms_dtm2_bigrams
sms_dtm_trigram

```{}
**OPTION**
library(tm); library(tau);

tokenize_ngrams <- function(x, n=3) return(rownames(as.data.frame(unclass(textcnt(x,method="string",n=n)))))

texts <- c("This is the first document.", "This is the second file.", "This is the third text.")
corpus <- Corpus(VectorSource(texts))
matrix <- DocumentTermMatrix(corpus,control=list(tokenize=tokenize_ngrams))

**OPTION**
library(RTextTools)
texts <- c("This is the first document.", "This is the second file.", "This is the third text.")
matrix <- create_matrix(texts,ngramLength=3)

**OPTION**
  library("RWeka")
  library("tm")

  data("crude")

  BigramTokenizer <- function(x) NGramTokenizer(x, Weka_control(min = 2, max = 2))
  tdm <- TermDocumentMatrix(crude, control = list(tokenize = BigramTokenizer))

  inspect(tdm[340:345,1:10])

**OPTION**
Can I use bigrams instead of single tokens in a term-document matrix?

Yes. Package NLP provides functionality to compute n-grams which can be used to construct a corresponding tokenizer. E.g.:

  library("tm")
  data("crude")

  BigramTokenizer <-
  function(x)
    unlist(lapply(ngrams(words(x), 2), paste, collapse = " "), use.names = FALSE)

  tdm <- TermDocumentMatrix(crude, control = list(tokenize = BigramTokenizer))
  inspect(removeSparseTerms(tdm[, 1:10], 0.7))
```
####\

##Data Preparation
Split data into training and testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data. Divide our data set into 75% training and 25% testing.

```{r}
.75 * 9129
.25 * 9129
```

This code assumes comments are random. **Probably want to try randomizing them.**
```{r}
sms_dtm_train2 <- sms_dtm2[1:6846, ]
sms_dtm_test2 <- sms_dtm2[6847:9129, ]
```

This code assumes comments are random. **Probably want to try randomizing them.**
```{r}
sms_dtm_train2 <- sms_dtm2_bigrams[1:6846, ]
sms_dtm_test2 <- sms_dtm2_bigrams[6847:9129, ]
```

Save vectors labeling rows in the training and testing vectors
```{r}
sms_train_labels2 <- sms_raw2[1:6846, ]$type
sms_test_labels2 <- sms_raw2[6847:9129,]$type
```

Make sure that the proportion of each sub-category is similar in the training and testing data set.
```{r}
prop.table(table(sms_train_labels2))
prop.table(table(sms_test_labels2))
```

####\

##Visualization
Create a wordcloud of the frequency of the words in the dataset using the package "wordcloud".
```{r}
wordcloud(sms_corpus_clean2, max.words = 50, random.order = FALSE)
```

Compare wordclouds between 3 groups.
```{r}
TWO <- subset(sms_raw2, type == "2_Writing")
THREE <- subset(sms_raw2, type == "3_Technical")
FOUR <- subset(sms_raw2, type == "4_Logic")

wordcloud(TWO$text, max.words = 50, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 50, scale = c(3, 0.5))
wordcloud(FOUR$text, max.words = 50, scale = c(3, 0.5))
```

####\

##Preparation for Naive Bayes
Remove words from the matrix that appear less than 5 times.
```{r}
sms_freq_words2 <- findFreqTerms(sms_dtm_train2, 5)
str(sms_freq_words2)
```

Data Tables for Analysis
sms_dtm2 = sms_dtm2_bigrams,  sms_dtm_trigram
sms_dtm_train2, _test2
sms_train_labels2, sms_test_labels2

Data Tables for Analysis
sms_dtm2
sms_dtm2_bigrams
sms_corpus_clean2

Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
```{r}
sms_dtm_freq_train2 <- sms_dtm_train2[ , sms_freq_words2]

sms_dtm_freq_test2 <- sms_dtm_test2[ , sms_freq_words2]
```

***
***

This code assumes comments are random. **Probably want to try randomizing them.**
```{}
sms_dtm_train2 <- sms_dtm2[1:6846, ]
sms_dtm_test2 <- sms_dtm2[6847:9129, ]
```

Save vectors labeling rows in the training and testing vectors
```{r}
sms_train_labels2 <- sms_raw2[1:6846, ]$type
sms_test_labels2 <- sms_raw2[6847:9129,]$type
```

Make sure that the proportion of each sub-category is similar in the training and testing data set.
```{r}
prop.table(table(sms_train_labels2))
prop.table(table(sms_test_labels2))
```

####\

##Visualization
Create a wordcloud of the frequency of the words in the dataset using the package "wordcloud".
```{r}
wordcloud(sms_corpus_clean2, max.words = 50, random.order = FALSE)
```

Compare wordclouds between 3 groups.
```{r}
TWO <- subset(sms_raw2, type == "2_Writing")
THREE <- subset(sms_raw2, type == "3_Technical")
FOUR <- subset(sms_raw2, type == "4_Logic")

wordcloud(TWO$text, max.words = 50, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 50, scale = c(3, 0.5))
wordcloud(FOUR$text, max.words = 50, scale = c(3, 0.5))
```

####\

##Preparation for Naive Bayes
Remove words from the matrix that appear less than 5 times.
```{r}
sms_freq_words2 <- findFreqTerms(sms_dtm_train2, 5)
str(sms_freq_words2)
```

Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
```{r}
sms_dtm_freq_train2 <- sms_dtm_train2[ , sms_freq_words2]
sms_dtm_freq_test2 <- sms_dtm_test2[ , sms_freq_words2]
```

The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables. To do this we'll build a convert_counts function and apply it to our data.
```{r}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}
```

This replaces values greater than 0 with yes, and values not greater than 0 with no. Let's apply it to our data.
```{r}
sms_train2 <- apply(sms_dtm_freq_train2, MARGIN = 2, convert_counts2)
sms_test2 <- apply(sms_dtm_freq_test2, MARGIN = 2, convert_counts2)
```

The resulting matrixes will be character type, with cells indicating "yes" or "no" if the word represented by the column appears in the message represented by the row.

####\

##Train Model, Predict, Evaluate
Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group TWO, THREE, or FOUR. Evaluate the prediction with the actual data using a crosstable from the gmodels package.
```{r}
sms_classifier2 <- naiveBayes(sms_train2, sms_train_labels2)
sms_test_pred2 <- predict(sms_classifier2, sms_test2)
CrossTable(sms_test_pred2, sms_test_labels2, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

Show the 5 most frequent words in the data:
```{r}
sack2 <- TermDocumentMatrix(sms_corpus_clean2)
pack2 <- as.matrix(sack2)
snack2 <- sort(rowSums(pack2), decreasing = TRUE)
hack2 <- data.frame(word = names(snack2), freq=snack2)
head(hack2, 10)
```

And the 5 most frequent words from each class:
```{r}
wordcloud(TWO$text, max.words = 10, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 10, scale = c(3, 0.5))
wordcloud(FOUR$text, max.words = 10, scale = c(3, 0.5))
```

***
***

The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables. To do this we'll build a convert_counts function and apply it to our data.
```{r}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}
```

This replaces values greater than 0 with yes, and values not greater than 0 with no. Let's apply it to our data.
```{r}
sms_train2 <- apply(sms_dtm_freq_train2, MARGIN = 2, convert_counts2)
sms_test2 <- apply(sms_dtm_freq_test2, MARGIN = 2, convert_counts2)
```

The resulting matrixes will be character type, with cells indicating "yes" or "no" if the word represented by the column appears in the message represented by the row.

####\

##Train Model, Predict, Evaluate
Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group TWO, THREE, or FOUR. Evaluate the prediction with the actual data using a crosstable from the gmodels package.
```{r}
sms_classifier2 <- naiveBayes(sms_train2, sms_train_labels2)
sms_test_pred2 <- predict(sms_classifier2, sms_test2)
CrossTable(sms_test_pred2, sms_test_labels2, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

Show the 5 most frequent words in the data:
```{r}
sack2 <- TermDocumentMatrix(sms_corpus_clean2)
pack2 <- as.matrix(sack2)
snack2 <- sort(rowSums(pack2), decreasing = TRUE)
hack2 <- data.frame(word = names(snack2), freq=snack2)
head(hack2, 10)
```

And the 5 most frequent words from each class:
```{r}
wordcloud(TWO$text, max.words = 10, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 10, scale = c(3, 0.5))
wordcloud(FOUR$text, max.words = 10, scale = c(3, 0.5))
```

***
***

**Permutation 1**
Did not remove numbers, did not remove stopwords or punctuation, did not stem the terms. Reduced word count limit from 5 to 4 for matrix word limit.

```{}
sms_freq_words2 <- findFreqTerms(sms_dtm_train2, 4)
str(sms_freq_words2)
```

This step raised accuracy for calling 3_Technical comments to 91.9%, but dropped overall accuracy to 79.0%.

***
***

**Permutation 2**
Removed stopwords, numbers and punctuation, but did not use stemming. Also dropped to comparing two groups. This approach raised the accuracy to:

*  87.7% for separating 2_writing and 3_technical. 
*  83.1% for separating 2_writing and 4_logic. 
*  91.2% for separating 3_technical and 4_logic.

```{r}
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
library(readr)
```

```{r}
#Read in CSV file named "coded_full_comments_dataset_Spring18anon.csv".
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')

#Isolate rows to compare. 
frequency_writing <- filter(base_data,code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking") #|code.subject=="2. Writing Quality"

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
sms_raw2 <- frequency_writing %>% select(23,22)

#Rename the columns.
names(sms_raw2)[1] <- "type"
names(sms_raw2)[2] <- "text"

#Simplify coding terms
sms_raw2[,1] <- ifelse(sms_raw2[,1] == "3. Technical and Scientific", "3_Technical", ifelse(sms_raw2[,1] == "4. Logic and Thinking","4_Logic", 99))# ifelse(sms_raw2[,1] == ,99)))"2. Writing Quality", "2_Writing"

#Change "type" element from character to a factor for analysis.
sms_raw2$type <- factor(sms_raw2$type)
str(sms_raw2$type)
table(sms_raw2$type)
```

```{r}
##Text Transformation using "tm" package in R to create a volitile coprus that contains the "text" vector from our data frame.
sms_corpus2 <- VCorpus(VectorSource(sms_raw2$text))
print(sms_corpus2)

#Check out the first few messages in the new corpus, which is basically a list that can be manipulated with list operations.
inspect(sms_corpus2[1:3])

#Use "as.character" function to see what a message looks like.
as.character(sms_corpus2[[3]])
```

In order to standardize the messages, the data set must be tranformed to all lower case letters. The words "Free", "free", and "FREE" should all be treated as the same word. Use the "tm_map"" funtion in R, and use the "content_transformer" function to transform the text.
```{r}
sms_corpus_clean2 <- tm_map(sms_corpus2, content_transformer(tolower))
```

Look at third message again to see if our data was transformed.
```{r}
as.character(sms_corpus2[[3]])

as.character((sms_corpus_clean2[[3]]))
```

This removes numbers. **May not want to do.**
```{r}
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removeNumbers)
```

Stopword removal takes out words that appear often but don't contribute to our objective. These words include "to", "and", "but" and "or". **One of variables to test out.**
```{r}
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removeWords, stopwords())
```

Remove punctuation as well using the "removePunctuation" function. **Try not doing this. Removes evidence of questions.**
```{r}
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removePunctuation)
as.character((sms_corpus_clean2[[3]]))
```

Perform "stemming" to the text data to strip the suffix from words. **Try NOT doing this.** 
```{}
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, stemDocument)
```

And now the final step in text mining is to remove white space from the document. 
```{r}
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, stripWhitespace)
```

```{r}
as.character(sms_corpus_clean2[[3]])
```

Perform tokenization using the "DocumentTermMatrix" function. This creates a matrix in which the rows indicate documents (SMS messages in this case) and the columns indicate words. It should be noted that the "DocumentTermMaxtrix" function has the power to do all of the text mining above in one command.
```{r}
sms_dtm2 <- DocumentTermMatrix(sms_corpus_clean2)
```

####\

##Data Preparation
Split data into training and testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data. Divide our data set into 75% training and 25% testing.

```{r}
.75 * 6551

.25 * 9129
```

This code assumes comments are random. **Probably want to try randomizing them.**
```{r}
sms_dtm_train2 <- sms_dtm2[1:4913, ]
sms_dtm_test2 <- sms_dtm2[4914:6551, ]
```

Save vectors labeling rows in the training and testing vectors
```{r}
sms_train_labels2 <- sms_raw2[1:4913, ]$type
sms_test_labels2 <- sms_raw2[4914:6551,]$type
```

Make sure that the proportion of each sub-category is similar in the training and testing data set.
```{r}
prop.table(table(sms_train_labels2))
prop.table(table(sms_test_labels2))
```

####\

##Visualization
Create a wordcloud of the frequency of the words in the dataset using the package "wordcloud".
```{r}
wordcloud(sms_corpus_clean2, max.words = 50, random.order = FALSE)
```

Compare wordclouds between 3 groups.
```{r}
#TWO <- subset(sms_raw2, type == "2_Writing")
THREE <- subset(sms_raw2, type == "3_Technical")
FOUR <- subset(sms_raw2, type == "4_Logic")

#wordcloud(TWO$text, max.words = 50, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 50, scale = c(3, 0.5))
wordcloud(FOUR$text, max.words = 50, scale = c(3, 0.5))
```

####\

##Preparation for Naive Bayes
Remove words from the matrix that appear less than 5 times.
```{r}
sms_freq_words2 <- findFreqTerms(sms_dtm_train2, 5)
str(sms_freq_words2)
```

Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
```{r}
sms_dtm_freq_train2 <- sms_dtm_train2[ , sms_freq_words2]
sms_dtm_freq_test2 <- sms_dtm_test2[ , sms_freq_words2]
```

The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables. To do this we'll build a convert_counts function and apply it to our data.
```{r}
convert_counts2 <- function(x) {
  x <- ifelse(x > 0, "Yes", "No")
}
```

This replaces values greater than 0 with yes, and values not greater than 0 with no. Let's apply it to our data.
```{r}
sms_train2 <- apply(sms_dtm_freq_train2, MARGIN = 2, convert_counts2)
sms_test2 <- apply(sms_dtm_freq_test2, MARGIN = 2, convert_counts2)
```

The resulting matrixes will be character type, with cells indicating "yes" or "no" if the word represented by the column appears in the message represented by the row.

####\

##Train Model, Predict, Evaluate
Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group TWO, THREE, or FOUR. Evaluate the prediction with the actual data using a crosstable from the gmodels package.
```{r}
sms_classifier2 <- naiveBayes(sms_train2, sms_train_labels2)
sms_test_pred2 <- predict(sms_classifier2, sms_test2)
CrossTable(sms_test_pred2, sms_test_labels2, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

Show the 10 most frequent words in the data:
```{r}
sack2 <- TermDocumentMatrix(sms_corpus_clean2)
pack2 <- as.matrix(sack2)
snack2 <- sort(rowSums(pack2), decreasing = TRUE)
hack2 <- data.frame(word = names(snack2), freq=snack2)
head(hack2, 10)
```

And the 10 most frequent words from each class:
```{r}
wordcloud(TWO$text, max.words = 10, scale = c(3, 0.5))
wordcloud(THREE$text, max.words = 10, scale = c(3, 0.5))
#wordcloud(FOUR$text, max.words = 10, scale = c(3, 0.5))
```

### 2019-02-25 hiding/Round_1_Data_intake_writing.Rmd
updates

---
title: "Round 2 for Data Intake Method"
author: "Dan Johnson"
date: "11/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Preparing and Pre-Checking the Initial Dataset

Call in tidyverse, other libraries.
```{r Libraries}
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
```

CSV file named "base_data.csv" already has correctly organized text and meta-data. First block reads in CSV, then filters and pulls rows and columns to a new working file.

```{r}
# Read CSV.
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18.csv')
head(base_data)

# Isolate table rows to compare.
frequency_writing <- filter(base_data, code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")

#Reduce larger dataframe to just 3 required columns of data  
frequency_writing.subcolumns <- frequency_writing %>% select(1, 22:24)
```

Next block tokenizes the TA comments as individual words and removes stop words.

```{r}
# Tokenize phrases, remove stop words listed in standard reference file.
base_data_tokenized_writing <- frequency_writing.subcolumns %>% 
  unnest_tokens(word,ta.comment) %>% 
  anti_join(stop_words)
```

At this point the four-column tibble has "UniqueID" in column 1, "code.subject" in column 2, "code.structure" in column 3, and the unnested "words" in column 4. All four columns are class = "character". Stop words from the R "stop_words" data file have been removed.

```{r}
#Groups by code.subject, calculated proportional frequency in each group.
#Final step creates 4 columns; 2 are for 1. Basic Criteria data, others for 
#remaining three data sets. Stats will compare 1. BC against others.
base_data_tokenized_writing_sorted <- base_data_tokenized_writing %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  spread(code.subject, proportion) %>% 
  gather(code.subject, proportion, `3. Technical and Scientific`:`4. Logic and Thinking`)
```

"base_data_tokenized_writing_sorted" is the final 4-column dataframe with individual words in column 1, "1. Basic Criteria" proportion in column 2, "code.subject" in column 3, and "proportion" in column 4. Columns 1 and 3 are class = "character", 2 and 4 are numeric ranges.

This structure is needed in order to compare word (in Column 1) frequencies for 1. Basic Criteria (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes. Which subset of the X values to use is coded in Column 3.

## Analysis
These three steps determine the Pearson correlation between the frequency of words in Basic Criteria versus the other three groups.

```{r}
cor.test(data = base_data_tokenized_writing_sorted[base_data_tokenized_writing_sorted$code.subject == "3. Technical and Scientific",], 
         ~ proportion + `2. Writing Quality`)
```

```{r}
cor.test(data = base_data_tokenized_writing_sorted[base_data_tokenized_writing_sorted$code.subject == "4. Logic and Thinking",], 
         ~ proportion + `2. Writing Quality`)
```

If the words used by TAs in comments are not different between the categories, there should be fairly high correlation.

The final step generates a graphical comparison of word frequencies.

```{r}
# Expect a warning about rows with missing values being removed
ggplot(base_data_tokenized_writing_sorted, aes(x = proportion, y = `2. Writing Quality`, color = abs(`2. Writing Quality` - proportion))) +
  geom_abline(color = "gray40", lty = 2) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
  facet_wrap(~code.subject, ncol = 1) +
  theme(legend.position="none") +
  labs(y = "Writing Quality", x = NULL)
  #ggsave("plot.png", width = 6, height = 10)
```

If the words used by TAs in comments are not different between the categories, the slopes of the lines in the three comparison graphs should be close to 1, with intercepts approaching zero.

### 2019-02-25 litreview.Rmd
update
updates

@@
 #Overview
-This page links out to sub-pages that describe various background topics. Follow the links for details.
+The supporting literature for this project is broad and convoluted.
+Rather than try to compress it to a single page, I have provided links out to sub-pages that describe various background topics.
+Follow the links for details.
 
-Description: https://adanieljohnson.github.io/default_website/pagename.html
+Fundamental Principles and Practices of Technical Writing Training:  https://adanieljohnson.github.io/default_website/pagename.html
 
-Technical Writing Training
-
-Text Classification Strategies
+Primer on Text Classification Strategies:  https://adanieljohnson.github.io/default_website/Text_classifier_models.html
 

### 2019-02-25 hiding/NB_2r2c.Rmd

--- 2019-02-22 NB_2r2c.Rmd
@@
 ---
-title: "Naive Bayes Analysis Ver 2r.2c"
+title: "Naive Bayes Classifier"
 author: "Dan Johnson"
@@
 
-##Summary
-This it the current Naive Bayes workflow used to classify TA comments based on Subject and Structure. Additions and revisions made to **Version 2r.2c** are:
+#Overview
+This is the current version of the Naive Bayes workflow used to classify TA comments based on Subject and Structure. Features of **Version 2r.2c**:
 
-*  Random selection of data rows used for training and test datasets. 
-*  Final output matrix tables are more informative. 
-*  Addition of a back-analysis step that reviews entire dataset, not just testing sub-set.
-*  Outputs a tidy data table in CSV format that contains the original comments (pre-clean-up) plus their hand-scored and predicted subject and structure.
+*  Data rows used for training and test datasets are assigned randomly but reproducibly using 'set.seed'. 
+*  Detailed final output matrix tables. 
+*  Back-analysis evaluates both a testing sub-set and the entire input dataset.
+*  Outputs a tidy data table in CSV format with unmodified original comments, and hand-scored and NB-predicted Subject and Structure assignments.
+
+This page outlines Ver 2r.2c of NB Classifier and the process for building various components used in the project. Details of each  component are provided in these sub-pages.
+
+For structure of the initial dataset see: https://adanieljohnson.github.io/default_website/datastructure.html
+
+For import and post-import data validation see: https://adanieljohnson.github.io/default_website/dataimport.html
 
@@
 
-##Initial Setup
-Calling the required libraries.
-
+#Initial Setup
+Calling required libraries.
 ```{r libraries}
@@
 
-####\
-
-##Version 2.0.r of Naive Bayes (NB) Protocol
-###Initial Data Input
+Import the initial dataset. Then extract comments and classification data as subsets of CSV.
 ```{r input data}
@@
 base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
-```
 
-###Extract, Organize Data Subset
-Extract comments and classification data as subsets of CSV.
-```{r subset data}
 #Select rows representing the sub-groups to compare. 
@@
 
+Randomize data used for training and testing sets. This block of code generates a reproducible vector of ~10,000 randomly assorted numbers using set.seed, then writes the vector back to data table.
 ```{r randomizer}
-#Generates a reproducible vector of ~10,000 randomly assorted numbers using set.seed 
 #Adjust the number so it matches the number of rows in the original data frame
@@
 
+Convert data set to a volative corpus using "tm" library.
 ```{r corpusmaker}
-#Convert data set to a volative corpus using "tm" library.
 #Create the volatile corpus that contains the "text" vector from data frame.
@@
 
-###Text Data Transformations
-The text data transforms are separated out as individual "tm_map"" and "content_transformer" functions to simplify the process of testing various combinations.
-Conversion to lower case and removing punctuation are the baseline set.
-Switch between "{r}" and "{}" to turn the other transformation blocks off or on.
-Removing extra white spaces must be the last transformation, and is required.
+#Text Data Transformations
+The text data transforms in this code block are intentially separate "tm_map"" and "content_transformer" functions.
+This simplifies the process of testing how various combinations affect classification.
+Converting the text to lower case and removing punctuation are the baseline transformations.
+Switch between "{r}" and "{}" to turn other transformation blocks off or on as desired.
+**Removing extra white spaces must be the last transformation, and is required.**
 
@@
 
-#The option to apply stemming has been removed from this version.
+#The option to apply stemming has been removed from this version. Stemming reduced accuracy in all trials. Future iterations should examine lemmatization instead.
 
@@
 
-#Look at an example of cleaned text comment to see if it fits what is expected.
+#Look at an example of cleaned text comment to see if cleaned comment matches what is expected. Change the value inside double brackets to look at a different comment.
 as.character((comments_corpus_cleanVERB_R[[5]]))
@@
 
-##Basic Naive Bayes (NB) Method Using N-grams
-This code block uses tm's NLP commands. The same block can be used to assess 1-, 2-, or 3-grams by modifying the "#" variable in "(words(x), #)".
+#Tokenize the N-Grams
+This code block uses tm's NLP commands to assess different n-grams.
+For single words, change the "#" variable in "(words(x), #)" to "1".
+Use other values to change the n-gram size.
+Data are stored in document-term matrix (dtm) format.
 
@@
 
-###Data Splitting
-Split data into 75% training and 25% testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data.
+#Split Data to Training, Testing Sets
+This code splits data into 75% training and 25% testing sets, so that after Naive Bayes algorithm is trained it can be tested on unseen data.
 
@@
 
-Saves 2 vectors that label the rows in the training and testing vectors
+This saves 3 vectors containing labels for the rows in the training and testing vectors
 ```{r labeler}
@@
 
+This checks the proportions of sub-categories in training and testing groups.
 The proportion of each sub-category should be similar (<1% difference) between the full dataset, and training and testing data subsets.
@@
 
-###Preparation for Naive Bayes
-Remove words from the matrix that appear less than 5 times.
+#Preparing Data for Naive Bayes Training
+Remove words from the document-term matrix that appear less than 5 times.
 ```{r common_words}
@@
 
-Limit Document Term Matrix to only include words in the comments_freq_words vector. Using all the rows, but we want to limit the columns to these words in the frequency vector.
+Limit document-term matrix to words in the comments_freq_words vector. We are using all of the rows, but we want to limit the columns to these words in the frequency vector.
 ```{r limit_dtm}
@@
 
-The basic e1071 naive bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables.
+The e1071 Naive Bayes classifier works with categorical features, so the matrix must be converted "yes" and "no" categorical variables.
 This is done using a **convert_counts function** and applying it to the data.
@@
 
+Below is an **alternate version** of convert_function code block that KEEPs word frequencies while still using the same variable names and code structure.
+The processed matrices have cells with word frequencies instead of categorical yes/no values.
+This block only works for binary classification (ingroup/outgroup).
+It breaks when applied to data with multiple categories.
 ```{}
-# Alternate version of convert_function code block.
-This version KEEPs word frequencies while still using the same variable names and code structure.
-The resulting matrices have cells with word frequencies instead.
-#Only works when the NB classifier is looking at ingroup/outgroup. Breaks on multiple categories.
-#
 convert_counts2 <- function(x) {
@@
 
+Applies whichever "convert_counts2" function is currently active.
 ```{r apply_conversion}
@@
 
-###Train Model, Predict, Evaluate for Subject
-This block uses the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic.
-It evaluates the prediction with the actual data using a crosstable from the gmodels package.
+#Train Model, Predict, Evaluate for Subject
+This block predicts whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. 
 ```{r bayes_training}
-# Train the classifier on training data, then test on test subset and full dataset.
+# Train the classifier on training data, then test on test dataset, and full dataset.
 comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject, laplace=1)
@@
 comments_all_pred_structure <- predict(comments_classifier_structure, comments_all)
+```
 
+A truth table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting Subject. 
+```{r}
 # Create a truth table by tabulating the predicted class labels with the actual class labels 
@@
 
-#The truth table lists how many of the predicted categories are in their correct categories.
-Adding diagonally and dividing by total count gives the percent accuracy at predicting subject.
-this is what the confusion matrix calculates.
-
-#The faster way is to calculate a confusion matrix, which includes BOTH the truth table, and the final overall accuracy.
-
-#Alternatively, the Cross Table command below (from "gmodels" package) generates a more detailed output table with frequency values for each individual cell.
+#An alternative is to use the Cross Table command (from "gmodels" package) to generate a more detailed output table with frequency values for each individual cell.
 #CrossTable(comments_test_pred_subject, comments_test_labels_subject, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
@@
 
+The faster method is to calculate a confusion matrix, which displays BOTH the truth table, and the final overall accuracy. A confusion matrix can be queried for specific values.
 ```{r reportOut}
-#"line" is just a spacer text for readability.
+#"line" is just spacer text for improving readability.
 line<-("**************************************************************")
@@
 
-# This print-out has both the overall and by-group accuracy values.
+#These commands show how to print out individual pieces of the confusion matrix. They show both the overall and by-group accuracy values for Subject x Test Subset.
 conf.mat1
@@
 print ("Test, Subject")
-# This print-out has by-group accuracy values in a single column. 
+
+# This block outputs the by-group accuracy values in a single column for Subject x Test Subset. 
 conf.mat1$byClass
@@
 
-# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
+# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Subject x Test Subset.
 conf.mat1$overall
@@
 
-# This outputs JUST the overall accuracy value. 
+# This block outputs JUST the overall accuracy value for Subject x Test Subset. 
 conf.mat1$overall['Accuracy']
@@
 
-# This print-out has both the overall and by-group accuracy values.
+# This block outputs both the overall and by-group accuracy values for Subject x All Data.
 conf.mat2
@@
 
-# This print-out has by-group accuracy values in a single column. 
+# This block outputs by-group accuracy values in a single column for Subject x All Data. 
 conf.mat2$byClass
@@
 
-# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
+# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Subject x All Data.
 conf.mat2$overall
@@
 
-# This outputs JUST the overall accuracy value. 
+# This block outputs JUST the overall accuracy value for Subject x All Data. 
 conf.mat2$overall['Accuracy']
@@
 
-# This print-out has both the overall and by-group accuracy values.
+# This block outputs both the overall and by-group accuracy values for Structure x Test Subset.
 conf.mat3
@@
 
-# This print-out has by-group accuracy values in a single column. 
+# This block outputs by-group accuracy values in a single column for Structure x Test Subset. 
 conf.mat3$byClass
@@
 
-# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
+# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Structure x Test Subset.
 conf.mat3$overall
@@
 
-# This outputs JUST the overall accuracy value. 
+# This block outputs JUST the overall accuracy value for Structure x Test Subset. 
 conf.mat3$overall['Accuracy']
@@
 
-# This print-out has both the overall and by-group accuracy values.
+# This block outputs both the overall and by-group accuracy values for Structure x All Data.
 conf.mat4
@@
 
-# This print-out has by-group accuracy values in a single column. 
+# This block outputs by-group accuracy values in a single column for Structure x All Data. 
 conf.mat4$byClass
@@
 
-# This print-out is a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value.
+# This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Structure x All Data.
 conf.mat4$overall
@@
 
-# This outputs JUST the overall accuracy value. 
+# This block outputs JUST the overall accuracy value for Structure x All Data. 
 print ("All, Structure")
 conf.mat4$overall['Accuracy']
-
 ```
@@
 
-###Output for Review, Analysis
+#Output for Review, Analysis
 The code block below creates a CSV file containing the original TA comments, the subject and structure labels assigned by a human observer, and the corresponding predicted categories for each comment based on the Naive Bayes estimator.
@@
 
-***
-***
+This is an alternate method to get the same data tables out using ```tidy``` format. It seems much longer to me, so likely is over-written.
+```{}
+tidy_prediction_table
 
-This is an alternate method to get the same data tables out using ```tidy``` format. It seems much longer to me, so likely is over-written.
-```{r tidy_prediction_table}
 #Use broom::tidy to reorganize into a data frame.
@@
 spread(prop_tally4,key=Bayes.Predicted.Subject, value=prop)
-```
 
-```{r}
 #Calculate the sum of values for each group.
@@
 proptallysum4
-
 ```
@@
 
+***
 Need to combine two datasets into one. 
+***
 

### 2019-02-25 docs/hiding/Check_in_Jan_14.Rmd
Web site update

--- 2019-02-25 hiding/Check_in_Feb_4.Rmd
@@
 ---
-title: "Feb 4 Check-In"
+title: "Jan 14 Check-In"
 author: "Dan Johnson"
-date: "2/3/19"
+date: "1/29/19"
 output: html_document
@@
 
-#New ideas From FLC Meeting #6
-
-Need help with:
-
-Coding issues
-
-*  How to ID, prune unneeded packages. NO SIMPLE SOLUTION. Cut and test seems best.
-
-*  When I join columns in correlation analysis, if a word is MISSING from **any subset** the data for that word is pruned from the full dataset. I am losing the unique words this way. 
-    +  Assume the NA would blow up correlations, which is why it is done this way.
-    +  Need to figure out how to join without pruning, then find the UNIQUE words.
-
-Post-Processing issues. Use broom package
-
-*  How to calculate the value automatically in the Naive Bayes final step.
-Data post-processing.
-I have stable code for my binomial naive Bayes classifier, and I am screening different permutations to see what the best approach is to refining the model.
-There is one step I cannot figure out how to code though.My primary output is a cross-comparison table showing what fraction of the test set comments is correctly identified.
-Adding the counts diagonally and dividing by the total number of test cases gives me the overall accuracy.The table has all of the data that I want to use for other calculations, but I cannot figure out how to either: 1) call values from the table, or 2) pull
-
- Output issues
-
-*  How to take the Excel/csv tables I have for permutations of naive Bayes, Pearson's r, and bring them in for display in summary page. broom
-
-Add to My Code:
-
-*  shell(cmd, shell, flag = "/c", intern = FALSE, wait = TRUE,
-      translate = FALSE, mustWork = FALSE, ...) runs UNIX commands
-https://www.rdocumentation.org/packages/base/versions/3.4.0/topics/shell
-
-*  Using csvy format. R supports .csvy using csvy package and read_csvy() and write_csvy()
-Reference Sites:
-
-*  Using broom for cleanup
-https://cran.r-project.org/web/packages/broom/vignettes/broom.html
-
-Packages to Explore:
-
-*  fs. https://fs.r-lib.org/index.html
-
-fs provides a cross-platform, uniform interface to file system operations.
-It shares the same back-end component as nodejs, the libuv C library, which brings the benefit of extensive real-world use and rigorous cross-platform testing.
-fs functions are divided into four main categories:
-
-    path_ for manipulating and constructing paths
-    file_ for files
-    dir_ for directories
-    link_ for links
-fs is designed to work well with the pipe, though because it is a minimal-dependency infrastructure package it doesn’t provide the pipe itself.
-You will need to attach magrittr or similar.
-fs functions also work well in conjunction with other tidyverse packages, like dplyr and purrr.
-
-To Do List:
-
-*  Best practices when writing R code
-https://swcarpentry.github.io/r-novice-inflammation/06-best-practices-R/
-
-#File Management
-
-```{r}
-headlines2 <- read_docx("R_0eqZDDQ0nG2obuh_text.docx") %>% head(9)
-View(headlines2)
-```
-
-R can manipulate files and folders. This is a primer:
-http://theautomatic.net/2018/07/11/manipulate-files-r/
-
-Get, change working directory
-Create files of any type based on extension
-Copy a file or folder
-List files in a directory, including by sub-type
-Use rowbind to join file contents
-Get snapshot data
-Remove files, folders
-Check for files, folders
-Get the basenames for files, folders
-Get file extensions
-Call shell to open the file
-Move a file
-
-# read in all the CSV files
-all_data_frames <- lapply(list.files(pattern = ".csv"), read.csv)
-
-# stack all data frames together
-single_data_frame <- Reduce(rbind, all_data_frames)
-
-Default
-draft
-General_workflow
-initial_exploration
-Naive_bayes_V2b, r, yn
-pre_work
-project
-questions
-reference_tables
-course_notes
-
-ta_comments_FLC
-Text Classifier Models
-Random_forest_notes
-Latent_dirichlet_allocation_notes
-  folder=data
-    codebook_for_comments.md
-    code_column_names.md
-    TA_and_Student_IDs.csv
-    ID_lookup.csv
-  folder=docs_and_notes
-    Grading Criteria.rmd
-    Raw_data_structure.rmd
-
-###Planned Project Pages
-It may seem excessive, but my thinking is that:
-1. Our group is the first so should establish some guidelines for creating tools.
-2. Work I put in here translates over to the final, shared product for NSF project.
-
-*  toc = Table of Contents
-*	 project = Project Summary
-
-*	 Background
-    +  Theoretical background on text classifiers
-    +  Practical background on project challenges
-
-*	 Project Methods and Guides
-    +  Anonymizing Data
-    +  Standard Workflow
-    +  Reference Tables
-        -  Recoding lists
-    +  Project Codebook
-
-*  Pages template:
-    +  Page Goals
-    +  Page Overview
-    +  Page Content
-
-*  Initial Exploration
-    +  Data Processing
-    +  Preliminary Visualization
-
-*  Deeper Analyses
-    +  Categorical Naive Bayes
-        -  Baseline cNB
-        -  Randomized cNB
-        -  Binary cNB
-        -  Summary of cNB Findings
-
-***
-***
-
-#Jan 14 Check-In, and New ideas From FLC Meeting #5
+#New ideas From FLC Meeting #5
 
@@
 
-Suggested: 
-Check my data for small errors that break correlations.
+Solved the data confidentiality problem. Turned out to be easier and more secure to do outside of R.
+
+Manipulating the column vectors.
+  Suggested: I need to recode some values to simplify. 
+Second, in this same correlation analysis, I would recommend recoding the values in the `code.subject` column.
+You have some type of error there and I tend to avoid calling values such as "2.
+Writing Quality" as they tend make code more difficult to visually parse (and therefore debug) and it's also easy to get something wrong with the value (even though it seems that you've directly copied and pasted the values to double check).
+
+For recoding comment labels, this is what I plan to use:
+
+Recoding Comment.subject
+0_null
+1_basic
+2_writing
+3_technical
+4_logic
+5_exposition
+6_misconduct
+7_administrative
+8_nobasis
+9_split
+10_context
+11_flag
+12_narrative
+13_sciname
+
+Recoding Comment.structure
+0_null
+1_pointer
+2_copyedit
+3_general
+4_specific
+5_rationale
+6_holistic
+7_idiomatic
+8_nobasis
+
+Calculation and plotting correlations and log-odds pairs is working (except for broken parts that recoding should solve hopefully.)
+
+Suggested: Check my data for small errors that break correlations.
 	"cor" blows up on missing data
@@
 
+Have an operational categorical naive Bayes workflow that can perform analysis using 1-, 2-, or 3-grams.
+
 ## Coding & Interpretation Questions
@@
 1.  Randomization. Does the straight split I used bias the data? 
-  Comment: YES, so instead use set.seed to pick more randomized dataset.
+  Comment: YES, so instead use the following code to pick more randomized dataset.
+
+```{r}
+library(dplyr)
+set.seed(1234) #randomization will be same set each time
+train <- sample_frac(mtcars, .8) #sets how many are in random training set
+test <- setdiff(mtcars, train) #chooses non-training for test set
+```
 
@@
 Anyway, things are looking good. Keep up the good work and great documentation!
+
+    FLC - Data Analysis (AB)
+
+My project is still taking shape (and isn't yet really research, per se), but a few things I've been working on:
+
+    Project Updates - Updating some content on my Project page
+    New Page - Learning Analytics. This is a proposal I have had accepted for an April conference, and I am hoping our group might help me develop some of my thinking on the issue.
+    Course Notes - I haven't kept up with note taking in this space, but added some of the detail from my work during Session 4.
+I am looking forward to engaging more actively with this space during our actual sessions, and I am interested in hearing about others' thoughts on this process.
+
+While there are many more updates I hope to make this term, I wanted to share some of this work. You might also check out Dan's page (as his is more robust than mine).
+
+To-Do
+
+    Prepare to present on your progress at Monday's Session
+    Update your website (or have a plan to do so).
+    Raise questions (technical or otherwise) in the TLC discussion board space.
+
+I am looking forward to seeing everyone on Monday (assuming we aren't snowed out).
+
+2.  How to ID, prune unneeded packages.
+
+3.
+When I join columns in correlation analysis, when a word is MISSING from **any subset** the data for that word is pruned from the full dataset.
+I am losing the unique words this way.
+Need to figure out how to join without pruning.
+
+6.  Broken correlation analyses. I have a problem buried in the coded data that is throwing an error in correlations. Jerid recommended recoding the values in the `code.subject` column.
 

### 2019-02-25 docs/hiding/Naive_bayes_notes3.Rmd
Web site update

--- 2019-02-25 hiding/Naive_bayes_notes4 (original).Rmd
@@
 
-##Starting Version for Bigrams, Trigrams
-The original code is in rows 22 to ~235 of of Version 2 of this document. Works as of 12/18/2018. This is modified for n-grams instead of single words.
+##Starting Version 1.0
+Rows 22 to ~235 of this code are original starting version. Works as of 12/18/2018
 
@@
 sms_corpus2 <- VCorpus(VectorSource(sms_raw2$text))
+print(sms_corpus2)
 
-#Convert to all lower case letters with "tm_map"" funtion in R, then use the "content_transformer" function to transform the text.
+#Check out the first few messages in the new corpus, which is basically a list that can be manipulated with list operations.
+inspect(sms_corpus2[1:3])
+
+#Use "as.character" function to see what a message looks like.
+as.character(sms_corpus2[[3]])
+```
+
+In order to standardize the messages, the data set must be tranformed to all lower case letters.
+The words "Free", "free", and "FREE" should all be treated as the same word.
+Use the "tm_map"" funtion in R, and use the "content_transformer" function to transform the text.
+```{r}
 sms_corpus_clean2 <- tm_map(sms_corpus2, content_transformer(tolower))
+```
 
-#This removes numbers. **May not want to do.**
+Look at third message again to see if our data was transformed.
+```{r}
+as.character(sms_corpus2[[3]])
+
+as.character((sms_corpus_clean2[[3]]))
+```
+
+This removes numbers. **May not want to do.**
+```{r}
 sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removeNumbers)
+```
 
-#Stopword removal. **One of variables to test out.**
+Stopword removal takes out words that appear often but don't contribute to our objective. These words include "to", "and", "but" and "or". **One of variables to test out.**
+```{r}
 sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removeWords, stopwords())
+```
 
-#Remove punctuation as well using the "removePunctuation" function. Try not doing this. Removes evidence of questions.
+Remove punctuation as well using the "removePunctuation" function. **Try not doing this. Removes evidence of questions.**
+```{r}
 sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removePunctuation)
+as.character((sms_corpus_clean2[[3]]))
 ```
@@
 Perform "stemming" to the text data to strip the suffix from words. **Try NOT doing this.** 
-```{}
+```{r}
 sms_corpus_clean2 <- tm_map(sms_corpus_clean2, stemDocument)
@@
 
-Final step in text mining is to remove white spaces. 
+And now the final step in text mining is to remove white space from the document. 
 ```{r}
@@
 
-Code to inspect a line of text.
 ```{r}
@@
 
-#Single word tokenization
-Block tokenizes using the "DocumentTermMatrix" function.
+**Original Version**
+Performs tokenization using the "DocumentTermMatrix" function.
 This creates a matrix in which the rows indicate documents (SMS messages in this case) and the columns indicate words.
 It should be noted that the "DocumentTermMaxtrix" function has the power to do all of the text mining above in one command.
-The "tidy" command is a hack that makes it easier to inspect the document terms matrix.
 ```{r}
@@
 
-#N-gram tokenization
-This code block uses tm's NLP functions to generate n-grams from the full corpus of TA comments
+**Quick code block to generate 2-grams and 3-grams from full corpus of TA comments**
 
 ```{r}
-
-NLP_Tokenizer <- function(x) {
-      unlist(lapply(ngrams(words(x), 2), paste, collapse = " "), use.names = FALSE)
-}
-
-sms_dtm2_bigrams <- TermDocumentMatrix(sms_corpus_clean2, control=list(tokenize = NLP_Tokenizer))
-
-inspect(sms_dtm2_bigrams)
-
-```
-
-This should build an inspectable doc-terms matrix using n-grams described in "sms_dtm2_bigrams"
-```{r}
-sms_dtm2_bigrams_tidy <- tidy(sms_dtm2_bigrams)
-
-#sms_dtm_trigram <- DocumentTermMatrix(sms_corpus_clean2, control = list(tokenize=tdm_NLP))
-#sms_dtm_trigram_tidy <- tidy(sms_dtm_trigram)
-```
-
-Data Tables for Analysis
-sms_dtm2
-sms_dtm2_bigrams
-sms_corpus_clean2
-
-**Along the way I picked up this quick code block for generating 2-grams and 3-grams from full corpus of TA comments using the "corpus"" library**
-
-```{}
 library(corpus)
@@
 
-Data Tables for Analysis
-sms_dtm2
-sms_dtm2_bigrams
-sms_dtm_trigram
+text.corpus = sms_corpus_clean2
 
-```{}
+**This code block used tm's NLP generate n-grams from full corpus of TA comments**
+
+```{r}
+
+NLPtrigramTokenizer <- function(x) {
+      unlist(lapply(ngrams(words(x), 3), paste, collapse = " "), use.names = FALSE)
+}
+
+tdm_NLP <- TermDocumentMatrix(sms_corpus_clean2, control=list(tokenize = NLPtrigramTokenizer))
+
+inspect(tdm_NLP)
+
+```
+
+This should build a doc-terms matrix tokenize using n-grams 
+```{r}
+library(topicmodels)
+library(RWeka)
+corpus <- sms_corpus_clean2 
+text_filter(corpus)$drop_punct==TRUE
+term_stats(corpus, ngrams = 2:3)
+
+sms_dtm_bigram <- DocumentTermMatrix(sms_corpus_clean2, control = list(tokenize=term_stats))
+sms_dtm_bigram_tidy <- tidy(sms_dtm_bigram)
+```
+
 **OPTION**
@@
   inspect(removeSparseTerms(tdm[, 1:10], 0.7))
-```
+
 ####\
@@
 ```{r}
-sms_dtm_train2 <- sms_dtm2[1:6846, ]
-sms_dtm_test2 <- sms_dtm2[6847:9129, ]
-```
-
-This code assumes comments are random. **Probably want to try randomizing them.**
-```{r}
-sms_dtm_train2 <- sms_dtm2_bigrams[1:6846, ]
-sms_dtm_test2 <- sms_dtm2_bigrams[6847:9129, ]
-```
-
-Save vectors labeling rows in the training and testing vectors
-```{r}
-sms_train_labels2 <- sms_raw2[1:6846, ]$type
-sms_test_labels2 <- sms_raw2[6847:9129,]$type
-```
-
-Make sure that the proportion of each sub-category is similar in the training and testing data set.
-```{r}
-prop.table(table(sms_train_labels2))
-prop.table(table(sms_test_labels2))
-```
-
-####\
-
-##Visualization
-Create a wordcloud of the frequency of the words in the dataset using the package "wordcloud".
-```{r}
-wordcloud(sms_corpus_clean2, max.words = 50, random.order = FALSE)
-```
-
-Compare wordclouds between 3 groups.
-```{r}
-TWO <- subset(sms_raw2, type == "2_Writing")
-THREE <- subset(sms_raw2, type == "3_Technical")
-FOUR <- subset(sms_raw2, type == "4_Logic")
-
-wordcloud(TWO$text, max.words = 50, scale = c(3, 0.5))
-wordcloud(THREE$text, max.words = 50, scale = c(3, 0.5))
-wordcloud(FOUR$text, max.words = 50, scale = c(3, 0.5))
-```
-
-####\
-
-##Preparation for Naive Bayes
-Remove words from the matrix that appear less than 5 times.
-```{r}
-sms_freq_words2 <- findFreqTerms(sms_dtm_train2, 5)
-str(sms_freq_words2)
-```
-
-Data Tables for Analysis
-sms_dtm2 = sms_dtm2_bigrams,  sms_dtm_trigram
-sms_dtm_train2, _test2
-sms_train_labels2, sms_test_labels2
-
-Data Tables for Analysis
-sms_dtm2
-sms_dtm2_bigrams
-sms_corpus_clean2
-
-Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
-```{r}
-sms_dtm_freq_train2 <- sms_dtm_train2[ , sms_freq_words2]
-
-sms_dtm_freq_test2 <- sms_dtm_test2[ , sms_freq_words2]
-```
-
-***
-***
-
-This code assumes comments are random. **Probably want to try randomizing them.**
-```{}
 sms_dtm_train2 <- sms_dtm2[1:6846, ]
@@
 ```
-
-The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables.
-To do this we'll build a convert_counts function and apply it to our data.
-```{r}
-convert_counts2 <- function(x) {
-  x <- ifelse(x > 0, "Yes", "No")
-}
-```
-
-This replaces values greater than 0 with yes, and values not greater than 0 with no. Let's apply it to our data.
-```{r}
-sms_train2 <- apply(sms_dtm_freq_train2, MARGIN = 2, convert_counts2)
-sms_test2 <- apply(sms_dtm_freq_test2, MARGIN = 2, convert_counts2)
-```
-
-The resulting matrixes will be character type, with cells indicating "yes" or "no" if the word represented by the column appears in the message represented by the row.
-
-####\
-
-##Train Model, Predict, Evaluate
-Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group TWO, THREE, or FOUR.
-Evaluate the prediction with the actual data using a crosstable from the gmodels package.
-```{r}
-sms_classifier2 <- naiveBayes(sms_train2, sms_train_labels2)
-sms_test_pred2 <- predict(sms_classifier2, sms_test2)
-CrossTable(sms_test_pred2, sms_test_labels2, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
-```
-
-Show the 5 most frequent words in the data:
-```{r}
-sack2 <- TermDocumentMatrix(sms_corpus_clean2)
-pack2 <- as.matrix(sack2)
-snack2 <- sort(rowSums(pack2), decreasing = TRUE)
-hack2 <- data.frame(word = names(snack2), freq=snack2)
-head(hack2, 10)
-```
-
-And the 5 most frequent words from each class:
-```{r}
-wordcloud(TWO$text, max.words = 10, scale = c(3, 0.5))
-wordcloud(THREE$text, max.words = 10, scale = c(3, 0.5))
-wordcloud(FOUR$text, max.words = 10, scale = c(3, 0.5))
-```
-
-***
-***
 

### 2019-02-25 docs/hiding/Naive_bayes_notes4.Rmd
Web site update

--- 2018-12-21 Naive_bayes.Rmd
@@
 ```
-
 ##Summary
 This notebook documents a basic workflow for applying a naive Bayes model to terms in TA comments.
-I based it on a tutorial [https://rpubs.com/Billyhansen6/318412](located here).
-This is quoted from Hansen's original tutorial:
->The objective of this project is to classify SMS messages as spam or ham (not spam).
-A Naive Bayes classifier approach will be used.
-This example is taken from chapter 4 of Machine Learning with R, Second Edition"
-
->An example of the conditional probability that will be computed is as follows:
-
-$$P(Spam|Hospital) = P(Hospital|Spam)P(Spam)/P(Hospital)$$
-
->which is the formula for determining the probability that a message is spam given that it contains the word "Hospital" in the message.
-
-Lines ~30-205 are the baseline version of the workflow using single words.
-It correctly classifies comments in the top 4 sub-categories  ~79% of the time.
-While that is close to minimum acceptable inter-rater reliability of two human readers, it is well below my goal of >95% for use in routine automated analysis.
-
-Lines ~208-290 describe an alternative method using n-grams. When I tested this approach using 2-grams, classification was slightly worse, at 68.3% correct.
-
-Lines ~295-335 show how to use word clouds to visualize single word frequencies. I did not find this approach very useful.
-
-To keep this code clean for now I will run various permutations in a scratch version, then summarize findings starting on line #345. In these subsequent iterations I will evaluate:
-
-*  Different text pre-processing strategies: stemmed versus unstemmed, numbers removed vs. not, etc.
-*  Using n-grams rather than single words
-*  Making pair-wise comparison versus comparing multiple groups simultaneously
-*  Modifying the variables in the NB analysis itself.
-
-Finally, there are several things still missing or needed for this workflow:
-
-*  How to feed a text dataset in, and get a table of the predicted categories back out.
-*  How to store the values from the Bayes prediction table so I can use them in calculations rather than type them in by hand.
-*  How to randomize which comments are used for training, testing datasets.
-
-####\
+Rows ~30-200 are the initial version that I based on this tutorial [https://rpubs.com/Billyhansen6/318412].
+Various permutation tests start around row #210.
 
@@
 ##Baseline Version 1.0 of Naive Bayes (NB) Protocol
-###Initial Data Input
+Initial data input.
 ```{r}
@@
 
-The OPTIONAL data transforms using "tm_map"" and "content_transformer" functions. Switch between "{r}" and "{}" to turn a transformation block off or on.
+These are the OPTIONAL data transforms using "tm_map"" and "content_transformer" functions.
 ```{r}
@@
 ```{r}
-#Remove punctuation using the "removePunctuation" function. 
+#Remove punctuation as well using the "removePunctuation" function. 
 #This step removes evidence of questions, so may remove data.
@@
 comments_corpus_clean <- tm_map(comments_corpus_clean, stripWhitespace)
+```
 
+```{r}
 #Look at an example of data.
@@
 
-###Basic Tokenizer
+**Original Tokenizer**
 This version uses the "DocumentTermMatrix" function to creates a matrix in which the rows indicate individual documents (individual TA comments) and the columns indicate presence or absence of the word that is the header for the column.
@@
 ```{r}
-.75 * 9340 #Number of documents/rows in the dataset; product is #rows for training
+.75 * 9340 #Number of rows in the dataset; product is #rows for training
 .25 * 9340 #Product is #rows for testing set
@@
 
-Calculate overall correct prediction rate.
 ```{r}
-(42+454+1186+185)/23.35
+(42+449+1185+186)/2335
 ```
@@
 ***
-##Modified Naive Bayes (NB) Method Using N-grams
-The tokenizer command block above only works for single words.
-The block below uses tm's NLP commands to generate either 1-, 2-, or 3-grams with the same command.
-The only change needed is "#" in "(words(x), #)" below.
-For simplicity I left off all of the preparatory code up leading up to the final cleaned corpus.
+***
+
+##Permutations of Naive Bayes (NB) Method
+Version 1.0 correctly identify comment subcategories ~80% of the time.
+This is better than the inter-rater reliability of two human raters, but still low.
+This notebook documents tests of various permutations of the NB method including:
+
+*  Compare 2-grams, 3-grams rather than single words
+*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
+    +  Ex.: Basic vs. Not Basic (1_Basic vs. not 1_Basic)
+*  Modify the variables within the NB analysis itself. 
+    +  Laplace constant (Y/N), Cutoff value for words removed from list.
+*  Do/Do not stem the terms
+    +  This may remove evidence of tense
+*  Remove/Leave stopwords
+*  Remove/Leave capitalization
+*  Remove/Leave numbers
+    +  This may remove evidence of technical descriptors
+*  Remove/Leave punctuation
+    +  This may remove evidence of questions vs statements
+
+Data set converted to a volative corpus using "tm" library then transformed.
+```{r}
+#Create the volatile coprus that contains the "text" vector from data frame.
+comments_corpus <- VCorpus(VectorSource(comments_raw$text))
+print(comments_corpus)
+
+#Check out the first few messages in the new corpus, which is basically a list that can be manipulated with list operations.
+inspect(comments_corpus[1:3])
+
+#Use "as.character" function to see what a message looks like.
+as.character(comments_corpus[[3]])
+```
+
+The OPTIONAL data transforms using "tm_map"" and "content_transformer" functions.
+```{r}
+#Convert to all lower case letters.
+comments_corpus_clean <- tm_map(comments_corpus, content_transformer(tolower))
+```
+
+```{r}
+#Remove numbers. 
+comments_corpus_clean <- tm_map(comments_corpus_clean, removeNumbers)
+```
+
+```{r}
+#Stopword removal. 
+comments_corpus_clean <- tm_map(comments_corpus_clean, removeWords, stopwords())
+```
+
+```{r}
+#Remove punctuation as well using the "removePunctuation" function. 
+#This step removes evidence of questions, so may remove data.
+comments_corpus_clean <- tm_map(comments_corpus_clean, removePunctuation)
+```
+
+```{r}
+#Stemming the text data to strip the suffix from words.  
+comments_corpus_clean <- tm_map(comments_corpus_clean, stemDocument)
+```
+
+Final prep and data check.
+```{r}
+#Final step removes white space from the document. This is NOT optional.
+comments_corpus_clean <- tm_map(comments_corpus_clean, stripWhitespace)
+```
+
+```{r}
+#Look at an example of data.
+as.character((comments_corpus_clean[[3]]))
+```
+
+####\
+
+**Alternative Tokenizer For N-grams**
+This code block uses tm's NLP commands to generate n-grams from full corpus of TA comments.
+It can create 1-, 2-, and 3-grams all with the same commands simply by changing "#" in "(words(x), #)" below.
 
@@
 ```{r}
-.75 * 9340 #Number of rows in the dataset; product is #rows for training
-.25 * 9340 #Product is #rows for testing set
+.75 * 9129
+.25 * 9129
 ```
@@
 ```{r}
-comments_dtm_train <- comments_dtm_2gram[1:7005, ]
-comments_dtm_test <- comments_dtm_2gram[7006:9340, ]
+comments_dtm_train <- comments_dtm_2gram[1:6846, ]
+comments_dtm_test <- comments_dtm_2gram[6847:9129, ]
 ```
@@
 ```{r}
-comments_train_labels <- comments_raw[1:7005, ]$type
-comments_test_labels <- comments_raw[7006:9340,]$type
+comments_train_labels <- comments_raw[1:6846, ]$type
+comments_test_labels <- comments_raw[6847:9129,]$type
 ```
@@
 ###Train Model, Predict, Evaluate
-Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group ONE, TWO, THREE, or FOUR.
+Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group TWO, THREE, or FOUR.
 Evaluate the prediction with the actual data using a crosstable from the gmodels package.
@@
 ```{r}
-(34+214+1254+94)/23.35
+(39+219+1233+90)/2283
 ```
@@
 ***
+
 ####\
@@
 ##Visualization
-This code generates a wordcloud of the frequency of single words in the dataset using the package "wordcloud".
+Create a wordcloud of the frequency of the words in the dataset using the package "wordcloud".
 ```{r}
@@
 
+***
+***
+
+**Permutation 1**
+Did not remove numbers, did not remove stopwords or punctuation, did not stem the terms. Reduced word count limit from 5 to 4 for matrix word limit.
+
+```{}
+comments_freq_words <- findFreqTerms(comments_dtm_train, 4)
+str(comments_freq_words)
+```
+
+This step raised accuracy for calling 3_Technical comments to 91.9%, but dropped overall accuracy to 79.0%.
+
+***
+***
+
+**Permutation 2**
+Removed stopwords, numbers and punctuation, but did not use stemming. Also dropped to comparing two groups. This approach raised the accuracy to:
+
+*  87.7% for separating 2_writing and 3_technical. 
+*  83.1% for separating 2_writing and 4_logic. 
+*  91.2% for separating 3_technical and 4_logic.
+
+```{r}
+library(tidyverse)
+library(tidytext)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(scales)
+library(tm)
+library(SnowballC)
+library(wordcloud)
+library(e1071)
+library(gmodels)
+library(readr)
+```
+
+```{r}
+#Read in CSV file named "coded_full_comments_dataset_Spring18anon.csv".
+base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')
+
+#Isolate rows to compare. 
+comments_subset <- filter(base_data,code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking") #|code.subject=="2. Writing Quality"
+
+#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
+comments_raw <- comments_subset %>% select(23,22)
+
+#Rename the columns.
+names(comments_raw)[1] <- "type"
+names(comments_raw)[2] <- "text"
+
+#Simplify coding terms
+comments_raw[,1] <- ifelse(comments_raw[,1] == "3.
+Technical and Scientific", "3_Technical", ifelse(comments_raw[,1] == "4.
+Logic and Thinking","4_Logic", 99))# ifelse(comments_raw[,1] == ,99)))"2.
+Writing Quality", "2_Writing"
+
+#Change "type" element from character to a factor for analysis.
+comments_raw$type <- factor(comments_raw$type)
+str(comments_raw$type)
+table(comments_raw$type)
+```
+
+```{r}
+##Text Transformation using "tm" package in R to create a volitile coprus that contains the "text" vector from our data frame.
+comments_corpus <- VCorpus(VectorSource(comments_raw$text))
+print(comments_corpus)
+
+#Check out the first few messages in the new corpus, which is basically a list that can be manipulated with list operations.
+inspect(comments_corpus[1:3])
+
+#Use "as.character" function to see what a message looks like.
+as.character(comments_corpus[[3]])
+```
+
+In order to standardize the messages, the data set must be tranformed to all lower case letters.
+The words "Free", "free", and "FREE" should all be treated as the same word.
+Use the "tm_map"" funtion in R, and use the "content_transformer" function to transform the text.
+```{r}
+comments_corpus_clean <- tm_map(comments_corpus, content_transformer(tolower))
+```
+
+Look at third message again to see if our data was transformed.
+```{r}
+as.character(comments_corpus[[3]])
+
+as.character((comments_corpus_clean[[3]]))
+```
+
+This removes numbers. **May not want to do.**
+```{r}
+comments_corpus_clean <- tm_map(comments_corpus_clean, removeNumbers)
+```
+
+Stopword removal takes out words that appear often but don't contribute to our objective. These words include "to", "and", "but" and "or". **One of variables to test out.**
+```{r}
+comments_corpus_clean <- tm_map(comments_corpus_clean, removeWords, stopwords())
+```
+
+Remove punctuation as well using the "removePunctuation" function. **Try not doing this. Removes evidence of questions.**
+```{r}
+comments_corpus_clean <- tm_map(comments_corpus_clean, removePunctuation)
+as.character((comments_corpus_clean[[3]]))
+```
+
+Perform "stemming" to the text data to strip the suffix from words. **Try NOT doing this.** 
+```{}
+comments_corpus_clean <- tm_map(comments_corpus_clean, stemDocument)
+```
+
+And now the final step in text mining is to remove white space from the document. 
+```{r}
+comments_corpus_clean <- tm_map(comments_corpus_clean, stripWhitespace)
+```
+
+```{r}
+as.character(comments_corpus_clean[[3]])
+```
+
+Perform tokenization using the "DocumentTermMatrix" function.
+This creates a matrix in which the rows indicate documents (SMS messages in this case) and the columns indicate words.
+It should be noted that the "DocumentTermMaxtrix" function has the power to do all of the text mining above in one command.
+```{r}
+sms_dtm2 <- DocumentTermMatrix(comments_corpus_clean)
+```
+
 ####\
 
-##Permutation Tests of Naive Bayes (NB) Method
-Version 1.0 correctly identifies comment subcategories ~79% of the time when using single words, and 68.3% of the time when using 2-grams. Permutations to test:
+##Data Preparation
+Split data into training and testing sets, so that after Naive Bayes spam filter algorithm is built it can be applied to unseen data. Divide our data set into 75% training and 25% testing.
 
-*  Compare 2-grams, 3-grams rather than single words
-*  Make pair-wise comparisons versus comparing multiple groups simultaneously. 
-    +  Ex.: Basic vs. Not Basic (1_Basic vs. not 1_Basic)
-*  Modify the variables within the NB analysis itself. 
-    +  Laplace constant (Y/N), Cutoff value for words removed from list.
-*  Do/Do not stem the terms
-    +  This may remove evidence of tense
-*  Remove/Leave stopwords
-*  Remove/Leave capitalization
-*  Remove/Leave numbers
-    +  This may remove evidence of technical descriptors
-*  Remove/Leave punctuation
-    +  This may remove evidence of questions vs statements
+```{r}
+.75 * 6551
+
+.25 * 9129
+```
+
+This code assumes comments are random. **Probably want to try randomizing them.**
+```{r}
+comments_dtm_train <- sms_dtm2[1:4913, ]
+comments_dtm_test <- sms_dtm2[4914:6551, ]
+```
+
+Save vectors labeling rows in the training and testing vectors
+```{r}
+comments_train_labels <- comments_raw[1:4913, ]$type
+comments_test_labels <- comments_raw[4914:6551,]$type
+```
+
+Make sure that the proportion of each sub-category is similar in the training and testing data set.
+```{r}
+prop.table(table(comments_train_labels))
+prop.table(table(comments_test_labels))
+```
+
+####\
+
+##Visualization
+Create a wordcloud of the frequency of the words in the dataset using the package "wordcloud".
+```{}
+wordcloud(comments_corpus_clean, max.words = 50, random.order = FALSE)
+```
+
+Compare wordclouds between 3 groups.
+```{}
+#TWO <- subset(comments_raw, type == "2_Writing")
+THREE <- subset(comments_raw, type == "3_Technical")
+FOUR <- subset(comments_raw, type == "4_Logic")
+
+#wordcloud(TWO$text, max.words = 50, scale = c(3, 0.5))
+wordcloud(THREE$text, max.words = 50, scale = c(3, 0.5))
+wordcloud(FOUR$text, max.words = 50, scale = c(3, 0.5))
+```
+
+####\
+
+##Preparation for Naive Bayes
+Remove words from the matrix that appear less than 5 times.
+```{r}
+comments_freq_words <- findFreqTerms(comments_dtm_train, 5)
+str(comments_freq_words)
+```
+
+Limit our Document Term Matrix to only include words in the sms_freq_vector. We want all the rows, but we want to limit the column to these words in the frequency vector.
+```{r}
+comments_dtm_freq_train <- comments_dtm_train[ , comments_freq_words]
+comments_dtm_freq_test <- comments_dtm_test[ , comments_freq_words]
+```
+
+The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables.
+To do this we'll build a convert_counts function and apply it to our data.
+```{r}
+convert_counts2 <- function(x) {
+  x <- ifelse(x > 0, "Yes", "No")
+}
+```
+
+This replaces values greater than 0 with yes, and values not greater than 0 with no. Let's apply it to our data.
+```{r}
+comments_train <- apply(comments_dtm_freq_train, MARGIN = 2, convert_counts2)
+comments_test <- apply(comments_dtm_freq_test, MARGIN = 2, convert_counts2)
+```
+
+The resulting matrixes will be character type, with cells indicating "yes" or "no" if the word represented by the column appears in the message represented by the row.
+
+####\
+
+##Train Model, Predict, Evaluate
+Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group TWO, THREE, or FOUR.
+Evaluate the prediction with the actual data using a crosstable from the gmodels package.
+```{r}
+comments_classifier <- naiveBayes(comments_train, comments_train_labels)
+comments_test_pred <- predict(comments_classifier, comments_test)
+CrossTable(comments_test_pred, comments_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
+```
+
+Show the 10 most frequent words in the data:
+```{r}
+sack2 <- TermDocumentMatrix(comments_corpus_clean)
+pack2 <- as.matrix(sack2)
+snack2 <- sort(rowSums(pack2), decreasing = TRUE)
+hack2 <- data.frame(word = names(snack2), freq=snack2)
+head(hack2, 10)
+```
+
+And the 10 most frequent words from each class:
+```{}
+wordcloud(TWO$text, max.words = 10, scale = c(3, 0.5))
+wordcloud(THREE$text, max.words = 10, scale = c(3, 0.5))
+#wordcloud(FOUR$text, max.words = 10, scale = c(3, 0.5))
+```

### 2019-02-25 NB_2r2c.Rmd
updates

@@
 ---
-title: "Naive Bayes Analysis Ver 2r.2c"
-author: "Dan Johnson"
-date: "2/21/2019"
-output: html_document
+title: "Naive Bayes Classifier"
+date: "2/24/2019"
+bibliography: library.bib
+biblio-style: apalike
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, child="_setup.Rmd"}
 ```
 
-##Summary
-This it the current Naive Bayes workflow used to classify TA comments based on Subject and Structure. Additions and revisions made to **Version 2r.2c** are:
+#Overview
+This is the current version of the Naive Bayes workflow used to classify TA comments based on Subject and Structure. Features of **Version 2r.2c**:
 
-*  Random selection of data rows used for training and test datasets. 
-*  Final output matrix tables are more informative. 
-*  Addition of a back-analysis step that reviews entire dataset, not just testing sub-set.
-*  Outputs a tidy data table in CSV format that contains the original comments (pre-clean-up) plus their hand-scored and predicted subject and structure.
+*  Data rows used for training and test datasets are assigned randomly but reproducibly using 'set.seed'. 
+*  Detailed final output matrix tables. 
+*  Back-analysis evaluates both a testing sub-set and the entire input dataset.
+*  Outputs a tidy data table in CSV format with unmodified original comments, and hand-scored and NB-predicted Subject and Structure assignments.
+
+This page outlines Ver 2r.2c of NB Classifier and the process for building various components used in the project. Details of each  component are provided in these sub-pages.
+
+For structure of the initial dataset see: https://adanieljohnson.github.io/default_website/datastructure.html
+
+For import and post-import data validation see: https://adanieljohnson.github.io/default_website/dataimport.html
 

### 2019-02-25 NB_2r2c.Rmd
update

@@
 
-#Overview
 This is the current version of the Naive Bayes workflow used to classify TA comments based on Subject and Structure. Features of **Version 2r.2c**:
@@
 
-```{r writeCSV}
+```{writeCSV}
 #Creates a data frame that compares predicted and actual labels for subject and structure
@@
 
+####\
+
 ##Graph Outputs
@@
 
-```{r}
+```{}
 ggplot(data=proptallysum1, aes(`Hand.Coded.Subject`, `total`,group=1))+

### 2019-02-25 Text_classifier_models.Rmd
update

@@
 ```
-
-#Overview
 This page provides general background on theory and approaches to text classification. Links below connect to sub-pages that describe more specific topics. Follow the links for details.
-
-Topic Categorization and Modeling Methods: https://adanieljohnson.github.io/default_website/Topic_categorization.html
-
-Choosing a Text Classifier Method: https://adanieljohnson.github.io/default_website/Choosing_classifier.html
-
-Designing a Classifier Evaluation Workflow: https://adanieljohnson.github.io/default_website/Classifier_comparison.html
-
-Strategies for Improving, Evaluating Text Classifiers: https://adanieljohnson.github.io/default_website/Improving_classifier.html
 
@@
 #Basic Workflow
-Text classification problems can be broken down into four overlapping sub-problems. Each piece may be more or less important to the overall problem, but all need to be considered:
+Text classification problems can be broken down into four overlapping sub-problems. Each piece may be more or less important to the overall problem, but all need to be considered.
 
-1.  Feature selection and engineering
-2.  Topic categorization and modeling
-3.  Selecting an appropriate text classifier method
-4.  Evaluating the outcomes
+1.  Feature selection and engineering (described here)
+2.  Topic categorization and modeling (https://adanieljohnson.github.io/default_website/Topic_categorization.html)
+3.  Selecting an appropriate text classifier method (https://adanieljohnson.github.io/default_website/Choosing_classifier.html)
+4.  Evaluating the outcomes, and improving the classifier ( https://adanieljohnson.github.io/default_website/Improving_classifier.html)
 
@@
 
-##Feature Selection and Engineering
+#Feature Selection and Engineering
 This is the process of selecting, tabulating, calculating, or creating text-based features for subsequent evaluation. This includes:
@@
 *  Engineering additional features using established NLP methods
-*	 Reverse engineering relevant features based on an empirically defined feature set.
+*	 Reverse engineering relevant features based on an empirically defined feature set
 
-The features listed below are used regularly for natural language processing.
-Not all will be appropriate for all situations though; selecting, evaluating, comparing, and validating the appropriate features for the classification task is an iterative process, not a one-time decision.
+The features listed below are used routinely in natural language processing.
+Not all will be appropriate for all situations though; **selecting, evaluating, comparing, and validating appropriate features for a classification task is an iterative process, not a one-time decision.**
 
@@
 	+  TF-IDF
-	+  Reading level estimates (described below)
+	+  Reading level estimates, readability (described below)
 *  Part of Speech Tagging and Frequency Distribution of Tags 
@@
 
-Additional features are available for longer texts such as multi-sentence paragraphs.
+Additional features are available for longer texts such as multi-sentence paragraphs. These features require analysis of sentence structure, syntax, parts of speech patterns, etc.
 
-*  Word choices
+*  Word choices (vocabulary range)
 *  Lexical diversity (described below)
@@
 
-###Lexical Diversity
-This is not the same as word choices.
-This explanation is an excerpt from "quanteda: Quantitative Analysis of Textual Data", which describes how to use ```textstat_lexdiv``` to calculate lexical diversity.
+#Some Feature Engineering Methods
+##Lexical Diversity
+This is different from simple word frequencies The following explanation is an excerpt from "quanteda: Quantitative Analysis of Textual Data", which describes how to use ```textstat_lexdiv``` to calculate lexical diversity.
 https://rdrr.io/cran/quanteda/man/textstat_lexdiv.html
@@
 
-koRpus package in R also provides functions to estimate lexical diversity or complexity.
+koRpus package in R also provides functions to estimate lexical diversity or complexity. For more about this package, go to 
 https://subscription.packtpub.com/book/big_data_and_business_intelligence/9781783551811/2/ch02lvl1sec16/lexical-diversity
@@
 
-###Readability metrics
+##Readability metrics
 These estimate the relative grade level or age of a typical reader who could comprehend 80+% (typically) of the text.
@@
 
-###Word Embedding
+##Word Embedding
 **Word embedding** is a set of language modeling and feature learning techniques in natural language processing (NLP) where words or phrases from the vocabulary are mapped to vectors of real numbers.
+The vectors then serve as engineered features for subsequent classifier functions.
+
 Conceptually it involves a mathematical embedding from a space with one dimension per word to a continuous vector space with a much lower dimension.
@@
 
-One of the main limitations of word embeddings (word vector space models in general) is that possible meanings of a word are conflated into a single representation (a single vector in the semantic space).
+One of the main limitations of word embeddings (and word vector space models in general) is that possible meanings of a word are conflated into a single representation (a single vector in the semantic space).
 Sense embeddings are a solution to this problem: individual meanings of words are represented as distinct vectors in the space.

### 2019-02-25 build.Rmd
updates
update

--- 2019-02-22 build.Rmd
@@
 title: "The Build"
-author: "Dan Johnson"
 date: "2/22/2019"
-output: html_document
+bibliography: library.bib
+biblio-style: apalike
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, child="_setup.Rmd"}
 ```
 
-#Overview
-This page outlines the process for building various components used in the project. Details of each  component are provided in these sub-pages.
+This page links to sub-pages describing how various components used in the overall project were constructed. Follow the links for details.
 
-Structure of the initial dataset: https://adanieljohnson.github.io/default_website/datastructure.html
+Extracting TAs' Comments (***Draft***): https://adanieljohnson.github.io/default_website/docxcommentextraction.html
 
-Import and post-import data validation: https://adanieljohnson.github.io/default_website/dataimport.html
+Codebook for Classifying TAs' Comments: https://adanieljohnson.github.io/default_website/codebook.html
+
+Structure of Development Dataset; Data Import & Validation: https://adanieljohnson.github.io/default_website/dataimport.html
+
+Initial Exploration of the Dataset: https://adanieljohnson.github.io/default_website/initial_exploration.html
+
+Current Version of the Naive Bayes Comment Classifier: https://adanieljohnson.github.io/default_website/NB_2r2c.html
+
+Embedding the Naive Bayes Classifier in an Ensemble Model (***Draft***): https://adanieljohnson.github.io/default_website/Ensemble_Notes.html
 
 ####\
-
-#Describe an Element
-
-####\
-
-#Describe an Element
-
-####\

### 2019-02-25 initial_exploration.Rmd
update

@@
 
-#Overview
-The first question to answer before making deeper comparisons is, **are word frequencies different between TAs' comments from different sub-categories?** This page describes the methods used for this initial exploration.
+The first question to answer before making deeper comparisons is, **are word frequencies different between TAs' comments from different sub-categories?** This page describes the methods used for initial exploration of the TA comments dataset.
 
-Results of these analyses are discussed on this page: https://adanieljohnson.github.io/default_website/findings.html
+Results of these analyses are discussed in depth on this page: https://adanieljohnson.github.io/default_website/findings.html
 
@@
 #The Analyses
-I ran three different analyses of word usage frequency.
+I ran three different analyses of word use frequency. For simplicity I reduced the analysis to just the four most common comment subjects.
 
-*  __Pearson correlations for word use frequencies__. This only compares words found in all of the Subject sub-categories
-*  __Word frequency differences based on log-odds ratios__. This highlights terms that are the most different between sub-categories.
-*  __N-grams mapping__ (*draft*). This identifies text strings, phrases, and terms that are unique (or nearly unique) to one sub-category.
+*  __Pearson correlations for word use frequencies__. This method compares the **absolute frequencies** for words found in **ALL** of the Subject sub-categories.
+*  __Word frequency differences based on log-odds ratios__.
+This method highlights terms that are the **most different between sub-categories**.
+Again, this method compares frequencies for words found in **ALL** of the Subject sub-categories.
+*  __N-grams mapping__ (***Draft***).
+This method identifies text strings, phrases, and terms that are unique (or nearly unique) to one sub-category.
+Unlike the other two methods, this one does not drop terms which are missing from one or more groups.
 
@@
 
-#ALTERNATIVE METHOD for pulling out stopwords.  
+#Stopwords removal.  
 sortedwords.comments.all <- base_data_tokenized %>%
@@
 
-The next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed. 
+The next block joins the individual data tables for each subset in a larger file using column named 'word'.
+Columns get confusing names so they are renamed.
+The graphs compare frequencies for ~770 words that appear in ALL three sub-categories: Writing, Technical, and Logic.
+
+Graphs 1-3 show frequencies in individual groups versus word frequency overall.
+Generally, frequency of terms in the whole set of comments should be similar to frequency in an individual group (X vs Y positions similar.) However, some terms associated with Logic subgroup (Graph 3) are far from the best-fit line, suggesting they are under-used for comments in that sub-category.
+
+Graphs 4-6 show pair-wise comparisons between sub-categories.
+Unlike overall frequency, most words are NOT clustered along the best-fit line, suggesting there is still more variation in word use between sub-categories.
+
 ```{r}
@@
 
-```
-
-The following six graphs compare frequencies for ~770 words that appear in ALL three sub-categories: Writing, Technical, and Logic.
-```{r}
+#Plot graphs
 ggplot(total6, aes(`all.fraction`,`writing.fraction`)) +
@@
 
-Graphs 1-3 show frequencies in individual groups versus word frequency overall.
-Generally, frequency of terms in the whole set of comments should be similar to frequency in an individual group (X vs Y positions similar.) However, some terms associated with Logic subgroup (Graph 3) are far from the best-fit line, suggesting they are under-used for comments in that sub-category.
-
-Graphs 4-6 show pair-wise comparisons between sub-categories.
-Unlike overall frequency, most words are NOT clustered along the best-fit line, suggesting there is still more variation in word use between sub-categories.
-
 ####\
@@
 ##Which Words Are Shared LEAST Between Categories?
-Unlike the preceding analysis, log-odds ratio comparison does not eliminate unshared terms.
-Instead, it identifies which words are MOST likely to be used in one sub-category, but not another.
-For simplicity I reduced the analysis to just the four most common comment subjects.
-This has been the most informative preliminary visualization I have found so far.
-
+Log-odds ratio comparison identifies the words that have the greatest difference in usage between the two groups being compared.
 This code block is based on a method for comparing word usage frequencies between two separate Twitter users [https://www.tidytextmining.com/twitter.html](described here.) If the words used by TAs in their comments differ between sub-categories, there should be fairly low Pearson correlations between word frequencies.

### 2019-02-25 litreview.Rmd
correction
update

@@
 #Overview
-The supporting literature for this project is broad and convoluted.
-Rather than try to compress it to a single page, I have provided links out to sub-pages that describe various background topics.
+This project brings together several traditionally separate fields of research:
+
+*  General learning theories
+*  Educational research specific to STEM fields
+*  Writing in the disciplines, and technical writing instruction
+*  Text analysis and classification
+
+Instead of trying to weave a coherent one-page narrative of these very different fields, I have provided links out to sub-pages that describe various background topics separately.
 Follow the links for details.
+
+General Learning Theories Relevant to the Current Challenge:
+https://adanieljohnson.github.io/default_website/pagename.html
+
+Educational Research Specific to STEM fields
+https://adanieljohnson.github.io/default_website/pagename.html
 
@@
 
-Primer on Text Classification Strategies:  https://adanieljohnson.github.io/default_website/Text_classifier_models.html
+Primer on Text Classification Strategies; Feature Selection and Engineering:  https://adanieljohnson.github.io/default_website/Text_Classifier_Models.html
+
+>Sub-pages:  
+>1.  Topic categorization and modeling (https://adanieljohnson.github.io/default_website/Topic_categorization.html)  
+>2.  Selecting an appropriate text classifier method (https://adanieljohnson.github.io/default_website/Choosing_classifier.html)  
+>3.  Evaluating the outcomes, and improving the classifier (https://adanieljohnson.github.io/default_website/Improving_classifier.html)
 

### 2019-02-25 Help.rmd
updates

Items to get help on:

#Index:
	Site map

#Background:

#Lit review:
	How to branch to Text Classification Primer, sub-pages look okay?
		Text_classifier_models	= 	Text Classification Primer
		Topic_categorization = Topic Categorization and Modeling Methods
		Choosing_classifier = Choosing a Text Classifier Method
		Classifier_comparison = Designing a Classifier Evaluation Workflow 
		Improving_classifier = Strategies for Improving Text Classifiers

#The Build:
	df
	  df

#Data Exploration:
	Help with aligning tables and keeping the NA values

#Findings:
	How to graph the different comparisons of NB values from different sources on same plot.

#Next Steps:
	Df
	  Df

#Etc.:
	Using a Bibliography in this format
	Can menus have sub-pages
	Adding automatic updates to top dates
	Moving top dates

Using NLP and text data mining

### 2019-02-25 notes.rmd
update

---
title: "Working Notes"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

#Items to get help on:

##Index:
	Site map

##Background:

##Lit review:
	How to branch to Text Classification Primer, sub-pages look okay?
		Text_classifier_models	= 	Text Classification Primer
		Topic_categorization = Topic Categorization and Modeling Methods
		Choosing_classifier = Choosing a Text Classifier Method
		Classifier_comparison = Designing a Classifier Evaluation Workflow 
		Improving_classifier = Strategies for Improving Text Classifiers

##The Build:
	df
	  df

##Data Exploration:
	Help with aligning tables and keeping the NA values

##Findings:
	How to graph the different comparisons of NB values from different sources on same plot.

##Next Steps:
	Df
	  Df

##Etc.:
	Using a Bibliography in this format
	Can menus have sub-pages
	Adding automatic updates to top dates
	Moving top dates

##Check These Out:
  Upload 'devtools::install_github("WFU-TLC/analyzr")'

####\

### 2019-02-25 notes.rmd
update

@@
 ##Background:
+	Df
+	  Df
 
@@
 ##The Build:
-	df
-	  df
+	Df
+	  Df
 
@@
 ##Next Steps:
-	Df
+	Still in draft stage
 	  Df
@@
 ##Etc.:
-	Using a Bibliography in this format
-	Can menus have sub-pages
-	Adding automatic updates to top dates
-	Moving top dates
+	Using a Bibliography in this format?
+	Can menus have sub-pages?
+	Adding automatic updates to top dates?
+	Moving top dates?
+	Fix GitHub auto-upload
 
@@
   Upload 'devtools::install_github("WFU-TLC/analyzr")'
+  Using Mike's Metadata packager to annotate our released dataset
+  Try using ```mutate(datacolumn=str_count(utterance_clean, "\\b(um|u=m)\\b"))``` to extract a string character and count.
 

### 2019-02-25 notes.rmd
update

@@
 ##Lit review:
-	How to branch to Text Classification Primer, sub-pages look okay?
+	How to branch to Text Classification Primer
+	  Use sub-menus
 		Text_classifier_models	= 	Text Classification Primer
@@
 ##Data Exploration:
-	Help with aligning tables and keeping the NA values
+	Help with aligning tables and keeping the NA values. 
+	  Use full_join, not merge.
 
 ##Findings:
-	How to graph the different comparisons of NB values from different sources on same plot.
+	How to graph the different comparisons of NB values from different sources on same plot? 
+	  Add ggplot (data=) value.
 
@@
 	Using a Bibliography in this format?
+
 	Can menus have sub-pages? 
+	  Yep, using sub-pages
 	Adding automatic updates to top dates? 
-	Moving top dates?
+	  `r_sys.date()`
 	Fix GitHub auto-upload

### 2019-02-25 Ensemble_Notes.Rmd
update

--- 2019-02-21 hiding/Ensemble_Notes.Rmd
@@
 ---
-title: "Notes on Building a Final Ensemble Protocol"
-author: "Dan Johnson"
-date: "2/14/2019"
-output: html_document
+title: "Final Ensemble Protocol"
+date: "`r format(Sys.time(), '%b %d, %Y')`"
+bibliography: library.bib
+biblio-style: apalike
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, child="_setup.Rmd"}
 ```

### 2019-02-25 docxcommentextraction.Rmd
Web site update
update

@@
 title: "DOCX Comment Extraction"
-author: "Dan Johnson"
-date: "2/22/2019"
-output: html_document
+date: "`r format(Sys.time(), '%b %d, %Y')`"
+bibliography: library.bib
+biblio-style: apalike
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, child="_setup.Rmd"}
 ```
@@
 #Sample Working Code
-```{r}
+```{}
 library(docxtractr)

### 2019-02-25 litreview.Rmd
update

--- 2019-02-23 litreview.Rmd
@@
 title: "Literature Review"
-date: "2/21/2019"
+date: "`r format(Sys.time(), '%b %d, %Y')`"
 bibliography: library.bib
@@
 #Overview
-This page links out to sub-pages that describe various background topics. Follow the links for details.
+This project brings together several traditionally separate fields of research:
 
-Description: https://adanieljohnson.github.io/default_website/pagename.html
+*  General learning theories
+*  Educational research specific to STEM fields
+*  Writing in the disciplines, and technical writing instruction
+*  Text analysis and classification
 
-Technical Writing Training
+Instead of trying to weave a coherent one-page narrative of these very different fields, I have provided links out to sub-pages for each of the various background topics. Follow the links for details.
 
-Text Classification Strategies
+General Learning Theories Relevant to the Current Challenge:
+https://adanieljohnson.github.io/default_website/pagename.html
+
+STEM Educational Research:
+https://adanieljohnson.github.io/default_website/pagename.html
+
+Foundations of Technical Writing Training:  https://adanieljohnson.github.io/default_website/pagename.html
+
+Text Classification 1 - Feature Selection and Engineering:  https://adanieljohnson.github.io/default_website/Text_Classifier_Models.html
+
+Text Classification 2 - Topic categorization and modeling (https://adanieljohnson.github.io/default_website/Topic_categorization.html)
+
+Text Classification 3 - Selecting an appropriate text classifier method (https://adanieljohnson.github.io/default_website/Choosing_classifier.html)
+
+Text Classification 4 - Evaluating the outcomes, and improving the classifier (https://adanieljohnson.github.io/default_website/Improving_classifier.html)
 

### 2019-02-25 notes.rmd
update

--- 2019-02-25 notes.rmd
@@
 title: "Working Notes"
+date: "`r format(Sys.time(), '%b %d, %Y')`"
 bibliography: library.bib
@@
 ```
+Items to get help on:
 
-#Items to get help on:
+##General/Site-Wide:
+  Embedding the bibliographic items  
+    Writing the bibtext  
+    Embedding using [@ BaseR] method.
 
 ##Index:
-	Site map
+*  Decide on bottom page site map strategy
 
 ##Background:
+*  Need to finish these pages:  
+    +  Df
+    +  Df
 
 ##Lit review:  
-	How to branch to Text Classification Primer, sub-pages look okay?
-		Text_classifier_models	= 	Text Classification Primer
-		Topic_categorization = Topic Categorization and Modeling Methods
-		Choosing_classifier = Choosing a Text Classifier Method
-		Classifier_comparison = Designing a Classifier Evaluation Workflow 
-		Improving_classifier = Strategies for Improving Text Classifiers
+*  Df  
+    +  Df
 
 ##The Build:
-	df
-	  df
+*  Find out from Melissa which method/model she is using. Useful for another project.  
+    +  Df
 
 ##Data Exploration:
-	Help with aligning tables and keeping the NA values
+*  Try fixing problem of aligning tables and keeping the NA values by switching from "merge" to "full_join"  
+*  Try modifying the comparison of word frequencies in all comments vs.
+subgroups to comparison between two subsamples from full group.
+How do the individual subgroups compare to a random sample?
+Is the variation I see just noise?
+*  For N-gram analysis, how to join tables with missing values?
+    +  Use full_join, not merge.
 
 ##Findings:
-	How to graph the different comparisons of NB values from different sources on same plot.
+*  How to graph the different comparisons of NB values from different sources on same plot? 
+    +  Add ggplot (data=) value.
+    +  Better to combine data into one dataset though.
 
-##Next Steps:
-	Df
-	  Df
+#Next Steps:
+*  Still in draft stage
+    +  Df
 
-##Etc.:
-	Using a Bibliography in this format
-	Can menus have sub-pages
-	Adding automatic updates to top dates
-	Moving top dates
+#Etc.:
+*  Using a Bibliography in this format  
+*  Adding automatic updates to top dates? 
+    +  ```date: "`r format(Sys.time(), '%b %d, %Y')`" ```
 
-##Check These Out:
-  Upload 'devtools::install_github("WFU-TLC/analyzr")'
+*  Fix GitHub auto-upload
+
+##Try These Out:
+*  Use Mike's Metadata packager to annotate our released dataset
+*  Use ```mutate(datacolumn=str_count(utterance_clean, "\\b(um|u=m)\\b"))``` to extract a string character and count. Use this to count punctuation only comments, or ID comments with questions.
 

### 2019-02-25 NB_2r2c.Rmd
update

--- 2019-02-25 NB_2r2c.Rmd
@@
 title: "Naive Bayes Classifier"
-date: "2/24/2019"
+date: "`r format(Sys.time(), '%b %d, %Y')`"
 bibliography: library.bib
@@
 
+####\
+
 #Text Data Transformations
@@
 ```{}
-#This is for convenience. It is not required if the number of training, test rows is known.
+#This block is for convenience. It is not required if the number of training, test rows is known.
 .75 * 9340 #Number of rows in the dataset; product is #rows for training
@@
 
-Below is an **alternate version** of convert_function code block that KEEPs word frequencies while still using the same variable names and code structure.
+The block below is an **alternate version** of the ```convert_function``` code block that KEEPs word frequencies while still using the same variable names and code structure.
 The processed matrices have cells with word frequencies instead of categorical yes/no values.
-This block only works for binary classification (ingroup/outgroup).
-It breaks when applied to data with multiple categories.
+This block **only** works when performing binary classification (ingroup/outgroup).
+The classifier fails when frequency values are used with data in multiple categories.
 ```{}
@@
 
-Applies whichever "convert_counts2" function is currently active.
+This block applies whichever "convert_counts2" function is currently active.
 ```{r apply_conversion}
@@
 
-#Train Model, Predict, Evaluate for Subject
+#Train Model, Predict & Evaluate for Subject
 This block predicts whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. 
@@
 ```{r reportOut}
-#"line" is just spacer text for improving readability.
-line<-("**************************************************************")
-
 # Prepare the confusion matrix. 
@@
 #These commands show how to print out individual pieces of the confusion matrix. They show both the overall and by-group accuracy values for Subject x Test Subset.
+print ("Test, Subject")
 conf.mat1
-print (line)
-print ("Test, Subject")
 
 # This block outputs the by-group accuracy values in a single column for Subject x Test Subset. 
+print ("Test, Subject")
 conf.mat1$byClass
-print (line)
-print ("Test, Subject")
 
 # This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Subject x Test Subset.
+print ("Test, Subject")
 conf.mat1$overall
-print (line)
-print ("Test, Subject")
 
 # This block outputs JUST the overall accuracy value for Subject x Test Subset. 
+print ("Test, Subject")
 conf.mat1$overall['Accuracy']
-print ("Test, Subject")
 
 # This block outputs both the overall and by-group accuracy values for Subject x All Data.
+print ("All, Subject")
 conf.mat2
-print (line)
-print ("All, Subject")
 
 # This block outputs by-group accuracy values in a single column for Subject x All Data. 
+print ("All, Subject")
 conf.mat2$byClass
-print (line)
-print ("All, Subject")
 
 # This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Subject x All Data.
+print ("All, Subject")
 conf.mat2$overall
-print (line)
-print ("All, Subject")
 
 # This block outputs JUST the overall accuracy value for Subject x All Data. 
+print ("All, Subject")
 conf.mat2$overall['Accuracy']
-print ("All, Subject")
 
 # This block outputs both the overall and by-group accuracy values for Structure x Test Subset.
+print ("Test, Structure")
 conf.mat3
-print (line)
-print ("Test, Structure")
 
 # This block outputs by-group accuracy values in a single column for Structure x Test Subset. 
+print ("Test, Structure")
 conf.mat3$byClass
-print (line)
-print ("Test, Structure")
 
 # This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Structure x Test Subset.
+print ("Test, Structure")
 conf.mat3$overall
-print (line)
-print ("Test, Structure")
 
@@
 # This block outputs both the overall and by-group accuracy values for Structure x All Data.
+print ("All, Structure")
 conf.mat4
-print (line)
-print ("All, Structure")
 
 # This block outputs by-group accuracy values in a single column for Structure x All Data. 
+print ("All, Structure")
 conf.mat4$byClass
-print (line)
-print ("All, Structure")
 
 # This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Structure x All Data.
+print ("All, Structure")
 conf.mat4$overall
-print (line)
-print ("All, Structure")
 
@@
 #Output for Review, Analysis
-The code block below creates a CSV file containing the original TA comments, the subject and structure labels assigned by a human observer, and the corresponding predicted categories for each comment based on the Naive Bayes estimator.
+The code block below creates a CSV file containing the **original** unmodified TA comments, the subject and structure labels assigned by a human observer, and the corresponding predicted categories for each comment based on the Naive Bayes estimator.
 Results are exported to a CSV file.
@@
 
-This is an alternate method to get the same data tables out using ```tidy``` format. It seems much longer to me, so likely is over-written.
-```{}
-tidy_prediction_table
+This block is an alternative method to get the same data tables out using ```tidy``` format. It seems much less direct, but may be over-written.
+```{r tidy_prediction_table}
 
@@
 
+####\
+
 ##Graph Outputs
-Comparing outcome from Hand-coding vs. Bayes
-
-***
-Need to combine two datasets into one. 
-***
+This code block generates two graphs comparing the sub-category frequencies obtained by hand-coding vs.
+predicted frequencies from the Naive Bayes classifier.
+There are separate comparisons for the Subject and Structure categories.
 
 ```{r}
-ggplot(data=proptallysum1, aes(`Hand.Coded.Subject`, `total`,group=1))+
-  geom_line()
-ggplot(data=proptallysum2, aes(x=Bayes.Predicted.Subject, y=total, group=1))+
-  geom_line()
-ggplot(data=proptallysum3, aes(x=Hand.Coded.Structure, y=total, group=1))+
-  geom_line()
-ggplot(data=proptallysum4, aes(x=Bayes.Predicted.Structure, y=total, group=1))+
-  geom_line()
+blendSubject = ggplot()+
+  geom_col(data=proptallysum1, aes(`Hand.Coded.Subject`, `total`, group=1), color="blue", fill="cornflowerblue", alpha=0.5, width=0.4) +
+  geom_col(data=proptallysum2, aes(`Bayes.Predicted.Subject`, `total`, group=1), color="brown", fill="indianred4", alpha=0.5, width=0.4,position = position_nudge(x=0.1))+
+  xlab("Subject Groups")+
+  ylab("Fraction of All Comments")+
+  ggtitle("Comparison of Coded (blue) vs Predicted (red) Subject")
+
+blendStructure = ggplot()+
+  geom_col(data=proptallysum3, aes(`Hand.Coded.Structure`, `total`,group=1), color="springgreen4", fill="springgreen3", alpha=0.5, width=0.4) +
+  geom_col(data=proptallysum4, aes(`Bayes.Predicted.Structure`, `total`, group=1), color="goldenrod3", fill="goldenrod2", alpha=0.5, width=0.4, position = position_nudge(x=0.1))+
+  xlab("Structure Groups")+
+  ylab("Fraction of All Comments")+
+  ggtitle("Comparison of Coded (green) vs Predicted (orange) Structure")
+
+print(blendSubject)
+print(blendStructure)
 ```

### 2019-02-26 NB_2r2c.Rmd
update

@@
 
-This page outlines Ver 2r.2c of NB Classifier and the process for building various components used in the project. Details of each  component are provided in these sub-pages.
-
-For structure of the initial dataset see: https://adanieljohnson.github.io/default_website/datastructure.html
-
-For import and post-import data validation see: https://adanieljohnson.github.io/default_website/dataimport.html
-
 ####\
@@
 library(caret)
+library(ggplot2)
 ```
@@
 
-####\
+#Tabular and Graphical Comparisons of Coded and Predicted Categories
+These code blocks compare the hand-coded versus NB-predicted frequencies. There are separate comparisons for the Subject and Structure categories, and both tabular and graphical comparisons.
 
-##Graph Outputs
-This code block generates two graphs comparing the sub-category frequencies obtained by hand-coding vs.
-predicted frequencies from the Naive Bayes classifier.
-There are separate comparisons for the Subject and Structure categories.
+```{r}
+colnames(proptallysum1)[colnames(proptallysum1)=="Hand.Coded.Subject"] <- "Subject"
+colnames(proptallysum1)[colnames(proptallysum1)=="total"] <- "Hand"
+colnames(proptallysum2)[colnames(proptallysum2)=="Bayes.Predicted.Subject"] <- "Subject"
+colnames(proptallysum2)[colnames(proptallysum2)=="total"] <- "Bayes" 
+propmerge12<-merge.data.frame(proptallysum1,proptallysum2, all=TRUE)
+propmerge12
+```
 
@@
 blendSubject = ggplot()+
-  geom_col(data=proptallysum1, aes(`Hand.Coded.Subject`, `total`, group=1), color="blue", fill="cornflowerblue", alpha=0.5, width=0.4) +
-  geom_col(data=proptallysum2, aes(`Bayes.Predicted.Subject`, `total`, group=1), color="brown", fill="indianred4", alpha=0.5, width=0.4,position = position_nudge(x=0.1))+
+  geom_col(data=proptallysum1, aes(`Subject`, `Hand`, group=1), color="blue", fill="cornflowerblue", alpha=0.5, width=0.4) +
+  geom_col(data=proptallysum2, aes(`Subject`, `Bayes`, group=1), color="brown", fill="indianred4", alpha=0.5, width=0.4,position = position_nudge(x=0.12))+
   xlab("Subject Groups")+
   ylab("Fraction of All Comments")+
-  ggtitle("Comparison of Coded (blue) vs Predicted (red) Subject")
+  ggtitle("Comment Subject Frequencies: Coded (blue) vs Predicted (red)")
 
+print(blendSubject)
+```
+
+```{r}
+colnames(proptallysum3)[colnames(proptallysum3)=="Hand.Coded.Structure"] <- "Structure"
+colnames(proptallysum3)[colnames(proptallysum3)=="total"] <- "Hand"
+colnames(proptallysum4)[colnames(proptallysum4)=="Bayes.Predicted.Structure"] <- "Structure"
+colnames(proptallysum4)[colnames(proptallysum4)=="total"] <- "Bayes"
+propmerge34<-merge.data.frame(proptallysum3,proptallysum4, all=TRUE)
+propmerge34
+```
+
+```{r}
 blendStructure = ggplot()+
-  geom_col(data=proptallysum3, aes(`Hand.Coded.Structure`, `total`,group=1), color="springgreen4", fill="springgreen3", alpha=0.5, width=0.4) +
-  geom_col(data=proptallysum4, aes(`Bayes.Predicted.Structure`, `total`, group=1), color="goldenrod3", fill="goldenrod2", alpha=0.5, width=0.4, position = position_nudge(x=0.1))+
+  geom_col(data=proptallysum3, aes(`Structure`, `Hand`,group=1), color="springgreen4", fill="springgreen3", alpha=0.5, width=0.4) +
+  geom_col(data=proptallysum4, aes(`Structure`, `Bayes`, group=1), color="goldenrod3", fill="goldenrod2", alpha=0.5, width=0.4, position = position_nudge(x=0.12))+
   xlab("Structure Groups")+
   ylab("Fraction of All Comments")+
-  ggtitle("Comparison of Coded (green) vs Predicted (orange) Structure")
+  ggtitle("Comment Structure Frequencies: Coded (green) vs Predicted (orange)")
 
-print(blendSubject)
 print(blendStructure)
 ```
+
+####\  
+***

### 2019-02-26 aacr.Rmd
update

---
title: "Automated Analysis of Student Writing Writing "
date: "`r format(Sys.time(), '%b %d, %Y')`"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

There is growing need and interest in incorporating automated feedback into writing training. This raises important questions that have not yet been adequately answered:

*  Does automated feedback identify similar issues with similar frequency as live instructors?
*  When instructors have access to automated feedback, do they trust it? Do they provide feedback on different issues, focus on higher-order problems, or continue grading as if they had no insights?
*  Do students respond to automated feedback in the same way as feedback from live instructors?

####\

#How Did Automated Feedback Begin?
Interest in technology-assisted automated feedback on writing began in the mid-1960s. Systems have progressed from basic spelling and grammar checkers to automated essay scoring (AES) systems that provide feedback on the content of open-response writing. AES systems use natural language processing and machine learning algorithms to produce a score that is based on syntactic and lexical patterns and features such as sentence length, word frequency and part of speech [@Shermis2010, @Balfour2013].

AES systems have the advantage of being scalable and are very attractive for large face-to-face courses, and for asynchronous courses like massive open online courses [@Balfour2013]. Demand for AES systems has spurred development of several commercial products including Criterion®, WriteToLearn™, Intelligent Essay Assessor™, IntelliMetric®, WriteLab, Inc. and Turnitin Feedback Studio.

AES systems are not without drawbacks. First, it is controversial how accurately unsupervised AES systems score student work. Perelman [@Perelman2013] highlighted the poor accuracy of extant AES by using obviously faked texts to demonstrate how easily current AES applications are fooled. Second, it is unclear to what extent students pay attention to automated feedback. One small study [@Heffernan2015] compared relative impacts of teacher feedback versus technology driven feedback from ETS' Criterion® tool on 12 Japanese English-foreign- language students' writing. Results showed automated feedback alone was less effective than feedback from the teacher and Criterion® combined. However, the authors only compared scores on three independent, consecutive writing assignments; they did not compare effects of feedback on draft versus final versions of the same assignment. Also, the authors failed to compare automated feedback alone to teachers' comments alone. The degree of authority that students attach to automated feedback may also be a confounding variable. Ruegg [@Ruegg2015] found that feedback from peer reviewers was not valued as much as teachers' feedback, even though teachers provided less explicit comments than peer reviewers. Ruegg argued that this is likely because teachers are the ones that ultimately grade the student's writing. More research is needed to improve AES systems and to test their effectiveness in the classroom.

Despite their limitations, demand remains high for AES systems, and there are ongoing efforts to integrate them into more traditional writing review models [@Balfour2013]. Constructed response assessments reveal more about student thinking and the persistence of misconceptions than do multiple-choice questions. However they require more analysis on the part of the educator, and so are less likely to be used in large-enrollment courses. Urban-Lurain and colleagues developed automated assessment of constructed response (AACR) as a way to give students constructed response test questions and other work that can be graded automatically [@Kaplan2014, @Weston2015, @Ha2015, @Urban-Lurain2015, @Ha2016]. AACR uses a robust and scalable platform for scoring open responses to pre-written prompts that probe important disciplinary concepts from biology and biological chemistry. The prompts are administered via online course management systems where students enter responses. Lexical and statistical analyses then predict what expert ratings would be for student responses. Work to date shows that automated analyses can predict expert ratings of students' work on these topics accurately and with higher inter-rater reliability than a group of trained human graders. The AACR model is growing rapidly to include additonal STEM disicplines and has a national presence (http://create4stem.msu.edu/project/aacr). 
The success of AACR would suggest that the general lexical analysis principles this project is based on can be applied to the analysis of student writing. Our NSF-sponsored propoosal is exploring whether automated feedback affect how graduate teaching assistants (TAs) grade writing-related activities. Using the text classifier being built for this FLC project, we will determine whether TAs provide students with more feedback focused on higher order thinking, and less on low-level copy editing.

####\  
***

### 2019-02-26 background.Rmd
update

@@
 ---
-title: "Background"
+title: "Background of This Project"
+date: "`r format(Sys.time(), '%b %d, %Y')`"
 bibliography: library.bib
@@
 Undergraduate introductory STEM courses often include technical writing training.
-Too often it is treated as an incidental skill that each student develops "along the way." At other times the instructional strategy is  haphazardly designed, implemented, and executed.
+Too often it is treated as an incidental skill that each student develops "along the way." At other times the instructional strategy is  haphazardly designed, implemented, and executed [@holder].
 
 Adding to the learning challenge, faculty may not be the ones helping students develop writing skills.
-If technical writing training is part of a lab course, graduate teaching assistants or even advanced undergraduates may be who teaches technical writing.
-TAs may have under-developed writing skills themselves, lack pedagogical content knowledge needed to guide students effectively, or not have been explicitly trained to teach writing in a particular way.
+If technical writing training is part of a lab course, graduate teaching assistants or even advanced undergraduates may be who teaches technical writing [@holder].
+TAs may have under-developed writing skills themselves, lack pedagogical content knowledge needed to guide students effectively, or not have been explicitly trained to teach writing in a particular way [@holder].
 
@@
 
-*  Embody best practices from 40+ years of research on "writing in the disciplines;"
+*  Embody best practices from 40+ years of research on "writing in the disciplines";
 *  Emphasize student development through holistic coaching;
@@
 ##Context & Prior Work
-From prior studies (*see the Literature Reveiw for details*), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead.
+From prior studies (*see the Literature Reveiw for details*), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead [@holder].
 To that end, our program trains graduate TAs to use a **global issues / biggest issues first** approach to correcting undergraduate technical writing.
@@
 
-Our bins-based report grading protocol reinforces this feedback strategy.
+Our bins-based report grading protocol [@Nilson] reinforces this feedback strategy.
 
-This holistic, coaching-oriented approach is well-supported by prior research, but we are uncertain how accurately TAs implement the recommended strategy.
-We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"** Self-reported TA effort is likely to be unreliable.
+This holistic, coaching-oriented approach is well-supported by prior research [@holder], but we are uncertain how accurately TAs implement the recommended strategy.
+We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"** Self-reported TA effort is likely to be unreliable [@holder].
 What is needed is a more direct measure of TA effort.
@@
 
-We next used a subset of 110 randomly selected comments to build an initial qualitative codebook for classifying comments based on subject and structure.
+We next used a subset of 110 randomly selected comments to build an initial qualitative codebook [@holder] for classifying comments based on subject and structure.
 This initial codebook was tested and revised using blocks of 1100 additional randomly selected comments until the categories and criteria stabilized.
@@
 ####\  
+***

### 2019-02-26 build.Rmd
updates
update

--- 2019-02-22 build.Rmd
@@
 ---
-title: "The Build"
-author: "Dan Johnson"
-date: "2/22/2019"
-output: html_document
+title: "Project Build-Out"
+date: "`r format(Sys.time(), '%b %d, %Y')`"
+bibliography: library.bib
+biblio-style: apalike
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r, child="_setup.Rmd"}
 ```
 
-#Overview
-This page links to sub-pages that describe how various components used in the project were built. Follow the links for details.
+This page links to sub-pages describing construction of the various components used in the overall project. Progress on each component is in parentheses.
 
-Extracting TAs' comments: https://adanieljohnson.github.io/default_website/docxcommentextraction.html
+**Extracting TAs' Comments** *(Early coding stage)*  
+Protocol used to extract TA comments from student lab reports.  
+https://adanieljohnson.github.io/default_website/docxcommentextraction.html
 
-Structure of development dataset; data import & validation: https://adanieljohnson.github.io/default_website/dataimport.html
+**Structure of Development Dataset; Data Import & Validation** *(Completed)*  
+Organization of the initial dataset, and how data were imported in R and validated.  
+https://adanieljohnson.github.io/default_website/dataimport.html
 
-Current Naive Bayes classifier: https://adanieljohnson.github.io/default_website/NB_2r2c.html
+**Codebook for Classifying TAs' Comments** *(Completed)*  
+Describes the process for developing the codes used to classify comments.  
+https://adanieljohnson.github.io/default_website/codebook.html
+
+**Initial Exploration of the Dataset** *(Partly completed)*  
+What is the distribution of words used in the TA comments?  
+https://adanieljohnson.github.io/default_website/initial_exploration.html
+
+**Current Version of the Naive Bayes Comment Classifier** *(Completed)*  
+Code for the most recent version of the optimized NB classifier.   
+https://adanieljohnson.github.io/default_website/NB_2r2c.html
+
+**Embedding the Naive Bayes Classifier in an Ensemble Model** *(Planning stage)*  
+The NB method alone does not classify TA comments with sufficient accuracy for the "Structure" category.
+This page outlines a vertically stacked "ensemble" classifier that uses pattern matching and other pre-processing steps to improve NB accuracy.
+https://adanieljohnson.github.io/default_website/Ensemble_Notes.html
 
 ####\  
-
-#Describe an Element
-
-####\
-
-#Describe an Element
-
-####\
+***

### 2019-02-26 futurefaculty.Rmd
update

---
title: "Preparing Future Technical Writing Teachers"
date: "`r format(Sys.time(), '%b %d, %Y')`"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```

Policymakers are calling for faculty development programs that specifically address STEM education priorities [@NRC2010, @NRC2012, @Starke-Meyerring2011] including development of faculty across career stages and preparation of doctoral students and postdocs to be effective teachers. This needs to include how to teach science writing.

A major barrier to routine writing in STEM courses is instructor knowledge. Many instructors are unaware that there are publicly available, evidence-based writing training models; instead of using these validated tools, they create their own materials from scratch. similarly, many college STEM teachers have never been introduced to proven writing training strategies, nor have they been trained to give effective, context-specific feedback [@Coil2010].

Knowing which instructional models are effective is not sufficient to ensure adoption. Faculty may have different priorities or hindering attitudes by the time they teach their first classes as assistant professors. These attitudes underlie comments like “I don’t have time for something new,” “I don’t have spare class time to spend teaching writing,” or “students should learn how to write elsewhere” [@Coil2010, @Clughe2012, @FSSE2016]. These attitudes are very resistant to change. For example, despite a nationwide multi-year effort promoting interactive, inquiry- and skills-centric teaching, many faculty still believe their main goal should be content delivery, not development of students’ disciplinary thinking and process skills [@DAvanzo2013, @Freeman2014, @Gottschalk2003, @NRC2012, @VandC2011, @FSSE2016].

Even when faculty are motivated to innovate, how they learn about alternative strategies affects the odds of adoption. For example, faculty who attend formal professional development activities that promote evidence- based practices are less likely to adopt those practices than if they learn about them from trusted colleagues or respected opinion leaders [@Andrews2016]. In contrast, graduate students who participate in formal professional development about learner-centered instruction readily adopt learner-centered methods into their personal teaching practice [(]@Ebert-May2015]. A more effective strategy for changing faculty instructional practices may be to target and train future faculty who then act as change agents [@Tanner2006].

####\  
***

### 2019-02-26 index.Rmd
update

@@
 ##Work Plan
-The work plan for this project is:
+The work outline for this project is:
 
-1.  Compile an initial dataset for testing, and de-identify both TAs and students by replacing names with anonymous IDs.
+1.  Compile an initial dataset of hand-coded TA comments on reports for testing. Data will be de-identified by replacing both TAs' and students' names with anonymous IDs.
 2.  Import TA comments and relevant metadata from student reports into R as a "tidy" dataframe.
 3.  Use basic text analysis methods to explore comment data structure, n-gram frequencies, etc., and identify potentially useful elements for feature engineering.
-4.  Write a baseline workflow using Naive Bayes for supervised text classification.
-5.
-Assign TA comments to categories using the NB classifier, then establish baseline accuracy values using the optimized NB classifier by comparing outcomes of automated classification to hand-coded results.
-6.  Using features (from Step #3) and comparison data from NB classification protocol (from Step #5) as guides, further optimize the NB classifier method. 
-7.  When accuracy stabilizes, apply the optimized NB classifier to the original FULL comment dataset, identify all comments that are being classified incorrectly, and look for potential patterns.
+4.  Write a workflow using Naive Bayes for supervised text classification, and test various permutations of analysis parameters to establish a baseline classifier.
+5.  Assign TA comments to categories using the baseline NB classifier, and establish baseline accuracy values by comparing outcomes of automated classification to hand-coded results.
+6.  Using features (from Step #3) and comparison data from NB classification protocol (from Step #5) as guides, further optimize the NB classifier method until predicted accuracy stabilizes.
+7.  Apply the optimized NB classifier to the original FULL comment dataset, identify all comments that are being classified incorrectly, and look for potential patterns.
 8.  Write a small set of rule- or REGEX pattern-based searches that can identify comments which are more likely to be classified incorrectly. 
-9.  Append rules/pattern-based pre-screening to the optimized Naive Bayes classifier to create a vertical Ensemble Classifier.
+9.  Combine the rules/pattern-based pre-screening process with the optimized Naive Bayes classifier to create a vertical Ensemble Classifier.
 10.
-Validate the Ensemble Classifier using a subset of ~2000 new hand-coded TA comments extracted from a non-overlapping set of student lab reports.
+Re-validate the Ensemble Classifier using a subset of ~2000 new hand-coded TA comments extracted from student lab reports for a different semester than the initial test dataset.
 Compare automated results for new data set to hand-coded categories.
 
-Looking beyond the current project, I will be using the NB-based Ensemble Classifier as the control method for evaluating other text classifier strategies.
+Looking beyond the current project, I will use the vertical Ensemble Classifier two ways:
+
+*  To monitor TA comment patterns in the ongoing NSF-funded project.
+*  As the control method for evaluating other potential text classifier strategies.
 
@@
 #### \   
+***

### 2019-02-26 litreview.Rmd
update

--- 2019-02-25 litreview.Rmd
@@
 
-#Overview
-The supporting literature for this project is broad and convoluted.
-Rather than try to compress it to a single page, I have provided links out to sub-pages that describe various background topics.
-Follow the links for details.
+This project brings together three disparate challenges:
 
-Fundamental Principles and Practices of Technical Writing Training:  https://adanieljohnson.github.io/default_website/pagename.html
+*  The challenges of teaching technical writing in the STEM disciplines
+*  Needs and barriers for training future faculty to teach technical writing
+*  Automated text analysis and classification
 
-Primer on Text Classification Strategies:  https://adanieljohnson.github.io/default_website/Text_classifier_models.html
+Follow the links below (or from top menu) to sub-pages that explore each of these challenges in more detail.
+
+**Challenges of Teaching Technical Writing**  
+More than 50% of colleges and universities use some form of Writing in the Disciplines, Write to Learn, and Writing Across the Curriculum (WID/WTL/WAC) as part of their general education programs.
+There are well-established WAC/WID teaching principles based on >40 years of evidence and experience, but much of this work focuses on non-technical styles of writing.
+We know much less about teaching technical writing specifically.
+What do we know, and what can we bring from the WAC/WID community?
+What do we NOT know?
+https://adanieljohnson.github.io/default_website/techwriting.html
+
+**Challenges of Training Future Faculty**  
+A major barrier to routine writing in STEM is limited instructor awareness of evidence-based training models.
+Many college STEM teachers have never been introduced to proven writing training strategies or trained to give effective, context-specific feedback.
+How does this impact instruction more generally?
+https://adanieljohnson.github.io/default_website/futurefaculty.html
+
+**Challenges of Applying Automated Text Analysis to Student Learning**  
+Interest in technology-assisted automated feedback on writing began in the mid-1960s.
+Systems have progressed from basic spelling and grammar checkers to automated essay scoring (AES) systems that provide feedback on the content of open-response writing.
+Can automated text analysis support student learning?
+https://adanieljohnson.github.io/default_website/aacr.html
+
+**Text Classification 1 - Feature Selection and Engineering**  
+What is the general strategy and steps for solving a text classification problem? What features are important? 
+https://adanieljohnson.github.io/default_website/Text_Classifier_Models.html
+
+**Text Classification 2 - Topic Categorization and Modeling**  
+Establishing meaningful text classification categories depends on the goals of the project or the question being asked. Are relevant categories known already? If not, how are they established?  
+https://adanieljohnson.github.io/default_website/Topic_categorization.html
+
+**Text Classification 3 - Selecting an Appropriate Classifier Method**  
+There are three general classes of text classification strategies: pattern matching, algorithmic, and neural networks. Each is described briefly.  
+https://adanieljohnson.github.io/default_website/Choosing_classifier.html
+
+**Text Classification 4 - Evaluating and Improving the Classifier**  
+If there are many ways to classify texts, how do we know which one is most accurate?
+How can we improve the classification process?
+How do we make comparisons and evaluate the different steps in a text classification problem?
+https://adanieljohnson.github.io/default_website/Improving_classifier.html
 
 ####\  
+***

### 2019-02-26 nextsteps.Rmd
update

@@
 
-Qualitative coding of TA comments is informative but unsustainable.
-Developing the codebook then rating the first set of 11,000 comments required nearly 2 months of investigator effort.
-Even with the codebook, individual coders will require training to achieve any accuracy.
-Given 10-12,000 comments are generate EACH SEMESTER, an automated sorting method is needed.
+DRAFT PLACEHOLDER PAGE
+
+Lorem ipsum dolor sit amet...
 
 ####\  
+***

### 2019-02-26 notes.rmd
update

@@
     Writing the bibtext  
-    Embedding using [@ BaseR] method.
+    Embedding using @baseR method, or [@holder].  
+  Modify the icons:
+
+  Set menu so landing page is an "Overview"
+  Append this code to end of all pages
+```
+      #### \   
+      ***
+```
 
 ##Index:
-*  Decide on bottom page site map strategy
+*  Decide on bottom page site map strategy, and attach
 
@@
 ##Lit review:  
-*  Df  
+*  Finalize what content pages will be
     +  Df
@@
 *  How to graph the different comparisons of NB values from different sources on same plot? 
-    +  Add ggplot (data=) value.
-    +  Better to combine data into one dataset though.
+    +  Add ggplot (data=) value. SOLVED 
+    +  Better to combine data into one dataset though. (REQUIRED for proper legend formats)
 
@@
 #Etc.:
-*  Using a Bibliography in this format  
+*  Adding the Bibliography in this format
+    +  Need to add the new citation entries to library.bib
 *  Adding automatic updates to top dates? 
@@
 
-*  Fix GitHub auto-upload
+#Miscellaneous:
+*  Fix GitHub link on laptop to auto-upload without passwords again
 
@@
 ####\  
+888

### 2019-02-26 techwriting.Rmd
update

---
title: "The Challenges of Teaching Technical Writing"
date: "`r format(Sys.time(), '%b %d, %Y')`"
bibliography: library.bib
biblio-style: apalike
---

```{r, child="_setup.Rmd"}
```
Strong scientific thinking & communication skills are essential for 21st century STEM workers [@Anderman2011, @CWPA2011, @PCAST2012], but there are not enough college graduates with these skills to meet projected US workforce demands [@AACU2007, @BHEF2011, @Carnevale2011]. Policy makers have called on educators to prioritize development of career-relevant “21st-century skills” [@CWPA2011, @NRC2010, @NRC2012, @Starke-Meyerring2011]. These include quantitative and information literacy, adaptability, novel problem-solving, self-regulation, systems thinking, analytical and critical thinking skills, and complex communication skills [@Anderman2011] A more scientifically and computationally literate citizenry also will be better prepared to evaluate personal choices, societal issues, and larger policies that have deep connections to science and technology [@PCAST2012].

All of the aforementioned 21st-century skills are vital for college graduates, so why focus specifically on improving written communication in the form of scientific writing? First, these skills do not develop in isolation. As students master complex communication, they also develop critical thinking skills [@Anderman2011, @Carpenter2001, @Gorzelski2016, @Rounsaville2008, @Reiff2011, @Rogers2011, @Starke-Meyerring2011, @White2009]. By improving written communication we also can impact other skills. Second, good communication skills cut across all STEM fields. Even individuals working in relative isolation routinely record data and observations, interpret evidence, and share findings informally or formally either orally or in writing. Scientific writing skills are also valued and highly sought by employers independent of discipline or field of study [@Treadwell1999, @Gray2005, @Fry2014]. Third, writing is not just a mean for communication. It is also a tool to teach and develop scientific thinking. Scientific writing proficiency helps students be more literate and focused, and move from intuition-based thinking towards evidence-based thinking [@Libarkin2012, @Quitadamo2007, @Fellows1994].

Given projected workforce needs and that strong communications skills are vital across the range of careers, teaching scientific writing should be an obvious priority in STEM courses. Ideally, instructors would give students regular opportunities to practice critical and applied thinking by assigning writing routinely [@Reynolds2012, @Holyoak1998, @Anderman2011], but there are significant logistical barriers to this pedagogical ideal [@Haswell2006a, @Haswell2006b, @Henderson2007].

The most obvious logistical barrier is workload. Good scientific writing is one of the most difficult and time-consuming skills to teach. It is an iterative cycle of writing drafts, receiving actionable feedback and coaching, then making further refinements until a final product emerges. This cycle is too time- and labor-intensive to be practical in high-enrollment introductory courses. Even in lower-enrollment courses with more than 10-20 students, multiple assignments pose significant logistical challenges [@Coil2010, @Fry2014]. When students do have an opportunity to write for class, circumstances may permit just one round of revisions. The feedback they receive needs to be as focused, individualized, and impactful as possible.

Giving effective feedback is a challenging and complex task that needs to be learned through specific training [@Bazerman1994, @Bazerman2006, @Gottschalk2003, @Graves2011, @Kiefer2002, @NCTENWP, @Perelman2009, @Szymanski2014, @Underwood2006]. Effective feedback will be process-oriented rather than focused on finding and correcting errors. Comments will be understandable and unambiguous. Effective feedback keeps the writer in control and encourages them to revise their draft thoughtfully and thoroughly, and keeps copy-editing feedback to a minimum [@Breidenbach2006, @Ruegg2015, @Trupiano2006]. Over-commenting or giving overly prescriptive comments can push a student writer into a passive role. They treat comments as a to-do list that needs to be completed rather than suggestions for reflection and self-assessment [@Anson2001, @CWPA2011, @Harris1979, @Haswell2006a, @Haswell2006b, @Reynolds2009, @Szymanski2015, @Swilky1991, @Underwood2006].

There is clear need for more robust student training and assessment models that make it easier to give students multiple cycles of “holistic feedback,” that is, feedback that helps undergraduates identify their current shortcomings as a writer and directs them towards a path to improvement. An ideal model would: 1) embody best practices for teaching writing; 2) reduce instructor workload; 3) stress holistic feedback rather than copy-editing; 4) limit the overall number of comments to what students can manage; and 5) help students understand feedback and prioritize revision efforts appropriately.

####\

#Future Faculty Need Training on How to Teach Scientific Writing
This project focuses specifically on improving the feedback process in two ways. First, the training program teaches instructors how to give holistic feedback that is individualized and context-specific. Second, students will receive a portion of their feedback from SAWHET, a technology-assisted, automated feedback system.

A major barrier to routine writing in STEM is limited instructor awareness of evidence-based training models. Many college STEM teachers have not been introduced to proven writing training strategies or trained to give effective, context-specific feedback [@Coil2010]. Policymakers are calling for faculty development programs that specifically address STEM education priorities [@NRC2010, @NRC2012, @Starke-Meyerring2011] including development of faculty across career stages and preparation of doctoral students and postdocs to be effective teachers. This needs to include how to teach science writing.

Knowing which instructional models are effective is not sufficient to ensure adoption. Faculty may have different priorities or hindering attitudes by the time they teach their first classes as assistant professors. These attitudes underlie comments like “I don’t have time for something new,” “I don’t have spare class time to spend teaching writing,” or “students should learn how to write elsewhere” [@Coil2010, @Clughe2012, @FSSE2016]. These attitudes are very resistant to change. For example, despite a nationwide multi-year effort promoting interactive, inquiry- and skills-centric teaching, many faculty still believe their main goal should be content delivery, not development of students’ disciplinary thinking and process skills [@DAvanzo2013, @Freeman2014, @Gottschalk2003, @NRC2012, @VandC2011, @FSSE2016].

Even when faculty are motivated to innovate, how they learn about alternative strategies affects the odds of adoption. For example, faculty who attend formal professional development activities that promote evidence- based practices are less likely to adopt those practices than if they learn about them from trusted colleagues or respected opinion leaders [@Andrews2016]. In contrast, graduate students who participate in formal professional development about learner-centered instruction readily adopt learner-centered methods into their personal teaching practice [@Ebert-May2015]. A more effective strategy for changing faculty instructional practices may be to target and train future faculty ([@Tanner2006] and references therein) who then act as change agents.

A central goal of this project is to test a scalable model for teaching GTAs about the importance of scientific writing in students’ intellectual development and training them to use best-practice models. Specifically, TAs will learn how to: 1) lead students through the instructional process, 2) use validated methods and rubrics, 3) provide effective, growth-oriented feedback, and 4) grade consistently. One of the questions to be tested in this project is whether GTA conceptual knowledge and self-efficacy related to teaching scientific writing training will increase over the course of the program described in Project Activities.

####\  
***

### 2019-02-26 NB_2r2c.Rmd
update

@@
 
-This block is an alternative method to get the same data tables out using ```tidy``` format. It seems much less direct, but may be over-written.
+This block is an **alternative method** to get the same data tables out using ```tidy``` format. It seems much less direct, but may be over-written.
 ```{r tidy_prediction_table}
@@
 
-#Tabular and Graphical Comparisons of Coded and Predicted Categories
+#Comparing Coded and Predicted Categories
 These code blocks compare the hand-coded versus NB-predicted frequencies. There are separate comparisons for the Subject and Structure categories, and both tabular and graphical comparisons.

### 2019-03-13 Choosing_classifier.Rmd
update

@@
 title: "Choosing a Text Classifier Method"
-date: "2/21/2019"
+date: "`r format(Sys.time(), '%b %d, %Y')`"
 bibliography: library.bib
@@
 
-#Choosing a Text Classifier Method
 The many methods for classifying texts fall into three general categories:
@@
 *  Algorithms (how closely does the query text meet pre-set criteria?)
-*  Neural networks (the network decides its own criteria, aka machine learning)
+*  Neural networks (the network decides its own criteria; aka, machine learning)
 
-This edited excerpt nicely describes the overall approach and problems of text classification. It is from: **xxxxx**
+This edited excerpt from a tutorial on chat-bots nicely describes the overall approach and problems of text classification, and the differences between pattern matching and algorithmic classifiers.
+For the original version go to: https://medium.com/@gk_/how-chat-bots-work-dfff656a35e2
 
@@
 
-This classifier is “naive” because each word is treated as having no connection with other words in the sentence being classified.
-In other words it assumes independence between "features," in this case, words.
+This classifier is "naive" because each word is treated as having no connection with other words in the sentence being classified. It assumes independence between "features," in this case, words.
 
-For example, in the sentence “the fox jumped over the log”, there is no relationship between “jumped” and “fox” or “log”.
-In NLP, this collection of words is referred to as “a bag of words”.
+For example, in the sentence "the fox jumped over the log", there is no relationship between "jumped" and "fox" or "log".
+In NLP, this collection of words is referred to as "a bag of words".
 While this is incredibly naive, the classifier isn’t attempting to understand the meaning of a sentence, it’s trying to classify it.
@@
 Pattern Matchers
-Early chatbots used **pattern matching** to classify text and produce a response.
-This is often referred to as “brute force” as the author of the system needs to describe every pattern for which there is a response.
+Early chatbots used pattern matching to classify text and produce a response.
+This is often referred to as "brute force" as the author of the system needs to describe every pattern for which there is a response.
 
-A standard structure for these patterns is “AIML” (artificial intelligence markup language). Its use of the term “artificial intelligence” is quite an embellishment, but that’s another story.
+A standard structure for these patterns is "AIML" (artificial intelligence markup language). Its use of the term "artificial intelligence" is quite an embellishment, but that’s another story.
 
@@
 Likewise it responds to anything solely because of an authored pattern.
-Given hundreds or thousands of patterns you might see a chatbot “persona” emerge.
+Given hundreds or thousands of patterns you might see a chatbot "persona" emerge.
 
-In 2000 a chatbot built by John Denning and colleagues using this approach was in the news for passing the "Turing test." It emulated the replies of a 13 year old boy from Ukraine (broken English and all).
+In 2000 a chatbot built by John Denning and colleagues using this approach was in the news for passing the "Turing test".
+It emulated the replies of a 13 year old boy from Ukraine (broken English and all).
 I met with John in 2015 and he made no false pretenses about the internal workings of this automaton.
-It may have been “brute force” but it proved a point: parts of a conversation can be made to appear “natural” using a sufficiently large definition of patterns.
-It proved Alan Turing’s assertion, that this question of a machine fooling humans was “meaningless”.
+It may have been "brute force" but it proved a point: parts of a conversation can be made to appear "natural" using a sufficiently large definition of patterns.
+It proved Alan Turing’s assertion, that this question of a machine fooling humans was "meaningless".
 
@@
 One text classification algorithm is "Multinomial Naive Bayes." Given a training set of sentences, each belonging to a class (that is, a particular group), we can count the occurrence of each word in each class, account for its commonality and assign each word a score for each class.
-Factoring for commonality is important: matching the word “it” is considerably less meaningful than a match for the word “cheese”.
+Factoring for commonality is important: matching the word "it" is considerably less meaningful than a match for the word "cheese".
 When we input a new sentence, scores for each word in the new sentence are compared to scores for those same words in the training set.
@@
 
-I will be using an algorithm-based classification strategy for this project.
-
 ####\
@@
 ###Neural Networks
-A neural network is a mathematical model that is designed to behave similar to biological neurons and nervous system.
-These models are used to recognize complex patterns and relationships that exists within a labelled data.
+A neural network is a mathematical model that is designed to behave similarly to biological neurons and nervous systems.
+These models are used to recognize complex patterns and relationships that exists within a labeled dataset.
 
-Artificial neural networks, invented in the 1940’s, are a way of calculating an output from an input (a classification) using weighted connections (“synapses”) that are calculated from repeated iterations through training data.
-Each pass through the training data alters the weights such that the neural network produces the output with greater “accuracy” (lower error rate).
-
-There’s not much new about these structures, except today’s software is using much faster processors and can work with a lot more memory.
-The combination of working memory and speed is crucial when you’re doing hundreds of thousands of matrix multiplications (the neural network’s essential math operation).
+Artificial neural networks, invented in the 1940’s, calculate an output from an input (a classification) using weighted connections ("synapses") that are calculated from repeated iterations through training data.
+Each pass through the training data alters the weights such that the neural network produces the output with greater "accuracy" (lower error rate).
 
 Each class is given with some number of example sentences.
-Once again each sentence is broken down by word (stemmed) and each word becomes an input for the neural network.
+Each sentence is broken down by word (stemmed) and each word becomes an input for the neural network.
 The synaptic weights are then calculated by iterating through the training data thousands of times, each time adjusting the weights slightly to greater accuracy.
-By recalculating back across multiple layers (“back-propagation”) the weights of all synapses are calibrated while the results are compared to the training data output.
+By recalculating back across multiple layers ("back-propagation") the weights of all synapses are calibrated while the results are compared to the training data output.
 These weights are like a ‘strength’ measure, in a neuron the synaptic weight is what causes something to be more memorable than not.
@@
 
-At some point the adjustment reaches a point of diminishing returns, this is called “over-fitting” and going beyond this is counter-productive.
+At some point the adjustment reaches a point of diminishing returns, this is called "over-fitting" and going beyond this is counter-productive.
 
-The trained neural network is less code than an comparable algorithm but it requires a potentially large matrix of “weights”.
+A fully trained neural network is less code than a comparable algorithm The neural network boils down to matrix multiplication and a formula for reducing values between -1 and 1 or some other minimal range.
+The challenge is that it requires a potentially large matrix of "weights".
 In a relatively small sample, where the training sentences have 150 unique words and 30 classes this would be a matrix of 150x30.
 Imagine multiplying a matrix of this size 100,000 times to establish a sufficiently low error rate.
+
 This is where processing speed comes in.
+A combination of working memory and processor speed is crucial when you’re doing hundreds of thousands of matrix multiplications (the neural network’s essential math operation).
+Today’s software takes full advantage of faster processors and can work with a lot more memory.
 
-If the neural network sounds magnificently sophisticated, relax, it boils down to matrix multiplication and a formula for reducing values between -1 and 1 or some other minimal range.
-A middle-school math student could learn this in a few hours.
-The hard work is achieving clean training data.
+The other hard part is obtaining clean training data, which is essential for proper classification.
 
-Just as there are variations in pattern matching code and in algorithms, there are variations in neural networks, some more complex than others.
+Just as there are different pattern matching models and algorithms, there are different types of neural networks, some more complex than others.
 The basic machinery is the same.
-The essential work is that of classification.
-
 Networks are looking for patterns in collections of terms, each term is reduced to a token.
 In this machine words have no meaning except for their patterned existence within training data.
-The label “artificial intelligence” applied to such machinery is mostly BS.
 

### 2019-03-13 aacr.Rmd
update

@@
 ---
-title: "Automated Analysis of Student Writing Writing "
+title: "Automated Feedback on Student Writing"
 date: "`r format(Sys.time(), '%b %d, %Y')`"
@@
 
-There is growing need and interest in incorporating automated feedback into writing training. This raises important questions that have not yet been adequately answered:
+There is growing need and interest in using automated feedback as part of writing training. This raises important questions that have not yet been adequately answered:
 
-*  Does automated feedback identify similar issues with similar frequency as live instructors?
+*  Can/do automated feedback systems identify similar issues as live instructors, with similar frequency?
 *  When instructors have access to automated feedback, do they trust it? Do they provide feedback on different issues, focus on higher-order problems, or continue grading as if they had no insights?
@@
 
-#How Did Automated Feedback Begin?
-Interest in technology-assisted automated feedback on writing began in the mid-1960s.
-Systems have progressed from basic spelling and grammar checkers to automated essay scoring (AES) systems that provide feedback on the content of open-response writing.
-AES systems use natural language processing and machine learning algorithms to produce a score that is based on syntactic and lexical patterns and features such as sentence length, word frequency and part of speech [@Shermis2010, @Balfour2013].
+#Origins of Automated Feedback
+Interest in automated feedback on writing emerged in the mid-1960s.
+Systems have since progressed from basic spelling checkers to syntax-aware spelling and grammar evaluation.
+More recently, automated essay scoring (AES) systems emerged that can provide feedback on open-response writing.
+AES systems use natural language processing and machine learning algorithms to produce a score based on syntactic and lexical patterns and features such as sentence length, word frequency and part of speech [@Shermis2010, @Balfour2013].
 
-AES systems have the advantage of being scalable and are very attractive for large face-to-face courses, and for asynchronous courses like massive open online courses [@Balfour2013].
 Demand for AES systems has spurred development of several commercial products including Criterion®, WriteToLearn™, Intelligent Essay Assessor™, IntelliMetric®, WriteLab, Inc.
@@
 
-AES systems are not without drawbacks.
-First, it is controversial how accurately unsupervised AES systems score student work.
-Perelman [@Perelman2013] highlighted the poor accuracy of extant AES by using obviously faked texts to demonstrate how easily current AES applications are fooled.
+AES systems have the advantage of being scalable and are very attractive for large face-to-face courses, and for asynchronous courses like massive open online courses [@Balfour2013].
+Still they are not without drawbacks.
+First, un-/minimally-supervised AES systems may not score student work consistently.
+Perelman [@Perelman2013] highlighted their poor accuracy, using obviously faked texts to demonstrate how easily current AES applications are fooled.
 Second, it is unclear to what extent students pay attention to automated feedback.
@@
 Also, the authors failed to compare automated feedback alone to teachers' comments alone.
-The degree of authority that students attach to automated feedback may also be a confounding variable.
+
+The degree of authority students attach to both automated and human feedback also is a confounding variable.
+Anecdotally, most instructors can cite many examples of students ignoring basic spelling and grammar feedback that is standard for word processors like MS.
+Word.
 Ruegg [@Ruegg2015] found that feedback from peer reviewers was not valued as much as teachers' feedback, even though teachers provided less explicit comments than peer reviewers.
 Ruegg argued that this is likely because teachers are the ones that ultimately grade the student's writing.
-More research is needed to improve AES systems and to test their effectiveness in the classroom.
 
@@
 The AACR model is growing rapidly to include additonal STEM disicplines and has a national presence (http://create4stem.msu.edu/project/aacr).
-The success of AACR would suggest that the general lexical analysis principles this project is based on can be applied to the analysis of student writing.
+
+The success of AACR **suggests** that the general lexical analysis principles this project is based on can be applied to the analysis of student writing.
+However, more research is needed before we can say with any confidence that automated feedback benefits students or their instructors.
+
 Our NSF-sponsored propoosal is exploring whether automated feedback affect how graduate teaching assistants (TAs) grade writing-related activities.

### 2019-03-13 build.Rmd
update

@@
 
-**Extracting TAs' Comments** *(Early coding stage)*  
+####\
+
+####Extracting TAs' Comments 
+*(Early coding stage)*  
 Protocol used to extract TA comments from student lab reports.  
@@
 
-**Structure of Development Dataset; Data Import & Validation** *(Completed)*  
+####Structure of Development Dataset; Data Import & Validation 
+*(Completed)*  
 Organization of the initial dataset, and how data were imported in R and validated.  
@@
 
-**Codebook for Classifying TAs' Comments** *(Completed)*  
+####Codebook for Classifying TAs' Comments 
+*(Completed)*  
 Describes the process for developing the codes used to classify comments.  
@@
 
-**Initial Exploration of the Dataset** *(Partly completed)*  
+####Initial Exploration of the Dataset 
+*(Partly completed)*  
 What is the distribution of words used in the TA comments?  
@@
 
-**Current Version of the Naive Bayes Comment Classifier** *(Completed)*  
+####Current Version of the Naive Bayes Comment Classifier 
+*(Completed)*  
 Code for the most recent version of the optimized NB classifier.   
@@
 
-**Embedding the Naive Bayes Classifier in an Ensemble Model** *(Planning stage)*  
+####Embedding the Naive Bayes Classifier in an Ensemble Model  
+*(Planning stage)*
 The NB method alone does not classify TA comments with sufficient accuracy for the "Structure" category.

### 2019-03-13 litreview.Rmd
update

@@
 title: "Literature Review"
-date: "`r format(Sys.time(), '%b %d, %Y')`"
+date: "Updated `r format(Sys.time(), '%b %d, %Y')`"
 bibliography: library.bib
@@
 
-**Challenges of Teaching Technical Writing**  
+####\
+
+####Challenges of Teaching Technical Writing
 More than 50% of colleges and universities use some form of Writing in the Disciplines, Write to Learn, and Writing Across the Curriculum (WID/WTL/WAC) as part of their general education programs.
@@
 
-**Challenges of Training Future Faculty**  
+####Challenges of Training Future Faculty
 A major barrier to routine writing in STEM is limited instructor awareness of evidence-based training models.
@@
 
-**Challenges of Applying Automated Text Analysis to Student Learning**  
+####Challenges of Applying Automated Text Analysis to Student Learning
 Interest in technology-assisted automated feedback on writing began in the mid-1960s.
@@
 
-**Text Classification 1 - Feature Selection and Engineering**  
+####Text Classification 1 - Feature Selection and Engineering
 What is the general strategy and steps for solving a text classification problem? What features are important? 
@@
 
-**Text Classification 2 - Topic Categorization and Modeling**  
+####Text Classification 2 - Topic Categorization and Modeling
 Establishing meaningful text classification categories depends on the goals of the project or the question being asked. Are relevant categories known already? If not, how are they established?  
@@
 
-**Text Classification 3 - Selecting an Appropriate Classifier Method**  
+####Text Classification 3 - Selecting an Appropriate Classifier Method
 There are three general classes of text classification strategies: pattern matching, algorithmic, and neural networks. Each is described briefly.  
@@
 
-**Text Classification 4 - Evaluating and Improving the Classifier**  
+####Text Classification 4 - Evaluating and Improving the Classifier
 If there are many ways to classify texts, how do we know which one is most accurate?

### 2019-03-13 techwriting.Rmd
update

@@
 Strong scientific thinking & communication skills are essential for 21st century STEM workers [@Anderman2011, @CWPA2011, @PCAST2012], but there are not enough college graduates with these skills to meet projected US workforce demands [@AACU2007, @BHEF2011, @Carnevale2011].
-Policy makers have called on educators to prioritize development of career-relevant “21st-century skills” [@CWPA2011, @NRC2010, @NRC2012, @Starke-Meyerring2011].
+Policy makers have called on educators to prioritize development of "career-relevant 21st-century skills” [@CWPA2011, @NRC2010, @NRC2012, @Starke-Meyerring2011].
 These include quantitative and information literacy, adaptability, novel problem-solving, self-regulation, systems thinking, analytical and critical thinking skills, and complex communication skills [@Anderman2011] A more scientifically and computationally literate citizenry also will be better prepared to evaluate personal choices, societal issues, and larger policies that have deep connections to science and technology [@PCAST2012].
@@
 
-#Future Faculty Need Training on How to Teach Scientific Writing
+#Faculty Need Training to Teach Technical Writing Well
 This project focuses specifically on improving the feedback process in two ways.

### 2019-03-13 litreview.Rmd
update

@@
 
-*  The challenges of teaching technical writing in the STEM disciplines
+*  The challenges of teaching technical writing to students in STEM disciplines
 *  Needs and barriers for training future faculty to teach technical writing
@@
 
-Follow the links below (or from top menu) to sub-pages that explore each of these challenges in more detail.
+Sub-pages (links below and in top menu) explore each of these challenges in detail.
 
@@
 ####Challenges of Teaching Technical Writing
-More than 50% of colleges and universities use some form of Writing in the Disciplines, Write to Learn, and Writing Across the Curriculum (WID/WTL/WAC) as part of their general education programs.
-There are well-established WAC/WID teaching principles based on >40 years of evidence and experience, but much of this work focuses on non-technical styles of writing.
-We know much less about teaching technical writing specifically.
-What do we know, and what can we bring from the WAC/WID community?
+Over 50% of colleges and universities have some form of Writing in the Disciplines, Write to Learn, or Writing Across the Curriculum (WID/WTL/WAC) as part of their general education programs.
+There are well-established WAC/WID teaching principles for non-technical writing based on >40 years of evidence and experience, but we know much less about teaching technical writing.
+What do we know, and what can we import from the WAC/WID community?
 What do we NOT know?
@@
 ####Challenges of Training Future Faculty
-A major barrier to routine writing in STEM is limited instructor awareness of evidence-based training models.
+One major barrier to incorporating more writing in STEM courses is limited instructor awareness of evidence-based training models.
 Many college STEM teachers have never been introduced to proven writing training strategies or trained to give effective, context-specific feedback.
@@
 ####Challenges of Applying Automated Text Analysis to Student Learning
-Interest in technology-assisted automated feedback on writing began in the mid-1960s.
-Systems have progressed from basic spelling and grammar checkers to automated essay scoring (AES) systems that provide feedback on the content of open-response writing.
-Can automated text analysis support student learning?
+How well does automated text analysis support student learning?
+Can it extend what instructors do beyond current capabiities, or does it only replicate what instructors already do?
+Students claim to want more opportunities for feedback and revision, but do they actually act on automated feedback?
 https://adanieljohnson.github.io/default_website/aacr.html
@@
 ####Text Classification 1 - Feature Selection and Engineering
-What is the general strategy and steps for solving a text classification problem? What features are important? 
+What is the general strategy for solving a text classification problem? What features of texts are important?  
 https://adanieljohnson.github.io/default_website/Text_Classifier_Models.html
@@
 How can we improve the classification process?
-How do we make comparisons and evaluate the different steps in a text classification problem?
+How do we make comparisons and evaluate the outcomes of a text classification problem?
 https://adanieljohnson.github.io/default_website/Improving_classifier.html

### 2019-03-13 futurefaculty.Rmd
update

@@
 biblio-style: apalike
+nocite: | 
+  @Nilson2014
 ---
@@
 For example, faculty who attend formal professional development activities that promote evidence- based practices are less likely to adopt those practices than if they learn about them from trusted colleagues or respected opinion leaders [@Andrews2016].
-In contrast, graduate students who participate in formal professional development about learner-centered instruction readily adopt learner-centered methods into their personal teaching practice [(]@Ebert-May2015].
+In contrast, graduate students who participate in formal professional development about learner-centered instruction readily adopt learner-centered methods into their personal teaching practice [@Ebert-May2015].
 A more effective strategy for changing faculty instructional practices may be to target and train future faculty who then act as change agents [@Tanner2006].
@@
 ***
+*Works Cited*

### 2019-03-13 Topic_categorization.Rmd
update

@@
 
-#Topic Categorization and Modeling Methods
-Establishing meaningful text classification categories depends on the goals of the project or the question being asked.
-Are relevant categories known already, or do they need to be established?
-What are the main topics
+Establishing meaningful text classification categories depends on the goals of the project or the question being asked. Are relevant categories known already, or do they need to be established?
 
@@
 
-LSA automatically adapts to new and changing terminology, and has been shown to be very tolerant of noise (i.e., misspelled words, typographical errors, unreadable characters, etc.).[24] This is especially important for applications using text derived from Optical Character Recognition (OCR) and speech-to-text conversion.
+LSA automatically adapts to new and changing terminology, and has been shown to be very tolerant of noise (i.e., misspelled words, typographical errors, unreadable characters, etc.).
+This is especially important for applications using text derived from Optical Character Recognition (OCR) and speech-to-text conversion.
 LSA also deals effectively with sparse, ambiguous, and contradictory data.

### 2019-03-13 background.Rmd
update

--- 2019-02-22 background.Rmd
@@
 Undergraduate introductory STEM courses often include technical writing training.
-Too often it is treated as an incidental skill that each student develops "along the way." At other times the instructional strategy is  haphazardly designed, implemented, and executed.
-
+Too often it is treated as an incidental skill that each student develops "along the way." At other times writing instruction is haphazardly designed, implemented, and executed.
 Adding to the learning challenge, faculty may not be the ones helping students develop writing skills.
@@
 TAs may have under-developed writing skills themselves, lack pedagogical content knowledge needed to guide students effectively, or not have been explicitly trained to teach writing in a particular way.
+
+The essential first step towards improving the current situation is widespread adoption of evidence-based writing training.
+Equally important is reducing the practical barriers that keep faculty from routinely including scientific writing in post-secondary STEM classes.
 
@@
 ##Context & Prior Work
-From prior studies (*see the Literature Reveiw for details*), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead.
+From prior studies (*see the Literature Review for details*), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead [@Bazerman1994; @Bazerman2006; @Breidenbach2006; @Gottschalk2003; @Graves2011; @Kiefer2002; @NCTENWP; @Perelman2009; @Ruegg2015; @Szymanski2014;@Trupiano2006; @Underwood2006; @Anson2001; @CWPA2011; @Harris1979; @Haswell2006a; @Haswell2006b; @Reynolds2009; @Szymanski2014; @Swilky1991].
 To that end, our program trains graduate TAs to use a **global issues / biggest issues first** approach to correcting undergraduate technical writing.
@@
 
-Our bins-based report grading protocol reinforces this feedback strategy.
+Our bins-based report grading protocol [@Nilson2014] reinforces this feedback strategy.
 
-This holistic, coaching-oriented approach is well-supported by prior research, but we are uncertain how accurately TAs implement the recommended strategy.
+This holistic, coaching-oriented approach is well-supported by prior research and is a recommended best practice by (among others) the Writing Across the Curriculum Clearinghouse.
+Yet we are uncertain how accurately TAs implement the recommended strategy.
 We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"** Self-reported TA effort is likely to be unreliable.
@@
 
-We next used a subset of 110 randomly selected comments to build an initial qualitative codebook for classifying comments based on subject and structure.
+We next used a subset of 110 randomly selected comments to build an initial qualitative codebook [@Elliott2018; @Saldana2015] for classifying comments based on subject and structure.
 This initial codebook was tested and revised using blocks of 1100 additional randomly selected comments until the categories and criteria stabilized.
-Using the final codebook, we then sorted all 11,000 comments into separate categories (see PAGE_LINK for complete codebook.)
+Using the final codebook, we then sorted all 11,000 comments into separate categories.
 
@@
 ####\  
+***
+*Work Cited*

### 2019-03-13 build.Rmd
update

@@
 The NB method alone does not classify TA comments with sufficient accuracy for the "Structure" category.
-This page outlines a vertically stacked "ensemble" classifier that uses pattern matching and other pre-processing steps to improve NB accuracy.
+This page outlines a draft plan for a vertically stacked "ensemble" classifier that uses pattern matching and other pre-processing steps to improve NB accuracy.
 https://adanieljohnson.github.io/default_website/Ensemble_Notes.html
@@
 ####\
-***

### 2019-03-13 index.Rmd
update

--- 2019-02-21 index.Rmd
@@
 
-In Summer 2018, we extracted and tabulated ~11,000 comments made by teaching assistants (TA) from >500 pairs (initial submissions and revised version) of short lab reports written by undergraduates in 3 different general biology lab courses during a single semester.
-Using an iteratively generated codebook, each of the comments was separately classified in two categories:
+Classifying TA comments has been informative, but doing it by hand is impractical: 10-12,000 comments need to be scored EACH semester.
+There also is a significant risk of "coding drift" over time, reducing accuracy.
 
-* __Subject__: did the comment focus mainly on **Basic Criteria, Technical Issues, Writing Quality, Logic and Thinking,** or **Other** issues
-* __Structure__: did the comment **Point** to an error without providing other help, provide simple **Copy Correction**, provide declarative **General or Specific Information** only, or provide **Holistic Coaching** that fosters student thinking and long-term improvement?
+My project for the **Faculty Learning Community for R** was to create an automated comment classifier that can sort TA comments from student report into the same categories that we use for hand-coded data.
+This automated classifier will be used within the larger NSF-funded research project to:
+
+*  Improve both speed and consistency when classifying TA comments across multiple semesters. 
+*  Monitor how the key concepts and points that TAs stress in comments on student writing change over time.
 
 #### \   
-Counts of coded TA comments were tabulated according to subject and structure, and frequencies for each subject/structure pair calculated as a fraction of all comments.
-This 2-way contingency table provided us with a direct estimate of the most common types of comments TAs made on student work, and a relative estimate of which elements of writing TAs emphasize most often.
+##Work Plan
+The plan for this project is to:
+
+1.  Compile an anonymized testing dataset of hand-coded TA comments from reports. 
+2.  Import TA comments and relevant metadata from student reports into R as a "tidy" dataframe.
+3.  Explore comment data structure, n-gram frequencies, etc., and identify potential elements for feature engineering.
+4.  Write a supervised text classification workflow that uses Naive Bayes to assign TA comments to our pre-defined categories. 
+5.  Test permutations of analysis parameters to optimize the classifier and establish baseline accuracy.
+6.  Apply the optimized NB classifier to the original FULL comment dataset, identify which comments are being classified incorrectly, and look for potential patterns.
+7.  Based on features identified in Steps #3 and #6, write a small set of rule- or REGEX pattern-based searches that can identify comments that are more likely to be classified incorrectly.
+
+After completing the FLC project, my goals are to:
+
+8.  Combine the rules/pattern-based pre-screening process with the optimized Naive Bayes classifier to create a vertically  Ensemble Classifier.
+9.  Re-validate the Ensemble Classifier against a subset of ~2000 independently hand-coded TA comments extracted from student lab reports for a different semester than the initial test dataset.
+10. Begin using the classifier to monitor TA comment patterns in the ongoing NSF-funded project.
+
+Looking further ahead, I hope to use the work presented here as a baseline for evaluating other potential text classification models that are described in the Background.
 
 #### \
-##Specific Goals
-Classifying TA comments this way is informative, but doing so by hand is impractical: 10-12,000 comments need to be scored EACH semester.
-There also is a significant risk of "coding drift" over time, reducing accuracy.
 
-My project for the **Faculty Learning Community for R** was to create an automated comment classifier that could generate two-axis contingency tables comparable to the one generated using hand-coded data.
-This automated classifier will used within the larger NSF-funded research project to:
-
-*  Improve both speed and consistency when classifying TA comments across multiple semesters. 
-*  Monitor how the types and focus of TAs comments on student writing change over time.
-
-#### \   
 ##Project Site Map

### 2019-03-13 notes.rmd
update

@@
 ##General/Site-Wide:
-  Embedding the bibliographic items  
-    Writing the bibtext  
-    Embedding using @baseR method, or [@holder].  
-  Modify the icons:
 
-  Set menu so landing page is an "Overview"
-  Append this code to end of all pages
++  Write bibliographic entries for missing citations. 
++  Set menu so landing page of each topic is an "Overview"
++  Modify bottom code of all pages. If has citations, make sure the page has:
 ```
@@
 ```
++  If a page has no citations, REMOVE the bottom code for the horizontal rule. It now is in the footer.
++  Beloe is the standard output format for citation tags in my copy of BibDesk. To make the tags work properly, need to search and replace the pre-string.
+
+```\@[Carpenter2001] is output. Change "\@[" to "[@"```
 
 ##Index:
-*  Decide on bottom page site map strategy, and attach
+*  Decide on bottom page site map strategy, and attach.
 
@@
 ####\
-888

### 2019-03-13 techwriting.Rmd
update

@@
 Over-commenting or giving overly prescriptive comments can push a student writer into a passive role.
-They treat comments as a to-do list that needs to be completed rather than suggestions for reflection and self-assessment [@Anson2001, @CWPA2011, @Harris1979, @Haswell2006a, @Haswell2006b, @Reynolds2009, @Szymanski2015, @Swilky1991, @Underwood2006].
+They treat comments as a to-do list that needs to be completed rather than suggestions for reflection and self-assessment [@Anson2001, @CWPA2011, @Harris1979, @Haswell2006a, @Haswell2006b, @Reynolds2009, @Swilky1991, @Underwood2006].
 
@@
 ***
+*Work Cited*

### 2019-03-16 aacr.Rmd
update

@@
 Word.
-Ruegg [@Ruegg2015] found that feedback from peer reviewers was not valued as much as teachers' feedback, even though teachers provided less explicit comments than peer reviewers.
+Ruegg [-@Ruegg2015] found that feedback from peer reviewers was not valued as much as teachers' feedback, even though teachers provided less explicit comments than peer reviewers.
 Ruegg argued that this is likely because teachers are the ones that ultimately grade the student's writing.

### 2019-03-16 background.Rmd
update

@@
 ##Context & Prior Work
-From prior studies (*see the Literature Review for details*), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead [@Bazerman1994; @Bazerman2006; @Breidenbach2006; @Gottschalk2003; @Graves2011; @Kiefer2002; @NCTENWP; @Perelman2009; @Ruegg2015; @Szymanski2014;@Trupiano2006; @Underwood2006; @Anson2001; @CWPA2011; @Harris1979; @Haswell2006a; @Haswell2006b; @Reynolds2009; @Szymanski2014; @Swilky1991].
-To that end, our program trains graduate TAs to use a **global issues / biggest issues first** approach to correcting undergraduate technical writing.
-Ideally:
+From prior studies (*see the Literature Review for details*), we know students make greater gains as writers if instructors avoid telling students what to correct ("copy editing"), and use a coaching-oriented approach instead [@Bazerman1994; @Bazerman2006; @Breidenbach2006; @Gottschalk2003; @Graves2011; @Kiefer2002; @NCTENWP; @Perelman2009; @Ruegg2015; @Szymanski2014; @Trupiano2006; @Underwood2006; @Anson2001; @CWPA2011; @Harris1979; @Haswell2006a; @Haswell2006b; @Reynolds2009; @Szymanski2014; @Swilky1991].
+
+To that end, our program trains graduate TAs to use a **global issues / biggest issues first** approach to correcting undergraduate technical writing. Ideally:
 

### 2019-03-16 build.Rmd
update

--- 2019-03-13 build.Rmd
@@
 ####Extracting TAs' Comments 
-*(Early coding stage)*  
-Protocol used to extract TA comments from student lab reports.  
+*(Preliminary coding)*  
+Fully R-based protocol for extracting TA comments from student lab reports. Will replace current process using a mix of Linux and Excel scripts.  
 https://adanieljohnson.github.io/default_website/docxcommentextraction.html
@@
 ####Structure of Development Dataset; Data Import & Validation 
-*(Completed)*  
+*(Complete)*  
 Organization of the initial dataset, and how data were imported in R and validated.  
@@
 ####Codebook for Classifying TAs' Comments 
-*(Completed)*  
+*(Complete)*  
 Describes the process for developing the codes used to classify comments.  
@@
 
-####Initial Exploration of the Dataset 
-*(Partly completed)*  
+####Initial Exploration of Dataset 
+*(Partly complete)*  
 What is the distribution of words used in the TA comments?  
@@
 
-####Current Version of the Naive Bayes Comment Classifier 
-*(Completed)*  
+####Naive Bayes Comment Classifier 
+*(Complete)*  
 Code for the most recent version of the optimized NB classifier.   
@@
 
-####Embedding the Naive Bayes Classifier in an Ensemble Model  
-*(Planning stage)*
-The NB method alone does not classify TA comments with sufficient accuracy for the "Structure" category.
-This page outlines a vertically stacked "ensemble" classifier that uses pattern matching and other pre-processing steps to improve NB accuracy.
+####Embedding Naive Bayes Classifier in an Ensemble Model  
+*(Initial design stage)*
+Outlines a strategy for combining pattern matching and other pre-processing steps with Naive Bayes Classifier to improve accuracy.
 https://adanieljohnson.github.io/default_website/Ensemble_Notes.html

### 2019-03-16 dataimport.Rmd
update

@@
 1.  Student and TA names have been recoded using pre-assigned anonymous IDs.
-2.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.rmd"
-3.  Data table headers match those listed below. The same list is in the file "code_column_names.rmd"
+2.  Vocabulary terms used in coding columns matches comment codebook criteria described at https://adanieljohnson.github.io/default_website/codebook.html.
+3.  Data table headers match those listed below.
 
@@
 
-Data were anonymized in Excel, replacing names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings **[LINK]**.
+Data were anonymized in Excel, replacing names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings.
 
@@
 
-Data were anonymized in Excel prior to conversion to CSV format. Names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings **[LINK]**.
+Data were anonymized in Excel prior to conversion to CSV format. Names, email addresses, and other confidential information was replaced with unique randomized alphanumeric strings.
 
@@
 1.  Student and TA names have been recoded using pre-assigned anonymous IDs.
-2.  Vocabulary terms used in coding columns matches criteria in "codebook_for_comments.rmd" [LINK]
-3.  Data table headers match those listed below. The same list is in the file "code_column_names.rmd" [LINK]
+2.  Vocabulary terms used in coding columns matches comment codebook criteria described at https://adanieljohnson.github.io/default_website/codebook.html.
+3.  Data table headers match those listed below.
 

### 2019-05-12 Classifier_comparison.Rmd
update
updates

@@
 title: "Designing a Classifier Evaluation Workflow"
-date: "2/21/2019"
+date: "`r format(Sys.time(), '%b %d, %Y')`"
 bibliography: library.bib
@@
 One of the many problems with this approach is that identifying and organizing a list of terms (preferred, alternate, etc.) is labor intensive (much like simple pattern matching), and the natural ambiguity of language (one keyword can have different meanings) causes lots of false positives.
-This makes the bag of keywords approach not only not scalable, but also inaccurate.
+These limitations make the bag of keywords approach both poorly scalable and less accurate.
 
@@
 In the past, he has been right 65% of times.
-
 Given the broad spectrum of access we have, we can probably combine all the information and make an informed decision.

### 2019-05-12 NB_2r2c.Rmd
update
updates

@@
 This block predicts whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. 
-```{r bayes_training}
+```{r bayes_training subject}
 # Train the classifier on training data, then test on test dataset, and full dataset.
@@
 comments_all_pred_subject <- predict(comments_classifier_subject, comments_all)
+```
 
+```{r bayes_training structure}
 comments_classifier_structure <- naiveBayes(comments_train, comments_train_labels_structure, laplace=1)
@@
 This block is an **alternative method** to get the same data tables out using ```tidy``` format. It seems much less direct, but may be over-written.
+
 ```{r tidy_prediction_table}
@@
 #Use mutate to add a proportion of total column. And yes, it calculates based on subgroups rather than whole column.
-prop_tally1<-mutate(tidy_tally1, prop = n/sum(n))
+prop_tally1<-mutate(tidy_tally1, prop = Freq/sum(Freq))
 prop_tally1<-select(prop_tally1, 1,2,4)
-prop_tally2<-mutate(tidy_tally2, prop = n/sum(n))
+prop_tally2<-mutate(tidy_tally2, prop = Freq/sum(Freq))
 prop_tally2<-select(prop_tally2, 1,2,4)
-prop_tally3<-mutate(tidy_tally3, prop = n/sum(n))
+prop_tally3<-mutate(tidy_tally3, prop = Freq/sum(Freq))
 prop_tally3<-select(prop_tally3, 1,2,4)
-prop_tally4<-mutate(tidy_tally4, prop = n/sum(n))
+prop_tally4<-mutate(tidy_tally4, prop = Freq/sum(Freq))
 prop_tally4<-select(prop_tally4, 1,2,4)

### 2019-05-12 N_gram_exploration.rmd
updates

--- 2019-02-24 initial_exploration.Rmd
@@
 ---
-title: "Initial Exploration"
-date: "2/21/2019"
+title: "N-gram Frequency Analysis"
+date: "`r format(Sys.time(), '%b %d, %Y')`"
 bibliography: library.bib
@@
 
-#Overview
-The first question to answer before making deeper comparisons is, **are word frequencies different between TAs' comments from different sub-categories?** This page describes the methods used for this initial exploration.
-
-Results of these analyses are discussed on this page: https://adanieljohnson.github.io/default_website/findings.html
-
-####\
-
-#The Analyses
-I ran three different analyses of word usage frequency.
-
-*  __Pearson correlations for word use frequencies__. This only compares words found in all of the Subject sub-categories
-*  __Word frequency differences based on log-odds ratios__. This highlights terms that are the most different between sub-categories.
-*  __N-grams mapping__ (*draft*). This identifies text strings, phrases, and terms that are unique (or nearly unique) to one sub-category.
-
-####\
-
-#Setup
-```{r Libraries_1}
-# Call in tidyverse, other required libraries.
-library(tidyverse)
-library(tidytext)
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-library(scales)
-
-#Read in CSV as a starting dataframe named "base_data". 
-base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
-```
-
-####\
-
-#Pearson Correlations of Word Frequencies
-##Are Shared Words Used At Different Frequencies?
-This analysis takes all comments from base_data, tokenizes and filters stopwords, then counts and sorts fractional frequencies for each word in each category.
-**NOTE**: this analysis only compares word use for terms present in ALL sub-categories.
-If a word is MISSING from **any subcategory**, the data row for that word is pruned from the dataset.
-This eliminates all unique words in favor of words found in all sub-categories.
-
-```{r}
-#Tokenize phrases, remove the stop words listed in R's standard reference file.
-base_data_tokenized <- base_data %>% 
-  unnest_tokens(word,ta.comment)
-
-#ALTERNATIVE METHOD for pulling out stopwords.  
-sortedwords.comments.all <- base_data_tokenized %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate total # of words
-total_words.all <- sortedwords.comments.all %>% 
-  summarize(total = sum(n))
-
-#Calculate then append fractional word frequencies.
-sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
-  all.fraction = (n*100)/total_words.all$total
-  )
-
-#Select SUBSETs of tokenized data based on code.subject: 
-base_data_tokenized.basic <- subset(base_data_tokenized, code.subject == "1_basic", select = (1:28))
-base_data_tokenized.writing <- subset(base_data_tokenized, code.subject == "2_writing", select = (1:28))
-base_data_tokenized.technical <- subset(base_data_tokenized, code.subject == "3_technical", select = (1:28))
-base_data_tokenized.logic <- subset(base_data_tokenized, code.subject == "4_logic", select = (1:28))
-
-#For each subset, filter stopwords, count words and total, then calculate frequencies.
-sortedwords.basic <- base_data_tokenized.basic %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-#Calculate the number of words in total in the subset
-total_words.basic <- sortedwords.basic %>% 
-  summarize(total = sum(n))
-
-#Mutate table to calculate then append fractional word frequency in the text.
-sortedwords.basic.fraction <- sortedwords.basic %>% as_tibble() %>% mutate(
-  basic.fraction = (n*100)/total_words.basic$total
-)
-#Repeats for subset 2
-sortedwords.writing <- base_data_tokenized.writing %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-total_words.writing <- sortedwords.writing %>% 
-  summarize(total = sum(n))
-
-sortedwords.writing.fraction <- sortedwords.writing %>% as_tibble() %>% mutate(
-  writing.fraction = (n*100)/total_words.writing$total
-)
-
-#Repeats for subset 3
-sortedwords.technical <- base_data_tokenized.technical %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-total_words.technical <- sortedwords.technical %>% 
-  summarize(total = sum(n))
-
-sortedwords.technical.fraction <- sortedwords.technical %>% as_tibble() %>% mutate(
-  technical.fraction = (n*100)/total_words.technical$total
-)
-
-#Repeats for subset 4
-sortedwords.logic <- base_data_tokenized.logic %>%
-  filter(!word %in% stop_words$word,
-         !word %in% str_remove_all(stop_words$word, "'"),
-         str_detect(word, "[a-z]")) %>%
-  count(word, word, sort = TRUE) %>%
-  ungroup()
-
-total_words.logic <- sortedwords.logic %>% 
-  summarize(total = sum(n))
-
-sortedwords.logic.fraction <- sortedwords.logic %>% as_tibble() %>% mutate(
-  logic.fraction = (n*100)/total_words.logic$total
-)
-```
-
-The next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed. 
-```{r}
-total <- merge(sortedwords.comments.all.fraction,sortedwords.technical.fraction,by="word") 
-total2 <- merge(total,sortedwords.writing.fraction,by="word")
-total3 <- merge(total2,sortedwords.logic.fraction,by="word")
-#rename lines that were possibly confusing
-names(total3)[2] <- "all.count"
-names(total3)[4] <- "technical.count"
-names(total3)[6] <- "writing.count"
-names(total3)[8] <- "logic.count"
-
-#Mutate table to calculate then append fractional word frequency in the text.
-total4 <- total3 %>% as_tibble() %>% mutate(
-  writing.wt = (writing.count *100)/all.count
-)
-
-total5 <- total4 %>% as_tibble() %>% mutate(
-  technical.wt = (technical.count *100)/all.count
-)
-
-total6 <- total5 %>% as_tibble() %>% mutate(
-  logic.wt = (logic.count *100)/all.count
-)
-
-```
-
-The following six graphs compare frequencies for ~770 words that appear in ALL three sub-categories: Writing, Technical, and Logic.
-```{r}
-ggplot(total6, aes(`all.fraction`,`writing.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-
-ggplot(total6, aes(`all.fraction`,`technical.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-
-ggplot(total6, aes(`all.fraction`,`logic.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-
-ggplot(total6, aes(`technical.fraction`,`writing.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-
-ggplot(total6, aes(`logic.fraction`,`writing.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-
-ggplot(total6, aes(`logic.fraction`,`technical.fraction`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-```
-
-Graphs 1-3 show frequencies in individual groups versus word frequency overall.
-Generally, frequency of terms in the whole set of comments should be similar to frequency in an individual group (X vs Y positions similar.) However, some terms associated with Logic subgroup (Graph 3) are far from the best-fit line, suggesting they are under-used for comments in that sub-category.
-
-Graphs 4-6 show pair-wise comparisons between sub-categories.
-Unlike overall frequency, most words are NOT clustered along the best-fit line, suggesting there is still more variation in word use between sub-categories.
-
-####\
-
-#Log-Odds Ratios of Word Frequencies
-##Which Words Are Shared LEAST Between Categories?
-Unlike the preceding analysis, log-odds ratio comparison does not eliminate unshared terms.
-Instead, it identifies which words are MOST likely to be used in one sub-category, but not another.
-For simplicity I reduced the analysis to just the four most common comment subjects.
-This has been the most informative preliminary visualization I have found so far.
-
-This code block is based on a method for comparing word usage frequencies between two separate Twitter users [https://www.tidytextmining.com/twitter.html](described here.) If the words used by TAs in their comments differ between sub-categories, there should be fairly low Pearson correlations between word frequencies.
-
-```{r}
-# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
-frequency_writing <- filter(base_data, code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")
-frequency_writing.subcolumns <- frequency_writing %>% select(1, 22:23)
-
-# Tokenize phrases, remove stop words listed in standard reference file.
-base_data_tokenized <- frequency_writing.subcolumns %>% 
-  unnest_tokens(word,ta.comment) %>% 
-  anti_join(stop_words)
-
-#At this point the dataframe should have "Unique.Record" in column 1, "code.subject" in column 2, and the unnested "word" in column 3.
-All columns are class = "character".
-Stop words listed in the "stop_words" R reference data file have been removed.
-
-#Block below re-organizes the data for comparisons. This pattern sets "2_writing" as the first dataset, then all other data in remaining columns.
-#Groups by code.subject, calculated proportional frequency in each group.
-#Final step creates 4 columns: word, 2_writing, code.subject, proportion
-#Stats need this format to be able to compare 2_writing against other values.
-base_data_tokenized_sorted2 <- base_data_tokenized %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  mutate(proportion = n / sum(n)) %>% 
-  select(-n) %>% 
-  spread(code.subject, proportion) %>% 
-  gather(code.subject, proportion, `3_technical`:`4_logic`)
-
-#The dataframe called "base_data_tokenized_sorted2" is the final 4-column dataframe for statistical analysis.
-Individual words are in column 1, "2_writing" frequency is in Column 2, "code.subject" is in column 3, and "proportion" is in column 4.
-Columns 1 and 3 are class = "character", 2 and 4 are numeric ranges.
-
-#This structure is needed in order to compare word (in Column 1) frequencies for 2.
-Writing Quality (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes.
-WHICH subset of X values to use is coded in Column 3.
-
-#Pearson correlation between frequency of words in 2_writing versus 3_technical and 2_writing versus 4_logic sub-categories.
-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3_technical",], 
-         ~ proportion + `2_writing`)
-
-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4_logic",], 
-         ~ proportion + `2_writing`)
-
-coroutput23 <-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3_technical",], 
-         ~ proportion + `2_writing`)
-coroutput24 <-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4_logic",], 
-         ~ proportion + `2_writing`)
-
-#Comparing sub-groups 3_technical and 4_logic requires recreating the dataframe in a slightly different format that excludes the data rows for 2_writing.
-
-# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
-frequency_writing3 <- filter(base_data, code.subject=="3_technical"|code.subject=="4_logic")
-frequency_writing.subcolumns3 <- frequency_writing3 %>% select(1, 22:23)
-
-# Tokenize phrases, remove stop words listed in standard reference file.
-base_data_tokenized3 <- frequency_writing.subcolumns3 %>% 
-  unnest_tokens(word,ta.comment) %>% 
-  anti_join(stop_words)
-
-#At this point the dataframe should have "Unique.Record" in column 1, "code.subject" in column 2, and the unnested "word" in column 3.
-All columns are class = "character".
-Stop words listed in the "stop_words" R reference data file have been removed.
-
-#Block below re-organizes the data for comparisons. This pattern sets "3_technical" as the first dataset, then all other data in remaining columns.
-#Groups by code.subject, calculated proportional frequency in each group.
-#Final step creates 4 columns: word, 3_technical, code.subject, proportion
-#Stats need this format to be able to compare 3_technical against other values.
-base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  mutate(proportion = n / sum(n)) %>% 
-  select(-n) %>% 
-  spread(code.subject, proportion) %>% 
-  gather(code.subject, proportion,`4_logic`)
-
-#The dataframe called "base_data_tokenized_sorted3" is the final 4-column dataframe for statistical analysis.
-Individual words are in column 1, "3_technical" frequency is in Column 2, "code.subject" is in column 3, and "proportion" is in column 4.
-Columns 1 and 3 are class = "character", 2 and 4 are numeric ranges.
-
-#This structure is needed in order to compare word (in Column 1) frequencies for 3_technical (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes.
-WHICH subset of X values to use is coded in Column 3.
-
-#Pearson correlation** between the frequency of words in 3_technical versus 4_logic sub-categories.
-coroutput34 <-cor.test(data = base_data_tokenized_sorted3[base_data_tokenized_sorted3$code.subject == "4_logic",], 
-         ~ proportion + `3_technical`)
-
-broom::tidy(coroutput23)
-broom::tidy(coroutput24)
-broom::tidy(coroutput34)
-```
-
-####\
-
-###Visualizing Frequency Differences Between Sub-Categories
-
-```{r}
-#Create groups by code.subject, calculates proportional frequency in each group.
-#Final step creates 5 columns: code.subject, n, word, total, frequency.
-base_data_tokenized_sorted2 <- base_data_tokenized %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  left_join(base_data_tokenized %>% 
-  group_by(code.subject) %>% 
-  summarise(total = n())) %>%
-  mutate(freq = n/total)
-
-base_data_tokenized_sorted2v3 <- base_data_tokenized_sorted2 %>% 
-  select(code.subject, word, freq) %>% 
-  spread(code.subject, freq) %>%
-  arrange(`2_writing`,`3_technical`)
-
-ggplot(base_data_tokenized_sorted2v3, aes(`2_writing`,`3_technical`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-  #ggsave("Freqplot_3vs2.png", width = 10, height = 6)
-
-#Comparing 2 vs. 4.
-base_data_tokenized_sorted2v4 <- base_data_tokenized_sorted2 %>% 
-  select(code.subject, word, freq) %>% 
-  spread(code.subject, freq) %>%
-  arrange(`2_writing`,`4_logic`)
-
-ggplot(base_data_tokenized_sorted2v3, aes(`2_writing`,`4_logic`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-  #ggsave("Freqplot_4vs2.png", width = 10, height = 6)
-
-#Create groups by code.subject, calculates proportional frequency in each group.
-#Final step creates 5 columns: code.subject, n, word, total, frequency.
-base_data_tokenized_sorted3 <- base_data_tokenized %>% 
-  group_by(code.subject) %>%
-  count(code.subject, word, sort=TRUE) %>% 
-  left_join(base_data_tokenized %>% 
-  group_by(code.subject) %>% 
-  summarise(total = n())) %>%
-  mutate(freq = n/total)
-
-#Set up comparison for 3 vs. 4.
-base_data_tokenized_sorted3v4 <- base_data_tokenized_sorted3 %>% 
-  select(code.subject, word, freq) %>% 
-  spread(code.subject, freq) %>%
-  arrange(`2_writing`,`4_logic`)
-
-ggplot(base_data_tokenized_sorted3v4, aes(`3_technical`,`4_logic`)) +
-  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
-  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
-  scale_x_log10(labels = percent_format()) +
-  scale_y_log10(labels = percent_format()) +
-  geom_abline(color = "red")
-  #ggsave("Freqplot_4vs3.png", width = 10, height = 6)
-```
-
-###Calculating and Plotting Log-Odds Pairs for Subject Sub-Categories
-
-```{r}
-word_ratios2v3 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`2_writing`/`3_technical`)) %>%
-  arrange(desc(logratio))
-
-word_ratios3v4 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`3_technical`/`4_logic`)) %>%
-  arrange(desc(logratio))
-
-word_ratios2v4 <- base_data_tokenized %>%
-  count(word, code.subject) %>%
-  group_by(word) %>%
-  filter(sum(n) >= 10) %>%
-  ungroup() %>%
-  spread(code.subject, n, fill = 0) %>%
-  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
-  mutate(logratio = log(`2_writing`/`4_logic`)) %>%
-  arrange(desc(logratio))
-
-word_ratios2v3 %>% 
-  arrange(abs(logratio))
-word_ratios3v4 %>% 
-  arrange(abs(logratio))
-word_ratios2v4 %>% 
-  arrange(abs(logratio))
-```
-
-```{r}
-word_ratios2v3 %>%
-  group_by(logratio < 0) %>%
-  top_n(25, abs(logratio)) %>%
-  ungroup() %>%
-  mutate(word = reorder(word, logratio)) %>%
-  ggplot(aes(word, logratio, fill = logratio < 0)) +
-  geom_col(show.legend = FALSE) +
-  coord_flip() +
-  ylab("log odds ratio (Writing-top/Technical-bottom)") +
-  scale_fill_discrete(name = "", labels = c("2_writing", "3_technical"))
-  #ggsave("Logplot_2vs3.png", width = 6, height = 10)
-
-word_ratios3v4 %>%
-  group_by(logratio < 0) %>%
-  top_n(25, abs(logratio)) %>%
-  ungroup() %>%
-  mutate(word = reorder(word, logratio)) %>%
-  ggplot(aes(word, logratio, fill = logratio < 0)) +
-  geom_col(show.legend = FALSE) +
-  coord_flip() +
-  ylab("log odds ratio (Technical-top/Logic-bottom)") +
-  scale_fill_discrete(name = "", labels = c("3_technical","4_logic"))
-  #ggsave("Logplot_3vs4.png", width = 6, height = 10)
-
-word_ratios2v4 %>%
-  group_by(logratio < 0) %>%
-  top_n(25, abs(logratio)) %>%
-  ungroup() %>%
-  mutate(word = reorder(word, logratio)) %>%
-  ggplot(aes(word, logratio, fill = logratio < 0)) +
-  geom_col(show.legend = FALSE) +
-  coord_flip() +
-  ylab("log odds ratio (Writing-top/Logic-bottom)") +
-  scale_fill_discrete(name = "", labels = c("2_writing", "4_logic"))
-  #ggsave("Logplot_2vs4.png", width = 6, height = 10)
-
-```
+Another comparison looks at frequencies of N-grams between different sub-categories. This page describes the process used for initial exploration of the TA comments dataset.
 
@@
 ```
+
+####\  
+***
+*Work Cited*

### 2019-05-12 Text_classifier_models.Rmd
update
updates

@@
 ---
-title: "Text Classification Primer"
+title: "Primer of Text Classification"
 date: "2/21/2019"
@@
 
-1.  Feature selection and engineering (described here)
+1.  Feature selection and engineering (*described below*)
 2.  Topic categorization and modeling (https://adanieljohnson.github.io/default_website/Topic_categorization.html)
 3.  Selecting an appropriate text classifier method (https://adanieljohnson.github.io/default_website/Choosing_classifier.html)
-4.  Evaluating the outcomes, and improving the classifier ( https://adanieljohnson.github.io/default_website/Improving_classifier.html)
+4.  Evaluating the outcomes, and improving the classifier (https://adanieljohnson.github.io/default_website/Improving_classifier.html)
 
@@
 	+  TF-IDF
-	+  Reading level estimates, readability (described below)
+	+  Reading level estimates, readability (*described below*)
 *  Part of Speech Tagging and Frequency Distribution of Tags 
@@
 *  Word choices (vocabulary range)
-*  Lexical diversity (described below)
+*  Lexical diversity (*described below*)
 *  Stance patterns - hedging, booster, generalization words or phrasing 
@@
 Altszyler et al.
-(2017) studied Word2vec performance in two semantic tests for different corpus size.
+[@Altszyler2017] studied Word2vec performance in two semantic tests for different corpus size.
 They found that Word2vec has a steep learning curve, outperforming another word-embedding technique (LSA) when it is trained with medium to large corpus size (more than 10 million words).

### 2019-05-12 Topic_categorization.Rmd
update
updates

@@
 *  Prior empirical studies
-*  The dataset itself, using *de novo* iterative coding. This process is described further here [LINK] and here [LINK].
+*  The dataset itself, using *de novo* iterative coding. This process is described further here: https://adanieljohnson.github.io/default_website/codebook.html
 
@@
 ####Latent Dirichlet Allocation
-(LDA) is a common method for generating topic model features.
+LDA is a common method for generating topic model features.
 LDA is an iterative model which starts from a fixed number of topics.

### 2019-05-12 aacr.Rmd
updates

@@
 
-There is growing need and interest in using automated feedback as part of writing training. This raises important questions that have not yet been adequately answered:
+There is growing need and interest in using automated feedback as part of writing training. This raises several questions that have not yet been adequately answered:
 
@@
 More recently, automated essay scoring (AES) systems emerged that can provide feedback on open-response writing.
-AES systems use natural language processing and machine learning algorithms to produce a score based on syntactic and lexical patterns and features such as sentence length, word frequency and part of speech [@Shermis2010, @Balfour2013].
+AES systems use natural language processing and machine learning algorithms to produce a score based on syntactic and lexical patterns and features such as sentence length, word frequency and part of speech [@Shermis2010; @Balfour2013].
 
@@
 However they require more analysis on the part of the educator, and so are less likely to be used in large-enrollment courses.
-Urban-Lurain and colleagues developed automated assessment of constructed response (AACR) as a way to give students constructed response test questions and other work that can be graded automatically [@Kaplan2014, @Weston2015, @Ha2015, @Urban-Lurain2015, @Ha2016].
+Urban-Lurain and colleagues developed automated assessment of constructed response (AACR) as a way to give students constructed response test questions and other work that can be graded automatically [@Kaplan2014; @Weston2015; @Ha2015; @Urban-Lurain2015; @Ha2016].
 AACR uses a robust and scalable platform for scoring open responses to pre-written prompts that probe important disciplinary concepts from biology and biological chemistry.

### 2019-05-12 background.Rmd
updates

@@
 Undergraduate introductory STEM courses often include technical writing training.
-Too often it is treated as an incidental skill that each student develops "along the way." At other times writing instruction is haphazardly designed, implemented, and executed.
-Adding to the learning challenge, faculty may not be the ones helping students develop writing skills.
-If technical writing training is part of a lab course, graduate teaching assistants or even advanced undergraduates may be who teaches technical writing.
+Often it is secondary to content coverage, and so treated as an incidental skill that each student develops "along the way." Alternatively, the writing instruction process may be haphazardly designed, implemented, and executed.
+Adding to the learning challenge, faculty may not be who is helping students develop writing skills.
+Often graduate teaching assistants or even advanced undergraduates are who teaches technical writing.
 TAs may have under-developed writing skills themselves, lack pedagogical content knowledge needed to guide students effectively, or not have been explicitly trained to teach writing in a particular way.
 
-The essential first step towards improving the current situation is widespread adoption of evidence-based writing training.
+The first step towards improving writing training is widespread adoption of evidence-based strategies.
 Equally important is reducing the practical barriers that keep faculty from routinely including scientific writing in post-secondary STEM classes.
 
-In 2014 we launched a report data intake process we call **SAWHET** that systematically collects writing and other data from student writers and graduate instructors.
-Using these data we are mapping key features of student writing, as well as connections and correlations between those features.
-These insights then inform how we teach technical writing and train graduate instructors.
+In 2014 we launched the **SAWHET** lab report collection system, to systematically collect report documents and metadata from student writers and graduate instructors.
+Using these data we are mapping key features of student writing, and connections and correlations between those features.
+These insights are informing both how we teach technical writing and how we train graduate instructors.
 
-Currently our structured database contains >5000 laboratory reports written by undergraduate students for introductory biology courses, all in standardized .docx and .txt formats.
+Currently our structured database contains >5000 laboratory reports written by undergraduate students for introductory biology courses in standardized .docx and .txt formats.
 Additional data and metadata available for each report includes:
@@
 
-To that end, our program trains graduate TAs to use a **global issues / biggest issues first** approach to correcting undergraduate technical writing. Ideally:
+To that end, we train graduate TAs to comment on undergraduate technical writing by addressing **global issues / biggest issues first**. Ideally, TAs will provide feedback within these parameters:
 
-*  TAs spend a shorter amount of time writing fewer comments that focus on broader issues.
-*  TAs provide no more than 3-5 comments per text page.
-*  Comments encourage student reflection and self-correction rather than prescribe specific changes. 
-*  As students' writing improves, TAs shift their focus to the next most-significant errors or issues.
+*  Spend less time writing fewer but more substantive comments, and provide at least two separate rounds of feedback on each document. 
+    +  In routine practice, limit feedback to 3-5 comments per text page. Students cannot process more than this easily.
+    +  Make recommendations that would result in the greatest improvement first.
+    +  As students' writing improves, TAs can shift their focus to the next most-significant errors or issues.
+*  Provide comments that encourage student reflection and self-correction rather than prescribe specific changes. 
+    +  Students tend to focus on the smallest, easiest corrections first, and avoid larger, more complex revisions.
+    +  Eliminating copy corrections leaves students with only the complex revisions to work on.
 
@@
 Yet we are uncertain how accurately TAs implement the recommended strategy.
-We have repeatedly asked ourselves, **"what do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"** Self-reported TA effort is likely to be unreliable.
-What is needed is a more direct measure of TA effort.
+We have repeatedly asked ourselves:
+
+**"What do TAs ACTUALLY spend most of their time and effort making comments about when they grade student writing?"**
+
+Self-reported TA effort is likely to be unreliable. What is needed is a more direct measure of TA effort.
 
@@
 * __Subject__: did the comment focus mainly on **Basic Criteria** (minimum pass/fail requirements), **Technical Issues, Writing Quality, Logic and Thinking,** or **Other** issues
-* __Structure__: did the comment **Point** to an error without providing other help, provide simple **Copy Correction**, provide declarative **General or Specific Information** only, or provide **Holistic Coaching** that fosters student thinking and long-term improvement?
+* __Structure__: did the comment simply **Point** to an error without providing other help, provide only **Copy Correction**, provide declarative **General or Specific Information**, or provide **Holistic Coaching** that fosters student thinking and long-term improvement?
 

### 2019-05-12 codebook.rmd
update
updates

@@
 title: "The TA Comments Codebook"
-date: "2/23/2019"
+date: "`r format(Sys.time(), '%b %d, %Y')`"
 bibliography: library.bib
@@
 #Subject Category
-Is primary emphasis of the comment on:
+Does the comment focus on or emphasize:
 
-*  Fundamental pieces
+*  Fundamental required components?
 *  How to communicate generally?
@@
 
-Below are the choices for identifying the subject of the comment.
+Final subject categories for comments are outlined below.
 Criteria for basic, technical, and writing choices are the same as used in the report grading rubric.
-To simplify sorting, administrative tags used to mark comments for further processing are included in this category.
+Administrative tags used to mark comments for further processing are included in this category.
 
@@
 #Structure Category
-Is the comment idiomatic or informative? Is there general or specific information contained in comment? Is it instructional only, or does the comment foster broader thinking?
+Is the comment idiomatic or informative? Is there general or specific information contained in the comment? Is it instructional only, or does the comment foster broader thinking?
 
@@
 *  ??
-*  Misspelling, typo, format, check, etc.
-*  *Repeated phrase. (Only points to immediate error; compare to below)
+*  One-word comments: misspelling, typo, format, check, etc.
+*  Simple phrases. (Only point to immediate error; compare to below)
 
@@
 
-####Contingency Table For Holistic Scores
+####Contingency Table For Deciding Whether to Assign Holistic Tag
 
-Scores need to have BOTH information and process (what and why) to be holistic
+A comment needs to have BOTH information and process (what and why) to be holistic.
 
@@
 *  Always use correct format for scientific names, i.e., italicize or underline. 
-(Idiomatic because not required in our grading scheme. Incidentally, this is not accurate form either, which is one answer to ongoing arguments.)
+(Idiomatic because this not required in our grading scheme. Incidentally, this is not accurate form either, which is one answer to ongoing arguments.)
 *  You provided both scientific and common names for your plant species, so can use common names throughout the rest of the report including the title and abstract.
 Revise accordingly.
-(Idiomatic because this is not standard to all fields)
+(Idiomatic because this is not standard to all fields, **and** is not a high-priority issue in our grading scheme.)
 
@@
 There are two questions in the comment, which suggests that it is a holistic comment to pull knowledge out through thinking.
-However, the TA comment provides some interpretation in the next sentence.
-So this comment is seeking or providing specific information, while also directly fostering student’s deeper thinking.
+The comment provides some interpretation in the next sentence, but is not just prescriptive.
+This comment is seeking or providing specific information, while also directly asking student to engage in deeper thinking.
 It would be tagged as “Holistic."
+
+####\

### 2019-05-12 findings.Rmd
update
updates

--- 2019-02-21 findings.Rmd
@@
 
-##Classifier Accuracy
+*Updated `r format(Sys.time(), '%b %d, %Y')`*
+
+#Vocabulary and Word Use Patterns
+##Results of Single Word Frequencies
+Pearson correlations for word use frequencies compares absolute use frequencies for words found in ALL THREE of the Subject sub-categories (excluding Basic Criteria).
+Results of one analysis are illustrated below.
+
+![*XY plot of relative word use frequency in technical comments vs.
+logic comments.
+*](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/docs/initial_exploration_files/figure-html/unnamed-chunk-3-6.png)
+
+####\
+
+Results are summarized below.
+
+Comparison|Correlation|95% CI|p-value
+----------|-----------|------|-------
+Writing vs. Technical|0.404|0.355 - 0.451|1.64e-47
+Writing vs. Logic|0.559|0.514 - 0.602|3.60e-79
+Logic vs. Technical|0.394|0.343 - 0.443|3.05e-42
+
+####\
+
+The comparatively low correlations of word frequencies between subject categories suggests that some terms can serve as diagnostic keywords for particular subjects.
+
+####\
+
+##Results of Log-Odds Pairs for Subject Sub-Categories
+Word frequency differences between sub-categories were calculated using log-odds ratios. One example analysis is shown below.
+
+![*Log-Odds ratio comparison of words most different between Writing comments (top) and Technical comments (bottom).
+*](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/docs/initial_exploration_files/figure-html/unnamed-chunk-7-1.png)
+
+####\
+
+Words identified by individual analyses were collated into two lists: highly predictive (those appearing in two comparisions) and moderately predictive (those found in one comparison).
+
+####\
+
+###Words for Subject = 2. Writing 
+__Highly predictive__
+Broad, broader, broadly; colloquial; language; quote, quotes, quotations; revise; tense
+
+__Moderately predictive__
+Abstract; avoid; belongs; concise; concluding; details; formal; grammar; hypothesize; ideas; implications; importance; incorrect; informal; judgement; language; locate; methods; paraphrase; precise; read; recipe; redundant; repetitive; salinity; save; starting; style; takeaway; title; typos; voice; word; wrote
+
+####\
+
+###Words for Subject = 3. Technical
+__Highly predictive__
+Average, averaged, averages; axis; bar, bars; caption, captions; deviation; figure, figures; format, formatting; graph, graphs; legend; paired; raw; represent; sd; standard; table, tables; units
+
+__Moderately predictive__
+Chloroplast; day; df; error; label; ml; properly; stat; total
+
+####\
+
+###Words for Subject = 4. Logic
+__Highly predictive__
+Connect, connection; expect, expected; relate; shelter
+
+__Moderately predictive__
+Affecting; benefits; biological; broader; build; calculated; chloroplast; concluding; costs; density; dots; error; evolution; expand; explore; finding; fitness; gain; grow; hypothesize; ideas; implications; importance; isolated; logical; mechanisms; prediction; procedural; relative; relevance; resources; salinity; sample; selection; sodium; temp; tie; weren't
+
+####\
+
+##Results of N-Gram Frequency Comparisons
+I also tried a limited scale test of an alternative approach, using 2-word phrases (2-grams) instead of single words to classify TA comments.
+Lists of 2-grams and relative frequencies of appearance were calculated for TA comments in all Subject categories (Basic, Writing, Technical, Logic).
+N-gram lists were sorted by appearance frequencies, and truncated at a within-subject frequency of <1%, producing a starting dataset of ~550 2-grams.
+
+All 2-grams occurring at similar rates in 3 out of 4 Subject categories were classified as too general and deleted, leaving a working dataset of ~338 2-grams (out of original 550).
+Remaining 2-grams were tagged as potential identifiers of 1-2 Subject areas.
+The individual 2-gram lists were merged back to the master set and sorted alphabetically.
+
+The compiled set was visually scanned to identify individual 2-grams and 2-gram patterns that:
+
+* Appeared in at least 1% of all comments in one sub-group
+* Appeared at a 10-fold greater frequency in one sub-group versus all others.
+
+Full analysis of the dataset requires writing a coded search algorithm; for now, I am treating these results as s "proof-of-concept" demonstration only.
+
+This is a partial list of the potentially useful 2-grams identified so far.
+
+***
+*  Subject = Basic
+    + "basic criteria/criterion" (present in 25.12% of all Basic comments)
+    + "missing outside"" (6.64% of all Basic comments)
+    + "outside source(s)"" (13.75% of all Basic comments)
+    + "criteria __"
+    + Single words (1-grams) strongly correlated with Basic category: basic, credit, grade, criteria, fail/failing, outside
+*  Subject = Writing
+    + "abstract __"
+    + "avoid __"
+    + "capitalize __" 
+    + "colloquial __"
+*  Subject = Technical
+    + amplitude
+    + analysis/analyze
+    + "a figure/graph/legend"
+    + "a paired"
+    + alpha level/value"
+    + "annotation exercise"
+    + "anova __"
+    + "average __"
+    + "bar(s) __"
+    + "caption(s) __"
+*  Subject = Logic
+    + "a biological"
+    + "affect __"
+    + "aggression/aggressive __"
+    + "allocation __"
+    + "axis __"
+    + "betta(s) __" (Is name of model system typically logic?)
+    + big(ger) picture/concept is logic
+    + connect "__"
+*  N-grams Identifying TWO Subjects
+    + "a citation" is basic or technical
+    + "a significant" is logic or technical
+    + "about __" is logic, possibly writing. Weaker example.
+    + "background __" is writing or logic
+    + "because that's" is writing/technical
+    + "(bio)core guide/resource" is technical or writing
+    + "biological __" is logic or writing
+    + "broader __" is logic or writing
+    + "data __" is most often technical, but occurs in others.
+
+####\
+
+#Classifier Accuracy
 The tables below show observed frequencies for 9,340 TA comments coded by hand, and frequencies for the same TA comments categorized using the optimized Naive Bayes classifier **without** pre-screening.
@@
 
-##Data Story 2
+##Test of Horizontal Stacked Naive Bayes
+One strategy to refine Naive Bayes was running the same classifier with different randomized datasets, then putting individual outputs side by side and assigning final categories based on most frequent classification in the set.
+
+To test this I ran 5 Subject classifier replicates using the code in *Naive_bayes_V2r_2b.rmd*, with 5 different ```set.seed``` values.
+The exported CSV files were combined to get a set of data with up to 5 classifier assignments for the same comment.
+
+Using multiple randomized NB training runs does not seem to help.
+First, it was rare for the  randomized runs to assign different incorrect classes.
+Second, when I compared 5 independently randomized runs by eye, incorrectly categorized items were repeatedly put into the SAME incorrect group each time.
+Given this, an Ensemble method of H-stacked categorical NB comparison seems unlikely to reduce overall errors in classification.
+
+Looking at the joined data DID provide other helpful insights into the overall classification procedure though. Naive Bayes does not seem to do well with comments that:
+
+*  Contain many or mostly punctuation symbols, such as:
+    +  : not ,
+    +  ?
+    +  ??
+    +  ???
+*  Are whole numbers
+    +  1
+    +  2
+    +  10
+*  Are one word questions or comments
+    +  because?
+    +  "and"
+*	 Are sentences enclosed in punctuation
+*  Are made up entirely of STOP/common words (I am not as sure on this one)
 
@@
 
-#Tool or Analysis 2
+##Overall Accuracy
+I looked at OVERALL accuracy of NB by recreating the 2-dimensional contingency table that I made when hand-coding, and comparing it side-by-side to NB classifications.
+From this I gained additional insights.
+Results are summarized below.
 
@@
 
-##Data Story 3
+![alt text](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/docs/NB_2r2c_files/figure-html/unnamed-chunk-4-1.png)
+
+![alt text](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/docs/NB_2r2c_files/figure-html/unnamed-chunk-6-1.png)
 
@@
 
-##Data Story 4
+OVERALL, the 2-D Naive Bayes model performed very well for classifying comment **Subject**.
+When I re-ran the full original dataset (9340 comments) through the prediction model, NB estimates for Subject were within 1.5% of my hand-scored assignments.
+Structure classification was not as robust.
+Other observations:
+
+*  NB over-classified comments as "copyediting of technical issues" by 8-10%, and under-classified comments that should be marked as "specific technical issues" by almost same amount.
+*  Comments that belonged to groups comprising <1-2% of all comments were more likely to be mis-classified. 
+*  Longer text comments were mis-classified more often. This suggests two follow-up methods to evaluate: 
+    +  Splitting longer comments into shorter phrases
+    +  Evaluating only the first 5-10 words of a comment
+
+It appears that Subject categories are assigned mainly on the terminology used, while Structure categories are assigned based more on the sentiment and sentence structure of a comment.
+This suggests that a sentiment classifier or PoS tagger step needs to be added.
 
@@
 
+#Looking Ahead
+##Mixed Ensemble Method
+Based on the results outlined above, it appears that maximum classification accuracy will be a mixed vertical Ensemble method. Going forward, the working strategy will be to:
+
+*  Pre-screen extracted TA feedback for "short-string" comments consisting of single words, two-word phrases, more punctuation than text, etc. These represent mostly pointers and copy-editing.
+*  Screen comments for 2-grams with high frequencies in one particular category. (Ex., "basic criteria" 2-gram.) 
+*  Use current Naive Bayes classifier to identify Subject.
+*  Evaluate whether a part-of-speech tagger can identify Structure of comments with higher accuracy, either alone when joined with NB preliminary classification.
+
 ####\

### 2019-05-12 futurefaculty.Rmd
update
updates

@@
 A major barrier to routine writing in STEM courses is instructor knowledge.
-Many instructors are unaware that there are publicly available, evidence-based writing training models; instead of using these validated tools, they create their own materials from scratch.
-similarly, many college STEM teachers have never been introduced to proven writing training strategies, nor have they been trained to give effective, context-specific feedback [@Coil2010].
+Many instructors are unaware that evidence-based writing training models are available, so create their own materials from scratch.
+Similarly, many college STEM teachers have never been introduced to proven writing training strategies, nor have they been trained to give effective, growth-oriented feedback [@Coil2010].
 
@@
 ***
-*Works Cited*
+*Work Cited*

### 2019-05-12 index.Rmd
updates

@@
 ##Overview
-I am using classroom research and data science to better understand how students develop technical writing skills in large-enrollment biology courses.
+The STEM Writing Project is using classroom research and data science to better understand how students develop technical writing skills in large-enrollment biology courses.
+Our current NSF-sponsored grant asks if student development as writers can be accelerated using a mix of scripted active instruction, automated support of writing in progress, and holistic feedback.
 
-My current NSF-sponsored grant asks whether we can accelerate student development as writers using a mix of scripted active instruction, automated support of writing in progress, and holistic feedback.
-One specific question we have is how **changing the structure and main focus of instructors' comments** affects students' writing development over time.
-To answer this we extract instructor comments from student reports, categorize the comments according to their subject and structure, then correlate the types of comments instructors make with student performance.
+One specific question we have is **does the structure and main focus of instructors' comments impact the rate of students' writing development over time?** To answer this we must extract instructor comments from student reports, categorize the comments according to their subject and structure, then correlate the types of comments instructors make with student performance.
 
-Classifying TA comments has been informative, but doing it by hand is impractical: 10-12,000 comments need to be scored EACH semester.
-There also is a significant risk of "coding drift" over time, reducing accuracy.
+Classifying TA comments has been informative, but doing it by hand is impractical.
+First, 10-12,000 comments must be scored EACH semester.
+Second, there is a significant risk of "coding drift" over time, reducing accuracy.
 
@@
 #### \   
-##Work Plan
-The plan for this project is to:
+##Outline of Work Plan
+The plan is to:
 
@@
 6.  Apply the optimized NB classifier to the original FULL comment dataset, identify which comments are being classified incorrectly, and look for potential patterns.
-7.  Based on features identified in Steps #3 and #6, write a small set of rule- or REGEX pattern-based searches that can identify comments that are more likely to be classified incorrectly.
 
@@
 
+7.  Based on features identified in Steps #3 and #6, write a small set of rule- or REGEX pattern-based searches that can identify comments that are more likely to be classified incorrectly. 
 8.  Combine the rules/pattern-based pre-screening process with the optimized Naive Bayes classifier to create a vertically  Ensemble Classifier.

### 2019-05-12 litreview.Rmd
update
updates

@@
 
-This project brings together three disparate challenges:
+This project raises three challenges:
 
-*  The challenges of teaching technical writing to students in STEM disciplines
-*  Needs and barriers for training future faculty to teach technical writing
+*  How best to teach students in STEM disciplines to write
+*  How to train future faculty to teach technical writing
 *  Automated text analysis and classification

### 2019-05-13 build.Rmd
update

@@
 ####Extracting TAs' Comments 
-*(Preliminary coding)*  
+*(Draft)*  
 Fully R-based protocol for extracting TA comments from student lab reports. Will replace current process using a mix of Linux and Excel scripts.  
@@
 ####Structure of Development Dataset; Data Import & Validation 
-*(Complete)*  
 Organization of the initial dataset, and how data were imported in R and validated.  
@@
 ####Codebook for Classifying TAs' Comments 
-*(Complete)*  
 Describes the process for developing the codes used to classify comments.  
@@
 ####Initial Exploration of Dataset 
-*(Partly complete)*  
 What is the distribution of words used in the TA comments?  
@@
 ####Naive Bayes Comment Classifier 
-*(Complete)*  
 Code for the most recent version of the optimized NB classifier.   
@@
 ####Embedding Naive Bayes Classifier in an Ensemble Model  
-*(Initial design stage)*
 Outlines a strategy for combining pattern matching and other pre-processing steps with Naive Bayes Classifier to improve accuracy.

### 2019-05-13 docxcommentextraction.Rmd
update

@@
 ---
-title: "DOCX Comment Extraction"
+title: "DRAFT: DOCX Comment Extraction"
 date: "`r format(Sys.time(), '%b %d, %Y')`"
@@
 
-The testing dataset used in this project was created using BaSH scripting and one-off tools. This page documents progress on a workflow to replicate that process using R tools.
+The testing dataset used in this project was created using BaSH scripting and one-off text processing. This page documents progress on building a workflow to replicate that process using R tools.
 
@@
 #General Approach
-Create a list of files to cycle through OR put all files to process in ONE folder.
-For each file, use 'read_docx', then 'docx_extract_all_cmnts'.
-Use 'do.call("rbind", list(dataframes to merge))' to cycle through and append the extracted comments to ONE dataframe
+*  Create a list of files to cycle through OR put all files to process in ONE folder. 
+*  For each file, use 'read_docx', then 'docx_extract_all_cmnts'. 
+*  Use 'do.call("rbind", list(dataframes to merge))' to cycle through and append the extracted comments to ONE dataframe
 
@@
 ```
+
+####\

### 2019-05-13 index.Rmd
update

@@
 ##Project Site Map
-Insert a text-based overview and site map of the major pages and branch points here.
-
-When appropriate, anchor pages branch to extended supporting documentation.
+![](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/images/Sitemap.png)
 

### 2019-05-13 findings.Rmd
update

@@
 
-![*XY plot of relative word use frequency in technical comments vs.
-logic comments.
-*](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/docs/initial_exploration_files/figure-html/unnamed-chunk-3-6.png)
+![*XY plot of relative word use frequency in technical comments vs. logic comments. *](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/images/unnamed-chunk-3-6.png)
 
@@
 ![*Log-Odds ratio comparison of words most different between Writing comments (top) and Technical comments (bottom).
-*](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/docs/initial_exploration_files/figure-html/unnamed-chunk-7-1.png)
+*](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/images/unnamed-chunk-7-1.png)
 
@@
 
-![alt text](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/docs/NB_2r2c_files/figure-html/unnamed-chunk-4-1.png)
+![](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/images/unnamed-chunk-4-1.png)
 
-![alt text](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/docs/NB_2r2c_files/figure-html/unnamed-chunk-6-1.png)
+![](/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/images/unnamed-chunk-6-1.png)
 

### 2019-05-13 docs/hiding/analyze_pkg_usage.R

@@ -1,0 +1,97 @@
+#!/usr/bin/env Rscript
+#Source is Ryan Thompson via Stack Overflow:
+#https://gist.github.com/DarwinAwardWinner/ed75d6d845975a577c8533ac060d6d34
+
+suppressPackageStartupMessages({
+  library(globals)
+  library(readr)
+  library(stringr)
+  library(rex)
+  library(magrittr)
+  library(rlang)
+  library(knitr)
+  library(assertthat)
+  library(withr)
+  library(glue)
+  library(BiocParallel)
+  register(MulticoreParam())
+})
+
+match_identical <- function(x, table, nomatch=NA_integer_) {
+  assert_that(is_list(table))
+  for (i in seq_along(table)) {
+    if (identical(x, table[[i]])) {
+      return(i)
+    }
+  }
+  return(nomatch)
+}
+
+get_Rmd_source <- function(fname) {
+  with_tempfile("temp", {
+    purl(fname, output=temp, quiet=TRUE)
+    read_file(temp)
+  })
+}
+
+analyze_pkg_usage <- function(fname) {
+  if (str_detect(fname, regex(rex(".Rmd", end), ignore_case=TRUE))) {
+    ## Rmd file
+    file_text <- get_Rmd_source(fname)
+  } else {
+    # R file
+    file_text <- read_file(fname)
+  }
+  lib_regex <- rex(
+    "library",
+    zero_or_more(space),
+    "(",
+    zero_or_more(space),
+    capture(one_or_more(any, type="lazy")),
+    zero_or_more(space),
+    one_of(",", ")"))
+  loaded_pkgs <- unique(do.call(rbind, str_match_all(file_text, lib_regex))[,2])
+  missing_pkgs <- character(0)
+  for (pkg in loaded_pkgs) {
+    if (!suppressMessages(suppressWarnings(
+      library(package=pkg, character.only=TRUE, logical.return = TRUE)))) {
+      missing_pkgs <- c(missing_pkgs, pkg)
+    }
+  }
+  loaded_pkgs %<>% setdiff(missing_pkgs)
+  package_envs <- loaded_pkgs %>% set_names %>%
+    lapply(. %>% str_c("package:", .) %>% as.environment)
+  package_namespaces <- loaded_pkgs %>% set_names %>% lapply(asNamespace)
+  
+  file_exprs <- parse(text=file_text)
[…]

### 2019-05-17 docxcommentextraction.Rmd
update

@@
 ---
-title: "DRAFT: DOCX Comment Extraction"
+title: "DRAFT: Extracting Comments From MS Word DOCX Files"
 date: "`r format(Sys.time(), '%b %d, %Y')`"
@@
 
-The testing dataset used in this project was created using BaSH scripting and one-off text processing. This page documents progress on building a workflow to replicate that process using R tools.
+The testing dataset used in to build this project was created using BaSH scripting and one-off text processing.
+This page documents progress on building an R-based repeatable workflow to simplify that process.
 
@@
 #General Approach
-*  Create a list of files to cycle through OR put all files to process in ONE folder. 
+*  Put all files to process in ONE folder. 
+*  Extract a list of file names to cycle through.
 *  For each file, use 'read_docx', then 'docx_extract_all_cmnts'. 
-*  Use 'do.call("rbind", list(dataframes to merge))' to cycle through and append the extracted comments to ONE dataframe
+*  Cycle through and append the extracted comments to ONE dataframe
 
-####\
-
-#Code Sources, Examples
-
-https://stackoverflow.com/questions/50439684/how-to-extract-plain-text-from-docx-file-using-r
-
-https://stackoverflow.com/questions/5965698/merge-unequal-dataframes-and-replace-missing-rows-with-0
-
-https://stat.ethz.ch/R-manual/R-devel/library/base/html/merge.html
-
-https://www.dummies.com/programming/r/how-to-use-the-merge-function-with-data-sets-in-r/
-
-https://datascienceplus.com/combining-data-in-r-the-skill-of-merging-joining-and-stacking/
-
-https://www.rdocumentation.org/packages/analogue/versions/0.17-1/topics/join
-Load required Libraries
-
-####\
-
-#Sample Working Code
-```{}
+```{r Libraries}
+library(plyr)
+library(tidyverse)
+library(reshape2)
+library(tidyr)
+library(readtext)
 library(docxtractr)
@@
 
-Call in Word files using ```read_docx```. Use ```docx_extract_all_cmts``` to pull out comments and ```rbind``` to combine them.
+```{r List_Files}
+#Generates a list of all files in a particular folder
+files_list <-list.files("~/Dropbox/Coding_Tools/R_Environment/R_Projects/DOCX_comment_extraction/Test_Set", full.names=TRUE)
+```
 
-To merge more than two dataframes within R, combine an ```rbind``` call with ```do.call```.
-The 'do.call' function uses any existing R functions, such as 'rbind', but instead of passing just two dataframes, it passes multiple data frames.
-Then use ```write.csv2``` to write the data to a CSV format.
+```{r Pull the comments}
+#### load data ####
+# Define path of directory
+# Get full path filename of every .docx-file in directory as 'filelist'
+# Get filenames without paths for same sequence as 'filelist_short'
+pathto <- "~/Dropbox/Coding_Tools/R_Environment/R_Projects/DOCX_comment_extraction/Test_Set/Hiding"
+filelist <- list.files(path = pathto, pattern = "*.docx", full.names = TRUE)
+filelist_short <- list.files(path = pathto, pattern = "*.docx", full.names = FALSE)
+
+# Read every file with docxtractr::read_docx()
+# Extract every comment from every file with docxtractr::docx_extract_all_cmnts()
+document_contents_list <- lapply(filelist, read_docx)
+comments_list <- lapply(document_contents_list, docx_extract_all_cmnts, include_text=TRUE)
+
+```
+
+The variable named 'comments_list' is a  list of length 6. Each item in the list is a tibble with variable number of rows, but all with the same 6 columns:
+
+*  id
+*  author
+*  date
+*  initials
+*  comment_text
+*  word_src
+
+I need to append each tibble to the next one for all items in the list.
+
+```{r Bind together the individual items in the comments_list List}
+
+# identify the filename of the docx file that was used to create the list of comments
+document_contents_list[[3]]$path
+
+# join the individual lists of comments into a larger list
+comments_list_joined <- bind_rows(comments_list, id=NULL)
+
+# problem is that I lose the filenames.
+
+```
+
+***
+Problem to solve now is the joined comments list contains no record of the source file's name. Need some way to attach the remaining report metadata.
+***
+
+```{r Write CSV}
+#Creates the output CSV file if needed for analysis outside R.
+write.csv(comments_list_joined, file = "~/Dropbox/Coding_Tools/R_Environment/R_Projects/DOCX_comment_extraction/Test_Set/Extracted_comments.csv")
+```
 
 ```{}
-Commentframe1<-read_docx("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/DOCX_comment_extraction/TestingReport.docx", track_changes = NULL)
-comments1<-docx_extract_all_cmnts(Commentframe1, include_text = FALSE)
 
-Commentframe2<-read_docx("/Users/danjohnson/Dropbox/Coding_Tools/R_Environment/R_Projects/DOCX_comment_extraction/TestingCommentExtraction.docx", track_changes = NULL)
-comments2<-docx_extract_all_cmnts(Commentframe2, include_text = FALSE)
-
-# comments1And2 <- rbind(comments1,comments2)
-# OR
-# comments1And2 <- do.call("rbind", list(comments1,comments2))
-
-#Check data have been successfully merged by checking the number of rows
-NROW(comments1And2)
-# [1] 32
-
-#When all lines are appended, write the CSV file 
-write.csv2(comments1And2, file = "TAcomments.csv")
+#Rename the columns in a data.frame. In this example, "name" has 2 columns. The command pipes new names to the 2 columns, using the vectorized text names.
+names(name_of_file_to_change) <- c("columns", "values")
 

### 2019-07-03 NB_v3_new_data_input.Rmd
Update

--- 2019-05-12 NB_2r2c.Rmd
@@
 
-#Train Model, Predict & Evaluate for Subject
-This block predicts whether a message is likely to be in group 1_basic, 2_writing, 3_technical, or 4_logic. 
-```{r bayes_training subject}
-# Train the classifier on training data, then test on test dataset, and full dataset.
-comments_classifier_subject <- naiveBayes(comments_train, comments_train_labels_subject, laplace=1)
-comments_test_pred_subject <- predict(comments_classifier_subject, comments_test)
-comments_all_pred_subject <- predict(comments_classifier_subject, comments_all)
-```
+Here is where I need to add code to format NEW datasets, and drop the text DTM into the trained classifier. Pseudo-code is:
+  Import Sp19_comments.csv
+  Extract ta.comments to a new data object
+  Run the ta.comments object through processing:
+
+{r input data}
+#Read in TA comments from NEW CSV file.
+base_dataVERB <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
+
+#Select rows representing the sub-groups to compare. 
+comments_subsetVERB <- filter(base_dataVERB,code.subject=="1_basic"|code.subject=="2_writing"|code.subject=="3_technical"|code.subject=="4_logic")
+
+#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
+comments_rawVERB <- comments_subsetVERB %>% select(23,24,22)
+
+#Rename the columns.
+names(comments_rawVERB)[1] <- "subject"
+names(comments_rawVERB)[2] <- "structure"
+names(comments_rawVERB)[3] <- "text"
+
+#Change "subject" element from character to a factor for analysis.
+comments_rawVERB$subject <- factor(comments_rawVERB$subject)
+str(comments_rawVERB$subject)
+table(comments_rawVERB$subject)
+
+#Change "structure" element from character to a factor for analysis.
+comments_rawVERB$structure <- factor(comments_rawVERB$structure)
+str(comments_rawVERB$structure)
+table(comments_rawVERB$structure)
+
+#Convert data set to a volative corpus using "tm" library.
+{r corpusmaker}
+#Create the volatile corpus that contains the "text" vector from data frame.
+comments_corpusVERB_R <- VCorpus(VectorSource(comments_rawVERB_R$text))
+print(comments_corpusVERB_R)
+
+#Check out the first few text comments in the new corpus, which is basically a list that can be manipulated with list operations.
+inspect(comments_corpusVERB_R[1:3])
+
+#Use "as.character" function to see what a single text comment looks like.
+as.character(comments_corpusVERB_R[[5]])
+
+#Text Data Transformations
+The text data transforms in this code block are intentially separate "tm_map"" and "content_transformer" functions.
+This simplifies the process of testing how various combinations affect classification.
+Converting the text to lower case and removing punctuation are the baseline transformations.
+Switch between "{r}" and "{}" to turn other transformation blocks off or on as desired.
+**Removing extra white spaces must be the last transformation, and is required.**
+
+{r textcleanup}
+#Convert text to all lower case letters.
+comments_corpus_cleanVERB_R <- tm_map(comments_corpusVERB_R, content_transformer(tolower))
+
+#Remove punctuation using the "removePunctuation" function. 
+#This step removes evidence of questions, so may remove data.
+#THIS ALSO REMOVES EVIDENCE OF POINTER ITEMS
+comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removePunctuation)
+
+#Remove numerals. 
+#This step may remove evidence indicating technical comments
+comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeNumbers)
+
+#Stopword removal. 
+#This is a standard cleanup step, but again may remove useful terms.
+#The "stopwords" file is a convention; other files can be substituted for it
+comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, removeWords, stopwords())
+
+#The option to apply stemming has been removed from this version. Stemming reduced accuracy in all trials. Future iterations should examine lemmatization instead.
+
+#Final step removes white space from the document. This is NOT an optional step.
+comments_corpus_cleanVERB_R <- tm_map(comments_corpus_cleanVERB_R, stripWhitespace)
+
+#Look at an example of cleaned text comment to see if cleaned comment matches what is expected. Change the value inside double brackets to look at a different comment.
+as.character((comments_corpus_cleanVERB_R[[5]]))
+
+#Tokenize the N-Grams
+This code block uses tm's NLP commands to assess different n-grams.
+For single words, change the "#" variable in "(words(x), #)" to "1".
+Use other values to change the n-gram size.
+Data are stored in document-term matrix (dtm) format.
+
+{r tokenizer}
+NLP_Tokenizer <- function(x) {
+      unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)
+}
+
+comments_dtm_1gram <- DocumentTermMatrix(comments_corpus_cleanVERB_R, control=list(tokenize = NLP_Tokenizer))
+comments_dtm_1gram_tidy <- tidy(comments_dtm_1gram)
+inspect(comments_dtm_1gram)
+
+Some subset of the above steps SHOULD convert the new dataset to a readable format for comparison. END GOAL is to have a data object that can be plugged into the "predict" command.
+
+***
+Break
+***
 

### 2019-11-15 NB_v4_new_data.Rmd
update

--- 2019-05-12 NB_2r2c.Rmd
@@
 
+####\
+
 A truth table lists how many of the predicted categories are in their correct categories. Adding diagonally and dividing by total count gives the percent accuracy at predicting Subject. 
@@
 table("Predictions"= comments_all_pred_subject,  "Actual Subject All" = comments_all_labels_subject )
-table("Predictions"= comments_test_pred_structure,  "Actual Structure Test" = comments_test_labels_structure )
-table("Predictions"= comments_all_pred_structure,  "Actual Structure All" = comments_all_labels_structure )
+#table("Predictions"= comments_test_pred_structure,  "Actual Structure Test" = comments_test_labels_structure )
+#table("Predictions"= comments_all_pred_structure,  "Actual Structure All" = comments_all_labels_structure )
 
@@
 conf.mat2 <- confusionMatrix(comments_all_pred_subject, comments_all_labels_subject)
-conf.mat3 <- confusionMatrix(comments_test_pred_structure, comments_test_labels_structure)
-conf.mat4 <- confusionMatrix(comments_all_pred_structure, comments_all_labels_structure)
+#conf.mat3 <- confusionMatrix(comments_test_pred_structure, comments_test_labels_structure)
+#conf.mat4 <- confusionMatrix(comments_all_pred_structure, comments_all_labels_structure)
 
@@
 # This block outputs both the overall and by-group accuracy values for Structure x Test Subset.
-print ("Test, Structure")
-conf.mat3
+#print ("Test, Structure")
+#conf.mat3
 
 # This block outputs by-group accuracy values in a single column for Structure x Test Subset. 
-print ("Test, Structure")
-conf.mat3$byClass
+#print ("Test, Structure")
+#conf.mat3$byClass
 
 # This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Structure x Test Subset.
-print ("Test, Structure")
-conf.mat3$overall
+#print ("Test, Structure")
+#conf.mat3$overall
 
 # This block outputs JUST the overall accuracy value for Structure x Test Subset. 
-conf.mat3$overall['Accuracy']
-print ("Test, Structure")
+#conf.mat3$overall['Accuracy']
+#print ("Test, Structure")
 
 # This block outputs both the overall and by-group accuracy values for Structure x All Data.
-print ("All, Structure")
-conf.mat4
+#print ("All, Structure")
+#conf.mat4
 
 # This block outputs by-group accuracy values in a single column for Structure x All Data. 
-print ("All, Structure")
-conf.mat4$byClass
+#print ("All, Structure")
+#conf.mat4$byClass
 
 # This block outputs a single line table with overall accuracy, kappa, lower and higher limits of accuracy, accuracy null value, P value for accuracy, and Mcnemar P value for Structure x All Data.
-print ("All, Structure")
-conf.mat4$overall
+#print ("All, Structure")
+#conf.mat4$overall
 
 # This block outputs JUST the overall accuracy value for Structure x All Data. 
-print ("All, Structure")
-conf.mat4$overall['Accuracy']
+#print ("All, Structure")
+#conf.mat4$overall['Accuracy']
 ```

### 2019-11-15 background.Rmd
update

@@
 Undergraduate introductory STEM courses often include technical writing training.
-Often it is secondary to content coverage, and so treated as an incidental skill that each student develops "along the way." Alternatively, the writing instruction process may be haphazardly designed, implemented, and executed.
+Often it is secondary to content coverage, and treated as an incidental skill that each student develops "along the way." Alternatively, the writing instruction process may be haphazardly designed, implemented, and executed.
 Adding to the learning challenge, faculty may not be who is helping students develop writing skills.
